### 2020-03-18 Analysis/pilot.Rmd
Commit set up of experiment pilot data, pilot analysis

---
title: "Provo Maze Pilot"
output: github_document
---

```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

A pilot for using Maze on Provo.

# Experiment overview
Materials are taken from the Provo corpus. The corpus contains 55 paragraphs each consisting of 1-4 sentences, about 50 words per paragraph. There are 134 sentences total, with 2689 words total. For this experiment, each participant sees a randomly selected 25 of the paragraphs, split into 5 blocks of 5. These are preceded by a practice block of 8 practice sentences.

I had intended to recruit 30 subjects, but stopped pushing HITs after 20 subjects worth had started because I was afraid results weren't writing; turned out that participants where slower to start/finish the experiment than I thought.

These were all presented using the Maze paradigm, with participants having to redo mistakes (ie, if they press the button for the wrong answer, an error message is displayed and they must press the other/correct key to continue with the sentence).

# Data prep

<!-- Something screwed up with group number on critical items (looks fine? for practice, def not otherwise). Looks like either in pushing the items, or in recording the results the groups for the ones that were pushed are just the first digit (not sure how that happened). I should try to trouble-shoot this, but for this it means we can't use that to join and have to use other sources of information. -->

```{r}
#file of sentences/words labelled with length, word frequency and surprisal
#created by prep.Rmd, note that 0's in surprisal are really NAs
surprisal <- read_rds("../Materials/surprisal.rds") %>% 
  mutate(word_num=Word_In_Paragraph_Number-1,
         )

data <- read_rds("../Data/provo_pilot_1.rds") %>% 
  filter(native=="yes" & residence=="yes") %>% 
  select(-native, -residence) %>% 
  rename(Text=sentence, Word=word) %>% 
  left_join(surprisal, by=c("word_num","Text","Word")) %>% 
  mutate(Text_ID=ifelse(type=="practice",group,Text_ID)) %>%  #for practice sentences, have 101 etc numbers 
  select(-group) %>% 
  mutate(word_num_mistake=ifelse(correct=="no", word_num,NA),word_num_mistake2=word_num_mistake) %>% 
  group_by(Text_ID, subject) %>% fill(word_num_mistake) %>% 
  mutate(after_mistake=word_num-word_num_mistake,
         after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
  ungroup() %>% 
  write_rds("../Data/provo_pilot_1_labelled.rds")
```

19 subjects remaining after non-native, non-resident exclusion

# Overall RT, errors

How many words did each participant see?

```{r}

 data %>% group_by(subject) %>% tally() #how many words / participant

by_part_error_rate <- data %>% group_by(subject) %>%
  mutate(corr.num=ifelse(correct=="yes", 1,0)) %>%
  summarize(fraction_correct=sum(corr.num)/n(), median_rt=median(rt), total_rt=sum(rt)/60000)

ggplot(by_part_error_rate, aes(x=fraction_correct, y=median_rt))+
  geom_point()+
  coord_cartesian(x=c(0,1), y=c(0,1000))+
  labs(title="By participant error rates versus median rt")

ggplot(by_part_error_rate, aes(x=fraction_correct, y=total_rt))+
  geom_point()+
  coord_cartesian(x=c(0,1), y=c(0,40))+
  labs(title="By participant error rates versus summed rt")

```

Each participant saw roughly 1300 words. All but two participants were attentive. Total reading time is the sum of recorded reading times in minutes, but does not include time to read instructions, do demographics, take pauses or correct mistakes. It thus should be taken as a lower bound on how long this took (but reading instructions, etc can reasonably be estimated to take no more than a couple minutes). Task was estimated to take 20 minutes, looks like that was about right, but given variance, we should pay fair wage for 25 or 30 minutes.

Would be an interesting question to look into how to get more of the sample from the attentive end and not the inattentive (in ORC pilot, qualitatively same distribution but the inattentive clump was much bigger). Could also be task effects, maybe this was less frustrating, easier to do "correctly"??

# Error rate

Label who's attentive v not for future sorting.

```{r}
data_sort <- data %>% left_join(by_part_error_rate, by="subject") %>% 
  mutate(is.attentive=ifelse(fraction_correct>.75,TRUE,F),
         is.correct=ifelse(correct=="yes",T,F)) %>% 
  filter(is.attentive)

```

Of attentive participants, what after mistake data can we trust? As a rough proxy, we'll consider it trustworthy if it had the same RT profile as elsewhere. (This may not be the best proxy for reliability.)

```{r}

after_mistake <- data_sort %>% 
  filter(type=="critical") %>% 
  filter(rt>100) %>% 
  filter(rt<5000) %>% 
  filter(is.correct) %>% 
  mutate(after_mistake_group=ifelse(after_mistake>10, 11, after_mistake),
         after_mistake_fct=as_factor(after_mistake_group),
         word_num_fct=as_factor(word_num))

ggplot(after_mistake, aes(x=after_mistake_fct,y=rt))+geom_boxplot()+coord_cartesian(ylim=c(500,1500))+labs(x="Position relative to last mistake, 0=before mistake, 1=word after mistake", title="Does being directly after a mistake inflate RT?")

#ggplot(after_mistake, aes(group=word_num_fct,x=word_num_fct, y=rt))+geom_boxplot()+coord_cartesian(ylim=c(0,2000))+labs(title="Position in sentence v RT")

after_mistake_summ <- after_mistake %>% 
  group_by(after_mistake_fct) %>% 
  summarize(mean_rt=mean(rt),
            median_rt=median(rt))

after_mistake_summ
```

It looks like the word after a mistake is higher, but after that it's fine. For the after mistake, the 11 category is really 11+ and contains words from farther along as well.

# Relevant analysis

The critical analysis here is whether RT reflects known gross RT effects, namely longer, less frequent, more surprising words reading more slowly.

For this pilot, we operationalize all of these in easy to calculate ways.

- frequency is  log10 of the number of occurances over 10^9 words in text (as calculated for the python wordfreq package). So,coefficient represents how RT changes  per going up a factor of 10 in word frequency.
- length is length in letters. Coefficient represents change in RT from increasing length by one letter.
- surprisal is determined by the Gulordava lstm model and represented in bits of surprisal. Coefficient represents how RT changes when expectation of the word is halved (i.e. going up one bit of surprisal).

All of these variables were centered, to make for a more interpretable intercept, but have their original scales.

While these do correlate with each other somewhat, we'd like to see them all acting independently at once.

Data exclusions:
Words where errors were made are excluded along with words immediately after errors. All other after mistake data is included.

In order to have reasonable analyses using RT as the dependent measure (untransformed), we want to do light exclusions so that outliers don't have too much pull. Here we exclude <100 ms (which is unreasonably short if you're doing any processing at all) and >5000 ms (which excludes things where someone paused to do somehthing else, but doesn't exclude any of the main curve of data). These two exclusions together exclude 50 data points, which isn't very many.

Data points are also implicitly excluded if we don't have frequency and surprisal values for them (formula implicitly excluded NAs). Among other things, this excludes the first word of every sentence.

```{r}
data_correct <- data_sort %>% 
  filter(is.correct==T) %>% 
  filter(type=="critical") %>% 
  filter(rt<5000) %>% #eliminates 40 data points
  filter(rt>100) %>% #eliminates 10 data points
  filter(after_mistake!=1) %>% 
  select(word_num, Word, rt, subject, Text_ID, Sentence_Number, Word_In_Sentence_Number, freq_cent, surp_cent, word_len_cent) %>%
  ungroup() %>% 
  mutate(word_num_fct=as_factor(word_num),
         Text_ID_fct=as_factor(Text_ID))

ggplot(data_correct) +geom_density(aes(rt))+labs("Density distribution of remaining RTs")

```

We do the models first with lm without hierarchical structure, and then repeat with appropriate mixed effects in brms.

```{r}

m1_all_no_interact <- lm(rt~ freq_cent+surp_cent+word_len_cent, data_correct)

m1_all <- lm(rt~ freq_cent*surp_cent*word_len_cent, data_correct)

summary(m1_all_no_interact)

summary(m1_all)

```

 Structure wise, we want to nest by subject (some could be more or less sensitive to any of this stuff), and word_num nested within Text_ID. Unclear if Text_ID is really a natural category, but words within the same passage have the same topic (which might contribute to difficulty?). Of course, individual words may have lots of unmeasured properties such as errors in the surprisal/etc or properties of their distractor.

```{r}

show_summary <- function(model){
  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
  stats <- gather_draws(model, `b_.*`, regex=T) %>% mutate(above_0=ifelse(.value>0, 1,0)) %>% group_by(.variable) %>% summarize(pct_above_0=mean(above_0)) %>% mutate(pval_equiv = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% left_join(intervals, by=".variable") %>% select(term=.variable, estimate=.value, lower=.lower, upper=.upper, pval_equiv)

  stats
}

```

```{r}
priors <- c(
    set_prior("normal(1000, 1000)", class="Intercept"),
    set_prior("normal(0, 500)", class="b"),
    set_prior("normal(0, 500)", class="sd"),
    set_prior("lkj(1)",       class="cor"))

m3_all_no_interact <- brm(rt ~ freq_cent+surp_cent+word_len_cent+
                            (freq_cent+surp_cent+word_len_cent|subject)+
                            (freq_cent+surp_cent+word_len_cent|Text_ID_fct/word_num_fct),
                          prior=priors, data=data_correct, file="m3_all_no_interact")

m3_all <- brm(rt ~ freq_cent*surp_cent*word_len_cent+
                (freq_cent*surp_cent*word_len_cent|subject)+
                (freq_cent*surp_cent*word_len_cent|Text_ID_fct/word_num_fct),
              prior=priors, data=data_correct, file="m3_all")
```

```{r} 
summary(m3_all_no_interact)

summary(m3_all)

a <- show_summary(m3_all_no_interact)
a

b <- show_summary(m3_all)
b
```

In general, there's reasonable agreement among the models. The intercept is somewhere around 900 ms, each additional bit of surprisal add about 20 ms. Frequency has a negative coefficient (although there's a lot of uncertainty about what the magnitude is), length has a positive effect of about 5 ms per character (although this isn't significant when interaction terms are present, and shows up in surprisal/length and three way interactions).

# Future TODOs, questions

- What is the right way to determine how much margin to leave after a mistake?

- TODO: use the error patterns to identify points of improvement for the auto-generation task.

- Should there be interaction terms?

- What other models of surprisal or frequency should I be testing?

- Should I try non-linear models of RT?

### 2020-03-18 Analysis/read_results.R

@@ -1,0 +1,72 @@
+library(tidyverse)
+library(stringr)
+library(urltools)
+#function reads in data from a maze file plus demographic, other questions.
+read_in_data <- function(filename){
+  #reads data in generically because different format for different controllers
+#filename="pilot_1_raw"
+  data <- read_csv(filename, comment="#", col_names=c("time", "MD5", "controller", "item", "elem",
+                                                      "type", "group", "col_8", "col_9", "col_10", "col_11", "col_12", "col_13", "col_14"), col_types=cols(
+                                                        time=col_integer(),
+                                                        MD5=col_character(),
+                                                        controller=col_character(),
+                                                        item=col_integer(),
+                                                        elem=col_integer(),
+                                                        type=col_character(),
+                                                        group=col_integer(),
+                                                        col_8=col_character(),
+                                                        col_9=col_character(),
+                                                        col_10=col_character(),
+                                                        col_11=col_character(),
+                                                        col_12=col_character(),
+                                                        col_13=col_character(),
+                                                        col_14=col_character()
+                                                      )) %>% mutate_all(url_decode) #deal with %2C issue
+  
+  order_seen <- data %>% 
+    select(time, MD5, group) %>% 
+    filter(!is.na(group)) %>% 
+    unique() %>% 
+    group_by(time, MD5) %>% 
+    mutate(trial_num = 1:n()) %>% 
+    type_convert()
+  
+  #split off non-maze, participant level stuff and process them
+  other <- filter(data, controller %in% c("Form", "Question")) %>% 
+    select(time, MD5, col_8, col_9) %>% 
+    group_by(time, MD5) %>%
+    pivot_wider(names_from=col_8, values_from=col_9) %>% 
+    rename(gender=`Please select your gender.`,
+           native=`Are you a native speaker of English?`,
+           residence=`Do you currently reside in the United States?`) %>% 
+    type_convert()
+  #TODO:read for comments
+  
+  #take the Maze task results, relabel and type appropriately
+  maze<- filter(data, controller=="Maze") %>% 
+    select(time, MD5, type, group, word_num=col_8, word=col_9, distractor=col_10, on_right=col_11, correct=col_12, rt=col_13, sentence=col_14) %>% 
[…]

### 2020-04-13 Analysis/nat_stories.Rmd
qualitative analysis

---
title: "Natural Stories"
output: github_document
---
```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

# Experiment

Motivation: The point of this experiment was to determine if Maze works on naturalistic text. Do participants comprehend what they are reading while doing this task? Do frequency, length and surprisal effects show up in the reading times?

To this end, we tested participants on stories from the Natural Stories corpus. The corpus consists of 10 stories (some fiction, some non-fiction) each about 1000 words, and has 6 comprehension questions for each. It has previously been run as SPR, where freq/surprisal/length show up and we have the comprehension question accuracies, so we have something to compare to.

1000 words is a lot for Maze especially when it needs to be done in one chunk (whereas we'd put pause screens in previous maze experiments, I didn't here because it could disrupt the story), so we had each participant only read one story. This was prefaced by a short practice item (~100 words) that I wrote, which had 2 practice questions.

At the end, we collected some optional demographics and debriefed participants.

We recruited 100 participants, across the 10 stories. (Stories were randomized, but not balanced, so we'll see how balanced the data is.) We estimated that this would take around 20 minutes (given how long past studies had taken) and paid 3.50.

# Assorted preliminaries

```{r participants}

data <- read_rds("../Data/cleaned.rds")

data_filt <- data %>% filter(native %in% c("ENG", "English", "ENGLISH", "english")) #I peeked at what people put that semantically maps to english
```

We got data from `r data %>% select(subject) %>% unique() %>% nrow()` subjects, after exluding those with a reported native language other than English, `r data_filt %>% select(subject) %>% unique() %>% nrow()` remain.

```{r stories}

data_stories <- data_filt %>% 
  select(type, subject) %>% 
  unique() %>% 
  group_by(type) %>% 
  tally() %>% 
  filter(type!="practice")

```

Within the filtered data, each story was read between `r min(data_stories$n)` and `r max(data_stories$n)` times. This seems like decent distribution.

## Error rates

There's lots of ways to try to get at participant diligence that we might want to use for exclusions. IN particular, it appears that mean accuracy is correlated with mean rt -- people with high accuracy take longer than people who look like they are randomly guessing. In this experiment, we have an additional source of insight into diligence -- comprehension question accuracy. In the SPR natural stories experiment, they used 5 out of 6 comprehension questions correct as their threshold for data inclusion.

The below plot shows how accurate participants were on the task (measured per word), versus mean reaction time (note that a filter of rt<5000 was used to exclude times when a participant got distracted or paused in doing the experiment). Coloring indicates how many of the comprehension questions (out of 6) they got right, and the grids are faceted on the SPR exclusion criteria for comparison. (Practice items and practice questions are not included here.)

```{r errors}

data_error_summ <- data_filt %>% 
  mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
  group_by(subject, critical, practice) %>%
  filter(type!="practice") %>% 
  filter(rt<5000) %>% 
  summarize(mean_rt=mean(rt),
            pct_correct=mean(correct.num)) %>% 
  mutate(critical_factor=as.character(critical*6),
         good_comp=ifelse(critical>.8,"5 or 6 correct","4 or fewer correct"))

ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=critical_factor))+
  geom_point()+
  facet_grid(.~good_comp)+
  labs(title="Relationship between error rate, RT, and comp ? accuracy",
       x="Percent maze correct",
       y="Mean reaction time",
       color="?s correct")

```

It appears that it is indeed possible to comprehend these stories while doing the maze task. There's clearly a correlation between accuracy on comprehension questions and accuracy on maze, although there is some variability. When the task is done well, average RT is between 600 and 1200 -- average RTs below 500 are unlikely to be attentive.

Unclear what the correct thresholds for exclusions should be, although I'm tempted by an 80% accuracy cut-off (for RT stuff at least). If measures of interest are about comprehension, comprehension questions seem reasonable.

## Experiment length, pay
The other point of interest is how long the experiment takes overall; we're curious about this for designing appropriate length experiments and compensating adequately in the future. This takes into account all the time spent doing maze, including practice items and correcting mistakes. It does not include time to read instructions, answer comprehension questions or fill out demographics.

```{r total-rt}

total_time <- data_filt %>% 
  mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
  group_by(subject, critical, practice) %>%
  summarize(total_rt=sum(total_rt)/60000,
            pct_correct=mean(correct.num)) %>% 
  mutate(critical_factor=as.character(critical*6),
         good_comp=ifelse(critical>.8,"5 or 6 correct","4 or fewer correct"))

ggplot(total_time, aes(x=pct_correct, y=total_rt, color=critical_factor))+
  geom_point()+
  facet_grid(.~good_comp)+
  labs(title="How long does this take",
       x="Percent maze correct",
       y="Total maze time",
       color="?s correct")

```

It seems like 20 minutes isn't a horribly off estimate, but that 25 minutes is more reasonable if we want an upper bound accounting for some time to do demographics.

In the random questions at the end section, particpants were asked how much they thought this should pay, common answers where 3.50, 4, and 5. (Experiment paid 3.50, maybe we should consider upping it a little, depending on what the timing data is.)

```{r pay}
pay <- data_filt %>% select(subject, pay) %>% unique() %>% 
  mutate(pay.numeric=as.numeric(str_remove_all(pay, "[^0-9.]"))) %>% 
  group_by(pay.numeric) %>% tally()

kable(pay) 
```

Taking these two pieces together, paying 4 or 5 dollars is probably better.

## Participants experience

What participants thought: participants were both asked what they thought the experiment was about and whether they used any strategies. (Second question was prompted from hearing that other people running maze get answers like "look at one side, only check the other if it doesn't seem plausible" and the like.)

```{r experience}
#TODO waiting for by participant error rate data

thoughts <- data_filt %>% select(subject, topic, strategy) %>% 
  unique() %>% 
  left_join(data_error_summ, by="subject") %>% 
  arrange(desc(pct_correct)) %>% select(pct_correct, topic, strategy)

kable(thoughts)
```

Mostly no one mentions strategies, some read out loud, stories are followable. (There's certainly some correlation between longer/more coherent/good faith answers and higher accuracies...)

# Comprehension question accuracy

We want to compare how good our participants were on the comprehension questions to how good people were on the SPR experiment reported in Futrell et al (2017). They had each story read about 100 times across 181 participants (participants generally read 5 stories). This means they have about 10 times as much data as we do. However, reading times in a story were excluded based on the comprehension of that story, so it makes sense to compare per-story accuracies with our per-story accuracies.

Here I plot the distribution of accuracies (number correct out of 6) for their participants (SPR), all of our participants (all Maze), and a restricted set of our participants who got at least 80% of the words right (Maze >80% correct), an exclusion threshold I would consider using.

```{r spr, include=F}
#want to compare to SPR data from paper

spr <- read_csv("../ns_spr_1.csv") %>% 
  union(read_csv("../ns_spr_2.csv")) %>% 
  select(WorkerId, correct, item) %>% 
  unique() %>% 
  group_by(correct) %>% 
  tally() %>% 
  mutate(pct=n/988) %>% 
  mutate(source="SPR") %>% 
  select(source,correct,pct)

all <- data_filt %>% select(subject, critical) %>% 
  unique() %>% 
  mutate(correct=critical*6) %>% 
   group_by(correct) %>% 
  tally() %>% 
  mutate(pct=n/95) %>% 
  mutate(source="all Maze") %>% 
  select(source,correct, pct)

some <- data_filt %>% select(subject, critical) %>% 
  unique() %>% 
  left_join(data_error_summ, by=c("subject", "critical")) %>% 
  filter(pct_correct>.8) %>% 
  mutate(correct=critical*6) %>% 
   group_by(correct) %>% 
  tally() %>% 
  mutate(pct=n/63) %>% 
  mutate(source="Maze >80% correct") %>% 
  select(source,correct, pct)

to_plot <- all %>% union(some) %>% union(spr)

ggplot(to_plot, aes(x=correct, y=pct, fill=source))+geom_col(position="dodge")+facet_grid(.~source)+labs(title="Comprehension question accuracy", x="Number correct", y="Percent of participants")

```

We're not getting as good comprehension as they did on SPR. They got 90% with 5 or 6 correct, we get 59%, which rises to 79% if we only include participants who were getting at least 80% of the words right. However, we also have to remember that it's about 3 times slower reading Maze well than doing SPR (they report average reading times around 330 ms), so the memory delay is greater with Maze. But we do still get some participants who do it. (There's also the fact that despite both studies sourcing participants from Mturk, the demographics have changed in between when the studies were run.)

# How do errors affect participants and when do they occur

From here on, we're only going to consider participants who got 80% correct on the maze.

```{r exclusion}
data_good <- data_filt %>% 
  left_join(data_error_summ, by=c("subject", "critical")) %>% 
  filter(type!="practice") %>% 
  filter(pct_correct>.8) %>% 
  mutate(is.correct=ifelse(correct=="yes",1,0))

```
This includes data from `r data_good %>% select(subject) %>% unique() %>% nrow()` participants.

When in sentences do errors occur?

```{r when-error}

when <- data_good %>% group_by(word_num) %>% 
  summarize(error_rate=1-mean(is.correct),
            count=n())

ggplot(when, aes(x=word_num, y=error_rate, size=count))+
  geom_point()+
  labs(title="Errors are relatively evenly distributed",
       x="Word number in sentence, 0 is not a real choice",
       y="By position error rate",
       size="Data quantity")
```

The 2nd and 3rd words in a sentence seem to be more error-prone than other words, perhaps due to worse distractors, but there isn't a huge difference.

We're also interested in how rapidly participants fix their mistakes, as this may give us insight into how many post-mistake words to exclude. In the below, each data point is a mistake, plotted on the x axis for the time to press the incorrect button and on the y on the *additional* time before pressing the correct button.

```{r after}

timing <- data_good %>% 
  filter(!is.correct) %>% 
  mutate(time_to_correct=total_rt-rt) %>% 
  filter(rt>100 & rt<5000& time_to_correct>0 & time_to_correct<5000)

ggplot(timing, aes(x=rt, y=time_to_correct))+geom_point(alpha=.1)+
  labs(title="How long does it take to correct a mistake?",
       x="Time to make mistake",
       y="Time to correct mistake")
```

It's quick to correct mistakes; it tends to take less time to correct a mistake than to read a new word, usually less than a second.

Given that, it seems like we don't need to exclude much data after a mistake. Two words seems safe. (This is totally arbitrary and hopeful no analysis is sensitive to it.)

### 2020-04-13 Analysis/read_results.R

@@ -1,0 +1,83 @@
+library(tidyverse)
+library(stringr)
+library(urltools)
+#function reads in data from a maze file plus demographic, other questions.
+read_in_data <- function(filename){
+  #reads data in generically because different format for different controllers
+#filename="../Data/raw_data"
+  data <- read_csv(filename, comment="#", col_names=c("time", "MD5", "controller", "item", "elem",
+                                                      "type", "group", "col_8", "col_9", "col_10", "col_11", "col_12", "col_13", "col_14", "col_15"), col_types=cols(
+                                                        time=col_integer(),
+                                                        MD5=col_character(),
+                                                        controller=col_character(),
+                                                        item=col_integer(),
+                                                        elem=col_integer(),
+                                                        type=col_character(),
+                                                        group=col_character(),
+                                                        col_8=col_character(),
+                                                        col_9=col_character(),
+                                                        col_10=col_character(),
+                                                        col_11=col_character(),
+                                                        col_12=col_character(),
+                                                        col_13=col_character(),
+                                                        col_14=col_character(),
+                                                        col_15=col_character()
+                                                      )) %>% mutate_all(url_decode) #deal with %2C issue
+  
+  #split off non-maze, participant level stuff and process them
+  other <- filter(data, type %in% c("questionnaire")) %>% 
+    select(time, MD5, col_8, col_9) %>% 
+    group_by(time, MD5) %>%
+    pivot_wider(names_from=col_8, values_from=col_9) %>% 
+    rename(gender=`Please select your gender.`) %>% 
+    type_convert()
+  #TODO:read for comments
+  
+  comp_q <- filter(data, type %in% c("practice_question","question_critical1","question_critical2",
+                                     "question_critical3","question_critical4","question_critical5",
+                                     "question_critical6","question_critical7","question_critical8","question_critical9","question_critical10")) %>% 
+    select(time, MD5, type, question=col_8, answer=col_9, correct_answer=col_10, q_time=col_11) %>% 
+    type_convert(col_types=cols(
+      time=col_integer(),
+      MD5=col_character(),
[…]

### 2020-04-14 Prep_code/nat_stories_prep.Rmd
code for getting surprisals, freqs, etc

---
title: "R Notebook"
output: github_document
---

```{r}
library(tidyverse)
library(readr)
```
# Initial
These are used to set things up that exist in other saved files.
```{r}
#Paragraph format it

# data <- read_delim("natural_stories_words.tsv",delim="\t", col_names = c("key","value")) %>% 
#   separate(key, into=c("Story", "word_num", "col3")) %>% 
#   filter(col3=="whole") %>% 
#   select(-col3) %>% 
#   separate("value", into=c("a","b","c","d","e","f","g"), sep=" ") %>% 
#   unite(col="Word",-Story,-word_num, sep="", na.rm=T) %>% 
#   group_by(Story) %>%
#   pivot_wider(names_from=word_num,values_from=Word) %>%
#   unite(col="Text", -Story, sep=" ", na.rm=T) %>%
#   write_delim("../Materials/natural_stories_draft.tsv", delim="\t")
```

```{r words}

words <- read_delim("natural_stories_words.tsv",delim="\t", col_names = c("key","value")) %>% 
  separate(key, into=c("Story_Num", "Word_In_Story_Num", "col3"), convert=T) %>% 
  filter(col3=="whole") %>% 
  select(-col3) %>% 
  separate("value", into=c("a","b","c","d","e","f","g"), sep=" ") %>% 
  unite(col="Word",-Story_Num,-Word_In_Story_Num, sep="", na.rm=T)

nums <- 1:64%>% as.character()

labelled <- read_delim("../Materials/natural_stories_sentences.tsv", delim="\t") %>% 
  separate(Sentence, into=nums, sep=" ") %>% 
  pivot_longer(cols=`1`:`64`, names_to="Word_In_Sentence_Num",values_to="Word") %>% 
  mutate(Word_In_Sentence_Num=as.integer(Word_In_Sentence_Num)-1) %>% 
  filter(!is.na(Word)) %>% 
  group_by(Story_Num) %>% 
  mutate("Word_In_Story_Num"=row_number()) %>% 
  left_join(words,by=c("Story_Num","Word","Word_In_Story_Num"))

```

# For Maze

```{r}
# takes sentences, writes into format to do maze (also other things)
for_maze <- read_delim("../Materials/natural_stories_sentences.tsv", delim="\t") %>%
  mutate(label=str_c(Story_Num, Sentence_Num, sep="_"), item=row_number()) %>% 
  select(label, item, Sentence) %>% 
   write_delim("ns_pre_maze.txt", delim=";", col_names=F, quote_escape = F)

```

# Parsing surprisals, etc

```{r grnn}

grnn <- read_delim("grnn_out_all.txt", delim="\t", col_names=c("model_token","grnn_surp")) %>% 
  filter(model_token!="<eos>") %>% 
  mutate(num=row_number())

grnn_translator <- read_delim("ns_tokens_trans.txt", delim="\t", col_names=c("Item","Word_In_Sentence_Num", "Word","token"), col_types=c(col_character(),col_integer(), col_character(),col_character())) %>% 
  mutate(num=row_number()) %>% 
  left_join(grnn, by="num") %>% 
  mutate(grnn_surp=ifelse(model_token==token & grnn_surp>0, grnn_surp, NA)) %>% 
  select(-model_token, -num) %>% 
  separate(Item, into=c("Story_Num","Sentence_Num"), convert=T) %>% 
  left_join(labelled, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word")) %>% 
  group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
  summarize(grnn_surp=sum(grnn_surp))

```

```{r ngram}
ngram <- read_delim("ngram_out_all.txt", skip=3,delim="\t", col_names=c("sentence_no","token_no","model_token","ngram_surp")) %>% 
  mutate(num=row_number())

ngram_translator <- read_delim("ns_lower_trans.txt", delim="\t", col_names=c("Item","Word_In_Sentence_Num", "Word","token"), col_types=c(col_character(),col_integer(), col_character(),col_character())) %>% 
  mutate(num=row_number()) %>% 
  left_join(ngram, by="num") %>% 
  mutate(ngram_surp=ifelse(model_token==token & ngram_surp>0, ngram_surp, NA)) %>% 
  select(-model_token, -num, -sentence_no, -token_no) %>% 
  separate(Item, into=c("Story_Num","Sentence_Num"), convert=T) %>% 
  left_join(labelled, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word")) %>% 
  group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
  summarize(ngram_surp=sum(ngram_surp))
```

```{r txl}
txl <- read_delim("txl_out_all.txt", delim="\t", col_names=c("model_token","txl_surp")) %>% 
  filter(model_token!="<eos>") %>% 
  mutate(num=row_number())

txl_translator <- read_delim("ns_tokens_moses_trans.txt", delim="\t", 
                             col_names=c("Item","Word_In_Sentence_Num", "Word","token"), 
                             col_types=c(col_character(),col_integer(), col_character(),col_character())) %>% 
  mutate(num=row_number()) %>% 
  left_join(txl, by="num") %>% 
  mutate(txl_surp=ifelse(model_token==token & txl_surp, txl_surp, NA)) %>% 
  select(-model_token, -num) %>% 
  separate(Item, into=c("Story_Num","Sentence_Num"), convert=T) %>% 
  left_join(labelled, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word")) %>% 
  group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
  summarize(txl_surp=sum(txl_surp))
```

```{r freq}

freq <- read_delim("ns_freq.txt", delim="\t",
                   col_names=c("Item","Word_In_Sentence_Num","Word","freq","length"), 
                   col_types=c(col_character(),col_integer(),col_character(),col_double(), col_integer())) %>% 
  mutate(freq=ifelse(freq>0, freq, NA)) %>% 
  separate(Item, into=c("Story_Num","Sentence_Num"), convert=T) %>% 
  left_join(labelled, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word"))

```

```{r conglomerate}
all <- txl_translator %>% 
  left_join(ngram_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
  left_join(grnn_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>%
  left_join(freq, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
  filter(Word_In_Sentence_Num!=0) %>% 
  ungroup() %>% 
  mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
         freq_center=freq-mean(freq, na.rm=T),
         length_center=length-mean(length, na.rm=T)) %>% 
  write_rds("natural_stories_surprisals.rds")

```

Note that word 0's of sentences will be removed b/c we definitely don't trust them

### 2020-04-14 Provo/provo_materials/prep.Rmd
Files that got deleted when they were moved in commit ea1e468

---
title: "Provo prep"
output: github_document
---

```{r}
library(tidyverse)
library(readr)
```
What we're trying to get out of this:
- List of words/sentences/paragraphs that are reconstructable to each other and labelled by word/sentence/item number.
- Surprisals, frequency, etc for each word

Read in a bunch of things.

Fundamentally want sentences which should match paragraphs, then we can check then line up and separate on spaces to make words. (will also need punct. stripped words for some purposes.)

# New attempt
Texts were copied from the predictability norms and then hand broken into sentences (because this is actually faster than anything I know how to do automatically). Some light editing, such as turning "and/or" into "or" and removing internal quote marks was done.
```{r}
sentences <- read_csv("edited_sentences.csv") %>% rename(Sentence=Text)

paragraphs <- sentences %>% 
  pivot_wider(names_from="Sentence_Number", values_from="Sentence") %>%
  unite("Text", `1`:`4`, sep=" ", na.rm=T)

nums <- 1:55 %>% as.character()

words <- sentences %>%
  separate(Sentence, into=nums, sep=fixed(" ")) %>%
  pivot_longer(cols=`1`:`55`,names_to = "Word_In_Sentence_Number", values_to="Word") %>% 
  filter(!is.na(Word)) %>% 
  left_join(sentences, by=c("Text_ID", "Sentence_Number")) %>% 
  left_join(paragraphs, by=c("Text_ID")) %>% 
  group_by(Text_ID) %>% 
  mutate(Word_In_Paragraph_Number=row_number()) %>% 
  type_convert() %>% 
  select(Text_ID, Sentence_Number, Word_In_Sentence_Number, Word_In_Paragraph_Number, Word, Sentence, Text) %>% 
  write_csv("words.csv")

```
## Get data
We want word frequency, surprisal, and length for each word.

For word frequency, want to take the words, and use word freq. Frequency is zipfian - log10 of number of occurances in 10^9 words.
```{r}

freq <- read_csv("freq.csv", col_names = c("Text_ID", "Sentence_Number", "Word_In_Sentence_Number","Word", "freq")) %>% 
  mutate(freq=ifelse(freq==0, NA, freq))

labelled <- words %>% left_join(freq)

```

For surprisal.
We tokenize in python using nltk. However, tokenize on words behaves differently than tokenize on sentences, so we have to hand fix two problems. (On sentences, "Mr." and "B." are tokenized as single characters, but if you tokenize by word it gets split b/c period is at end?).

This whole adventure means we can add up the surprisals for multi-token words.

```{r}

sents <- words %>% select(Text_ID, Sentence_Number, Sentence) %>% unique() %>% write_csv("sents.csv")

tokens <- read_csv("tokens2.csv", col_names=c("word","1","2")) %>% 
  bind_cols(freq) %>% 
  pivot_longer(cols="1":"2",names_to="position", values_to="token") %>% 
  filter(!is.na(token))

surp <- read_tsv("surp.txt", col_names=c("token_unk","surp")) %>%
  filter(token_unk!="<eos>") %>%
  bind_cols(tokens) %>% 
  mutate(surp=ifelse(surp==0, NA, surp)) %>% 
  mutate(surp=ifelse(token_unk=="<unk>", NA, surp)) %>% 
  select(surp,Word, Text_ID, Sentence_Number, Word_In_Sentence_Number, freq) %>% 
  group_by(Text_ID, Sentence_Number, Word_In_Sentence_Number, Word, freq) %>% 
  summarize(surp=sum(surp)) %>% 
  ungroup() %>% 
  mutate(word_len=str_length(Word)) %>% 
  select(Text_ID, Sentence_Number, Word_In_Sentence_Number, Word, freq, surp, word_len) %>%
  left_join(words, by=c("Text_ID", "Sentence_Number", "Word_In_Sentence_Number", "Word")) %>% 
    mutate(surp_cent=surp-mean(surp, na.rm=T),
           word_len_cent=word_len-mean(word_len,na.rm=T),
           freq_cent=freq-mean(freq, na.rm=T)) %>% #centered
  write_rds("surprisal.rds")

```
For both surprisal and frequency, values of 0 were converted to NAs because the sources use 0 as a unknown marker, so those values are untrustworthy and should not be included.

Both centered and non-centered versions of the variables are included because the non-centered are more interpretable for graphs (potentially), but the analyses want centered for interpretable intercepts.
## Maze prep & re-prep

```{r}

for_maze <- sents %>% ungroup() %>%  mutate(label=str_c(Text_ID, Sentence_Number, sep="_"), num=row_number()) %>% select(label, num, Sentence) %>% write_delim("provo_pre_maze.txt", delim=";", col_names=F, quote_escape = F)

```

Run them using  ./automate_new.py provo_pre_maze.txt provo_out.txt --model gulordava --freq wordfreq --format basic --num_to_test 100 --min_abs 25 --min_rel 15

Takes maze materials and writes them into paragraphs. 
```{r}
maze <- read_delim("provo_out.txt", delim=";", col_names = c("label", "number", "sentence", "distractors", "blah")) %>%
  select(-blah, -number) %>%
  separate(label, into=c("Text_ID", "Sentence_Number")) %>%
  pivot_wider(names_from=c("Sentence_Number"), values_from=c("sentence", "distractors"), values_fill=list(sentence="",distractors=""))%>% 
  unite("Text", sentence_1:sentence_4, sep=" ") %>% 
  unite("Distractor", distractors_1:distractors_4, sep=" ") %>%
  mutate(Text=trimws(Text), Distractor=trimws(Distractor))

write_lines(with(maze, sprintf('["critical","%s","%s","%s"],', Text_ID, Text, Distractor)), path="critical.txt")

```

### 2020-04-14 Provo/provo_materials/experiment.js

@@ -1,0 +1,151 @@
+var shuffleSequence = seq("code", "setcounter", "welcome", "intro", "intro-practice",  followEachWith("practice", "sep"), "end-practice",randomize(anyOf(startsWith("critical")),
+"instructions2", anyOf("questionnaire"),"topic","debriefing");
+
+var showProgressBar =false;
+
+var defaults = [
+		"Question", {
+				as: ["yes", "no"],
+        presentAsScale: false,
+        presentHorizontally: false,
+    },
+    "Maze", {redo: true},
+];
+
+var code = Math.floor(Math.random()*100000000);
+var sendingResultsMessage = "The results are now being transferred.  Please wait.";
+var completionMessage = "Thank you for your participation.  The results were successfully transmitted.  Your participation code is: " + code.toString();
+var completionErrorMessage = "The transmission of the results failed.  Please contact online_experiment@mit.edu and retry the transmission again by clicking the link.  Your participation code is: " + code.toString();
+
+var items = [
+    ["code", "DashedSentence", {s:code.toString(), mode:"speeded acceptability", wordTime:1}],
+	["setcounter", "__SetCounter__", { }],
+    ["welcome", "Message", {html:'<table width="100%"><tr><td valign="top" align="right">Department of Brain and Cognitive Sciences<br>Massachusetts Institute of Technology<br>77 Massachusetts Avenue<br>Cambridge, MA 02139-4307, USA</tr></table>\
+<h2>Thank you very much for your participation!</h2><p>This task is part of a MIT scientific research project. Your decision to complete this task is voluntary. There is no way for us to identify you. The only information we will have, in addition to your responses, is the time at which you completed the survey. The results of the research may be presented at scientific meetings or published in scientific journals. Clicking on the link below indicates that you are at least 18 years of age and agree to complete this task voluntarily.'}],
+    ["topic", "Form", {html:'Thank you for completing the experiment. Very briefly, what do you think this study is about?<br/><textarea name="topic" rows="3" cols="50" autofocus="true"></textarea>'}],
+    ["debriefing", "Message", {html:'<p>Thank you.  You will receive the participation code on the next page.</p>\n\n<p>Purpose of this study (feel free to skip): We’re generally interested in how the human brain processes language. The present study is testing out a new method for studying what types of sentence constructions are easier or harder to read. Your data will help us to answer these questions.</p>'}],
[…]

### 2020-04-17 Analysis/nat_stories.Rmd
more analysis, start of gams, lms

@@
 Given that, it seems like we don't need to exclude much data after a mistake. Two words seems safe. (This is totally arbitrary and hopeful no analysis is sensitive to it.)
+
+While we are planning on analysing data after mistakes, it might be interesting to know how much of the data is pre-mistake, and how many of the sentences are completed error-free.
+
+```{r before}
+
+data_non_error <- data_good %>% 
+  filter(is.correct==1) %>% 
+  filter(word_num>0)
+
+data_before <- data_good %>% 
+  mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
+  group_by(sentence, subject) %>% 
+  fill(word_num_mistake) %>% ungroup() %>% 
+  mutate(after_mistake=word_num-word_num_mistake,
+         after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
+  filter(after_mistake==0) %>% 
+  filter(is.correct==1) %>% 
+  filter(word_num>0)
+
+```
+
+If we only exclude the errors and the first words of sentences, there are `r nrow(data_non_error)` words.
+If we also exclude all words after mistakes, there are `r nrow(data_before)` words.
+We will exclude only a couple of words after each mistake, but we could also opt to analyse only the pre-error sections.
+
+```{r complete}
+
+data_sentence <- data_before <- data_good %>% 
+  mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
+  group_by(sentence, subject) %>% 
+  fill(word_num_mistake) %>% ungroup() %>% 
+  mutate(after_mistake=word_num-word_num_mistake,
+         after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
+  group_by(sentence, subject) %>% 
+  summarize(total_mistake=sum(after_mistake)) %>% 
+  mutate(sent.correct=ifelse(total_mistake==0,1,0))
+```
+
+Of the `r nrow(data_sentence)` sentences that good participants completed, `r filter(data_sentence, sent.correct==1) %>% nrow()` were completed entirely correctly.

### 2020-04-17 Analysis/nat_stories_lms.Rmd
more analysis, start of gams, lms

---
title: "Natural Stories LM analysis"
output: github_document
---
```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
library(mgcv)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

```{r participants}

data <- read_rds("../Data/cleaned.rds")

data_filt <- data %>% filter(native %in% c("ENG", "English", "ENGLISH", "english")) #I peeked at what people put that semantically maps to english

data_error_summ <- data_filt %>% 
  mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
  group_by(subject, critical, practice) %>%
  filter(type!="practice") %>% 
  filter(rt<5000) %>% 
  summarize(pct_correct=mean(correct.num)) %>% 
  ungroup() %>% 
  mutate(is.attentive=ifelse(pct_correct>.8, T,F)) %>% 
  select(subject, is.attentive)

data_low_error <- data_filt %>% 
  left_join(data_error_summ, by="subject") %>% 
  filter(is.attentive) %>% 
  filter(type!="practice")

data_error_free <- data_low_error %>% 
  mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
  group_by(sentence, subject) %>% fill(word_num_mistake) %>% ungroup() %>% 
  mutate(after_mistake=word_num-word_num_mistake,
         after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
  filter(correct=="yes") %>% 
  filter(!after_mistake %in% c(1,2))

data_no_first <- data_error_free %>% filter(word_num!=0)

data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  select(subject, word_num, word, rt, sentence, type)

data_stories <- data_ready %>% select(type, subject) %>% 
  unique() %>% 
  group_by(type) %>% 
  tally()
```

# Overview

`r data %>% select(subject) %>% unique() %>% nrow()` participants read naturalistic stories from the natural stories corpus. Each participant read 1 story.

We exclude

- participants who do not report English as a native language (`r data_filt %>% select(subject) %>% unique() %>% nrow()` remaining)
- participants who do not get 80% of the words correct (`r data_low_error %>% select(subject) %>% unique() %>% nrow()` remaining)
- practice items (`r data_low_error %>% nrow()` words remaining)
- words that were wrong or were within two after a mistake (`r data_error_free %>% nrow()` words remaining)
- the first word of every sentence (didn't have a real distractor, RT is measured slightly differently) (`r data_no_first %>% nrow()` words remaining)
- words with RTs <100 or >5000 (<100 we think is likely a recording error, or at least not reading the words at all, >5000 is likely getting distracted) (`r data_ready %>% nrow()` words remaining)

Within the filtered data, each story was read between `r min(data_stories$n)` and `r max(data_stories$n)` times, for an average of `r mean(data_stories$n)`.

We use as predictors:

Definition: "stripped word" refers to the word after start and end punctuation (commas, periods, quotes, parens, etc) have been removed.

- length in characters of stripped word
- unigram frequency of stripped word. Frequencies for words are calculated using word_tokenize on the gulordava train data and counting up instances. (This tends to tokenize off punctuation, but is capitalization sensitive). Frequencies are represented as log2 of the expected occurances in 1 billion words.

Surprisals are measured in bits TODO: double check with Jenn about these being in bits, model specs/cites!

- ngram (5-gram KN smoothed)
- GRNN
- Transformer-XL

All predictors are centered at 0, but they are not rescaled, so they are still interpretable.

Centering differences:

- txl -8.66
- ngram -10.32
- grnn -7.94
- freq -18.75
- length -4.49

```{r labels}
labs <- read_rds("../Prep_code/natural_stories_surprisals.rds") %>% left_join(read_delim("../Materials/natural_stories_sentences.tsv", delim="\t")) %>% 
  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())

past_word <- labs %>% 
  mutate(word_num=word_num+1) %>% 
  rename(past_txl=txl_center, past_ngram=ngram_center, past_grnn=grnn_center, past_freq=freq, past_length=length) %>% 
  select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
  left_join(labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))

labelled <- data_ready %>% left_join(past_word, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
  select(rt, subject, Word_In_Story_Num, Story_Num,txl_center,ngram_center,grnn_center, freq_center,length_center, starts_with("past")) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("../Data/for_modelling.Rds")

```

# Rough pass with lmer

Given past experience, lmer probably will choke on the full heirarchical effects structure, but we can do a first pass with lm anyway as a test.

I'm copying model formula from the syntax gym paper, so frequency and length interact but surprisal doesn't, and both current and past word predictors are considered.

```{r lm}

lm_ngram_model <- lm(rt ~ ngram_center + freq_center * length_center+ past_ngram + past_freq*past_length, data=labelled)

lm_txl_model <- lm(rt ~ txl_center + freq_center * length_center+ past_txl + past_freq*past_length, data=labelled)

lm_grnn_model <- lm(rt ~ grnn_center + freq_center * length_center+ past_grnn + past_freq*past_length, data=labelled)

```

```{r lm_summary}

summary(lm_ngram_model)

summary(lm_txl_model)

summary(lm_grnn_model)

```
<!--
# Now with heirarchy

Not yet run in most cases, will be computation-time consuming!

Questions:

 - should there be interaction terms
 - should there be previous word terms
 - how important is Word_ID intercept? (We have up to 8 observations per each of 10000ish)

We'll run these models again using brms with full mixed effects.

The mixed effects are

- everything gets to vary by subject
- intercept  by Word_ID (where word_id is different for every token in the texts)

Priors:

- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
- normal(0,500) for beta and sd -- we don't really know what effects are
- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way -->

```{r}
  priors <- c(
      set_prior("normal(1000, 1000)", class="Intercept"),
      set_prior("normal(0, 500)", class="b"),
      set_prior("normal(0, 500)", class="sd"),
      set_prior("lkj(1)",       class="cor"))

  # brm_ngram_model <- brm(rt ~ ngram_center * freq_center * length_center +
  #                          (ngram_center * freq_center * length_center|subject)+
  #                          (1|Word_ID), data=labelled, file="brm_ngram", prior=priors)
  # 
  # brm_txl_model <- brm(rt ~ txl_center * freq_center * length_center +
  #                          (txl_center * freq_center * length_center|subject)+
  #                          (1|Word_ID), data=labelled, file="brm_txl", prior=priors)
  # 
  # brm_grnn_model <- brm(rt ~ grnn_center * freq_center * length_center +
  #                          (grnn_center * freq_center * length_center|subject)+
  #                          (1|Word_ID), data=labelled, file="brm_grnn", prior=priors)

```

```{r}

# show_summary <- function(model){
#   intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
#   stats <- gather_draws(model, `b_.*`, regex=T) %>% mutate(above_0=ifelse(.value>0, 1,0)) %>% group_by(.variable) %>% summarize(pct_above_0=mean(above_0)) %>% mutate(pval_equiv = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% left_join(intervals, by=".variable") %>% select(term=.variable, estimate=.value, lower=.lower, upper=.upper, pval_equiv)
#   
#   stats
# }

```

```{r}
# brm_ngram_model <- brm(rt ~ ngram_center * freq_center * length_center +
#                          (ngram_center * freq_center * length_center|subject), data=labelled, file="brm_ngram", prior=priors)
# 
# summary(brm_ngram_model)
# 
# kable(show_summary(brm_ngram_model))
```

# GAMs

This is still using centered values, although we may want to switch back to non-centered eventually (translate by about 8-10 for surprisals).

```{r practice}

gam_ngram <- gam(rt ~ s(ngram_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)

gam_txl <- gam(rt ~ s(txl_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)

gam_grnn <- gam(rt ~ s(grnn_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=filter(labelled))

```

```{r}
 plot.gam(gam_ngram, select =1)
plot.gam(gam_ngram, select =3)

 plot.gam(gam_txl, select =1)
plot.gam(gam_txl, select =3)

 plot.gam(gam_grnn, select =1)
plot.gam(gam_grnn, select =3)

```

Some of these seem surprising, especially in relation to the lm output. One important check on the odd behavior for high surprisals is how reliable these might be, so I checked what the density of surprisals is and compared it with surprisals reported elsewhere (syntax gym reports up to 15 bits, Smith and Levy down to 10^-6, which is around 20 bits). Note that some of these high surprisals may come from adding together surprisals from multi-token word with punctuation (like quotemarks).

So, we see a linear pattern of same-word surprisal and little effect of previous word surprisal over the same interval (remember due to centering, that's looking at up to around 10 on these scales), but then weird things happen after that, when there is little data and it might be unreliable.

```{r density}

ggplot(labelled, aes(x=ngram_center))+geom_density()

ggplot(labelled, aes(x=txl_center))+geom_density()

ggplot(labelled, aes(x=grnn_center))+geom_density()

```

### 2020-04-22 Prep_code/nat_stories_prep.Rmd
label when words are multi-token

@@
   group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
-  summarize(grnn_surp=sum(grnn_surp))
+  summarize(grnn_surp=sum(grnn_surp),
+            grnn_token_count=n())
 
@@
   group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
-  summarize(ngram_surp=sum(ngram_surp))
+  summarize(ngram_surp=sum(ngram_surp),
+            ngram_token_count=n())
 ```
@@
   group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
-  summarize(txl_surp=sum(txl_surp))
+  summarize(txl_surp=sum(txl_surp),
+            txl_token_count=n())
 ```

### 2020-04-24 Analysis/nat_stories.Rmd
changes from adding abstract graphs

@@
 library(knitr)
+library(viridis)
 theme_set(theme_bw())
@@
        color="?s correct")
+
+```
+## Graphs for abstract
+```{r error_abstract}
+for_plot <- data_error_summ %>% 
+  mutate(num_correct=case_when(
+    critical_factor %in% c("1","2","3") ~ "3 or fewer",
+    T ~ critical_factor)) %>% 
+    mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct"))
+
+ggplot(for_plot, aes(x=pct_correct, y=mean_rt, color=accurate))+
+  geom_point(size=1)+
+  scale_color_viridis(discrete=T)+
+  labs(x="Fraction words selected correctly",
+       y="Mean reaction time",
+       color="?s correct")+
+  coord_cartesian(xlim=c(0.4,1), ylim=c(0,1600))+
+    scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
+  guides(color=FALSE)
+
+ggsave("../Papers/error.pdf", width=3, height=2, unit="in")
+
+some <- data_filt %>% select(subject, critical) %>% 
+  unique() %>% 
+  left_join(data_error_summ, by=c("subject", "critical")) %>% 
+  mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct")) %>% 
+  mutate(correct=critical*6) %>% 
+  group_by(correct,accurate)
+
+ggplot(some, aes(x=correct,fill=accurate))+
+  geom_bar(position="dodge")+
+  facet_grid(.~accurate)+
+  labs(x="Comp. questions correct (out of 6)", y="Participants")+
+  scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
+guides(fill=FALSE)
+
+ggsave("../Papers/comp.pdf", width=3, height=1.5, unit="in")
 

### 2020-04-24 Analysis/nat_stories_lms.Rmd
changes from adding abstract graphs

@@
 library(mgcv)
+library(mgcViz)
 theme_set(theme_bw())
@@
 
-past_word <- labs %>% 
-  mutate(word_num=word_num+1) %>% 
-  rename(past_txl=txl_center, past_ngram=ngram_center, past_grnn=grnn_center, past_freq=freq, past_length=length) %>% 
-  select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
-  left_join(labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))
+# past_word <- labs %>% 
+#   mutate(word_num=word_num+1) %>% 
+#   rename(past_txl=txl_center, past_ngram=ngram_center, past_grnn=grnn_center, past_freq=freq, past_length=length) %>% 
+#   select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
+#   left_join(labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))
 
-labelled <- data_ready %>% left_join(past_word, by=c("word_num", "word", "sentence")) %>%
-  filter(word_num>1) %>% 
-  select(rt, subject, Word_In_Story_Num, Story_Num,txl_center,ngram_center,grnn_center, freq_center,length_center, starts_with("past")) %>% 
-  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
-  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("../Data/for_modelling.Rds")
+# labelled <- data_ready %>% left_join(past_word, by=c("word_num", "word", "sentence")) %>%
+#   filter(word_num>1) %>% 
+#   select(rt, subject, Word_In_Story_Num, Story_Num,txl_center,ngram_center,grnn_center, freq_center,length_center, starts_with("past")) %>% 
+#   mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
+#   select(-Word_In_Story_Num, -Story_Num) %>% write_rds("../Data/for_modelling.Rds")
 
@@
 
-```{r lm}
-
-lm_ngram_model <- lm(rt ~ ngram_center + freq_center * length_center+ past_ngram + past_freq*past_length, data=labelled)
-
-lm_txl_model <- lm(rt ~ txl_center + freq_center * length_center+ past_txl + past_freq*past_length, data=labelled)
-
-lm_grnn_model <- lm(rt ~ grnn_center + freq_center * length_center+ past_grnn + past_freq*past_length, data=labelled)
-
-```
-
-```{r lm_summary}
-
-summary(lm_ngram_model)
-
-summary(lm_txl_model)
-
-summary(lm_grnn_model)
-
-```
 <!--
@@
 
-# show_summary <- function(model){
-#   intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
-#   stats <- gather_draws(model, `b_.*`, regex=T) %>% mutate(above_0=ifelse(.value>0, 1,0)) %>% group_by(.variable) %>% summarize(pct_above_0=mean(above_0)) %>% mutate(pval_equiv = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% left_join(intervals, by=".variable") %>% select(term=.variable, estimate=.value, lower=.lower, upper=.upper, pval_equiv)
-#   
-#   stats
-# }
+show_summary <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% mutate(above_0=ifelse(.value>0, 1,0)) %>% group_by(.variable) %>% summarize(pct_above_0=mean(above_0)) %>% mutate(pval_equiv = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% left_join(intervals, by=".variable") %>% select(term=.variable, estimate=.value, lower=.lower, upper=.upper, pval_equiv)
+
+  stats
+}
 
@@
 
-gam_ngram <- gam(rt ~ s(ngram_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)
-
-gam_txl <- gam(rt ~ s(txl_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)
-
-gam_grnn <- gam(rt ~ s(grnn_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=filter(labelled))
+# gam_ngram <- gam(rt ~ s(ngram_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)
+# 
+# gam_txl <- gam(rt ~ s(txl_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)
+# 
+# gam_grnn <- gam(rt ~ s(grnn_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=filter(labelled))
 
@@
 ```{r}
- plot.gam(gam_ngram, select =1)
-plot.gam(gam_ngram, select =3)
-
- plot.gam(gam_txl, select =1)
-plot.gam(gam_txl, select =3)
-
- plot.gam(gam_grnn, select =1)
-plot.gam(gam_grnn, select =3)
+#  plot.gam(gam_ngram, select =1)
+# plot.gam(gam_ngram, select =3)
+# 
+#  plot.gam(gam_txl, select =1)
+# plot.gam(gam_txl, select =3)
+# 
+#  plot.gam(gam_grnn, select =1)
+# plot.gam(gam_grnn, select =3)
 
@@
 ```{r density}
-
-ggplot(labelled, aes(x=ngram_center))+geom_density()
-
-ggplot(labelled, aes(x=txl_center))+geom_density()
-
-ggplot(labelled, aes(x=grnn_center))+geom_density()
+# 
+# ggplot(labelled, aes(x=ngram_center))+geom_density()
+# 
+# ggplot(labelled, aes(x=txl_center))+geom_density()
+# 
+# ggplot(labelled, aes(x=grnn_center))+geom_density()
 
 ```
+
+# Second approach, being more selective
+This time around we only include words that are single tokens in *all* the models and that have full data. We need the full data both for previous word and current word for the models.
+
+We try both on pre-error only data and on all data (modulo exclusions). This is all uncentered.
+
+```{r labs2}
+# A more selective approach
+
+select_labs <- labs %>% 
+  filter(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1) %>% 
+  filter(!is.na(txl_surp) & !is.na(ngram_surp) & !is.na(grnn_surp) & !is.na(freq) & !is.na(length)) %>% 
+    mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
+         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
+         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
+         freq_center=freq-mean(freq, na.rm=T),
+         length_center=length-mean(length, na.rm=T))
+
+past_word <- select_labs %>% 
+  mutate(word_num=word_num+1) %>% 
+  rename(past_txl=txl_surp, past_ngram=ngram_surp, past_grnn=grnn_surp, past_freq=freq, past_length=length,past_c_txl=txl_center, past_c_ngram=ngram_center, past_c_grnn=grnn_center, past_c_freq=freq_center, past_c_length=length_center) %>%
+  select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
+  left_join(select_labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))
+
+data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  filter(after_mistake==0) %>% 
+  select(subject, word_num, word, rt, sentence, type)
+
+data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  select(subject, word_num, word, rt, sentence, type)
+
+labelled_pre_error <- data_pre_error %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
+  filter(word_num>1) %>% 
+  select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
+         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
+  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
+  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("pre_error.rds")
+
+labelled_post_error <- data_ready %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
+  filter(word_num>1) %>% 
+select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
+         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
+  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
+  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("post_error.rds")
+
+```
+
+## Linear models
+
+Using pre-error data only.
+
+```{r lm-2}
+
+lm_ngram_model_1 <- lm(rt ~ ngram_surp + freq * length+ past_ngram + past_freq*past_length, data=labelled_pre_error)
+
+lm_txl_model_1 <- lm(rt ~ txl_surp + freq * length+ past_txl + past_freq*past_length, data=labelled_pre_error)
+
+lm_grnn_model_1 <- lm(rt ~ grnn_surp + freq * length + past_grnn + past_freq*past_length, data=labelled_pre_error)
+
+```
+
+```{r lm_summary-2}
+
+summary(lm_ngram_model_1)
+
+summary(lm_txl_model_1)
+
+summary(lm_grnn_model_1)
+
+```
+Including post-error data.
+
+```{r lm-3}
+
+lm_ngram_model_2 <- lm(rt ~ ngram_surp + freq * length+ past_ngram + past_freq*past_length, data=labelled_post_error)
+
+lm_txl_model_2 <- lm(rt ~ txl_surp + freq * length+ past_txl + past_freq*past_length, data=labelled_post_error)
+
+lm_grnn_model_2 <- lm(rt ~ grnn_surp + freq * length + past_grnn + past_freq*past_length, data=labelled_post_error)
+
+```
+
+```{r lm_summary-3}
+
+summary(lm_ngram_model_2)
+
+summary(lm_txl_model_2)
+
+summary(lm_grnn_model_2)
+
+```
+
+## GAMs
+
+Using only pre-error data.
+
+```{r gams-2}
+gam_ngram_1 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)
+
+gam_txl_1 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)
+
+gam_grnn_1 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)
+
+```
+
+```{r plot_gam-2}
+ plot.gam(gam_ngram_1, select =1)
+plot.gam(gam_ngram_1, select =3)
+
+ plot.gam(gam_txl_1, select =1)
+plot.gam(gam_txl_1, select =3)
+
+ plot.gam(gam_grnn_1, select =1)
+plot.gam(gam_grnn_1, select =3)
+
+```
+
+Including post-error data.
+
+```{r gams-3}
+gam_ngram_2 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)
+
+gam_txl_2 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)
+
+gam_grnn_2 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)
+
+```
+
+```{r plot_gam-3}
+ plot.gam(gam_ngram_2, select =1)
+plot.gam(gam_ngram_2, select =3)
+
+plot_smooths(model=gam_ngram_2,series=ngram_surp)
+plot_smooths(model=gam_ngram_2,series=past_ngram)
+
+ plot.gam(gam_txl_2, select =1)
+plot.gam(gam_txl_2, select =3)
+
+plot_smooths(model=gam_txl_2,series=txl_surp)
+plot_smooths(model=gam_txl_2,series=past_txl)
+
+ plot.gam(gam_grnn_2, select =1)
+plot.gam(gam_grnn_2, select =3)
+
+ v <- seq(0,25, by=1)
+a <- predict(gam_ngram_2, type="terms")
+
+b <- predict_gam(gam_ngram_2, exclude_terms="te(freq,length)",length_out=2)
+
+summary(gam_ngram_2)
+```
+
+This is much more reassuring. We do want to check with brms though.
+
+```{r}
+
+priors <- c(
+  set_prior("normal(1000, 1000)", class="Intercept"),
+  set_prior("normal(0, 500)", class="b"),
+  set_prior("normal(0, 500)", class="sd"),
+  set_prior("lkj(1)",       class="cor"))
+
+brm_ngram_model <- brm(rt ~ ngram_surp + freq * length + past_ngram + past_freq*past_length + 
+                         (ngram_surp + freq * length + past_ngram + past_freq*past_length|subject), data=labelled_pre_error, file="brm_ngram_pre", prior=priors)
+
+summary(brm_ngram_model)
+
+show_summary(brm_ngram_model)
+```

### 2020-04-24 Prep_code/nat_stories_prep.Rmd
changes from adding abstract graphs

@@
   ungroup() %>% 
-  mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
-         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
-         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
-         freq_center=freq-mean(freq, na.rm=T),
-         length_center=length-mean(length, na.rm=T)) %>% 
   write_rds("natural_stories_surprisals.rds")

### 2020-04-24 Papers/amlap2020_abstract.tex
Abstract draft

\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{helvet}
\pagestyle{empty}
\usepackage[kerning=true]{microtype}
\usepackage{parskip}
\usepackage{sansmath}
\usepackage[font={small, bf}]{caption}
\usepackage[font={small}]{subcaption}
\usepackage{graphicx}
\usepackage{multicol}
\setlength{\abovecaptionskip}{0pt}
\setlength{\floatsep}{10pt}
\setlength{\textfloatsep}{0pt}
\setlength{\intextsep}{0pt}
\setlength{\belowcaptionskip}{0pt}
\setlength{\parindent}{5ex}
\setlength{\parskip}{0pt}
% Feel free to use additional packages for glosses, figures, whatnot.

% The next bit is for reserving sufficient space for authors,
% affiliations, and e-mail address.  No need to change for initial
% anonymous version.  For the final version, replace the
% \toggletrue{anonymous} with \togglefalse{anonymous} to de-anonymize.
\usepackage{etoolbox}
\newtoggle{anonymous}
\toggletrue{anonymous}

\renewcommand{\title}[1]{\textbf{#1}\\}
\newcommand{\authors}[1]{\iftoggle{anonymous}{\phantom{#1}}{#1}\\}
\newcommand{\email}[1]{\iftoggle{anonymous}{\phantom{#1}}{#1}}

\begin{document}

% First page:

% Insert title, authors, affiliations, and e-mail address in the next three lines:

\title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
\authors{Veronica Boyce (MIT), Roger Levy (MIT)} %TODO, check what affiliation I have at conference time
\email{vboyce@mit.edu}

% Here goes the main text of your abstract:
%Top overview
A-maze is a promising new method for measuring incremental sentence processing (Boyce et al. 2020; Sloggett et al. 2020). However, so far A-maze has only been used on constructed psycholinguistic items, so it is unknown how fluently people can comprehend and remember what they read during the Maze task. %TODO do I need to motivate why people might not comprehend?
%TODO do I need to introduce what the Maze task is earlier?
We address this question by running A-maze on the stories from the Natural Stories Corpus (Futrell et al. 2017). We find that some participants can follow the stories and correctly answer comprehension questions even when reading with Maze. We also find a linear effect of a word's surprisal on reaction time, but little effect of the surprisal of the previous word, indicating a lack of spill-over effects.

% How does A-maze work 
In the Maze task, participants read a sentence word by word (see Figure~1). For each word position, they see two words, one of which is the next word in the sentence and one of which is a distractor. Participants press a key to indicate which word continues the sentence; the time between key presses (reaction time, or RT) is the dependent measure.  Usually in the Maze task, when a participant makes a mistake, the sentence stops and they move on to the next item. In order to present coherent texts, we instead have participants correct their mistakes (as shown in Figure~1). When a participant makes a mistake, they see an error message and must press the correct key to continue with the sentence. This way, participants see all the content of each sentence, and so can follow the story as it continues in the next sentence.

 The Natural Stories Corpus contains 10 passages, each about 1000 words long and 6 comprehension questions per passage (Futrell et al. 2017). We used the A-maze framework from Boyce et al. (2020) and the language model from Gulordava et al. (2017) to generate distractor words for the texts. We recruited 100 participants from Amazon Mechanical Turk; each participant did a short practice passage to become familiar with the Maze task, followed by reading one of the Natural Stories in the Maze task, and answering the corresponding comprehension questions.

 We excluded 5 participants who did not report being native English speakers. Some participants appear to be randomly pressing keys to get through the task quickly, while others seem to do the task more slowly and diligently (Figure~2). Task accuracy is correlated with performance on the comprehension questions ($r^2=.47$); of the 63 participants who had at least 80\% accuracy on the Maze task, 50 got 5 or 6 of the 6 comprehension questions correct (Figure~4). We exclude participants with less than 80\% accuracy from analyses.

We wanted to confirm whether a word's surprisal, unigram frequency, and length are predictive of RT. We get surprisal estimates from  3 models: a Kneser-Ney smoothed 5-gram model, a recurrent neural network model (GRNN) from Gulordava et al. (2018), and a transformer model (TXL) from Dai et al. (2019). We fit BRMS models using surprisal, frequency, length, and frequency:length interaction of the word and the previous word as predictors of RT.  We only include words that occurred before any mistakes in their sentence. We only analyse words that the models treat as single tokens.  Results are shown in Table 1; across all models, surprisal and length of the current word show sizeable effects, and properties of the previous word show smaller effects, suggesting a lack of spillover. To confirm the shape of relationship between surprisal and RT, we fit GAMs; as shown in Figure 3, the relationship between surprisal and RT is roughly linear at the current word, and there is little effect for the prior word's surprisal.  %TODO this result also holds? if we include RT's from after a mistake

%TODO am I allowed to use the word surprisal?
% Take away
Our results provide evidence that the A-maze task can be used effectively on naturalistic texts where comprehension is important. We present a variation on the Maze task, where participants correct their mistakes and continue reading, so they don't miss parts of the story. This also makes it easy to exclude participants who are not doing the task attentively. We find that attentive participants can generally follow the story they are reading, and we confirm that A-maze RTs show a strong relationship with the surprisal of the current, but not the past, word. We make our code and data available at [link to repo redacted for anonymity].

\newpage
\begin{figure}
	\begin{minipage}{.5\textwidth}
		\captionof{figure}{Maze with Error Correction}
		{\center\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=.8\textwidth]{maze_diagram.pdf}\\} 
		\begin{small}
			Participants see two words at a time and try to select the correct word.  When the participant makes a mistake, they must correct it to continue. Blue circles indicate selected words.

		\end{small}

	\end{minipage}
	~~~
	\begin{minipage}{.5\textwidth}	
				\captionof{figure}{Accuracy versus RT}
		{\center\includegraphics[width=\textwidth]{error.pdf}\\} 
		\begin{small}
			Correlation between participant's accuracy on the Maze task and RT. Participants with less than 80\% accuracy (in red) were excluded from analyses.

		\end{small}
	\end{minipage}
\end{figure}
\vspace{1em}
\begin{figure}
		\begin{minipage}{.5\textwidth}
		\captionof{figure}{Surprisal and RT\\
		}
		{\center\includegraphics[width=\textwidth]{gam.pdf}\\} 
		\begin{small}
			Current word surprisal is linearly predictive of RT, but previous word surprisal is not predictive of RT.

		\end{small}
	\end{minipage}	
	~~~
\begin{minipage}{.5\textwidth}
	\captionof{figure}{Comprehension question accuracy}
	{\center\includegraphics[width=\textwidth]{comp.pdf}\\} 
	\begin{small}
		Participants with higher accuracy on the Maze task also did better on comprehension questions.

	\end{small}
\end{minipage}

\end{figure}

	\setlength{\tabcolsep}{4pt}
		\captionof{table}{Regression Coefficients}
\begin{small}%TODO do I even include p-values??
%\centering
\begin{tabular}{l|rlr|rlr|rlr}
	\hline
	&\multicolumn{3}{c|}{5-gram}&\multicolumn{3}{c|}{GRNN}&\multicolumn{3}{c}{TXL}\\
		 & Est & CI & $p$ & Est & CI & $p$ &Est & CI & $p$ \\ 
		 \hline
Intercept & 867.7 & [829.6, 904.3] & 0.00 & 868.5 & [835.4, 903] & 0.00 & 866.5 & [830.4, 900.7] & 0.00 \\
Surprisal & 10.6 & [8.6, 12.6] & 0.00 & 22.1 & [19.8, 24.4] & 0.00 & 17.0 & [14.8, 19.2] & 0.00 \\
Frequency & -4.2 & [-7.3, -0.9] & 0.01 & 1.7 & [-1.2, 4.6] & 0.25 & -0.8 & [-3.7, 2.2] & 0.59 \\ 
Length & 20.1 & [15.3, 24.7] & 0.00 & 18.0 & [13.6, 22.4] & 0.00 & 20.8 & [16.4, 25.3] & 0.00 \\ 
Freq : Length & 1.1 & [0.4, 1.8] & 0.00 & 1.2 & [0.4, 1.9] & 0.00 & 1.2 & [0.5, 2] & 0.00 \\ 
\hline
Past Surprisal & 1.9 & [0.3, 3.4] & 0.02 & 1.9 & [0.6, 3.3] & 0.01 & 0.1 & [-1.2, 1.3] & 0.95 \\ 
Past Freq & 3.0 & [0.8, 5.2] & 0.01 & 1.5 & [-0.2, 3.3] & 0.09 & 0.9 & [-0.9, 2.8] & 0.35 \\ 
Past Length & -5.0 & [-8.7, -1.5] & 0.01 & -6.9 & [-10.5, -3.3] & 0.00 & -5.3 & [-9, -1.7] & 0.00 \\ 
Past Freq : Length& -1.0 & [-1.7, -0.3] & 0.01 & -1.3 & [-1.9, -0.6] & 0.00 & -1.2 & [-1.9, -0.5] & 0.00 \\ 
\hline
\end{tabular}
\vspace{.5em}

\end{small}
		\begin{small}
			Point estimates, credible intervals, and p-value equivalents. Models were run with centered predictors and full mixed effects by subject. Surprisal is in bits, frequency is in $log_2$ occurances per billion words. 
		\end{small}
\vspace{.5em}

	\begin{small}{
			\noindent\textbf{References:}
			Boyce et al (2020). \textit{J Mem. Lang.} $\bullet$
			Dai et al (2019) arXiv:1901.02860 $\bullet$
			Forster et al (2009). \textit{Behav. Res. Methods} $\bullet$
			Futrell et al (2017). arXiv:1708.05763 $\bullet$
			Gulordava et al (2018).  \textit{NAACL-HLT 2018} $\bullet$
			Sloggett et al (2020). \textit{CUNY 2020} $\bullet$
			Witzel et al (2012).  \textit{J Psycholinguist Res} 
			%TODO add cites for length, freq, surprisal justification
		}
	\end{small}

\end{document}

### 2020-04-24 Analysis/nat_stories.Rmd
minor changes for abstract

@@
 ```
+
+It appears that it is indeed possible to comprehend these stories while doing the maze task.
+There's clearly a correlation between accuracy on comprehension questions and accuracy on maze, although there is some variability.
+When the task is done well, average RT is between 600 and 1200 -- average RTs below 500 are unlikely to be attentive.
+
+Unclear what the correct thresholds for exclusions should be, although I'm tempted by an 80% accuracy cut-off (for RT stuff at least).
+If measures of interest are about comprehension, comprehension questions seem reasonable.
+
 ## Graphs for abstract
+
 ```{r error_abstract}
+
 for_plot <- data_error_summ %>% 
@@
 
+b <- cor(some$pct_correct, some$correct)
+
 ```
-
-It appears that it is indeed possible to comprehend these stories while doing the maze task.
-There's clearly a correlation between accuracy on comprehension questions and accuracy on maze, although there is some variability.
-When the task is done well, average RT is between 600 and 1200 -- average RTs below 500 are unlikely to be attentive.
-
-Unclear what the correct thresholds for exclusions should be, although I'm tempted by an 80% accuracy cut-off (for RT stuff at least).
-If measures of interest are about comprehension, comprehension questions seem reasonable.
 

### 2020-04-24 Analysis/nat_stories_lms.Rmd
minor changes for abstract

@@
 library(mgcViz)
+library(tidymv)
 theme_set(theme_bw())
@@
 
+data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  filter(after_mistake==0) %>% 
+  select(subject, word_num, word, rt, sentence, type)
+
 data_stories <- data_ready %>% select(type, subject) %>% 
@@
 
+We also do the analyses on only the words before mistakes (per sentence) (`r data_pre_error %>% nrow()` words)
+
+From the modelling side:
+(After attempts without doing this filtering) we only include words which are single token and known words in each of the models vocabularies.
+We also only include words with frequencies.
+This is roughly equivalent to excluding words with punctuation.
+
 We use as predictors:
 
-Definition: "stripped word" refers to the word after start and end punctuation (commas, periods, quotes, parens, etc) have been removed.
-
 - length in characters of stripped word
-- unigram frequency of stripped word.
+- unigram frequency of word.
 Frequencies for words are calculated using word_tokenize on the gulordava train data and counting up instances.
@@
 
-Surprisals are measured in bits TODO: double check with Jenn about these being in bits, model specs/cites!
+Surprisals are measured in bits (TODO: double check with Jenn about these being in bits, model specs/cites!)
 
@@
 
-All predictors are centered at 0, but they are not rescaled, so they are still interpretable.
+We sometimes center the predictors, but don't rescale them. (We center them for the brm models, but don't center surprisal for the GAM model).
 
-Centering differences:
-
-- txl -8.66
-- ngram -10.32
-- grnn -7.94
-- freq -18.75
-- length -4.49
+The model formula we use is rt ~ surp + length * freq + past_surp + past_length * past_freq, so we're allowing frequency and length to interact and including the same terms for the current and past word.
 
@@
   select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())
-
-# past_word <- labs %>% 
-#   mutate(word_num=word_num+1) %>% 
-#   rename(past_txl=txl_center, past_ngram=ngram_center, past_grnn=grnn_center, past_freq=freq, past_length=length) %>% 
-#   select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
-#   left_join(labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))
-
-# labelled <- data_ready %>% left_join(past_word, by=c("word_num", "word", "sentence")) %>%
-#   filter(word_num>1) %>% 
-#   select(rt, subject, Word_In_Story_Num, Story_Num,txl_center,ngram_center,grnn_center, freq_center,length_center, starts_with("past")) %>% 
-#   mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
-#   select(-Word_In_Story_Num, -Story_Num) %>% write_rds("../Data/for_modelling.Rds")
-
-```
-
-# Rough pass with lmer
-
-Given past experience, lmer probably will choke on the full heirarchical effects structure, but we can do a first pass with lm anyway as a test.
-
-I'm copying model formula from the syntax gym paper, so frequency and length interact but surprisal doesn't, and both current and past word predictors are considered.
-
-<!--
-# Now with heirarchy
-
-Not yet run in most cases, will be computation-time consuming!
-
-Questions:
-
- - should there be interaction terms
- - should there be previous word terms
- - how important is Word_ID intercept? (We have up to 8 observations per each of 10000ish)
-
-We'll run these models again using brms with full mixed effects.
-
-The mixed effects are
-
-- everything gets to vary by subject
-- intercept  by Word_ID (where word_id is different for every token in the texts)
-
-Priors:
-
-- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
-- normal(0,500) for beta and sd -- we don't really know what effects are
-- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way -->
-
-```{r}
-  priors <- c(
-      set_prior("normal(1000, 1000)", class="Intercept"),
-      set_prior("normal(0, 500)", class="b"),
-      set_prior("normal(0, 500)", class="sd"),
-      set_prior("lkj(1)",       class="cor"))
-
-  # brm_ngram_model <- brm(rt ~ ngram_center * freq_center * length_center +
-  #                          (ngram_center * freq_center * length_center|subject)+
-  #                          (1|Word_ID), data=labelled, file="brm_ngram", prior=priors)
-  # 
-  # brm_txl_model <- brm(rt ~ txl_center * freq_center * length_center +
-  #                          (txl_center * freq_center * length_center|subject)+
-  #                          (1|Word_ID), data=labelled, file="brm_txl", prior=priors)
-  # 
-  # brm_grnn_model <- brm(rt ~ grnn_center * freq_center * length_center +
-  #                          (grnn_center * freq_center * length_center|subject)+
-  #                          (1|Word_ID), data=labelled, file="brm_grnn", prior=priors)
-
-```
-
-```{r}
-
-show_summary <- function(model){
-  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
-  stats <- gather_draws(model, `b_.*`, regex=T) %>% mutate(above_0=ifelse(.value>0, 1,0)) %>% group_by(.variable) %>% summarize(pct_above_0=mean(above_0)) %>% mutate(pval_equiv = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% left_join(intervals, by=".variable") %>% select(term=.variable, estimate=.value, lower=.lower, upper=.upper, pval_equiv)
-
-  stats
-}
-
-```
-
-```{r}
-# brm_ngram_model <- brm(rt ~ ngram_center * freq_center * length_center +
-#                          (ngram_center * freq_center * length_center|subject), data=labelled, file="brm_ngram", prior=priors)
-# 
-# summary(brm_ngram_model)
-# 
-# kable(show_summary(brm_ngram_model))
-```
-
-# GAMs
-
-This is still using centered values, although we may want to switch back to non-centered eventually (translate by about 8-10 for surprisals).
-
-```{r practice}
-
-# gam_ngram <- gam(rt ~ s(ngram_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)
-# 
-# gam_txl <- gam(rt ~ s(txl_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)
-# 
-# gam_grnn <- gam(rt ~ s(grnn_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=filter(labelled))
-
-```
-
-```{r}
-#  plot.gam(gam_ngram, select =1)
-# plot.gam(gam_ngram, select =3)
-# 
-#  plot.gam(gam_txl, select =1)
-# plot.gam(gam_txl, select =3)
-# 
-#  plot.gam(gam_grnn, select =1)
-# plot.gam(gam_grnn, select =3)
-
-```
-
-Some of these seem surprising, especially in relation to the lm output.
-One important check on the odd behavior for high surprisals is how reliable these might be, so I checked what the density of surprisals is and compared it with surprisals reported elsewhere (syntax gym reports up to 15 bits, Smith and Levy down to 10^-6, which is around 20 bits).
-Note that some of these high surprisals may come from adding together surprisals from multi-token word with punctuation (like quotemarks).
-
-So, we see a linear pattern of same-word surprisal and little effect of previous word surprisal over the same interval (remember due to centering, that's looking at up to around 10 on these scales), but then weird things happen after that, when there is little data and it might be unreliable.
-
-```{r density}
-# 
-# ggplot(labelled, aes(x=ngram_center))+geom_density()
-# 
-# ggplot(labelled, aes(x=txl_center))+geom_density()
-# 
-# ggplot(labelled, aes(x=grnn_center))+geom_density()
-
-```
-
-# Second approach, being more selective
-This time around we only include words that are single tokens in *all* the models and that have full data. We need the full data both for previous word and current word for the models.
-
-We try both on pre-error only data and on all data (modulo exclusions). This is all uncentered.
-
-```{r labs2}
-# A more selective approach
 
@@
   left_join(select_labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))
-
-data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
-  filter(after_mistake==0) %>% 
-  select(subject, word_num, word, rt, sentence, type)
-
-data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
-  select(subject, word_num, word, rt, sentence, type)
 
@@
 
+These use centered predictors.
+
 Using pre-error data only.
@@
 
-lm_ngram_model_1 <- lm(rt ~ ngram_surp + freq * length+ past_ngram + past_freq*past_length, data=labelled_pre_error)
+lm_ngram_model_1 <- lm(rt ~ ngram_center + freq_center * length_center+ past_c_ngram + past_c_freq*past_c_length, data=labelled_pre_error)
 
-lm_txl_model_1 <- lm(rt ~ txl_surp + freq * length+ past_txl + past_freq*past_length, data=labelled_pre_error)
+lm_txl_model_1 <- lm(rt ~ txl_center + freq_center * length_center+ past_c_txl + past_c_freq*past_c_length, data=labelled_pre_error)
 
-lm_grnn_model_1 <- lm(rt ~ grnn_surp + freq * length + past_grnn + past_freq*past_length, data=labelled_pre_error)
+lm_grnn_model_1 <- lm(rt ~ grnn_center + freq_center * length_center+ past_c_grnn + past_c_freq*past_c_length, data=labelled_pre_error)
 
@@
 
-lm_ngram_model_2 <- lm(rt ~ ngram_surp + freq * length+ past_ngram + past_freq*past_length, data=labelled_post_error)
+lm_ngram_model_2 <- lm(rt ~ ngram_center + freq_center * length_center+ past_c_ngram + past_c_freq*past_c_length, data=labelled_post_error)
 
-lm_txl_model_2 <- lm(rt ~ txl_surp + freq * length+ past_txl + past_freq*past_length, data=labelled_post_error)
+lm_txl_model_2 <- lm(rt ~ txl_center + freq_center * length_center+ past_c_txl + past_c_freq*past_c_length, data=labelled_post_error)
 
-lm_grnn_model_2 <- lm(rt ~ grnn_surp + freq * length + past_grnn + past_freq*past_length, data=labelled_post_error)
+lm_grnn_model_2 <- lm(rt ~ grnn_center + freq_center * length_center+ past_c_grnn + past_c_freq*past_c_length, data=labelled_post_error)
 
@@
 
+## BRM models
+
+We include a by-subject effect for everything.
+Technically speaking, we do have repeated measures at the word level, but there are 10K or so words and at most a handful of copies of each word, so that's likely to get us not very much at the cost of making the model rather intractable (and so I'm not going to deal with it, at least not right now).
+
+Priors:
+
+- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
+- normal(0,500) for beta and sd -- we don't really know what effects are
+- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way
+
+```{r}
+  priors <- c(
+      set_prior("normal(1000, 1000)", class="Intercept"),
+      set_prior("normal(0, 500)", class="b"),
+      set_prior("normal(0, 500)", class="sd"),
+      set_prior("lkj(1)",       class="cor"))
+
+ brm_ngram_model <- brm(rt ~ ngram_center + freq_center * length_center +
+                         past_c_ngram + past_c_freq*past_c_length +
+                         (ngram_center + freq_center * length_center +
+                            past_c_ngram + past_c_freq*past_c_length|subject), data=labelled_pre_error,
+                       file="brm_ngram_pre_center", prior=priors, control=list(adapt_delta=.95))
+
+brm_grnn_model <- brm(rt ~ grnn_center + freq_center * length_center +
+                         past_c_grnn + past_c_freq*past_c_length +
+                         (grnn_center + freq_center * length_center +
+                            past_c_grnn + past_c_freq*past_c_length|subject), data=labelled_pre_error,
+                       file="brm_grnn_pre_center", prior=priors, control=list(adapt_delta=.95))
+
+brm_txl_model <- brm(rt ~ txl_center + freq_center * length_center +
+                         past_c_txl + past_c_freq*past_c_length +
+                         (txl_center + freq_center * length_center +
+                            past_c_txl + past_c_freq*past_c_length|subject), data=labelled_pre_error,
+                       file="brm_txl_pre_center", prior=priors, control=list(adapt_delta=.95))
+
+```
+
+```{r}
+
+show_summary <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=1),
+           upper=round(.upper, digits=1),
+           E=round(.value, digits=1),
+           `CI`=str_c("[",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           ) %>% 
+    select(Term, `E`, `CI`,`P`)
+
+  stats
+}
+```
+```{r}
+summary(brm_txl_model)
+summary(brm_grnn_model)
+summary(brm_ngram_model)
+```
+
+```{r}
+a <- show_summary(brm_txl_model) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
+b <- show_summary(brm_grnn_model) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
+c <- show_summary(brm_ngram_model) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
+
+summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`E`,`CI`,`P`)) %>% 
+  select(Term, `E_5-gram`, `CI_5-gram`, `P_5-gram`, `E_GRNN`, CI_GRNN, P_GRNN, E_TXL, CI_TXL, P_TXL)
+
+library(xtable)
+print(xtable(summ,digits=c(0,0,1,0,2,1,0,2,1,0,2)), include.rownames=FALSE)
+```
+
 ## GAMs
+
+I think there might be ways to have mixed effects for this, but for now we're not.
+Also, couldn't figure out good ways of getting graphs from the full models, so just fitting the pieces within ggplot for graphs.
 
@@
 
-```{r plot_gam-2}
- plot.gam(gam_ngram_1, select =1)
-plot.gam(gam_ngram_1, select =3)
+```{r}
+for_graph <- labelled_pre_error %>% 
+  select(rt, txl_surp, ngram_surp, grnn_surp, past_txl, past_grnn, past_ngram) %>%
+  pivot_longer(cols=`txl_surp`:`past_ngram`,names_to="label", values_to="data") %>% 
+  mutate(type=case_when(
+         label %in% c("past_ngram", "past_txl", "past_grnn")~"Previous",
+         T~"Current"),
+         model=case_when(
+           label %in% c("txl_surp", "past_txl")~"TXL",
+           label %in% c("grnn_surp", "past_grnn")~"GRNN",
+           T ~ "5-gram"
+         ))
 
- plot.gam(gam_txl_1, select =1)
-plot.gam(gam_txl_1, select =3)
+ggplot(for_graph, aes(x=data, y=rt))+
+         geom_smooth(method="gam",formula= y~s(x, bs="cr", k=20),color="black")+
+         facet_grid(type~model)+
+        labs(x="Surprisal (bits)", y="Reaction Time (ms)")
 
- plot.gam(gam_grnn_1, select =1)
-plot.gam(gam_grnn_1, select =3)
+ggsave("../Papers/gam.pdf", width=3.5, height=2, unit="in")
 
@@
 ```{r plot_gam-3}
- plot.gam(gam_ngram_2, select =1)
-plot.gam(gam_ngram_2, select =3)
+for_graph_2 <- labelled_post_error %>% 
+  select(rt, txl_surp, ngram_surp, grnn_surp, past_txl, past_grnn, past_ngram) %>%
+  pivot_longer(cols=`txl_surp`:`past_ngram`,names_to="label", values_to="data") %>% 
+  mutate(type=case_when(
+         label %in% c("past_ngram", "past_txl", "past_grnn")~"Previous",
+         T~"Current"),
+         model=case_when(
+           label %in% c("txl_surp", "past_txl")~"TXL",
+           label %in% c("grnn_surp", "past_grnn")~"GRNN",
+           T ~ "5-gram"
+         ))
 
-plot_smooths(model=gam_ngram_2,series=ngram_surp)
-plot_smooths(model=gam_ngram_2,series=past_ngram)
+ggplot(for_graph_2, aes(x=data, y=rt))+
+         geom_smooth(method="gam",formula= y~s(x, bs="cr", k=20),color="black")+
+         facet_grid(type~model)+
+        labs(x="Surprisal (bits)", y="Reaction Time (ms)")
 
- plot.gam(gam_txl_2, select =1)
-plot.gam(gam_txl_2, select =3)
-
-plot_smooths(model=gam_txl_2,series=txl_surp)
-plot_smooths(model=gam_txl_2,series=past_txl)
-
- plot.gam(gam_grnn_2, select =1)
-plot.gam(gam_grnn_2, select =3)
-
- v <- seq(0,25, by=1)
-a <- predict(gam_ngram_2, type="terms")
-
-b <- predict_gam(gam_ngram_2, exclude_terms="te(freq,length)",length_out=2)
-
-summary(gam_ngram_2)
 ```
-
-This is much more reassuring. We do want to check with brms though.
-
-```{r}
-
-priors <- c(
-  set_prior("normal(1000, 1000)", class="Intercept"),
-  set_prior("normal(0, 500)", class="b"),
-  set_prior("normal(0, 500)", class="sd"),
-  set_prior("lkj(1)",       class="cor"))
-
-brm_ngram_model <- brm(rt ~ ngram_surp + freq * length + past_ngram + past_freq*past_length + 
-                         (ngram_surp + freq * length + past_ngram + past_freq*past_length|subject), data=labelled_pre_error, file="brm_ngram_pre", prior=priors)
-
-summary(brm_ngram_model)
-
-show_summary(brm_ngram_model)
-```

### 2020-04-30 Papers/amlap2020_abstract.tex
revisions to abstract

@@
 \email{vboyce@mit.edu}
+\newline
+%
+% Here goes the main text of your abstract:
+%Comment: I feel like you could make a bigger deal of the current-word previous-word / RT correlation.
+Given how there are spillover effects for SPR and also for eye-tracking (if I’m not mistaken), it’s crazy that the relationship between RT/Surprisal for the previous word is basically flat, when it comes to Maze!
+%Comment: I’m confused with Table 1: Doesn’t length have a bigger effect size than surprisal?
+On a related note: does it matter which has a bigger effect?
+Why is it important to say that surprisal has a bigger effect, compared to just a robust, significant effect?
 
-% Here goes the main text of your abstract:
-%Top overview
-A-maze is a promising new method for measuring incremental sentence processing (Boyce et al.
+A-maze is a new method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities (Boyce et al.
 2020; Sloggett et al.
 2020).
-However, so far A-maze has only been used on constructed psycholinguistic items, so it is unknown how fluently people can comprehend and remember what they read during the Maze task.
-%TODO do I need to motivate why people might not comprehend?
-%TODO do I need to introduce what the Maze task is earlier?
-We address this question by running A-maze on the stories from the Natural Stories Corpus (Futrell et al.
-2017).
-We find that some participants can follow the stories and correctly answer comprehension questions even when reading with Maze.
-We also find a linear effect of a word's surprisal on reaction time, but little effect of the surprisal of the previous word, indicating a lack of spill-over effects.
+We test A-maze on the Natural Stories corpus (Futrell et al.
+2017) and find that people can comprehend what they read during the Maze task.
+Moreover, the Maze task yields useable reaction time data with word predictability effects that are linear in the surprisal of the current word, with little spillover effect from the surprisal of the previous word.
 
-% How does A-maze work 
-In the Maze task, participants read a sentence word by word (see Figure~1).
-For each word position, they see two words, one of which is the next word in the sentence and one of which is a distractor.
+The Maze task (Forster et al.
+2009; Witzel et al.
+2012) is an incremental processing method where participants read a sentence word by word  (Figure~1).
+For each word position, participants see two words, one of which is the next word in the sentence and one of which is a distractor.
 Participants press a key to indicate which word continues the sentence; the time between key presses (reaction time, or RT) is the dependent measure.
-Usually in the Maze task, when a participant makes a mistake, the sentence stops and they move on to the next item.
-In order to present coherent texts, we instead have participants correct their mistakes (as shown in Figure~1).
+Traditionally, when a participant makes a mistake, the sentence stops and they move on to the next item.
+In order to present coherent texts, we instead have participants correct their mistakes (Figure~1).
 When a participant makes a mistake, they see an error message and must press the correct key to continue with the sentence.
-This way, participants see all the content of each sentence, and so can follow the story as it continues in the next sentence.
+This way, participants see all the content and can follow the story, allowing us to test multi-sentence materials.
 
- The Natural Stories Corpus contains 10 passages, each about 1000 words long and 6 comprehension questions per passage (Futrell et al.
+The Natural Stories Corpus consists of 10 passages ($\sim$1000 words each) with 6 comprehension questions per passage (Futrell et al.
 2017).
@@
 (2020) and the language model from Gulordava et al.
-(2017) to generate distractor words for the texts.
-We recruited 100 participants from Amazon Mechanical Turk; each participant did a short practice passage to become familiar with the Maze task, followed by reading one of the Natural Stories in the Maze task, and answering the corresponding comprehension questions.
+(2018) to generate distractor words for the texts.
+We recruited 100 participants from Amazon Mechanical Turk; each participant read one passage in the Maze task and answered the corresponding comprehension questions.
 
- We excluded 5 participants who did not report being native English speakers.
-Some participants appear to be randomly pressing keys to get through the task quickly, while others seem to do the task more slowly and diligently (Figure~2).
+We exclude 5 participants who do not report being native English speakers.
+Participants clustered into those who appear to randomly press keys to get through the task quickly and those who do the task more slowly and accurately (Figure~2); we exclude participants with less than 80\% accuracy from analyses.
 Task accuracy is correlated with performance on the comprehension questions ($r^2=.47$); of the 63 participants who had at least 80\% accuracy on the Maze task, 50 got 5 or 6 of the 6 comprehension questions correct (Figure~4).
-We exclude participants with less than 80\% accuracy from analyses.
+People can understand and remember what they read while doing the Maze task successfully.
 
-We wanted to confirm whether a word's surprisal, unigram frequency, and length are predictive of RT.
-We get surprisal estimates from  3 models: a Kneser-Ney smoothed 5-gram model, a recurrent neural network model (GRNN) from Gulordava et al.
+We also investigated the relationship between a word's predictability and its RT.
+We fit Bayesian regression models using suprisal, frequency, length, and surprisal:length and frequency:length interactions at both the current and previous word as predictors of RT (see Table~1 for details).
+We use 3 models for surprisals: a Kneser-Ney smoothed 5-gram model, a recurrent neural network model (GRNN) from Gulordava et al.
 (2018), and a transformer model (TXL) from Dai et al.
 (2019).
-We fit BRMS models using surprisal, frequency, length, and frequency:length interaction of the word and the previous word as predictors of RT.
-We only include words that occurred before any mistakes in their sentence.
-We only analyse words that the models treat as single tokens.
-Results are shown in Table 1; across all models, surprisal and length of the current word show sizeable effects, and properties of the previous word show smaller effects, suggesting a lack of spillover.
-To confirm the shape of relationship between surprisal and RT, we fit GAMs; as shown in Figure 3, the relationship between surprisal and RT is roughly linear at the current word, and there is little effect for the prior word's surprisal.
-%TODO this result also holds?
-if we include RT's from after a mistake
+Across all models, we see large effects of current word surprisal and length, which noticeably exceed effects of previous word predictors (Table~1).
+We also fit GAMs to look at the shape of the relationship between surprisal and RT; we find a linear relationship between RT and current word surprisal and no relationship with previous word surprisal (Figure~3).
+The linear relationship matches known effects from eye-tracking and self-paced reading (Smith \& Levy 2013), but the lack of spillover indicates a potential advantage of Maze over eye-tracking and self-paced reading.
 
-%TODO am I allowed to use the word surprisal?
-% Take away
-Our results provide evidence that the A-maze task can be used effectively on naturalistic texts where comprehension is important.
-We present a variation on the Maze task, where participants correct their mistakes and continue reading, so they don't miss parts of the story.
-This also makes it easy to exclude participants who are not doing the task attentively.
-We find that attentive participants can generally follow the story they are reading, and we confirm that A-maze RTs show a strong relationship with the surprisal of the current, but not the past, word.
-We make our code and data available at [link to repo redacted for anonymity].
+In summary, we provide an adaptation of the Maze task for long naturalistic texts, and show that participants can comprehend what they read during A-maze.
+Additionally, we find robust, localized surprisal effects, supporting Forster et al.'s (2009) argument that Maze forces highly incremental processing.
+Overall, this suggests that A-maze is a versatile alternative to eye-tracking and self-paced reading.
+
+Our code and data are available at [link to repo redacted for anonymity].
+
+%It’s worth thinking about “why should AMLaP attendees care about this work?” and building an answer to that into the abstract.  Minimally, the work shows
+%
+%- Maze is really a flexible method
+%- you can have people read naturalistic text passages with Maze, they can understand the passages and you wind up with usable data
+%- word predictability effects in Maze are linear in surprisal, as they are with SPR and eye tracking
+%- potentially unlike SPR or eye tracking, surprisal seems to be far more important than word length or word frequency as a determinant of processing time in Maze (see comments/questions below about the model)
+%- also potentially unlike SPR or eye tracking, there isn’t much spillover in Maze for word predictability effects (though there is a little for word length).
+All this seems to support Forster et al.
+2009’s argument that Maze “forces highly incremental processing”.
 
@@
 \end{figure}
-\vspace{1em}
 \begin{figure}
@@
 	\begin{small}
-		Participants with higher accuracy on the Maze task also did better on comprehension questions.
+		Participants with high accuracy on the Maze task also perform well on comprehension questions.
 
@@
 
-	\setlength{\tabcolsep}{4pt}
+	\setlength{\tabcolsep}{3pt}
 		\captionof{table}{Regression Coefficients}
@@
 		 \hline
-Intercept & 867.7 & [829.6, 904.3] & 0.00 & 868.5 & [835.4, 903] & 0.00 & 866.5 & [830.4, 900.7] & 0.00 \\
-Surprisal & 10.6 & [8.6, 12.6] & 0.00 & 22.1 & [19.8, 24.4] & 0.00 & 17.0 & [14.8, 19.2] & 0.00 \\
-Frequency & -4.2 & [-7.3, -0.9] & 0.01 & 1.7 & [-1.2, 4.6] & 0.25 & -0.8 & [-3.7, 2.2] & 0.59 \\ 
-Length & 20.1 & [15.3, 24.7] & 0.00 & 18.0 & [13.6, 22.4] & 0.00 & 20.8 & [16.4, 25.3] & 0.00 \\ 
-Freq : Length & 1.1 & [0.4, 1.8] & 0.00 & 1.2 & [0.4, 1.9] & 0.00 & 1.2 & [0.5, 2] & 0.00 \\ 
+Intercept & 865.3 & [829.9, 902.9] & 0.00 & 871.1 & [837.9, 905.3] & 0.00 & 870.8 & [832.5, 907.8] & 0.00 \\ 
+Surprisal & 11.7 & [9.3, 14.1] & 0.00 & 23.7 & [21, 26.5] & 0.00 & 18.5 & [16.1, 21.1] & 0.00 \\ 
+Frequency & -2.9 & [-6.3, 0.5] & 0.10 & 2.9 & [-0.2, 6] & 0.06 & 0.4 & [-2.7, 3.5] & 0.79 \\ 
+		Length & 20.5 & [15.4, 25.6] & 0.00 & 18.5 & [13.3, 23.7] & 0.00 & 21.4 & [16.2, 26.6] & 0.00 \\ 
+		Surprisal:Length & -2.0 & [-3, -1] & 0.00 & -1.8 & [-2.7, -0.9] & 0.00 & -1.4 & [-2.2, -0.6] & 0.00 \\
+		Freq:Length & -1.0 & [-2.5, 0.4] & 0.16 & -0.1 & [-1.2, 1] & 0.82 & 0.2 & [-0.9, 1.2] & 0.76 \\ 
 		\hline
-Past Surprisal & 1.9 & [0.3, 3.4] & 0.02 & 1.9 & [0.6, 3.3] & 0.01 & 0.1 & [-1.2, 1.3] & 0.95 \\ 
-Past Freq & 3.0 & [0.8, 5.2] & 0.01 & 1.5 & [-0.2, 3.3] & 0.09 & 0.9 & [-0.9, 2.8] & 0.35 \\ 
-Past Length & -5.0 & [-8.7, -1.5] & 0.01 & -6.9 & [-10.5, -3.3] & 0.00 & -5.3 & [-9, -1.7] & 0.00 \\ 
-Past Freq : Length& -1.0 & [-1.7, -0.3] & 0.01 & -1.3 & [-1.9, -0.6] & 0.00 & -1.2 & [-1.9, -0.5] & 0.00 \\ 
+			Past Surprisal & 1.6 & [-0.5, 3.6] & 0.14 & 2.7 & [0.8, 4.5] & 0.00 & 0.8 & [-0.9, 2.5] & 0.40 \\ 
+		Past Freq & 2.6 & [-0.1, 5.4] & 0.06 & 1.9 & [-0.2, 4.2] & 0.08 & 1.2 & [-1.1, 3.6] & 0.30 \\ 
+		Past Length & -4.8 & [-9, -0.1] & 0.04 & -6.6 & [-10.9, -2.1] & 0.00 & -5.2 & [-9.3, -0.7] & 0.03 \\ 
+				Past Surp:Length & -0.2 & [-1.2, 0.8] & 0.72 & -0.9 & [-1.7, -0.2] & 0.01 & -0.6 & [-1.3, 0.2] & 0.13 \\ 
+		Past Freq:Length & -1.0 & [-2.3, 0.3] & 0.15 & -1.8 & [-2.9, -0.8] & 0.00 & -1.5 & [-2.6, -0.5] & 0.01 \\
+
 \hline
 \end{tabular}
-\vspace{.5em}
+\vspace{.1em}
 
@@
 			Point estimates, credible intervals, and p-value equivalents.
-Models were run with centered predictors and full mixed effects by subject.
-Surprisal is in bits, frequency is in $log_2$ occurances per billion words.
+Surprisal was measured in bits, frequency in $log_2$ occurrences per billion words, and length in characters.
+All predictors were centered, and only single token words were included.
+Models were fit in BRMS with formula \textit{rt $\sim$ surprisal*length + freq*length + past\_surp*past\_length + past\_freq*past\_length +
+			(surprisal*length + freq*length + past\_surp*past\_length + past\_freq*past\_length | participant)+(1|word\_id)}. Results shown for regression on pre-error data only.
 		\end{small}
-\vspace{.5em}
 
+\vspace{.1em}
 	\begin{small}{
@@
 			Sloggett et al (2020). \textit{CUNY 2020} $\bullet$
+			Smith \& Levy (2013). \textit{Cognition} $\bullet$
 			Witzel et al (2012).  \textit{J Psycholinguist Res} 

### 2020-04-30 Analysis/nat_stories.Rmd
minor changes

@@
 library(viridis)
+library(cowplot)
 theme_set(theme_bw())
@@
 There's lots of ways to try to get at participant diligence that we might want to use for exclusions.
-IN particular, it appears that mean accuracy is correlated with mean rt -- people with high accuracy take longer than people who look like they are randomly guessing.
+In particular, it appears that mean accuracy is correlated with mean rt -- people with high accuracy take longer than people who look like they are randomly guessing.
 In this experiment, we have an additional source of insight into diligence -- comprehension question accuracy.
@@
   labs(x="Fraction words selected correctly",
-       y="Mean reaction time",
+       y="Mean Reaction Time (ms)",
        color="?s correct")+
@@
 
-ggsave("../Papers/error.pdf", width=3, height=2, unit="in")
+ggsave("../Papers/error.pdf", width=3, height=2.1, unit="in")
 
@@
 
+# Error analysis
+
+```{r}
+
+b <- data_filt %>% filter(correct=="no") %>%
+  filter(type!="practice") %>%
+  select(word_num, subject, sentence) %>%
+  group_by(subject, sentence) %>%
+  mutate(row_num=row_number()) %>%
+  filter(row_num==1) %>%
+  select(-row_num) %>%
+  left_join(for_plot, by="subject")
+
+c <- data_filt %>% filter(type!=practice) %>%
+  group_by(subject, sentence) %>%
+  mutate(is.error=ifelse(correct=="no",1,0)) %>%
+  summarize(mistakes=sum(is.error)) %>%
+  mutate(no_mistake=ifelse(mistakes==0,"Error-free","Errorful")) %>%
+  select(subject, sentence, no_mistake) %>%
+  left_join(for_plot, by="subject")
+
+a <- ggplot(b, aes(x=word_num, fill=accurate))+geom_bar()+facet_grid(.~accurate)+scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
+  labs(title="Locations of first mistakes", x="Word position (0 is a fake choice), x-axis cut at 30")+coord_cartesian(xlim=c(0,30))+
+  guides(fill=F)
+
+d <- ggplot(c, aes(x=no_mistake, fill=accurate))+geom_bar()+facet_grid(.~accurate)+scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
+  labs(title="Number of sentences completed correctly", x=NULL)+
+  guides(fill=F)
+
+plot_grid(a,d,ncol=1, rel_heights=c(1,.75))
+
+```
+
+Even more evidence some participants are taking the random button pressing; also more evidence that some others are doing quite well.
+
 # How do errors affect participants and when do they occur

### 2020-04-30 Analysis/nat_stories_lms.Rmd
cleaned up with only current analysis

@@
 
-Surprisals are measured in bits (TODO: double check with Jenn about these being in bits, model specs/cites!)
+Surprisals are measured in bits.
 
@@
 
-The model formula we use is rt ~ surp + length * freq + past_surp + past_length * past_freq, so we're allowing frequency and length to interact and including the same terms for the current and past word.
+The model formula we use is rt ~ surp* length +  freq * length + past_surp * past_length + past_freq * past_length.
 
@@
 ```
+```{r}
 
-## Linear models
+lab_summary <- select_labs %>% summarize_at(c("length","freq","txl_surp","grnn_surp","ngram_surp"), list(mean=mean, stdev=sd, min=min, max=max)) %>% 
+  pivot_longer(everything()) %>% arrange(name)
 
-These use centered predictors.
-
-Using pre-error data only.
-
-```{r lm-2}
-
-lm_ngram_model_1 <- lm(rt ~ ngram_center + freq_center * length_center+ past_c_ngram + past_c_freq*past_c_length, data=labelled_pre_error)
-
-lm_txl_model_1 <- lm(rt ~ txl_center + freq_center * length_center+ past_c_txl + past_c_freq*past_c_length, data=labelled_pre_error)
-
-lm_grnn_model_1 <- lm(rt ~ grnn_center + freq_center * length_center+ past_c_grnn + past_c_freq*past_c_length, data=labelled_pre_error)
-
-```
-
-```{r lm_summary-2}
-
-summary(lm_ngram_model_1)
-
-summary(lm_txl_model_1)
-
-summary(lm_grnn_model_1)
-
-```
-Including post-error data.
-
-```{r lm-3}
-
-lm_ngram_model_2 <- lm(rt ~ ngram_center + freq_center * length_center+ past_c_ngram + past_c_freq*past_c_length, data=labelled_post_error)
-
-lm_txl_model_2 <- lm(rt ~ txl_center + freq_center * length_center+ past_c_txl + past_c_freq*past_c_length, data=labelled_post_error)
-
-lm_grnn_model_2 <- lm(rt ~ grnn_center + freq_center * length_center+ past_c_grnn + past_c_freq*past_c_length, data=labelled_post_error)
-
-```
-
-```{r lm_summary-3}
-
-summary(lm_ngram_model_2)
-
-summary(lm_txl_model_2)
-
-summary(lm_grnn_model_2)
+ggplot(select_labs, aes(length))+geom_bar()
 
@@
 
-We include a by-subject effect for everything.
-Technically speaking, we do have repeated measures at the word level, but there are 10K or so words and at most a handful of copies of each word, so that's likely to get us not very much at the cost of making the model rather intractable (and so I'm not going to deal with it, at least not right now).
+We include a by-subject effect for everything, and a by_word random intercept (full mixed effects).
 
@@
 
- brm_ngram_model <- brm(rt ~ ngram_center + freq_center * length_center +
-                         past_c_ngram + past_c_freq*past_c_length +
-                         (ngram_center + freq_center * length_center +
-                            past_c_ngram + past_c_freq*past_c_length|subject), data=labelled_pre_error,
-                       file="brm_ngram_pre_center", prior=priors, control=list(adapt_delta=.95))
+```
 
-brm_grnn_model <- brm(rt ~ grnn_center + freq_center * length_center +
-                         past_c_grnn + past_c_freq*past_c_length +
-                         (grnn_center + freq_center * length_center +
-                            past_c_grnn + past_c_freq*past_c_length|subject), data=labelled_pre_error,
-                       file="brm_grnn_pre_center", prior=priors, control=list(adapt_delta=.95))
+### On pre-error data only 
+```{r}
+brm_ngram_interact <- brm(rt ~ ngram_center* length_center + freq_center * length_center + 
+                         past_c_ngram*past_c_length + past_c_freq*past_c_length + 
+                         (ngram_center* length_center + freq_center * length_center + 
+                            past_c_ngram *past_c_length+ past_c_freq*past_c_length|subject)+
+                         (1|Word_ID), data=labelled_pre_error, 
+                       file="brm_ngram_interact", prior=priors, control=list(adapt_delta=.95))
 
-brm_txl_model <- brm(rt ~ txl_center + freq_center * length_center +
-                         past_c_txl + past_c_freq*past_c_length +
-                         (txl_center + freq_center * length_center +
-                            past_c_txl + past_c_freq*past_c_length|subject), data=labelled_pre_error,
-                       file="brm_txl_pre_center", prior=priors, control=list(adapt_delta=.95))
+brm_grnn_interact <- brm(rt ~ grnn_center * length_center+ freq_center * length_center + 
+                         past_c_grnn *past_c_length+ past_c_freq*past_c_length + 
+                         (grnn_center* length_center + freq_center * length_center + 
+                            past_c_grnn*past_c_length + past_c_freq*past_c_length|subject)+
+                        (1|Word_ID), data=labelled_pre_error, 
+                       file="brm_grnn_interact", prior=priors, control=list(adapt_delta=.95))
+
+brm_txl_interact <- brm(rt ~ txl_center* length_center + freq_center * length_center + 
+                         past_c_txl*past_c_length + past_c_freq*past_c_length + 
+                         (txl_center* length_center + freq_center * length_center + 
+                            past_c_txl *past_c_length+ past_c_freq*past_c_length|subject)+
+                       (1|Word_ID), data=labelled_pre_error, 
+                       file="brm_txl_interact", prior=priors, control=list(adapt_delta=.95))
 
@@
 }
-```
-```{r}
-summary(brm_txl_model)
-summary(brm_grnn_model)
-summary(brm_ngram_model)
-```
-
-```{r}
-a <- show_summary(brm_txl_model) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
-b <- show_summary(brm_grnn_model) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
-c <- show_summary(brm_ngram_model) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
+a <- show_summary(brm_txl_interact) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
+b <- show_summary(brm_grnn_interact) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
+c <- show_summary(brm_ngram_interact) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
 
@@
 print(xtable(summ,digits=c(0,0,1,0,2,1,0,2,1,0,2)), include.rownames=FALSE)
+```
+
+### On post-error as well
+
+```{r}
+brm_ngram_interact_post <- brm(rt ~ ngram_center* length_center + freq_center * length_center + 
+                         past_c_ngram*past_c_length + past_c_freq*past_c_length + 
+                         (ngram_center* length_center + freq_center * length_center + 
+                            past_c_ngram *past_c_length+ past_c_freq*past_c_length|subject)+
+                         (1|Word_ID), data=labelled_post_error, 
+                       file="brm_ngram_interact_post", prior=priors, control=list(adapt_delta=.95))
+
+brm_grnn_interact_post <- brm(rt ~ grnn_center * length_center+ freq_center * length_center + 
+                         past_c_grnn *past_c_length+ past_c_freq*past_c_length + 
+                         (grnn_center* length_center + freq_center * length_center + 
+                            past_c_grnn*past_c_length + past_c_freq*past_c_length|subject)+
+                        (1|Word_ID), data=labelled_post_error, 
+                       file="brm_grnn_interact_post", prior=priors, control=list(adapt_delta=.95))
+
+brm_txl_interact_post <- brm(rt ~ txl_center* length_center + freq_center * length_center + 
+                         past_c_txl*past_c_length + past_c_freq*past_c_length + 
+                         (txl_center* length_center + freq_center * length_center + 
+                            past_c_txl *past_c_length+ past_c_freq*past_c_length|subject)+
+                       (1|Word_ID), data=labelled_post_error, 
+                       file="brm_txl_interact_post", prior=priors, control=list(adapt_delta=.95))
+
+```
+```{r}
+show_summary_2 <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P-value equivalent` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=2),
+           upper=round(.upper, digits=2),
+           `Credible Interval`=str_c("[",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           Estimate=round(.value, digits=2)) %>% 
+    select(Term, Estimate, `Credible Interval`, `P-value equivalent`)
+
+  stats
+}
+```
+
+### Summaries
+
+```{r}
+summary(brm_ngram_interact_post)
+summary(brm_ngram_interact)
+
+summary(brm_grnn_interact_post)
+summary(brm_grnn_interact)
+
+summary(brm_txl_interact_post)
+summary(brm_txl_interact)
 ```
@@
 ```
+
+# TODOs for future
+
+In terms of surprisal analyses, the possible next steps are doing GAM analysis more robustly (with mixed effects, presenting results from the models) and doing DTSR.
+
+We could also try looking at the specific syntactic features that are in the dataset (the point of natural stories was to have a bunch of idioms and weird syntax); we could maybe analyse those?

### 2020-05-22 Papers/lab_meeting_2020.tex
more presentation work

\documentclass[12pt, xcolor=beamer,table,dvipsnames, ignorenonframetext, ngerman]{beamer}
\usetheme{Frankfurt}
\usecolortheme{dove}
\usepackage{appendixnumberbeamer}
%\setbeamersize{text margin left=20pt,text margin right=20pt,}

\beamertemplatenavigationsymbolsempty 
\setbeamertemplate{mini frames}{}
\setbeamertemplate{itemize item}{\textbullet}

\addtobeamertemplate{navigation symbols}{}{
	\ifnum\insertframenumber>\inserttotalframenumber%
	\relax
	\else%
	\usebeamerfont{footline}%
	\usebeamercolor[fg]{footline}%
	\hspace{1em}%
	\insertframenumber
	\fi%
}
\setbeamercolor{footline}{fg=black}
\usepackage{soul}
\makeatletter
\let\HL\hl
\renewcommand\hl{%
	\let\set@color\beamerorig@set@color
	\let\reset@color\beamerorig@reset@color
	\HL}

\usepackage{tipa}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\mode<presentation>

  \setbeamercovered{invisible}
\usepackage{multicol}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{ulem}
\usepackage{tipa}
\usepackage{qtree}
\usepackage{phonrule}
\usepackage{graphicx}
\usepackage{apacite}
\usepackage{xcolor}
\setlength\parindent{0pt}
\usepackage{natbib}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{tcolorbox}
\tcbuselibrary{raster}

\title{TODO A-maze works pretty well, actually}
\author{Veronica Boyce}
\date{3 June 2020}

\begin{document}

\begin{frame}
\maketitle
\end{frame}

%Notes for potential Maze presentation:
%
%Goal: introduce A-maze as an option to people, including for people who don't normally do processing; why it might be good
%
%Background:
%- why reaction time
%- existing methods
%- unsatisfied with web results
%
\section{Why RT?}
\begin{frame}{Why measure reading time?}
Linguistic and psycholinguistic theories make predictions about processing difficulty.
\medskip

We assume that harder processing manifests in longer reading/reaction time (RT).
\medskip

RT patterns may be phenomena that theories need to explain.

Examples of harder processing
\begin{itemize}
	\item Less frequent words are harder to retrieve
	\item Reparsing/reanalysis
	\item Less grammatical sentences
\end{itemize}
%certainly not the only measure
%ex. Noisy channel, syntactic preferences, surprisal theory, what's hard in L2.
\end{frame}

\begin{frame}{Two common methods}
%can't get at the internals of processing, so want to use time as proxy for how much is going on 
%but literate humans are good at reading 
\begin{columns}[t]\pause
	\begin{column}{.5\textwidth}
		\begin{center}
			\textbf{\Large Eye-tracking}

			\medskip

			\includegraphics[width=.9\textwidth]{eye-tracker.jpeg}
			\begin{itemize}
				\item Expensive
				\item Hard to analyse
			\end{itemize}
		\end{center}
	\end{column}\pause
	\begin{column}{.5\textwidth}
		\begin{center}
\textbf{\Large Self-pace reading}

\medskip

\begin{tikzpicture} %TODO on slide etc
\draw (-.5,-.5) rectangle (3.2,.5);
\onslide<3>{\node at (0,0) {The};}
\onslide<3>{\node at (1.7,0) {- - - - - - - - - - -};}
\onslide<4>{\node at (0,0) {- - - };}
\onslide<4>{\node at (.7,0) {cat};}
\onslide<4>{\node at (2,0) {- - - - - - - - };}
\onslide<5>{\node at (0.3,0) {- - - - - - };}
\onslide<5>{\node at (1.6,0) {drank};}
\onslide<5>{\node at (2.6,0) {- - - -};}
\onslide<6->{\node at (.9,0) {- - - - - - - - - - -};}
\onslide<6->{\node at (2.7,0) {milk.};}
\end{tikzpicture}
\end{center}
			\begin{itemize}
	\item Lots of spillover
	\item Messy data 
\end{itemize}
\end{column}
\end{columns}

\end{frame}

\begin{frame}{Maze Task}
\centering
\tikzset{
	font={\fontsize{14pt}{12}\selectfont}}
\begin{tikzpicture}
\onslide<1->{\node[draw] (a) at (-1,-.5) {e};}
\onslide<1->{\node[draw] (b) at (1,-.5) {i};}
\onslide<1-2>{\node  (c) at (-1,.5) {The};}
\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
\onslide<3-4>{\node  (d) at (1,.5) {dog};}
\onslide<5-6>{\node  (c) at (-1,.5) {revise};}
\onslide<5-6>{\node  (d) at (1,.5) {chased};}
\onslide<7-8>{\node  (c) at (-1,.5) {the};}
\onslide<7-8>{\node  (d) at (1,.5) {wish};}
\onslide<9-10>{\node  (c) at (-1,.5) {mitigate.};}
\onslide<9-10>{\node  (d) at (1,.5) {squirrel.};}
\onslide<2,8>{\node[draw, fill=pink!50] at (a) {e};}
\onslide<4,6,10>{\node[draw, fill=pink!50] at (b) {i};}
\end{tikzpicture}
\end{frame}

\begin{frame}{A third option: Maze}
\begin{columns}
	\begin{column}{0.5\textwidth}
		\begin{center}
		\textbf{\large G-maze}
		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{gmaze.pdf}
		\end{center}

	\end{column}
	\begin{column}{0.5\textwidth} 
		\begin{center}
		\textbf{\large L-maze}
			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{lmaze.pdf}
		\end{center}
	\end{column}
\end{columns}

\medskip

Sentence ends when a mistake is made

Central claim: forces extremely incremental processing (no spillover)

\begin{flushright}
	{\small(Forster et al. 2009; Witzel et al. 2012)}
\end{flushright}
\end{frame}

\section{Experiment 1a}
\begin{frame}{Web Maze}
Maze well suited to run on the web.
\begin{itemize}
	\item Implement in Ibex using js
\end{itemize}

Test by replicating Witzel et al. (2012)
\begin{itemize}
	\item Comparison of eye-tracking, SPR, L-maze, G-maze (all in-lab)
	\item Got materials and data from original
	\item Run SPR, L-maze, and G-maze on MTurk
\end{itemize}
\end{frame}

\begin{frame}{Materials}

\textbf{Relative Clause}

\sethlcolor{green}
\textit{Low:} The son of the \uline{lady} who politely introduced \hl{herself} was popular at the party.
\sethlcolor{pink}

 \textit{High:} The \uline{son} of the lady who politely introduced \hl{himself} was popular at the party.

\textbf{Adverb Clause}

\sethlcolor{green}\textit{Low:} James will fix the car he \uline{drove} \hl{yesterday}, but he will need some help.

\sethlcolor{pink} \textit{High:} James \uline{will fix} the car he drove \hl{tomorrow}, but he will need some help.

\textbf{Sentence v Noun Phrase conjunction}

\sethlcolor{green} \textit{Comma:} The swimmer disappointed her \uline{coach}, and her mother \hl{tried} to console her.

\sethlcolor{pink}
\textit{No comma:} The swimmer disappointed her \uline{coach} and her mother \hl{tried} to console her.
\end{frame}

\begin{frame}{Results}
\begin{small}	
	\sethlcolor{green}
The son of the lady who politely introduced \hl{herself}\sethlcolor{pink} / \hl{himself} was popular at the party.

\end{small}
\includegraphics[width=\textwidth]{g_rel.pdf}
\end{frame}

\begin{frame}{Results}
\begin{small}

\sethlcolor{green}James will fix the car he drove \hl{yesterday}\sethlcolor{pink} / \hl{tomorrow},  but he will need some help.

\end{small}
\includegraphics[width=\textwidth]{g_adv.pdf}
\end{frame}

\begin{frame}{Results}
\begin{small}	
\sethlcolor{green}The swimmer disappointed her coach\hl{,} and her mother \hl{tried} / \sethlcolor{pink}\hl{tried} to console her.
\end{small}
\includegraphics[width=\textwidth]{g_svnp.pdf}
\end{frame}
\begin{frame}
Good news: G-maze work over the web

Bad news: G-maze materials are a huge pain to write
\end{frame}

\section{A-maze}

\begin{frame}{Meanwhile in NLP}
Language models (LM)
\begin{itemize}
	\item Trained on large corpora to predict the next word
	\item Learn the distribution over next word given a partial sentence
\end{itemize}
Surprisal: negative log probability
\begin{itemize}
	\item 2 bits of surprisal = 1/4
	\item 10 bits of surprisal $\approx$ 1/1000 
	\item +1 surprisal = half as likely
\end{itemize}
\begin{tikzpicture}
\node (A) at (0,0) {"The dog chased"};
\node  (B) at (4,0) [draw, minimum width=1.5cm, minimum height=1.5cm, fill=black] {\textcolor{white}{LM}} ;
\draw [->, thick] (A) -- (B); 
\node[align=left] (C) at (7,0) {"the" 1.2 bits \\"a" 3.1 bits\\"cats" 8 bits\\...\\"wish" 25 bits\\...};
\draw [->, thick] (B) -- (C); 
\end{tikzpicture}
\end{frame}

\begin{frame}{LMs as RAs}
Feed items to LM, get out distribution and choose a high-surprisal word for the distractor.

Want to avoid random nonsense, so we restrict distractor options
\begin{itemize}
	\item Use a dictionary, do some manual exclusions
	\item Match distractor's length, frequency to correct word
	\item Go down list of potential distractors until we find one that's high enough surprisal
\end{itemize}

\end{frame}

\section{Experiment 1b}
\begin{frame}{But does it work?}
Foreshadowing: Yes.
\end{frame}

\begin{frame}{But does it work?}
Test with 2 pre-trained LSTM models (Gulordava 
Go back to these syntactic things. Add A-maze results. Ran them didn't look at them. 
\end{frame}

\begin{frame}{But does it work?}
\includegraphics[width=.95\textwidth]{a_all.pdf}
\end{frame}
%
%\begin{frame}{Caveats}
%there were a lot of mistakes on word 2
% - can't tell if this is bad participants or bad distractors
% - presumably some of each
% - data loss, want to fix this
% - can't really tell how good participants are,
%\end{frame}
%
%\begin{frame}
%Solution 1:
%make it better
%
%we did a little of this, added some tweaks, made is easier for users to specify badnesses
%
%but this is hard, it's unclear what to expect, and we haven't done grid search or anything
%\end{frame}
%
%\begin{frame}
%Solution 2: `redo mode`
%Just have people correct their mistakes and keep going. 
%\end{frame}
%
%\begin{frame}{Maze with Error Correction}
%\centering
%\tikzset{
%	font={\fontsize{14pt}{12}\selectfont}}
%\begin{tikzpicture}
%\onslide<1->{\node[draw] (a) at (-1,-.5) {e};}
%\onslide<1->{\node[draw] (b) at (1,-.5) {i};}
%\onslide<1-2>{\node  (c) at (-1,.5) {The};}
%\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
%\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
%\onslide<3-4>{\node  (d) at (1,.5) {dog};}
%\onslide<5-8>{\node  (c) at (-1,.5) {revise};}
%\onslide<5-8>{\node  (d) at (1,.5) {chased};}
%\onslide<9-10>{\node  (c) at (-1,.5) {the};}
%\onslide<9-10>{\node  (d) at (1,.5) {wish};}
%\onslide<11-12>{\node  (c) at (-1,.5) {mitigate.};}
%\onslide<11-12>{\node  (d) at (1,.5) {squirrel.};}
%\onslide<2,6,10>{\node[draw, fill=pink!50] at (a) {e};}
%\onslide<4,8,12>{\node[draw, fill=pink!50] at (b) {i};}
%\onslide<7,8>{\node[text=red] at (0,-1.5) {\small Incorrect. Please try again.};}
%\end{tikzpicture}
%\end{frame}
%
%\begin{frame}
%\includegraphics{maze_diagram.pdf} %TODO trim
%- really easy to do in our framework
%and now we have a lot of advantages
%- have all the data
%so we can tell the bad participant from the bad item
%and the bad distractor doesn't screw us over
%
%Convenient side effect: we can run multi-sentence items. Hadn't been done before b/c even low error rates add up.
%\end{frame}
%
%\section{Experiment 2}
%
%\begin{frame}{Natural Stories}
%Natural stories corpus
%10 stories, each about 1000 words
%read fluently to native speakers
%sample,
%and they come with comprehension questions
%\end{frame}
%
%\begin{frame}{Natural Stories}
%Methods
%- read a short practice w comp q
%- read 1 story and questions
%\end{frame}
%
%\begin{frame}{Natural Stories}
%Methods
%- read a short practice w comp q
%- read 1 story and questions
%\end{frame}
%
%\begin{frame}{Natural Stories}
%Participant quality
%- mean RT versus how many they got right
%- select this part and not those
%\includegraphics{error.pdf}
%\end{frame}
%\begin{frame}{Natural Stories}
%Participant quality and comprehension questions
%\includegraphics{comp.pdf}
%\end{frame}
%
%\begin{frame}{Natural Stories}
%It's working
%- people will put up with reading this for 15 minutes
%- some of them can understand the story
%- our distractors are actually good enough (most of the time)
%\end{frame}
%
%\begin{frame}{Natural Stories}
%Analysis time -- we check for known surprisal effects
%length*surprisal (? frequency)
%Copy over table 
%spillover
%is it linear
%what we find
%\includegraphics[]{gam.pdf}	
%note that it's crucial to use single token words -- predictions on add punct together aren't good
%\end{frame}
%
%\begin{frame}{Natural Stories}
%And it same result holds on after-mistake data 
%skip a couple words where people might be recovering (double check how much we skip), but then it's fine
%\end{frame}
%
%\section{Conclusion}
%\begin{frame}
%Potential and Unknowns
%- easily extendable to other language models, languages 
%- easy to use; command line runnable, backed by Python
%
%Unknowns:
%- only as good as the lms are, might not be great for triple center embeddings ... but you can also fix a few words
%\end{frame}

%

\appendix
%Back pocket slide on matching procedure
%back pocket on more on model results??
%do we even need back pockets??

\end{document}

### 2020-05-27 Papers/lab_meeting_2020.tex
work on slides

@@
-\documentclass[12pt, xcolor=beamer,table,dvipsnames, ignorenonframetext, ngerman]{beamer}
+\documentclass[12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman]{beamer}
 \usetheme{Frankfurt}
@@
 
-\title{TODO A-maze works pretty well, actually}
+\title{A-maze: Easier measurement of incremental processing}
 \author{Veronica Boyce}
@@
 
-%Notes for potential Maze presentation:
 %
-%Goal: introduce A-maze as an option to people, including for people who don't normally do processing; why it might be good
-%
-%Background:
-%- why reaction time
-%- existing methods
-%- unsatisfied with web results
-%
+
 \section{Why RT?}
 \begin{frame}{Why measure reading time?}
+\pause
 Linguistic and psycholinguistic theories make predictions about processing difficulty.
+\medskip
+\pause
+
+Examples of increased difficulty
+\begin{itemize}\pause
+	\item Constructions not in the grammar \pause
+	\item Lexical items that are harder to retrieve \pause
+	\item Words that force reparsing or reanalysis
+\end{itemize}
+\pause
 \medskip
@@
 We assume that harder processing manifests in longer reading/reaction time (RT).
+\pause
 \medskip
@@
 
-Examples of harder processing
-\begin{itemize}
-	\item Less frequent words are harder to retrieve
-	\item Reparsing/reanalysis
-	\item Less grammatical sentences
-\end{itemize}
-%certainly not the only measure
-%ex. Noisy channel, syntactic preferences, surprisal theory, what's hard in L2.
 \end{frame}
@@
 			\includegraphics[width=.9\textwidth]{eye-tracker.jpeg}
+			\pause
 			\begin{itemize}
@@
 		\begin{center}
-\textbf{\Large Self-pace reading}
+\textbf{\Large Self-paced reading}
 
@@
 \draw (-.5,-.5) rectangle (3.2,.5);
-\onslide<3>{\node at (0,0) {The};}
-\onslide<3>{\node at (1.7,0) {- - - - - - - - - - -};}
-\onslide<4>{\node at (0,0) {- - - };}
-\onslide<4>{\node at (.7,0) {cat};}
-\onslide<4>{\node at (2,0) {- - - - - - - - };}
-\onslide<5>{\node at (0.3,0) {- - - - - - };}
-\onslide<5>{\node at (1.6,0) {drank};}
-\onslide<5>{\node at (2.6,0) {- - - -};}
-\onslide<6->{\node at (.9,0) {- - - - - - - - - - -};}
-\onslide<6->{\node at (2.7,0) {milk.};}
+\onslide<4>{\node at (0,0) {The};}
+\onslide<4>{\node at (1.7,0) {- - - - - - - - - - -};}
+\onslide<5>{\node at (0,0) {- - - };}
+\onslide<5>{\node at (.7,0) {cat};}
+\onslide<5>{\node at (2,0) {- - - - - - - - };}
+\onslide<6>{\node at (0.3,0) {- - - - - - };}
+\onslide<6>{\node at (1.6,0) {drank};}
+\onslide<6>{\node at (2.6,0) {- - - -};}
+\onslide<7->{\node at (.9,0) {- - - - - - - - - - -};}
+\onslide<7->{\node at (2.7,0) {milk.};}
 \end{tikzpicture}
 \end{center}
+\onslide<8->{
 			\begin{itemize}
@@
 \end{itemize}
+}
 \end{column}
 \end{columns}
-
 \end{frame}
 
+\begin{frame}{A third option: Maze}
+\end{frame}
 \begin{frame}{Maze Task}
@@
 		\end{center}
-
+	\pause
 	\end{column}
@@
 \medskip
-
-Sentence ends when a mistake is made
+\pause
+Sentence ends when a mistake is made.
+\pause
 
@@
 \section{Experiment 1a}
-\begin{frame}{Web Maze}
-Maze well suited to run on the web.
+\begin{frame}{Web Implementation}
+Maze well suited to run on the web
 \begin{itemize}
-	\item Implement in Ibex using js
+	\item Implement in Ibex
 \end{itemize} 
+\pause
 
-Test by replicating Witzel et al. (2012)
+In Boyce et al. (2020) we replicate Witzel et al. (2012)
 \begin{itemize}
-	\item Comparison of eye-tracking, SPR, L-maze, G-maze (all in-lab)
-	\item Got materials and data from original
-	\item Run SPR, L-maze, and G-maze on MTurk
+	\item Witzel et al (2012): Comparison of eye-tracking, SPR, L-maze, G-maze (all in-lab)
+	\item Got materials and data from Witzel
+	\item We run SPR, L-maze, and G-maze on MTurk
 \end{itemize}
@@
 \end{frame}
-\begin{frame}
-Good news: G-maze work over the web
 
-Bad news: G-maze materials are a huge pain to write
+\begin{frame}{Interim Summary}
+\textcolor{PineGreen}{\Large \textbf{Good news:} G-maze works well over the web (better than L-maze or SPR)}
+
+\pause
+
+\bigskip
+
+\textcolor{Red}{\Large \textbf{Bad news:} I do not want to write G-maze materials.}
+
 \end{frame}
@@
 
-\begin{frame}{Meanwhile in NLP}
-Language models (LM)
+\begin{frame}{Meanwhile in Natural Language Processing}
+\pause
+Language models (LMs)
 \begin{itemize}
 	\item Trained on large corpora to predict the next word
-	\item Learn the distribution over next word given a partial sentence
+	\item Given a partial sentence, return probabilities of the next word
 \end{itemize}
+\pause
 Surprisal: negative log probability
@@
 \end{itemize}
+\pause
 \begin{tikzpicture}
-\node (A) at (0,0) {"The dog chased"};
+\node (A) at (0,0) {"The dog chased ..."};
 \node  (B) at (4,0) [draw, minimum width=1.5cm, minimum height=1.5cm, fill=black] {\textcolor{white}{LM}} ;
@@
 
-\begin{frame}{LMs as RAs}
-Feed items to LM, get out distribution and choose a high-surprisal word for the distractor.
+\begin{frame}{Can we use LMs to choose distractors?} Idea: Use high surprisal according to LM as a proxy for really bad in context
+\medskip
 
-Want to avoid random nonsense, so we restrict distractor options
+\pause
+
 \begin{itemize}
-	\item Use a dictionary, do some manual exclusions
-	\item Match distractor's length, frequency to correct word
-	\item Go down list of potential distractors until we find one that's high enough surprisal
+	\item Model the target sentence word by word \pause
+	\item At each position, choose a high surprisal word
+\end{itemize}
+\medskip
+\pause
+Want quality control on distractors
+\pause
+\begin{itemize}
+	\item Restrict to a list of possible distractors
+	\item Only consider distractors of same length, frequency as target wrd
+	\item Check distractors until we find one with high surprisal
+	\item Match to real word punctuation
 \end{itemize}
@@
 
+\begin{frame}
+\centering
+\includegraphics[clip, trim=3.25cm 5cm 2.5cm 2.5cm,width=.9\textwidth]{flow_2.pdf}
+\end{frame}
 \section{Experiment 1b}
-\begin{frame}{But does it work?}
-Foreshadowing: Yes.
+
+\begin{frame}{Does A-maze work? }
+
+Test it on materials from Witzel et al. (2012)
+\medskip
+
+Try with two pre-trained LSTM models (Gulordava 2018, Jozefowicz 2016)
+
 \end{frame}
 
-\begin{frame}{But does it work?}
-Test with 2 pre-trained LSTM models (Gulordava 
-Go back to these syntactic things. Add A-maze results. Ran them didn't look at them. 
+\begin{frame}{Does A-maze work?  Yes.} \pause
+\includegraphics[width=1\textwidth]{a_all.pdf}
 \end{frame}
 
-\begin{frame}{But does it work?}
-\includegraphics[width=.95\textwidth]{a_all.pdf}
+\begin{frame}{Caveats}
+A lot of mistakes on the second word of the sentence ("The \ul{dog} ...") \pause
+\includegraphics{../graph_errors2.pdf} \pause
+Bad participants or bad distractors?
+
+Either way, lots of data loss.
 \end{frame}
-%
-%\begin{frame}{Caveats}
-%there were a lot of mistakes on word 2
-% - can't tell if this is bad participants or bad distractors
-% - presumably some of each
-% - data loss, want to fix this
-% - can't really tell how good participants are,
-%\end{frame}
-%
-%\begin{frame}
-%Solution 1:
-%make it better
-%
-%we did a little of this, added some tweaks, made is easier for users to specify badnesses
-%
-%but this is hard, it's unclear what to expect, and we haven't done grid search or anything
-%\end{frame}
-%
-%\begin{frame}
-%Solution 2: `redo mode`
-%Just have people correct their mistakes and keep going. 
-%\end{frame}
-%
-%\begin{frame}{Maze with Error Correction}
-%\centering
-%\tikzset{
-%	font={\fontsize{14pt}{12}\selectfont}}
-%\begin{tikzpicture}
-%\onslide<1->{\node[draw] (a) at (-1,-.5) {e};}
-%\onslide<1->{\node[draw] (b) at (1,-.5) {i};}
-%\onslide<1-2>{\node  (c) at (-1,.5) {The};}
-%\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
-%\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
-%\onslide<3-4>{\node  (d) at (1,.5) {dog};}
-%\onslide<5-8>{\node  (c) at (-1,.5) {revise};}
-%\onslide<5-8>{\node  (d) at (1,.5) {chased};}
-%\onslide<9-10>{\node  (c) at (-1,.5) {the};}
-%\onslide<9-10>{\node  (d) at (1,.5) {wish};}
-%\onslide<11-12>{\node  (c) at (-1,.5) {mitigate.};}
-%\onslide<11-12>{\node  (d) at (1,.5) {squirrel.};}
-%\onslide<2,6,10>{\node[draw, fill=pink!50] at (a) {e};}
-%\onslide<4,8,12>{\node[draw, fill=pink!50] at (b) {i};}
-%\onslide<7,8>{\node[text=red] at (0,-1.5) {\small Incorrect. Please try again.};}
-%\end{tikzpicture}
-%\end{frame}
-%
-%\begin{frame}
-%\includegraphics{maze_diagram.pdf} %TODO trim
-%- really easy to do in our framework
-%and now we have a lot of advantages
-%- have all the data
-%so we can tell the bad participant from the bad item
-%and the bad distractor doesn't screw us over
-%
-%Convenient side effect: we can run multi-sentence items. Hadn't been done before b/c even low error rates add up.
-%\end{frame}
-%
-%\section{Experiment 2}
-%
-%\begin{frame}{Natural Stories}
-%Natural stories corpus
-%10 stories, each about 1000 words
-%read fluently to native speakers
-%sample,
-%and they come with comprehension questions
-%\end{frame}
-%
-%\begin{frame}{Natural Stories}
-%Methods
-%- read a short practice w comp q
-%- read 1 story and questions
-%\end{frame}
-%
-%\begin{frame}{Natural Stories}
-%Methods
-%- read a short practice w comp q
-%- read 1 story and questions
-%\end{frame}
-%
-%\begin{frame}{Natural Stories}
-%Participant quality
-%- mean RT versus how many they got right
-%- select this part and not those
-%\includegraphics{error.pdf}
-%\end{frame}
-%\begin{frame}{Natural Stories}
-%Participant quality and comprehension questions
-%\includegraphics{comp.pdf}
-%\end{frame}
-%
-%\begin{frame}{Natural Stories}
-%It's working
-%- people will put up with reading this for 15 minutes
-%- some of them can understand the story
-%- our distractors are actually good enough (most of the time)
-%\end{frame}
-%
-%\begin{frame}{Natural Stories}
-%Analysis time -- we check for known surprisal effects
-%length*surprisal (? frequency)
-%Copy over table 
-%spillover
-%is it linear
-%what we find
-%\includegraphics[]{gam.pdf}	
-%note that it's crucial to use single token words -- predictions on add punct together aren't good
-%\end{frame}
-%
-%\begin{frame}{Natural Stories}
-%And it same result holds on after-mistake data 
-%skip a couple words where people might be recovering (double check how much we skip), but then it's fine
-%\end{frame}
-%
-%\section{Conclusion}
-%\begin{frame}
-%Potential and Unknowns
-%- easily extendable to other language models, languages 
-%- easy to use; command line runnable, backed by Python
-%
-%Unknowns:
-%- only as good as the lms are, might not be great for triple center embeddings ... but you can also fix a few words
-%\end{frame}
 
-%
+\begin{frame}{Caveats}
+
+{\large Definitely some bad distractors}
+\begin{table}
+
+		\begin{tabular}{rlll}
+			Prefix & Correct & Distractor & Error Rate \\
+			\hline
+			\hline
+			Gulordava&&&\\
+			\hline
+			The & niece & cooks & 44\%\\
+			The swimmer & disappointed & propositions & 30\%\\
+			The & semester & steroids & 29\%\\
+			\hline
+			\hline
+			 Jozefowicz&&&\\
+			\hline
+			The & husband & authors & 46\%\\
+			Jim & listened & survived & 43\%\\
+			The & uncle & roads & 42\%\\
+			The & knight & saints & 40\%\\
+		\end{tabular}
+\end{table}
+
+\end{frame}
+
+\begin{frame}{Potential improvements}
+
+{\large Option 1: Choose better distractors.}
+\begin{itemize}
+	\item Tweak thresholds of distractor selection.
+	\item Could search for optimal thresholds.
+	\item Could manually check and fix distractors.
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Potential improvements}
+
+{\large Option 2: Be forgiving.}
+\begin{itemize}
+	\item Just don't terminate on mistakes.
+	\item Instead have participant correct mistake and continue. 
+	\item Easy to implement in Ibex framework.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Maze with Error Correction}
+\centering
+\tikzset{
+	font={\fontsize{14pt}{12}\selectfont}}
+\begin{tikzpicture}
+\onslide<1->{\node[draw] (a) at (-1,-.5) {e};}
+\onslide<1->{\node[draw] (b) at (1,-.5) {i};}
+\onslide<1-2>{\node  (c) at (-1,.5) {The};}
+\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
+\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
+\onslide<3-4>{\node  (d) at (1,.5) {dog};}
+\onslide<5-8>{\node  (c) at (-1,.5) {revise};}
+\onslide<5-8>{\node  (d) at (1,.5) {chased};}
+\onslide<9-10>{\node  (c) at (-1,.5) {the};}
+\onslide<9-10>{\node  (d) at (1,.5) {wish};}
+\onslide<11-12>{\node  (c) at (-1,.5) {mitigate.};}
+\onslide<11-12>{\node  (d) at (1,.5) {squirrel.};}
+\onslide<2,6,10>{\node[draw, fill=pink!50] at (a) {e};}
+\onslide<4,8,12>{\node[draw, fill=pink!50] at (b) {i};}
+\onslide<7,8>{\node[text=red] at (0,-1.5) {\small Incorrect. Please try again.};}
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}{Maze with Error Correction}
+
+{\large The ``solution'' to our problems!}
+
+\begin{columns}
+	\begin{column}{0.3\textwidth}
+		\begin{center}
+		\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=\textwidth]{maze_diagram.pdf}
+		\end{center}
+
+	\end{column} \pause 
+	\begin{column}{0.7\textwidth} 
+		\begin{center}
+			\begin{itemize}
+				\item Can be toggled in Ibex
+				\item Have all the data
+				\item Identify bad participants v bad distractors
+			\end{itemize}
+		\end{center}
+	\end{column}
+\end{columns}
+\medskip
+\pause
+
+Side effect: Can now run multi-sentence items!
+\end{frame}
+
+\section{Experiment 2} %TODO: Needs work from here on!
+
+\begin{frame}{Natural Stories}
+Natural stories corpus (Futrell et al. 2017)
+\begin{itemize}
+	\item 10 stories, each about 1000 words
+	\item Some unusual constructions, but read fluently
+	\item 6 comprehension questions per story
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Natural Stories}
+
+\begin{small}Tulip mania was a period in the Dutch Golden Age during which contract prices for bulbs of the recently introduced tulip reached extraordinarily high levels and then suddenly collapsed.
+At the peak of tulip mania in February sixteen thirty-seven, tulip contracts sold for more than ten times the annual income of a skilled craftsman.
+It is generally considered the first recorded economic bubble.
+[...]
+\medskip
+
+Q: When did tulip mania reach its peak?
+
+A: \hspace{3em} 1630's\hspace{3em} 1730's \end{small}
+
+\end{frame}
+
+\begin{frame}{Experiment Plan}
+Experiment questions:
+\begin{itemize}
+	\item Will people do this task?
+	\item Can participants comprehend the stories?
+\end{itemize}
+
+\pause
+
+\medskip
+
+Checks on RT data
+\begin{itemize}
+	\item Check if RT is linear in surprisal (Smith \& Levy 2013)
+	\item Check for known effects of word length, frequency
+	\item Check for spillover effects
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Participant accuracy}
+100 participants each read 1 story \pause
+\begin{center}
+\includegraphics[width=.8\textwidth]{error.pdf}
+\end{center}
+\end{frame}
+\begin{frame}{Comprehension questions}
+\begin{center} \pause
+\includegraphics[width=.8\textwidth]{comp.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}{Surprisal Effects}
+
+Use 3 LMs to estimate surprisal: smoothed 5-gram, Gulordava RNN, Transformer-XL (Gulordava et al. 2018, Dai et al. 2019)
+
+\includegraphics[width=.9\textwidth]{gam.pdf}
+
+Only on single token words.
+
+\end{frame}
+
+\begin{frame}{What about post-mistake data?}
+Exclude data from mistakes or the two words after a mistake.
+
+\includegraphics[width=.9\textwidth]{gam2.pdf}	
+\end{frame}
+
+\begin{frame}{Result Summary}
+\begin{itemize}
+	\item People will read in the Maze task for 15-20 minutes
+	\item It's possible to comprehend during Maze
+	\item Distractors are generally good enough
+	\item Find expected RT patterns
+	\item Very little spillover
+\end{itemize}
+\end{frame}
+\section{Conclusion}
+
+\begin{frame}{Use A-maze!}
+Easy to use!
+\begin{itemize}
+	\item Runs on command line 
+	\item Match distractors across minimal pair sentences
+	\item Customize surprisal thresholds, vocabulary lists
+	\item Can output pre-formatted for Ibex
+	\item Lots of documentation at vboyce@github.io/Maze
+\end{itemize}
+\medskip
+\pause
+\end{frame}
+
+\begin{frame}{Use A-maze!}
+Contribute to A-maze!:
+\begin{itemize}
+	\item Written in Python 3
+	\item Add more language models
+	\item Add more frequency sources
+	\item Extend to non-English languages
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{}
+
+\textcolor{ForestGreen}{\large Documentation: vboyce@github.io/Maze}
+
+with links to the following:
+\begin{itemize}
+
+\item A-maze code: github.com/vboyce/Maze
+
+\item Web-maze code: github.com/vboyce/Ibex-with-Maze
+
+\item Sample task: syntaxgym.org:666
+\end{itemize}
+\end{frame}
 
 \appendix
-%Back pocket slide on matching procedure
+
+\begin{frame}{Matching distractors}
+If unspecified: Match by position
+\begin{itemize}
+	\item The son of the lady who politely introduced \sethlcolor{green} \hl{herself}\sethlcolor{pink} / \hl{himself} was popular at the party.
+\end{itemize}
+Can specify labels for each word to pair (within item)
+\begin{itemize}
+	\item The cat who the dog scared hid in a box.\\pre-1 pre-2 who art noun verb main-verb post-1 post-2 post-3
+\item The dog who scared the cat sniffed around the couch.\\ pre-1 pre-2 who verb art noun main-verb post-1 post-2 post-3
+\end{itemize}
+\end{frame}
 %back pocket on more on model results??
@@
 
+\begin{frame}{Post-mistake only}
+Exclude data from mistakes or the two words after a mistake.
+
+\includegraphics[]{gam3.pdf}	
+\end{frame}
+
 \end{document}

### 2020-05-27 Papers/flow_2.tex
add gitignore, work on paper

\documentclass[10pt]{minimal}

\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows, fit}
\begin{document}
\pagestyle{empty}

% Define block styles
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=10em, text centered, minimum height=3em,line width=.3mm]
\tikzstyle{line} = [draw, -latex', line width=.5mm]
\tikzstyle{cloud} = [draw, rounded rectangle,fill=red!20, text width=7em, text centered,
    minimum height=4em,line width=.3mm]
    \tikzstyle{input} = [draw, rounded rectangle,fill=green!20, text width=7em, text centered,
    minimum height=4em, line width=.3mm]

\begin{tikzpicture}[node distance =3cm]
\node [input] (freq) {Frequency source};
\node [input, right of=freq, xshift=0cm] (dict) {Dictionary};
\node [block, text width=13em, below of=freq, yshift=0cm, xshift=2cm] (dist) {Filter for valid distractors:\\ a) in frequency source \\ b) only lowercase alpha characters \\ c) not on manual exclude list \\ d) in model vocab };
\node [cloud, text width =12em, below of=dist, xshift=0cm, yshift=0cm] (distlist) {(Length, frequency) : potential distractors\\ lookup };

 \node [input, left of=freq, xshift=-6.5cm, yshift=2.5cm, line width=.5mm] (mat) {\textbf{Materials}};
   \node [block, below of =mat, xshift=.5cm, yshift=1cm] (prep) {Read in materials};
 \node [block, below right of=prep,xshift=-.4cm, yshift=0cm] (sentence) {Incrementally model sentence};     
   \node [input, right of=mat,xshift=1cm, yshift=-.5cm, text width=10em] (NN) {Neural Net model\\{\includegraphics[width=3cm]{lstm.png}}};

   \node [cloud, below right of=sentence, xshift=.5cm,  yshift=0cm] (word2) {True next word};  
\node [cloud, below left of =sentence, xshift=.5cm,  yshift=0cm] (surp) {Probability distribution over next word};
\node [block, below  of =word2, yshift=1cm, xshift=-.5cm] (potdist) {Generate length, frequency matched potential distractors};
\node [block, text width=12em,  below of =potdist, yshift=1cm, xshift=-3cm] (choose) {Choose  distractor word with high  surprisal};
\node [cloud, below of =choose, xshift=-1cm, yshift=1cm] (distractor) {Distractor};

\node [block, right of =distractor, xshift=1cm] (collect) {Collect distractor words into a sentence};

\node [block, right of =choose, xshift=5cm] (format) {Write items and distractors to file};

\node [input, line width= .5mm, below of=format, yshift=1cm] (final) {\textbf{Materials for Maze experiment}};
\node[draw,dotted, line width= .5mm, fit=(freq) (distlist) (dict) (dist)] (box1) { };
\node[draw, draw=none, above of=box1, xshift=0cm,  yshift=1cm] {Set up};
\node[draw,dashed,blue,line width=.5mm, inner sep=10pt, fit= (surp) (word2)(choose)(potdist)] (box2){ };
\node[draw, draw=none, above of=box2, xshift=0cm,  yshift=.2cm] {Do for each word in sentence};
	\node[draw,dashed,red,line width=.5mm, inner sep=13pt, fit= (word2) (surp) (choose) (distractor) (collect)  (sentence)] (box3) { };
\node[draw, draw=none, above of=box3, xshift=0cm,  yshift=2.4cm] {Do for each item in materials};
    % Draw edges
    \path [line] (dict) -- (dist);
    \path [line] (dist.south) -- (distlist.north);
	\path [line] (word2) -- (potdist);
	\path [line] (freq) -- ([yshift=.2cm]potdist.east);
		\path [line] (freq) -- (dist);
	\path [line] (distlist) -- (potdist);
	\path [line] (potdist) -- (choose);	
	\path [line] (surp) -- (choose);
	\path [line] (choose) -- (distractor);
	\path [line] ([yshift=-.2cm]sentence) -| ([xshift=-.7cm]surp.north);
	\path [line] ([yshift=-.2cm]sentence) -| ([xshift=.7cm]word2.north);
	\path [line] ([xshift=-1cm]prep.south) |- ([yshift=.2cm]sentence.west);
	\path [line] ([xshift=1.5cm]NN.south) |- ([yshift=.2cm]sentence.east);
	\path [line] (distractor) -- (collect);
	\path [line] (collect) -- (format.west);
	\path [line] (format) -- (final);
	\path [line] (mat) -- (prep);

\end{tikzpicture}

\end{document}

### 2020-05-27 Analysis/nat_stories.Rmd
gitignore

@@
 ggsave("../Papers/error.pdf", width=3, height=2.1, unit="in")
+
+ggplot(for_plot, aes(x=pct_correct, y=mean_rt))+
+  geom_point()+
+  scale_color_viridis(discrete=T)+
+  labs(x="Fraction words selected correctly",
+       y="Mean Reaction Time (ms)",
+       color="?s correct")+
+  coord_cartesian(xlim=c(0.4,1), ylim=c(0,1600))+
+    #scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
+  guides(color=FALSE)
+
+ggsave("../error.jpg", width=5, height=3, unit="in")
 

### 2020-05-27 Analysis/nat_stories_lms.Rmd
gitignore

@@
 data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  select(subject, word_num, word, rt, sentence, type)
+
+data_post_only <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  filter(!after_mistake==0) %>% 
     select(subject, word_num, word, rt, sentence, type)
@@
 
+labelled_post_only <- data_post_only %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
+  filter(word_num>1) %>% 
+select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
+         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
+  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
+  select(-Word_In_Story_Num, -Story_Num)
 ```
+
 ```{r}
@@
 
+ggsave("../Papers/gam2.pdf", width=3.5, height=2, unit="in")
+
+for_graph_3 <- labelled_post_only %>% 
+  select(rt, txl_surp, ngram_surp, grnn_surp, past_txl, past_grnn, past_ngram) %>%
+  pivot_longer(cols=`txl_surp`:`past_ngram`,names_to="label", values_to="data") %>% 
+  mutate(type=case_when(
+         label %in% c("past_ngram", "past_txl", "past_grnn")~"Previous",
+         T~"Current"),
+         model=case_when(
+           label %in% c("txl_surp", "past_txl")~"TXL",
+           label %in% c("grnn_surp", "past_grnn")~"GRNN",
+           T ~ "5-gram"
+         ))
+
+ggplot(for_graph_3, aes(x=data, y=rt))+
+         geom_smooth(method="gam",formula= y~s(x, bs="cr", k=20),color="black")+
+         facet_grid(type~model)+
+        labs(x="Surprisal (bits)", y="Reaction Time (ms)")
+
+ggsave("../Papers/gam3.pdf", width=3.5, height=2, unit="in")
+
 ```

### 2020-05-27 Papers/lab_meeting_2020.tex
gitignore

@@
 
-In Boyce et al. (2020) we replicate Witzel et al. (2012)
+Test by replicating Witzel et al. (2012)
 \begin{itemize}
@@
 
-\begin{frame}{Can we use LMs to choose distractors?} Idea: Use high surprisal according to LM as a proxy for really bad in context
+\begin{frame}{Can we use LMs to choose distractors?}
+
+Use high surprisal according to LM as a proxy for bad in context
 \medskip
@@
 \begin{itemize}
-	\item Restrict to a list of possible distractors
-	\item Only consider distractors of same length, frequency as target wrd
-	\item Check distractors until we find one with high surprisal
-	\item Match to real word punctuation
+	\item Restrict to a list of possible distractors \pause
+	\item Only consider distractors of same length, frequency as target word \pause
+	\item Check distractors until we find one with high surprisal \pause
 \end{itemize}
@@
 
+\section{Experiment 2}
+
 \begin{frame}{Potential improvements}
 
-{\large Option 1: Choose better distractors.}
+{\large Option 1: Choose better distractors.} \pause
 \begin{itemize}
-	\item Tweak thresholds of distractor selection.
-	\item Could search for optimal thresholds.
+	\item Tweak thresholds of distractor selection. \pause
+	\item Could search for optimal thresholds. \pause
 	\item Could manually check and fix distractors. 
@@
 
-{\large Option 2: Be forgiving.}
+{\large Option 2: Be forgiving.} \pause
 \begin{itemize}
-	\item Just don't terminate on mistakes.
+	\item Just don't terminate on mistakes. \pause
 	\item Instead have participant correct mistake and continue.
-	\item Easy to implement in Ibex framework.
 	\end{itemize}
@@
 
-{\large The ``solution'' to our problems!}
+{\large The ``solution'' to our problems!} \pause
 
@@
 			\begin{itemize}
-				\item Can be toggled in Ibex
-				\item Have all the data
+				\item Easily implemented as a toggle in Ibex \pause
+				\item Have all the data \pause
 				\item Identify bad participants v bad distractors 
@@
 
-\section{Experiment 2} %TODO: Needs work from here on!
+ %TODO: Needs work from here on!
 
@@
 Use 3 LMs to estimate surprisal: smoothed 5-gram, Gulordava RNN, Transformer-XL (Gulordava et al. 2018, Dai et al. 2019)
+\pause
 
-\includegraphics[width=.9\textwidth]{gam.pdf}
-
-Only on single token words.
+Fit GAMs on pre-error data.
+\begin{itemize}
+	\item Limit to single-token words.
+	\item Fit to both current and past word surprisal.
+\end{itemize}
 
@@
 
+\begin{frame}{Surprisal Effects}
+\includegraphics[width=.9\textwidth]{gam.pdf}	
+\end{frame}
 \begin{frame}{What about post-mistake data?}
@@
 
-\begin{frame}{Result Summary}
-\begin{itemize}
-	\item People will read in the Maze task for 15-20 minutes
-	\item It's possible to comprehend during Maze
-	\item Distractors are generally good enough
-	\item Find expected RT patterns
+\begin{frame}{Summary}
+\begin{itemize}\pause
+	\item People will read in the Maze task for 15-20 minutes \pause
+	\item It's possible to comprehend during Maze \pause
+	\item Distractors are generally good enough \pause
+	\item Find expected RT patterns \pause
 	\item Very little spillover
@@
 \begin{frame}{Use A-maze!}
-Easy to use!
+Easy to use! \pause
 \begin{itemize}
-	\item Runs on command line 
-	\item Match distractors across minimal pair sentences
-	\item Customize surprisal thresholds, vocabulary lists
+	\item Runs on command line \pause
+	\item Match distractors across minimal pair sentences \pause
+	\item Customize surprisal thresholds, vocabulary lists \pause
 	\item Can output pre-formatted for Ibex
-	\item Lots of documentation at vboyce@github.io/Maze
 \end{itemize}
@@
 
-\begin{frame}{Use A-maze!}
-Contribute to A-maze!:
+\begin{frame}{Use A-maze!} \pause
+Contribute to A-maze:
 \begin{itemize}
-	\item Written in Python 3
-	\item Add more language models
-	\item Add more frequency sources
+	\item Written in Python 3 \pause
+	\item Interface with other language models \pause
+	\item Add more frequency sources \pause
 	\item Extend to non-English languages
@@
 
-\textcolor{ForestGreen}{\large Documentation: vboyce@github.io/Maze}
+\textcolor{ForestGreen}{\large Documentation: vboyce.github.io/Maze}
 
@@
 \item Sample task: syntaxgym.org:666
+
+\item Paper: psyarxiv.com/b7nqd/
 \end{itemize}

### 2020-05-28 Papers/lab_meeting_2020.tex
more slides

@@
 	\item Only consider distractors of same length, frequency as target word \pause
-	\item Check distractors until we find one with high surprisal \pause
+	\item Check distractors until we find one with high surprisal
 \end{itemize}
@@
 \begin{frame}{Does A-maze work? }
-
+\pause
 Test it on materials from Witzel et al. (2012)
@@
 
+\pause
 Try with two pre-trained LSTM models (Gulordava 2018, Jozefowicz 2016)
@@
 A lot of mistakes on the second word of the sentence ("The \ul{dog} ...") \pause
-\includegraphics{../graph_errors2.pdf} \pause
+\includegraphics{graph_errors2.pdf} \pause
 Bad participants or bad distractors?
@@
 \medskip
-\pause
 \end{frame}
@@
 
+\begin{frame}{Regression coefficients}
+\begin{tiny}
+\begin{tabular}{l|rlr|rlr|rlr}
+	\hline
+	&\multicolumn{3}{c|}{5-gram}&\multicolumn{3}{c|}{GRNN}&\multicolumn{3}{c}{TXL}\\
+	& Est & CI & $p$ & Est & CI & $p$ &Est & CI & $p$ \\ 
+	\hline
+	Intercept & 865.3 & [829.9, 902.9] & 0.00 & 871.1 & [837.9, 905.3] & 0.00 & 870.8 & [832.5, 907.8] & 0.00 \\ 
+	Surprisal & 11.7 & [9.3, 14.1] & 0.00 & 23.7 & [21, 26.5] & 0.00 & 18.5 & [16.1, 21.1] & 0.00 \\ 
+	Frequency & -2.9 & [-6.3, 0.5] & 0.10 & 2.9 & [-0.2, 6] & 0.06 & 0.4 & [-2.7, 3.5] & 0.79 \\ 
+	Length & 20.5 & [15.4, 25.6] & 0.00 & 18.5 & [13.3, 23.7] & 0.00 & 21.4 & [16.2, 26.6] & 0.00 \\ 
+	Surprisal:Length & -2.0 & [-3, -1] & 0.00 & -1.8 & [-2.7, -0.9] & 0.00 & -1.4 & [-2.2, -0.6] & 0.00 \\
+	Freq:Length & -1.0 & [-2.5, 0.4] & 0.16 & -0.1 & [-1.2, 1] & 0.82 & 0.2 & [-0.9, 1.2] & 0.76 \\ 
+	\hline
+	Past Surprisal & 1.6 & [-0.5, 3.6] & 0.14 & 2.7 & [0.8, 4.5] & 0.00 & 0.8 & [-0.9, 2.5] & 0.40 \\ 
+	Past Freq & 2.6 & [-0.1, 5.4] & 0.06 & 1.9 & [-0.2, 4.2] & 0.08 & 1.2 & [-1.1, 3.6] & 0.30 \\ 
+	Past Length & -4.8 & [-9, -0.1] & 0.04 & -6.6 & [-10.9, -2.1] & 0.00 & -5.2 & [-9.3, -0.7] & 0.03 \\ 
+	Past Surp:Length & -0.2 & [-1.2, 0.8] & 0.72 & -0.9 & [-1.7, -0.2] & 0.01 & -0.6 & [-1.3, 0.2] & 0.13 \\ 
+	Past Freq:Length & -1.0 & [-2.3, 0.3] & 0.15 & -1.8 & [-2.9, -0.8] & 0.00 & -1.5 & [-2.6, -0.5] & 0.01 \\
+
+	\hline
+\end{tabular}
+\end{tiny}
+\end{frame}
+
 \end{document}

### 2020-06-08 Analysis/nat_stories.Rmd
checked how long comp question took to answer (involved redoing how compq data was arranged)

--- 2020-04-30 Analysis/nat_stories.Rmd
@@
 library(knitr)
-library(viridis)
 library(cowplot)
@@
   mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
-  group_by(subject, critical, practice) %>%
+  group_by(subject, num_correct) %>%
   filter(type!="practice") %>% 
@@
             pct_correct=mean(correct.num)) %>% 
-  mutate(critical_factor=as.character(critical*6),
-         good_comp=ifelse(critical>.8,"5 or 6 correct","4 or fewer correct"))
+  mutate(good_comp=ifelse(num_correct>4,"5 or 6 correct","4 or fewer correct"))
 
-ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=critical_factor))+
+ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=as.factor(num_correct)))+
   geom_point()+
@@
 for_plot <- data_error_summ %>% 
-  mutate(num_correct=case_when(
-    critical_factor %in% c("1","2","3") ~ "3 or fewer",
-    T ~ critical_factor)) %>% 
+  # mutate(num_correct=case_when(
+  #   critical_factor %in% c("1","2","3") ~ "3 or fewer",
+  #   T ~ critical_factor)) %>% 
     mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct"))
@@
   geom_point(size=1)+
-  scale_color_viridis(discrete=T)+
   labs(x="Fraction words selected correctly",
@@
 
-some <- data_filt %>% select(subject, critical) %>% 
+ggplot(for_plot, aes(x=pct_correct, y=mean_rt))+
+  geom_point()+
+  labs(x="Fraction words selected correctly",
+       y="Mean Reaction Time (ms)",
+       color="?s correct")+
+  coord_cartesian(xlim=c(0.4,1), ylim=c(0,1600))+
+    #scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
+  guides(color=FALSE)
+
+ggsave("../error.jpg", width=5, height=3, unit="in")
+
+some <- data_filt %>% select(subject, num_correct) %>% 
   unique() %>% 
-  left_join(data_error_summ, by=c("subject", "critical")) %>% 
+  left_join(data_error_summ, by=c("subject", "num_correct")) %>% 
   mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct")) %>% 
-  mutate(correct=critical*6) %>% 
-  group_by(correct,accurate)
+  group_by(num_correct,accurate)
 
-ggplot(some, aes(x=correct,fill=accurate))+
+ggplot(some, aes(x=num_correct,fill=accurate))+
   geom_bar(position="dodge")+
@@
 
-b <- cor(some$pct_correct, some$correct)
+```
 
+## Are comprehension questions working
+
+As was raised when I presented this at lab meeting, it could be that participants are looking up answers to comprehension questions, so we want to check based on length.
+
+```{r}
+
+correct <- data_filt %>% select(subject, num_correct, correct_answer_1:correct_answer_6) %>% unique() %>% 
+  pivot_longer(correct_answer_1:correct_answer_6, names_to=c("q_num"), values_to=c("is.correct"), names_prefix="correct_answer_")
+
+time <- data_filt %>% select(subject, num_correct, q_time_1:q_time_6) %>% unique() %>% 
+  pivot_longer(q_time_1:q_time_6, names_to=c("q_num"), values_to=c("q_time"), names_prefix="q_time_") %>% 
+  left_join(correct, by=c("subject", "num_correct","q_num")) %>% 
+  left_join(data_error_summ, by=c("subject","num_correct")) %>% 
+      mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct")) %>% 
+  filter(q_time<20000)
+
+ggplot(time, aes(x=q_time))+geom_density()+facet_grid(.~accurate)
+
+time_filt <- time %>%  filter(q_time<10000) %>% 
+  group_by(subject, accurate) %>% summarize(mod_pct_correct=sum(is.correct), mean_pct_correct=mean(is.correct))
+
+ggplot(time_filt, aes(x=mod_pct_correct))+geom_histogram()+facet_grid(.~accurate)
 ```
+
+Can play around with cutoffs etc, but the net result is that most of the answering is happening within 10 (much within 5) seconds, and it doesn't look like people are looking up answers, so comprehension is probably reliable.
 
@@
   mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
-  group_by(subject, critical, practice) %>%
+  group_by(subject, num_correct) %>%
   summarize(total_rt=sum(total_rt)/60000,
             pct_correct=mean(correct.num)) %>% 
-  mutate(critical_factor=as.character(critical*6),
-         good_comp=ifelse(critical>.8,"5 or 6 correct","4 or fewer correct"))
+  mutate(         good_comp=ifelse(num_correct>4,"5 or 6 correct","4 or fewer correct"))
 
-ggplot(total_time, aes(x=pct_correct, y=total_rt, color=critical_factor))+
+ggplot(total_time, aes(x=pct_correct, y=total_rt, color=as.factor(num_correct)))+
   geom_point()+
@@
 
-all <- data_filt %>% select(subject, critical) %>% 
+all <- data_filt %>% select(subject, num_correct) %>% 
   unique() %>% 
-  mutate(correct=critical*6) %>% 
+  mutate(correct=num_correct) %>% 
    group_by(correct) %>% 
@@
 
-some <- data_filt %>% select(subject, critical) %>% 
+some <- data_filt %>% select(subject, num_correct) %>% 
   unique() %>% 
-  left_join(data_error_summ, by=c("subject", "critical")) %>% 
+  left_join(data_error_summ, by=c("subject", "num_correct")) %>% 
   filter(pct_correct>.8) %>% 
-  mutate(correct=critical*6) %>% 
+  mutate(correct=num_correct) %>% 
    group_by(correct) %>% 
@@
 
-c <- data_filt %>% filter(type!=practice) %>%
+c <- data_filt %>% filter(type!="practice") %>%
   group_by(subject, sentence) %>%
@@
 data_good <- data_filt %>% 
-  left_join(data_error_summ, by=c("subject", "critical")) %>% 
+  left_join(data_error_summ, by=c("subject", "num_correct")) %>% 
   filter(type!="practice") %>% 

### 2020-06-08 Analysis/read_results.R

@@ -1,5 +1,4 @@
-library(tidyverse)
-library(stringr)
+ library(tidyverse)
 library(urltools)
 #function reads in data from a maze file plus demographic, other questions.
 read_in_data <- function(filename){
@@ -46,13 +45,14 @@ read_in_data <- function(filename){
       correct_answer=col_integer(),
       q_time=col_integer()
     )) %>% 
-    group_by(time, MD5, type) %>% 
-    summarize(pct_correct=mean(correct_answer)) %>% 
-    ungroup() %>% 
-    mutate(question_type=ifelse(type=="practice_question", "practice", "critical")) %>% 
-    select(-type) %>% 
-    pivot_wider(names_from="question_type", values_from="pct_correct")
-  
+     group_by(time, MD5, type) %>%
+       mutate(question_num=row_number(),
+              question_num=ifelse(type=="practice_question", str_c("prac_",question_num),str_c("",question_num))) %>% 
+              ungroup() %>% 
+    select(-question, -answer, -type) %>% 
+    pivot_wider(names_from=question_num, values_from=c(correct_answer,q_time)) %>% 
+    mutate(num_correct=(correct_answer_1+correct_answer_2+correct_answer_3+correct_answer_4+correct_answer_5+correct_answer_6))
+    
   
   #take the Maze task results, relabel and type appropriately
   maze<- filter(data, controller=="Maze") %>% 
@@ -75,9 +75,9 @@ read_in_data <- function(filename){
   maze
 }
 
-#Note: more comprehension questions data, such as time to answer exists, but we don't write it here b/c I don't care about it
-natural_stories_1 <- read_in_data("../Data/raw_data") %>% 
+#note: more comprehension questions data, such as time to answer exists, but we don't write it here b/c i don't care about it
+natural_stories_1 <- read_in_data("../Data/raw_data") %>%
   mutate(subject=paste(MD5, time),
-         subject=factor(subject, levels=unique(subject), labels=1:length(unique(subject)))) %>% 
-  select(-MD5, -time) %>% 
+         subject=factor(subject, levels=unique(subject), labels=1:length(unique(subject)))) %>%
+  select(-MD5, -time) %>%
   write_rds("../Data/cleaned.rds")

### 2020-07-20 Analysis/nat_stories_gam.Rmd
with new gam

--- 2020-04-24 Analysis/nat_stories_lms.Rmd
@@
 ---
-title: "Natural Stories LM analysis"
+title: "Natural Stories GAM analysis"
 output: github_document
@@
 library(tidymv)
+library(rsample)
 theme_set(theme_bw())
@@
   mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
-  group_by(subject, critical, practice) %>%
+  group_by(subject) %>%
   filter(type!="practice") %>% 
@@
 
-We sometimes center the predictors, but don't rescale them. (We center them for the brm models, but don't center surprisal for the GAM model).
-
-The model formula we use is rt ~ surp + length * freq + past_surp + past_length * past_freq, so we're allowing frequency and length to interact and including the same terms for the current and past word.
+For GAM models, we center length and frequency but not surprisal.
+We want surprisal interpretable, but we also will be plotting it (at least for the bootstrapping) at length and frequencies set to 0, so they need to be centered.
+(Not sure this last piece is actually true/matters).
 
@@
 
-```
-
-## Linear models
-
-These use centered predictors.
-
-Using pre-error data only.
-
-```{r lm-2}
-
-lm_ngram_model_1 <- lm(rt ~ ngram_center + freq_center * length_center+ past_c_ngram + past_c_freq*past_c_length, data=labelled_pre_error)
-
-lm_txl_model_1 <- lm(rt ~ txl_center + freq_center * length_center+ past_c_txl + past_c_freq*past_c_length, data=labelled_pre_error)
-
-lm_grnn_model_1 <- lm(rt ~ grnn_center + freq_center * length_center+ past_c_grnn + past_c_freq*past_c_length, data=labelled_pre_error)
-
-```
-
-```{r lm_summary-2}
-
-summary(lm_ngram_model_1)
-
-summary(lm_txl_model_1)
-
-summary(lm_grnn_model_1)
-
-```
-Including post-error data.
-
-```{r lm-3}
-
-lm_ngram_model_2 <- lm(rt ~ ngram_center + freq_center * length_center+ past_c_ngram + past_c_freq*past_c_length, data=labelled_post_error)
-
-lm_txl_model_2 <- lm(rt ~ txl_center + freq_center * length_center+ past_c_txl + past_c_freq*past_c_length, data=labelled_post_error)
-
-lm_grnn_model_2 <- lm(rt ~ grnn_center + freq_center * length_center+ past_c_grnn + past_c_freq*past_c_length, data=labelled_post_error)
-
-```
-
-```{r lm_summary-3}
-
-summary(lm_ngram_model_2)
-
-summary(lm_txl_model_2)
-
-summary(lm_grnn_model_2)
-
-```
-
-## BRM models
-
-We include a by-subject effect for everything.
-Technically speaking, we do have repeated measures at the word level, but there are 10K or so words and at most a handful of copies of each word, so that's likely to get us not very much at the cost of making the model rather intractable (and so I'm not going to deal with it, at least not right now).
-
-Priors:
-
-- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
-- normal(0,500) for beta and sd -- we don't really know what effects are
-- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way
-
-```{r}
-  priors <- c(
-      set_prior("normal(1000, 1000)", class="Intercept"),
-      set_prior("normal(0, 500)", class="b"),
-      set_prior("normal(0, 500)", class="sd"),
-      set_prior("lkj(1)",       class="cor"))
-
- brm_ngram_model <- brm(rt ~ ngram_center + freq_center * length_center +
-                         past_c_ngram + past_c_freq*past_c_length +
-                         (ngram_center + freq_center * length_center +
-                            past_c_ngram + past_c_freq*past_c_length|subject), data=labelled_pre_error,
-                       file="brm_ngram_pre_center", prior=priors, control=list(adapt_delta=.95))
-
-brm_grnn_model <- brm(rt ~ grnn_center + freq_center * length_center +
-                         past_c_grnn + past_c_freq*past_c_length +
-                         (grnn_center + freq_center * length_center +
-                            past_c_grnn + past_c_freq*past_c_length|subject), data=labelled_pre_error,
-                       file="brm_grnn_pre_center", prior=priors, control=list(adapt_delta=.95))
-
-brm_txl_model <- brm(rt ~ txl_center + freq_center * length_center +
-                         past_c_txl + past_c_freq*past_c_length +
-                         (txl_center + freq_center * length_center +
-                            past_c_txl + past_c_freq*past_c_length|subject), data=labelled_pre_error,
-                       file="brm_txl_pre_center", prior=priors, control=list(adapt_delta=.95))
-
-```
-
-```{r}
-
-show_summary <- function(model){
-  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
-
-  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
-    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
-    group_by(.variable) %>% 
-    summarize(pct_above_0=mean(above_0)) %>% 
-    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
-    left_join(intervals, by=".variable") %>% 
-    mutate(lower=round(.lower, digits=1),
-           upper=round(.upper, digits=1),
-           E=round(.value, digits=1),
-           `CI`=str_c("[",lower,", ", upper,"]"),
-           Term=str_sub(.variable, 3, -1),
-           ) %>% 
-    select(Term, `E`, `CI`,`P`)
-
-  stats
-}
-```
-```{r}
-summary(brm_txl_model)
-summary(brm_grnn_model)
-summary(brm_ngram_model)
-```
-
-```{r}
-a <- show_summary(brm_txl_model) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
-b <- show_summary(brm_grnn_model) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
-c <- show_summary(brm_ngram_model) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
-
-summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`E`,`CI`,`P`)) %>% 
-  select(Term, `E_5-gram`, `CI_5-gram`, `P_5-gram`, `E_GRNN`, CI_GRNN, P_GRNN, E_TXL, CI_TXL, P_TXL)
-
-library(xtable)
-print(xtable(summ,digits=c(0,0,1,0,2,1,0,2,1,0,2)), include.rownames=FALSE)
+labelled_post_only <- data_post_only %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
+  filter(word_num>1) %>% 
+select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
+         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
+  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
+  select(-Word_In_Story_Num, -Story_Num)
 ```
@@
 
-I think there might be ways to have mixed effects for this, but for now we're not.
-Also, couldn't figure out good ways of getting graphs from the full models, so just fitting the pieces within ggplot for graphs.
-
 Using only pre-error data.
 
-```{r gams-2}
-gam_ngram_1 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)
+Does not have hierarchical effects.
 
-gam_txl_1 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)
+```{r for gam}
 
-gam_grnn_1 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)
+ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            )
+
+grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            )
+
+txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_surp, prev_surp=past_txl, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            )
 
@@
 
-```{r}
-for_graph <- labelled_pre_error %>% 
-  select(rt, txl_surp, ngram_surp, grnn_surp, past_txl, past_grnn, past_ngram) %>%
-  pivot_longer(cols=`txl_surp`:`past_ngram`,names_to="label", values_to="data") %>% 
-  mutate(type=case_when(
-         label %in% c("past_ngram", "past_txl", "past_grnn")~"Previous",
-         T~"Current"),
-         model=case_when(
-           label %in% c("txl_surp", "past_txl")~"TXL",
-           label %in% c("grnn_surp", "past_grnn")~"GRNN",
-           T ~ "5-gram"
-         ))
+```{r gams-2}
+gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")
 
-ggplot(for_graph, aes(x=data, y=rt))+
-         geom_smooth(method="gam",formula= y~s(x, bs="cr", k=20),color="black")+
-         facet_grid(type~model)+
+gam_grnn_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_data, method="REML")
+
+gam_txl_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_data, method="REML")
+
+```
+
+```{r plot-gam}
+
+a <- get_gam_predictions(model=gam_ngram_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
+b <- get_gam_predictions(model=gam_ngram_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
+c <- get_gam_predictions(model=gam_grnn_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
+d <- get_gam_predictions(model=gam_grnn_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
+e <- get_gam_predictions(model=gam_txl_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
+f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
+
+all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f)
+
+ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+  geom_line()+
+  geom_ribbon(alpha=.3)+
+  facet_grid(s~model)+
   labs(x="Surprisal (bits)", y="Reaction Time (ms)")
@@
 
-```
-
-Including post-error data.
-
-```{r gams-3}
-gam_ngram_2 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)
-
-gam_txl_2 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)
-
-gam_grnn_2 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)
+#confirm that this really is just a prettied up version of what the usual output is
+#plot(gam_txl_1, seWithMean = TRUE, shift = coef(gam_txl_1)[1], select=1)
 
 ```
-
-```{r plot_gam-3}
-for_graph_2 <- labelled_post_error %>% 
-  select(rt, txl_surp, ngram_surp, grnn_surp, past_txl, past_grnn, past_ngram) %>%
-  pivot_longer(cols=`txl_surp`:`past_ngram`,names_to="label", values_to="data") %>% 
-  mutate(type=case_when(
-         label %in% c("past_ngram", "past_txl", "past_grnn")~"Previous",
-         T~"Current"),
-         model=case_when(
-           label %in% c("txl_surp", "past_txl")~"TXL",
-           label %in% c("grnn_surp", "past_grnn")~"GRNN",
-           T ~ "5-gram"
-         ))
-
-ggplot(for_graph_2, aes(x=data, y=rt))+
-         geom_smooth(method="gam",formula= y~s(x, bs="cr", k=20),color="black")+
-         facet_grid(type~model)+
-        labs(x="Surprisal (bits)", y="Reaction Time (ms)")
-
-```

### 2020-07-20 Analysis/nat_stories_gam.Rmd
new gams

@@
 
-## GAMs
+### GAMs
 
+## Try 1
 Using only pre-error data.
 
-Does not have hierarchical effects.
+Does not have hierarchical effects. Has smooths for the surprisals, and tensor effects/interactions for the freq x length
 
@@
 
-```{r gams-2}
+```{r gams}
 gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")

### 2020-07-20 Analysis/nat_stories_lms.Rmd
new gams

--- 2020-04-30 Analysis/nat_stories_lms.Rmd
@@
 
-We sometimes center the predictors, but don't rescale them. (We center them for the brm models, but don't center surprisal for the GAM model).
+We center the predictors for the brms models, but don't rescale them.
 
@@
 
-ggplot(select_labs, aes(length))+geom_bar()
-
+lab_summary
 ```
@@
 
-## GAMs
-
-I think there might be ways to have mixed effects for this, but for now we're not.
-Also, couldn't figure out good ways of getting graphs from the full models, so just fitting the pieces within ggplot for graphs.
-
-Using only pre-error data.
-
-```{r gams-2}
-gam_ngram_1 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)
-
-gam_txl_1 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)
-
-gam_grnn_1 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)
-
-```
+## Log likelihoods
 
 ```{r}
-for_graph <- labelled_pre_error %>% 
-  select(rt, txl_surp, ngram_surp, grnn_surp, past_txl, past_grnn, past_ngram) %>%
-  pivot_longer(cols=`txl_surp`:`past_ngram`,names_to="label", values_to="data") %>% 
-  mutate(type=case_when(
-         label %in% c("past_ngram", "past_txl", "past_grnn")~"Previous",
-         T~"Current"),
-         model=case_when(
-           label %in% c("txl_surp", "past_txl")~"TXL",
-           label %in% c("grnn_surp", "past_grnn")~"GRNN",
-           T ~ "5-gram"
-         ))
-
-ggplot(for_graph, aes(x=data, y=rt))+
-         geom_smooth(method="gam",formula= y~s(x, bs="cr", k=20),color="black")+
-         facet_grid(type~model)+
-        labs(x="Surprisal (bits)", y="Reaction Time (ms)")
-
-ggsave("../Papers/gam.pdf", width=3.5, height=2, unit="in")
-
+# a <- log_lik(brm_ngram_interact)
+# 
+# a2 <- add_criterion(brm_ngram_interact, "waic")
+# 
+# b <- log_lik(brm_grnn_interact)
+# 
+# c <- log_lik(brm_txl_interact)
 ```
-
-Including post-error data.
-
-```{r gams-3}
-gam_ngram_2 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)
-
-gam_txl_2 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)
-
-gam_grnn_2 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)
-
-```
-
-```{r plot_gam-3}
-for_graph_2 <- labelled_post_error %>% 
-  select(rt, txl_surp, ngram_surp, grnn_surp, past_txl, past_grnn, past_ngram) %>%
-  pivot_longer(cols=`txl_surp`:`past_ngram`,names_to="label", values_to="data") %>% 
-  mutate(type=case_when(
-         label %in% c("past_ngram", "past_txl", "past_grnn")~"Previous",
-         T~"Current"),
-         model=case_when(
-           label %in% c("txl_surp", "past_txl")~"TXL",
-           label %in% c("grnn_surp", "past_grnn")~"GRNN",
-           T ~ "5-gram"
-         ))
-
-ggplot(for_graph_2, aes(x=data, y=rt))+
-         geom_smooth(method="gam",formula= y~s(x, bs="cr", k=20),color="black")+
-         facet_grid(type~model)+
-        labs(x="Surprisal (bits)", y="Reaction Time (ms)")
-
-```
-
-# TODOs for future
-
-In terms of surprisal analyses, the possible next steps are doing GAM analysis more robustly (with mixed effects, presenting results from the models) and doing DTSR.
-
-We could also try looking at the specific syntactic features that are in the dataset (the point of natural stories was to have a bunch of idioms and weird syntax); we could maybe analyse those?

### 2020-07-21 Analysis/nat_stories_gam.Rmd
add density to the gam plot

@@
 library(rsample)
+library(cowplot)
+library(scales)
 theme_set(theme_bw())
@@
                                             prev_freq=past_c_freq, prev_len=past_c_length
-                                            )
+                                            ) %>% mutate(model="5-gram")
 
@@
                                             prev_freq=past_c_freq, prev_len=past_c_length
-                                            )
+                                            )%>% mutate(model="GRNN")
 
@@
                                             prev_freq=past_c_freq, prev_len=past_c_length
-                                            )
+                                            ) %>% mutate(model="TXL")
 
+all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% 
+  select(surprisal, prev_surp,model) %>% 
+  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
+  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
 ```
@@
 f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
+```
 
+```{r}
 all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f)
 
-ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+gam1 <-  ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+
   geom_ribbon(alpha=.3)+
+   #geom_density(data=all_data, aes(x=value), fill="gray",)+
   facet_grid(s~model)+
-  labs(x="Surprisal (bits)", y="Reaction Time (ms)")
+  coord_cartesian(ylim=c(700,1300), xlim=c(0,28))+
+  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
 
-ggsave("../Papers/gam.pdf", width=3.5, height=2, unit="in")
+dens2 <-   ggplot(all_data, aes(x=value))+
+  geom_density(fill="gray",)+
+  facet_grid(.~model)+
+  labs(x="Surprisal (bits)", y="")+
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+bot <- plot_grid(NA, dens2, NA, nrow=1, rel_widths = c(.16, 1, .08))
+
+ plot_grid(gam1, bot, nrow=2, rel_heights = c(1, .3))
+
+ggsave("../Papers/gam.pdf", width=4, height=2.3, unit="in")
 

### 2020-07-21 Analysis/nat_stories_lms.Rmd
add density to the gam plot

@@
 ```{r}
-# a <- log_lik(brm_ngram_interact)
-# 
+ a <- log_lik(brm_ngram_interact)
+
+a2 <- exp(a*-1) %>% mean()
 # a2 <- add_criterion(brm_ngram_interact, "waic")
 # 
-# b <- log_lik(brm_grnn_interact)
-# 
-# c <- log_lik(brm_txl_interact)
+ b <- log_lik(brm_grnn_interact)
+
+ b2 <- exp(a*-1) %>% mean()
+ c <- log_lik(brm_txl_interact)
+ c2 <- exp(c*-1) %>% mean()
+
+ product <- sum(a)
+
+ a2 <- a*-1+product
 ```

### 2020-07-23 Analysis/nat_stories_lms.Rmd
add attempt at loglik

--- 2020-07-20 Analysis/nat_stories_lms.Rmd
@@
 library(knitr)
-library(mgcv)
-library(mgcViz)
-library(tidymv)
+library(matrixStats)
 theme_set(theme_bw())
@@
 ```{r}
-summary(brm_ngram_interact_post)
-summary(brm_ngram_interact)
-
-summary(brm_grnn_interact_post)
-summary(brm_grnn_interact)
-
-summary(brm_txl_interact_post)
-summary(brm_txl_interact)
+# summary(brm_ngram_interact_post)
+# summary(brm_ngram_interact)
+# 
+# summary(brm_grnn_interact_post)
+# summary(brm_grnn_interact)
+# 
+# summary(brm_txl_interact_post)
+# summary(brm_txl_interact)
 ```
@@
 ```{r}
-# a <- log_lik(brm_ngram_interact)
-# 
-# a2 <- add_criterion(brm_ngram_interact, "waic")
-# 
-# b <- log_lik(brm_grnn_interact)
-# 
-# c <- log_lik(brm_txl_interact)
+
+#given a vector of logs, find harmonic mean
+#uses one of the formulae from wikipedia page so we have underflow and not overflow issues
+#then use logSumExp to reduce underflow
+get_harmonic <- function(d){
+  l <- length(d)
+  log_prod <- sum(d) #this is the log of the product 
+  log_inv <- d*-1 # this is list of logs of inverses
+  log_inv_prod <- log_inv+log_prod #we "multiply" by the product, all of these are now teeny
+  log_sum <- logSumExp(log_inv_prod) #we sum these up, still logged
+  overall <- l * exp(log_prod-log_sum)
+
+  overall
+}
+
+ a <- log_lik(brm_ngram_interact)
+
+ b <- log_lik(brm_grnn_interact)
+
+ c <- log_lik(brm_txl_interact)
+
+ get_harmonic(a)
+
+ get_harmonic(b)
+
+ get_harmonic(c)
 ```
+
+I get 9.087e-32 for Ngram, 3.602e-32 for GRNN, and 2.24e-33 for TXL. I'm not sure this is the order we expect, and not sure how to interpret it. (Also not sure I did things right.)

### 2020-07-28 Analysis/nat_stories_lms.Rmd
add nested model comparison, anovas

@@
   filter(word_num>1) %>% 
-  select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
+  select(rt, subject, Word_In_Story_Num, word, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
          txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
@@
 
- a <- log_lik(brm_ngram_interact)
-
- b <- log_lik(brm_grnn_interact)
-
- c <- log_lik(brm_txl_interact)
-
- get_harmonic(a)
-
- get_harmonic(b)
-
- get_harmonic(c)
+ # a <- log_lik(brm_ngram_interact)
+ # 
+ # b <- log_lik(brm_grnn_interact)
+ # 
+ # c <- log_lik(brm_txl_interact)
+ # 
+ # get_harmonic(a)
+ # 
+ # get_harmonic(b)
+ # 
+ # get_harmonic(c)
 ```
@@
 I get 9.087e-32 for Ngram, 3.602e-32 for GRNN, and 2.24e-33 for TXL. I'm not sure this is the order we expect, and not sure how to interpret it. (Also not sure I did things right.)
+
+# Frequentist LMs
+
+```{r}
+#mean by subject for each word
+d_lm <- labelled_pre_error %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
+                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
+  summarize(mean_rt=mean(rt))
+```
+
+## Models
+
+We want to be able to do nested model comparision, so we want models with 
+- only length, frequency fx
+- each 1 surprisal model as predictor
+- all the surprisals
+
+For now, no interactions between surprisal and others, and if we include a predictor, include both current and lagged of it.
+
+```{r}
+
+no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+all_surp <- lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   grnn_center + past_c_grnn +
+                   txl_center + past_c_txl +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+ngram_grnn <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   grnn_center + past_c_grnn +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+ngram_txl <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   txl_center + past_c_txl +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+grnn_txl <-  lmer(mean_rt ~ grnn_center + past_c_grnn +
+                   txl_center + past_c_txl +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+no_lag <-  lmer(mean_rt ~ ngram_center +
+                   grnn_center +
+                   txl_center + 
+                   freq_center*length_center+(1|word), data=d_lm)
+
+no_surp_lag <-  lmer(mean_rt ~ ngram_center +  
+                   grnn_center + 
+                   txl_center +                    freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+  only_surp_lag <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   grnn_center + past_c_grnn +
+                   txl_center + past_c_txl +
+                   freq_center*length_center+(1|word), data=d_lm)
+```
+
+```{r}
+
+anova(all_surp, txl_only)
+anova(all_surp, grnn_only)
+anova(all_surp, ngram_only)
+
+anova(txl_only, no_surp)
+anova(grnn_only, no_surp)
+anova(ngram_only, no_surp)
+
+anova(all_surp, ngram_grnn)
+anova(all_surp, ngram_txl)
+anova(all_surp, grnn_txl)
+
+anova(ngram_grnn, ngram_only)
+anova(ngram_grnn, grnn_only)
+anova(ngram_txl, ngram_only)
+anova(ngram_txl, txl_only)
+anova(grnn_txl, grnn_only)
+anova(grnn_txl, txl_only)
+
+anova(all_surp, no_lag)
+
+anova(all_surp, no_surp_lag)
+
+anova(no_surp_lag, no_lag)
+
+anova(all_surp, only_surp_lag)
+
+anova(only_surp_lag, no_lag)
+```
+In general, adding to the model makes it better.
+
+The model with all 3 surprisal predictors is better than any model with only one. Any one surprisal predictor is better than no surprisals predictors.
+
+However, adding ngram predictors to a model that already has txl & grnn does not help. 
+In other cases, adding the 3rd surprisal source to the other two does help.
+
+Ngram+Grnn is not better than Grnn only. Otherwise, pairs are better than singletons. This suggests that Ngram's info is a subset of GRNN, but not a subset of TXL.
+
+Past surprisal predictors don't help (with or without past freq,length effects in the models), but past freq,length do (with or without past surprisal predictors).

### 2020-07-28 Papers/amlap2020_abstract.tex
new gams
add nested model comparison, anovas

@@
 \newtoggle{anonymous}
-\toggletrue{anonymous}
+\togglefalse{anonymous}
 
@@
 \title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
-\authors{Veronica Boyce (MIT), Roger Levy (MIT)} %TODO, check what affiliation I have at conference time
-\email{vboyce@mit.edu}
+\authors{Veronica Boyce (MIT/Stanford), Roger Levy (MIT)} %TODO, check what affiliation I have at conference time
+\email{vboyce@stanford.edu}
 \newline
@@
 Across all models, we see large effects of current word surprisal and length, which noticeably exceed effects of previous word predictors (Table~1).
+We see minimal to no frequency effects (consistent with Shain 2019).
+In a nested model comparison, we found GRNN and TXL encoded complementary information, but adding 5-gram surprisals did not improve a regression that already had GRNN.
 We also fit GAMs to look at the shape of the relationship between surprisal and RT; we find a linear relationship between RT and current word surprisal and no relationship with previous word surprisal (Figure~3).
@@
 Overall, this suggests that A-maze is a versatile alternative to eye-tracking and self-paced reading.
-
-Our code and data are available at [link to repo redacted for anonymity].
+Code and data at github.com/vboyce/amaze-natural-stories.
 
@@
 			Gulordava et al (2018).  \textit{NAACL-HLT 2018} $\bullet$
+			Shain (2019). \textit{NAACL-HLT 2019} $\bullet$
 			Sloggett et al (2020). \textit{CUNY 2020} $\bullet$
 			Smith \& Levy (2013). \textit{Cognition} $\bullet$
-			Witzel et al (2012).  \textit{J Psycholinguist Res} 
-			%TODO add cites for length, freq, surprisal justification
+			Witzel et al (2012).  \textit{J Psycholinguist Res.} 
 		}

### 2020-08-25 Analysis/nat_stories_gam.Rmd
try models with weak data exclusions

@@
 ```
+
+## Permissive
+
+```{r }
+
+labelled_any <- read_rds("non-selective.rds")
+
+ngram_any<- labelled_any %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            ) %>% mutate(model="5-gram")
+
+grnn_any <- labelled_any %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            )%>% mutate(model="GRNN")
+
+txl_any <- labelled_any %>% select(rt, surprisal=txl_surp, prev_surp=past_txl, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            ) %>% mutate(model="TXL")
+
+all_any <- ngram_any %>% union(grnn_any) %>% union(txl_any) %>% 
+  select(surprisal, prev_surp,model) %>% 
+  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
+  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
+```
+
+```{r }
+gam_ngram_2 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_any, method="REML")
+
+gam_grnn_2 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_any, method="REML")
+
+gam_txl_2 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_any, method="REML")
+
+```
+
+```{r}
+
+a2 <- get_gam_predictions(model=gam_ngram_2, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
+b2 <- get_gam_predictions(model=gam_ngram_2, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
+c2 <- get_gam_predictions(model=gam_grnn_2, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
+d2 <- get_gam_predictions(model=gam_grnn_2, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
+e2 <- get_gam_predictions(model=gam_txl_2, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
+f2 <- get_gam_predictions(model=gam_txl_2, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
+```
+
+```{r}
+all_any_result <- a2 %>% union(b2) %>% union(c2) %>% union(d2) %>% union(e2) %>% union(f2)
+
+gam2 <-  ggplot(all_any_result, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+  geom_line()+
+  geom_ribbon(alpha=.3)+
+   #geom_density(data=all_data, aes(x=value), fill="gray",)+
+  facet_grid(s~model)+
+  coord_cartesian(ylim=c(700,1300), xlim=c(0,28))+
+  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
+
+dens_any <-   ggplot(all_any, aes(x=value))+
+  geom_density(fill="gray",)+
+  facet_grid(.~model)+
+  labs(x="Surprisal (bits)", y="")+
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+bot <- plot_grid(NA, dens_any, NA, nrow=1, rel_widths = c(.16, 1, .08))
+
+ plot_grid(gam2, bot, nrow=2, rel_heights = c(1, .3))
+
+ggsave("../Papers/gam-any.pdf", width=4, height=2.3, unit="in")
+
+#confirm that this really is just a prettied up version of what the usual output is
+#plot(gam_txl_1, seWithMean = TRUE, shift = coef(gam_txl_1)[1], select=1)
+
+```

### 2020-08-25 Analysis/nat_stories_lms.Rmd
try models with weak data exclusions

@@
   tally()
+
+data_anything_goes <- data_filt %>% 
+  filter(type!="practice") %>% 
+  filter(correct=="yes") %>% 
+    select(subject, word_num, word, rt, sentence, type)
+
 ```
@@
   select(-Word_In_Story_Num, -Story_Num)
+
+labelled_anything_goes <- data_anything_goes %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
+  filter(word_num>1) %>% 
+select(rt, subject, Word_In_Story_Num, word, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
+         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
+  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
+  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("non-selective.rds")
 ```
@@
 Past surprisal predictors don't help (with or without past freq,length effects in the models), but past freq,length do (with or without past surprisal predictors).
+
+## What if we don't exclude participants
+
+One potentially useful comparison with past SPR/eye-tracking is if we are more inclusive of participants, do we get data that looks more like theirs. (Perhaps our exclusion criteria are more robust?)
+
+```{r freq_all}
+#mean by subject for each word
+d_lm2 <- labelled_anything_goes %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
+                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
+  summarize(mean_rt=mean(rt))
+```
+
+```{r}
+
+no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
+
+ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
+
+grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
+
+txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
+
+```

### 2020-08-25 Papers/Amlap_2020_talk/amlap_2020_talk.tex
draft of talk

--- 2020-05-28 Papers/lab_meeting_2020.tex
@@
 \beamertemplatenavigationsymbolsempty 
-\setbeamertemplate{mini frames}{}
+\setbeamertemplate{headline}{}
 \setbeamertemplate{itemize item}{\textbullet}
@@
 }
+\addtobeamertemplate{frametitle}{\vspace*{.5cm}}{\vspace*{.5cm}}
+
 \setbeamercolor{footline}{fg=black}
@@
 
-\title{A-maze: Easier measurement of incremental processing}
-\author{Veronica Boyce}
-\date{3 June 2020}
+\title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
+\author{Veronica Boyce, Roger Levy}
+\date{AMLaP 2020}
 
@@
 \maketitle
+%TODO maybe put a picture of Roger's face??, at least acknowledge him 
+%Pre-amble -- thank Roger, outline with be methods heavy, and then talk about the specific experiment & how it's relevant to how we think about surprisal 
 \end{frame}
 
-%
+% But first -- Incremental processing methods
 
-\section{Why RT?}
-\begin{frame}{Why measure reading time?}
+\begin{frame}{Incremental processing methods}
 \pause
-Linguistic and psycholinguistic theories make predictions about processing difficulty.
-\medskip
-\pause
+Goal: Measure processing difficulty
 
-Examples of increased difficulty
-\begin{itemize}\pause
-	\item Constructions not in the grammar \pause
-	\item Lexical items that are harder to retrieve \pause
-	\item Words that force reparsing or reanalysis
-\end{itemize}
-\pause
+Assumption: Longer RT = more difficulty
 \medskip
 
-We assume that harder processing manifests in longer reading/reaction time (RT).
 \pause
-\medskip
-
-RT patterns may be phenomena that theories need to explain.
-
 \end{frame}
 
-\begin{frame}{Two common methods}
-%can't get at the internals of processing, so want to use time as proxy for how much is going on 
-%but literate humans are good at reading 
+\begin{frame}{Incremental processing methods}
+	How to measure RT?
+
+Usual Methods
 \begin{columns}[t]\pause
@@
 
-			\includegraphics[width=.9\textwidth]{eye-tracker.jpeg}
+			\includegraphics[width=.9\textwidth]{../eye-tracker.jpeg}
 			\pause
-			\begin{itemize}
-				\item Expensive
-				\item Hard to analyse
-			\end{itemize}
 		\end{center}
@@
 \onslide<8->{
-			\begin{itemize}
-	\item Lots of spillover
-	\item Messy data 
-\end{itemize}
 }
@@
 \end{columns}
+\bigskip
+Different methods have different trade-offs
 \end{frame}
 
-\begin{frame}{A third option: Maze}
-\end{frame}
-\begin{frame}{Maze Task}
+\begin{frame}{An alternative: Maze}
+%TODO First mentioned, then more elaborated in 2009 paper, has been used ~dozen times 
 \centering
@@
 	font={\fontsize{14pt}{12}\selectfont}}
+%TODO Maybe update with circles to show the selections. 
 \begin{tikzpicture}
-\onslide<1->{\node[draw] (a) at (-1,-.5) {e};}
-\onslide<1->{\node[draw] (b) at (1,-.5) {i};}
 \onslide<1-2>{\node  (c) at (-1,.5) {The};}
@@
 \onslide<9-10>{\node  (d) at (1,.5) {squirrel.};}
-\onslide<2,8>{\node[draw, fill=pink!50] at (a) {e};}
-\onslide<4,6,10>{\node[draw, fill=pink!50] at (b) {i};}
+\onslide<2,8>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
+\onslide<4,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
+align=center] at (d) {};}
 \end{tikzpicture}
@@
 
-\begin{frame}{A third option: Maze}
+\begin{frame}{An alternative: Maze}
 \begin{columns}
@@
 		\textbf{\large G-maze}
-		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{gmaze.pdf}
+		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../gmaze.pdf}
 		\end{center}
@@
 		\textbf{\large L-maze}
-			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{lmaze.pdf}
+			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../lmaze.pdf}
 		\end{center}
@@
 \pause
-Sentence ends when a mistake is made.
+Sentence ends if a mistake is made.
 \pause
 
-Central claim: forces extremely incremental processing (no spillover)
+Claim: forces incremental processing (no spillover)
 
@@
 \end{flushright}
+% Compelling results especially for G-maze, but 
+% was run in the lab, and generating G-maze materials is a huge amount of work
 \end{frame}
 
-\section{Experiment 1a}
-\begin{frame}{Web Implementation}
-Maze well suited to run on the web
+\begin{frame}{Using Maze}
+Is Maze a viable alternative to web SPR?
+
+Needs:
 \begin{itemize}
-	\item Implement in Ibex
+	\item Run on web 
+	\item Easy generation of distractor words
+	\item Work for long or multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
 \end{itemize} 
@@
 
-Test by replicating Witzel et al. (2012)
-\begin{itemize}
-	\item Witzel et al (2012): Comparison of eye-tracking, SPR, L-maze, G-maze (all in-lab)
-	\item Got materials and data from Witzel
-	\item We run SPR, L-maze, and G-maze on MTurk
-\end{itemize}
 \end{frame}
 
-\begin{frame}{Materials}
+\begin{frame}{Run on web}
+	%TODO include screenshot
+	Wrote an Ibex module
+	\begin{itemize}
+		\item Based off SPR module
+		\item Fits in with existing workflow
+	\end{itemize}
+ Replicated Witzel et al. (2012) results (see Boyce et al. 2020)
+\end{frame}
+%
+\begin{frame}{Using Maze}
+	Is Maze a viable alternative to web SPR?
 
-\textbf{Relative Clause}
-
-\sethlcolor{green}
-\textit{Low:} The son of the \uline{lady} who politely introduced \hl{herself} was popular at the party.
-\sethlcolor{pink}
-
- \textit{High:} The \uline{son} of the lady who politely introduced \hl{himself} was popular at the party.
-
-\textbf{Adverb Clause}
-
-\sethlcolor{green}\textit{Low:} James will fix the car he \uline{drove} \hl{yesterday}, but he will need some help.
-
-\sethlcolor{pink} \textit{High:} James \uline{will fix} the car he drove \hl{tomorrow}, but he will need some help.
-
-\textbf{Sentence v Noun Phrase conjunction}
-
-\sethlcolor{green} \textit{Comma:} The swimmer disappointed her \uline{coach}, and her mother \hl{tried} to console her.
-
-\sethlcolor{pink}
-\textit{No comma:} The swimmer disappointed her \uline{coach} and her mother \hl{tried} to console her.
-\end{frame}
-
-\begin{frame}{Results}
-\begin{small}	
-	\sethlcolor{green}
-The son of the lady who politely introduced \hl{herself}\sethlcolor{pink} / \hl{himself} was popular at the party.
-
-\end{small}
-\includegraphics[width=\textwidth]{g_rel.pdf}
-\end{frame}
-
-\begin{frame}{Results}
-\begin{small}
-
-\sethlcolor{green}James will fix the car he drove \hl{yesterday}\sethlcolor{pink} / \hl{tomorrow},  but he will need some help.
-
-\end{small}
-\includegraphics[width=\textwidth]{g_adv.pdf}
-\end{frame}
-
-\begin{frame}{Results}
-\begin{small}	
-\sethlcolor{green}The swimmer disappointed her coach\hl{,} and her mother \hl{tried} / \sethlcolor{pink}\hl{tried} to console her.
-\end{small}
-\includegraphics[width=\textwidth]{g_svnp.pdf}
-\end{frame}
-
-\begin{frame}{Interim Summary}
-\textcolor{PineGreen}{\Large \textbf{Good news:} G-maze works well over the web (better than L-maze or SPR)}
-
+	Needs:
+	\begin{itemize}
+		\item Run on web 
+		\item Easy generation of distractor words
+		\item Work for long or multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
+	\end{itemize} 
 	\pause
 
-\bigskip
+\end{frame}
+%
+\begin{frame}{Generating distractors}
+\pause
+Task: Given  \textit{The dog chased} Find a word that can't be a continuation.
+\begin{itemize}
+	\item Tedious (and hard!) to do by hand. 
+\end{itemize}
 
-\textcolor{Red}{\Large \textbf{Bad news:} I do not want to write G-maze materials.}
+What makes something an unacceptable continuation?
+\begin{itemize}
+	\item Ungrammatical
+	\item ...or otherwise really unlikely 
+	\item $\approx$ high surprisal 
+\end{itemize}
 
+Can we use Neural Language Models?
 \end{frame}
-
-\section{A-maze}
-
-\begin{frame}{Meanwhile in Natural Language Processing}
-\pause
+%
+\begin{frame}{Can we use LMs?}
 Language models (LMs)
@@
 \end{itemize}
-\pause
-Surprisal: negative log probability
+
+LSTM RNNs (ex. Gulordava 2018) show some syntactic knowledge %TODO cite people
+
+Add some controls:
 \begin{itemize}
-	\item 2 bits of surprisal = 1/4
-	\item 10 bits of surprisal $\approx$ 1/1000 
-	\item +1 surprisal = half as likely
-\end{itemize}
-\pause
-\begin{tikzpicture}
-\node (A) at (0,0) {"The dog chased ..."};
-\node  (B) at (4,0) [draw, minimum width=1.5cm, minimum height=1.5cm, fill=black] {\textcolor{white}{LM}} ;
-\draw [->, thick] (A) -- (B); 
-\node[align=left] (C) at (7,0) {"the" 1.2 bits \\"a" 3.1 bits\\"cats" 8 bits\\...\\"wish" 25 bits\\...};
-\draw [->, thick] (B) -- (C); 
-\end{tikzpicture}
-\end{frame}
-
-\begin{frame}{Can we use LMs to choose distractors?}
-
-Use high surprisal according to LM as a proxy for bad in context
-\medskip
-
-\pause
-
-\begin{itemize}
-	\item Model the target sentence word by word \pause
-	\item At each position, choose a high surprisal word
-\end{itemize}
-\medskip
-\pause
-Want quality control on distractors
-\pause
-\begin{itemize}
-	\item Restrict to a list of possible distractors \pause
-	\item Only consider distractors of same length, frequency as target word \pause
-	\item Check distractors until we find one with high surprisal
+\item Restrict to a list of possible distractors
+\item Only consider length, frequency matches
 \end{itemize}
@@
 \end{frame}
+%
+\begin{frame}{Can we use LMs to choose distractors?}
 
-\begin{frame}
-\centering
-\includegraphics[clip, trim=3.25cm 5cm 2.5cm 2.5cm,width=.9\textwidth]{flow_2.pdf}
-\end{frame}
-\section{Experiment 1b}
+Does it work?
 
-\begin{frame}{Does A-maze work? }
-\pause
-Test it on materials from Witzel et al. (2012)
+Yes, at least well enough.
+\begin{itemize}
+	\item Sometimes generates plausible distractors.
+	\item Overall results comparable with G-maze %TODO citations to me, Sloggett
+	%TODO show graph from Boyce et al
+\end{itemize} 
 \medskip
 
+Parameter-tweaking may help, as would hand-checking. 
+\end{frame}
+%
+\begin{frame}{Using Maze}
+	Is Maze a viable alternative to web SPR?
+
+	Needs:
+	\begin{itemize}
+		\item Run on web %TODO figure out how to make checkmarks again
+		\item Easy generation of distractor words 
+		\item Work for long or multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
+	\end{itemize} 
 	\pause
-Try with two pre-trained LSTM models (Gulordava 2018, Jozefowicz 2016)
 
 \end{frame}
+%
+%
+\begin{frame}{Long items}
 
-\begin{frame}{Does A-maze work?  Yes.} \pause
-\includegraphics[width=1\textwidth]{a_all.pdf}
-\end{frame}
+Errors terminate sentences. 
+Some error rate is inevitable.
 
-\begin{frame}{Caveats}
-A lot of mistakes on the second word of the sentence ("The \ul{dog} ...") \pause
-\includegraphics{graph_errors2.pdf} \pause
-Bad participants or bad distractors?
+Want to run multi-sentence items. 
+\begin{itemize}
+	\item Treat whole story as a unit: Almost no one makes it to the end.
+	\item Treat each sentence as a unit:
+	Many participants miss key context. 
+\end{itemize}
 
-Either way, lots of data loss.
-\end{frame}
-
-\begin{frame}{Caveats}
-
-{\large Definitely some bad distractors}
-\begin{table}
-
-		\begin{tabular}{rlll}
-			Prefix & Correct & Distractor & Error Rate \\
-			\hline
-			\hline
-			Gulordava&&&\\
-			\hline
-			The & niece & cooks & 44\%\\
-			The swimmer & disappointed & propositions & 30\%\\
-			The & semester & steroids & 29\%\\
-			\hline
-			\hline
-			 Jozefowicz&&&\\
-			\hline
-			The & husband & authors & 46\%\\
-			Jim & listened & survived & 43\%\\
-			The & uncle & roads & 42\%\\
-			The & knight & saints & 40\%\\
-		\end{tabular}
-\end{table}
+What if after an error, participants corrected errors and the sentence continued?
 
 \end{frame}
-
-\section{Experiment 2}
-
-\begin{frame}{Potential improvements}
-
-{\large Option 1: Choose better distractors.} \pause
-\begin{itemize}
-	\item Tweak thresholds of distractor selection. \pause
-	\item Could search for optimal thresholds. \pause
-	\item Could manually check and fix distractors. 
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}{Potential improvements}
-
-{\large Option 2: Be forgiving.} \pause
-\begin{itemize}
-	\item Just don't terminate on mistakes. \pause
-	\item Instead have participant correct mistake and continue.
-	\end{itemize}
-\end{frame}
-
+%
 \begin{frame}{Maze with Error Correction}
@@
 	font={\fontsize{14pt}{12}\selectfont}}
+
 \begin{tikzpicture}
-\onslide<1->{\node[draw] (a) at (-1,-.5) {e};}
-\onslide<1->{\node[draw] (b) at (1,-.5) {i};}
 \onslide<1-2>{\node  (c) at (-1,.5) {The};}
@@
 \onslide<11-12>{\node  (d) at (1,.5) {squirrel.};}
-\onslide<2,6,10>{\node[draw, fill=pink!50] at (a) {e};}
-\onslide<4,8,12>{\node[draw, fill=pink!50] at (b) {i};}
+\onslide<2,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
+\onslide<4,8,12>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
+	align=center] at (d) {};}
 \onslide<7,8>{\node[text=red] at (0,-1.5) {\small Incorrect. Please try again.};}
 \end{tikzpicture}
+
 \end{frame}
@@
 
-{\large The ``solution'' to our problems!} \pause
-
 \begin{columns}
@@
 		\begin{center}
-		\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=\textwidth]{maze_diagram.pdf}
+		\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=\textwidth]{../maze_diagram.pdf}
 		\end{center}
@@
 		\begin{center}
+			Features
 			\begin{itemize}
 				\item Easily implemented as a toggle in Ibex \pause
+				\item Can look at long materials (in principle)
 				\item Have all the data \pause
 				\item Identify bad participants v bad distractors 
+				%easy to weed out participants who aren't paying attention 
+				\item Makes it less of a big deal when we have bad distractors
 			\end{itemize}
@@
 \end{frame}
-
- %TODO: Needs work from here on!
+%
+%%So now that I've explained the method, going to switch over to current experiment & what we hoped to get from it
+%
+\begin{frame}
+	Various open questions we want to address -- really test suitability of Maze
+	\begin{itemize}
+		\item Do people comprehend/remember what they read?
+		\item Does this error correction thing work? 
+		\item Are people willing to read texts in this way? 
+		\item Do we get predictability effects? (Both SPR, eye-tracking show certain surprisal effects, does Maze?)
+	\end{itemize}
+\end{frame}
 
@@
 
-\end{frame}
-
-\begin{frame}{Experiment Plan}
-Experiment questions:
-\begin{itemize}
-	\item Will people do this task?
-	\item Can participants comprehend the stories?
-\end{itemize}
-
-\pause
-
-\medskip
-
-Checks on RT data
-\begin{itemize}
-	\item Check if RT is linear in surprisal (Smith \& Levy 2013)
-	\item Check for known effects of word length, frequency
-	\item Check for spillover effects
-\end{itemize}
+%HARD mode
 
@@
 \begin{frame}{Participant accuracy}
-100 participants each read 1 story \pause
+100 participants from MTurk each read 1 story (20 minutes) \pause
 \begin{center}
-\includegraphics[width=.8\textwidth]{error.pdf}
+\includegraphics[width=.8\textwidth]{../error.pdf}
 \end{center}
@@
 \begin{center} \pause
-\includegraphics[width=.8\textwidth]{comp.pdf}
+\includegraphics[width=.8\textwidth]{../comp.pdf}
 \end{center}
+% exclusion options: use comprehension checks as participant filter.
+Here we can use overall accuracy as the "paying attention" filter.
+Don't have to worry about how to write good comprehension questions.
+May still need to do RT exclusions for double-presses or getting distracted.
 \end{frame}
@@
 \begin{frame}{Surprisal Effects}
-
+%TODO copy from updated abstract
 Use 3 LMs to estimate surprisal: smoothed 5-gram, Gulordava RNN, Transformer-XL (Gulordava et al. 2018, Dai et al. 2019)
 \pause
+
+%TODO make sure that updated correct gams are used!!!
 
@@
 
+\begin{frame}
+	Implications for surprisal theory -- size of the effects is different
+	include effect sizes
+	Is it about participant quality?
+\end{frame}
+
 \begin{frame}{Surprisal Effects}
-\includegraphics[width=.9\textwidth]{gam.pdf}	
+\includegraphics[width=.9\textwidth]{../gam.pdf}	
 \end{frame}
@@
 
-\includegraphics[width=.9\textwidth]{gam2.pdf}	
+\includegraphics[width=.9\textwidth]{../gam2.pdf}	
 \end{frame}
@@
 \end{frame}
-\section{Conclusion}
 
@@
 \end{frame}
-%back pocket on more on model results??
-%do we even need back pockets??
+%%back pocket on more on model results??
+%%do we even need back pockets??
 
@@
 
-\includegraphics[]{gam3.pdf}	
+\includegraphics[]{../gam3.pdf}	
 \end{frame}

### 2020-08-26 Papers/Amlap_2020_talk/amlap_2020_talk.tex
more work on talk

@@
-\documentclass[12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman]{beamer}
+\documentclass[12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman,t]{beamer}
 \usetheme{Frankfurt}
@@
 
+\newcommand{\bluecheck}{}%
+\DeclareRobustCommand{\greencheck}{%
+	\tikz\fill[scale=0.6, color=ForestGreen]
+	(0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;%
+}
+
 \title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
@@
 %TODO maybe put a picture of Roger's face??, at least acknowledge him 
-%Pre-amble -- thank Roger, outline with be methods heavy, and then talk about the specific experiment & how it's relevant to how we think about surprisal 
 \end{frame}
-
-% But first -- Incremental processing methods
 
@@
 \pause
-Goal: Measure processing difficulty
+Want a measure of processing difficulty \greencheck\pause
 
-Assumption: Longer RT = more difficulty
-\medskip
+Assume that longer RT = more difficulty \pause
+\smallskip
+\begin{center}
+{\large How to measure RT?}
+\end{center}
+\end{frame}
+\begin{frame}{Incremental processing methods}
 
-\pause
-\end{frame}
-
-\begin{frame}{Incremental processing methods}
-	How to measure RT?
-
-Usual Methods
+{\large Common ways to measure RT}
 \begin{columns}[t]\pause
@@
 		\begin{center}
-			\textbf{\Large Eye-tracking}
+			\textbf{\large Eye-tracking}
 
@@
 		\begin{center}
-\textbf{\Large Self-paced reading}
+\textbf{\large Self-paced reading}
 
@@
 
-\begin{tikzpicture} %TODO on slide etc
+\begin{tikzpicture}
 \draw (-.5,-.5) rectangle (3.2,.5);
@@
 \end{center}
-\onslide<8->{
-}
 \end{column}
@@
 \bigskip
-Different methods have different trade-offs
+\onslide<8->{Different methods have different trade-offs}
 \end{frame}
 
-\begin{frame}{An alternative: Maze}
-%TODO First mentioned, then more elaborated in 2009 paper, has been used ~dozen times 
+\begin{frame}[c]{An alternative: Maze}
+
+\pause 
 \centering
+
+\bigskip
+
 \tikzset{
 	font={\fontsize{14pt}{12}\selectfont}}
-%TODO Maybe update with circles to show the selections. 
 \begin{tikzpicture}
@@
 \begin{frame}{An alternative: Maze}
+\vskip -1em
+{\small(Forster et al. 2009; Witzel et al. 2012)}
+	\vskip -1em
 \begin{columns}
@@
 		\begin{center}
-		\textbf{\large G-maze}
+		\textbf{\large G-maze}\\
+		`Grammatical' choices\\
 		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../gmaze.pdf}
@@
 		\begin{center}
-		\textbf{\large L-maze}
+		\textbf{\large L-maze}\\
+		`Lexical' choices\\
 			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../lmaze.pdf}
@@
 Claim: forces incremental processing (no spillover)
-
-\begin{flushright}
-	{\small(Forster et al. 2009; Witzel et al. 2012)}
-\end{flushright}
 % Compelling results especially for G-maze, but 
@@
 
-\begin{frame}{Using Maze}
-Is Maze a viable alternative to web SPR?
+\begin{frame}{Maze Made Easy}
 
-Needs:
+	Want to use Maze instead of web SPR\pause
+
+	\medskip
+
+	Needs some tweaks:\pause
 	\begin{itemize}
-	\item Run on web 
-	\item Easy generation of distractor words
-	\item Work for long or multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
+		\item Run on web \pause
+		\item Easily generate distractors \pause
+		\item Work for multi-sentence items \pause
 	\end{itemize}
-\pause
 
@@
 \begin{frame}{Run on web}
-	%TODO include screenshot
-	Wrote an Ibex module
-	\begin{itemize}
-		\item Based off SPR module
-		\item Fits in with existing workflow
-	\end{itemize}
- Replicated Witzel et al. (2012) results (see Boyce et al. 2020)
+	Wrote an Ibex module \pause
+
+	\begin{center}\includegraphics[width=.7\textwidth]{screenshot.png} \end{center}\pause
+
+ Replicated Witzel et al. (2012) results (Boyce et al. 2020)
 \end{frame}
 %
-\begin{frame}{Using Maze}
-	Is Maze a viable alternative to web SPR?
+\begin{frame}{Maze Made Easy}
 
-	Needs:
+	Want to use Maze instead of web SPR
+
+	\medskip
+
+	Needs some tweaks:
 	\begin{itemize}
-		\item Run on web 
-		\item Easy generation of distractor words
-		\item Work for long or multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
+		\item Run on web \greencheck
+		\item Easily generate distractors
+		\item Work for multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
 	\end{itemize}
-	\pause
 
@@
 \pause
-Task: Given  \textit{The dog chased} Find a word that can't be a continuation.
+Goal: Find a word that can't continue a partial sentence
 \begin{itemize}
-	\item Tedious (and hard!) to do by hand. 
-\end{itemize}
+	\item Ex.  \textit{The dog chased} \pause
+	\item Tedious (and hard!) to do by hand
+\end{itemize} \pause
 
-What makes something an unacceptable continuation?
+What makes something an unacceptable continuation? \pause
 \begin{itemize}
-	\item Ungrammatical
-	\item ...or otherwise really unlikely 
+	\item Ungrammatical \pause
+	\item ...or otherwise really unlikely \pause 
 	\item $\approx$ high surprisal 
-\end{itemize}
+\end{itemize} \pause
 
@@
 \begin{frame}{Can we use LMs?}
+	\pause
 Language models (LMs)
 \begin{itemize}
-	\item Trained on large corpora to predict the next word
+	\item Trained to predict the next word
 	\item Given a partial sentence, return probabilities of the next word
-\end{itemize}
+\end{itemize}\pause
 
-LSTM RNNs (ex. Gulordava 2018) show some syntactic knowledge %TODO cite people
+LSTM RNNs show some syntactic knowledge %TODO cite people 
+\pause
 
-Add some controls:
+Add some restrictions:
 \begin{itemize}
@@
 %
-\begin{frame}{Can we use LMs to choose distractors?}
-
-Does it work?
-
-Yes, at least well enough.
+\begin{frame}{Does it work?} 
+\pause
+\vskip -2em
+{\large Yes, at least well enough.} \pause
 \begin{itemize}
-	\item Sometimes generates plausible distractors.
-	\item Overall results comparable with G-maze %TODO citations to me, Sloggett
-	%TODO show graph from Boyce et al
+	\item Sometimes generates plausible distractors. \pause
+	\item A-maze results comparable with G-maze {\small(Boyce et al 2020, Sloggett et al 2020)} \pause 
 \end{itemize} 
 \medskip
+\centering
 
-Parameter-tweaking may help, as would hand-checking. 
+\includegraphics[width=.7\textwidth]{../a_all.pdf}
 \end{frame}
 %
-\begin{frame}{Using Maze}
-	Is Maze a viable alternative to web SPR?
+\begin{frame}{Maze Made Easy}
 
-	Needs:
+	Want to use Maze instead of web SPR
+
+	\medskip
+
+	Needs some tweaks:
 	\begin{itemize}
-		\item Run on web %TODO figure out how to make checkmarks again
-		\item Easy generation of distractor words 
-		\item Work for long or multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
+		\item Run on web \greencheck
+		\item Easily generate distractors \greencheck
+		\item Work for multi-sentence items 
 	\end{itemize}
-	\pause
 
@@
 
-Errors terminate sentences. 
-Some error rate is inevitable.
+Want to run multi-sentence items. \pause
 
-Want to run multi-sentence items. 
+Problem: Errors terminate sentences. \pause
 \begin{itemize}
-	\item Treat whole story as a unit: Almost no one makes it to the end.
-	\item Treat each sentence as a unit:
-	Many participants miss key context. 
+	\item Treat whole story as a unit: ~ No one makes it to the end. \pause
+	\item Treat each sentence as a unit: Some participants miss key context. \pause
 \end{itemize}
@@
 %
-\begin{frame}{Maze with Error Correction}
+\begin{frame}[c]{Maze with Error Correction}
 \centering
@@
 \begin{columns}
-	\begin{column}{0.3\textwidth}
+	\begin{column}{0.4\textwidth}
 		\begin{center}
@@
 	\end{column} \pause 
-	\begin{column}{0.7\textwidth} 
+	\begin{column}{0.6\textwidth} 
 		\begin{center}
-			Features
 			\begin{itemize}
-				\item Easily implemented as a toggle in Ibex \pause
-				\item Can look at long materials (in principle)
+				\item Can be toggled in Ibex \pause
+				\item Long materials feasible \pause
 				\item Have all the data \pause
-				\item Identify bad participants v bad distractors 
-				%easy to weed out participants who aren't paying attention 
-				\item Makes it less of a big deal when we have bad distractors
+				\item Compensates for bad distractors
 			\end{itemize}
@@
 \pause
-
-Side effect: Can now run multi-sentence items!
 \end{frame}
 %
+
+\begin{frame}{Maze Made Easy}
+
+	Want to use Maze instead of web SPR
+
+	\medskip
+
+	Needs some tweaks:
+	\begin{itemize}
+		\item Run on web \greencheck
+		\item Easily generate distractors \greencheck
+		\item Work for multi-sentence items \greencheck ? 
+	\end{itemize}
+
+\end{frame}
 %%So now that I've explained the method, going to switch over to current experiment & what we hoped to get from it
 %
-\begin{frame}
-	Various open questions we want to address -- really test suitability of Maze
+\begin{frame}{Current experiment}
+	Various open questions to address \pause
 	\begin{itemize}
-		\item Do people comprehend/remember what they read?
-		\item Does this error correction thing work? 
-		\item Are people willing to read texts in this way? 
-		\item Do we get predictability effects? (Both SPR, eye-tracking show certain surprisal effects, does Maze?)
+		\item Are people willing to read long texts in this way? \pause
+		\item Do people comprehend what they read? \pause
+		\item Does the error correction variant work? \pause
+		\item Do we get predictability effects? 
 	\end{itemize}
@@
 \begin{frame}{Natural Stories}
-Natural stories corpus (Futrell et al. 2017)
+Natural stories corpus (Futrell et al. 2017) \pause
 \begin{itemize}
-	\item 10 stories, each about 1000 words
-	\item Some unusual constructions, but read fluently
-	\item 6 comprehension questions per story
+	\item 10 stories, each about 1000 words \pause
+	\item Some unusual constructions, but read fluently \pause
+	\item 6 comprehension questions per story \pause
 \end{itemize}
@@
 \end{frame}
+
 \begin{frame}{Comprehension questions}
@@
 
+%HERE%
 \begin{frame}{Surprisal Effects}

### 2020-08-26 Analysis/nat_stories_lms.Rmd
add analysis of all data

@@
 
+This is all data from participants who speak English that wasn't a mistake. (So, includes bad participants and post-mistake data.)
+
 ```{r freq_all}
@@
 
+summary(no_surp)
+summary(ngram_only)
+summary(grnn_only)
+summary(txl_only)
+
 ```
+
+BRM models from this same very permissive data. 
+```{r brm-models}
+
+any_grnn <- read_rds("brm_grnn_any_center.rds")
+any_txl <- read_rds("brm_txl_any_center.rds")
+any_ngram <- read_rds("brm_ngram_any_center.rds")
+
+summary(any_grnn)
+summary(any_txl)
+summary(any_ngram)
+
+```
+
+```{r}
+
+show_summary <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=1),
+           upper=round(.upper, digits=1),
+           E=round(.value, digits=1),
+           `CI`=str_c("[",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           ) %>% 
+    select(Term, `E`, `CI`,`P`)
+
+  stats
+}
+a <- show_summary(any_txl) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
+b <- show_summary(any_grnn) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
+c <- show_summary(any_ngram) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
+
+summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`E`,`CI`,`P`)) %>% 
+  select(Term, `E_5-gram`, `CI_5-gram`, `P_5-gram`, `E_GRNN`, CI_GRNN, P_GRNN, E_TXL, CI_TXL, P_TXL)
+
+library(xtable)
+print(xtable(summ,digits=c(0,0,1,0,2,1,0,2,1,0,2)), include.rownames=FALSE)
+```

### 2020-08-27 Papers/Amlap_2020_talk/amlap_2020_talk.tex
work on talk

@@
 
-\begin{frame}
-\maketitle
+\begin{frame}{ }
+\vskip 6em
+
+\begin{center}
+{\LARGE A-maze of Natural Stories:}
+
+\smallskip
+
+{\Large Texts are comprehensible using\\ \smallskip the Maze task}
+
+\bigskip
+
+{\large Veronica Boyce, Roger Levy}
+
+\bigskip
+
+{\large AMLaP 2020}
+
+\begin{tikzpicture}[remember picture,overlay]
+\node[xshift=-1.8cm,yshift=-1cm] at (current page.east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{roger.jpg}};
+\end{tikzpicture}
+
+\begin{tikzpicture}[remember picture,overlay]
+\node[xshift=2.5cm,yshift=-1.5cm] at (current page.north west) {\includegraphics[width=.3\textwidth]{cpl_header.png}};
+\end{tikzpicture}
+
 %TODO maybe put a picture of Roger's face??, at least acknowledge him 
+\end{center}
 \end{frame}
@@
 \pause
-Want a measure of processing difficulty \greencheck\pause
+Want a measure of processing difficulty\pause
 
@@
 		\end{center}
-	\end{column}\pause
+	\end{column}
 	\begin{column}{.5\textwidth}
@@
 
-\pause 
 \centering
@@
 
-	Want to use Maze instead of web SPR\pause
+	Can we use Maze instead of web SPR?\pause
 
@@
 		\item Easily generate distractors \pause
-		\item Work for multi-sentence items \pause
+		\item Work for multi-sentence items
 	\end{itemize}
@@
 \begin{frame}{Run on web}
+
+	\pause
 	Wrote an Ibex module \pause
@@
 
-	Want to use Maze instead of web SPR
+	Can we use Maze instead of web SPR?
 
@@
 
-LSTM RNNs show some syntactic knowledge %TODO cite people 
+Run items through LM, choose high surprisal words as distractors
 \pause
@@
 \begin{itemize}
-	\item Treat whole story as a unit: ~ No one makes it to the end. \pause
-	\item Treat each sentence as a unit: Some participants miss key context. \pause
+	\item Treat whole story as a unit: \pause ~ No one makes it to the end. \pause
+	\item Treat each sentence as a unit: \pause Some participants miss key context. \pause
 \end{itemize}
@@
 			\begin{itemize}
-				\item Can be toggled in Ibex \pause
+				\item Can be toggled in Ibex Maze \pause
 				\item Long materials feasible \pause
@@
 \end{columns}
-\medskip
-\pause
 \end{frame}
@@
 	\item 10 stories, each about 1000 words \pause
-	\item Some unusual constructions, but read fluently \pause
-	\item 6 comprehension questions per story \pause
+%	\item Some unusual constructions, but read fluently \pause
+	\item 6 comprehension questions per story
 \end{itemize}
@@
 \medskip
+
+\pause
 
@@
 \begin{frame}{Surprisal Effects}
-%TODO copy from updated abstract
-Use 3 LMs to estimate surprisal: smoothed 5-gram, Gulordava RNN, Transformer-XL (Gulordava et al. 2018, Dai et al. 2019)
+\large{Is RT linear in terms of surprisal?}\pause
+\medskip
+
+Estimate surprisal from 3 models:
+\begin{itemize}
+	\item smoothed 5-gram
+	\item LSTM-RNN (Gulordava et al. 2018)
+	\item Transformer-XL (Dai et al. 2019)
+\end{itemize}
+
 \pause
 
-%TODO make sure that updated correct gams are used!!!
-
-Fit GAMs on pre-error data.
+Fit GAMs
 \begin{itemize}
-	\item Limit to single-token words.
-	\item Fit to both current and past word surprisal.
+	\item Limit to single-token words
+	\item Fit to both current and past word surprisal
+	\item Include frequency, length as predictors
 \end{itemize}
 
-\end{frame}
-
-\begin{frame}
-	Implications for surprisal theory -- size of the effects is different
-	include effect sizes
-	Is it about participant quality?
 \end{frame}
@@
 \end{frame}
-\begin{frame}{What about post-mistake data?}
-Exclude data from mistakes or the two words after a mistake.
 
-\includegraphics[width=.9\textwidth]{../gam2.pdf}	
+\begin{frame}{Surprisal Effects}
+	Linear Models \\
+	\pause
+
+	\begin{small}
+		\begin{tabular}{l|r|r|r}
+			\hline
+			&5-gram & GRNN & TXL\\
+			\hline
+			Intercept & \textbf{865.3} & \textbf{871.1} & \textbf{870.8} \\ 
+			Surprisal & \textbf{11.7}  & \textbf{23.7} & \textbf{18.5}  \\ 
+			Frequency & -2.9  & 2.9  & 0.4  \\ 
+			Length & \textbf{20.5} & \textbf{18.5}  & \textbf{21.4} \\ 
+			Surprisal:Length & \textbf{-2.0}  & \textbf{-1.8}  & \textbf{-1.4} \\
+			Freq:Length & -1.0  & -0.1 & 0.2  \\ 
+			\hline
+			Past Surprisal & 1.6  & \textbf{2.7} & 0.8 \\ 
+			Past Freq & 2.6  & 1.9  & 1.2  \\ 
+			Past Length & \textbf{-4.8}  & \textbf{-6.6 }&\textbf{ -5.2} \\ 
+			Past Surp:Length & -0.2 & \textbf{-0.9} & -0.6  \\ 
+			Past Freq:Length & -1.0  & \textbf{-1.8 }&   \textbf{-1.5}  \\ 
+			\hline
+		\end{tabular}
+
+	Surprisal in bits, Length in characters, \\
+	Frequency in $log_2$ occurrences/billion words
+	\end{small}
+
+\end{frame}
+\begin{frame}{Surprisal Effects}
+	Takeaways: \pause
+	\begin{itemize}
+		\item Minimal frequency effects (consistent with Shain 2019) \pause
+		\item Long RTs \pause
+		\item Large effects of Length, Surprisal \pause
+		\item Very little spillover \pause
+	\end{itemize}
+
+	Model comparison: GRNN is best, but TXL complementary
 \end{frame}
 
-\begin{frame}{Summary}
-\begin{itemize}\pause
-	\item People will read in the Maze task for 15-20 minutes \pause
-	\item It's possible to comprehend during Maze \pause
-	\item Distractors are generally good enough \pause
-	\item Find expected RT patterns \pause
-	\item Very little spillover
-\end{itemize}
-\end{frame}
+\begin{frame}{Why such large effects?}
+	\pause
+	Framework: Look at words long enough to ID with some threshold of certainty \pause
 
-\begin{frame}{Use A-maze!}
-Easy to use! \pause
+Could hypothesize mechanisms for difference: \pause
 \begin{itemize}
-	\item Runs on command line \pause
-	\item Match distractors across minimal pair sentences \pause
-	\item Customize surprisal thresholds, vocabulary lists \pause
-	\item Can output pre-formatted for Ibex
-\end{itemize}
-\medskip
-\end{frame}
-
-\begin{frame}{Use A-maze!} \pause
-Contribute to A-maze:
-\begin{itemize}
-	\item Written in Python 3 \pause
-	\item Interface with other language models \pause
-	\item Add more frequency sources \pause
-	\item Extend to non-English languages
+	\item Task leads to higher threshold \pause 
+	\item Task demands reduce available resources for processing \pause
+	\item Presence of second word 
 \end{itemize}
@@
 
-\begin{frame}{}
+\begin{frame}{Conclusion}
 
-\textcolor{ForestGreen}{\large Documentation: vboyce.github.io/Maze}
+	\pause
 
-with links to the following:
+Consider A-maze! 
+\begin{itemize}\pause
+	\item Documentation:  \textcolor{ForestGreen}{vboyce.github.io/Maze} \pause
+	\item Versatile \pause
+	\item Low spillover \pause 
+\end{itemize}
+
+Natural Stories A-maze: \pause 
 \begin{itemize}
+	\item Participants comprehend what they read \pause
+	\item Find linear, large surprisal effects
+\end{itemize}
 
-\item A-maze code: github.com/vboyce/Maze
-
-\item Web-maze code: github.com/vboyce/Ibex-with-Maze
-
-\item Sample task: syntaxgym.org:666
-
-\item Paper: psyarxiv.com/b7nqd/
-\end{itemize}
 \end{frame}
@@
 
-\begin{frame}{Matching distractors}
-If unspecified: Match by position
-\begin{itemize}
-	\item The son of the lady who politely introduced \sethlcolor{green} \hl{herself}\sethlcolor{pink} / \hl{himself} was popular at the party.
-\end{itemize}
-Can specify labels for each word to pair (within item)
-\begin{itemize}
-	\item The cat who the dog scared hid in a box.\\pre-1 pre-2 who art noun verb main-verb post-1 post-2 post-3
-\item The dog who scared the cat sniffed around the couch.\\ pre-1 pre-2 who verb art noun main-verb post-1 post-2 post-3
-\end{itemize}
-\end{frame}
-%%back pocket on more on model results??
-%%do we even need back pockets??
-
-\begin{frame}{Post-mistake only}
-Exclude data from mistakes or the two words after a mistake.
-
-\includegraphics[]{../gam3.pdf}	
-\end{frame}
-
-\begin{frame}{Regression coefficients}
-\begin{tiny}
-\begin{tabular}{l|rlr|rlr|rlr}
-	\hline
-	&\multicolumn{3}{c|}{5-gram}&\multicolumn{3}{c|}{GRNN}&\multicolumn{3}{c}{TXL}\\
-	& Est & CI & $p$ & Est & CI & $p$ &Est & CI & $p$ \\ 
-	\hline
-	Intercept & 865.3 & [829.9, 902.9] & 0.00 & 871.1 & [837.9, 905.3] & 0.00 & 870.8 & [832.5, 907.8] & 0.00 \\ 
-	Surprisal & 11.7 & [9.3, 14.1] & 0.00 & 23.7 & [21, 26.5] & 0.00 & 18.5 & [16.1, 21.1] & 0.00 \\ 
-	Frequency & -2.9 & [-6.3, 0.5] & 0.10 & 2.9 & [-0.2, 6] & 0.06 & 0.4 & [-2.7, 3.5] & 0.79 \\ 
-	Length & 20.5 & [15.4, 25.6] & 0.00 & 18.5 & [13.3, 23.7] & 0.00 & 21.4 & [16.2, 26.6] & 0.00 \\ 
-	Surprisal:Length & -2.0 & [-3, -1] & 0.00 & -1.8 & [-2.7, -0.9] & 0.00 & -1.4 & [-2.2, -0.6] & 0.00 \\
-	Freq:Length & -1.0 & [-2.5, 0.4] & 0.16 & -0.1 & [-1.2, 1] & 0.82 & 0.2 & [-0.9, 1.2] & 0.76 \\ 
-	\hline
-	Past Surprisal & 1.6 & [-0.5, 3.6] & 0.14 & 2.7 & [0.8, 4.5] & 0.00 & 0.8 & [-0.9, 2.5] & 0.40 \\ 
-	Past Freq & 2.6 & [-0.1, 5.4] & 0.06 & 1.9 & [-0.2, 4.2] & 0.08 & 1.2 & [-1.1, 3.6] & 0.30 \\ 
-	Past Length & -4.8 & [-9, -0.1] & 0.04 & -6.6 & [-10.9, -2.1] & 0.00 & -5.2 & [-9.3, -0.7] & 0.03 \\ 
-	Past Surp:Length & -0.2 & [-1.2, 0.8] & 0.72 & -0.9 & [-1.7, -0.2] & 0.01 & -0.6 & [-1.3, 0.2] & 0.13 \\ 
-	Past Freq:Length & -1.0 & [-2.3, 0.3] & 0.15 & -1.8 & [-2.9, -0.8] & 0.00 & -1.5 & [-2.6, -0.5] & 0.01 \\
-
-	\hline
-\end{tabular}
-\end{tiny}
+\begin{frame}{Surprisal Effects}
+	GAM if we only exclude mistakes (all participants, post-mistake data)
+	\includegraphics[width=.9\textwidth]{../gam-any.pdf}	
 \end{frame}

### 2020-08-28 Papers/Amlap_2020_talk/amlap_2020_talk.tex
tweaks to talk

@@
 
-%TODO maybe put a picture of Roger's face??, at least acknowledge him 
 \end{center}
@@
 
+%\begin{frame}{Incremental processing methods}
+%\pause
+%Want a measure of processing difficulty\pause
+%
+%Assume that longer RT = more difficulty \pause
+%\smallskip
+%\begin{center}
+%{\large How to measure RT?}
+%\end{center}
+%\end{frame}
 \begin{frame}{Incremental processing methods}
 	\pause
-Want a measure of processing difficulty\pause
-
-Assume that longer RT = more difficulty \pause
-\smallskip
-\begin{center}
-{\large How to measure RT?}
-\end{center}
-\end{frame}
-\begin{frame}{Incremental processing methods}
-
 {\large Common ways to measure RT}
@@
 Run items through LM, choose high surprisal words as distractors
-\pause
 
-Add some restrictions:
-\begin{itemize}
-\item Restrict to a list of possible distractors
-\item Only consider length, frequency matches
-\end{itemize}
-
+%Add some restrictions:
+%\begin{itemize}
+%\item Restrict to a list of possible distractors
+%\item Only consider length, frequency matches
+%\end{itemize}
+%	
 \end{frame}
@@
 \begin{itemize}
-	\item Treat whole story as a unit: \pause ~ No one makes it to the end. \pause
+	\item Treat whole story as a unit: \pause Few participants make it to the end. \pause
 	\item Treat each sentence as a unit: \pause Some participants miss key context. \pause
@@
 	\begin{itemize}
-		\item Are people willing to read long texts in this way? \pause
-		\item Do people comprehend what they read? \pause
-		\item Does the error correction variant work? \pause
+		\item Will people read long texts in Maze? \pause
+		\item Will they comprehend what they read? \pause
+		\item Does error correction Maze work? \pause
 		\item Do we get predictability effects? 
@@
 \begin{itemize}
-	\item Limit to single-token words
+	%\item Limit to single-token words
 	\item Fit to both current and past word surprisal
@@
 \begin{frame}{Surprisal Effects}
+	\vskip -2em
 	Linear Models \\
@@
 \end{frame}
-\begin{frame}{Surprisal Effects}
-	Takeaways: \pause
-	\begin{itemize}
-		\item Minimal frequency effects (consistent with Shain 2019) \pause
-		\item Long RTs \pause
-		\item Large effects of Length, Surprisal \pause
-		\item Very little spillover \pause
-	\end{itemize}
-
-	Model comparison: GRNN is best, but TXL complementary
-\end{frame}
+%\begin{frame}{Surprisal Effects}
+%	Takeaways: \pause
+%	\begin{itemize}
+%		\item Minimal frequency effects (consistent with Shain 2019) \pause
+%		\item Long RTs \pause
+%		\item Large effects of Length, Surprisal \pause
+%		\item Very little spillover \pause
+%	\end{itemize}
+%	
+%	Model comparison: GRNN is best, but TXL complementary
+%\end{frame}
 
@@
 	\pause
-	Framework: Look at words long enough to ID with some threshold of certainty \pause
+	Bayesian reader framework: Look at words long enough to ID with some threshold of certainty \pause
 
-Could hypothesize mechanisms for difference: \pause
+Possible mechanisms for difference: \pause
 \begin{itemize}
-	\item Task leads to higher threshold \pause 
-	\item Task demands reduce available resources for processing \pause
+	\item Higher threshold \pause 
+	\item Fewer available resources for processing \pause
 	\item Presence of second word 
@@
 
+\begin{frame}
+	\centering
+	\includegraphics[clip, trim=3.25cm 5cm 2.5cm 2.5cm,width=.9\textwidth]{../flow_2.pdf}
+\end{frame}
+
+\begin{frame}{Caveats}
+
+	{\large Definitely some bad distractors}
+	\begin{table}
+
+		\begin{tabular}{rlll}
+			Prefix & Correct & Distractor & Error Rate \\
+			\hline
+			\hline
+			Gulordava&&&\\
+			\hline
+			The & niece & cooks & 44\%\\
+			The swimmer & disappointed & propositions & 30\%\\
+			The & semester & steroids & 29\%\\
+			\hline
+			\hline
+			Jozefowicz&&&\\
+			\hline
+			The & husband & authors & 46\%\\
+			Jim & listened & survived & 43\%\\
+			The & uncle & roads & 42\%\\
+			The & knight & saints & 40\%\\
+		\end{tabular}
+	\end{table}
+
+\end{frame}
+
 \begin{frame}{Surprisal Effects}
@@
 
+\begin{frame}{Links}
+
+	\textcolor{ForestGreen}{\large Documentation: vboyce.github.io/Maze}
+
+	with links to the following:
+	\begin{itemize}
+
+		\item A-maze code: github.com/vboyce/Maze
+
+		\item Web-maze code: github.com/vboyce/Ibex-with-Maze
+
+		\item Sample task: syntaxgym.org:666
+
+		\item Paper: psyarxiv.com/b7nqd/
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Matching distractors}
+	If unspecified: Match by position
+	\begin{itemize}
+		\item The son of the lady who politely introduced \sethlcolor{green} \hl{herself}\sethlcolor{pink} / \hl{himself} was popular at the party.
+	\end{itemize}
+	Can specify labels for each word to pair (within item)
+	\begin{itemize}
+		\item The cat who the dog scared hid in a box.\\pre-1 pre-2 who art noun verb main-verb post-1 post-2 post-3
+		\item The dog who scared the cat sniffed around the couch.\\ pre-1 pre-2 who verb art noun main-verb post-1 post-2 post-3
+	\end{itemize}
+\end{frame}
+
 \end{document}

### 2020-09-01 Analysis/nat_stories_spr.Rmd
add analysis of original SPR nat stories

---
title: "Natural Stories LM analysis"
output: github_document
---
```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
library(matrixStats)
library(mgcv)
library(mgcViz)
library(tidymv)
library(rsample)
library(cowplot)
library(scales)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

```{r read spr}

spr_info <- read_tsv("../Data/SPR/processed_wordinfo.tsv") %>%
  select(word, zone, item) %>% arrange(item, zone) %>% 
  mutate(row=row_number())

labs <- read_rds("../Prep_code/natural_stories_surprisals2.rds") %>% left_join(read_delim("../Materials/natural_stories_sentences.tsv", delim="\t")) %>% 
  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything()) %>% 
  mutate(row=row_number())

spr_labels <- spr_info %>% inner_join(labs, by=c("row", "word")) %>% 
  filter(word_num!=0) %>% 
  filter(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1) %>% 
  filter(!is.na(txl_surp) & !is.na(ngram_surp) & !is.na(grnn_surp) & !is.na(freq) & !is.na(length)) %>% 
    mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
         freq_center=freq-mean(freq, na.rm=T),
         length_center=length-mean(length, na.rm=T))

past_word <- spr_labels %>% 
  mutate(word_num=word_num+1,
         zone=zone+1) %>% 
  rename(past_txl=txl_surp, past_ngram=ngram_surp, past_grnn=grnn_surp, past_freq=freq, past_length=length,past_c_txl=txl_center, past_c_ngram=ngram_center, past_c_grnn=grnn_center, past_c_freq=freq_center, past_c_length=length_center) %>% 
  select(sentence,Story_Num,Sentence_Num, word_num, zone, item, starts_with("past")) %>% 
  left_join(spr_labels, by=c("Story_Num","Sentence_Num","sentence", "word_num", "zone", "item")) %>% 
  filter(!is.na(word)) %>% select(-word)

```

```{r data}
spr_data <- read_tsv("../Data/SPR/processed_RTs.tsv") %>% 
  inner_join(past_word, by=c("zone", "item")) %>% 
  select(word, rt=RT, subject=WorkerId, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("spr.rds")
```

## BRM models

(not yet done)

We include a by-subject effect for everything, and a by_word random intercept (full mixed effects).

Priors:

- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
- normal(0,500) for beta and sd -- we don't really know what effects are
- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way

```{r}
  priors <- c(
      set_prior("normal(1000, 1000)", class="Intercept"),
      set_prior("normal(0, 500)", class="b"),
      set_prior("normal(0, 500)", class="sd"),
      set_prior("lkj(1)",       class="cor"))

```

# Frequentist LMs

```{r}
#mean by subject for each word
d_lm <- spr_data %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
  summarize(mean_rt=mean(rt))
```

## Models

We want to be able to do nested model comparision, so we want models with 
- only length, frequency fx
- each 1 surprisal model as predictor
- all the surprisals

For now, no interactions between surprisal and others, and if we include a predictor, include both current and lagged of it.

```{r}

no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

all_surp <- lmer(mean_rt ~ ngram_center + past_c_ngram + 
                   grnn_center + past_c_grnn +
                   txl_center + past_c_txl +
                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

ngram_grnn <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
                   grnn_center + past_c_grnn +
                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

ngram_txl <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
                   txl_center + past_c_txl +
                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

grnn_txl <-  lmer(mean_rt ~ grnn_center + past_c_grnn +
                   txl_center + past_c_txl +
                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

no_lag <-  lmer(mean_rt ~ ngram_center +
                   grnn_center +
                   txl_center + 
                   freq_center*length_center+(1|word), data=d_lm)

no_surp_lag <-  lmer(mean_rt ~ ngram_center +  
                   grnn_center + 
                   txl_center +                    freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

  only_surp_lag <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
                   grnn_center + past_c_grnn +
                   txl_center + past_c_txl +
                   freq_center*length_center+(1|word), data=d_lm)
```

```{r}

anova(all_surp, txl_only)
anova(all_surp, grnn_only)
anova(all_surp, ngram_only)

anova(txl_only, no_surp)
anova(grnn_only, no_surp)
anova(ngram_only, no_surp)

anova(all_surp, ngram_grnn)
anova(all_surp, ngram_txl)
anova(all_surp, grnn_txl)

anova(ngram_grnn, ngram_only)
anova(ngram_grnn, grnn_only)
anova(ngram_txl, ngram_only)
anova(ngram_txl, txl_only)
anova(grnn_txl, grnn_only)
anova(grnn_txl, txl_only)

anova(all_surp, no_lag)

anova(all_surp, no_surp_lag)

anova(no_surp_lag, no_lag)

anova(all_surp, only_surp_lag)

anova(only_surp_lag, no_lag)
```
## Try 1
Using only pre-error data.

Does not have hierarchical effects. Has smooths for the surprisals, and tensor effects/interactions for the freq x length

```{r for gam}

ngram_data <- spr_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_c_freq, prev_len=past_c_length
                                            ) %>% mutate(model="5-gram")

grnn_data <- spr_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_c_freq, prev_len=past_c_length
                                            )%>% mutate(model="GRNN")

txl_data <- spr_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_c_freq, prev_len=past_c_length
                                            ) %>% mutate(model="TXL")

all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% 
  select(surprisal, prev_surp,model) %>% 
  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
```

```{r gams}
gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")

gam_grnn_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_data, method="REML")

gam_txl_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_data, method="REML")

```

```{r plot-gam}

a <- get_gam_predictions(model=gam_ngram_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
b <- get_gam_predictions(model=gam_ngram_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
c <- get_gam_predictions(model=gam_grnn_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
d <- get_gam_predictions(model=gam_grnn_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
e <- get_gam_predictions(model=gam_txl_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
```

```{r}
all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f)

gam1 <-  ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
  geom_line()+
  geom_ribbon(alpha=.3)+
   #geom_density(data=all_data, aes(x=value), fill="gray",)+
  facet_grid(s~model)+
  coord_cartesian(ylim=c(700,1300), xlim=c(0,28))+
  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))

dens2 <-   ggplot(all_data, aes(x=value))+
  geom_density(fill="gray",)+
  facet_grid(.~model)+
  labs(x="Surprisal (bits)", y="")+
  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))

bot <- plot_grid(NA, dens2, NA, nrow=1, rel_widths = c(.16, 1, .08))

 plot_grid(gam1, bot, nrow=2, rel_heights = c(1, .3))

ggsave("../Papers/gam.pdf", width=4, height=2.3, unit="in")

#confirm that this really is just a prettied up version of what the usual output is
#plot(gam_txl_1, seWithMean = TRUE, shift = coef(gam_txl_1)[1], select=1)

```

### 2020-09-01 Prep_code/nat_stories_prep.Rmd
try models with weak data exclusions
add analysis of original SPR nat stories

@@
 
+all <- txl_translator %>% 
+  left_join(ngram_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
+  left_join(grnn_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>%
+  left_join(freq, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
+  #filter(Word_In_Sentence_Num!=0) %>% 
+  ungroup() %>% 
+  write_rds("natural_stories_surprisals2.rds")
 ```

### 2020-09-04 Analysis/nat_stories_spr.Rmd
talk edits

@@
 ```{r}
+spr_data <- read_rds("spr.rds")
 #mean by subject for each word
@@
 
-ngram_data <- spr_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
-                                            ) %>% mutate(model="5-gram")
+# trying with semi-summarized data, b/c otherwise it takes forever
 
-grnn_data <- spr_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
-                                            )%>% mutate(model="GRNN")
-
-txl_data <- spr_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
-                                            ) %>% mutate(model="TXL")
+# ngram_data <- spr_data%>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram, 
+#                                             freq=freq_center, len=length_center,
+#                                             prev_freq=past_c_freq, prev_len=past_c_length
+#                                             ) %>% mutate(model="5-gram")
+# 
+# grnn_data <- spr_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn, 
+#                                             freq=freq_center, len=length_center,
+#                                             prev_freq=past_c_freq, prev_len=past_c_length
+#                                             )%>% mutate(model="GRNN")
+# 
+# txl_data <- spr_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl, 
+#                                             freq=freq_center, len=length_center,
+#                                             prev_freq=past_c_freq, prev_len=past_c_length
+#                                             ) %>% mutate(model="TXL")
 
@@
 ```{r gams}
-gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")
-
-gam_grnn_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_data, method="REML")
-
-gam_txl_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_data, method="REML")
+# gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")
+# 
+# gam_grnn_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_data, method="REML")
+# 
+# gam_txl_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_data, method="REML")
 
@@
 
-ggsave("../Papers/gam.pdf", width=4, height=2.3, unit="in")
+ggsave("../Papers/gam-spr.pdf", width=4, height=2.3, unit="in")
 

### 2020-09-04 Papers/Amlap_2020_talk/amlap_2020_talk.tex
modify graph
talk edits

@@
 \pause
-\vskip -2em
+%\vskip -2em
 {\large Yes, at least well enough.} \pause
@@
 	\item Sometimes generates plausible distractors. \pause
-	\item A-maze results comparable with G-maze {\small(Boyce et al 2020, Sloggett et al 2020)} \pause 
+	\item A-maze results comparable with G-maze {\small(Boyce et al 2020, Sloggett et al 2020)}
 \end{itemize} 
+\end{frame}
+
+\begin{frame}{Does it work?} 
+	%\vskip -2em
+		\small
+	From ``Maze Made Easy'' (Boyce et al 2020)
 	\medskip
+
 	\centering
 
-\includegraphics[width=.7\textwidth]{../a_all.pdf}
+	\includegraphics[width=\textwidth]{../spr_g_amaze.pdf}
+
+		{\footnotesize Error bars: 95\% CI}
+
 \end{frame}
@@
 \end{frame}
-%\begin{frame}{Surprisal Effects}
-%	Takeaways: \pause
-%	\begin{itemize}
-%		\item Minimal frequency effects (consistent with Shain 2019) \pause
-%		\item Long RTs \pause
-%		\item Large effects of Length, Surprisal \pause
-%		\item Very little spillover \pause
-%	\end{itemize}
+\begin{frame}{Surprisal Effects}
+	Takeaways: \pause
+	\begin{itemize}
+		\item Minimal frequency effects (consistent with Shain 2019) \pause
+		\item Large effects of Length, Surprisal \pause
+		\item Very little spillover \pause
+	\end{itemize}
 %	
-%	Model comparison: GRNN is best, but TXL complementary
-%\end{frame}
+	Model comparison: GRNN is best, but TXL complementary
+\end{frame}
 
@@
 	\pause
-	Bayesian reader framework: Look at words long enough to ID with some threshold of certainty \pause
+	Bayesian Reader (Norris 2006): Look at words long enough to ID with some threshold of certainty \pause
 

### 2020-09-16 Papers/Paper/manuscript.Rmd
init blobs of words

TODO: read Natural Stories paper, crib some of intro from our old paper, wordcount old paper -- how long is 4000 words??

? how much do I re-explain Maze here? briefly explain how it works?, but maybe not in the introduction??

Outline:
- Introduction
 - hook/frame
 - sufficient set-up to explain why this is useful
- Methodology
- Experiment methods/set-up
- Results: what we have already, (additional analyses??)
- Discussion (why we are awesome, other stuff)
- ?? Supplement ??

# Abstract
TODO

# Introduction

## Hook 
TODO: make not suck
We all want to have more methods for doing research. A diversity of robust methods provides for converging evidence from different methods (either within or between types), and choices allow for a good fit to the scope and goals of research.

## Introduce SPR, eye-tracking
The two major incremental processing methods are eye-tracking and self-paced reading. The field of incremental processing research has long been dominated by eye-tracking and self-paced reading, both of which have advantages, but they present a serious tradeoff between ease of running an experiment and quality of the data. There's also analytic flexibility issues. There are other methods, and one less common method is attracting some more attention.

## But sometimes, you want different trade-offs, so Maze.

The Maze method (CITATIONS) provides a different set of tradeoffs than other methods. It promises good localization, but at what seems like an exhorbitant cost for widespread use. In the Maze task, participants see two words at a time and have to choose which word continues the sentence. The time to choose the correct word is taken as the dependent measure and interpreted as representing the difficulty of integrating the word into the sentence so far. If participants make a mistake, the sentence discontinues, so participants are incentivized to be careful (not sure this is a real incentive here). This is claimed to force extremely incremental processing, where participants must fully integrate each word into the sentence in order to be confident selecting it. In contrast, both eye-tracking and SPR provide some more flexibility in when participants integrate words, which can lead to spillover and thus analytic flexibility. In eye-tracking, participants read naturally, and so have the option to look back at a previous span when clarification is needed (thus, processing difficulty may be reflected in various measures). In SPR, the time it takes participants to initiate a button click is relatively long compared to how long it takes to integrate a word, so participants may click ahead, identifying each word before they move on, but not fully integrating it until they are looking at a later word. This results in spillover effects and messiness.

There's times when you don't want to deal with spillover effects, and want things to be very incremental.

## but Maze is really not useable, except there's this new version
However, Maze had some drawbacks, namely the time- and effort- intensive process of creating distractors probably dissuaded people from using it widely. An alternative version of the Maze task, dubbed A-maze, was recently introduced that gets around this problem by automatically generating distrators using a program and neural nets (Boyce et al, 2020). Boyce et al provide an interface for running the Maze task in a browser and a way of generating distractors that are usually, although not always, sufficiently bad. In a comparison between these A-maze materials and traditional hand-crafted distractors (from Witzel 2012), they find that A-maze yields comparable results to G-maze (either lab or web), and in one condition recovers effects that G-maze did not. A-maze also outperforms self-paced reading on this syntactic disambiguation paradigm. Sloggett (cite) also directly compared A-maze and G-maze on a disambiguaton paradigm and similarly found that A-maze and G-maze yielded similar results. 
One evident shortcoming of the Maze task is that it has been used on only a limited set of materials. Some of this is due to the fact that it isn't used as widely as other incremental processing methods, and some is due to inherent limitations of the paradigm. In particular, because Maze tasks discontinue after participants make mistakes, the farther into an item a word is, the fewer participants see it. This makes it hard to run long materials using Maze, and prevents Maze from being used on long text passages where SPR or eye-tracking could be used.

## What we do is offer an expansion on new Maze

We find a way around this problem by creating a version of the Maze task interface where participants correct their mistakes and continue with the sentence rather than being discontinued. We use this method with A-maze generated distractors on the Natural Stories corpus.

With new methods, its useful to try to establish how widely they work. We hope to test A-maze on an experiment that is very different from the syntactic disambiguation, in order to learn whether it works for a different type of psycholinguistic material. 
While Maze has traditionally been limited to single sentence items; we innovate on how the materials are presented and show that it is possible for participants to read naturalistic passages that are presented in the Maze paradigm. We confirm that participants can understand what they are reading, and that Maze shows linear surprisal effects. This opens up another area of research to being amenable to the new A-maze paradigm.

This is just one other data point on A-maze, so we still don't know how widely the method holds up; however the fact that it works on such disparate materials gives us hope that it will be useful for a variety of paradigms.

# Error-correction maze

The Maze tasked as presented in Forster et al 2009 consists of showing participants individual sentences for materials that are controlled (in a Latin Square type design). Other Maze experiments have also worked like this. One issue with Maze is data loss -- whenever a participant makes a mistake, they don't see the rest of the sentence and thus they do not contribute RT data for any words in the rest of the sentence. While this hasn't been a problem with in lab G-maze (error rates were fairly low and items not too long), it's more of a problem with A-maze on crowd-sourced participants, due to both compontents. One of the issues with A-maze is that some of the distractors, especially early in the sentence are problematic, and therefore data-loss is high.

However, even with G-maze quality materials, and high quality subjects, the Maze paradigm with eventually run into trouble as materials get longer. Even assuming a very low per word error rate (TODO: check what would be reasonable, but could also use 1%), the errors will build up as materials get longer, and few participants will make it to the end. This makes it hard to run long materials, such as could be done with eye-tracking or SPR. One potential option is to have participants skip to the next sentence (rather than next item) when they make a mistake, but having missed some context from the last sentence, this may be jarring and they may be confused (and have unreliable RT) due to lack of context and discourse cues.

To solve this conundrum, we decided to try letting participant continue after their mistakes. To do this, we presented them with an error message, and waited until they selected the correct option, then the sentence continued as normal. This was easily implemented as a variant on the Maze Ibex thing. We record both the RT to the first click and also the total RT until they select the correct answer separately.

This in principle solves several problems. Firstly, it allows for running longer materials, potentially expanding the range of questions Maze is suitable for. Secondly, it compensates for some of the shortcomings of Maze identified in Boyce et al 2020 -- distractors might still cause participants to make unavoidable mistakes, but at least they still see the rest of the sentence (perhaps reducing frustration), and we get their data. Whether RT data from after mistakes is useable is a question of potential interest.

Lastly, this helps to address the "distractors or participants" question around errors mentioned in Boyce. When participants continue with the sentence, we can see what fraction of time they made mistakes and do participant level selection on the basis of this, without having to account for the censoring due to lack of data. This provides a convenient way of controlling data quality after the fact.

It also starts to reduce the perverse incentives. With error-terminates Maze, the fastest way to get through the task is to select randomly, and it's quite quick because mistakes skip you to the next sentence. With error-correction Maze, randomly jamming buttons takes more effort, but is still an effective strategy. [Footnote: it has been suggested that one way to disincentivize this is to force a pause when a participant makes a mistake, forcing them to wait some short period of time (500ms or 1 sec) before being able to correct their mistake. This seems like a promising idea that could be implemented.]

# Natural Stories corpus

We wanted to try Maze on a naturalistic materials, to see if error-correction Maze would work, whether surprisal and other coarse-level processing effects would appear (as a sanity check) -- eye-tracking and SPR show these effects robustly and in relatively well established patterns, and for Maze to be trusted as an alternative, it would need to as well. Maze typically doesn't use comprehension questions because comprehending the local sentence structure is required to do the task. Thus, there isn't much information on how well participants comprehend and remember what they read. We wanted to know if participants understood and internalized the gist of what they were reading with Maze.

To satisfy these various desires, we decided to run a Maze task on the Natural Stories corpus (Futrell et al 2020). This corpus consists of 10 passages, each about 1000 words long. They are naturalistic, although they were designed to contain a fair number of low frequency constructions such as BLAH and BLAH. (We do not specifically analyse this.) They also come with comprehension questions, and SPR data from the corpus is available. THis seemed like an appropriate severe test, and the corpus was available. SEE FOOBAR for examples of the materials and questions. 
# Actual methods junk

# Materials
We took the corpus and split it up into sentences (this involved some judgement calls due to quoted speech) because the Maze material generation works at the sentence level. We ran the sentences through A-maze generation using TODO foo, bar parameters. Materials are available in TODO: anon repo. We set up the materials in Ibex, using Maze to display things.

We also wrote a short passage to use as practice, with 2 comprehension questions (double check!!).

Participants gave their informed consent, saw instructions, saw the practice item and practice comprehension questions, and then completed the story using the Maze task, then the 6 main comprehension questions. At the end of the experiment, we asked how their experience was, gathered demographic questions (which ones), and debriefed participants.

# Participants
We recruited XX participants from Mechanical Turk, and paid each participant YY dollars.

# Analysis stuff

We excluded participants who did not indicate that their native language was English. This left us with YY participants.

## Qualitative stuff

Our first points of interest were how well participants would do on the Maze task and on the comprehension questions. Figure blah shows the distributions and relationship between participants accuracy on the Maze task (i.e. fraction of the time they selected the correct continuation) and the time to took for them to make selections (mean RT; RT's above ZZ were assumed to be indicative of a break and excluded). Participant dots are color coded based on whether they were at least 80% accurate; this is the threshold we use for including data in future analysis. A number of participants did quite well at the task (accuracies reached (double check!) 99%). This shows that it's possible for participants to focus on the task for the whole time. (While this is not overall much longer than previous studies such as Witzel 2012 and Boyce et al 2020, it had to be continuous, and so we couldn't use a blocking structure to create pauses.) It also shows that A-maze distractors are high-enough quality that participants are able to get them right most of the time. Given the clump of participants at around 50% accuracy, which is consistent with randomly clicking, we assume they are not taking the task seriously. In terms of RT, we see that attentive participants take on average %FOOBAR seconds, while participants who are clicking randomly can do so much faster.

In terms of comprehension questions, we see relatively good performance by participants who were attentive to the task. These answers were also provided quickly, which mostly rules out the (define quickly) possibility that they were googling for the answers. There's a chance that some knew the answers already, so this isn't super conclusive. Because the stories take a long time to read (~15 minutes+), participants are having to remember the story/answers to questions for quite a while. This performance is less goood than that for SPR; but it's hard to know comparing between methods with different task demands and subject pools. Participants who didn't do well also didn't do well here.

## Quantitative stuff

There's commonly observed effects of surprisal, frequency, and length on stuff, so we looked at that. TODO: how we got estimates of surprisal, frequency, and length. TODO: word exclusions

- GAMs
- LMER/BRMS

*write up what we have, then we can consider adding comparison and ex-gaussians and stuff when we do it*

TODO

# Discussion

We've shown another way to adapt the Maze task to make it suitable for a wider range of experiments. While we think that error-correction Maze is generally a good idea, especially when using A-maze, this is an orthagonal adjustment to the task. By using the Natural Stories corpus, we establish that Maze works for longer naturalistic tasks, with participants able to do the task and understand what they are reading. Additionally, their RTs reflect expected patterns in terms of length and surprisal. While there are some differences, especially in scale, to that found with other incremental processing paradigms, we think this is a reason to explore more. Could be a statistical thing (???) or due to task differences. Comparisons between methods could be very useful for identifying how task demands influence processing.

All the code is available, and we encourage researchers to consider trying out Maze if they think it's appropriate to their experiments.

This is just one other data point on A-maze, so we still don't know how widely the method holds up; however the fact that it works on such disparate materials gives us hope that it will be useful for a variety of paradigms.

### 2020-09-17 Papers/Paper/manuscript.Rmd
progress on beginning

@@
-TODO: read Natural Stories paper, crib some of intro from our old paper, wordcount old paper -- how long is 4000 words??
+---
+title             : "The title"
+shorttitle        : "Title"
 
-? how much do I re-explain Maze here? briefly explain how it works?, but maybe not in the introduction??
+author: 
+  - name          : "Veronica Boyce"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "Postal address"
+    email         : "vboyce@stanford.edu"
+    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+      - Conceptualization
+      - Writing - Original Draft Preparation
+      - Writing - Review & Editing
+  - name          : "Roger Levy"
+    affiliation   : "2"
+    role:
+      - Writing - Review & Editing
 
-Outline:
-- Introduction
- - hook/frame
- - sufficient set-up to explain why this is useful
-- Methodology
-- Experiment methods/set-up
-- Results: what we have already, (additional analyses??)
-- Discussion (why we are awesome, other stuff)
-- ?? Supplement ??
+affiliation:
+  - id            : "1"
+    institution   : "Stanford University"
+  - id            : "2"
+    institution   : "Massachusetts Institute of Technology"
 
-# Abstract
-TODO
+authornote:
 
-# Introduction
+abstract: |
+  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
 
-## Hook 
-TODO: make not suck
-We all want to have more methods for doing research.
+  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
+
+  One sentence clearly stating the **general problem** being addressed by  this particular study.
+
+  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
+
+  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
+
+  One or two sentences to put the results into a more **general context**.
+
+  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
+
+  <!-- https://tinyurl.com/ybremelq -->
+
+keywords          : "keywords"
+wordcount         : "X"
+
+bibliography      : ["r-references.bib"]
+
+floatsintext      : yes
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+linenumbers       : yes
+mask              : no
+draft             : no
+
+documentclass     : "apa6"
+classoption       : "man"
+output            : papaja::apa6_pdf
+---
+
+```{r setup, include = FALSE}
+library("papaja")
+r_refs("r-references.bib")
+```
+
+```{r analysis-preferences}
+# Seed for random number generation
+set.seed(42)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+```
+
 A diversity of robust methods provides for converging evidence from different methods (either within or between types), and choices allow for a good fit to the scope and goals of research.
+But we also want these methods to be reliable.
+Some of that is theoretical, but we also want empirical evidence that what the method captures is relevant to what we want to study, and that this connection is true across as broad a range of materials as we hope to use it for.
+Robustness is important to establishing how good a new method is.
+What range of tasks can this new tool accomplish and how well?
+(Could really extend on a toolbox analogy.)
 
-## Introduce SPR, eye-tracking
-The two major incremental processing methods are eye-tracking and self-paced reading.
-The field of incremental processing research has long been dominated by eye-tracking and self-paced reading, both of which have advantages, but they present a serious tradeoff between ease of running an experiment and quality of the data.
-There's also analytic flexibility issues.
-There are other methods, and one less common method is attracting some more attention.
+Incremental processing -- it's really important and we want different ways of measuring it.
+The field of incremental processing research is dominated by eye-tracking and self-paced reading.
+In eye-tracking ...
+Eye-tracking is natural and yields fine grained data, but it's expensive in terms of researcher time and equipment.
+Self-paced reading is quicker and cheaper to run because it can be done online without specialized equipment or supervision; however, the results are plagued by spillover effects.
+Both methods introduce analytic flexibility in which eye-tracking measures or what region of spillover you consider, and how data is residualized.
+In SPR, the time it takes participants to initiate a button click is relatively long compared to how long it takes to integrate a word, so participants may click ahead, identifying each word before they move on, but not fully integrating it until they are looking at a later word.
+There seems to be a gap for a method that is cheap to run, but has straightforwardly localized (and thus easy to analyse data).
+This may seem like too much to hope for, but the Maze task appears to fill this gap.
 
-## But sometimes, you want different trade-offs, so Maze.
-
-The Maze method (CITATIONS) provides a different set of tradeoffs than other methods.
-It promises good localization, but at what seems like an exhorbitant cost for widespread use.
 In the Maze task, participants see two words at a time and have to choose which word continues the sentence.
@@
 This is claimed to force extremely incremental processing, where participants must fully integrate each word into the sentence in order to be confident selecting it.
-In contrast, both eye-tracking and SPR provide some more flexibility in when participants integrate words, which can lead to spillover and thus analytic flexibility.
-In eye-tracking, participants read naturally, and so have the option to look back at a previous span when clarification is needed (thus, processing difficulty may be reflected in various measures).
-In SPR, the time it takes participants to initiate a button click is relatively long compared to how long it takes to integrate a word, so participants may click ahead, identifying each word before they move on, but not fully integrating it until they are looking at a later word.
-This results in spillover effects and messiness.
+The Maze task is even more unnatural than SPR, but in some experiments, this might not be a real downside.
+The actually prohibitive fine print on the Maze task is the need for distractors; while cheap and easy to run, traditional Maze materials are effort-intensive to construct, which may have kept this method from widespread adoption.
 
-There's times when you don't want to deal with spillover effects, and want things to be very incremental.
-
-## but Maze is really not useable, except there's this new version
-However, Maze had some drawbacks, namely the time- and effort- intensive process of creating distractors probably dissuaded people from using it widely.
+However, a solution to the issue of generating Maze materials has been proposed.
 An alternative version of the Maze task, dubbed A-maze, was recently introduced that gets around this problem by automatically generating distrators using a program and neural nets (Boyce et al, 2020).
+While the quality of the automatically generated distractors is not quite up that of hand-generated (as judged by their being occassional distractors that seem completely acceptable continuations), the results from A-maze match those of G-maze.
 Boyce et al provide an interface for running the Maze task in a browser and a way of generating distractors that are usually, although not always, sufficiently bad.
@@
 Sloggett (cite) also directly compared A-maze and G-maze on a disambiguaton paradigm and similarly found that A-maze and G-maze yielded similar results.
+
+These demonstrations make A-maze seem quite promising, as it has the advantages of the traditional Maze task, without it's drawback. This method seems worth exploring further.
+
 One evident shortcoming of the Maze task is that it has been used on only a limited set of materials.
@@
 
-## What we do is offer an expansion on new Maze
-
 We find a way around this problem by creating a version of the Maze task interface where participants correct their mistakes and continue with the sentence rather than being discontinued.
-We use this method with A-maze generated distractors on the Natural Stories corpus.
-
-With new methods, its useful to try to establish how widely they work.
-We hope to test A-maze on an experiment that is very different from the syntactic disambiguation, in order to learn whether it works for a different type of psycholinguistic material.
-While Maze has traditionally been limited to single sentence items; we innovate on how the materials are presented and show that it is possible for participants to read naturalistic passages that are presented in the Maze paradigm.
-We confirm that participants can understand what they are reading, and that Maze shows linear surprisal effects.
-This opens up another area of research to being amenable to the new A-maze paradigm.
-
-This is just one other data point on A-maze, so we still don't know how widely the method holds up; however the fact that it works on such disparate materials gives us hope that it will be useful for a variety of paradigms.
+This allows for multi-sentence items to be run using Maze, which we verify by running Maze on the Natural Stories corpus.
+With this data, we are able to confirm that participants can comprehend what they read using Maze and investigate the effects of surprisal on RT in Maze.
 
@@
 
-The Maze tasked as presented in Forster et al 2009 consists of showing participants individual sentences for materials that are controlled (in a Latin Square type design).
-Other Maze experiments have also worked like this.
-One issue with Maze is data loss -- whenever a participant makes a mistake, they don't see the rest of the sentence and thus they do not contribute RT data for any words in the rest of the sentence.
+The Maze tasked as historically done involves presenting in-lab participants (often college students) with materials that consist of individual, unrelated sentences, some of which exemplify the relevant properties (often with a Latin Square design).
+While this is a common psycholinguistic paradigm, it isn't the only one, and Maze would be more versatile (and thus more appealing as an alternative to eye-tracking and SPR) if it also worked on other types of materials, such as longer texts used in research in discourse or surprisal.
+There's also interest in comparing neural net performance to that of human behavioral data on naturalistic corpora, which necessitates behavioral corpora.
+
+One major concern with Maze is data loss -- whenever a participant makes a mistake, they don't see the rest of the sentence and thus they do not contribute RT data for any words in the rest of the sentence.
+TODO cite how much this happened in old paper.
 While this hasn't been a problem with in lab G-maze (error rates were fairly low and items not too long), it's more of a problem with A-maze on crowd-sourced participants, due to both compontents.
-One of the issues with A-maze is that some of the distractors, especially early in the sentence are problematic, and therefore data-loss is high.
+In particular, Boyce et al 2020 found that some of the distractors, especially early in the sentence are problematic, and therefore data-loss is high.
 
 However, even with G-maze quality materials, and high quality subjects, the Maze paradigm with eventually run into trouble as materials get longer.
-Even assuming a very low per word error rate (TODO: check what would be reasonable, but could also use 1%), the errors will build up as materials get longer, and few participants will make it to the end.
-This makes it hard to run long materials, such as could be done with eye-tracking or SPR.
-One potential option is to have participants skip to the next sentence (rather than next item) when they make a mistake, but having missed some context from the last sentence, this may be jarring and they may be confused (and have unreliable RT) due to lack of context and discourse cues.
+Even assuming a very low per word error rate , the errors will build up as materials get longer, and few participants will make it to the end.
+For instance, with a 1% error rate, `r round(.99**15*100)`% would complete each 15-word sentence, but only `r round(.99**50*100)`% of participants would complete each 50 word vignette, and `r round(.99**200*100)`% of 200 word passages.
+One potential option would be to have participants skip to the next sentence (rather than next item) when they make a mistake.
+However, having missed some context from the last sentence, this may be jarring and and confusing for participants.
+Their RTs may also be unreliable  due to having missed context and discourse cues.
 
-To solve this conundrum, we decided to try letting participant continue after their mistakes.
-To do this, we presented them with an error message, and waited until they selected the correct option, then the sentence continued as normal.
-This was easily implemented as a variant on the Maze Ibex thing.
-We record both the RT to the first click and also the total RT until they select the correct answer separately.
+To solve this conundrum, we decide to get rid of the source of the problem: the terminating on errors.
+Instead, we let participants continue after they make a mistake.
+To do this, we present them with an error message, and wait until they select the correct option, before continuing the sentence as normal.
+This was easily implemented as a new option built on the Ibex Maze code used in Boyce et al 2020.
+We record both the RT to the first click and also the total RT until they select the correct answer as separate values.
 
@@
 Firstly, it allows for running longer materials, potentially expanding the range of questions Maze is suitable for.
-Secondly, it compensates for some of the shortcomings of Maze identified in Boyce et al 2020 -- distractors might still cause participants to make unavoidable mistakes, but at least they still see the rest of the sentence (perhaps reducing frustration), and we get their data.
-Whether RT data from after mistakes is useable is a question of potential interest.
+Secondly, it compensates for some of the shortcomings of Maze identified in Boyce et al 2020 -- distractors might still cause participants to make unavoidable mistakes, but at least they still see the rest of the sentence and we get their data.
+We believe (although don't have evidence) that being able to correct mistakes may reduce frustration related to unavoidable mistakes from impossible choices and accidental misclicks.
+Whether this also solves the data loss problem from the researcher perspective within a sentence depends on whether post-mistake data are high-quality and trustworthy; this is a hard-to-assess question of potential interest.
 
-Lastly, this helps to address the "distractors or participants" question around errors mentioned in Boyce.
-When participants continue with the sentence, we can see what fraction of time they made mistakes and do participant level selection on the basis of this, without having to account for the censoring due to lack of data.
-This provides a convenient way of controlling data quality after the fact.
+Even if all post-mistake data are discarded before analysis, they can still help address the "distractors or participants" question around errors mentioned in Boyce.
+When participants continue with the sentence, it is easy to calculate the per-word error rate for participants, without having to account for the censoring present in traditional Maze (if you already made a mistake, we can't know if you would have made another one; with error-correction, we do know if you did).
+This allows us to exclude participants on the basis of their task accuracy, ensuring RT data comes from participants who were paying attention and doing the task.
+This provides a convenient and clear-cut way of controlling data quality after the fact.
 
@@
 With error-correction Maze, randomly jamming buttons takes more effort, but is still an effective strategy.
-[Footnote: it has been suggested that one way to disincentivize this is to force a pause when a participant makes a mistake, forcing them to wait some short period of time (500ms or 1 sec) before being able to correct their mistake.
-This seems like a promising idea that could be implemented.]
+[Footnote: In discussing this work, we recieved the suggestion that one way to disincentivize random clicking is to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms or 1 sec) before being able to correct their mistake.
+This seems like a promising improvement that could be worth implementing and testing.]
 
@@
 
-We wanted to try Maze on a naturalistic materials, to see if error-correction Maze would work, whether surprisal and other coarse-level processing effects would appear (as a sanity check) -- eye-tracking and SPR show these effects robustly and in relatively well established patterns, and for Maze to be trusted as an alternative, it would need to as well.
+As our test of the Maze method, we use the Natural Stories corpus, which consists of 10 passages, each about 1000 words long (around 50 sentences), a mix of fiction and non-fiction.
+We think this is an appropriately severe test: the passages are long, and they contain a number of low-frequency syntactic constructions.
+In addition, they have a lot of punctuation (such as from quoted speech) and proper nouns (which are likely to be out of vocabulary for the autogeneration model).
+If people can make it through these passages while doing the Maze task accurately, we think they are likely to be able to on other naturalistic passages as well (many of which may be shorter or simpler).
+
+The Natural Stories corpus also comes with comprehension questions, 6 binary-choice questions per story.
 Maze typically doesn't use comprehension questions because comprehending the local sentence structure is required to do the task.
@@
 
-To satisfy these various desires, we decided to run a Maze task on the Natural Stories corpus (Futrell et al 2020).
-This corpus consists of 10 passages, each about 1000 words long.
-They are naturalistic, although they were designed to contain a fair number of low frequency constructions such as BLAH and BLAH.
-(We do not specifically analyse this.) They also come with comprehension questions, and SPR data from the corpus is available.
-THis seemed like an appropriate severe test, and the corpus was available.
-SEE FOOBAR for examples of the materials and questions.
-# Actual methods junk
+Our last check point of interest was whether and how coarse phenomena like the effect of word length and surprisal, which are well established for eye-tracking and SPR would manifest with Maze.
+If they show similar patterns to that found in other methods, that's more promising.
 
-# Materials
+Another consideration is the ability to compare our data with other analyses and data obtained from different methods.
+This corpus has SPR data available with we can compare to.
+Using a standard data set also means that this set of results will be easy to compare to in the future.
+
+# Methods
+We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
+<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
+
+## Participants
+We recruited XX participants from Mechanical Turk, and paid each participant YY dollars.
+
+## Material
+
 We took the corpus and split it up into sentences (this involved some judgement calls due to quoted speech) because the Maze material generation works at the sentence level.
@@
 
+## Procedure
+
 Participants gave their informed consent, saw instructions, saw the practice item and practice comprehension questions, and then completed the story using the Maze task, then the 6 main comprehension questions.
@@
 
-# Participants
-We recruited XX participants from Mechanical Turk, and paid each participant YY dollars.
+## Data analysis
+We used `r cite_r("r-references.bib")` for all our analyses.
+We excluded participants who did not indicate that their native language was English. This left us with YY participants.
 
-# Analysis stuff
-
-We excluded participants who did not indicate that their native language was English. This left us with YY participants.
+# Results
 
@@
 
-*write up what we have, then we can consider adding comparison and ex-gaussians and stuff when we do it*
-
-TODO
-
 # Discussion
-
 We've shown another way to adapt the Maze task to make it suitable for a wider range of experiments.
@@
 This is just one other data point on A-maze, so we still don't know how widely the method holds up; however the fact that it works on such disparate materials gives us hope that it will be useful for a variety of paradigms.
+
+With new methods, its useful to try to establish how widely they work.
+We hope to test A-maze on an experiment that is very different from the syntactic disambiguation, in order to learn whether it works for a different type of psycholinguistic material.
+While Maze has traditionally been limited to single sentence items; we innovate on how the materials are presented and show that it is possible for participants to read naturalistic passages that are presented in the Maze paradigm.
+We confirm that participants can understand what they are reading, and that Maze shows linear surprisal effects.
+This opens up another area of research to being amenable to the new A-maze paradigm.
+
+This is just one other data point on A-maze, so we still don't know how widely the method holds up; however the fact that it works on such disparate materials gives us hope that it will be useful for a variety of paradigms.
+
+The Maze task might be espeically good for comparing with NNs because of the forced incrementality.
+While eventually we do want to figure out models of more natural reading, this might be a useful complement b/c we don't have to deal with nasty spillover.
+
+\newpage
+
+# References
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+<div id="refs" custom-style="Bibliography"></div>
+\endgroup

### 2020-09-22 Papers/Paper/manuscript.Rmd
clean up intro a bit, draft abstract

@@
 ---
-title             : "The title"
-shorttitle        : "Title"
+title             : "A-maze of Natural Stories: Texts are comprehensible during the Maze task"
+shorttitle        : "A-maze of Natural Stories"
 
@@
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
+    address       : "TODO"
     email         : "vboyce@stanford.edu"
@@
       - Conceptualization
+      - Formal Analysis
+      - Investigation
+      - Software
       - Writing - Original Draft Preparation
@@
     role:
+      - Conceptualization
+      - Formal Analysis
       - Writing - Review & Editing
@@
 
-authornote:
+authornote: |
+  TODO
 
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
+  Easy to use, reliable methods are important to science.
+In studying how people understand language in real time, we use incremental processing methods to measure word-by-word reading time.
+However, neither of the common methods (eye-tracking and self-paced reading) can run over the web and produce localized effects.
+Another method, called the Maze task, seems to be able to do this, especially since a technique called A-maze makes the task much faster to construct.
+Due to task limitations, Maze has not been used on long, naturalistic passages, only short targeted materials.
+Here we present an adaptation of the Maze method suitable for long materials; we test this method on the Natural Stories corpus and find that participants can comprehend what they read while doing the task and that the reading time patterns are broadly similar to those found with other methods.
+We find support for the localization of reading time effects during the Maze task, as well as extending the range of materials Maze is suitable for.
 
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "keywords"
+keywords          : "TODO"
 wordcount         : "X"
 
-bibliography      : ["r-references.bib"]
+bibliography      : ["r-references.bib","refs.bib"]
 
@@
 
-Incremental processing -- it's really important and we want different ways of measuring it.
-The field of incremental processing research is dominated by eye-tracking and self-paced reading.
-In eye-tracking ...
-Eye-tracking is natural and yields fine grained data, but it's expensive in terms of researcher time and equipment.
-Self-paced reading is quicker and cheaper to run because it can be done online without specialized equipment or supervision; however, the results are plagued by spillover effects.
-Both methods introduce analytic flexibility in which eye-tracking measures or what region of spillover you consider, and how data is residualized.
-In SPR, the time it takes participants to initiate a button click is relatively long compared to how long it takes to integrate a word, so participants may click ahead, identifying each word before they move on, but not fully integrating it until they are looking at a later word.
-There seems to be a gap for a method that is cheap to run, but has straightforwardly localized (and thus easy to analyse data).
+When we hear or read language, we incorporate new words into our model of the unfolding context as we go.
+To understand this real-time process and how it copes with unexpected words, psycholinguists use incremental processing methods.
+This set of methods involves measuring how long participants choose to spend reading and integrating a word before moving on, although the tasks and measures of reading/reaction time differ.
+
+The two most commonly used methods are eye-tracking and self-paced reading.
+In eye-tracking, participants read text freely from a screen, while the eye movements and fixations are tracked with an infrared camera CITE, check for truth.
+Eye-tracking is naturalistic and yields fine-grained data about how long participants spent on each word when, but it's expensive to run requiring both specialized equipment and researcher time to run experiments.
+
+Self-paced reading (commonly abbreviated SPR) in contrast can be run online with no specialzed equipment or direct researcher supervision, and so it's possible to collect data quickly and cheaply.
+In SPR, participants see one word of the text on screen at a time and press a button to contine to the next word instead; the time between button presses when a word was on screen is used as the reading time.
+The problem with SPR, especially when run online, is that effects from one word tend to show up as longer reading times spread out over several later words, in what is known as spillover effects.
+Both of these methods suffer somewhat from analytic flexibilty as processing difficulty may manifest in several ways in eye-tracking and is hard to localize in SPR.
+
+For some experimental paradigms, it would be idea to have a readily useable method that localizes processing effects in a straightforward to analyse manner.
 This may seem like too much to hope for, but the Maze task appears to fill this gap.
 
-In the Maze task, participants see two words at a time and have to choose which word continues the sentence.
+In the Maze task, participants see two words at a time, one of which continues the sentence, and one of which does not, and they have to choose which word is the correct continuation.
 The time to choose the correct word is taken as the dependent measure and interpreted as representing the difficulty of integrating the word into the sentence so far.
-If participants make a mistake, the sentence discontinues, so participants are incentivized to be careful (not sure this is a real incentive here).
-This is claimed to force extremely incremental processing, where participants must fully integrate each word into the sentence in order to be confident selecting it.
-The Maze task is even more unnatural than SPR, but in some experiments, this might not be a real downside.
-The actually prohibitive fine print on the Maze task is the need for distractors; while cheap and easy to run, traditional Maze materials are effort-intensive to construct, which may have kept this method from widespread adoption.
+If participants make a mistake, the sentence discontinues.
+Because of the required choices, the Maze task is claimed to  force extremely incremental processing, where participants must fully integrate each word into the sentence in order to confidently select it and move on.
+Experiments using  The actually prohibitive fine print on the Maze task is the need for distractors; while cheap and easy to run, traditional Maze materials are effort-intensive to construct, which may have kept this method from widespread adoption.
 
@@
 Boyce et al provide an interface for running the Maze task in a browser and a way of generating distractors that are usually, although not always, sufficiently bad.
+
 In a comparison between these A-maze materials and traditional hand-crafted distractors (from Witzel 2012), they find that A-maze yields comparable results to G-maze (either lab or web), and in one condition recovers effects that G-maze did not.
@@
 
-As our test of the Maze method, we use the Natural Stories corpus, which consists of 10 passages, each about 1000 words long (around 50 sentences), a mix of fiction and non-fiction.
-We think this is an appropriately severe test: the passages are long, and they contain a number of low-frequency syntactic constructions.
-In addition, they have a lot of punctuation (such as from quoted speech) and proper nouns (which are likely to be out of vocabulary for the autogeneration model).
-If people can make it through these passages while doing the Maze task accurately, we think they are likely to be able to on other naturalistic passages as well (many of which may be shorter or simpler).
+We test this method using the Natural Stories corpus, which consists of 10 passages, some fiction and some non-fiction.
+We think this is a fairly severe test: the passages are long (about 1000 words each), they contain a lot of punctuation (for instance quoted speech), a number of proper nouns (which are likely out of vocabulary for the autogeneration model), and a number of low frequency grammatical constructions.
+However, they are intended to read naturally to native English speakers.
+We choose this because if participants can succeed at doing the Maze task and understanding it on this set of materials, we think they are likely to on basically any naturalistic text (many of which might be shorter or easier to follow).
+The corpus also comes with binary-choice comprehension questions, 6 per story.
 
-The Natural Stories corpus also comes with comprehension questions, 6 binary-choice questions per story.
+THis corpus has been used for SPR, so there's the possiblity of comparison.
+
+In addition to testing that Maze works at all on this type of material, we also wanted to check for two other desired characteristics.
+That the materials can be comprehended even when read with Maze, and that commonly found psycholinguistic effects of length and surprisal show.
+whether surprisal and other coarse-level processing effects would appear (as a sanity check) -- eye-tracking and SPR show these effects robustly and in relatively well established patterns, and for Maze to be trusted as an alternative, it would need to as well.
 Maze typically doesn't use comprehension questions because comprehending the local sentence structure is required to do the task.
@@
 
-Our last check point of interest was whether and how coarse phenomena like the effect of word length and surprisal, which are well established for eye-tracking and SPR would manifest with Maze.
-If they show similar patterns to that found in other methods, that's more promising.
-
-Another consideration is the ability to compare our data with other analyses and data obtained from different methods.
-This corpus has SPR data available with we can compare to.
-Using a standard data set also means that this set of results will be easy to compare to in the future.
+To satisfy these various desires, we decided to run a Maze task on the Natural Stories corpus (Futrell et al 2020).
+This corpus consists of 10 passages, each about 1000 words long.
+They are naturalistic, although they were designed to contain a fair number of low frequency constructions such as BLAH and BLAH.
+(We do not specifically analyse this.) They also come with comprehension questions, and SPR data from the corpus is available.
+THis seemed like an appropriate severe test, and the corpus was available.
+SEE FOOBAR for examples of the materials and questions.
 

### 2021-01-26 Papers/Paper/manuscript.Rmd
work on manuscript

@@
     institution   : "Massachusetts Institute of Technology"
+header-includes:
+ - \usepackage{setspace}\singlespacing
 
@@
 abstract: |
-  Easy to use, reliable methods are important to science.
+  TODO Easy to use, reliable methods are important to science.
 In studying how people understand language in real time, we use incremental processing methods to measure word-by-word reading time.
@@
 library("papaja")
+library("citr")
+library(patchwork)
+library(here)
+library(tidyverse)
+library(readr)
+library(brms)
+library(lme4)
+library(rstan)
+library(tidybayes)
+library(knitr)
+library(matrixStats)
 r_refs("r-references.bib")
-```
-
-```{r analysis-preferences}
-# Seed for random number generation
+theme_set(theme_bw())
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+options(knitr.table.format = "pdf")
 set.seed(42)
@@
 ```
-
+# Intro
+[TODO: hook about toolbox of methods]
 A diversity of robust methods provides for converging evidence from different methods (either within or between types), and choices allow for a good fit to the scope and goals of research.
@@
 
-When we hear or read language, we incorporate new words into our model of the unfolding context as we go.
-To understand this real-time process and how it copes with unexpected words, psycholinguists use incremental processing methods.
-This set of methods involves measuring how long participants choose to spend reading and integrating a word before moving on, although the tasks and measures of reading/reaction time differ.
+[incrementality; so need incremental methods] 
+One question in psycholinguistics is how real-time comprehension works; as we read or listen, we process language incrementally, building and changing our mental model of a sentence.
+How long we spend on each word is due to both qualities of the word itself and how it fits with the sentence context so far TODO citations
 
-The two most commonly used methods are eye-tracking and self-paced reading.
+We process language incrementally, building a model of the sentence as we hear or read word.
+When a word is unexpected and does not fit the model of the sentence so far, our understanding of the sentence may need to be revised.
+Incremental processing methods measure this process, tracking how long each word takes to read as a proxy for how hard it is to process and incoroporate into the mental conception of meaning.
+Fluent language uses process language rapidly and flexibly adapt to many unexpected words.
+This makes it harder to capture reading time effects because they may be obscured by the overall robust reading process.
+Different techniques handle this localization problem differently, and present different trade-offs.
+The two most commonly used methods are eye-tracking and self-paced reading, but neither of these techniques are ideal at localization.
+One measure of localization is whether current or past word reading time is more predictive.
+SPR and eye-tarcking both don't seem great.
+
+[eye-track and SPR, failure to localize] TODO: possibly shorten?
 In eye-tracking, participants read text freely from a screen, while the eye movements and fixations are tracked with an infrared camera CITE, check for truth.
 Eye-tracking is naturalistic and yields fine-grained data about how long participants spent on each word when, but it's expensive to run requiring both specialized equipment and researcher time to run experiments.
-
 Self-paced reading (commonly abbreviated SPR) in contrast can be run online with no specialzed equipment or direct researcher supervision, and so it's possible to collect data quickly and cheaply.
@@
 
-For some experimental paradigms, it would be idea to have a readily useable method that localizes processing effects in a straightforward to analyse manner.
-This may seem like too much to hope for, but the Maze task appears to fill this gap.
-
+[Maze]An alternative that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing.
 In the Maze task, participants see two words at a time, one of which continues the sentence, and one of which does not, and they have to choose which word is the correct continuation.
@@
 If participants make a mistake, the sentence discontinues.
-Because of the required choices, the Maze task is claimed to  force extremely incremental processing, where participants must fully integrate each word into the sentence in order to confidently select it and move on.
-Experiments using  The actually prohibitive fine print on the Maze task is the need for distractors; while cheap and easy to run, traditional Maze materials are effort-intensive to construct, which may have kept this method from widespread adoption.
+Theoretically, participants must fully integrate each word into the sentence in order to confidently select it and move on.
+This is suppported by studies finding localized effects without much spillover.
 
-However, a solution to the issue of generating Maze materials has been proposed.
+[A-Maze] However, traditional Maze materials are effort-instensive to construct, because of the need to select infelicitious words as distractors for each spot of each sentence; this may have kept this method from widespread adoption.
+
+CITE US propose a solution to this problem of generating Maze distractors.
+In A-maze, distractors are automatically generated by feeding the correct sentences through a program that uses NLP language models to select high-surprisal distractors.
+The quality of the automatically generated distractors is not up to that of hand-generated distractors, in that the distractors ocassionally are good fits in context.
+Nonetheless, Boyce et al found that materials with A-maze distractors had similar results to G-maze distractors.
+TODO: talk about the paradigm more
+[Summary of 2020 paper]
+
+[Summary other uses of A-maze - Slogett]
+Slogett also found that it works similarly.
+
+While new and untested, A-maze seems like a potentially powerful addition to the toolkit.
+However, like the other Maze tasks it has been limited in its application to single-sentence items probing minimal comparisons in constructed sentences.
+It's only been used on minimal pair types of sentences where each sentence is independent.
+In particular, it hasn't been used on longer, naturalistic texts, so we don't know how in compares to other methods here.
+And hasn't been used on multi-sentence issue of generating Maze materials has been proposed.
 An alternative version of the Maze task, dubbed A-maze, was recently introduced that gets around this problem by automatically generating distrators using a program and neural nets (Boyce et al, 2020).
@@
 
-One evident shortcoming of the Maze task is that it has been used on only a limited set of materials.
+[why do longer Maze] While some psycholinguistic phenomena are well-explored within individual targeted sentences, there's also a need for reading time data of more naturalistic passages.
+This is important as a comparison for models of language understanding, for correlating with NLP language models.
+In addition, some phenomena such as discourse effects, only work in the context of passages.
+THus, it would be useful/versatile if Maze could be used in these other contexts.
+
+[short-coming, where's the question we need to answer] One evident shortcoming of the Maze task is that it has been used on only a limited set of materials.
 Some of this is due to the fact that it isn't used as widely as other incremental processing methods, and some is due to inherent limitations of the paradigm.
@@
 
-We find a way around this problem by creating a version of the Maze task interface where participants correct their mistakes and continue with the sentence rather than being discontinued.
+[Our contribution is...] We find a way around this problem by creating a version of the Maze task interface where participants correct their mistakes and continue with the sentence rather than being discontinued.
 This allows for multi-sentence items to be run using Maze, which we verify by running Maze on the Natural Stories corpus.
@@
 
-The Maze tasked as historically done involves presenting in-lab participants (often college students) with materials that consist of individual, unrelated sentences, some of which exemplify the relevant properties (often with a Latin Square design).
-While this is a common psycholinguistic paradigm, it isn't the only one, and Maze would be more versatile (and thus more appealing as an alternative to eye-tracking and SPR) if it also worked on other types of materials, such as longer texts used in research in discourse or surprisal.
-There's also interest in comparing neural net performance to that of human behavioral data on naturalistic corpora, which necessitates behavioral corpora.
+[review the terminate thing, and why it was there] The advantage of the Maze task is that it forces incremental processing and automatically excludes inattentive participants by terminating a sentence when a participant makes an error.
+Thus, only participants who have processed the sentence up that point and are paying attention contribute data at critical regions later in the sentence.
+However, this causes data loss ecause a participant does not see the rest of the sentence after a mistake, and therefore does not contribute reading time.
+This data loss problem was relatively minor in traditional G-maze tasks, which had hand-crafted (and thus usually obviously wrong) distractors and attentive, in-person participants.
+This leads to a low per-word error rate, which does not compound too badly over the individual sentences, but it becomes prohibitive if items grow longer than one or two sentences.
+TODO: look up what the loss rates were!
 
-One major concern with Maze is data loss -- whenever a participant makes a mistake, they don't see the rest of the sentence and thus they do not contribute RT data for any words in the rest of the sentence.
-TODO cite how much this happened in old paper.
-While this hasn't been a problem with in lab G-maze (error rates were fairly low and items not too long), it's more of a problem with A-maze on crowd-sourced participants, due to both compontents.
+[ how this causes problems for 2020 and theoretically] 
+However, as @boyceMazeMadeEasy2020 found, this data loss is much more extreme with A-maze materials and crowd-sourced participants.
+It's unclear to what extent the errors are from participants were guessing randomly or from auto-generated distractors that in fact fit the sentence.
+Both of these are clearly contributing somewhat.
 In particular, Boyce et al 2020 found that some of the distractors, especially early in the sentence are problematic, and therefore data-loss is high.
 
-However, even with G-maze quality materials, and high quality subjects, the Maze paradigm with eventually run into trouble as materials get longer.
-Even assuming a very low per word error rate , the errors will build up as materials get longer, and few participants will make it to the end.
+Even if we had attentive participants and good distractors, the Maze paradigm will eventually run into problems with data loss as materials get longer: the errors will build up and few participants will make it to the end.
 For instance, with a 1% error rate, `r round(.99**15*100)`% would complete each 15-word sentence, but only `r round(.99**50*100)`% of participants would complete each 50 word vignette, and `r round(.99**200*100)`% of 200 word passages.
-One potential option would be to have participants skip to the next sentence (rather than next item) when they make a mistake.
-However, having missed some context from the last sentence, this may be jarring and and confusing for participants.
-Their RTs may also be unreliable  due to having missed context and discourse cues.
+In order to run longer materials, we need something to do when participants make a mistake, other than terminate the entire item.
+One option would be skip to the next sentence (rather than next item) when a participant makes a mistake.
+It could be jarring and confusing to be jumped from the middle of one sentence to start of the next and to be missing some context from the end of the sentence.
+In addition, this skips over context and potential discourse cues which makes the RTs potentially unreliable.
 
 To solve this conundrum, we decide to get rid of the source of the problem: the terminating on errors.
-Instead, we let participants continue after they make a mistake.
+Instead, we let participants continue on when they make a mistake.
 To do this, we present them with an error message, and wait until they select the correct option, before continuing the sentence as normal.
-This was easily implemented as a new option built on the Ibex Maze code used in Boyce et al 2020.
-We record both the RT to the first click and also the total RT until they select the correct answer as separate values.
+We added this "error-correction" option to the Ibex Maze implementation (TODO repo link); which can be toggled on or off.
+The code records both the RT to the first click and also the total RT until the correct answer is selected as separate values.
+TODO diagram goes here
 
-This in principle solves several problems.
+[ why this error thing seems really cool and could work] 
+This simple tweak in principle solves several problems.
 Firstly, it allows for running longer materials, potentially expanding the range of questions Maze is suitable for.
-Secondly, it compensates for some of the shortcomings of Maze identified in Boyce et al 2020 -- distractors might still cause participants to make unavoidable mistakes, but at least they still see the rest of the sentence and we get their data.
+Now, when participants make the occassional mistake or misclick, they can continue with the story, seeing every word of it.
+Thus, they provide reading times in later sentences and see all the context.
+Seeing the error message and resolving it is quick and does not disrupt the (already slow)  reading process too much.
+
+Secondly, this error-correction compensates for some of the shortcomings of A-Maze identified in @boyceMazeMadeEasy2020; distractors might still cause participants to make unavoidable mistakes, but at least the still see the rest of the sentence and we get their data.
 We believe (although don't have evidence) that being able to correct mistakes may reduce frustration related to unavoidable mistakes from impossible choices and accidental misclicks.
@@
 
-Even if all post-mistake data are discarded before analysis, they can still help address the "distractors or participants" question around errors mentioned in Boyce.
-When participants continue with the sentence, it is easy to calculate the per-word error rate for participants, without having to account for the censoring present in traditional Maze (if you already made a mistake, we can't know if you would have made another one; with error-correction, we do know if you did).
-This allows us to exclude participants on the basis of their task accuracy, ensuring RT data comes from participants who were paying attention and doing the task.
+Even if researchers exclude all post-mistake data from analysis, this process can still address the question of whether errors are due to inattentive participants or bad distractors (as discussed in @boyceMazeMadeEasy2020, section XX).
+When we have a participants selection for every word, we can easy calculate a per-word error rate for each participant, without having to address the censoring present in error-terminating Maze (where we can't know if a participant would have made more mistakes after the first; with error correction, we know how may more they did make).
+If participants are guessing, we'll see high per-word error rates, if the early distractors are bad, we'll see low per-word error rates (with errors concentrated at the start of the sentences).
+This per-word error rate metric measures participants task accuracy, and allows us to exclude data from inattentive participants.
 This provides a convenient and clear-cut way of controlling data quality after the fact.
@@
 
-We test this method using the Natural Stories corpus, which consists of 10 passages, some fiction and some non-fiction.
-We think this is a fairly severe test: the passages are long (about 1000 words each), they contain a lot of punctuation (for instance quoted speech), a number of proper nouns (which are likely out of vocabulary for the autogeneration model), and a number of low frequency grammatical constructions.
+We test this error-correction Maze on the Natural Stories corpus.
+The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages, some fiction and some non-fiction, that are designed to read fluently to native speakers.
+At the same time, the passages are long, and contain copious punctuation (including quoted speech), many proper nouns, and low frequency grammatical constructions.
 However, they are intended to read naturally to native English speakers.
+Taken together, these properties make this a severe test of our process.
+Long passages test the error-correction process and how well participants can track a story.
+Punctuation, proper nouns, and infrequent constructions all might stress the NN model generating distractors.
 We choose this because if participants can succeed at doing the Maze task and understanding it on this set of materials, we think they are likely to on basically any naturalistic text (many of which might be shorter or easier to follow).
-The corpus also comes with binary-choice comprehension questions, 6 per story.
 
-THis corpus has been used for SPR, so there's the possiblity of comparison.
+The corpus also comes with binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
+@futrellNaturalStoriesCorpus2020 ran SPR experiments on the corpus, so we can also compare patterns between Maze and SPR.
 
-In addition to testing that Maze works at all on this type of material, we also wanted to check for two other desired characteristics.
-That the materials can be comprehended even when read with Maze, and that commonly found psycholinguistic effects of length and surprisal show.
+TODO fix this part:
  whether surprisal and other coarse-level processing effects would appear (as a sanity check) -- eye-tracking and SPR show these effects robustly and in relatively well established patterns, and for Maze to be trusted as an alternative, it would need to as well.
@@
 
-To satisfy these various desires, we decided to run a Maze task on the Natural Stories corpus (Futrell et al 2020).
-This corpus consists of 10 passages, each about 1000 words long.
-They are naturalistic, although they were designed to contain a fair number of low frequency constructions such as BLAH and BLAH.
-(We do not specifically analyse this.) They also come with comprehension questions, and SPR data from the corpus is available.
-THis seemed like an appropriate severe test, and the corpus was available.
-SEE FOOBAR for examples of the materials and questions.
+This corpus allows us to address a set of sequential questions.
+- Firstly, whether this method words at all and whether participants can read stories in the odd paradigm.
+- Secondly, that participants can comprehend the stories (as measured by the comprehension questions). 
+- Thirdly, that the RT profiles show surprisal and other course-level processing effects.
+- Fourthly, although we do not examine it here, the data set can be examined at the targetted syntactic phenomena or compared with NN models. (TODO: unclear whether we should include parts of this).
 
 # Methods
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
 
+## Procedure
+
+Participants gave their informed consent, saw instructions, and saw the practice item and practice comprehension questions before reading the main story using the Maze task.
+After they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions and were debriefed participants.
 ## Participants
@@
 
-We took the corpus and split it up into sentences (this involved some judgement calls due to quoted speech) because the Maze material generation works at the sentence level.
+We took the corpus and split it up into sentences (this involved some judgment calls due to quoted speech) because the Maze material generation works at the sentence level.
 We ran the sentences through A-maze generation using TODO foo, bar parameters.
@@
 
-We also wrote a short passage to use as practice, with 2 comprehension questions (double check!!).
+We also wrote a short passage to use as practice, with 2 comprehension questions (double check!!). All of these materials are available at TODO
+We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
+<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
 
-## Procedure
+1000 words is a lot for Maze especially when it needs to be done in one chunk (whereas we'd put pause screens in previous maze experiments, I didn't here because it could disrupt the story), so we had each participant only read one story.
+This was prefaced by a short practice item (~100 words) that I wrote, which had 2 practice questions.
 
-Participants gave their informed consent, saw instructions, saw the practice item and practice comprehension questions, and then completed the story using the Maze task, then the 6 main comprehension questions.
-At the end of the experiment, we asked how their experience was, gathered demographic questions (which ones), and debriefed participants.
+At the end, we collected some optional demographics and debriefed participants.
+
+We recruited 100 participants, across the 10 stories.
+(Stories were randomized, but not balanced, so we'll see how balanced the data is.) We estimated that this would take around 20 minutes (given how long past studies had taken) and paid 3.50.
+In fact, it appears to have taken most attentive participants around 20 minutes to read the story, and thus closer to 25 minutes for the overall study.
 
 ## Data analysis
+
+```{r participants}
+
+data <- read_rds(paste0(here(),"/Data/cleaned.rds"))
+
+data_filt <- data %>% 
+  filter(native %in% c("ENG", "English", "ENGLISH", "english")) #I peeked at what people put that semantically maps to english
+
+data_stories <- data_filt %>% 
+  select(type, subject) %>% 
+  unique() %>% 
+  group_by(type) %>% 
+  tally() %>% 
+  filter(type!="practice")
+
+data_error_summ <- data_filt %>% 
+  mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
+  group_by(subject, num_correct) %>%
+  filter(type!="practice") %>% 
+  filter(rt<5000) %>% 
+  summarize(mean_rt=mean(rt),
+            pct_correct=mean(correct.num)) %>% 
+  mutate(good_comp=ifelse(num_correct>4,"5 or 6 correct","4 or fewer correct")) %>% 
+    mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct"),
+           is.attentive=ifelse(pct_correct>.8,1,0))
+
+some <- data_filt %>% select(subject, num_correct) %>% 
+  unique() %>% 
+  left_join(data_error_summ, by=c("subject", "num_correct")) %>% 
+  mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct")) %>% 
+  group_by(num_correct,accurate)
+
+data_good <- data_filt %>% 
+  left_join(data_error_summ, by=c("subject", "num_correct")) %>% 
+  filter(type!="practice") %>% 
+  filter(pct_correct>.8) %>% 
+  mutate(is.correct=ifelse(correct=="yes",1,0))
+
+data_non_error <- data_good %>% 
+  filter(is.correct==1) %>% 
+  filter(word_num>0)
+
+data_before <- data_good %>% 
+  mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
+  group_by(sentence, subject) %>% 
+  fill(word_num_mistake) %>% ungroup() %>% 
+  mutate(after_mistake=word_num-word_num_mistake,
+         after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
+  filter(after_mistake==0) %>% 
+  filter(is.correct==1) %>% 
+  filter(word_num>0)
+
+data_sentence <- data_before <- data_good %>% 
+  mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
+  group_by(sentence, subject) %>% 
+  fill(word_num_mistake) %>% ungroup() %>% 
+  mutate(after_mistake=word_num-word_num_mistake,
+         after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
+  group_by(sentence, subject) %>% 
+  summarize(total_mistake=sum(after_mistake)) %>% 
+  mutate(sent.correct=ifelse(total_mistake==0,1,0))
+```
+
 We used `r cite_r("r-references.bib")` for all our analyses.
-We excluded participants who did not indicate that their native language was English. This left us with YY participants.
+Of the `r data %>% select(subject) %>% unique() %>% nrow()` participants who each read one story, we excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
+After this exclusion, each of the 10 stories was read between `r min(data_stories$n)` and `r max(data_stories$n)` times.
 
@@
 
-## Qualitative stuff
+## Can people read the stories?
+
+(ref:error-cap) A.
+Correlation between a participant's accuracy on the Maze task (fraction of words selected correctly) and their average reaction time (in ms).
+Many participants (marked in green) attained accuracies of >80%; others (in red) appear to be randomly guessing and were excluded from further analysis.
+B.
+Performance on comprehension questions.
+Participants who had >80% task accuracy tended to do well on comprehension questions; those who were at chance on the task were also at chance on comprehension questions.
+
+```{r errors, out.width="100%", out.height="25%", fig.pos='t', fig.cap="(ref:error-cap)"}
+error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
+  geom_point(size=1)+
+  labs(x="Fraction words selected correctly",
+       y="Mean Reaction Time (ms)")+
+  coord_cartesian(xlim=c(0.4,1), ylim=c(0,1600), expand=F)+
+    scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
+  guides(color=FALSE)
+
+comp_plot <- ggplot(some, aes(x=num_correct,fill=accurate))+
+  geom_bar(position="dodge")+
+  facet_grid(.~accurate)+
+  labs(x="Comprehension questions correct (out of 6)", y="Participants")+
+  scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
+guides(fill=FALSE)
+
+error_plot+comp_plot+plot_annotation(tag_levels="A")
+
+```
+
+The first question of interest is whether participants can do the Maze task on such a long text and whether the auto-generated distractors are clear enough to allow high rates of accuracy.
+To examine this, we calculate each participants by-word error rate and compare to their average RT (\@ref(fig:errors)).
+(To avoid biases the average RT estimate if a participant got distracted or stepped away from the computer, RTs greater than 5 seconds were excluded.) A number of participants do the task well, with high per-word accuracies reaching 99%.
+This indicates that the distractors are mostly not a problem since some people can reliably choose the right option.
+It also indicates that some participants can maintain focus on the task for the whole story.
+These participants who are careful tend to take around 1 second per word selection.
+In contrast, other participants are going much faster and appear to be randomly guessing.
+We believe this bimodal distribution is a reflection of the mixture of workers on Mturk -- some sped through the task to get paid for as little time as possible, while others took the task seriously.
+This comparision allows us to easily see who was paying attention to the task; we  only analyse data from participants with at least 80% accuracy.
+
+When participants do make errors, they correct them quickly, generally within one second, which means it's probably not disrupting their reading much.
+
+The next question qualitative check is whether participants are able to comprehend the story.
+We tested this using 6 binary-choice questions after each story.
+As seen in @ref(fig:comp-q), most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the maze task were also at chance on the comprehension questions.
+One concern is that they comprehension questions may have been too easy, as some of the stories are about factual events (ex.
+tulip mania) or folk stories.
+We can't rule out that some participants may have already known some of the answers, but participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet.
+
+This provides some evidence that people are able to understand and remember details of what they are reading during the Maze task.
+While the comprehension scores are worse than those from SPR (cite), we must keep in mind that Maze is a much slower task, and therefore participants have more of a delay between the start of the story and answering comprehension questions.
+We take comprehension question performance as a less reliable indicator of attention that maze task peformance, and so use maze task performance as our exclusion metric.
+
+While we are planning on analysing data after mistakes, it might be interesting to know how much of the data is pre-mistake, and how many of the sentences are completed error-free.
+If we only exclude the errors and the first words of sentences, there are `r nrow(data_non_error)` words.
+If we also exclude all words after mistakes, there are `r nrow(data_before)` words.
+We will exclude only a couple of words after each mistake, but we could also opt to analyse only the pre-error sections.
+Of the `r nrow(data_sentence)` sentences that good participants completed, `r filter(data_sentence, sent.correct==1) %>% nrow()` were completed entirely correctly.
 
@@
 
+## LMER stuff
+
+```{r}
+subject_att <- data_error_summ %>% 
+  select(subject, is.attentive)
+
+data_low_error <- data_filt %>% 
+  left_join(subject_att, by="subject") %>% 
+  filter(is.attentive==1) %>% 
+  filter(type!="practice")
+
+  data_error_free <- data_low_error %>% 
+    mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
+    group_by(sentence, subject) %>% fill(word_num_mistake) %>% ungroup() %>% 
+    mutate(after_mistake=word_num-word_num_mistake,
+           after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
+    filter(correct=="yes") %>% 
+    filter(!after_mistake %in% c(1,2))
+
+data_no_first <- data_error_free %>% filter(word_num!=0)
+
+data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  select(subject, word_num, word, rt, sentence, type)
+
+data_post_only <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  filter(!after_mistake==0) %>% 
+    select(subject, word_num, word, rt, sentence, type)
+
+data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  filter(after_mistake==0) %>% 
+  select(subject, word_num, word, rt, sentence, type)
+
+data_stories <- data_ready %>% select(type, subject) %>% 
+  unique() %>% 
+  group_by(type) %>% 
+  tally()
+
+data_anything_goes <- data_filt %>% 
+  filter(type!="practice") %>% 
+  filter(correct=="yes") %>% 
+    select(subject, word_num, word, rt, sentence, type)
+
+```
+
+`r data %>% select(subject) %>% unique() %>% nrow()` participants read naturalistic stories from the natural stories corpus. Each participant read 1 story.
+
+We exclude
+
+- participants who do not report English as a native language (`r data_filt %>% select(subject) %>% unique() %>% nrow()` remaining)
+- participants who do not get 80% of the words correct (`r data_low_error %>% select(subject) %>% unique() %>% nrow()` remaining)
+- practice items (`r data_low_error %>% nrow()` words remaining)
+- words that were wrong or were within two after a mistake (`r data_error_free %>% nrow()` words remaining)
+- the first word of every sentence (didn't have a real distractor, RT is measured slightly differently) (`r data_no_first %>% nrow()` words remaining)
+- words with RTs <100 or >5000 (<100 we think is likely a recording error, or at least not reading the words at all, >5000 is likely getting distracted) (`r data_ready %>% nrow()` words remaining)
+
+Within the filtered data, each story was read between `r min(data_stories$n)` and `r max(data_stories$n)` times, for an average of `r mean(data_stories$n)`.
+
+We also do the analyses on only the words before mistakes (per sentence) (`r data_pre_error %>% nrow()` words)
+
+From the modelling side:
+(After attempts without doing this filtering) we only include words which are single token and known words in each of the models vocabularies.
+We also only include words with frequencies.
+This is roughly equivalent to excluding words with punctuation.
+
+We use as predictors:
+
+- length in characters of stripped word
+- unigram frequency of word.
+Frequencies for words are calculated using word_tokenize on the gulordava train data and counting up instances.
+(This tends to tokenize off punctuation, but is capitalization sensitive).
+Frequencies are represented as log2 of the expected occurances in 1 billion words.
+
+Surprisals are measured in bits.
+
+- ngram (5-gram KN smoothed)
+- GRNN
+- Transformer-XL
+
+We center the predictors for the brms models, but don't rescale them.
+
+The model formula we use is rt ~ surp* length +  freq * length + past_surp * past_length + past_freq * past_length.
+
+# Frequentist LMs
+
+```{r}
+#mean by subject for each word
+labelled_pre_error <- read_rds(paste0(here(),"/Analysis/pre_error.rds"))
+d_lm <- labelled_pre_error %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
+                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
+  summarize(mean_rt=mean(rt))
+```
+
+## Models
+
+We want to be able to do nested model comparision, so we want models with 
+- only length, frequency fx
+- each 1 surprisal model as predictor
+- all the surprisals
+
+For now, no interactions between surprisal and others, and if we include a predictor, include both current and lagged of it.
+
+```{r}
+
+no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+all_surp <- lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   grnn_center + past_c_grnn +
+                   txl_center + past_c_txl +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+ngram_grnn <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   grnn_center + past_c_grnn +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+ngram_txl <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   txl_center + past_c_txl +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+grnn_txl <-  lmer(mean_rt ~ grnn_center + past_c_grnn +
+                   txl_center + past_c_txl +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+no_lag <-  lmer(mean_rt ~ ngram_center +
+                   grnn_center +
+                   txl_center + 
+                   freq_center*length_center+(1|word), data=d_lm)
+
+no_surp_lag <-  lmer(mean_rt ~ ngram_center +  
+                   grnn_center + 
+                   txl_center +                    freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
+
+  only_surp_lag <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   grnn_center + past_c_grnn +
+                   txl_center + past_c_txl +
+                   freq_center*length_center+(1|word), data=d_lm)
+```
+
+```{r}
+
+anova(all_surp, txl_only)
+anova(all_surp, grnn_only)
+anova(all_surp, ngram_only)
+
+anova(txl_only, no_surp)
+anova(grnn_only, no_surp)
+anova(ngram_only, no_surp)
+
+anova(all_surp, ngram_grnn)
+anova(all_surp, ngram_txl)
+anova(all_surp, grnn_txl)
+
+anova(ngram_grnn, ngram_only)
+anova(ngram_grnn, grnn_only)
+anova(ngram_txl, ngram_only)
+anova(ngram_txl, txl_only)
+anova(grnn_txl, grnn_only)
+anova(grnn_txl, txl_only)
+
+anova(all_surp, no_lag)
+
+anova(all_surp, no_surp_lag)
+
+anova(no_surp_lag, no_lag)
+
+anova(all_surp, only_surp_lag)
+
+anova(only_surp_lag, no_lag)
+```
+In general, adding to the model makes it better.
+
+The model with all 3 surprisal predictors is better than any model with only one. Any one surprisal predictor is better than no surprisals predictors.
+
+However, adding ngram predictors to a model that already has txl & grnn does not help. 
+In other cases, adding the 3rd surprisal source to the other two does help.
+
+Ngram+Grnn is not better than Grnn only. Otherwise, pairs are better than singletons. This suggests that Ngram's info is a subset of GRNN, but not a subset of TXL.
+
+Past surprisal predictors don't help (with or without past freq,length effects in the models), but past freq,length do (with or without past surprisal predictors).
+
+## What if we don't exclude participants
+
+One potentially useful comparison with past SPR/eye-tracking is if we are more inclusive of participants, do we get data that looks more like theirs. (Perhaps our exclusion criteria are more robust?)
+
+This is all data from participants who speak English that wasn't a mistake. (So, includes bad participants and post-mistake data.)
+
+```{r freq_all}
+#mean by subject for each word
+labelled_anything_goes <- read_rds(paste0(here(),"/Analysis/non-selective.rds"))
+d_lm2 <- labelled_anything_goes %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
+                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
+  summarize(mean_rt=mean(rt))
+```
+
+```{r}
+
+no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
+
+ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
+
+grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
+
+txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
+
+#summary(no_surp)
+#summary(ngram_only)
+#summary(grnn_only)
+#summary(txl_only)
+
+```
+
+BRM models from this same very permissive data. 
+```{r brm-models}
+
+any_grnn <- read_rds(paste0(here(),"/Analysis/brm_grnn_any_center.rds"))
+any_txl <- read_rds(paste0(here(),"/Analysis/brm_txl_any_center.rds"))
+any_ngram <- read_rds(paste0(here(),"/Analysis/brm_ngram_any_center.rds"))
+
+#summary(any_grnn)
+#summary(any_txl)
+#summary(any_ngram)
+
+```
+
+```{r}
+
+show_summary <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=1),
+           upper=round(.upper, digits=1),
+           E=round(.value, digits=1),
+           `CI`=str_c("[",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           ) %>% 
+    select(Term, `E`, `CI`,`P`)
+
+  stats
+}
+a <- show_summary(any_txl) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
+b <- show_summary(any_grnn) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
+c <- show_summary(any_ngram) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
+
+summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`E`,`CI`,`P`)) %>% 
+  select(Term, `E_5-gram`, `CI_5-gram`, `P_5-gram`, `E_GRNN`, CI_GRNN, P_GRNN, E_TXL, CI_TXL, P_TXL)
+
+library(xtable)
+print(xtable(summ,digits=c(0,0,1,0,2,1,0,2,1,0,2)), include.rownames=FALSE)
+```
+## BRM models
+
+We include a by-subject effect for everything, and a by_word random intercept (full mixed effects).
+
+Priors:
+
+- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
+- normal(0,500) for beta and sd -- we don't really know what effects are
+- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way
+
+```{r}
+  priors <- c(
+      set_prior("normal(1000, 1000)", class="Intercept"),
+      set_prior("normal(0, 500)", class="b"),
+      set_prior("normal(0, 500)", class="sd"),
+      set_prior("lkj(1)",       class="cor"))
+
+```
+
+### On pre-error data only
+
+```{r}
+
+show_summary <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=1),
+           upper=round(.upper, digits=1),
+           E=round(.value, digits=1),
+           `CI`=str_c("[",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           ) %>% 
+    select(Term, `E`, `CI`,`P`)
+
+  stats
+}
+a <- show_summary(brm_txl_interact) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
+b <- show_summary(brm_grnn_interact) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
+c <- show_summary(brm_ngram_interact) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
+
+summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`E`,`CI`,`P`)) %>% 
+  select(Term, `E_5-gram`, `CI_5-gram`, `P_5-gram`, `E_GRNN`, CI_GRNN, P_GRNN, E_TXL, CI_TXL, P_TXL)
+
+library(xtable)
+print(xtable(summ,digits=c(0,0,1,0,2,1,0,2,1,0,2)), include.rownames=FALSE)
+```
+
+### On post-error as well
+
+```{r}
+show_summary_2 <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P-value equivalent` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=2),
+           upper=round(.upper, digits=2),
+           `Credible Interval`=str_c("[",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           Estimate=round(.value, digits=2)) %>% 
+    select(Term, Estimate, `Credible Interval`, `P-value equivalent`)
+
+  stats
+}
+```
+
+## Comparision with SPR
+
+We want to compare how good our participants were on the comprehension questions to how good people were on the SPR experiment reported in Futrell et al (2017).
+They had each story read about 100 times across 181 participants (participants generally read 5 stories).
+This means they have about 10 times as much data as we do.
+However, reading times in a story were excluded based on the comprehension of that story, so it makes sense to compare per-story accuracies with our per-story accuracies.
+
 # Discussion
@@
 While eventually we do want to figure out models of more natural reading, this might be a useful complement b/c we don't have to deal with nasty spillover.
-
+Might be extra good to do with pauses for mistakes to fix the incentives. 
 \newpage

### 2021-01-27 Papers/Paper/manuscript.Rmd
more writing!

@@
 # Methods
+We constructed A-maze distractors for the Natural stories corpus (CITE) and recruited 100 MTurkers to each read one the stories in the Maze paradigm.
+On the basis of previous uses of the Maze paradigm, we estimated that participants would do the task at roughly 1 second per word, or around 17 minutes for a 1000 word passage, so we only gave each participant one passage to read to keep the task manageable.
+
+## Participants
+We recruited 100 participants from Amazon Mechanical Turk, and paid each participant 3.50 dollars for what we estimated was a 20 minute task.
+In fact, it appears to have taken most attentive participants around 20 minutes to read the story, and thus closer to 25 minutes for the overall study.
 
 ## Procedure
+Participants first gave their informed consent and saw task instructions.
+Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading the main story in the A-maze task.
+After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, saw an debriefing and were given a code to enter into Mturk for payment.
+The experiment was implemented in Ibex (CITE) and the experimental code is availabe at REPO.
 
-Participants gave their informed consent, saw instructions, and saw the practice item and practice comprehension questions before reading the main story using the Maze task.
-After they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions and were debriefed participants.
-## Participants
-We recruited XX participants from Mechanical Turk, and paid each participant YY dollars.
-
-## Material
-
-We took the corpus and split it up into sentences (this involved some judgment calls due to quoted speech) because the Maze material generation works at the sentence level.
-We ran the sentences through A-maze generation using TODO foo, bar parameters.
-Materials are available in TODO: anon repo.
-We set up the materials in Ibex, using Maze to display things.
-
-We also wrote a short passage to use as practice, with 2 comprehension questions (double check!!). All of these materials are available at TODO
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
-
-1000 words is a lot for Maze especially when it needs to be done in one chunk (whereas we'd put pause screens in previous maze experiments, I didn't here because it could disrupt the story), so we had each participant only read one story.
-This was prefaced by a short practice item (~100 words) that I wrote, which had 2 practice questions.
-
-At the end, we collected some optional demographics and debriefed participants.
-
-We recruited 100 participants, across the 10 stories.
-(Stories were randomized, but not balanced, so we'll see how balanced the data is.) We estimated that this would take around 20 minutes (given how long past studies had taken) and paid 3.50.
-In fact, it appears to have taken most attentive participants around 20 minutes to read the story, and thus closer to 25 minutes for the overall study.
+## Materials
+We took the texts of the Natural Stories corpus (CITE) and split it into sentences because the A-Maze process works on the sentence level.
+(This involved some judgment calls due to quoted speech).
+We used the original comprehension questions provided in the Natural Stories corpus.
+To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions.
+We then ran the sentences of the items through the A-maze generation process.
+We updated the code of the generation process to fix some issues identified in that paper; the code used is available at REPO.
+The underlying method did not change, but features such as appropriately handling more types of punctuation (needed for the Natural Stories corpus) were added.
+We ran the materials using FOOBAR paramters.
+We used the auto-generated distractors as they were, without checking them for quality.
+Materials are available at REPO.
+TODO other notable changes to the codebase??
 
@@
 
+Of the `r data %>% select(subject) %>% unique() %>% nrow()` participants who each read one story, we excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
 We used `r cite_r("r-references.bib")` for all our analyses.
-Of the `r data %>% select(subject) %>% unique() %>% nrow()` participants who each read one story, we excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
-After this exclusion, each of the 10 stories was read between `r min(data_stories$n)` and `r max(data_stories$n)` times.
 
@@
 
-## Can people read the stories?
+## Reading stories in the Maze task
 
@@
 Correlation between a participant's accuracy on the Maze task (fraction of words selected correctly) and their average reaction time (in ms).
-Many participants (marked in green) attained accuracies of >80%; others (in red) appear to be randomly guessing and were excluded from further analysis.
+Many participants (marked in green) chose the correct word >80% of the time; others (in red) appear to be randomly guessing and were excluded from further analysis.
 B.
@@
 
-```{r errors, out.width="100%", out.height="25%", fig.pos='t', fig.cap="(ref:error-cap)"}
+```{r errors, out.height="25%", fig.pos='t', fig.cap="(ref:error-cap)"}
 error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
@@
 
-The first question of interest is whether participants can do the Maze task on such a long text and whether the auto-generated distractors are clear enough to allow high rates of accuracy.
-To examine this, we calculate each participants by-word error rate and compare to their average RT (\@ref(fig:errors)).
-(To avoid biases the average RT estimate if a participant got distracted or stepped away from the computer, RTs greater than 5 seconds were excluded.) A number of participants do the task well, with high per-word accuracies reaching 99%.
-This indicates that the distractors are mostly not a problem since some people can reliably choose the right option.
-It also indicates that some participants can maintain focus on the task for the whole story.
-These participants who are careful tend to take around 1 second per word selection.
-In contrast, other participants are going much faster and appear to be randomly guessing.
-We believe this bimodal distribution is a reflection of the mixture of workers on Mturk -- some sped through the task to get paid for as little time as possible, while others took the task seriously.
-This comparision allows us to easily see who was paying attention to the task; we  only analyse data from participants with at least 80% accuracy.
+Our first concern was whether the paradigm would work at all. To our relief, it did. Some participants were able to complete the Maze task with high accuracy and understand the story while doing so.
 
+We first wanted to check for participant accuracy, which reflects both how well participants can focus on and navigate the task and whether the auto-generated distractors are of acceptable quality.
+We calculated the per-word error rate for each participant and compared it against their average reaction time.
+(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.) As seen in Figure \@ref(fig:errors), there is a cluster of participants who make relatively few errors (marked in green), with some reaching 99% accuracy.
+This indicates that the distractors were generally appropriate and that some participants can maintain focus on the task for the whole story.
+These careful participants tend to take around 1 second for each word selection, which is much slower than other paradigms CITE SPR something.
+
+In contrast, we also have a cluster of participants (in red) who we infer sped through the task clicking randomly.
+This distribution is likely due to the mix of workers on Mturk; as we did not use qualification cutoffs.
+
+Another important check is whether participants were able to comprehend the story,which we tested using 6 binary-choice comprehension questions after each story.
+We count how many questions each participant got right, and group them by their performance on the Maze task.
+As seen in @ref(fig:comp-q), most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions.
+Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet.These are not especially hard questions, and we can't rule out that some participants may have been able to guess the answers without reading the story.
+Nonetheless, this provides preliminary evidence that people can understand and remember details of stories they read during the Maze task.
+
+Given that Maze task accuracy and comprehension question accuracy are correlated, but task accuracy is finer-grained and less noisy we use task performance as our exclusion metric.
+This easy to calculate speed and accuracy comparision makes it easy to see who was paying attention to the task; we exclude the random guessers and only analyse data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
+
+## Error patterns
 When participants do make errors, they correct them quickly, generally within one second, which means it's probably not disrupting their reading much.
-
-The next question qualitative check is whether participants are able to comprehend the story.
-We tested this using 6 binary-choice questions after each story.
-As seen in @ref(fig:comp-q), most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the maze task were also at chance on the comprehension questions.
-One concern is that they comprehension questions may have been too easy, as some of the stories are about factual events (ex.
-tulip mania) or folk stories.
-We can't rule out that some participants may have already known some of the answers, but participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet.
-
-This provides some evidence that people are able to understand and remember details of what they are reading during the Maze task.
-While the comprehension scores are worse than those from SPR (cite), we must keep in mind that Maze is a much slower task, and therefore participants have more of a delay between the start of the story and answering comprehension questions.
-We take comprehension question performance as a less reliable indicator of attention that maze task peformance, and so use maze task performance as our exclusion metric.
 
@@
 
-```{r}
+```{r, include=F}
 
@@
 
-library(xtable)
-print(xtable(summ,digits=c(0,0,1,0,2,1,0,2,1,0,2)), include.rownames=FALSE)
 ```
@@
 }
+brm_txl_interact <- read_rds(paste0(here(),"/Analysis/brm_txl_interact.rds"))
+brm_grnn_interact <- read_rds(paste0(here(),"/Analysis/brm_grnn_interact.rds"))
+brm_ngram_interact <- read_rds(paste0(here(),"/Analysis/brm_ngram_interact.rds"))
 a <- show_summary(brm_txl_interact) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))

### 2021-01-29 Papers/Paper/manuscript.Rmd
work on manuscript

@@
 library(matrixStats)
+library(tidyverse)
+library(readr)
+library(brms)
+library(lme4)
+library(rstan)
+library(tidybayes)
+library(knitr)
+library(mgcv)
+library(mgcViz)
+library(tidymv)
+library(rsample)
+library(cowplot)
+library(scales)
 r_refs("r-references.bib")
@@
 Thus, only participants who have processed the sentence up that point and are paying attention contribute data at critical regions later in the sentence.
-However, this causes data loss ecause a participant does not see the rest of the sentence after a mistake, and therefore does not contribute reading time.
-This data loss problem was relatively minor in traditional G-maze tasks, which had hand-crafted (and thus usually obviously wrong) distractors and attentive, in-person participants.
-This leads to a low per-word error rate, which does not compound too badly over the individual sentences, but it becomes prohibitive if items grow longer than one or two sentences.
-TODO: look up what the loss rates were!
+However, this means we don't have data after a participant makes a mistake in an item.
+In traditional G-maze tasks, with hand-crafted distractors and attentive participants, this is a small issue as the low per-word error rate does not compound too badly over the individual sentences.
 
 [ how this causes problems for 2020 and theoretically] 
-However, as @boyceMazeMadeEasy2020 found, this data loss is much more extreme with A-maze materials and crowd-sourced participants.
-It's unclear to what extent the errors are from participants were guessing randomly or from auto-generated distractors that in fact fit the sentence.
-Both of these are clearly contributing somewhat.
-In particular, Boyce et al 2020 found that some of the distractors, especially early in the sentence are problematic, and therefore data-loss is high.
+However, this data loss was worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020].
+The high errors could be from participants guessing randomly or from auto-generated distractors that in fact fit the sentence.
+@boyceMazeMadeEasy2020 noted that some distractors, especially early in the sentence were problematic and caused considerable data loss.
 
-Even if we had attentive participants and good distractors, the Maze paradigm will eventually run into problems with data loss as materials get longer: the errors will build up and few participants will make it to the end.
-For instance, with a 1% error rate, `r round(.99**15*100)`% would complete each 15-word sentence, but only `r round(.99**50*100)`% of participants would complete each 50 word vignette, and `r round(.99**200*100)`% of 200 word passages.
+Given the appeals of using A-maze, it would be adventageous to get around this problem.
+Additionally, even with attentive participants and good distractors, the Maze paradigm will eventually run into problems as error rates compound over long materials, and few participants will make it to the end.
+For instance, with a 1% error rate, `r round(.99**15*100)`% would complete each 15-word sentence, but only `r round(.99**50*100)`% of participants would complete each 50 word vignette, and `r round(.99**200*100)`% each 200 word passages.
 In order to run longer materials, we need something to do when participants make a mistake, other than terminate the entire item.
-One option would be skip to the next sentence (rather than next item) when a participant makes a mistake.
-It could be jarring and confusing to be jumped from the middle of one sentence to start of the next and to be missing some context from the end of the sentence.
-In addition, this skips over context and potential discourse cues which makes the RTs potentially unreliable.
 
-To solve this conundrum, we decide to get rid of the source of the problem: the terminating on errors.
-Instead, we let participants continue on when they make a mistake.
+To resolve this, we introduce an *error-correction* variant of Maze, where we let participants continue on when they make a mistake.
 To do this, we present them with an error message, and wait until they select the correct option, before continuing the sentence as normal.
-We added this "error-correction" option to the Ibex Maze implementation (TODO repo link); which can be toggled on or off.
+We make this "error-correction" Maze available as an option in a modification of the Ibex Maze implementation introduced in CITE ME (TODO repo link).
 The code records both the RT to the first click and also the total RT until the correct answer is selected as separate values.
@@
 [ why this error thing seems really cool and could work] 
-This simple tweak in principle solves several problems.
-Firstly, it allows for running longer materials, potentially expanding the range of questions Maze is suitable for.
-Now, when participants make the occassional mistake or misclick, they can continue with the story, seeing every word of it.
-Thus, they provide reading times in later sentences and see all the context.
-Seeing the error message and resolving it is quick and does not disrupt the (already slow)  reading process too much.
-
-Secondly, this error-correction compensates for some of the shortcomings of A-Maze identified in @boyceMazeMadeEasy2020; distractors might still cause participants to make unavoidable mistakes, but at least the still see the rest of the sentence and we get their data.
-We believe (although don't have evidence) that being able to correct mistakes may reduce frustration related to unavoidable mistakes from impossible choices and accidental misclicks.
-Whether this also solves the data loss problem from the researcher perspective within a sentence depends on whether post-mistake data are high-quality and trustworthy; this is a hard-to-assess question of potential interest.
-
-Even if researchers exclude all post-mistake data from analysis, this process can still address the question of whether errors are due to inattentive participants or bad distractors (as discussed in @boyceMazeMadeEasy2020, section XX).
-When we have a participants selection for every word, we can easy calculate a per-word error rate for each participant, without having to address the censoring present in error-terminating Maze (where we can't know if a participant would have made more mistakes after the first; with error correction, we know how may more they did make).
-If participants are guessing, we'll see high per-word error rates, if the early distractors are bad, we'll see low per-word error rates (with errors concentrated at the start of the sentences).
-This per-word error rate metric measures participants task accuracy, and allows us to exclude data from inattentive participants.
-This provides a convenient and clear-cut way of controlling data quality after the fact.
-
-It also starts to reduce the perverse incentives.
-With error-terminates Maze, the fastest way to get through the task is to select randomly, and it's quite quick because mistakes skip you to the next sentence.
-With error-correction Maze, randomly jamming buttons takes more effort, but is still an effective strategy.
-[Footnote: In discussing this work, we recieved the suggestion that one way to disincentivize random clicking is to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms or 1 sec) before being able to correct their mistake.
-This seems like a promising improvement that could be worth implementing and testing.]
+Our goal is that this error-correction will minimize the problems of having occassional poor auto-generated distractors by letting participants continue on and still read the whole item.
+It will also enable us to run long materials on Maze, thus expanding the range of questions that Maze can be used on.
 
@@
 We test this error-correction Maze on the Natural Stories corpus.
-The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages, some fiction and some non-fiction, that are designed to read fluently to native speakers.
-At the same time, the passages are long, and contain copious punctuation (including quoted speech), many proper nouns, and low frequency grammatical constructions.
-However, they are intended to read naturally to native English speakers.
+The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers.
+At the same time, the passages contain copious punctuation (including quoted speech), many proper nouns, and low frequency grammatical constructions.
 Taken together, these properties make this a severe test of our process.
-Long passages test the error-correction process and how well participants can track a story.
-Punctuation, proper nouns, and infrequent constructions all might stress the NN model generating distractors.
-We choose this because if participants can succeed at doing the Maze task and understanding it on this set of materials, we think they are likely to on basically any naturalistic text (many of which might be shorter or easier to follow).
+The materials are on the difficult side for our participants and for the language model we use to generate distractors.
+If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on basically any naturalistic text.
 
 The corpus also comes with binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
-@futrellNaturalStoriesCorpus2020 ran SPR experiments on the corpus, so we can also compare patterns between Maze and SPR.
-
-TODO fix this part:
- whether surprisal and other coarse-level processing effects would appear (as a sanity check) -- eye-tracking and SPR show these effects robustly and in relatively well established patterns, and for Maze to be trusted as an alternative, it would need to as well.
-Maze typically doesn't use comprehension questions because comprehending the local sentence structure is required to do the task.
-Thus, there isn't much information on how well participants comprehend and remember what they read.
-We wanted to know if participants understood and internalized the gist of what they were reading with Maze.
-
-This corpus allows us to address a set of sequential questions.
-- Firstly, whether this method words at all and whether participants can read stories in the odd paradigm.
-- Secondly, that participants can comprehend the stories (as measured by the comprehension questions). 
-- Thirdly, that the RT profiles show surprisal and other course-level processing effects.
-- Fourthly, although we do not examine it here, the data set can be examined at the targetted syntactic phenomena or compared with NN models. (TODO: unclear whether we should include parts of this).
+Using this corpus on the A-maze task allows us to address, first whether participants will read and understand long passages using A-maze with error correction, and second, whether the resulting RTs profiles will show expected patterns such as a linear relationship with surprisal.
 
 # Methods
-We constructed A-maze distractors for the Natural stories corpus (CITE) and recruited 100 MTurkers to each read one the stories in the Maze paradigm.
-On the basis of previous uses of the Maze paradigm, we estimated that participants would do the task at roughly 1 second per word, or around 17 minutes for a 1000 word passage, so we only gave each participant one passage to read to keep the task manageable.
-
-## Participants
-We recruited 100 participants from Amazon Mechanical Turk, and paid each participant 3.50 dollars for what we estimated was a 20 minute task.
-In fact, it appears to have taken most attentive participants around 20 minutes to read the story, and thus closer to 25 minutes for the overall study.
-
-## Procedure
-Participants first gave their informed consent and saw task instructions.
-Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading the main story in the A-maze task.
-After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, saw an debriefing and were given a code to enter into Mturk for payment.
-The experiment was implemented in Ibex (CITE) and the experimental code is availabe at REPO.
-
-## Materials
-We took the texts of the Natural Stories corpus (CITE) and split it into sentences because the A-Maze process works on the sentence level.
-(This involved some judgment calls due to quoted speech).
-We used the original comprehension questions provided in the Natural Stories corpus.
-To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions.
-We then ran the sentences of the items through the A-maze generation process.
-We updated the code of the generation process to fix some issues identified in that paper; the code used is available at REPO.
-The underlying method did not change, but features such as appropriately handling more types of punctuation (needed for the Natural Stories corpus) were added.
-We ran the materials using FOOBAR paramters.
-We used the auto-generated distractors as they were, without checking them for quality.
-Materials are available at REPO.
-TODO other notable changes to the codebase??
-
-## Data analysis
 
@@
 
-Of the `r data %>% select(subject) %>% unique() %>% nrow()` participants who each read one story, we excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
-We used `r cite_r("r-references.bib")` for all our analyses.
+We constructed A-maze distractors for the Natural stories corpus (CITE) and recruited 100 MTurkers to each read a story in the Maze paradigm.
+
+## Materials (good)
+We took the texts of the Natural Stories corpus (CITE) and split it into sentences because the A-Maze process works on the sentence level.
+(This involved some judgment calls due to quoted speech).
+We used the original comprehension questions provided in the Natural Stories corpus.
+To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions.
+We then ran the sentences of the items through the A-maze generation process.
+We updated the code of the generation process to fix some issues identified in that paper; the code used is available at REPO.
+The underlying method did not change, but features such as appropriately handling more types of punctuation (needed for the Natural Stories corpus) were added.
+We ran the materials using FOOBAR paramters.
+We used the auto-generated distractors as they were, without checking them for quality.
+Materials are available at REPO.
+TODO other notable changes to the codebase??
+
+## Participants
+We recruited 100 participants from Amazon Mechanical Turk, and paid each participant 3.50 dollars.
+We estimated that reading a 1000 word story would take 17 minutes at 1 word/second (based on previous work), so we expected the task to take around 20 minutes total.
+In fact, it appears to have taken most attentive participants around 20 minutes to read the story, and thus closer to 25 minutes for the overall study.
+We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
+
+## Procedure
+Participants first gave their informed consent and saw task instructions.
+Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading the main story in the A-maze task.
+After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, saw an debriefing and were given a code to enter into Mturk for payment.
+The experiment was implemented in Ibex (CITE) and the experimental code is availabe at REPO.
+
+## Data analysis
+Analyses were done in R (CITE), using the following CURATE packages. We used `r cite_r("r-references.bib")` for all our analyses.
 
@@
 
-## Reading stories in the Maze task
+## Reading stories in the Maze task (good)
 
@@
 
-```{r errors, out.height="25%", fig.pos='t', fig.cap="(ref:error-cap)"}
+```{r errors, out.height="25%", fig.width=6, fig.height=2, fig.pos='t', fig.cap="(ref:error-cap)"}
 error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
@@
 
-Our first concern was whether the paradigm would work at all. To our relief, it did. Some participants were able to complete the Maze task with high accuracy and understand the story while doing so.
+Our first concern was whether this error-correction paradigm on long passages would work.
+It did; some participants were able to complete the Maze task with high accuracy and understand the story while doing so.
 
-We first wanted to check for participant accuracy, which reflects both how well participants can focus on and navigate the task and whether the auto-generated distractors are of acceptable quality.
-We calculated the per-word error rate for each participant and compared it against their average reaction time.
-(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.) As seen in Figure \@ref(fig:errors), there is a cluster of participants who make relatively few errors (marked in green), with some reaching 99% accuracy.
+Participant accuracy reflects how well participants can navigate the task and whether the auto-generated distractors are of acceptable quality.
+We calculated the per-word error rate for each participant and graphed it against their average reaction time.
+(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.) As seen in Figure \@ref(fig:errors)A, there is a cluster of participants who make relatively few errors (marked in green), with some reaching 99% accuracy.
 This indicates that the distractors were generally appropriate and that some participants can maintain focus on the task for the whole story.
 These careful participants tend to take around 1 second for each word selection, which is much slower than other paradigms CITE SPR something.
-
-In contrast, we also have a cluster of participants (in red) who we infer sped through the task clicking randomly.
+We also have a cluster of participants (in red) who we infer sped through the task clicking randomly.
 This distribution is likely due to the mix of workers on Mturk; as we did not use qualification cutoffs.
 
-Another important check is whether participants were able to comprehend the story,which we tested using 6 binary-choice comprehension questions after each story.
-We count how many questions each participant got right, and group them by their performance on the Maze task.
-As seen in @ref(fig:comp-q), most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions.
-Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet.These are not especially hard questions, and we can't rule out that some participants may have been able to guess the answers without reading the story.
+Another check is whether participants comprehended the story.
+We counted how many of the binary-choice comprehension questions each participant got right (out of 6), and group them by their performance on the Maze task.
+As seen in @ref(fig:errors)B, most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions.
+Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet.
+These are not especially hard questions, and we can't rule out that some participants may have been able to guess the answers without reading the story.
 Nonetheless, this provides preliminary evidence that people can understand and remember details of stories they read during the Maze task.
 
-Given that Maze task accuracy and comprehension question accuracy are correlated, but task accuracy is finer-grained and less noisy we use task performance as our exclusion metric.
+Given that Maze task accuracy and comprehension question accuracy are correlated, but task accuracy is finer-grained, we use task performance as our exclusion metric.
 This easy to calculate speed and accuracy comparision makes it easy to see who was paying attention to the task; we exclude the random guessers and only analyse data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
 
-## Error patterns
-When participants do make errors, they correct them quickly, generally within one second, which means it's probably not disrupting their reading much.
-
-While we are planning on analysing data after mistakes, it might be interesting to know how much of the data is pre-mistake, and how many of the sentences are completed error-free.
-If we only exclude the errors and the first words of sentences, there are `r nrow(data_non_error)` words.
-If we also exclude all words after mistakes, there are `r nrow(data_before)` words.
-We will exclude only a couple of words after each mistake, but we could also opt to analyse only the pre-error sections.
-Of the `r nrow(data_sentence)` sentences that good participants completed, `r filter(data_sentence, sent.correct==1) %>% nrow()` were completed entirely correctly.
-
-Our first points of interest were how well participants would do on the Maze task and on the comprehension questions.
-Figure blah shows the distributions and relationship between participants accuracy on the Maze task (i.e.
-fraction of the time they selected the correct continuation) and the time to took for them to make selections (mean RT; RT's above ZZ were assumed to be indicative of a break and excluded).
-Participant dots are color coded based on whether they were at least 80% accurate; this is the threshold we use for including data in future analysis.
-A number of participants did quite well at the task (accuracies reached (double check!) 99%).
-This shows that it's possible for participants to focus on the task for the whole time.
-(While this is not overall much longer than previous studies such as Witzel 2012 and Boyce et al 2020, it had to be continuous, and so we couldn't use a blocking structure to create pauses.) It also shows that A-maze distractors are high-enough quality that participants are able to get them right most of the time.
-Given the clump of participants at around 50% accuracy, which is consistent with randomly clicking, we assume they are not taking the task seriously.
-In terms of RT, we see that attentive participants take on average %FOOBAR seconds, while participants who are clicking randomly can do so much faster.
-
-In terms of comprehension questions, we see relatively good performance by participants who were attentive to the task.
-These answers were also provided quickly, which mostly rules out the (define quickly) possibility that they were googling for the answers.
-There's a chance that some knew the answers already, so this isn't super conclusive.
-Because the stories take a long time to read (~15 minutes+), participants are having to remember the story/answers to questions for quite a while.
-This performance is less goood than that for SPR; but it's hard to know comparing between methods with different task demands and subject pools.
-Participants who didn't do well also didn't do well here.
-
-## Quantitative stuff
-
-There's commonly observed effects of surprisal, frequency, and length on stuff, so we looked at that. TODO: how we got estimates of surprisal, frequency, and length. TODO: word exclusions
-
-- GAMs
-- LMER/BRMS
-
-## LMER stuff
+## RT and surprisal
 
@@
 
-`r data %>% select(subject) %>% unique() %>% nrow()` participants read naturalistic stories from the natural stories corpus. Each participant read 1 story.
+The relationship between a word's predictability and it's reading time is of interest.
+CITE It's well-established that for eye-tracking and SPR, RTs are roughly linear in term of a words surprisal (negative log probability).
+We wanted to check if this linear relationship also holds for Maze.
+Two other common word-features that affect RT are word length and a word's overall frequency.
+CITE
 
-We exclude
-
-- participants who do not report English as a native language (`r data_filt %>% select(subject) %>% unique() %>% nrow()` remaining)
-- participants who do not get 80% of the words correct (`r data_low_error %>% select(subject) %>% unique() %>% nrow()` remaining)
-- practice items (`r data_low_error %>% nrow()` words remaining)
-- words that were wrong or were within two after a mistake (`r data_error_free %>% nrow()` words remaining)
-- the first word of every sentence (didn't have a real distractor, RT is measured slightly differently) (`r data_no_first %>% nrow()` words remaining)
-- words with RTs <100 or >5000 (<100 we think is likely a recording error, or at least not reading the words at all, >5000 is likely getting distracted) (`r data_ready %>% nrow()` words remaining)
-
-Within the filtered data, each story was read between `r min(data_stories$n)` and `r max(data_stories$n)` times, for an average of `r mean(data_stories$n)`.
-
-We also do the analyses on only the words before mistakes (per sentence) (`r data_pre_error %>% nrow()` words)
-
-From the modelling side:
-(After attempts without doing this filtering) we only include words which are single token and known words in each of the models vocabularies.
-We also only include words with frequencies.
-This is roughly equivalent to excluding words with punctuation.
-
-We use as predictors:
-
-- length in characters of stripped word
-- unigram frequency of word.
-Frequencies for words are calculated using word_tokenize on the gulordava train data and counting up instances.
-(This tends to tokenize off punctuation, but is capitalization sensitive).
-Frequencies are represented as log2 of the expected occurances in 1 billion words.
-
-Surprisals are measured in bits.
-
-- ngram (5-gram KN smoothed)
-- GRNN
-- Transformer-XL
-
-We center the predictors for the brms models, but don't rescale them.
-
+To assess this, we create a set of predictors of frequency, word length, and surprisals from several models.
+For length, we use the length of characters (excluding end punctuation).
+For a measure of unigram frequency, we tokenize the CITE gulordava training data (this tokenizes off punctuation, but preserves case) and tally up instances.
+For the models, we use the log2 frequency of the expected occurances in 1 billion words (thus frequency is log, but higher values indicate higher frequency words).
+For surprisals, we get per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN CITE, and Transformer-XL CITE.
+We center, but do not rescale these predictors.
+For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
+CITE some SPR stuff This is an additional test of localization.
 The model formula we use is rt ~ surp* length +  freq * length + past_surp * past_length + past_freq * past_length.
 
-# Frequentist LMs
+We only include words that were in the vocabularies of all three models as a single token and for which we have frequency information.
+This has the effect of excluding words that had punctuation as well as some uncommon or proper nouns.
+We also exclude the first word of every sentence (which had a dummy distractor).
+We do not analyse RTs for words where the RT was <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
+We also omit words where the participant made a mistake and the two words after a mistake.
+
+### GAM
+For GAM models, we center length and frequency but not surprisal.
+We want surprisal interpretable, but we also will be plotting it (at least for the bootstrapping) at length and frequencies set to 0, so they need to be centered.
+(Not sure this last piece is actually true/matters).
+
+### GAMs
+
+## Try 1
+Using only pre-error data.
+
+Does not have hierarchical effects. Has smooths for the surprisals, and tensor effects/interactions for the freq x length
 
@@
 ```
+
+```{r for gam}
+
+ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            ) %>% mutate(model="5-gram")
+
+grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            )%>% mutate(model="GRNN")
+
+txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_surp, prev_surp=past_txl, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            ) %>% mutate(model="TXL")
+
+all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% 
+  select(surprisal, prev_surp,model) %>% 
+  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
+  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
+```
+
+```{r gams, eval=F}
+gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")
+
+gam_grnn_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_data, method="REML")
+
+gam_txl_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_data, method="REML")
+
+```
+
+```{r plot-gam, eval=F}
+
+a <- get_gam_predictions(model=gam_ngram_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
+b <- get_gam_predictions(model=gam_ngram_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
+c <- get_gam_predictions(model=gam_grnn_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
+d <- get_gam_predictions(model=gam_grnn_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
+e <- get_gam_predictions(model=gam_txl_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
+f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
+```
+
+```{r, eval=F}
+all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f)
+
+gam1 <-  ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+  geom_line()+
+  geom_ribbon(alpha=.3)+
+   #geom_density(data=all_data, aes(x=value), fill="gray",)+
+  facet_grid(s~model)+
+  coord_cartesian(ylim=c(700,1300), xlim=c(0,28))+
+  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
+
+dens2 <-   ggplot(all_data, aes(x=value))+
+  geom_density(fill="gray",)+
+  facet_grid(.~model)+
+  labs(x="Surprisal (bits)", y="")+
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+bot <- plot_grid(NA, dens2, NA, nrow=1, rel_widths = c(.16, 1, .08))
+
+ plot_grid(gam1, bot, nrow=2, rel_heights = c(1, .3))
+
+#confirm that this really is just a prettied up version of what the usual output is
+#plot(gam_txl_1, seWithMean = TRUE, shift = coef(gam_txl_1)[1], select=1)
+
+```
+
+### Frequentist LMs
 
@@
 
+## Error patterns
+When participants do make errors, they correct them quickly, generally within one second, which means it's probably not disrupting their reading much.
+
+While we are planning on analysing data after mistakes, it might be interesting to know how much of the data is pre-mistake, and how many of the sentences are completed error-free.
+If we only exclude the errors and the first words of sentences, there are `r nrow(data_non_error)` words.
+If we also exclude all words after mistakes, there are `r nrow(data_before)` words.
+We will exclude only a couple of words after each mistake, but we could also opt to analyse only the pre-error sections.
+Of the `r nrow(data_sentence)` sentences that good participants completed, `r filter(data_sentence, sent.correct==1) %>% nrow()` were completed entirely correctly.
 # Discussion
@@
 Might be extra good to do with pauses for mistakes to fix the incentives. 
+This simple tweak in principle solves several problems.
+Firstly, it allows for running longer materials, potentially expanding the range of questions Maze is suitable for.
+Now, when participants make the occassional mistake or misclick, they can continue with the story, seeing every word of it.
+Thus, they provide reading times in later sentences and see all the context.
+Seeing the error message and resolving it is quick and does not disrupt the (already slow)  reading process too much.
+
+Secondly, this error-correction compensates for some of the shortcomings of A-Maze identified in @boyceMazeMadeEasy2020; distractors might still cause participants to make unavoidable mistakes, but at least the still see the rest of the sentence and we get their data.
+We believe (although don't have evidence) that being able to correct mistakes may reduce frustration related to unavoidable mistakes from impossible choices and accidental misclicks.
+Whether this also solves the data loss problem from the researcher perspective within a sentence depends on whether post-mistake data are high-quality and trustworthy; this is a hard-to-assess question of potential interest.
+
+Even if researchers exclude all post-mistake data from analysis, this process can still address the question of whether errors are due to inattentive participants or bad distractors (as discussed in @boyceMazeMadeEasy2020, section XX).
+When we have a participants selection for every word, we can easy calculate a per-word error rate for each participant, without having to address the censoring present in error-terminating Maze (where we can't know if a participant would have made more mistakes after the first; with error correction, we know how may more they did make).
+If participants are guessing, we'll see high per-word error rates, if the early distractors are bad, we'll see low per-word error rates (with errors concentrated at the start of the sentences).
+This per-word error rate metric measures participants task accuracy, and allows us to exclude data from inattentive participants.
+This provides a convenient and clear-cut way of controlling data quality after the fact.
+
+It also starts to reduce the perverse incentives.
+With error-terminates Maze, the fastest way to get through the task is to select randomly, and it's quite quick because mistakes skip you to the next sentence.
+With error-correction Maze, randomly jamming buttons takes more effort, but is still an effective strategy.
+[Footnote: In discussing this work, we recieved the suggestion that one way to disincentivize random clicking is to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms or 1 sec) before being able to correct their mistake.
+This seems like a promising improvement that could be worth implementing and testing.]
 \newpage

### 2021-02-04 Papers/Paper/manuscript.Rmd
edits to paper

@@
  library("citr")
-library(patchwork)
  library(here)
  library(tidyverse)
-library(readr)
+library(patchwork)
  library(brms)
  library(lme4)
-library(rstan)
  library(tidybayes)
-library(knitr)
-library(matrixStats)
-library(tidyverse)
-library(readr)
-library(brms)
-library(lme4)
-library(rstan)
-library(tidybayes)
-library(knitr)
  library(mgcv)
-library(mgcViz)
-library(tidymv)
-library(rsample)
  library(cowplot)
-library(scales)
-r_refs("r-references.bib")
+ library(gridExtra)
+r_refs("r-references.bib", append=F)
 theme_set(theme_bw())
-rstan_options(auto_write = TRUE)
-options(mc.cores = parallel::detectCores())
 options(knitr.table.format = "pdf")
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
 ```
 # Intro
-[TODO: hook about toolbox of methods]
-A diversity of robust methods provides for converging evidence from different methods (either within or between types), and choices allow for a good fit to the scope and goals of research.
-But we also want these methods to be reliable.
-Some of that is theoretical, but we also want empirical evidence that what the method captures is relevant to what we want to study, and that this connection is true across as broad a range of materials as we hope to use it for.
-Robustness is important to establishing how good a new method is.
-What range of tasks can this new tool accomplish and how well?
-(Could really extend on a toolbox analogy.)
-
-[incrementality; so need incremental methods] 
-One question in psycholinguistics is how real-time comprehension works; as we read or listen, we process language incrementally, building and changing our mental model of a sentence.
-How long we spend on each word is due to both qualities of the word itself and how it fits with the sentence context so far TODO citations
-
 We process language incrementally, building a model of the sentence as we hear or read word.
 When a word is unexpected and does not fit the model of the sentence so far, our understanding of the sentence may need to be revised.
-Incremental processing methods measure this process, tracking how long each word takes to read as a proxy for how hard it is to process and incoroporate into the mental conception of meaning.
-Fluent language uses process language rapidly and flexibly adapt to many unexpected words.
-This makes it harder to capture reading time effects because they may be obscured by the overall robust reading process.
-Different techniques handle this localization problem differently, and present different trade-offs.
-The two most commonly used methods are eye-tracking and self-paced reading, but neither of these techniques are ideal at localization.
-One measure of localization is whether current or past word reading time is more predictive.
-SPR and eye-tarcking both don't seem great.
+Incremental processing methods measure how long each word takes to read as a proxy for how hard it is to incoroporate into the mental model of the sentence.
+Fluent language uses process language rapidly and flexibly adapt to many unexpected words, which makes it hard to pinpoint slow downs related to difficult-to-process words.
 
-[eye-track and SPR, failure to localize] TODO: possibly shorten?
-In eye-tracking, participants read text freely from a screen, while the eye movements and fixations are tracked with an infrared camera CITE, check for truth.
-Eye-tracking is naturalistic and yields fine-grained data about how long participants spent on each word when, but it's expensive to run requiring both specialized equipment and researcher time to run experiments.
-Self-paced reading (commonly abbreviated SPR) in contrast can be run online with no specialzed equipment or direct researcher supervision, and so it's possible to collect data quickly and cheaply.
-In SPR, participants see one word of the text on screen at a time and press a button to contine to the next word instead; the time between button presses when a word was on screen is used as the reading time.
-The problem with SPR, especially when run online, is that effects from one word tend to show up as longer reading times spread out over several later words, in what is known as spillover effects.
-Both of these methods suffer somewhat from analytic flexibilty as processing difficulty may manifest in several ways in eye-tracking and is hard to localize in SPR.
+TODO motivate why we care about localization
 
-[Maze]An alternative that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing.
-In the Maze task, participants see two words at a time, one of which continues the sentence, and one of which does not, and they have to choose which word is the correct continuation.
-The time to choose the correct word is taken as the dependent measure and interpreted as representing the difficulty of integrating the word into the sentence so far.
+The two most commonly used methods are eye-tracking and self-paced reading.
+Both these methods are known to have spillover effects, which makes localization harder.
+CITE some things TODO might be worth bringing up RT/surprisal/spillover findings hereOur next question is what the relationship between a word's predictability and it's reading time is for Maze.
+It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability).
+Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization.
+In addition to suprisal predicting RT, word length and word's overall frequency are also often found to be predictive.
+TODO many CITES
+
+An alternative method that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing [@forsterMazeTaskMeasuring2009].
+In the Maze task, participants see two words at a time, a correct word that continues the sentence, and a distractor which does not.
+Participants must choose the correct word, and their reaction time (RT) is the dependent measure.
 If participants make a mistake, the sentence discontinues.
-Theoretically, participants must fully integrate each word into the sentence in order to confidently select it and move on.
-This is suppported by studies finding localized effects without much spillover.
+Theoretically, participants must fully integrate each word into the sentence in order to confidently select it.
+This idea is supported by studies finding strongly localized effects [@witzel12].
 
-[A-Maze] However, traditional Maze materials are effort-instensive to construct, because of the need to select infelicitious words as distractors for each spot of each sentence; this may have kept this method from widespread adoption.
+The downside of Maze is that materials are effort-intensive to construct because of the need to select infelicitious words as distractors for each spot of each sentence; this may explain why the Maze task was not widely adopted.
+@boyceMazeMadeEasy2020 demonstrate a way to automatically generate Maze distractors by using NLP language models to find words that are high-surprisal in the context of the target sentence.
+The quality of these A-Maze distractors is not up to that of hand-generated distractors, but @boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzel12.
+@sloggettAmazeAnyOther also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.
 
-CITE US propose a solution to this problem of generating Maze distractors.
-In A-maze, distractors are automatically generated by feeding the correct sentences through a program that uses NLP language models to select high-surprisal distractors.
-The quality of the automatically generated distractors is not up to that of hand-generated distractors, in that the distractors ocassionally are good fits in context.
-Nonetheless, Boyce et al found that materials with A-maze distractors had similar results to G-maze distractors.
-TODO: talk about the paradigm more
-[Summary of 2020 paper]
+While new, A-maze seems like a potentially powerful addition to the psycholinguists toolkit.
+However, like the other Maze tasks it has been limited in its application to single-sentence items probing minimal comparisons in constructed sentences.
+While some psycholinguistic phenomena are well-explored within individual targeted sentences, there's also a need for reading time data on longer passages, for instance in order to compare to language models or study discourse effects.
+Therefore, it's important to test whether the Maze task works on these types of materials.
 
-[Summary other uses of A-maze - Slogett]
-Slogett also found that it works similarly.
-
-While new and untested, A-maze seems like a potentially powerful addition to the toolkit.
-However, like the other Maze tasks it has been limited in its application to single-sentence items probing minimal comparisons in constructed sentences.
-It's only been used on minimal pair types of sentences where each sentence is independent.
-In particular, it hasn't been used on longer, naturalistic texts, so we don't know how in compares to other methods here.
-And hasn't been used on multi-sentence issue of generating Maze materials has been proposed.
-An alternative version of the Maze task, dubbed A-maze, was recently introduced that gets around this problem by automatically generating distrators using a program and neural nets (Boyce et al, 2020).
-While the quality of the automatically generated distractors is not quite up that of hand-generated (as judged by their being occassional distractors that seem completely acceptable continuations), the results from A-maze match those of G-maze.
-Boyce et al provide an interface for running the Maze task in a browser and a way of generating distractors that are usually, although not always, sufficiently bad.
-
-In a comparison between these A-maze materials and traditional hand-crafted distractors (from Witzel 2012), they find that A-maze yields comparable results to G-maze (either lab or web), and in one condition recovers effects that G-maze did not.
-A-maze also outperforms self-paced reading on this syntactic disambiguation paradigm.
-Sloggett (cite) also directly compared A-maze and G-maze on a disambiguaton paradigm and similarly found that A-maze and G-maze yielded similar results.
-
-These demonstrations make A-maze seem quite promising, as it has the advantages of the traditional Maze task, without it's drawback. This method seems worth exploring further.
-
-[why do longer Maze] While some psycholinguistic phenomena are well-explored within individual targeted sentences, there's also a need for reading time data of more naturalistic passages.
-This is important as a comparison for models of language understanding, for correlating with NLP language models.
-In addition, some phenomena such as discourse effects, only work in the context of passages.
-THus, it would be useful/versatile if Maze could be used in these other contexts.
-
-[short-coming, where's the question we need to answer] One evident shortcoming of the Maze task is that it has been used on only a limited set of materials.
-Some of this is due to the fact that it isn't used as widely as other incremental processing methods, and some is due to inherent limitations of the paradigm.
+While the issue of needing to generate distractors for long passages is solved with A-maze, another problem with Maze remains.
 In particular, because Maze tasks discontinue after participants make mistakes, the farther into an item a word is, the fewer participants see it.
@@
 
-[Our contribution is...] We find a way around this problem by creating a version of the Maze task interface where participants correct their mistakes and continue with the sentence rather than being discontinued.
+We address this problem by creating a Maze task interface where participants correct their mistakes and continue with the sentence rather than terminating on mistakes.
 This allows for multi-sentence items to be run using Maze, which we verify by running Maze on the Natural Stories corpus.
@@
 # Error-correction maze
+(ref:diagram-cap) Schematic of error-correction Maze.
+A participant reads a sentence word by word, choosing the correct word at each time point (selections marked in blue).
+When they make a mistake, an error message is displayed, they try again, and continue with the sentence.
 
-[review the terminate thing, and why it was there] The advantage of the Maze task is that it forces incremental processing and automatically excludes inattentive participants by terminating a sentence when a participant makes an error.
-Thus, only participants who have processed the sentence up that point and are paying attention contribute data at critical regions later in the sentence.
+```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:diagram-cap)"}
+knitr::include_graphics(paste0(here(),"/Papers/maze_diagram_cropped.pdf"))
+```
+One advantage of the Maze task is that it forces incremental processing and automatically excludes inattentive participants by terminating a sentence when a participant makes an error.
+Thus, only participants who have processed the sentence and are paying attention contribute data at critical regions later in the sentence.
 However, this means we don't have data after a participant makes a mistake in an item.
-In traditional G-maze tasks, with hand-crafted distractors and attentive participants, this is a small issue as the low per-word error rate does not compound too badly over the individual sentences.
+In traditional G-maze tasks, with hand-crafted distractors and attentive participants, this is a small issue.
+However, this data loss is much worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020].
+The high errors are likely from some combination of participants guessing randomly and from auto-generated distractors that in fact fit the sentence; as @boyceMazeMadeEasy2020 noted, some distractors, especially early in the sentence, were problematic and caused considerable data loss.
 
-[ how this causes problems for 2020 and theoretically] 
-However, this data loss was worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020].
-The high errors could be from participants guessing randomly or from auto-generated distractors that in fact fit the sentence.
-@boyceMazeMadeEasy2020 noted that some distractors, especially early in the sentence were problematic and caused considerable data loss.
-
-Given the appeals of using A-maze, it would be adventageous to get around this problem.
-Additionally, even with attentive participants and good distractors, the Maze paradigm will eventually run into problems as error rates compound over long materials, and few participants will make it to the end.
-For instance, with a 1% error rate, `r round(.99**15*100)`% would complete each 15-word sentence, but only `r round(.99**50*100)`% of participants would complete each 50 word vignette, and `r round(.99**200*100)`% each 200 word passages.
+We could try to improve the situation by auto-generating better distractors or hand-replacing problematic ones.
+However, the inherent data losses poses a more fundamental problem: even with attentive participants and good distractors, the Maze paradigm will eventually run into problems as error rates compound over long materials where few participants will make it to the end.
+For instance, with a 1% error rate, `r round(.99**15*100)`% would complete each 15-word sentence, but only `r round(.99**50*100)`% of participants would complete a 50 word vignette, and `r round(.99**200*100)`% a 200 word passages.
 In order to run longer materials, we need something to do when participants make a mistake, other than terminate the entire item.
@@
 To resolve this, we introduce an *error-correction* variant of Maze, where we let participants continue on when they make a mistake.
-To do this, we present them with an error message, and wait until they select the correct option, before continuing the sentence as normal.
-We make this "error-correction" Maze available as an option in a modification of the Ibex Maze implementation introduced in CITE ME (TODO repo link).
+We present them with an error message and wait until they select the correct option, before continuing the sentence as normal.
+This *error-correction* variant is shown in Figure \@ref(fig:diagram).
+We make this "error-correction" Maze available as an option in a modification of the Ibex Maze implementation introduced in @boyceMazeMadeEasy2020 (TODO repo link).
 The code records both the RT to the first click and also the total RT until the correct answer is selected as separate values.
-TODO diagram goes here
 
-[ why this error thing seems really cool and could work] 
-Our goal is that this error-correction will minimize the problems of having occassional poor auto-generated distractors by letting participants continue on and still read the whole item.
+Our goal is that this error-correction will minimize the problems of having occasional poor auto-generated distractors by letting participants continue on and still read the whole item.
 It will also enable us to run long materials on Maze, thus expanding the range of questions that Maze can be used on.
@@
 
-We constructed A-maze distractors for the Natural stories corpus (CITE) and recruited 100 MTurkers to each read a story in the Maze paradigm.
+We constructed A-maze distractors for the Natural stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 MTurkers to each read a story in the Maze paradigm.
 
-## Materials (good)
-We took the texts of the Natural Stories corpus (CITE) and split it into sentences because the A-Maze process works on the sentence level.
-(This involved some judgment calls due to quoted speech).
+## Materials
+We took the texts of the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and split it into sentences because the A-Maze process works on the sentence level.
 We used the original comprehension questions provided in the Natural Stories corpus.
@@
 The underlying method did not change, but features such as appropriately handling more types of punctuation (needed for the Natural Stories corpus) were added.
-We ran the materials using FOOBAR paramters.
+We ran the materials using FOOBAR parameters.
 We used the auto-generated distractors as they were, without checking them for quality.
@@
 We recruited 100 participants from Amazon Mechanical Turk, and paid each participant 3.50 dollars.
-We estimated that reading a 1000 word story would take 17 minutes at 1 word/second (based on previous work), so we expected the task to take around 20 minutes total.
-In fact, it appears to have taken most attentive participants around 20 minutes to read the story, and thus closer to 25 minutes for the overall study.
 We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
@@
 
-## Reading stories in the Maze task (good)
+## Reading stories in the Maze task
 
@@
 B.
-Performance on comprehension questions.
+Performance on the comprehension questions.
 Participants who had >80% task accuracy tended to do well on comprehension questions; those who were at chance on the task were also at chance on comprehension questions.
 
-```{r errors, out.height="25%", fig.width=6, fig.height=2, fig.pos='t', fig.cap="(ref:error-cap)"}
+```{r errors, out.height="25%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:error-cap)"}
 error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
@@
 We counted how many of the binary-choice comprehension questions each participant got right (out of 6), and group them by their performance on the Maze task.
-As seen in @ref(fig:errors)B, most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions.
+As seen in Figure \@ref(fig:errors)B, most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions.
 Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet.
@@
 
-The relationship between a word's predictability and it's reading time is of interest.
-CITE It's well-established that for eye-tracking and SPR, RTs are roughly linear in term of a words surprisal (negative log probability).
-We wanted to check if this linear relationship also holds for Maze.
-Two other common word-features that affect RT are word length and a word's overall frequency.
-CITE
+### GAMs
 
-To assess this, we create a set of predictors of frequency, word length, and surprisals from several models.
+We wanted to see if Maze also showed a linear relationship between surprisal and RT, and whether this was localized at the current word or whether previous word surprisals were also influential.
+To do this, we created a set of predictors of frequency, word length, and surprisals from several different langauge models, and fit generalized additive models to check for a linear relationship.
  For length, we use the length of characters (excluding end punctuation).
-For a measure of unigram frequency, we tokenize the CITE gulordava training data (this tokenizes off punctuation, but preserves case) and tally up instances.
+For a measure of unigram frequency, we tokenize the training data from @gulordava18 (this tokenizes off punctuation, but preserves case) and tally up instances.
 For the models, we use the log2 frequency of the expected occurances in 1 billion words (thus frequency is log, but higher values indicate higher frequency words).
-For surprisals, we get per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN CITE, and Transformer-XL CITE.
-We center, but do not rescale these predictors.
+For surprisals, we get per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava], and Transformer-XL [@daiTransformerXLAttentiveLanguage2019a].
+We centered but did not rescale the length and frequency predictors; for interpretability, we kept surprisal uncentered.
 For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
-CITE some SPR stuff This is an additional test of localization.
-The model formula we use is rt ~ surp* length +  freq * length + past_surp * past_length + past_freq * past_length.
+CITE some papers that did the same thing .
+We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions.
 
@@
 
-### GAM
-For GAM models, we center length and frequency but not surprisal.
-We want surprisal interpretable, but we also will be plotting it (at least for the bootstrapping) at length and frequencies set to 0, so they need to be centered.
-(Not sure this last piece is actually true/matters).
+(ref:gam-cap) GAM predictions of reaction time (RT) as a function of either current word surprisal (top) or previous word surprisal (bottom).
+Density of data is shown along the x-axis.
+For each of the 3 language models used, there is a linear relationship between current word surprisal and RT (at least when there is enough data).
+There is no relationship between previous word surprisal and RT.
 
-### GAMs
-
-## Try 1
-Using only pre-error data.
-
-Does not have hierarchical effects. Has smooths for the surprisals, and tensor effects/interactions for the freq x length
-
-```{r}
-#mean by subject for each word
+```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:gam-cap)"}
 labelled_pre_error <- read_rds(paste0(here(),"/Analysis/pre_error.rds"))
-d_lm <- labelled_pre_error %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
-                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
-  summarize(mean_rt=mean(rt))
-```
-
-```{r for gam}
 
@@
   mutate(s=ifelse(name=="surprisal", "Current","Previous"))
-```
 
-```{r gams, eval=F}
-gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")
-
-gam_grnn_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_data, method="REML")
-
-gam_txl_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_data, method="REML")
-
-```
-
-```{r plot-gam, eval=F}
-
-a <- get_gam_predictions(model=gam_ngram_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
-b <- get_gam_predictions(model=gam_ngram_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
-c <- get_gam_predictions(model=gam_grnn_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
-d <- get_gam_predictions(model=gam_grnn_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
-e <- get_gam_predictions(model=gam_txl_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
-f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
-```
-
-```{r, eval=F}
-all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f)
-
+all <-read_rds(paste0(here(),"/Analysis/gam_predictions.rds"))
 gam1 <-  ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
@@
   labs(x="Surprisal (bits)", y="")+
-  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+  theme(strip.text.x=element_blank(),axis.text.y=element_blank(), panel.grid=element_blank(), axis.ticks.y=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
 
-bot <- plot_grid(NA, dens2, NA, nrow=1, rel_widths = c(.16, 1, .08))
-
- plot_grid(gam1, bot, nrow=2, rel_heights = c(1, .3))
-
-#confirm that this really is just a prettied up version of what the usual output is
-#plot(gam_txl_1, seWithMean = TRUE, shift = coef(gam_txl_1)[1], select=1)
+p2 = align_plots(gam1, dens2, align = "v", axis="lr")
+plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))
 
@@
 
-### Frequentist LMs
+The smooths for the current and previous words surprisals are shown in Figure \@ref(fig:gam).
+Note that for each of the models, high-surprisal words at relatively rare, with much of the data for words between 0 and 15 bits of surprisal.
+All of the models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data.
+All of the models show a flat relationship between previous word surprisal and RT.
+This is a sign of localization as the previous word's surprisal is not affecting RT, only the word's own suprisal is.
+The linear relationship matches that found with other methodologies.
 
-## Models
+Given the linear relationship, we next fit linear models to look at coefficients and model fit. We fit BRMS models.
+
+### BRM models
+
+We include a by-subject effect for everything, and a by_word random intercept (full mixed effects).
+
+Priors:
+
+- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
+- normal(0,500) for beta and sd -- we don't really know what effects are
+- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way
+
+```{r}
+
+show_summary <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=1),
+           upper=round(.upper, digits=1),
+           E=round(.value, digits=1),
+           `Estimate`=str_c(E," [",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           ) %>% 
+    select(Term, `Estimate`)
+
+  stats
+}
+```
+
+```{r pre-error}
+brm_txl_interact <- read_rds(paste0(here(),"/Analysis/brm_txl_interact.rds"))
+brm_grnn_interact <- read_rds(paste0(here(),"/Analysis/brm_grnn_interact.rds"))
+brm_ngram_interact <- read_rds(paste0(here(),"/Analysis/brm_ngram_interact.rds"))
+a <- show_summary(brm_txl_interact) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
+b <- show_summary(brm_grnn_interact) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
+c <- show_summary(brm_ngram_interact) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
+
+summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
+  mutate(Term=factor(Term, 
+                        levels=c("Intercept", 
+                                 "surp_center",
+                                 "length_center",
+                                 "freq_center",
+                                 "surp_center:length_center",
+                                 "length_center:freq_center",
+                                 "past_c_surp",
+                                 "past_c_length",
+                                 "past_c_freq",
+                                 "past_c_surp:past_c_length",
+                                 "past_c_length:past_c_freq")
+                                 )) %>% 
+  arrange(Term) %>% 
+  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency",
+                  "Surp x Length", "Freq x Length", "Past Surprisal",
+                  "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length")) %>% 
+  select(Term, `5-gram`, `GRNN`, TXL)
+
+knitr::kable(summ, format="latex", position="t",caption="TODO")
+```
+
+```{r post-error}
+brm_txl_interact_post <- read_rds(paste0(here(),"/Analysis/brm_txl_interact_post.rds"))
+brm_grnn_interact_post <- read_rds(paste0(here(),"/Analysis/brm_grnn_interact_post.rds"))
+brm_ngram_interact_post <- read_rds(paste0(here(),"/Analysis/brm_ngram_interact_post.rds"))
+
+d <- show_summary(brm_txl_interact_post) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
+e <- show_summary(brm_grnn_interact_post) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
+f <- show_summary(brm_ngram_interact_post) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
+
+summ <- d %>% union(e) %>% union(f) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
+   mutate(Term=factor(Term, 
+                        levels=c("Intercept", 
+                                 "surp_center",
+                                 "length_center",
+                                 "freq_center",
+                                 "surp_center:length_center",
+                                 "length_center:freq_center",
+                                 "past_c_surp",
+                                 "past_c_length",
+                                 "past_c_freq",
+                                 "past_c_surp:past_c_length",
+                                 "past_c_length:past_c_freq")
+                                 )) %>% 
+  arrange(Term) %>% 
+  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency",
+                  "Surp x Length", "Freq x Length", "Past Surprisal",
+                  "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length")) %>% 
+  select(Term, `5-gram`, `GRNN`, TXL)
+
+knitr::kable(summ, format="latex", position="t",caption="TODO")
+
+```
+
+As we can see in Table \@ref(tab:pre-error) and Table \@ref(tab:post-error).
+
+### Model comparison
 
@@
 ```{r}
+d_lm <- labelled_pre_error %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
+                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
+  summarize(mean_rt=mean(rt))
 
@@
 
-## What if we don't exclude participants
-
-One potentially useful comparison with past SPR/eye-tracking is if we are more inclusive of participants, do we get data that looks more like theirs. (Perhaps our exclusion criteria are more robust?)
-
-This is all data from participants who speak English that wasn't a mistake. (So, includes bad participants and post-mistake data.)
-
-```{r freq_all}
-#mean by subject for each word
-labelled_anything_goes <- read_rds(paste0(here(),"/Analysis/non-selective.rds"))
-d_lm2 <- labelled_anything_goes %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
-                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
-  summarize(mean_rt=mean(rt))
-```
-
-```{r}
-
-no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
-
-ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
-
-grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
-
-txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm2)
-
-#summary(no_surp)
-#summary(ngram_only)
-#summary(grnn_only)
-#summary(txl_only)
-
-```
-
-BRM models from this same very permissive data. 
-```{r brm-models}
-
-any_grnn <- read_rds(paste0(here(),"/Analysis/brm_grnn_any_center.rds"))
-any_txl <- read_rds(paste0(here(),"/Analysis/brm_txl_any_center.rds"))
-any_ngram <- read_rds(paste0(here(),"/Analysis/brm_ngram_any_center.rds"))
-
-#summary(any_grnn)
-#summary(any_txl)
-#summary(any_ngram)
-
-```
-
-```{r}
-
-show_summary <- function(model){
-  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
-
-  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
-    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
-    group_by(.variable) %>% 
-    summarize(pct_above_0=mean(above_0)) %>% 
-    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
-    left_join(intervals, by=".variable") %>% 
-    mutate(lower=round(.lower, digits=1),
-           upper=round(.upper, digits=1),
-           E=round(.value, digits=1),
-           `CI`=str_c("[",lower,", ", upper,"]"),
-           Term=str_sub(.variable, 3, -1),
-           ) %>% 
-    select(Term, `E`, `CI`,`P`)
-
-  stats
-}
-a <- show_summary(any_txl) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
-b <- show_summary(any_grnn) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
-c <- show_summary(any_ngram) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
-
-summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`E`,`CI`,`P`)) %>% 
-  select(Term, `E_5-gram`, `CI_5-gram`, `P_5-gram`, `E_GRNN`, CI_GRNN, P_GRNN, E_TXL, CI_TXL, P_TXL)
-
-```
-## BRM models
-
-We include a by-subject effect for everything, and a by_word random intercept (full mixed effects).
-
-Priors:
-
-- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
-- normal(0,500) for beta and sd -- we don't really know what effects are
-- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way
-
-```{r}
-  priors <- c(
-      set_prior("normal(1000, 1000)", class="Intercept"),
-      set_prior("normal(0, 500)", class="b"),
-      set_prior("normal(0, 500)", class="sd"),
-      set_prior("lkj(1)",       class="cor"))
-
-```
-
-### On pre-error data only
-
-```{r}
-
-show_summary <- function(model){
-  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
-
-  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
-    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
-    group_by(.variable) %>% 
-    summarize(pct_above_0=mean(above_0)) %>% 
-    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
-    left_join(intervals, by=".variable") %>% 
-    mutate(lower=round(.lower, digits=1),
-           upper=round(.upper, digits=1),
-           E=round(.value, digits=1),
-           `CI`=str_c("[",lower,", ", upper,"]"),
-           Term=str_sub(.variable, 3, -1),
-           ) %>% 
-    select(Term, `E`, `CI`,`P`)
-
-  stats
-}
-brm_txl_interact <- read_rds(paste0(here(),"/Analysis/brm_txl_interact.rds"))
-brm_grnn_interact <- read_rds(paste0(here(),"/Analysis/brm_grnn_interact.rds"))
-brm_ngram_interact <- read_rds(paste0(here(),"/Analysis/brm_ngram_interact.rds"))
-a <- show_summary(brm_txl_interact) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
-b <- show_summary(brm_grnn_interact) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
-c <- show_summary(brm_ngram_interact) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
-
-summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`E`,`CI`,`P`)) %>% 
-  select(Term, `E_5-gram`, `CI_5-gram`, `P_5-gram`, `E_GRNN`, CI_GRNN, P_GRNN, E_TXL, CI_TXL, P_TXL)
-
-```
-
-### On post-error as well
-
-```{r}
-show_summary_2 <- function(model){
-  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
-
-  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
-    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
-    group_by(.variable) %>% 
-    summarize(pct_above_0=mean(above_0)) %>% 
-    mutate(`P-value equivalent` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
-    left_join(intervals, by=".variable") %>% 
-    mutate(lower=round(.lower, digits=2),
-           upper=round(.upper, digits=2),
-           `Credible Interval`=str_c("[",lower,", ", upper,"]"),
-           Term=str_sub(.variable, 3, -1),
-           Estimate=round(.value, digits=2)) %>% 
-    select(Term, Estimate, `Credible Interval`, `P-value equivalent`)
-
-  stats
-}
-```
-
 ## Comparision with SPR
@@
 Of the `r nrow(data_sentence)` sentences that good participants completed, `r filter(data_sentence, sent.correct==1) %>% nrow()` were completed entirely correctly.
+
 # Discussion
+[summarize results] 
+- error-correction paradigm
+- test on natural stories
+- it works
+
 We've shown another way to adapt the Maze task to make it suitable for a wider range of experiments.
@@
 
-This is just one other data point on A-maze, so we still don't know how widely the method holds up; however the fact that it works on such disparate materials gives us hope that it will be useful for a variety of paradigms.
-
-With new methods, its useful to try to establish how widely they work.
-We hope to test A-maze on an experiment that is very different from the syntactic disambiguation, in order to learn whether it works for a different type of psycholinguistic material.
 While Maze has traditionally been limited to single sentence items; we innovate on how the materials are presented and show that it is possible for participants to read naturalistic passages that are presented in the Maze paradigm.
@@
 
-This is just one other data point on A-maze, so we still don't know how widely the method holds up; however the fact that it works on such disparate materials gives us hope that it will be useful for a variety of paradigms.
-
 The Maze task might be espeically good for comparing with NNs because of the forced incrementality.
@@
 Might be extra good to do with pauses for mistakes to fix the incentives.
-This simple tweak in principle solves several problems.
-Firstly, it allows for running longer materials, potentially expanding the range of questions Maze is suitable for.
-Now, when participants make the occassional mistake or misclick, they can continue with the story, seeing every word of it.
-Thus, they provide reading times in later sentences and see all the context.
-Seeing the error message and resolving it is quick and does not disrupt the (already slow)  reading process too much.
 
 Secondly, this error-correction compensates for some of the shortcomings of A-Maze identified in @boyceMazeMadeEasy2020; distractors might still cause participants to make unavoidable mistakes, but at least the still see the rest of the sentence and we get their data.
-We believe (although don't have evidence) that being able to correct mistakes may reduce frustration related to unavoidable mistakes from impossible choices and accidental misclicks.
 Whether this also solves the data loss problem from the researcher perspective within a sentence depends on whether post-mistake data are high-quality and trustworthy; this is a hard-to-assess question of potential interest.

### 2021-02-05 Papers/Paper/manuscript.Rmd
more edits

@@
 We do not analyse RTs for words where the RT was <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
-We also omit words where the participant made a mistake and the two words after a mistake.
+We also omit words where the participant made a mistake and had made a mistake earlier in the sentence (Although see supplement for results including post-mistake TODO).
 
@@
 
-We include a by-subject effect for everything, and a by_word random intercept (full mixed effects).
-
-Priors:
-
-- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
-- normal(0,500) for beta and sd -- we don't really know what effects are
-- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way
+We built models using BRMS, with surprisal, frequency, length and surprisal x length and frequency x length effects.
+Predictors were as described above, except all were centered.
+We used full mixed effects, including by-subject effects of everything and a per-word random intercept.
+We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations).
+TODO CITE also maybe this is too details-y methods.
 
@@
 
-knitr::kable(summ, format="latex", position="t",caption="TODO")
+knitr::kable(summ, format="latex", position="t",caption="Predictions from fitted Bayesian regression models.
+All terms were centered, but not rescaled.
+Units are in ms.
+Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")
 ```
 
-```{r post-error}
-brm_txl_interact_post <- read_rds(paste0(here(),"/Analysis/brm_txl_interact_post.rds"))
-brm_grnn_interact_post <- read_rds(paste0(here(),"/Analysis/brm_grnn_interact_post.rds"))
-brm_ngram_interact_post <- read_rds(paste0(here(),"/Analysis/brm_ngram_interact_post.rds"))
-
-d <- show_summary(brm_txl_interact_post) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
-e <- show_summary(brm_grnn_interact_post) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
-f <- show_summary(brm_ngram_interact_post) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
-
-summ <- d %>% union(e) %>% union(f) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
-   mutate(Term=factor(Term, 
-                        levels=c("Intercept", 
-                                 "surp_center",
-                                 "length_center",
-                                 "freq_center",
-                                 "surp_center:length_center",
-                                 "length_center:freq_center",
-                                 "past_c_surp",
-                                 "past_c_length",
-                                 "past_c_freq",
-                                 "past_c_surp:past_c_length",
-                                 "past_c_length:past_c_freq")
-                                 )) %>% 
-  arrange(Term) %>% 
-  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency",
-                  "Surp x Length", "Freq x Length", "Past Surprisal",
-                  "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length")) %>% 
-  select(Term, `5-gram`, `GRNN`, TXL)
-
-knitr::kable(summ, format="latex", position="t",caption="TODO")
-
-```
-
-As we can see in Table \@ref(tab:pre-error) and Table \@ref(tab:post-error).
+As we can see in Table \@ref(tab:pre-error), we find large effects of surprisal and length, but minimal effects of frequency.
+These effects are larger than what is usually reported in other methods CITE, but this could be due to the overall slowness of the method; the intercepts are also much bigger.
+The lack of frequency effects is somewhat surprising, and some CITE argue that they aren't real anyway.
+Notably the coefficients for the lagged terms are smaller than surprisal and length of the current word.
 
@@
 
-We want to be able to do nested model comparision, so we want models with 
-- only length, frequency fx
-- each 1 surprisal model as predictor
-- all the surprisals
-
-For now, no interactions between surprisal and others, and if we include a predictor, include both current and lagged of it.
+We also ran a nested model comparison with models fit in lmer.
+For this, we fit models with only frequency and length as predictors, as well as models that also had one or more sources of surprisal.
+For all effects, we included both current and lagged versions of it.
+The model with all 3 surprisal predictors is better than any model with only one.
+Any one surprisal predictor is better than no surprisals predictors.
+However, adding ngram predictors to a model that already has txl & grnn does not help.
+In other cases, adding the 3rd surprisal source to the other two does help.
+Ngram+Grnn is not better than Grnn only.
+Otherwise, pairs are better than singletons.
+This suggests that Ngram's info is a subset of GRNN, but not a subset of TXL.
 
@@
 ```
-In general, adding to the model makes it better.
 
-The model with all 3 surprisal predictors is better than any model with only one. Any one surprisal predictor is better than no surprisals predictors.
-
-However, adding ngram predictors to a model that already has txl & grnn does not help. 
-In other cases, adding the 3rd surprisal source to the other two does help.
-
-Ngram+Grnn is not better than Grnn only. Otherwise, pairs are better than singletons. This suggests that Ngram's info is a subset of GRNN, but not a subset of TXL.
-
-Past surprisal predictors don't help (with or without past freq,length effects in the models), but past freq,length do (with or without past surprisal predictors).
-
-## Comparision with SPR
-
+<!-- ## Comparision with SPR
 We want to compare how good our participants were on the comprehension questions to how good people were on the SPR experiment reported in Futrell et al (2017).
@@
 However, reading times in a story were excluded based on the comprehension of that story, so it makes sense to compare per-story accuracies with our per-story accuracies.
+-->
 
-## Error patterns
+<!-- ## Error patterns
 When participants do make errors, they correct them quickly, generally within one second, which means it's probably not disrupting their reading much.
@@
 Of the `r nrow(data_sentence)` sentences that good participants completed, `r filter(data_sentence, sent.correct==1) %>% nrow()` were completed entirely correctly.
+-->
 

### 2021-04-19 Papers/UCI_2021/uci_2021.tex
start talk

--- 2020-05-28 Papers/lab_meeting_2020.tex
@@
 
+\DeclareRobustCommand{\greencheck}{%
+	\tikz\fill[scale=0.6, color=ForestGreen]
+	(0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;%
+}
 \title{A-maze: Easier measurement of incremental processing}
 \author{Veronica Boyce}
-\date{3 June 2020}
+\date{26 April 2021}
 
@@
 
+\begin{frame}{Road map} %Acknowledge collaborators 
+	%TODO: find a picture of Richard!!
+	Methods talk 
+	- briefly summarize RT methods, introduce Maze task
+	- talk through study 1 (w/ Roger and Richard) where we created and tested this variant method
+	- talk through study 2 where I tweaked the method and did another study
+	- have a little historical revisionism -- there were some boring implementational tweaks, so what I'll present is a little different that what was used in the first study
+	\begin{tikzpicture}[remember picture,overlay]
+	\node[xshift=-1.8cm,yshift=-1cm] at (current page.east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{../Images/roger.jpg}};
+	\end{tikzpicture}
+
+	\begin{tikzpicture}[remember picture,overlay]
+	\node[xshift=2.5cm,yshift=-1.5cm] at (current page.north west) {\includegraphics[width=.3\textwidth]{../Images/cpl_header.png}};
+	\end{tikzpicture}
+
+\end{frame}
 %
@@
 
-			\includegraphics[width=.9\textwidth]{eye-tracker.jpeg}
+			\includegraphics[width=.9\textwidth]{../Images/eye-tracker.jpeg}
 			\pause
@@
 
-\begin{tikzpicture} %TODO on slide etc
+\begin{tikzpicture} 
 \draw (-.5,-.5) rectangle (3.2,.5);
@@
 \end{columns}
+	\bigskip
+\onslide<8->{Different methods have different trade-offs}
 \end{frame}
 
-\begin{frame}{A third option: Maze}
-\end{frame}
-\begin{frame}{Maze Task}
+\begin{frame}[c]{An alternative: Maze}
+
 	\centering
+
+	\bigskip
+
 	\tikzset{
@@
 	\begin{tikzpicture}
-\onslide<1->{\node[draw] (a) at (-1,-.5) {e};}
-\onslide<1->{\node[draw] (b) at (1,-.5) {i};}
 	\onslide<1-2>{\node  (c) at (-1,.5) {The};}
@@
 	\onslide<9-10>{\node  (d) at (1,.5) {squirrel.};}
-\onslide<2,8>{\node[draw, fill=pink!50] at (a) {e};}
-\onslide<4,6,10>{\node[draw, fill=pink!50] at (b) {i};}
+	\onslide<2,8>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
+	\onslide<4,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
+		align=center] at (d) {};}
 	\end{tikzpicture}
@@
 		\textbf{\large G-maze}
-		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{gmaze.pdf}
+		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../Images/gmaze.pdf}
 		\end{center}
@@
 		\textbf{\large L-maze}
-			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{lmaze.pdf}
+			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../Images/lmaze.pdf}
 		\end{center}
@@
 
+\begin{frame}{Maze Made Easy}
+
+	Can we use Maze instead of web SPR?\pause
+
+	\medskip
+
+	Needs some tweaks:\pause
+	\begin{itemize}
+		\item Run on web \pause
+		\item Easily generate distractors \pause
+		\item Work for multi-sentence items
+	\end{itemize}
+
+\end{frame}
+
 \section{Experiment 1a}
-\begin{frame}{Web Implementation}
-Maze well suited to run on the web
-\begin{itemize}
-	\item Implement in Ibex
-\end{itemize} 
+\begin{frame}{Run on web}
+
 	\pause
+	Wrote an Ibex module \pause
+
+	\begin{center}\includegraphics[width=.7\textwidth]{../Images/screenshot.png} \end{center}\pause
 
@@
 \end{small}
-\includegraphics[width=\textwidth]{g_rel.pdf}
+\includegraphics[width=\textwidth]{../Images/g_rel.pdf}
 \end{frame}
@@
 \end{small}
-\includegraphics[width=\textwidth]{g_adv.pdf}
+\includegraphics[width=\textwidth]{../Images/g_adv.pdf}
 \end{frame}
@@
 \end{small}
-\includegraphics[width=\textwidth]{g_svnp.pdf}
+\includegraphics[width=\textwidth]{../Images/g_svnp.pdf}
 \end{frame}
 
-\begin{frame}{Interim Summary}
-\textcolor{PineGreen}{\Large \textbf{Good news:} G-maze works well over the web (better than L-maze or SPR)}
+\begin{frame}{Maze Made Easy}
 
-\pause
+	Can we use Maze instead of web SPR?
 
-\bigskip
+	\medskip
 
-\textcolor{Red}{\Large \textbf{Bad news:} I do not want to write G-maze materials.}
+	Needs some tweaks:
+	\begin{itemize}
+		\item Run on web \greencheck
+		\item Easily generate distractors
+		\item Work for multi-sentence items 
+	\end{itemize}
 
@@
 \section{A-maze}
+
+\begin{frame}{Generating distractors}
+	\pause
+	Goal: Find a word that can't continue a partial sentence
+	\begin{itemize}
+		\item Ex.  \textit{The dog chased} \pause
+		\item Tedious (and hard!) to do by hand
+	\end{itemize} \pause
+
+	What makes something an unacceptable continuation? \pause
+	\begin{itemize}
+		\item Ungrammatical \pause
+		\item ...or otherwise really unlikely \pause 
+		\item $\approx$ high surprisal 
+	\end{itemize} \pause
+
+	Can we use Neural Language Models?
+\end{frame}
 
@@
 	\centering
-\includegraphics[clip, trim=3.25cm 5cm 2.5cm 2.5cm,width=.9\textwidth]{flow_2.pdf}
+	\includegraphics[clip, trim=3.25cm 5cm 2.5cm 2.5cm,width=.9\textwidth]{../Images/flow_2.pdf}
 \end{frame}
+
 \section{Experiment 1b}
 
-\begin{frame}{Does A-maze work? }
-\pause
-Test it on materials from Witzel et al. (2012)
+\begin{frame}{Does it work?} 
+	%\vskip -2em
+	\small
+	From ``Maze Made Easy'' (Boyce et al 2020)
 	\medskip
 
-\pause
-Try with two pre-trained LSTM models (Gulordava 2018, Jozefowicz 2016)
+	\centering
+
+	\includegraphics[width=\textwidth]{../Images/spr_g_amaze.pdf}
+
+	{\footnotesize Error bars: 95\% CI}
 
@@
 
-\begin{frame}{Does A-maze work?  Yes.} \pause
-\includegraphics[width=1\textwidth]{a_all.pdf}
+\begin{frame}{Does it work?} 
+	\pause
+	%\vskip -2em
+	{\large Yes, at least well enough.} \pause
+	\begin{itemize}
+		\item Sometimes generates plausible distractors. \pause
+		\item Sloggett et al 2020 also found A-maze results comparable with G-maze
+	\end{itemize} 
 \end{frame}
 
-\begin{frame}{Caveats}
-A lot of mistakes on the second word of the sentence ("The \ul{dog} ...") \pause
-\includegraphics{graph_errors2.pdf} \pause
-Bad participants or bad distractors?
+\section{Error-correction A-Maze}
 
-Either way, lots of data loss.
+\begin{frame}{Maze Made Easy}
+
+	Can we use Maze instead of web SPR?
+
+	\medskip
+
+	Needs some tweaks:
+	\begin{itemize}
+		\item Run on web \greencheck
+		\item Easily generate distractors \greencheck
+		\item Work for multi-sentence items 
+	\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Long items}
+
+	Want to run multi-sentence items. \pause
+
+	Problem: Errors terminate sentences. \pause
+	\begin{itemize}
+		\item Treat whole story as a unit: \pause Few participants make it to the end. \pause
+		\item Treat each sentence as a unit: \pause Some participants miss key context. \pause
+
+	\end{itemize}
+
+	What if after an error, participants corrected errors and the sentence continued?
+
+\end{frame}
+
+\begin{frame}[c]{Maze with Error Correction}
+	\centering
+	\tikzset{
+		font={\fontsize{14pt}{12}\selectfont}}
+
+	\begin{tikzpicture}
+	\onslide<1-2>{\node  (c) at (-1,.5) {The};}
+	\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
+	\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
+	\onslide<3-4>{\node  (d) at (1,.5) {dog};}
+	\onslide<5-8>{\node  (c) at (-1,.5) {revise};}
+	\onslide<5-8>{\node  (d) at (1,.5) {chased};}
+	\onslide<9-10>{\node  (c) at (-1,.5) {the};}
+	\onslide<9-10>{\node  (d) at (1,.5) {wish};}
+	\onslide<11-12>{\node  (c) at (-1,.5) {mitigate.};}
+	\onslide<11-12>{\node  (d) at (1,.5) {squirrel.};}
+	\onslide<2,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
+	\onslide<4,8,12>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
+		align=center] at (d) {};}
+	\onslide<7,8>{\node[text=red] at (0,-1.5) {\small Incorrect. Please try again.};}
+	\end{tikzpicture}
+
+\end{frame}
+
+\begin{frame}{Maze with Error Correction}
+
+	Incidentally, also ``solves'' problem of bad distractors!
+
+	\begin{columns}
+		\begin{column}{0.4\textwidth}
+			\begin{center}
+				\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=\textwidth]{../Images/maze_diagram.pdf}
+			\end{center}
+
+		\end{column} \pause 
+		\begin{column}{0.6\textwidth} 
+			\begin{center}
+				\begin{itemize}
+					\item Can be toggled in Ibex Maze \pause
+					\item Long materials feasible \pause
+					\item Have all the data \pause
+					\item Compensates for bad distractors
+				\end{itemize}
+			\end{center}
+		\end{column}
+	\end{columns}
+\end{frame}
+%
+\begin{frame}{Maze Made Easy}
+
+	Can we use Maze instead of web SPR?
+
+	\medskip
+
+	Needs some tweaks:
+	\begin{itemize}
+		\item Run on web \greencheck
+		\item Easily generate distractors \greencheck
+		\item Work for multi-sentence items \greencheck ? 
+	\end{itemize}
+
+\end{frame}
+
+\section{Experiment 2}
+\begin{frame}{}
+	Various open questions to address \pause
+	\begin{itemize}
+		\item Will people read long texts in Maze? \pause
+		\item Will they comprehend what they read? \pause
+		\item Does error correction Maze work? \pause
+		\item Do we get predictability effects? 
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Natural Stories}
+	Natural stories corpus (Futrell et al. 2017) \pause
+	\begin{itemize}
+		\item 10 stories, each about 1000 words \pause
+		%	\item Some unusual constructions, but read fluently \pause
+		\item 6 comprehension questions per story
+	\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Natural Stories}
+
+\begin{small}Tulip mania was a period in the Dutch Golden Age during which contract prices for bulbs of the recently introduced tulip reached extraordinarily high levels and then suddenly collapsed.
+At the peak of tulip mania in February sixteen thirty-seven, tulip contracts sold for more than ten times the annual income of a skilled craftsman.
+It is generally considered the first recorded economic bubble.
+[...]
+\medskip
+
+Q: When did tulip mania reach its peak?
+
+A: \hspace{3em} 1630's\hspace{3em} 1730's \end{small}
+
+\end{frame}
+
+\begin{frame}{Participant accuracy}
+100 participants each read 1 story \pause
+\begin{center}
+\includegraphics[width=.8\textwidth]{../Images/error.pdf}
+\end{center}
+\end{frame}
+\begin{frame}{Comprehension questions}
+\begin{center} \pause
+\includegraphics[width=.8\textwidth]{../Images/comp.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}{Surprisal Effects}
+	\large{Is RT linear in terms of surprisal?}\pause
+	\medskip
+
+	Estimate surprisal from 3 models:
+	\begin{itemize}
+		\item smoothed 5-gram
+		\item LSTM-RNN (Gulordava et al. 2018)
+		\item Transformer-XL (Dai et al. 2019)
+	\end{itemize}
+
+	\pause
+
+	Fit GAMs
+	\begin{itemize}
+		%\item Limit to single-token words
+		\item Fit to both current and past word surprisal
+		\item Include frequency, length as predictors
+	\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Surprisal Effects}
+	\includegraphics[width=.9\textwidth]{../Images/amaze_gam.pdf}	
+\end{frame}
+
+\begin{frame}{SPR comparison}
+
+	\includegraphics[width=\textwidth]{../Images/spr_gam.pdf}
+
+\end{frame}
+
+\begin{frame}{Surprisal Effects}
+	\vskip -2em
+	Linear Models \\ 
+	\pause
+
+	\begin{small}
+		\begin{tabular}{l|r|r|r}
+			\hline
+			&5-gram & GRNN & TXL\\
+			\hline
+			Intercept & \textbf{865.3} & \textbf{871.1} & \textbf{870.8} \\ 
+			Surprisal & \textbf{11.7}  & \textbf{23.7} & \textbf{18.5}  \\ 
+			Frequency & -2.9  & 2.9  & 0.4  \\ 
+			Length & \textbf{20.5} & \textbf{18.5}  & \textbf{21.4} \\ 
+			Surprisal:Length & \textbf{-2.0}  & \textbf{-1.8}  & \textbf{-1.4} \\
+			Freq:Length & -1.0  & -0.1 & 0.2  \\ 
+			\hline
+			Past Surprisal & 1.6  & \textbf{2.7} & 0.8 \\ 
+			Past Freq & 2.6  & 1.9  & 1.2  \\ 
+			Past Length & \textbf{-4.8}  & \textbf{-6.6 }&\textbf{ -5.2} \\ 
+			Past Surp:Length & -0.2 & \textbf{-0.9} & -0.6  \\ 
+			Past Freq:Length & -1.0  & \textbf{-1.8 }&   \textbf{-1.5}  \\ 
+			\hline
+		\end{tabular}
+
+		Surprisal in bits, Length in characters, \\
+		Frequency in $log_2$ occurrences/billion words
+	\end{small}
+
+\end{frame}
+\begin{frame}{Surprisal Effects}
+	Takeaways: \pause
+	\begin{itemize}
+		\item Minimal frequency effects (consistent with Shain 2019) \pause
+		\item Large effects of Length, Surprisal \pause
+		\item Very little spillover \pause
+	\end{itemize}
+	%	
+	Model comparison: GRNN is best, but TXL complementary
+\end{frame}
+
+\begin{frame}{Summary}
+	\begin{itemize}\pause
+		\item People will read in the Maze task for 15-20 minutes \pause
+		\item It's possible to comprehend during Maze \pause
+		\item Distractors are generally good enough \pause
+		\item Find expected RT patterns \pause
+		\item Very little spillover
+	\end{itemize}
+\end{frame}
+
+\section{Conclusion}
+
+\begin{frame}{Consider A-maze!}
+Easy to use! \pause
+\begin{itemize}
+	\item Runs on command line \pause
+	\item Match distractors across minimal pair sentences \pause
+	\item Customize surprisal thresholds, vocabulary lists \pause
+	\item Can output pre-formatted for Ibex
+\end{itemize}
+\medskip
+\end{frame}
+
+\begin{frame}{Consider A-maze!} \pause
+Adapt A-maze to your projects:
+\begin{itemize}
+	\item Written in Python 3 
+	\item Interface with other language models 
+	\item Add more frequency sources 
+	\item Extend to non-English languages
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{}
+
+\textcolor{ForestGreen}{\large Documentation: vboyce.github.io/Maze}
+
+with links to the following:
+\begin{itemize}
+
+\item A-maze code: github.com/vboyce/Maze
+
+\item Web-maze code: github.com/vboyce/Ibex-with-Maze
+
+\item Paper: psyarxiv.com/b7nqd/
+\end{itemize}
+\end{frame}
+
+\appendix
+
+\begin{frame}{Matching distractors}
+If unspecified: Match by position
+\begin{itemize}
+	\item The son of the lady who politely introduced \sethlcolor{green} \hl{herself}\sethlcolor{pink} / \hl{himself} was popular at the party.
+\end{itemize}
+Can specify labels for each word to pair (within item)
+\begin{itemize}
+	\item The cat who the dog scared hid in a box.\\pre-1 pre-2 who art noun verb main-verb post-1 post-2 post-3
+\item The dog who scared the cat sniffed around the couch.\\ pre-1 pre-2 who verb art noun main-verb post-1 post-2 post-3
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Regression coefficients}
+\begin{tiny}
+\begin{tabular}{l|rlr|rlr|rlr}
+	\hline
+	&\multicolumn{3}{c|}{5-gram}&\multicolumn{3}{c|}{GRNN}&\multicolumn{3}{c}{TXL}\\
+	& Est & CI & $p$ & Est & CI & $p$ &Est & CI & $p$ \\ 
+	\hline
+	Intercept & 865.3 & [829.9, 902.9] & 0.00 & 871.1 & [837.9, 905.3] & 0.00 & 870.8 & [832.5, 907.8] & 0.00 \\ 
+	Surprisal & 11.7 & [9.3, 14.1] & 0.00 & 23.7 & [21, 26.5] & 0.00 & 18.5 & [16.1, 21.1] & 0.00 \\ 
+	Frequency & -2.9 & [-6.3, 0.5] & 0.10 & 2.9 & [-0.2, 6] & 0.06 & 0.4 & [-2.7, 3.5] & 0.79 \\ 
+	Length & 20.5 & [15.4, 25.6] & 0.00 & 18.5 & [13.3, 23.7] & 0.00 & 21.4 & [16.2, 26.6] & 0.00 \\ 
+	Surprisal:Length & -2.0 & [-3, -1] & 0.00 & -1.8 & [-2.7, -0.9] & 0.00 & -1.4 & [-2.2, -0.6] & 0.00 \\
+	Freq:Length & -1.0 & [-2.5, 0.4] & 0.16 & -0.1 & [-1.2, 1] & 0.82 & 0.2 & [-0.9, 1.2] & 0.76 \\ 
+	\hline
+	Past Surprisal & 1.6 & [-0.5, 3.6] & 0.14 & 2.7 & [0.8, 4.5] & 0.00 & 0.8 & [-0.9, 2.5] & 0.40 \\ 
+	Past Freq & 2.6 & [-0.1, 5.4] & 0.06 & 1.9 & [-0.2, 4.2] & 0.08 & 1.2 & [-1.1, 3.6] & 0.30 \\ 
+	Past Length & -4.8 & [-9, -0.1] & 0.04 & -6.6 & [-10.9, -2.1] & 0.00 & -5.2 & [-9.3, -0.7] & 0.03 \\ 
+	Past Surp:Length & -0.2 & [-1.2, 0.8] & 0.72 & -0.9 & [-1.7, -0.2] & 0.01 & -0.6 & [-1.3, 0.2] & 0.13 \\ 
+	Past Freq:Length & -1.0 & [-2.3, 0.3] & 0.15 & -1.8 & [-2.9, -0.8] & 0.00 & -1.5 & [-2.6, -0.5] & 0.01 \\ 	
+	\hline
+\end{tabular}
+\end{tiny}
 \end{frame}
@@
 
-\section{Experiment 2}
+\begin{frame}{What about post-mistake data?}
+	Exclude data from mistakes or the two words after a mistake.
 
-\begin{frame}{Potential improvements}
+	\includegraphics[width=.9\textwidth]{../Images/gam2.pdf}	
+\end{frame}
 
-{\large Option 1: Choose better distractors.} \pause
+\begin{frame}{Why such large effects?}
+	\pause
+	Bayesian Reader (Norris 2006): Look at words long enough to ID with some threshold of certainty \pause
+
+	Possible mechanisms for difference: \pause
 	\begin{itemize}
-	\item Tweak thresholds of distractor selection. \pause
-	\item Could search for optimal thresholds. \pause
-	\item Could manually check and fix distractors. 
+		\item Higher threshold \pause 
+		\item Fewer available resources for processing \pause
+		\item Presence of second word 
 	\end{itemize}
@@
 \end{frame}
-
-\begin{frame}{Potential improvements}
-
-{\large Option 2: Be forgiving.} \pause
-\begin{itemize}
-	\item Just don't terminate on mistakes. \pause
-	\item Instead have participant correct mistake and continue.
-	\end{itemize}
-\end{frame}
-
-\begin{frame}{Maze with Error Correction}
-\centering
-\tikzset{
-	font={\fontsize{14pt}{12}\selectfont}}
-\begin{tikzpicture}
-\onslide<1->{\node[draw] (a) at (-1,-.5) {e};}
-\onslide<1->{\node[draw] (b) at (1,-.5) {i};}
-\onslide<1-2>{\node  (c) at (-1,.5) {The};}
-\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
-\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
-\onslide<3-4>{\node  (d) at (1,.5) {dog};}
-\onslide<5-8>{\node  (c) at (-1,.5) {revise};}
-\onslide<5-8>{\node  (d) at (1,.5) {chased};}
-\onslide<9-10>{\node  (c) at (-1,.5) {the};}
-\onslide<9-10>{\node  (d) at (1,.5) {wish};}
-\onslide<11-12>{\node  (c) at (-1,.5) {mitigate.};}
-\onslide<11-12>{\node  (d) at (1,.5) {squirrel.};}
-\onslide<2,6,10>{\node[draw, fill=pink!50] at (a) {e};}
-\onslide<4,8,12>{\node[draw, fill=pink!50] at (b) {i};}
-\onslide<7,8>{\node[text=red] at (0,-1.5) {\small Incorrect. Please try again.};}
-\end{tikzpicture}
-\end{frame}
-
-\begin{frame}{Maze with Error Correction}
-
-{\large The ``solution'' to our problems!} \pause
-
-\begin{columns}
-	\begin{column}{0.3\textwidth}
-		\begin{center}
-		\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=\textwidth]{maze_diagram.pdf}
-		\end{center}
-
-	\end{column} \pause 
-	\begin{column}{0.7\textwidth} 
-		\begin{center}
-			\begin{itemize}
-				\item Easily implemented as a toggle in Ibex \pause
-				\item Have all the data \pause
-				\item Identify bad participants v bad distractors 
-			\end{itemize}
-		\end{center}
-	\end{column}
-\end{columns}
-\medskip
-\pause
-
-Side effect: Can now run multi-sentence items!
-\end{frame}
-
- %TODO: Needs work from here on!
-
-\begin{frame}{Natural Stories}
-Natural stories corpus (Futrell et al. 2017)
-\begin{itemize}
-	\item 10 stories, each about 1000 words
-	\item Some unusual constructions, but read fluently
-	\item 6 comprehension questions per story
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}{Natural Stories}
-
-\begin{small}Tulip mania was a period in the Dutch Golden Age during which contract prices for bulbs of the recently introduced tulip reached extraordinarily high levels and then suddenly collapsed.
-At the peak of tulip mania in February sixteen thirty-seven, tulip contracts sold for more than ten times the annual income of a skilled craftsman.
-It is generally considered the first recorded economic bubble.
-[...]
-\medskip
-
-Q: When did tulip mania reach its peak?
-
-A: \hspace{3em} 1630's\hspace{3em} 1730's \end{small}
-
-\end{frame}
-
-\begin{frame}{Experiment Plan}
-Experiment questions:
-\begin{itemize}
-	\item Will people do this task?
-	\item Can participants comprehend the stories?
-\end{itemize}
-
-\pause
-
-\medskip
-
-Checks on RT data
-\begin{itemize}
-	\item Check if RT is linear in surprisal (Smith \& Levy 2013)
-	\item Check for known effects of word length, frequency
-	\item Check for spillover effects
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}{Participant accuracy}
-100 participants each read 1 story \pause
-\begin{center}
-\includegraphics[width=.8\textwidth]{error.pdf}
-\end{center}
-\end{frame}
-\begin{frame}{Comprehension questions}
-\begin{center} \pause
-\includegraphics[width=.8\textwidth]{comp.pdf}
-\end{center}
-\end{frame}
-
-\begin{frame}{Surprisal Effects}
-
-Use 3 LMs to estimate surprisal: smoothed 5-gram, Gulordava RNN, Transformer-XL (Gulordava et al. 2018, Dai et al. 2019)
-\pause
-
-Fit GAMs on pre-error data.
-\begin{itemize}
-	\item Limit to single-token words.
-	\item Fit to both current and past word surprisal.
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}{Surprisal Effects}
-\includegraphics[width=.9\textwidth]{gam.pdf}	
-\end{frame}
-\begin{frame}{What about post-mistake data?}
-Exclude data from mistakes or the two words after a mistake.
-
-\includegraphics[width=.9\textwidth]{gam2.pdf}	
-\end{frame}
-
-\begin{frame}{Summary}
-\begin{itemize}\pause
-	\item People will read in the Maze task for 15-20 minutes \pause
-	\item It's possible to comprehend during Maze \pause
-	\item Distractors are generally good enough \pause
-	\item Find expected RT patterns \pause
-	\item Very little spillover
-\end{itemize}
-\end{frame}
-\section{Conclusion}
-
-\begin{frame}{Use A-maze!}
-Easy to use! \pause
-\begin{itemize}
-	\item Runs on command line \pause
-	\item Match distractors across minimal pair sentences \pause
-	\item Customize surprisal thresholds, vocabulary lists \pause
-	\item Can output pre-formatted for Ibex
-\end{itemize}
-\medskip
-\end{frame}
-
-\begin{frame}{Use A-maze!} \pause
-Contribute to A-maze:
-\begin{itemize}
-	\item Written in Python 3 \pause
-	\item Interface with other language models \pause
-	\item Add more frequency sources \pause
-	\item Extend to non-English languages
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}{}
-
-\textcolor{ForestGreen}{\large Documentation: vboyce.github.io/Maze}
-
-with links to the following:
-\begin{itemize}
-
-\item A-maze code: github.com/vboyce/Maze
-
-\item Web-maze code: github.com/vboyce/Ibex-with-Maze
-
-\item Sample task: syntaxgym.org:666
-
-\item Paper: psyarxiv.com/b7nqd/
-\end{itemize}
-\end{frame}
-
-\appendix
-
-\begin{frame}{Matching distractors}
-If unspecified: Match by position
-\begin{itemize}
-	\item The son of the lady who politely introduced \sethlcolor{green} \hl{herself}\sethlcolor{pink} / \hl{himself} was popular at the party.
-\end{itemize}
-Can specify labels for each word to pair (within item)
-\begin{itemize}
-	\item The cat who the dog scared hid in a box.\\pre-1 pre-2 who art noun verb main-verb post-1 post-2 post-3
-\item The dog who scared the cat sniffed around the couch.\\ pre-1 pre-2 who verb art noun main-verb post-1 post-2 post-3
-\end{itemize}
-\end{frame}
-%back pocket on more on model results??
-%do we even need back pockets??
-
-\begin{frame}{Post-mistake only}
-Exclude data from mistakes or the two words after a mistake.
-
-\includegraphics[]{gam3.pdf}	
-\end{frame}
-
-\begin{frame}{Regression coefficients}
-\begin{tiny}
-\begin{tabular}{l|rlr|rlr|rlr}
-	\hline
-	&\multicolumn{3}{c|}{5-gram}&\multicolumn{3}{c|}{GRNN}&\multicolumn{3}{c}{TXL}\\
-	& Est & CI & $p$ & Est & CI & $p$ &Est & CI & $p$ \\ 
-	\hline
-	Intercept & 865.3 & [829.9, 902.9] & 0.00 & 871.1 & [837.9, 905.3] & 0.00 & 870.8 & [832.5, 907.8] & 0.00 \\ 
-	Surprisal & 11.7 & [9.3, 14.1] & 0.00 & 23.7 & [21, 26.5] & 0.00 & 18.5 & [16.1, 21.1] & 0.00 \\ 
-	Frequency & -2.9 & [-6.3, 0.5] & 0.10 & 2.9 & [-0.2, 6] & 0.06 & 0.4 & [-2.7, 3.5] & 0.79 \\ 
-	Length & 20.5 & [15.4, 25.6] & 0.00 & 18.5 & [13.3, 23.7] & 0.00 & 21.4 & [16.2, 26.6] & 0.00 \\ 
-	Surprisal:Length & -2.0 & [-3, -1] & 0.00 & -1.8 & [-2.7, -0.9] & 0.00 & -1.4 & [-2.2, -0.6] & 0.00 \\
-	Freq:Length & -1.0 & [-2.5, 0.4] & 0.16 & -0.1 & [-1.2, 1] & 0.82 & 0.2 & [-0.9, 1.2] & 0.76 \\ 
-	\hline
-	Past Surprisal & 1.6 & [-0.5, 3.6] & 0.14 & 2.7 & [0.8, 4.5] & 0.00 & 0.8 & [-0.9, 2.5] & 0.40 \\ 
-	Past Freq & 2.6 & [-0.1, 5.4] & 0.06 & 1.9 & [-0.2, 4.2] & 0.08 & 1.2 & [-1.1, 3.6] & 0.30 \\ 
-	Past Length & -4.8 & [-9, -0.1] & 0.04 & -6.6 & [-10.9, -2.1] & 0.00 & -5.2 & [-9.3, -0.7] & 0.03 \\ 
-	Past Surp:Length & -0.2 & [-1.2, 0.8] & 0.72 & -0.9 & [-1.7, -0.2] & 0.01 & -0.6 & [-1.3, 0.2] & 0.13 \\ 
-	Past Freq:Length & -1.0 & [-2.3, 0.3] & 0.15 & -1.8 & [-2.9, -0.8] & 0.00 & -1.5 & [-2.6, -0.5] & 0.01 \\
-
-	\hline
-\end{tabular}
-\end{tiny}
-\end{frame}
-
 \end{document}

### 2021-04-21 Papers/UCI_2021/uci_2021.tex
updates

@@
-\documentclass[12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman]{beamer}
+\documentclass[ 12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman]{beamer}
 \usetheme{Frankfurt}
@@
 
-\begin{frame}{Road map} %Acknowledge collaborators 
-	%TODO: find a picture of Richard!!
-	Methods talk 
-	- briefly summarize RT methods, introduce Maze task
-	- talk through study 1 (w/ Roger and Richard) where we created and tested this variant method
-	- talk through study 2 where I tweaked the method and did another study
-	- have a little historical revisionism -- there were some boring implementational tweaks, so what I'll present is a little different that what was used in the first study
+\begin{frame}{Plan} %Acknowledge collaborators 
+	\pause
+%Methods \& experiments to test them
+\bigskip
+\begin{itemize}
+	\item Maze task
+	\item A-maze
+	\item Study 1: methods comparison
+	\item Variant of A-maze
+	\item Study 2: test on Natural Stories Corpus
+\end{itemize}\pause 
 	\begin{tikzpicture}[remember picture,overlay]
-	\node[xshift=-1.8cm,yshift=-1cm] at (current page.east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{../Images/roger.jpg}};
+	\node[xshift=-2cm,yshift=0cm] at (current page.east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{../Images/roger.jpg}};
 	\end{tikzpicture}
-
 		\begin{tikzpicture}[remember picture,overlay]
-	\node[xshift=2.5cm,yshift=-1.5cm] at (current page.north west) {\includegraphics[width=.3\textwidth]{../Images/cpl_header.png}};
+	\node[xshift=-2cm,yshift=2cm] at (current page.south east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{../Images/richard.jpg}};
+	\end{tikzpicture}
+	\begin{tikzpicture}[remember picture,overlay]
+	\node[xshift=-2cm,yshift=-2cm] at (current page.north east) {\includegraphics[width=.3\textwidth]{../Images/cpl_header.png}};
 	\end{tikzpicture}
@@
 Examples of increased difficulty
-\begin{itemize}\pause
-	\item Constructions not in the grammar \pause
-	\item Lexical items that are harder to retrieve \pause
+\begin{itemize}
+	\item Constructions not in the grammar 
+	\item Lexical items that are harder to retrieve 
 	\item Words that force reparsing or reanalysis
@@
 	\bigskip
-\onslide<8->{Different methods have different trade-offs}
+\onslide<9->{Different methods have different trade-offs}
 \end{frame}
@@
 	\pause
-	Wrote an Ibex module \pause
-
-	\begin{center}\includegraphics[width=.7\textwidth]{../Images/screenshot.png} \end{center}\pause
+	\begin{columns}
+		\begin{column}{0.5\textwidth}
+				Wrote an Ibex module
+		\end{column}
+		\begin{column}{0.5\textwidth} 
+			\begin{center}\includegraphics[width=\textwidth]{../Images/screenshot.png} \end{center}\pause
+		\end{column}
+	\end{columns}
 
@@
 \end{itemize}
-\pause
 Surprisal: negative log probability
@@
 \end{itemize}
-\pause
 \begin{tikzpicture}
@@
 	%\vskip -2em
-	\small
-	From ``Maze Made Easy'' (Boyce et al 2020)
 	\medskip
@@
 	\begin{itemize}
-		\item Sometimes generates plausible distractors. \pause
-		\item Sloggett et al 2020 also found A-maze results comparable with G-maze
+		\item Caveat: Sometimes generates plausible distractors. \pause
+		\item Sloggett et al (2020) also found A-maze results comparable with G-maze
 	\end{itemize} 
@@
 
-	Incidentally, also ``solves'' problem of bad distractors!
+%	Incidentally, also ``solves'' problem of bad distractors!
 
@@
 \begin{frame}{Surprisal Effects}
-	\includegraphics[width=.9\textwidth]{../Images/amaze_gam.pdf}	
+	\includegraphics[width=\textwidth]{../Images/amaze_gam.pdf}	
 \end{frame}
@@
 \begin{frame}{SPR comparison}
+	Using SPR data collected by Futrell et al. (2017)
 
@@
 \begin{frame}{Surprisal Effects}
-	Takeaways: \pause
+	Takeaways: 
 	\begin{itemize}
-		\item Minimal frequency effects (consistent with Shain 2019) \pause
-		\item Large effects of Length, Surprisal \pause
-		\item Very little spillover \pause
+		\item Minimal frequency effects (consistent with Shain 2019) 
+		\item Large effects of Length, Surprisal
+		\item Very little spillover 
 	\end{itemize}
 	%	
-	Model comparison: GRNN is best, but TXL complementary
+%	Model comparison: GRNN is best, but TXL complementary
 \end{frame}
@@
 	\item Customize surprisal thresholds, vocabulary lists \pause
-	\item Can output pre-formatted for Ibex
+	\item Can output pre-formatted for Ibex\pause
 \end{itemize}
-\medskip
-\end{frame}
 
-\begin{frame}{Consider A-maze!} \pause
 Adapt A-maze to your projects:
@@
 \end{itemize}
-
+\medskip
 \end{frame}
@@
 
-\item Paper: psyarxiv.com/b7nqd/
+\item Exp 1 Paper: psyarxiv.com/b7nqd/
 \end{itemize}
@@
 \begin{frame}{Why such large effects?}
-	\pause
-	Bayesian Reader (Norris 2006): Look at words long enough to ID with some threshold of certainty \pause
+	Bayesian Reader (Norris 2006): Look at words long enough to ID with some threshold of certainty
 
-	Possible mechanisms for difference: \pause
+	Possible mechanisms for difference: 
 	\begin{itemize}
-		\item Higher threshold \pause 
-		\item Fewer available resources for processing \pause
+		\item Higher threshold 
+		\item Fewer available resources for processing
 		\item Presence of second word 

### 2021-06-17 Papers/Paper/manuscript.Rmd
start talk
writing

@@
     corresponding : yes    # Define only one corresponding author
-    address       : "TODO"
+    address       : "450 Jane Stanford Way, Building 420, Stanford University, Stanford, CA 94305"
     email         : "vboyce@stanford.edu"
@@
 abstract: |
-  TODO Easy to use, reliable methods are important to science.
-In studying how people understand language in real time, we use incremental processing methods to measure word-by-word reading time.
-However, neither of the common methods (eye-tracking and self-paced reading) can run over the web and produce localized effects.
-Another method, called the Maze task, seems to be able to do this, especially since a technique called A-maze makes the task much faster to construct.
-Due to task limitations, Maze has not been used on long, naturalistic passages, only short targeted materials.
-Here we present an adaptation of the Maze method suitable for long materials; we test this method on the Natural Stories corpus and find that participants can comprehend what they read while doing the task and that the reading time patterns are broadly similar to those found with other methods.
   We find support for the localization of reading time effects during the Maze task, as well as extending the range of materials Maze is suitable for.
+How long it takes to read a word in a sentence is reflective of how hard it is to identify and integrate the word in the surrounding context.
+Techniques that slow down the reading process and localize the processing time for each word are useful to understanding the time course of language processing.
+A-maze is a new method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities.
+We adapt A-maze for use on longer passages and test it on the Natural Stories corpus.
+We find that people can comprehend what they read during the Maze task.
+Moreover, the  task yields useable reaction time data with word predictability effects that are linear in the surprisal of the current word, with little spillover effect from the surprisal of the previous word.
+This expands the types of effects that can be studied with A-maze, showing it to be a a versatile alternative to eye-tracking and self-paced reading.
 
@@
  library("papaja")
- library("citr")
  library(here)
@@
 # Intro
-We process language incrementally, building a model of the sentence as we hear or read word.
-When a word is unexpected and does not fit the model of the sentence so far, our understanding of the sentence may need to be revised.
-Incremental processing methods measure how long each word takes to read as a proxy for how hard it is to incoroporate into the mental model of the sentence.
-Fluent language uses process language rapidly and flexibly adapt to many unexpected words, which makes it hard to pinpoint slow downs related to difficult-to-process words.
+It's remarkable how flexible we are when reading; while we do occasionally stumble when we read something unexpected, we often are able to read slightly unexpected things without a problem.
+However, these expectations shape how fast we read, even if we don't notice a stumble, unexpected words take longer to process as they force us to rebuild our burgeoning mental model of the sentence.
+Fortunately for fluent readers and unfortunately for studying language, this process of adjustment is very quick, which makes measures of reading time messy.
 
-TODO motivate why we care about localization
+Measures of online reading are one way to understand language and how the mind processes language.
+Many theories of language structure and language processing ground out in predictions about the difficulty of processing words.
+For instance, the subject v object relative debate includes theories that make fine-grained predictions about which word is how slow -- this needs localized methods to adjudicate it.
+Other theories such as noisy channel processing also need support from localized word-by-word results.
+TODO WHY *DO* we care about localization?
 
-The two most commonly used methods are eye-tracking and self-paced reading.
-Both these methods are known to have spillover effects, which makes localization harder.
-CITE some things TODO might be worth bringing up RT/surprisal/spillover findings hereOur next question is what the relationship between a word's predictability and it's reading time is for Maze.
+Incremental processing methods such as self-paced reading or eye-tracking measure how long someone spends looking at one word before moving on and use that as a proxy for how difficult word was in context.
+Across multiple methods for measuring per-word processing and reading time, how unexpected a word is correlates with how long it takes to read it and move on (CITE).
+However, the two major methods of measuring incremental processing both suffer from imprecise localization.
+In eye-tracking, people read naturally which involves skipping words, jumping ahead and looking back, the dynamics of which make it hard to isolate effects.
+Even when reading order is controlled in self-paced reading, readers are moving quickly through the text and it may take multiple words for slowdowns to catch up with them.
+
+One effect of this lack of localization is that reading time for a word is dependent not only on how unexpected a word is, but also how unexpected the previous word is.
+This is an indication of spillover from the previous word.
+
 It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability).
@@
 The quality of these A-Maze distractors is not up to that of hand-generated distractors, but @boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzel12.
+A-maze out performed a SPR control in detecting and localizing expected slowdown effects.
 @sloggettAmazeAnyOther also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.
 
-While new, A-maze seems like a potentially powerful addition to the psycholinguists toolkit.
+A-maze is a potentially powerful addition to the psycholinguists toolkit.
 However, like the other Maze tasks it has been limited in its application to single-sentence items probing minimal comparisons in constructed sentences.
-While some psycholinguistic phenomena are well-explored within individual targeted sentences, there's also a need for reading time data on longer passages, for instance in order to compare to language models or study discourse effects.
-Therefore, it's important to test whether the Maze task works on these types of materials.
+This limits its useful, as some important questions such as comparing human data to language models or studying discourse effects require processing times over multi-sentence passages.
+Therefore, it's important to expand the Maze task to these types of materials and verify that it finds comparable patterns to other methods.
 
@@
 # Error-correction maze
+<!-- GOOD -->
 (ref:diagram-cap) Schematic of error-correction Maze.
-A participant reads a sentence word by word, choosing the correct word at each time point (selections marked in blue).
-When they make a mistake, an error message is displayed, they try again, and continue with the sentence.
+A participant reads a sentence word by word, choosing the correct word at each time point (selections marked in blue ovals).
+When they make a mistake, an error message is displayed, so they try again and continue with the sentence.
 
@@
 
-We could try to improve the situation by auto-generating better distractors or hand-replacing problematic ones.
-However, the inherent data losses poses a more fundamental problem: even with attentive participants and good distractors, the Maze paradigm will eventually run into problems as error rates compound over long materials where few participants will make it to the end.
+This situation could be improved by auto-generating better distractors or hand-replacing problematic ones, but that does not solve the fundamental problem.
+Well-chosen distractors and attentive participants will reduce the error rate, but the error rate will still compound over long materials.
 For instance, with a 1% error rate, `r round(.99**15*100)`% would complete each 15-word sentence, but only `r round(.99**50*100)`% of participants would complete a 50 word vignette, and `r round(.99**200*100)`% a 200 word passages.
@@
 
-To resolve this, we introduce an *error-correction* variant of Maze, where we let participants continue on when they make a mistake.
-We present them with an error message and wait until they select the correct option, before continuing the sentence as normal.
-This *error-correction* variant is shown in Figure \@ref(fig:diagram).
-We make this "error-correction" Maze available as an option in a modification of the Ibex Maze implementation introduced in @boyceMazeMadeEasy2020 (TODO repo link).
+To resolve this, we introduce an *error-correction* variant of Maze shown in Figure \@ref(fig:diagram).
+When a participant makes an error, we present them with an error message and wait until they select the correct option, before continuing the sentence as normal.
+We make this "error-correction" Maze available as an option in a modification of the Ibex Maze implementation introduced in @boyceMazeMadeEasy2020 (https://github.com/vboyce/Ibex-with-Maze).
 The code records both the RT to the first click and also the total RT until the correct answer is selected as separate values.
 
-Our goal is that this error-correction will minimize the problems of having occasional poor auto-generated distractors by letting participants continue on and still read the whole item.
-It will also enable us to run long materials on Maze, thus expanding the range of questions that Maze can be used on.
+This variant of Maze expands the types of materials that can be used with maze to include arbitrarily long passages and cushions the impact of occasional problematic distractors.
 
@@
 The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers.
-At the same time, the passages contain copious punctuation (including quoted speech), many proper nouns, and low frequency grammatical constructions.
-Taken together, these properties make this a severe test of our process.
-The materials are on the difficult side for our participants and for the language model we use to generate distractors.
+At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions.
+Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants.
 If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on basically any naturalistic text.
 
-The corpus also comes with binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
+The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
 Using this corpus on the A-maze task allows us to address, first whether participants will read and understand long passages using A-maze with error correction, and second, whether the resulting RTs profiles will show expected patterns such as a linear relationship with surprisal.
@@
 
-We constructed A-maze distractors for the Natural stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 MTurkers to each read a story in the Maze paradigm.
+We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the Maze paradigm.
 
 ## Materials
-We took the texts of the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and split it into sentences because the A-Maze process works on the sentence level.
+We took the texts of the Natural Stories corpus [@futrellNaturalStoriesCorpus2020], split them into sentences, and ran the sentences through the A-maze generation process.
+We follow the A-maze generation process outlined in @boyceMazeMadeEasy2020, although we use an updated version of the codebase with fixes some issues identified in that paper (code at https://github.com/vboyce/Maze).
+Additionally, the capability to appropriately handle a wider variety of punctuation (needed for this corpus) was added.
+We took the auto-generated distractors as they were, without checking them for quality.
 We used the original comprehension questions provided in the Natural Stories corpus.
 To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions.
-We then ran the sentences of the items through the A-maze generation process.
-We updated the code of the generation process to fix some issues identified in that paper; the code used is available at REPO.
-The underlying method did not change, but features such as appropriately handling more types of punctuation (needed for the Natural Stories corpus) were added.
-We ran the materials using FOOBAR parameters.
-We used the auto-generated distractors as they were, without checking them for quality.
-Materials are available at REPO.
-TODO other notable changes to the codebase??
+All materials are available at TODO NEW REPO.
 
 ## Participants
-We recruited 100 participants from Amazon Mechanical Turk, and paid each participant 3.50 dollars.
+We recruited 100 participants from Amazon Mechanical Turk, and paid each participant 3.50 dollars, for roughly 20 minutes of work.
 We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
@@
 Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading the main story in the A-maze task.
-After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, saw an debriefing and were given a code to enter into Mturk for payment.
-The experiment was implemented in Ibex (CITE) and the experimental code is availabe at REPO.
+After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, saw an debriefing and were given a code to enter for payment.
+The experiment was implemented in Ibex (https://github.com/addrummond/ibex) and the experimental code is available at REPO.
 
 ## Data analysis
-Analyses were done in R (CITE), using the following CURATE packages. We used `r cite_r("r-references.bib")` for all our analyses.
+
+```{r}
+package_list <- c("tidyverse","brms","rstan","papaja", "here",
+                  "lme4", "mgcv", "tidybayes", "patchwork","cowplot", "gridExtra")
+```
+
+We conducted data processing and analyses using `r cite_r("r-references.bib", withhold=F, pkgs=package_list)`.
+
+To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.
+For length, we used the length in characters excluding end punctuation.
+For unigram frequency, we tokenized the training data from @gulordava18 and tallied up instances.
+We then used the log2 frequency of the expected occurances in 1 billion words as the model predictor, so higher values indicate higher log frequencies.
+We got per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava], and Transformer-XL [@daiTransformerXLAttentiveLanguage2019a].
+For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
+
+We only included words that were a single token in each of the model vocabularies and for which we had frequency information.
+In practice, this excluded words with punctuation as well as uncommon or proper nouns.
+We also excluded the first word of every sentence (which had a dummy distractor).
+We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
+We exclude words where mistakes occurred or which occurred after a mistake in the same sentence.
+
+For generalized additive models, we centered but did not rescale the length and frequency predictors, but left surprisal uncentered for interpretability.
+We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions.
+
+For linear models, we centered all predictors.
+We used full mixed effects, including by-subject slopes and a per-word random intercept.
+We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations).
+Models were run in BRM.
+
+For model comparison, we fit models with only frequency and length as predictors, as well as models that also had one or more sources of surprisal. We centered all effects.
+
+### TODO SPR!!!
 
@@
 Performance on the comprehension questions.
-Participants who had >80% task accuracy tended to do well on comprehension questions; those who were at chance on the task were also at chance on comprehension questions.
+Participants who had >80% task accuracy tended to do well; those who were at chance on the task were also at chance on the questions.
 
@@
 
-Our first concern was whether this error-correction paradigm on long passages would work.
-It did; some participants were able to complete the Maze task with high accuracy and understand the story while doing so.
+Many participants completed the Maze task with a high degree of accuracy and also answers the comprehension questions correctly.
 
-Participant accuracy reflects how well participants can navigate the task and whether the auto-generated distractors are of acceptable quality.
+Participant accuracy reflects both how well participants can navigate the task and what quality the auto-generated distractors are.
 We calculated the per-word error rate for each participant and graphed it against their average reaction time.
-(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.) As seen in Figure \@ref(fig:errors)A, there is a cluster of participants who make relatively few errors (marked in green), with some reaching 99% accuracy.
-This indicates that the distractors were generally appropriate and that some participants can maintain focus on the task for the whole story.
-These careful participants tend to take around 1 second for each word selection, which is much slower than other paradigms CITE SPR something.
-We also have a cluster of participants (in red) who we infer sped through the task clicking randomly.
-This distribution is likely due to the mix of workers on Mturk; as we did not use qualification cutoffs.
+(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) make relatively few errors, with some reaching 99% accuracy.
+This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story.
+These careful participants took around 1 second for each word selection, which is much slower than other paradigms CITE.
+Another cluster of participants (in red) sped through the task, seemingly clicking randomly.
+This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.
 
 Another check is whether participants comprehended the story.
-We counted how many of the binary-choice comprehension questions each participant got right (out of 6), and group them by their performance on the Maze task.
+We counted how many of the binary-choice comprehension questions each participant got right (out of 6).
 As seen in Figure \@ref(fig:errors)B, most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions.
 Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet.
-These are not especially hard questions, and we can't rule out that some participants may have been able to guess the answers without reading the story.
+We can't rule out that some participants may have been able to guess the answers without understanding the story.
 Nonetheless, this provides preliminary evidence that people can understand and remember details of stories they read during the Maze task.
 
-Given that Maze task accuracy and comprehension question accuracy are correlated, but task accuracy is finer-grained, we use task performance as our exclusion metric.
-This easy to calculate speed and accuracy comparision makes it easy to see who was paying attention to the task; we exclude the random guessers and only analyse data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
+We use task performance as our exclusion metric and only analyze data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
 
@@
 
-### GAMs
-
-We wanted to see if Maze also showed a linear relationship between surprisal and RT, and whether this was localized at the current word or whether previous word surprisals were also influential.
-To do this, we created a set of predictors of frequency, word length, and surprisals from several different langauge models, and fit generalized additive models to check for a linear relationship.
- For length, we use the length of characters (excluding end punctuation).
-For a measure of unigram frequency, we tokenize the training data from @gulordava18 (this tokenizes off punctuation, but preserves case) and tally up instances.
-For the models, we use the log2 frequency of the expected occurances in 1 billion words (thus frequency is log, but higher values indicate higher frequency words).
-For surprisals, we get per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava], and Transformer-XL [@daiTransformerXLAttentiveLanguage2019a].
-We centered but did not rescale the length and frequency predictors; for interpretability, we kept surprisal uncentered.
-For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
-CITE some papers that did the same thing .
-We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions.
-
-We only include words that were in the vocabularies of all three models as a single token and for which we have frequency information.
-This has the effect of excluding words that had punctuation as well as some uncommon or proper nouns.
-We also exclude the first word of every sentence (which had a dummy distractor).
-We do not analyse RTs for words where the RT was <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
-We also omit words where the participant made a mistake and had made a mistake earlier in the sentence (Although see supplement for results including post-mistake TODO).
+We fitted generalized additive models to test whether the RTs from the Maze experiment showed a linear relationship with surprisal and whether the effect was limited the current word or had spilled over from the prior word.
+For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.
 
@@
 plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))
+ggsave("sample_plot.pdf")
 
@@
 The smooths for the current and previous words surprisals are shown in Figure \@ref(fig:gam).
-Note that for each of the models, high-surprisal words at relatively rare, with much of the data for words between 0 and 15 bits of surprisal.
+Note that for each of the models, high-surprisal words are rare, with much of the data for words between 0 and 15 bits of surprisal.
 All of the models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data.
@@
 
-Given the linear relationship, we next fit linear models to look at coefficients and model fit. We fit BRMS models.
+### TODO comparison with SPR !!!!
 
-### BRM models
-
-We built models using BRMS, with surprisal, frequency, length and surprisal x length and frequency x length effects.
-Predictors were as described above, except all were centered.
-We used full mixed effects, including by-subject effects of everything and a per-word random intercept.
-We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations).
-TODO CITE also maybe this is too details-y methods.
+Given that the GAM models show a roughly linear relationship, we fit mixed linear models to quantify the influence of surprisal, frequency and word length. 
+We built linear models with surprisal, frequency, length and surprisal x length and frequency x length effects from the current and previous words as predictors.
 
@@
 As we can see in Table \@ref(tab:pre-error), we find large effects of surprisal and length, but minimal effects of frequency.
-These effects are larger than what is usually reported in other methods CITE, but this could be due to the overall slowness of the method; the intercepts are also much bigger.
-The lack of frequency effects is somewhat surprising, and some CITE argue that they aren't real anyway.
-Notably the coefficients for the lagged terms are smaller than surprisal and length of the current word.
+These effects are larger than what is usually reported in other methods CITE, but this could be due to the overall slowness of the method.
+The lack of frequency effects is somewhat surprising, but consistent with CITE shain.
+Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word.
 
-### Model comparison
-
-We also ran a nested model comparison with models fit in lmer.
-For this, we fit models with only frequency and length as predictors, as well as models that also had one or more sources of surprisal.
-For all effects, we included both current and lagged versions of it.
-The model with all 3 surprisal predictors is better than any model with only one.
-Any one surprisal predictor is better than no surprisals predictors.
-However, adding ngram predictors to a model that already has txl & grnn does not help.
-In other cases, adding the 3rd surprisal source to the other two does help.
-Ngram+Grnn is not better than Grnn only.
-Otherwise, pairs are better than singletons.
-This suggests that Ngram's info is a subset of GRNN, but not a subset of TXL.
+As a last analysis, we checked which of our surprisal models had the best fit using a nested model comparison. 
+We found that all surprisal sources provide predictive value over none, but that the information provided by the Ngram model does not provide additional value to a model that already has GRNN in it.
+TXL and GRNN appear to contain some complementary predictive value.
 
@@
 ```{r, include=F}
-
 anova(all_surp, txl_only)
@@
 
-<!-- ## Comparision with SPR
-We want to compare how good our participants were on the comprehension questions to how good people were on the SPR experiment reported in Futrell et al (2017).
-They had each story read about 100 times across 181 participants (participants generally read 5 stories).
-This means they have about 10 times as much data as we do.
-However, reading times in a story were excluded based on the comprehension of that story, so it makes sense to compare per-story accuracies with our per-story accuracies.
--->
+# Discussion
 
-<!-- ## Error patterns
-When participants do make errors, they correct them quickly, generally within one second, which means it's probably not disrupting their reading much.
-
-While we are planning on analysing data after mistakes, it might be interesting to know how much of the data is pre-mistake, and how many of the sentences are completed error-free.
-If we only exclude the errors and the first words of sentences, there are `r nrow(data_non_error)` words.
-If we also exclude all words after mistakes, there are `r nrow(data_before)` words.
-We will exclude only a couple of words after each mistake, but we could also opt to analyse only the pre-error sections.
-Of the `r nrow(data_sentence)` sentences that good participants completed, `r filter(data_sentence, sent.correct==1) %>% nrow()` were completed entirely correctly.
--->
-
-# Discussion
+Between auto-generating Maze distractors and the error-correction paradigm, Maze is a good complement to existing incremental processing methods.
+It localizes effects better that SPR does while showing similar patterns in terms of surprisal.
+In additon, despite the oddness of the task participants can succeed at it while understanding what they are reading.
 [summarize results] 
@@
 
-We've shown another way to adapt the Maze task to make it suitable for a wider range of experiments.
-While we think that error-correction Maze is generally a good idea, especially when using A-maze, this is an orthagonal adjustment to the task.
-By using the Natural Stories corpus, we establish that Maze works for longer naturalistic tasks, with participants able to do the task and understand what they are reading.
-Additionally, their RTs reflect expected patterns in terms of length and surprisal.
  While there are some differences, especially in scale, to that found with other incremental processing paradigms, we think this is a reason to explore more.
-Could be a statistical thing (???) or due to task differences.
 Comparisons between methods could be very useful for identifying how task demands influence processing.
@@
 All the code is available, and we encourage researchers to consider trying out Maze if they think it's appropriate to their experiments.
-
-While Maze has traditionally been limited to single sentence items; we innovate on how the materials are presented and show that it is possible for participants to read naturalistic passages that are presented in the Maze paradigm.
-We confirm that participants can understand what they are reading, and that Maze shows linear surprisal effects.
 This opens up another area of research to being amenable to the new A-maze paradigm.
@@
 While eventually we do want to figure out models of more natural reading, this might be a useful complement b/c we don't have to deal with nasty spillover.
-Might be extra good to do with pauses for mistakes to fix the incentives.
 
@@
 This seems like a promising improvement that could be worth implementing and testing.]
+
 \newpage

### 2021-06-17 Papers/Paper/manuscript.Rmd
more writing

@@
  library(gridExtra)
+library(broom.mixed)
+library(tidymv)
+
 r_refs("r-references.bib", append=F)
@@
 Theoretically, participants must fully integrate each word into the sentence in order to confidently select it.
-This idea is supported by studies finding strongly localized effects [@witzel12].
+This idea is supported by studies finding strongly localized effects [@witzelComparisonsOnlineReading2012a].
 
@@
 @boyceMazeMadeEasy2020 demonstrate a way to automatically generate Maze distractors by using NLP language models to find words that are high-surprisal in the context of the target sentence.
-The quality of these A-Maze distractors is not up to that of hand-generated distractors, but @boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzel12.
+The quality of these A-Maze distractors is not up to that of hand-generated distractors, but @boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzelComparisonsOnlineReading2012a.
 A-maze out performed a SPR control in detecting and localizing expected slowdown effects.
@@
 Using this corpus on the A-maze task allows us to address, first whether participants will read and understand long passages using A-maze with error correction, and second, whether the resulting RTs profiles will show expected patterns such as a linear relationship with surprisal.
+
+In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran.
+They recruited 181 participants, most of whom read 5 o the stories for a total of 988 story reads, or close to a million RT measures.
+After reading each story, each participant answered 6 binary-choice comprehension questions.
 
@@
 ## Participants
-We recruited 100 participants from Amazon Mechanical Turk, and paid each participant 3.50 dollars, for roughly 20 minutes of work.
+We recruited 100 participants from Amazon Mechanical Turk, and paid each participant $3.50 for roughly 20 minutes of work.
 We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
@@
 ```{r}
-package_list <- c("tidyverse","brms","rstan","papaja", "here",
-                  "lme4", "mgcv", "tidybayes", "patchwork","cowplot", "gridExtra")
+package_list <- c("tidyverse","brms","rstan","papaja",
+                  "lme4", "mgcv", "tidybayes")
 ```
@@
 
+### Model specification
 To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.
@@
 We then used the log2 frequency of the expected occurances in 1 billion words as the model predictor, so higher values indicate higher log frequencies.
-We got per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava], and Transformer-XL [@daiTransformerXLAttentiveLanguage2019a].
+We got per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], and Transformer-XL [@daiTransformerXLAttentiveLanguage2019].
 For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
@@
 For linear models, we centered all predictors.
-We used full mixed effects, including by-subject slopes and a per-word random intercept.
+We used full mixed effects, including by-subject slopes and a per-word-token random intercept.
 We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations).
@@
 
-For model comparison, we fit models with only frequency and length as predictors, as well as models that also had one or more sources of surprisal. We centered all effects.
+For model comparison, we fit models with only frequency and length as predictors, as well as models that also had one or more sources of surprisal.
+We centered all effects.
+We included a random intercept for word identity, but no further random effects so the models would fit in LME4.
 
-### TODO SPR!!!
+### Self-paced reading comparison
+
+As a comparison we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020].
+Data exclusions are the same as for A-maze; we do not make any exclusions based on participants comprehension accuracy.
+Generalized additive models are run with the same structure and predictors as for A-maze.
+Due to the size of the data, we were unable to run the full Bayesian mixed model; instead we estimate effect sizes uses a model with only random intercept for word identity fit in lmer, like that used for model comparison.
+We also run model comparison models similar to those described above.
 
@@
 
-### TODO comparison with SPR !!!!
-
 Given that the GAM models show a roughly linear relationship, we fit mixed linear models to quantify the influence of surprisal, frequency and word length. 
@@
 These effects are larger than what is usually reported in other methods CITE, but this could be due to the overall slowness of the method.
-The lack of frequency effects is somewhat surprising, but consistent with CITE shain.
+The lack of frequency effects is somewhat surprising, but consistent with @shainLargescaleStudyEffects2019.
 Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word.
@@
 
+## Comparison with SPR
+
+```{r read_data}
+d.raw.a <- here("Data/SPR/batch1_pro.csv") %>% 
+  read_csv()
+
+d.raw.b <- here("Data/SPR/batch2_pro.csv") %>% 
+  read_csv()
+
+offset=230 # following note in repo, we have to correct alignment on item 3 b/c an empty token was displayed
+
+d.raw <- d.raw.a %>% union(d.raw.b) %>% 
+  mutate(zone=if_else(item == 3 & zone > offset, zone - 3, zone - 2))
+
+# spot check that this matches number wise with their processed words
+
+```
+
+```{r surprisals}
+
+tokenization <- here("Data/SPR/all_stories.tok.txt") %>% 
+  read_delim(delim="\t") %>% 
+  mutate(Word_In_Story_Num=zone,
+         Story_Num=item) %>%
+  rename(Word=word)
+
+#something is going on is story 2, sentence 4 His brother had blatantly peeked..." 
+#they have it as "peaked", not sure what was displayed, but we can inner join
+surprisals <- here("Data/SPR/natural_stories_surprisals2.rds") %>% 
+  read_rds() %>% 
+  inner_join(tokenization, by=c("Word_In_Story_Num", "Story_Num", "Word")) %>% 
+  filter(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1) %>% 
+  filter(!is.na(txl_surp) & !is.na(ngram_surp) & !is.na(grnn_surp) & !is.na(freq) & !is.na(length)) %>% 
+    mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
+         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
+         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
+         freq_center=freq-mean(freq, na.rm=T),
+         length_center=length-mean(length, na.rm=T))
+
+past_word <- surprisals %>% 
+  mutate(Word_In_Sentence_Num=Word_In_Sentence_Num+1) %>% 
+  rename(past_txl=txl_surp, past_ngram=ngram_surp, past_grnn=grnn_surp, past_freq=freq, past_length=length,past_c_txl=txl_center, past_c_ngram=ngram_center, past_c_grnn=grnn_center, past_c_freq=freq_center, past_c_length=length_center) %>%
+  select(Story_Num,Sentence_Num, Word_In_Sentence_Num, starts_with("past")) %>% 
+  inner_join(surprisals, by=c("Story_Num","Sentence_Num", "Word_In_Sentence_Num"))
+
+```
+
+```{r filter}
+
+d.filter <- d.raw %>% 
+  inner_join(past_word, by=c("zone", "item")) %>% 
+  filter(RT>100 & RT<5000) %>% 
+  filter(Word_In_Sentence_Num!=0)
+
+#subset for testing
+d.subset <- d.filter #%>% filter(Story_Num==1)
+```
+
+```{r pre_gam}
+ngram_data <- d.subset %>% select(rt=RT, surprisal=ngram_surp, prev_surp=past_ngram, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            ) %>% mutate(model="5-gram")
+
+grnn_data <- d.subset %>% select(rt=RT, surprisal=grnn_surp, prev_surp=past_grnn, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            )%>% mutate(model="GRNN")
+
+txl_data <- d.subset %>% select(rt=RT, surprisal=txl_surp, prev_surp=past_txl, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            ) %>% mutate(model="TXL")
+
+all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% 
+  select(surprisal, prev_surp,model) %>% 
+  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
+  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
+
+```
+
+```{r gams, eval=F}
+gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")
+
+gam_grnn_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_data, method="REML")
+
+gam_txl_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_data, method="REML")
+
+```
+
+```{r predict-gam, eval=F}
+
+a <- get_gam_predictions(model=gam_ngram_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
+b <- get_gam_predictions(model=gam_ngram_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
+c <- get_gam_predictions(model=gam_grnn_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
+d <- get_gam_predictions(model=gam_grnn_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
+e <- get_gam_predictions(model=gam_txl_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
+f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
+all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% write_rds("SPR_gam_predictions.rds")
+
+```
+
+```{r prep-brms}
+d.for.lm <- d.subset %>%
+  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
+  mutate(subject=factor(WorkerId),
+         rt=RT) %>% write_rds("for_brms.rds")
+```
+
+```{r}
+#mean by subject for each word
+d.byword <- d.for.lm %>% rename(word=Word) %>% 
+  group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
+                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
+  summarize(mean_rt=mean(rt))
+```
+
+```{r}
+
+ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+```
+
+(ref:spr-gam-cap) GAM predictions of reaction time (RT) for SPR data from @futrellNaturalStoriesCorpus2020 as a function of either current word surprisal (top) or previous word surprisal (bottom).
+Density of data is shown along the x-axis.
+
+```{r spr-gam, out.height="40%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:spr-gam-cap)"}
+all <- read_rds(here("Analysis/SPR_gam_predictions.rds"))
+
+gam1 <-  ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+  geom_line()+
+  geom_ribbon(alpha=.3)+
+  facet_grid(s~model)+
+  coord_cartesian(ylim=c(300,400), xlim=c(0,28))+
+  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+
+  theme(axis.ticks.x=element_blank(), 
+        axis.title.x=element_blank(), 
+        axis.text.x=element_blank(), 
+        plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
+
+dens2 <-   ggplot(all_data, aes(x=value))+
+  geom_density(fill="gray",)+
+  facet_grid(.~model)+
+  labs(x="Surprisal (bits)", y="")+
+  theme(strip.text.x=element_blank(),
+        axis.text.y=element_blank(),
+        panel.grid=element_blank(),
+        axis.ticks.y=element_blank(), 
+        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+p2 = align_plots(gam1, dens2, align = "v", axis="lr")
+plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))
+
+```
+
+```{r spr-table}
+ngram_only_tidy <- tidy(ngram_only, effects="fixed") %>% mutate(type="5-gram") %>% 
+  mutate(term=str_replace(term, "ngram", "surp"))
+
+txl_only_tidy <- tidy(txl_only, effects="fixed") %>% mutate(type="TXL") %>% 
+   mutate(term=str_replace(term, "txl", "surp"))
+
+grnn_only_tidy <- tidy(grnn_only, effects="fixed") %>% mutate(type="GRNN") %>% 
+   mutate(term=str_replace(term, "grnn", "surp"))
+
+all_tidy <- ngram_only_tidy %>% union(txl_only_tidy) %>% union(grnn_only_tidy) %>% 
+  select(Term=term, estimate, std.error, type) %>% 
+  mutate(Ci_high=estimate+1.96*std.error,
+         Ci_low=estimate-1.96*std.error,
+         Range=str_c(round(estimate,1)," [",round(Ci_low, 1),", ",round(Ci_high,1),"]")) %>% 
+  select(Term, type, Range) %>% 
+  pivot_wider(id_cols=Term, names_from=type, values_from=Range) %>% 
+    mutate(Term=factor(Term, 
+                        levels=c("(Intercept)", 
+                                 "surp_center",
+                                 "length_center",
+                                 "freq_center",
+                                 "freq_center:length_center",
+                                 "past_c_surp",
+                                 "past_c_length",
+                                 "past_c_freq",
+                                 "past_c_freq:past_c_length")
+                                 )) %>% 
+  arrange(Term) %>% 
+  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency",
+                   "Freq x Length", "Past Surprisal",
+                  "Past Length", "Past Freq", "Past Freq x Length"))
+
+knitr::kable(all_tidy, format="latex", position="t",caption="Predictions from fitted linear regression models.
+All terms were centered, but not rescaled.
+Units are in ms.
+Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")
+```
+
+As a comparison, we ran the same types of models on the Self-Paced Reading data collected by @futrellNaturalStoriesCorpus2020.
+As shown in Figure \@ref(fig:spr-gam), there is a roughly linear but fairly flat relationship between RT and surprisal.
+Note that the y-axis is fairly narrow, and so the predicted effect of changes in surprisal is fairly small.
+This is confirmed by linear models (see Table \@ref(tab:spr-table)); note that due to the size of the data, these models were fit differently than the A-maze models due to computational constraints.
+Surprisal and length effects are evident for both the current and past word, but we do not see frequency effects.
+The effects are much smaller than for A-maze, although some of this may be due to noise from spillover from prior words or differences in model structure.
+
+Lastly, we are interested to know which of the models is better at predicting human behavior data, and whether they are complementary.
+I fit a series of models including the frequency and length predictors, and some number of past or present surprisal sources.
+This allows us to tell whether different models of surprisal are "picking up" on different things that affect reading time.
+
+We ran nested model comparison on the SPR data as well, and found that all sources of surprisal have predictive value, but the GRNN model provides the most benefit.
+Adding other models to GRNN is marginally beneficial, but TXL and N-gram together are more predictive than either alone.
+
+```{r, eval=F}
+
+no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+all_surp <- lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   grnn_center + past_c_grnn +
+                   txl_center + past_c_txl +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+ngram_grnn <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   grnn_center + past_c_grnn +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+ngram_txl <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   txl_center + past_c_txl +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+grnn_txl <-  lmer(mean_rt ~ grnn_center + past_c_grnn +
+                   txl_center + past_c_txl +
+                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+no_lag <-  lmer(mean_rt ~ ngram_center +
+                   grnn_center +
+                   txl_center + 
+                   freq_center*length_center+(1|word), data=d.byword)
+
+no_surp_lag <-  lmer(mean_rt ~ ngram_center +  
+                   grnn_center + 
+                   txl_center +                    freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+
+  only_surp_lag <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
+                   grnn_center + past_c_grnn +
+                   txl_center + past_c_txl +
+                   freq_center*length_center+(1|word), data=d.byword)
+```
+
+```{r, eval=F}
+anova(txl_only, no_surp)
+anova(grnn_only, no_surp)
+anova(ngram_only, no_surp)
+
+anova(all_surp, ngram_grnn)
+anova(all_surp, ngram_txl)
+anova(all_surp, grnn_txl)
+
+anova(ngram_grnn, ngram_only)
+anova(ngram_grnn, grnn_only)
+anova(ngram_txl, ngram_only)
+anova(ngram_txl, txl_only)
+anova(grnn_txl, grnn_only)
+anova(grnn_txl, txl_only)
+```
+
 # Discussion
 
-Between auto-generating Maze distractors and the error-correction paradigm, Maze is a good complement to existing incremental processing methods.
-It localizes effects better that SPR does while showing similar patterns in terms of surprisal.
-In additon, despite the oddness of the task participants can succeed at it while understanding what they are reading.
-[summarize results] 
-- error-correction paradigm
-- test on natural stories
-- it works
+We introduced a tweak to the paradigm for displaying Maze tasks that makes it useable for multi-sentence materials.
+In this error-correction Maze paradigm, participants read all the words because they can correct their mistakes and move on.
+We tested this method on the Natural Stories corpus, showing that, despite the oddness of the task, participants can read and understand a 1000 word story in this method.
+We additionally showed that the RTs generated in the Maze task show a linear relationship between the RT of a word and it's surprisal, but no relationship with the surprisal of the previous word.
+This provides additional evidence for the argument that the Maze task forces very incremental processing [@forsterMazeTaskMeasuring2009].
 
- While there are some differences, especially in scale, to that found with other incremental processing paradigms, we think this is a reason to explore more.
-Comparisons between methods could be very useful for identifying how task demands influence processing.
+This extreme incrementality may make the Maze task a good target for comparisons between human RTs and predictions of surprisal from neural language models because spillover effects would not need to be considered as much.
+It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more.
+Comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing.
 
-All the code is available, and we encourage researchers to consider trying out Maze if they think it's appropriate to their experiments.
-This opens up another area of research to being amenable to the new A-maze paradigm.
-
-The Maze task might be espeically good for comparing with NNs because of the forced incrementality.
-While eventually we do want to figure out models of more natural reading, this might be a useful complement b/c we don't have to deal with nasty spillover.
-
-Secondly, this error-correction compensates for some of the shortcomings of A-Maze identified in @boyceMazeMadeEasy2020; distractors might still cause participants to make unavoidable mistakes, but at least the still see the rest of the sentence and we get their data.
+While the error-correction paradigm is crucial to running long materials, it also provides some benefits even on shorter materials when using A-maze with variable participant populations.
+The error-correction compensates for some of the shortcomings of A-Maze identified in @boyceMazeMadeEasy2020; poor distractors might still cause participants to make unavoidable mistakes, but they will still see the rest of the sentence, which may reduce frustration.
 Whether this also solves the data loss problem from the researcher perspective within a sentence depends on whether post-mistake data are high-quality and trustworthy; this is a hard-to-assess question of potential interest.
 
-Even if researchers exclude all post-mistake data from analysis, this process can still address the question of whether errors are due to inattentive participants or bad distractors (as discussed in @boyceMazeMadeEasy2020, section XX).
-When we have a participants selection for every word, we can easy calculate a per-word error rate for each participant, without having to address the censoring present in error-terminating Maze (where we can't know if a participant would have made more mistakes after the first; with error correction, we know how may more they did make).
-If participants are guessing, we'll see high per-word error rates, if the early distractors are bad, we'll see low per-word error rates (with errors concentrated at the start of the sentences).
-This per-word error rate metric measures participants task accuracy, and allows us to exclude data from inattentive participants.
-This provides a convenient and clear-cut way of controlling data quality after the fact.
+Even if post-mistake data is not analysed, it can be used to distinguish between errors due to inattentive participants or merely specific bad distractors early in the sentence.
+Researchers can calculate a per-word error rate for each participant; high per-word error rates are consistent with guessing, low error rates with errors clustered early the sentence are consistent with poor distractors.
+This per-word error rate metric measures participants task accuracy and provides a convenient and clear-cut way of controlling data quality after the fact.
 
-It also starts to reduce the perverse incentives.
-With error-terminates Maze, the fastest way to get through the task is to select randomly, and it's quite quick because mistakes skip you to the next sentence.
-With error-correction Maze, randomly jamming buttons takes more effort, but is still an effective strategy.
-[Footnote: In discussing this work, we recieved the suggestion that one way to disincentivize random clicking is to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms or 1 sec) before being able to correct their mistake.
-This seems like a promising improvement that could be worth implementing and testing.]
+Error-correction Maze also starts to reduce perverse incentives from the desire to complete the task quickly.
+With traditional Maze, clicking randomly will likely lead to a mistake, which will cause a participant to skip ahead to the next sentence.
+With error-correction Maze, randomly jamming buttons takes more effort, but is still faster than doing the task.
+In discussing this work, we received the suggestion that one way to disincentivize random clicking is to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms or 1 sec) before being able to correct their mistake.
+This seems like a promising improvement that could be worth implementing and testing.
 
+Between the distractor auto-generation process introduced in @boyceMazeMadeEasy2020 and the error-correction paradigm introduced here, the Maze paradigm is now easy to use on a wide range of materials.
+The ease of use, large effects, and forced incrementality make Maze a good complement to existing incremental processing methods.
+We encourage researchers to consider Maze as an option for doing incremental processing work.
 \newpage

### 2021-06-18 Papers/Paper/manuscript.Rmd
add some citations

@@
 
-keywords          : "TODO"
-wordcount         : "X"
+keywords: TODO
+wordcount: TODO
 
@@
 ```{r setup, include = FALSE}
+knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
  library("papaja")
@@
 Measures of online reading are one way to understand language and how the mind processes language.
-Many theories of language structure and language processing ground out in predictions about the difficulty of processing words.
-For instance, the subject v object relative debate includes theories that make fine-grained predictions about which word is how slow -- this needs localized methods to adjudicate it.
-Other theories such as noisy channel processing also need support from localized word-by-word results.
-TODO WHY *DO* we care about localization?
+Many theories of language structure and language processing ground out in predictions about the difficulty of processing words [@bartekSearchOnlineLocality2011].
+For instance, the subject v object relative debate includes theories that make fine-grained predictions about which word is how slow -- this needs localized methods to adjudicate it [@grodnerConsequencesSerialNature2005; @staubEyeMovementsProcessing2010; @traxlerProcessingSubjectObject2002].
+Other theories such as noisy channel processing also need support from localized word-by-word results [@levyIntegratingSurprisalUncertaininput2011].
 
 Incremental processing methods such as self-paced reading or eye-tracking measure how long someone spends looking at one word before moving on and use that as a proxy for how difficult word was in context.
-Across multiple methods for measuring per-word processing and reading time, how unexpected a word is correlates with how long it takes to read it and move on (CITE).
+How unexpected a word is correlates with how long it takes to read it and move on [@raynerEffectsFrequencyPredictability2004; @klieglLengthFrequencyPredictability2004].
 However, the two major methods of measuring incremental processing both suffer from imprecise localization.
-In eye-tracking, people read naturally which involves skipping words, jumping ahead and looking back, the dynamics of which make it hard to isolate effects.
-Even when reading order is controlled in self-paced reading, readers are moving quickly through the text and it may take multiple words for slowdowns to catch up with them.
+In eye-tracking, people read naturally which involves skipping words, jumping ahead and looking back, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982; @raynerEyeMovementsReading1998].
+Even when reading order is controlled in self-paced reading, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006].
 
@@
 
-It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability).
+It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural].
 Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization.
-In addition to suprisal predicting RT, word length and word's overall frequency are also often found to be predictive.
-TODO many CITES
+In addition to suprisal predicting RT, word length and word's overall frequency are also often found to be predictive [@klieglLengthFrequencyPredictability2004].
 
-An alternative method that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing [@forsterMazeTaskMeasuring2009].
+An alternative method that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing [@freedman85; @forsterMazeTaskMeasuring2009].
 In the Maze task, participants see two words at a time, a correct word that continues the sentence, and a distractor which does not.
@@
 ### Model specification
+
 To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.
@@
 For linear models, we centered all predictors.
-We used full mixed effects, including by-subject slopes and a per-word-token random intercept.
+We used full mixed effects, including by-subject slopes and a per-word-token random intercept [@barrRandomEffectsStructure2013].
 We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations).
-Models were run in BRM.
+Models were run in brm [@burknerAdvancedBayesianMultilevel2018].
 
@@
 We centered all effects.
-We included a random intercept for word identity, but no further random effects so the models would fit in LME4.
+We included a random intercept for word identity, but no further random effects so the models would fit in lme4 [@batesFittingLinearMixedeffects2015].
 
@@
     scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
-  guides(color=FALSE)
+  guides(color="none")
 
@@
   scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
-guides(fill=FALSE)
+guides(fill="none")
 
@@
 This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story.
-These careful participants took around 1 second for each word selection, which is much slower than other paradigms CITE.
+These careful participants took around 1 second for each word selection, which is much slower than other paradigms.
 Another cluster of participants (in red) sped through the task, seemingly clicking randomly.
@@
 
-This extreme incrementality may make the Maze task a good target for comparisons between human RTs and predictions of surprisal from neural language models because spillover effects would not need to be considered as much.
+This extreme incrementality may make the Maze task a good target for comparisons between human RTs and predictions of surprisal from neural language models because spillover effects would not need to be considered as much [@wilcoxPredictivePowerNeural].
 It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more.
-Comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing.
+Comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex.
+@bartekSearchOnlineLocality2011].
 

### 2021-06-23 Papers/Paper/manuscript.Rmd
work on manuscript

@@
 This is an indication of spillover from the previous word.
-
 It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural].
@@
 A-maze out performed a SPR control in detecting and localizing expected slowdown effects.
-@sloggettAmazeAnyOther also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.
+@sloggettAmazeAnyOther2020 also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.
 
@@
 This makes it hard to run long materials using Maze, and prevents Maze from being used on long text passages where SPR or eye-tracking could be used.
+We resolve this issue by introducing a new paradigm for running Maze tasks where participants correct their errors and continue reading rather than discontinuing after a mistake.
 
-We address this problem by creating a Maze task interface where participants correct their mistakes and continue with the sentence rather than terminating on mistakes.
-This allows for multi-sentence items to be run using Maze, which we verify by running Maze on the Natural Stories corpus.
-With this data, we are able to confirm that participants can comprehend what they read using Maze and investigate the effects of surprisal on RT in Maze.
+Using this tweak, we test A-maze on the passages from the Natural Stories corpus.
+The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers.
+At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions.
+Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants.
+If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on basically any naturalistic text.
+The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
+@futrellNaturalStoriesCorpus2020 collected self-paced reading time data on this corpus, which we compare the A-maze results to.
+Participants were able to read and understand these long passages using A-maze with the error correction paradigm, and their RT profiles showed a linear relationship to the surprisal of the words.
 
@@
 This variant of Maze expands the types of materials that can be used with maze to include arbitrarily long passages and cushions the impact of occasional problematic distractors.
-
-# Natural Stories corpus
-
-We test this error-correction Maze on the Natural Stories corpus.
-The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers.
-At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions.
-Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants.
-If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on basically any naturalistic text.
-
-The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
-Using this corpus on the A-maze task allows us to address, first whether participants will read and understand long passages using A-maze with error correction, and second, whether the resulting RTs profiles will show expected patterns such as a linear relationship with surprisal.
-
-In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran.
-They recruited 181 participants, most of whom read 5 o the stories for a total of 988 story reads, or close to a million RT measures.
-After reading each story, each participant answered 6 binary-choice comprehension questions.
 
@@
 
-## Data analysis
+## Models
 
@@
 We conducted data processing and analyses using `r cite_r("r-references.bib", withhold=F, pkgs=package_list)`.
-
-### Model specification
 
@@
 
-### Self-paced reading comparison
+## Self-paced reading comparison
 
-As a comparison we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020].
+In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran.
+They recruited 181 participants, most of whom read 5 of the stories for a total of 988 story reads, or close to a million RT measures.
+After reading each story, each participant answered 6 binary-choice comprehension questions.
+As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020].
 Data exclusions are the same as for A-maze; we do not make any exclusions based on participants comprehension accuracy.

### 2021-06-23 Papers/Paper/manuscript.Rmd
files to make the manuscript happy

@@
 
-```{r pre-error}
+```{r pre-error-prep, eval=F}
 brm_txl_interact <- read_rds(paste0(here(),"/Analysis/brm_txl_interact.rds"))
@@
 
-summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
+summ <- a %>% union(b) %>% union(c) %>% write_rds(here("Analysis/brms_summary.rds"))
+
+```
+
+```{r pre-error}
+summ <- read_rds(here("Analysis/brms_summary.rds")) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
   mutate(Term=factor(Term, 
@@
 
-```{r}
+```{r,include=F, eval=F}
 d_lm <- labelled_pre_error %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
@@
 
-```{r, include=F}
+```{r, include=F, eval=F}
 anova(all_surp, txl_only)
@@
 
-```{r read_data}
+```{r read_data, eval=F}
 d.raw.a <- here("Data/SPR/batch1_pro.csv") %>% 
@@
 
-```{r surprisals}
+```{r surprisals, eval=F}
 
@@
 
-```{r filter}
+```{r filter, eval=F}
 
@@
 
-```{r pre_gam}
+```{r pre_gam, eval=F}
 ngram_data <- d.subset %>% select(rt=RT, surprisal=ngram_surp, prev_surp=past_ngram, 
@@
 
-```{r prep-brms}
+```{r prep-brms, eval=F}
 d.for.lm <- d.subset %>%
@@
   mutate(subject=factor(WorkerId),
-         rt=RT) %>% write_rds("for_brms.rds")
+         rt=RT) 
 ```
 
-```{r}
+```{r, eval=F}
 #mean by subject for each word
@@
 
-```{r}
+```{r, eval=F}
 
@@
 ```
-
-```{r spr-table}
+```{r spr-prep, eval=F}
 ngram_only_tidy <- tidy(ngram_only, effects="fixed") %>% mutate(type="5-gram") %>% 
@@
 
-all_tidy <- ngram_only_tidy %>% union(txl_only_tidy) %>% union(grnn_only_tidy) %>% 
+all_tidy <- ngram_only_tidy %>% union(txl_only_tidy) %>% union(grnn_only_tidy) %>% write_rds(here("Analysis/SPR_lmer_summ.rds"))
+```
+```{r spr-table}
+
+all_tidy <- read_rds(here("Analysis/SPR_lmer_summ.rds")) %>% 
   select(Term=term, estimate, std.error, type) %>% 

### 2021-07-01 Analysis/nat_stories_gam.Rmd
edits to paper
assorted changes working on gpt and gams

--- 2020-07-21 Analysis/nat_stories_gam.Rmd
@@
 title: "Natural Stories GAM analysis"
-output: github_document
+output: html_document
 ---
@@
 
-data_post_only <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
-  filter(!after_mistake==0) %>% 
-    select(subject, word_num, word, rt, sentence, type)
-
 data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
@@
 ```{r labels}
-labs <- read_rds("../Prep_code/natural_stories_surprisals.rds") %>% left_join(read_delim("../Materials/natural_stories_sentences.tsv", delim="\t")) %>% 
+labs <- read_rds("../Prep_code/natural_stories_surprisals.rds") %>% 
+  left_join(read_delim("../Materials/natural_stories_sentences.tsv", delim="\t")) %>% 
   select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())
@@
 select_labs <- labs %>% 
-  filter(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1) %>% 
-  filter(!is.na(txl_surp) & !is.na(ngram_surp) & !is.na(grnn_surp) & !is.na(freq) & !is.na(length)) %>% 
+  mutate(across(ends_with("surp"),~ifelse(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1 & gpt_token_count==1 & word_num>0, .x, NA))) %>% 
+  select(-ends_with("token_count")) %>% 
     mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
@@
          freq_center=freq-mean(freq, na.rm=T),
-         length_center=length-mean(length, na.rm=T))
+         length_center=length-mean(length, na.rm=T),
+         gpt_center=gpt_surp-mean(gpt_surp, na.rm=T)) %>% 
+  group_by(sentence,Story_Num,Sentence_Num) %>% 
+  mutate(across(txl_surp:gpt_center,.names="past_{.col}", lag),
+         across(txl_surp:gpt_center,.names="past2_{.col}", ~lag(.x,n=2)),
+         across(txl_surp:gpt_center,.names="past3_{.col}", ~lag(.x,n=3)))
 
-past_word <- select_labs %>% 
-  mutate(word_num=word_num+1) %>% 
-  rename(past_txl=txl_surp, past_ngram=ngram_surp, past_grnn=grnn_surp, past_freq=freq, past_length=length,past_c_txl=txl_center, past_c_ngram=ngram_center, past_c_grnn=grnn_center, past_c_freq=freq_center, past_c_length=length_center) %>%
-  select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
-  left_join(select_labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))
+labelled_pre_error <- data_pre_error %>% inner_join(select_labs, by=c("word_num", "word", "sentence")) %>%
+  filter(word_num>1) %>% 
+  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_")))%>% write_rds("pre_error.rds")
 
-labelled_pre_error <- data_pre_error %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
-  filter(word_num>1) %>% 
-  select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
-         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
-  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
-  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("pre_error.rds")
+```
 
-labelled_post_error <- data_ready %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
-  filter(word_num>1) %>% 
-select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
-         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
-  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
-  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("post_error.rds")
+```{r item-means}
 
-labelled_post_only <- data_post_only %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
-  filter(word_num>1) %>% 
-select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
-         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
-  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
-  select(-Word_In_Story_Num, -Story_Num)
+select_data <- labelled_pre_error %>%   select(-starts_with("past2"), -starts_with("past3")) %>% 
+  filter(across(everything(),~!is.na(.x)))
+
+item_mean_data <- select_data %>% 
+  group_by(across(c(-subject, -rt))) %>% 
+  summarize(rt=mean(rt)) %>% 
+  ungroup()
+
 ```
@@
 
-## Try 1
-Using only pre-error data.
-
-Does not have hierarchical effects. Has smooths for the surprisals, and tensor effects/interactions for the freq x length
-
 ```{r for gam}
 
-ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram, 
+ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            prev_freq=past_freq_center, prev_len=past_length_center
+                                            ) %>% mutate(model="5-gram") 
+grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center
+                                            )%>% mutate(model="GRNN")
+
+txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center
+                                            ) %>% mutate(model="TXL")
+
+gpt_data <- labelled_pre_error %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center
+                                            ) %>% mutate(model="GPT-2")
+
+ngram_mean_data <- item_mean_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center
                                             ) %>% mutate(model="5-gram")
 
-grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn, 
+grnn_mean_data <- item_mean_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            prev_freq=past_freq_center, prev_len=past_length_center
                                             )%>% mutate(model="GRNN")
 
-txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_surp, prev_surp=past_txl, 
+txl_mean_data <- item_mean_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            prev_freq=past_freq_center, prev_len=past_length_center
                                             ) %>% mutate(model="TXL")
 
-all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% 
+gpt_mean_data <- item_mean_data %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center
+                                            ) %>% mutate(model="GPT-2")
+
+all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
+  select(surprisal, prev_surp,model) %>% 
+  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
+  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
+
+all_mean_data <- ngram_mean_data %>% union(grnn_mean_data) %>% union(txl_mean_data) %>% union(gpt_mean_data) %>% 
   select(surprisal, prev_surp,model) %>% 
@@
 
-```{r gams}
-gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")
+```{r formulae, echo=T}
+#no hierarchical, has freq x len interaction
+formula_interact <- rt ~ s(surprisal, bs="cr", k=20)+
+                     ti(freq, bs="cr") + 
+                     ti(len, bs="cr")+
+                     ti(freq,len, bs="cr")+
+                     s(prev_surp, bs="cr", k=20)+
+                     ti(prev_freq, bs="cr")+
+                     ti(prev_len, bs="cr")+
+                     ti(prev_freq, prev_len, bs="cr")
 
-gam_grnn_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_data, method="REML")
+# no hierarchical, NO freq x len interaction
+formula_no_interact <- rt ~ s(surprisal, bs="cr", k=20)+
+                     ti(freq, bs="cr") + 
+                     ti(len, bs="cr")+
+                     s(prev_surp, bs="cr", k=20)+
+                     ti(prev_freq, bs="cr")+
+                     ti(prev_len, bs="cr")
+```
 
-gam_txl_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_data, method="REML")
+All of this is on by-item mean data.
+
+```{r meaned data}
+gam_ngram_1 <- gam(formula_interact, data=ngram_mean_data, method="REML")
+gam_grnn_1 <- gam(formula_interact, data=grnn_mean_data, method="REML")
+gam_txl_1 <- gam(formula_interact, data=txl_mean_data, method="REML")
+gam_gpt_1 <- gam(formula_interact, data=gpt_mean_data, method="REML")
+
+gam_ngram_2 <- gam(formula_no_interact, data=ngram_mean_data, method="REML")
+gam_grnn_2 <- gam(formula_no_interact, data=grnn_mean_data, method="REML")
+gam_txl_2 <- gam(formula_no_interact, data=txl_mean_data, method="REML")
+gam_gpt_2 <- gam(formula_no_interact, data=gpt_mean_data, method="REML")
 
@@
 
-```{r plot-gam}
+```{r anovas}
+
+anova(gam_ngram_1,gam_ngram_2)
+anova(gam_grnn_1,gam_grnn_2)
+anova(gam_txl_1,gam_txl_2)
+anova(gam_gpt_1,gam_gpt_2)
+```
+
+I don't know how to interpret the above.
+
+I also don't know how to interpret the plots.
+
+```{r plots}
+plot(gam_gpt_1)
+```
+
+```{r plot-gam, eval=F}
 
@@
 
-```{r}
-all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f)
+```{r, eval=F}
+all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% write_rds("gam_predictions.rds")
 

### 2021-07-01 Papers/Paper/manuscript.Rmd
assorted changes working on gpt and gams

@@
 #they have it as "peaked", not sure what was displayed, but we can inner join
-surprisals <- here("Data/SPR/natural_stories_surprisals2.rds") %>% 
+surprisals <- here("Data/SPR/natural_stories_surprisals.rds") %>% 
   read_rds() %>% 

### 2021-07-01 Prep_code/nat_stories_prep.Rmd
assorted changes working on gpt and gams

--- 2020-04-24 Prep_code/nat_stories_prep.Rmd
@@
 library(readr)
+library(here)
+
+prep_dir <- "Prep_code"
+materials <- "Materials"
 ```
@@
    write_delim("ns_pre_maze.txt", delim=";", col_names=F, quote_escape = F)
+```
+
+# Surprisals
+For GPT (and some others), run get_surps.sh.
+```{r}
+# For models
+sentences <- read_delim(here(materials,"natural_stories_sentences.tsv"),delim="\t") %>% 
+  pull(Sentence) %>% 
+  write_lines(here(prep_dir,"sentences.txt"))
+
+words <-labelled %>% write_tsv(here(prep_dir,"words.txt"))
 
@@
 # Parsing surprisals, etc
+
+```{r gpt}
+
+gpt_translator <- read_tsv(here(prep_dir,"gpt_surp_trans.txt"), col_names=c("Story_Num","Sentence_Num","Word_In_Sentence_Num","Word_In_Story_Num","Word","Token", "Surprisal")) %>% 
+  group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
+  summarize(gpt_surp=sum(Surprisal),
+            gpt_token_count=n())
+```
 
@@
 ```{r conglomerate}
+
 all <- txl_translator %>% 
@@
   left_join(grnn_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>%
+  left_join(gpt_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
   left_join(freq, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
-  filter(Word_In_Sentence_Num!=0) %>% 
   ungroup() %>% 
   write_rds("natural_stories_surprisals.rds")
-
 ```
-
-Note that word 0's of sentences will be removed b/c we definitely don't trust them

### 2021-07-01 Prep_code/wrangle_gpt.py

@@ -1,0 +1,41 @@
+import csv
+
+def do_thing(surpfile,wordfile,outfile):
+    words=[]
+    story_num=[]
+    sentence_num=[]
+    word_in_sentence_num=[]
+    word_in_story_num=[]
+    with open(wordfile, 'r') as f:
+        reader = csv.reader(f, delimiter="\t")
+        next(reader,None)
+        for row in reader:
+            story_num.append(row[0])
+            sentence_num.append(row[1])
+            word_in_sentence_num.append(row[2])
+            words.append(row[3])
+            word_in_story_num.append(row[4])
+    with open(surpfile, 'r', newline="") as f:
+        reader = csv.reader(f, delimiter="\t")
+        with open(outfile, 'w') as out:
+                writer = csv.writer(out, delimiter="\t")
+                i=0
+                word=words[i]
+                token_str=""
+                next(reader,None)
+                for row in reader:
+                    if(token_str==word):
+                        i+=1
+                        word=words[i]
+                        token_str=""
+                    if(len(token_str)>len(word)):
+                        print(token_str)
+                        print(word)
+                        raise AssertionError
+                    token=row[2].replace("Ġ","")
+                    token_str+=token
+                    writer.writerow([story_num[i],sentence_num[i],word_in_sentence_num[i],word_in_story_num[i],word, token,row[3]])
+
+
+do_thing("gpt_surp.txt","words.txt","gpt_surp_trans.txt")
+

### 2021-07-02 Analysis/nat_stories.Rmd
start repo cleaning

@@
 library(cowplot)
+library(here)
 theme_set(theme_bw())
@@
 options(mc.cores = parallel::detectCores())
+
+data_dir <- "Data"
 ```
@@
 
-data <- read_rds("../Data/cleaned.rds")
+data <- read_rds(here(data_dir,"cleaned.rds"))
 
@@
 
-ggsave("../Papers/error.pdf", width=3, height=2.1, unit="in")
+#ggsave("../Papers/error.pdf", width=3, height=2.1, unit="in")
 
@@
 
-ggsave("../error.jpg", width=5, height=3, unit="in")
+#ggsave("../error.jpg", width=5, height=3, unit="in")
 
@@
 
-ggsave("../Papers/comp.pdf", width=3, height=1.5, unit="in")
+#ggsave("../Papers/comp.pdf", width=3, height=1.5, unit="in")
 
@@
 
-spr <- read_csv("../ns_spr_1.csv") %>% 
-  union(read_csv("../ns_spr_2.csv")) %>% 
-  select(WorkerId, correct, item) %>% 
-  unique() %>% 
+d.raw.a <- here("Data/SPR/batch1_pro.csv") %>% 
+  read_csv()
+
+d.raw.b <- here("Data/SPR/batch2_pro.csv") %>% 
+  read_csv()
+
+offset=230 # following note in repo, we have to correct alignment on item 3 b/c an empty token was displayed
+
+d.raw <- d.raw.a %>% union(d.raw.b) %>% 
+  mutate(zone=if_else(item == 3 & zone > offset, zone - 3, zone - 2)) %>% 
+  select(WorkerId,correct,item) %>% unique()
+
+# spot check that this matches number wise with their processed words
+
+all_spr <- d.raw %>% group_by(correct) %>% 
+  tally() %>% 
+  mutate(pct=n/sum(n)) %>% 
+  mutate(source="SPR") %>% 
+  select(source,correct,pct)
+
+first <- d.raw %>% 
+  group_by(WorkerId,correct) %>% 
+  mutate(num=row_number()) %>% 
+  filter(num==1) %>% 
+  select(-num) %>% 
   group_by(correct) %>% 
   tally() %>% 
-  mutate(pct=n/988) %>% 
-  mutate(source="SPR") %>% 
+  mutate(pct=n/sum(n)) %>% 
+  mutate(source="SPR 1st") %>% 
   select(source,correct,pct)
@@
   tally() %>% 
-  mutate(pct=n/95) %>% 
+  mutate(pct=n/sum(n)) %>% 
   mutate(source="all Maze") %>% 
@@
   tally() %>% 
-  mutate(pct=n/63) %>% 
+  mutate(pct=n/sum(n)) %>% 
   mutate(source="Maze >80% correct") %>% 
@@
 
-to_plot <- all %>% union(some) %>% union(spr)
+to_plot <- all %>% union(some) %>% union(all_spr) %>% union(first)
 

### 2021-07-02 Papers/Paper/manuscript.Rmd
start repo cleaning

@@
 ```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:diagram-cap)"}
-knitr::include_graphics(paste0(here(),"/Papers/maze_diagram_cropped.pdf"))
+knitr::include_graphics(here("Papers/maze_diagram_cropped.pdf"))
 ```
@@
 
-data <- read_rds(paste0(here(),"/Data/cleaned.rds"))
+data <- read_rds(here("Data/cleaned.rds"))
 
@@
 ```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:gam-cap)"}
-labelled_pre_error <- read_rds(paste0(here(),"/Analysis/pre_error.rds"))
+labelled_pre_error <- read_rds(here("Data/maze_pre_error.rds"))
 
-ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram, 
+ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                             ) %>% mutate(model="5-gram") 
-
-grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn, 
+grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                             )%>% mutate(model="GRNN")
 
-txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_surp, prev_surp=past_txl, 
+txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
+                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
                                             ) %>% mutate(model="TXL")
@@
 
-all <-read_rds(paste0(here(),"/Analysis/gam_predictions.rds"))
+all <-read_rds(here("Analysis/gam_predictions.rds"))
 gam1 <-  ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
@@
 
-p2 = align_plots(gam1, dens2, align = "v", axis="lr")
+p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
 plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))

### 2021-07-02 Prep_code/nat_stories_prep.Rmd
start repo cleaning

--- 2020-04-24 Prep_code/nat_stories_prep.Rmd
@@
 library(readr)
+library(here)
+
+prep_dir <- "Prep_code"
+materials <- "Materials"
+predictors <- "Prep_code/predictors"
 ```
@@
 
-words <- read_delim("natural_stories_words.tsv",delim="\t", col_names = c("key","value")) %>% 
+words <- read_delim(here(prep_dir,"natural_stories_words.tsv"),delim="\t", col_names = c("key","value")) %>% 
   separate(key, into=c("Story_Num", "Word_In_Story_Num", "col3"), convert=T) %>% 
@@
 
-labelled <- read_delim("../Materials/natural_stories_sentences.tsv", delim="\t") %>% 
+labelled <- read_delim(here(materials,"natural_stories_sentences.tsv"), delim="\t") %>% 
   separate(Sentence, into=nums, sep=" ") %>% 
@@
 # takes sentences, writes into format to do maze (also other things)
-for_maze <- read_delim("../Materials/natural_stories_sentences.tsv", delim="\t") %>%
+for_maze <- read_delim(here(materials,"natural_stories_sentences.tsv"), delim="\t") %>%
   mutate(label=str_c(Story_Num, Sentence_Num, sep="_"), item=row_number()) %>% 
   select(label, item, Sentence) %>% 
-   write_delim("ns_pre_maze.txt", delim=";", col_names=F, quote_escape = F)
+   write_delim(here(prep_dir,"ns_pre_maze.txt"), delim=";", col_names=F, quote_escape = F)
+```
+
+```{r}
+# For gpt model models
+sentences <- read_delim(here(materials,"natural_stories_sentences.tsv"),delim="\t") %>% 
+  pull(Sentence) %>% 
+  write_lines(here(prep_dir,"sentences.txt"))
+
+words <-labelled %>% write_tsv(here(prep_dir,"words.txt"))
 
@@
 
+```{r gpt}
+
+gpt_translator <- read_tsv(here(predictors,"gpt_surp_trans.txt"), col_names=c("Story_Num","Sentence_Num","Word_In_Sentence_Num","Word_In_Story_Num","Word","Token", "Surprisal")) %>% 
+  group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
+  summarize(gpt_surp=sum(Surprisal),
+            gpt_token_count=n())
+```
+
 ```{r grnn}
 
-grnn <- read_delim("grnn_out_all.txt", delim="\t", col_names=c("model_token","grnn_surp")) %>% 
+grnn <- read_delim(here(predictors,"grnn_out_all.txt"), delim="\t", col_names=c("model_token","grnn_surp")) %>% 
   filter(model_token!="<eos>") %>% 
@@
 
-grnn_translator <- read_delim("ns_tokens_trans.txt", delim="\t", col_names=c("Item","Word_In_Sentence_Num", "Word","token"), col_types=c(col_character(),col_integer(), col_character(),col_character())) %>%
+grnn_translator <- read_delim(here(predictors,"ns_tokens_trans.txt"), delim="\t", col_names=c("Item","Word_In_Sentence_Num", "Word","token"), col_types=c(col_character(),col_integer(), col_character(),col_character())) %>%
   mutate(num=row_number()) %>% 
@@
 ```{r ngram}
-ngram <- read_delim("ngram_out_all.txt", skip=3,delim="\t", col_names=c("sentence_no","token_no","model_token","ngram_surp")) %>% 
+ngram <- read_delim(here(predictors,"ngram_out_all.txt"), skip=3,delim="\t", col_names=c("sentence_no","token_no","model_token","ngram_surp")) %>% 
   mutate(num=row_number())
 
-ngram_translator <- read_delim("ns_lower_trans.txt", delim="\t", col_names=c("Item","Word_In_Sentence_Num", "Word","token"), col_types=c(col_character(),col_integer(), col_character(),col_character())) %>%
+ngram_translator <- read_delim(here(predictors,"ns_lower_trans.txt"), delim="\t", col_names=c("Item","Word_In_Sentence_Num", "Word","token"), col_types=c(col_character(),col_integer(), col_character(),col_character())) %>%
   mutate(num=row_number()) %>% 
@@
 ```{r txl}
-txl <- read_delim("txl_out_all.txt", delim="\t", col_names=c("model_token","txl_surp")) %>% 
+txl <- read_delim(here(predictors,"txl_out_all.txt"), delim="\t", col_names=c("model_token","txl_surp")) %>% 
   filter(model_token!="<eos>") %>% 
@@
 
-txl_translator <- read_delim("ns_tokens_moses_trans.txt", delim="\t", 
+txl_translator <- read_delim(here(predictors,"ns_tokens_moses_trans.txt"), delim="\t", 
                              col_names=c("Item","Word_In_Sentence_Num", "Word","token"), 
@@
 
-freq <- read_delim("ns_freq.txt", delim="\t",
+freq <- read_delim(here(predictors,"ns_freq.txt"), delim="\t",
                    col_names=c("Item","Word_In_Sentence_Num","Word","freq","length"), 
@@
   left_join(grnn_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>%
+  left_join(gpt_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
   left_join(freq, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
-  filter(Word_In_Sentence_Num!=0) %>% 
   ungroup() %>% 
-  write_rds("natural_stories_surprisals.rds")
-
+  write_rds(here(prep_dir,"natural_stories_surprisals.rds"))
 ```
-
-Note that word 0's of sentences will be removed b/c we definitely don't trust them

### 2021-07-02 Prep_code/make_unigrams.py

@@ -1,20 +1,0 @@
-import nltk
-import math
-import json
-from collections import Counter
-
-with open('gulo_train.txt', 'r') as file:
-    data = file.read()
-
-tokens = nltk.word_tokenize(data)
-length = len(tokens)
-print(length)
-
-c=Counter(tokens)
-
-logprobs={}
-for t,count in c.items():
-    logprobs[t]=math.log2(count)-math.log2(length)+math.log2(10**9)
-
-with open('freqs.txt','w') as f:
-    json.dump(logprobs, f)

### 2021-07-02 Prep_code/useful.py

@@ -109,8 +109,45 @@ def tokenize_file(infile, outfile, translate_file, lower=False, eos=True, moses=
                     else:
                         out.write(sent + "\n")
 
+def wrangle_gpt(surpfile,wordfile,outfile):
+    words=[]
+    story_num=[]
+    sentence_num=[]
+    word_in_sentence_num=[]
+    word_in_story_num=[]
+    with open(wordfile, 'r') as f:
+        reader = csv.reader(f, delimiter="\t")
+        next(reader,None)
+        for row in reader:
+            story_num.append(row[0])
+            sentence_num.append(row[1])
+            word_in_sentence_num.append(row[2])
+            words.append(row[3])
+            word_in_story_num.append(row[4])
+    with open(surpfile, 'r', newline="") as f:
+        reader = csv.reader(f, delimiter="\t")
+        with open(outfile, 'w') as out:
+                writer = csv.writer(out, delimiter="\t")
+                i=0
+                word=words[i]
+                token_str=""
+                next(reader,None)
+                for row in reader:
+                    if(token_str==word):
+                        i+=1
+                        word=words[i]
+                        token_str=""
+                    if(len(token_str)>len(word)):
+                        print(token_str)
+                        print(word)
+                        raise AssertionError
+                    token=row[2].replace("Ġ","")
+                    token_str+=token
+                    writer.writerow([story_num[i],sentence_num[i],word_in_sentence_num[i],word_in_story_num[i],word, token,row[3]])
 
-tokenize_file("ns_pre_maze.txt", "ns_lower_tokens.txt", "ns_lower_trans.txt", lower=True, eos=False)
-tokenize_file("ns_pre_maze.txt", "ns_tokens.txt", "ns_tokens_trans.txt")
-tokenize_file("ns_pre_maze.txt", "ns_tokens_moses.txt","ns_tokens_moses_trans.txt", moses=True)
-get_freq("ns_pre_maze.txt","ns_freq.txt")
+
+wrangle_gpt("predictors/gpt_surp.txt","words.txt","predictors/gpt_surp_trans.txt")
+tokenize_file("ns_pre_maze.txt", "predictors/ns_lower_tokens.txt", "ns_lower_trans.txt", lower=True, eos=False)
+tokenize_file("ns_pre_maze.txt", "predictors/ns_tokens.txt", "ns_tokens_trans.txt")
[…]

### 2021-07-02 Prep_code/wrangle_gpt.py

@@ -1,41 +1,0 @@
-import csv
-
-def do_thing(surpfile,wordfile,outfile):
-    words=[]
-    story_num=[]
-    sentence_num=[]
-    word_in_sentence_num=[]
-    word_in_story_num=[]
-    with open(wordfile, 'r') as f:
-        reader = csv.reader(f, delimiter="\t")
-        next(reader,None)
-        for row in reader:
-            story_num.append(row[0])
-            sentence_num.append(row[1])
-            word_in_sentence_num.append(row[2])
-            words.append(row[3])
-            word_in_story_num.append(row[4])
-    with open(surpfile, 'r', newline="") as f:
-        reader = csv.reader(f, delimiter="\t")
-        with open(outfile, 'w') as out:
-                writer = csv.writer(out, delimiter="\t")
-                i=0
-                word=words[i]
-                token_str=""
-                next(reader,None)
-                for row in reader:
-                    if(token_str==word):
-                        i+=1
-                        word=words[i]
-                        token_str=""
-                    if(len(token_str)>len(word)):
-                        print(token_str)
-                        print(word)
-                        raise AssertionError
-                    token=row[2].replace("Ġ","")
-                    token_str+=token
-                    writer.writerow([story_num[i],sentence_num[i],word_in_sentence_num[i],word_in_story_num[i],word, token,row[3]])
-
-
-do_thing("gpt_surp.txt","words.txt","gpt_surp_trans.txt")
-

### 2021-07-02 Analysis/models.Rmd
more renaming and moving

--- 2021-07-01 Analysis/nat_stories_gam.Rmd
@@
 ---
-title: "Natural Stories GAM analysis"
+title: "Natural Stories model analyses"
 output: html_document
@@
 library(scales)
+library(here)
 theme_set(theme_bw())
@@
 
-data <- read_rds("../Data/cleaned.rds")
+data <- read_rds(here("Data/cleaned.rds"))
 
@@
 
-# Overview
+<!--# Overview
 
@@
 (Not sure this last piece is actually true/matters).
+-->
 
+# Maze Prep
 ```{r labels}
-labs <- read_rds("../Prep_code/natural_stories_surprisals.rds") %>% 
-  left_join(read_delim("../Materials/natural_stories_sentences.tsv", delim="\t")) %>% 
+labs <- read_rds(here("Prep_code/natural_stories_surprisals.rds")) %>% 
+  left_join(read_delim(here("Materials/natural_stories_sentences.tsv"), delim="\t")) %>% 
   select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())
@@
 
+```
+
+```{r maze}
+
 labelled_pre_error <- data_pre_error %>% inner_join(select_labs, by=c("word_num", "word", "sentence")) %>%
   filter(word_num>1) %>% 
-  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_")))%>% write_rds("pre_error.rds")
+  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_")))%>% write_rds(here("Data/maze_pre_error.rds"))
 
@@
 
-select_data <- labelled_pre_error %>%   select(-starts_with("past2"), -starts_with("past3")) %>% 
+select_data <- read_rds(here("Data/maze_pre_error.rds")) %>%  select(-starts_with("past2"), -starts_with("past3")) %>% 
   filter(across(everything(),~!is.na(.x)))
@@
 
-### GAMs
+# Maze GAMs
 
@@
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                             ) %>% mutate(model="5-gram") 
@@
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                             )%>% mutate(model="GRNN")
@@
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center
+                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
                                             ) %>% mutate(model="TXL")
@@
                                             freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center
+                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
                                             ) %>% mutate(model="GPT-2")
@@
                      ti(prev_len, bs="cr")
+
+formula_heirarchy <- rt ~ s(surprisal, bs="cr", k=20)+
+                    s(surprisal, by=subject, bs="cr", k=20)+
+                     ti(freq, bs="cr") + 
+                     ti(len, bs="cr")+
+                     s(prev_surp, bs="cr", k=20)+
+                      s(prev_surp, by=subject, bs="cr", k=20)+
+                     ti(prev_freq, bs="cr")+
+                     ti(prev_len, bs="cr")
+
 ```
@@
 ```
+
+By-factor GAMs fail to run overnight / crash the R session.
 
@@
 ```
+
+# SPR Prep
+```{r read_data, eval=F}
+d.raw.a <- here("Data/SPR/batch1_pro.csv") %>% 
+  read_csv()
+
+d.raw.b <- here("Data/SPR/batch2_pro.csv") %>% 
+  read_csv()
+
+offset=230 # following note in repo, we have to correct alignment on item 3 b/c an empty token was displayed
+
+d.raw <- d.raw.a %>% union(d.raw.b) %>% 
+  mutate(zone=if_else(item == 3 & zone > offset, zone - 3, zone - 2))
+
+# spot check that this matches number wise with their processed words
+
+```
+
+```{r first-story}
+# to make things more comparable, only look at the first story each participant read
+
+first <- d.raw %>% select(WorkerId,correct,item) %>% unique() %>% 
+  group_by(WorkerId,correct) %>% 
+  mutate(num=row_number()) %>% 
+  filter(num==1) %>% 
+  select(-num)
+
+d.limited <- d.raw %>% inner_join(first, by=c("WorkerId","correct","item"))
+
+#d.limited%>% select(WorkerId,correct,item) %>% unique() %>% group_by(correct) %>% tally()  %>% summarize(pct=n/sum(n),correct=correct)
+
+d.good <- d.limited %>% filter(correct>4) %>% rename(rt=RT) %>% filter(rt>100 & rt<5000)
+```
+
+```{r surprisals, eval=F}
+
+tokenization <- here("Data/SPR/all_stories.tok.txt") %>% 
+  read_delim(delim="\t") %>% 
+  mutate(Word_In_Story_Num=zone,
+         Story_Num=item)
+
+#something is going on is story 2, sentence 4 His brother had blatantly peeked..." 
+#they have it as "peaked", not sure what was displayed, but we can inner join
+spr_labelled <- select_labs %>% inner_join(tokenization, by=c("word","Word_In_Story_Num","Story_Num")) %>% 
+  left_join(d.good, by=c("item","zone")) %>% 
+  rename(subject=WorkerId) %>% 
+  select(-WorkTimeInSeconds) %>% 
+  filter(word_num>1) %>% 
+  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_")))%>% write_rds(here("Data/clean_spr.rds"))
+```
+
+```{r spr item-means}
+
+select_spr <- read_rds(here("Data/clean_spr.rds")) %>% 
+  filter(across(everything(),~!is.na(.x))) %>% 
+  ungroup()
+
+item_mean_spr <- select_spr %>% 
+  group_by(across(c(-subject, -rt))) %>% 
+  summarize(rt=mean(rt)) %>% 
+  ungroup()
+
+```
+
+# SPR GAMs
+TODO
+
+# Maze LMs
+
+```{r for brms}
+
+ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_center, prev_surp=past_ngram_center, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
+                                            ) %>% mutate(model="5-gram") 
+grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_center, prev_surp=past_grnn_center, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
+                                            )%>% mutate(model="GRNN")
+
+txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_center, prev_surp=past_txl_center, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
+                                            ) %>% mutate(model="TXL")
+
+gpt_data <- labelled_pre_error %>% select(rt, surprisal=gpt_center, prev_surp=past_gpt_center, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
+                                            ) %>% mutate(model="GPT-2")
+
+all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
+  write_rds("maze_pre_brms.rds")
+
+```
+
+```{r, eval=F}
+
+data=readRDS("maze_pre_brms.rds")
+ priors <- c(
+      set_prior("normal(1000, 1000)", class="Intercept"),
+      set_prior("normal(0, 500)", class="b"),
+      set_prior("normal(0, 500)", class="sd"),
+      set_prior("lkj(1)",       class="cor"))
+run_model <- function(model_spec,data,file_name){
+  d <- data %>% filter(model==model_spec)
+  m1 <- brm(rt ~ surprisal * len + freq * len +
+              prev_surp * prev_len + prev_freq * prev_len+
+              (surprisal * len + freq * len +
+              prev_surp * prev_len + prev_freq * prev_len|subject)+(1|Word_ID),
+            data=d,
+            file=file_name,
+            prior=priors,
+            control=list(adapt_delta=.95))
+}
+
+run_model("5-gram",data,"5-gram_maze")
+run_model("GRNN",data,"grnn_maze")
+run_model("TXL",data,"txl_maze")
+run_model("GPT-2",data,"gpt_maze")
+
+```
+
+# SPR LMs
+
+```{r spr brms}
+
+ngram_data <- select_spr %>% rename_with(~str_replace_all(.x, "ngram", "surp"))%>% 
+  select(rt, surp_center, freq_center, length_center,
+         past_surp_center, past_freq_center, past_length_center,
+         past2_surp_center, past2_freq_center, past2_length_center,
+         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID) %>% mutate(model="5-gram")
+
+grnn_data <- select_spr %>% rename_with(~str_replace_all(.x, "grnn", "surp"))%>% 
+  select(rt, surp_center, freq_center, length_center,
+         past_surp_center, past_freq_center, past_length_center,
+         past2_surp_center, past2_freq_center, past2_length_center,
+         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID)%>% mutate(model="GRNN")
+
+txl_data <- select_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
+  select(rt, surp_center, freq_center, length_center,
+         past_surp_center, past_freq_center, past_length_center,
+         past2_surp_center, past2_freq_center, past2_length_center,
+         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID) %>% mutate(model="TXL")
+
+gpt_data <-select_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
+  select(rt, surp_center, freq_center, length_center,
+         past_surp_center, past_freq_center, past_length_center,
+         past2_surp_center, past2_freq_center, past2_length_center,
+         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID) %>% mutate(model="GPT-2")
+
+all_spr <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
+  write_rds("spr_pre_brms.rds")
+```
+
+```{r spr lms, eval=F}
+
+run_lm <- function(data,modelspec){
+  d <- data %>% filter(model==modelspec)
+model_1 <- lmer(rt ~ surp_center * length_center + length_center * freq_center +
+                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
+                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
+                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
+                  (1|subject)+(1|Word_ID), data=d)
+return(model_1)
+}
+
+m_ngram <- run_lm(all_spr,"5-gram")
+m_grnn <- run_lm(all_spr, "GRNN")
+m_txl <- run_lm(all_spr,"TXL")
+m_gpt <- run_lm(all_spr,"GPT-2")
+
+```
+
+```{r, eval=F}
+
+data=readRDS("spr_pre_brms.rds")
+ priors <- c(
+      set_prior("normal(300, 100)", class="Intercept"),
+      set_prior("normal(0, 5)", class="b"),
+      set_prior("normal(0, 5)", class="sd"),
+      set_prior("lkj(1)",       class="cor"))
+run_model <- function(model_spec,data,file_name){
+  d <- data %>% filter(model==model_spec)
+  m1 <- brm(rt ~ surp_center * length_center + length_center * freq_center +
+                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
+                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
+                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
+              (surp_center * length_center + length_center * freq_center +
+                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
+                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
+                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center |subject)+(1|Word_ID),
+            data=d,
+            file=file_name,
+            prior=priors,
+            control=list(adapt_delta=.95))
+}
+
+run_model("5-gram",data,"5-gram_spr")
+run_model("GRNN",data,"grnn_spr")
+run_model("TXL",data,"txl_spr")
+run_model("GPT-2",data,"gpt_spr")
+
+```

### 2021-07-02 Papers/Amlap_2020_talk/amlap2020_abstract.tex
rearrange files

--- 2020-07-28 Papers/amlap2020_abstract.tex
@@
 We see minimal to no frequency effects (consistent with Shain 2019).
-In a nested model comparison, we found GRNN and TXL encoded complementary information, but adding 5-gram surprisals did not improve a regression that already had GRNN.
+In a nested model comparison, we found the GRNN fit human reading times the best, although TXL seems to encode some complementary information.
 We also fit GAMs to look at the shape of the relationship between surprisal and RT; we find a linear relationship between RT and current word surprisal and no relationship with previous word surprisal (Figure~3).
@@
 Overall, this suggests that A-maze is a versatile alternative to eye-tracking and self-paced reading.
-Code and data at github.com/vboyce/amaze-natural-stories.
+Code and data at \url{github.com/vboyce/amaze-natural-stories}.
 

### 2021-07-05 Analysis/models.Rmd
more gam models

@@
   pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
-  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
+  mutate(s=ifelse(name=="surprisal", "Current","Previous")) %>% write_rds(here("Analysis/models/meaned_maze.rds"))
 ```
@@
 
-```{r meaned data}
+```{r meaned data, eval=F}
 gam_ngram_1 <- gam(formula_interact, data=ngram_mean_data, method="REML")
@@
 
-```{r anovas}
+```{r anovas, eval=F}
 
@@
 
-```{r plots}
+```{r plots, eval=F}
 plot(gam_gpt_1)
@@
 f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
+g <- get_gam_predictions(model=gam_gpt_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Current")
+h <- get_gam_predictions(model=gam_gpt_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Previous")
+all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% union(g) %>% union(h) %>% 
+  mutate(model=factor(model, levels=c("5-gram", "GRNN", "TXL","GPT-2"))) %>% 
+                        write_rds(here("Analysis/models/gam_predictions.rds"))
 ```
@@
 ```{r, eval=F}
-all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% write_rds("gam_predictions.rds")
 
-gam1 <-  ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+gam1 <- read_rds(here("Analysis/models/gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+
   geom_ribbon(alpha=.3)+
-   #geom_density(data=all_data, aes(x=value), fill="gray",)+
+  #facet_grid(s~model, scales="free_y")+
   facet_grid(s~model)+
-  coord_cartesian(ylim=c(700,1300), xlim=c(0,28))+
+  coord_cartesian(xlim=c(0,28))+
   labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
 
-dens2 <-   ggplot(all_data, aes(x=value))+
+dens2 <-   ggplot(all_mean_data, aes(x=value))+
   geom_density(fill="gray",)+
@@
   labs(x="Surprisal (bits)", y="")+
+    coord_cartesian(xlim=c(0,28))+
   theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
 
-bot <- plot_grid(NA, dens2, NA, nrow=1, rel_widths = c(.16, 1, .08))
-
- plot_grid(gam1, bot, nrow=2, rel_heights = c(1, .3))
-
-ggsave("../Papers/gam.pdf", width=4, height=2.3, unit="in")
+p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
+plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))
 
@@
 
-```{r first-story}
+```{r first-story, eval=F}
 # to make things more comparable, only look at the first story each participant read
@@
 # SPR GAMs
-TODO
+
+```{r for spr gam, eval=F}
+
+ngram_mean_spr <- item_mean_spr %>% rename_with(~str_replace_all(.x, "ngram_", ""))%>%
+  select(rt, surp, freq_center, length_center,
+         past_surp, past_freq_center, past_length_center,
+         past2_surp, past2_freq_center, past2_length_center,
+         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="5-gram")
+
+grnn_mean_spr <- item_mean_spr %>% rename_with(~str_replace_all(.x, "grnn_", ""))%>%
+  select(rt, surp, freq_center, length_center,
+         past_surp, past_freq_center, past_length_center,
+         past2_surp, past2_freq_center, past2_length_center,
+         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="GRNN")
+
+txl_mean_spr <- item_mean_spr %>% rename_with(~str_replace_all(.x, "txl_", ""))%>%
+  select(rt, surp, freq_center, length_center,
+         past_surp, past_freq_center, past_length_center,
+         past2_surp, past2_freq_center, past2_length_center,
+         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="TXL")
+
+gpt_mean_spr<-item_mean_spr %>% rename_with(~str_replace_all(.x, "gpt_", ""))%>%
+  select(rt, surp, freq_center, length_center,
+         past_surp, past_freq_center, past_length_center,
+         past2_surp, past2_freq_center, past2_length_center,
+         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="GPT-2")
+
+all_mean_spr <- ngram_mean_spr %>% union(grnn_mean_spr) %>% union(txl_mean_spr) %>% union(gpt_mean_spr) %>% 
+  select(surp, past_surp, past2_surp, past3_surp,model) %>% 
+  pivot_longer(cols=`surp`:`past3_surp`) %>% 
+  mutate(s=ifelse(name=="surprisal", "Current","Previous")) %>% write_rds(here("Analysis/models/meaned_spr.rds"))
+```
+
+```{r formulae, echo=T}
+#no hierarchical, has freq x len interaction
+spr_formula_interact <- rt ~ s(surp, bs="cr", k=20)+
+                     ti(length_center, bs="cr")+
+                     ti(freq_center,length_center, bs="cr")+
+                     s(past_surp, bs="cr", k=20)+
+                     ti(past_freq_center, bs="cr")+
+                     ti(past_length_center, bs="cr")+
+                     ti(past_freq_center, past_length_center, bs="cr")+
+                     s(past2_surp, bs="cr", k=20)+
+                     ti(past2_freq_center, bs="cr")+
+                     ti(past2_length_center, bs="cr")+
+                     ti(past2_freq_center, past2_length_center, bs="cr")+
+                     s(past3_surp, bs="cr", k=20)+
+                     ti(past3_freq_center, bs="cr")+
+                     ti(past3_length_center, bs="cr")+
+                     ti(past3_freq_center, past3_length_center, bs="cr")
+
+```
+
+All of this is on by-item mean data.
+
+```{r spr meaned data, eval=F}
+gam_ngram_1 <- gam(spr_formula_interact, data=ngram_mean_spr, method="REML")
+gam_grnn_1 <- gam(spr_formula_interact, data=grnn_mean_spr, method="REML")
+gam_txl_1 <- gam(spr_formula_interact, data=txl_mean_spr, method="REML")
+gam_gpt_1 <- gam(spr_formula_interact, data=gpt_mean_spr, method="REML")
+
+```
+
+```{r plot-gam-spr, eval=F}
+
+a <- get_gam_predictions(model=gam_ngram_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
+b <- get_gam_predictions(model=gam_ngram_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
+c <- get_gam_predictions(model=gam_ngram_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="2Previous")
+d <- get_gam_predictions(model=gam_ngram_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="3Previous")
+
+e <- get_gam_predictions(model=gam_grnn_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
+f <- get_gam_predictions(model=gam_grnn_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
+g <- get_gam_predictions(model=gam_grnn_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="2Previous")
+h <- get_gam_predictions(model=gam_grnn_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="3Previous")
+
+i <- get_gam_predictions(model=gam_txl_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
+j <- get_gam_predictions(model=gam_txl_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
+k <- get_gam_predictions(model=gam_txl_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="2Previous")
+l <- get_gam_predictions(model=gam_txl_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="3Previous")
+
+m <- get_gam_predictions(model=gam_gpt_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Current")
+n <- get_gam_predictions(model=gam_gpt_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Previous")
+o <- get_gam_predictions(model=gam_gpt_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="2Previous")
+p <- get_gam_predictions(model=gam_gpt_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="3Previous")
+
+all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% union(g) %>% union(h) %>% 
+   union(i) %>% union(j) %>% union(k) %>% union(l) %>% union(m) %>% union(n) %>% union(o) %>% union(p) %>% 
+  mutate(model=factor(model, levels=c("5-gram", "GRNN", "TXL","GPT-2")),
+         s=factor(s, levels=c("Current", "Previous", "2Previous", "3Previous"))) %>% 
+                        write_rds(here("Analysis/models/spr_gam_predictions.rds"))
+```
+
+```{r, eval=T}
+
+gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+  geom_line()+
+  geom_ribbon(alpha=.3)+
+  facet_grid(s~model)+
+  coord_cartesian(xlim=c(0,28))+
+  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
+
+dens2 <-   ggplot(all_mean_spr, aes(x=value))+
+  geom_density(fill="gray",)+
+  facet_grid(.~model)+
+  labs(x="Surprisal (bits)", y="")+
+    coord_cartesian(xlim=c(0,28))+
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
+plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))
+
+```
 
@@
 all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
-  write_rds("maze_pre_brms.rds")
+  write_rds(here("Analysis/models/maze_pre_brms.rds"))
 
@@
 
-```{r, eval=F}
-
+```{r run brms, eval=F}
+# Note this was actually run on a cluster with a separate script that just had this
 data=readRDS("maze_pre_brms.rds")
@@
 run_model("GPT-2",data,"gpt_maze")
+
+```
+
+```{r summary}
+
+show_summary <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=1),
+           upper=round(.upper, digits=1),
+           E=round(.value, digits=1),
+           `Estimate`=str_c(E," [",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           ) %>% 
+    select(Term, `Estimate`)
+
+  stats
+}
+```
+
+```{r process-maze-models, eval=F}
+ngram <- read_rds(here("Analysis/models/5-gram_maze.rds")) %>% show_summary() %>% mutate(model="5-gram")
+txl <- read_rds(here("Analysis/models/txl_maze.rds")) %>% show_summary() %>% mutate(model="TXL")
+grnn <- read_rds(here("Analysis/models/grnn_maze.rds")) %>% show_summary() %>% mutate(model="GRNN")
+gpt <- read_rds(here("Analysis/models/gpt_maze.rds")) %>% show_summary() %>% mutate(model="GPT-2")
+
+summ <- ngram %>% union(txl) %>% union(grnn) %>% union(gpt) %>% write_rds(here("Analysis/models/brms_maze_summ.rds"))
 
@@
 all_spr <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
-  write_rds("spr_pre_brms.rds")
+  write_rds(here("Analysis/models/spr_pre_brms.rds"))
+
+ngram_mean_data <- item_mean_spr %>% rename_with(~str_replace_all(.x, "ngram", "surp"))%>% 
+  select(rt, surp_center, freq_center, length_center,
+         past_surp_center, past_freq_center, past_length_center,
+         past2_surp_center, past2_freq_center, past2_length_center,
+         past3_surp_center, past3_freq_center, past3_length_center, Word_ID) %>% mutate(model="5-gram")
+
+grnn_mean_data <- item_mean_spr %>% rename_with(~str_replace_all(.x, "grnn", "surp"))%>% 
+  select(rt, surp_center, freq_center, length_center,
+         past_surp_center, past_freq_center, past_length_center,
+         past2_surp_center, past2_freq_center, past2_length_center,
+         past3_surp_center, past3_freq_center, past3_length_center, Word_ID)%>% mutate(model="GRNN")
+
+txl_mean_data <- item_mean_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
+  select(rt, surp_center, freq_center, length_center,
+         past_surp_center, past_freq_center, past_length_center,
+         past2_surp_center, past2_freq_center, past2_length_center,
+         past3_surp_center, past3_freq_center, past3_length_center, Word_ID) %>% mutate(model="TXL")
+
+gpt_mean_data <-item_mean_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
+  select(rt, surp_center, freq_center, length_center,
+         past_surp_center, past_freq_center, past_length_center,
+         past2_surp_center, past2_freq_center, past2_length_center,
+         past3_surp_center, past3_freq_center, past3_length_center, Word_ID) %>% mutate(model="GPT-2")
+
+all_mean_spr <- ngram_mean_data %>% union(grnn_mean_data) %>% union(txl_mean_data) %>% union(gpt_mean_data)
+
 ```
 
-```{r spr lms, eval=F}
+```{r spr lms-mean, eval=T}
 
@@
                  past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
-                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
-                  (1|subject)+(1|Word_ID), data=d)
+                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center +(1|Word_ID), data=d)
 return(model_1)
@@
 
-m_ngram <- run_lm(all_spr,"5-gram")
-m_grnn <- run_lm(all_spr, "GRNN")
-m_txl <- run_lm(all_spr,"TXL")
-m_gpt <- run_lm(all_spr,"GPT-2")
+m_ngram <- run_lm(all_mean_spr,"5-gram")
+m_grnn <- run_lm(all_mean_spr, "GRNN")
+m_txl <- run_lm(all_mean_spr,"TXL")
+m_gpt <- run_lm(all_mean_spr,"GPT-2")
 
+```
+
+```{r lmsumm}
+summary(m_ngram)
+summary(m_grnn)
+summary(m_txl)
+summary(m_gpt)
 ```

### 2021-07-05 Analysis/nat_stories.Rmd
more gam models

@@
 
-first <- d.raw %>% 
+first_spr <- d.raw %>% 
   group_by(WorkerId,correct) %>% 
@@
   mutate(pct=n/sum(n)) %>% 
-  mutate(source="SPR 1st") %>% 
+  mutate(source="SPR 1st story") %>% 
   select(source,correct,pct)
 
-all <- data_filt %>% select(subject, num_correct) %>% 
+all_maze <- data_filt %>% select(subject, num_correct) %>% 
   unique() %>% 
@@
 
-some <- data_filt %>% select(subject, num_correct) %>% 
+maze_bad <- data_filt %>% select(subject, num_correct) %>% 
+  unique() %>% 
+  left_join(data_error_summ, by=c("subject", "num_correct")) %>% 
+  filter(pct_correct<.8) %>% 
+  mutate(correct=num_correct) %>% 
+   group_by(correct) %>% 
+  tally() %>% 
+  mutate(pct=n/sum(n)) %>% 
+  mutate(source="<80% correct") %>% 
+  select(source,correct, pct)
+
+maze_good <- data_filt %>% select(subject, num_correct) %>% 
   unique() %>% 
@@
   mutate(pct=n/sum(n)) %>% 
-  mutate(source="Maze >80% correct") %>% 
+  mutate(source=">80% correct") %>% 
   select(source,correct, pct)
 
-to_plot <- all %>% union(some) %>% union(all_spr) %>% union(first)
+to_plot <- all_maze %>% union(maze_bad) %>% union(maze_good)  %>% union(all_spr) %>% union(first_spr) %>% write_rds(here("Analysis/comp.rds"))
 

### 2021-07-05 Papers/Paper/manuscript.Rmd
more gam models

@@
 This is an indication of spillover from the previous word.
-It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural].
+It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020].
 Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization.
@@
 We then used the log2 frequency of the expected occurances in 1 billion words as the model predictor, so higher values indicate higher log frequencies.
-We got per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], and Transformer-XL [@daiTransformerXLAttentiveLanguage2019].
+We got per-word surprisals for each of 4 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].
 For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
 
-We only included words that were a single token in each of the model vocabularies and for which we had frequency information.
-In practice, this excluded words with punctuation as well as uncommon or proper nouns.
-We also excluded the first word of every sentence (which had a dummy distractor).
+We only included words that were treated as a single token by every model and for which we had frequency information.
+This mainly excluded words with punctuation as well as uncommon or proper nouns.
 We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
 We exclude words where mistakes occurred or which occurred after a mistake in the same sentence.
+We only analysed words where we had values for all predictors.
 
@@
 We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions.
+We collapse the reading times across subjects by modelling the mean RT for each word.
 
@@
 We included a random intercept for word identity, but no further random effects so the models would fit in lme4 [@batesFittingLinearMixedeffects2015].
+TODO!!!
 
@@
 In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran.
-They recruited 181 participants, most of whom read 5 of the stories for a total of 988 story reads, or close to a million RT measures.
+They recruited 181 participants, most of whom read 5 of the stories.
 After reading each story, each participant answered 6 binary-choice comprehension questions.
 As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020].
-Data exclusions are the same as for A-maze; we do not make any exclusions based on participants comprehension accuracy.
-Generalized additive models are run with the same structure and predictors as for A-maze.
-Due to the size of the data, we were unable to run the full Bayesian mixed model; instead we estimate effect sizes uses a model with only random intercept for word identity fit in lmer, like that used for model comparison.
-We also run model comparison models similar to those described above.
+For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct.
+TODO model differences!!!
 
@@
 Performance on the comprehension questions.
-Participants who had >80% task accuracy tended to do well; those who were at chance on the task were also at chance on the questions.
+Participants with low accuracy also performed poorly on comprehension questions; Participants with >80% task accuracy tended to do well; their performance was roughly comparable to the performance of SPR participants from  @futrellNaturalStoriesCorpus2020 on their first stories.
 
@@
 
-comp_plot <- ggplot(some, aes(x=num_correct,fill=accurate))+
-  geom_bar(position="dodge")+
-  facet_grid(.~accurate)+
-  labs(x="Comprehension questions correct (out of 6)", y="Participants")+
-  scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
+comp_comp <- read_rds(here("Analysis/comp.rds")) %>% 
+  filter(source %in% c(">80% correct","<80% correct", "SPR 1st story"))
+
+comp_tally <- read_rds(here("Analysis/comp.rds")) %>% group_by(source) %>% filter(correct>4) %>% 
+  summarize(good=round(sum(pct)*100))
+comp_plot <- ggplot(comp_comp, aes(x=correct,y=pct,fill=source))+
+  geom_col(position="dodge")+
+  facet_grid(.~source)+
+  labs(x="Comprehension questions correct (out of 6)", y="Percent participants")+
+  scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred",  "SPR 1st story"="grey30"))+
 guides(fill="none")
@@
 Nonetheless, this provides preliminary evidence that people can understand and remember details of stories they read during the Maze task.
+The comprehension question performance of accurate Maze participants is broadly similar to the performance of SPR participants from @futrellNaturalStoriesCorpus2020 on the first story read.
+Overall, `r comp_tally %>% filter(source=="all Maze") %>% pull(good)`% of Maze participants got 5 or 6 questions right (`r comp_tally %>% filter(source=="<80% correct") %>% pull(good)`% of low-accuracy participants and `r comp_tally %>% filter(source==">80% correct") %>% pull(good)`% of high-accuracy participants) compared to `r comp_tally %>% filter(source=="SPR") %>% pull(good)`% of all SPR reads and `r comp_tally %>% filter(source=="SPR 1st story") %>% pull(good)`% of 1st SPR reads.
+Note that these differences cannot be directly attributed to methods, as the participant populations differed.
 
-We use task performance as our exclusion metric and only analyze data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
+We use task performance as our exclusion metric because it is more fine-grained and only analyze data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
+For the SPR comparison, we follow @futrellNaturalStoriesCorpus2020's criteria and exclude participants who got less than 5 of the questions correct.
 
@@
 Density of data is shown along the x-axis.
-For each of the 3 language models used, there is a linear relationship between current word surprisal and RT (at least when there is enough data).
+For each of the 4 language models used, there is a linear relationship between current word surprisal and RT (at least when there is enough data).
 There is no relationship between previous word surprisal and RT.
+TODO replace graph!
 
@@
 
-all <-read_rds(here("Analysis/gam_predictions.rds"))
-gam1 <-  ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+gam1 <- read_rds(here("Analysis/models/gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+
   geom_ribbon(alpha=.3)+
-   #geom_density(data=all_data, aes(x=value), fill="gray",)+
+  #facet_grid(s~model, scales="free_y")+
   facet_grid(s~model)+
-  coord_cartesian(ylim=c(700,1300), xlim=c(0,28))+
+  coord_cartesian(xlim=c(0,28))+
   labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
@@
   labs(x="Surprisal (bits)", y="")+
-  theme(strip.text.x=element_blank(),axis.text.y=element_blank(), panel.grid=element_blank(), axis.ticks.y=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+    coord_cartesian(xlim=c(0,28))+
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
 
@@
 plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))
-ggsave("sample_plot.pdf")
 
@@
 
-```{r}
-
-show_summary <- function(model){
-  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
-
-  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
-    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
-    group_by(.variable) %>% 
-    summarize(pct_above_0=mean(above_0)) %>% 
-    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
-    left_join(intervals, by=".variable") %>% 
-    mutate(lower=round(.lower, digits=1),
-           upper=round(.upper, digits=1),
-           E=round(.value, digits=1),
-           `Estimate`=str_c(E," [",lower,", ", upper,"]"),
-           Term=str_sub(.variable, 3, -1),
-           ) %>% 
-    select(Term, `Estimate`)
-
-  stats
-}
-```
-
-```{r pre-error-prep, eval=F}
-brm_txl_interact <- read_rds(paste0(here(),"/Analysis/brm_txl_interact.rds"))
-brm_grnn_interact <- read_rds(paste0(here(),"/Analysis/brm_grnn_interact.rds"))
-brm_ngram_interact <- read_rds(paste0(here(),"/Analysis/brm_ngram_interact.rds"))
-a <- show_summary(brm_txl_interact) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
-b <- show_summary(brm_grnn_interact) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
-c <- show_summary(brm_ngram_interact) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))
-
-summ <- a %>% union(b) %>% union(c) %>% write_rds(here("Analysis/brms_summary.rds"))
-
-```
-
 ```{r pre-error}
-summ <- read_rds(here("Analysis/brms_summary.rds")) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
+summ <- read_rds(here("Analysis/models/brms_maze_summ.rds")) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
   mutate(Term=factor(Term, 
                         levels=c("Intercept", 
-                                 "surp_center",
-                                 "length_center",
-                                 "freq_center",
-                                 "surp_center:length_center",
-                                 "length_center:freq_center",
-                                 "past_c_surp",
-                                 "past_c_length",
-                                 "past_c_freq",
-                                 "past_c_surp:past_c_length",
-                                 "past_c_length:past_c_freq")
+                                 "surprisal",
+                                 "len",
+                                 "freq",
+                                 "surprisal:len",
+                                 "len:freq",
+                                 "prev_surp",
+                                 "prev_len",
+                                 "prev_freq",
+                                 "prev_surp:prev_len",
+                                 "prev_len:prev_freq")
                                  )) %>% 
@@
                   "Surp x Length", "Freq x Length", "Past Surprisal",
-                  "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length")) %>% 
-  select(Term, `5-gram`, `GRNN`, TXL)
+                  "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length"))
 
@@
 
-```{r read_data, eval=F}
-d.raw.a <- here("Data/SPR/batch1_pro.csv") %>% 
-  read_csv()
-
-d.raw.b <- here("Data/SPR/batch2_pro.csv") %>% 
-  read_csv()
-
-offset=230 # following note in repo, we have to correct alignment on item 3 b/c an empty token was displayed
-
-d.raw <- d.raw.a %>% union(d.raw.b) %>% 
-  mutate(zone=if_else(item == 3 & zone > offset, zone - 3, zone - 2))
-
-# spot check that this matches number wise with their processed words
-
-```
-
-```{r surprisals, eval=F}
-
-tokenization <- here("Data/SPR/all_stories.tok.txt") %>% 
-  read_delim(delim="\t") %>% 
-  mutate(Word_In_Story_Num=zone,
-         Story_Num=item) %>%
-  rename(Word=word)
-
-#something is going on is story 2, sentence 4 His brother had blatantly peeked..." 
-#they have it as "peaked", not sure what was displayed, but we can inner join
-surprisals <- here("Data/SPR/natural_stories_surprisals.rds") %>% 
-  read_rds() %>% 
-  inner_join(tokenization, by=c("Word_In_Story_Num", "Story_Num", "Word")) %>% 
-  filter(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1) %>% 
-  filter(!is.na(txl_surp) & !is.na(ngram_surp) & !is.na(grnn_surp) & !is.na(freq) & !is.na(length)) %>% 
-    mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
-         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
-         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
-         freq_center=freq-mean(freq, na.rm=T),
-         length_center=length-mean(length, na.rm=T))
-
-past_word <- surprisals %>% 
-  mutate(Word_In_Sentence_Num=Word_In_Sentence_Num+1) %>% 
-  rename(past_txl=txl_surp, past_ngram=ngram_surp, past_grnn=grnn_surp, past_freq=freq, past_length=length,past_c_txl=txl_center, past_c_ngram=ngram_center, past_c_grnn=grnn_center, past_c_freq=freq_center, past_c_length=length_center) %>%
-  select(Story_Num,Sentence_Num, Word_In_Sentence_Num, starts_with("past")) %>% 
-  inner_join(surprisals, by=c("Story_Num","Sentence_Num", "Word_In_Sentence_Num"))
-
-```
-
-```{r filter, eval=F}
-
-d.filter <- d.raw %>% 
-  inner_join(past_word, by=c("zone", "item")) %>% 
-  filter(RT>100 & RT<5000) %>% 
-  filter(Word_In_Sentence_Num!=0)
-
-#subset for testing
-d.subset <- d.filter #%>% filter(Story_Num==1)
-```
-
-```{r pre_gam, eval=F}
-ngram_data <- d.subset %>% select(rt=RT, surprisal=ngram_surp, prev_surp=past_ngram, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
-                                            ) %>% mutate(model="5-gram")
-
-grnn_data <- d.subset %>% select(rt=RT, surprisal=grnn_surp, prev_surp=past_grnn, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
-                                            )%>% mutate(model="GRNN")
-
-txl_data <- d.subset %>% select(rt=RT, surprisal=txl_surp, prev_surp=past_txl, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_c_freq, prev_len=past_c_length
-                                            ) %>% mutate(model="TXL")
-
-all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% 
-  select(surprisal, prev_surp,model) %>% 
-  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
-  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
-
-```
-
-```{r gams, eval=F}
-gam_ngram_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=ngram_data, method="REML")
-
-gam_grnn_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=grnn_data, method="REML")
-
-gam_txl_1 <- gam(rt ~ s(surprisal, bs="cr", k=20)+ te(freq, len, bs="cr")+s(prev_surp, bs="cr", k=20)+te(prev_freq, prev_len, bs="cr"), data=txl_data, method="REML")
-
-```
-
-```{r predict-gam, eval=F}
-
-a <- get_gam_predictions(model=gam_ngram_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
-b <- get_gam_predictions(model=gam_ngram_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
-c <- get_gam_predictions(model=gam_grnn_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
-d <- get_gam_predictions(model=gam_grnn_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
-e <- get_gam_predictions(model=gam_txl_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
-f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
-all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% write_rds("SPR_gam_predictions.rds")
-
-```
-
-```{r prep-brms, eval=F}
-d.for.lm <- d.subset %>%
-  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
-  mutate(subject=factor(WorkerId),
-         rt=RT) 
-```
-
-```{r, eval=F}
-#mean by subject for each word
-d.byword <- d.for.lm %>% rename(word=Word) %>% 
-  group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
-                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
-  summarize(mean_rt=mean(rt))
-```
-
-```{r, eval=F}
-
-ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-```
-
 (ref:spr-gam-cap) GAM predictions of reaction time (RT) for SPR data from @futrellNaturalStoriesCorpus2020 as a function of either current word surprisal (top) or previous word surprisal (bottom).
 Density of data is shown along the x-axis.
+TODO replace graph!
 
-```{r spr-gam, out.height="40%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:spr-gam-cap)"}
-all <- read_rds(here("Analysis/SPR_gam_predictions.rds"))
-
-gam1 <-  ggplot(all, aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos='t', fig.cap="(ref:spr-gam-cap)"}
+gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+
@@
   facet_grid(s~model)+
-  coord_cartesian(ylim=c(300,400), xlim=c(0,28))+
-  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+
-  theme(axis.ticks.x=element_blank(), 
-        axis.title.x=element_blank(), 
-        axis.text.x=element_blank(), 
-        plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
+  coord_cartesian(xlim=c(0,28))+
+  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
 
-dens2 <-   ggplot(all_data, aes(x=value))+
+ all_mean_spr <-  read_rds(here("Analysis/models/meaned_spr.rds"))
+
+dens2 <-   ggplot(all_mean_spr, aes(x=value))+
   geom_density(fill="gray",)+
@@
   labs(x="Surprisal (bits)", y="")+
-  theme(strip.text.x=element_blank(),
-        axis.text.y=element_blank(),
-        panel.grid=element_blank(),
-        axis.ticks.y=element_blank(), 
-        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+    coord_cartesian(xlim=c(0,28))+
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
 
@@
 ```
-```{r spr-prep, eval=F}
-ngram_only_tidy <- tidy(ngram_only, effects="fixed") %>% mutate(type="5-gram") %>% 
-  mutate(term=str_replace(term, "ngram", "surp"))
 
-txl_only_tidy <- tidy(txl_only, effects="fixed") %>% mutate(type="TXL") %>% 
-   mutate(term=str_replace(term, "txl", "surp"))
-
-grnn_only_tidy <- tidy(grnn_only, effects="fixed") %>% mutate(type="GRNN") %>% 
-   mutate(term=str_replace(term, "grnn", "surp"))
-
-all_tidy <- ngram_only_tidy %>% union(txl_only_tidy) %>% union(grnn_only_tidy) %>% write_rds(here("Analysis/SPR_lmer_summ.rds"))
-```
-```{r spr-table}
+```{r spr-table, eval=F}
 
@@
 
-This extreme incrementality may make the Maze task a good target for comparisons between human RTs and predictions of surprisal from neural language models because spillover effects would not need to be considered as much [@wilcoxPredictivePowerNeural].
+This extreme incrementality may make the Maze task a good target for comparisons between human RTs and predictions of surprisal from neural language models because spillover effects would not need to be considered as much [@wilcoxPredictivePowerNeural2020].
 It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more.

### 2021-07-14 Analysis/models.Rmd
work on nested model comparisons

@@
 ---
+
+# Prep
+
 ```{r, include=F}
-knitr::opts_chunk$set(echo = FALSE, warning=F)
+knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
 options(knitr.table.format = "html")
+knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
 library(tidyverse)
@@
 
-ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
+ngram_data <- select_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                             freq=freq_center, len=length_center,
@@
                                             ) %>% mutate(model="5-gram") 
-grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
+grnn_data <- select_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
                                             freq=freq_center, len=length_center,
@@
 
-txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
+txl_data <- select_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
                                             freq=freq_center, len=length_center,
@@
 
-gpt_data <- labelled_pre_error %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
+gpt_data <- select_data %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
                                             freq=freq_center, len=length_center,
@@
 
-```{r meaned data, eval=F}
+```{r meaned data, eval=T, cache=T}
 gam_ngram_1 <- gam(formula_interact, data=ngram_mean_data, method="REML")
@@
 
-```{r anovas, eval=F}
+```{r anovas, eval=T}
 
-anova(gam_ngram_1,gam_ngram_2)
-anova(gam_grnn_1,gam_grnn_2)
-anova(gam_txl_1,gam_txl_2)
-anova(gam_gpt_1,gam_gpt_2)
+anova(gam_ngram_1,gam_ngram_2, test="Chisq")
+anova(gam_grnn_1,gam_grnn_2, test="Chisq")
+anova(gam_txl_1,gam_txl_2, test="Chisq")
+anova(gam_gpt_1,gam_gpt_2, test="Chisq")
 ```
 
-I don't know how to interpret the above.
+```{r plots, eval=T}
+plot(gam_gpt_1, select=4)
+plot(gam_grnn_1, select=4)
 
-I also don't know how to interpret the plots.
+plot(gam_txl_1, select=4)
+plot(gam_ngram_1, select=4)
 
-```{r plots, eval=F}
-plot(gam_gpt_1)
 ```
 
-By-factor GAMs fail to run overnight / crash the R session.
+It looks like the effect is ~ 0 for most of the space where there's actually data, and it only goes off the rails in the region where there isn't data.
+
+Below is all words with centered frequency < -6 and centered_length > 4. Most of these are I think are recognizable as words. 
+```{r}
+select_data %>% select(word,  freq_center, length_center, sentence) %>% unique() %>% 
+  arrange(freq_center) %>% 
+  filter(freq_center< -6 & length_center > 4) %>% knitr::kable()
+```
 
@@
 
-```{r, eval=F}
+```{r, eval=T}
 
@@
 #plot(gam_txl_1, seWithMean = TRUE, shift = coef(gam_txl_1)[1], select=1)
+
+```
+
+# Maze LMs
+
+```{r for brms}
+
+ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_center, prev_surp=past_ngram_center, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
+                                            ) %>% mutate(model="5-gram") 
+grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_center, prev_surp=past_grnn_center, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
+                                            )%>% mutate(model="GRNN")
+
+txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_center, prev_surp=past_txl_center, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
+                                            ) %>% mutate(model="TXL")
+
+gpt_data <- labelled_pre_error %>% select(rt, surprisal=gpt_center, prev_surp=past_gpt_center, 
+                                            freq=freq_center, len=length_center,
+                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
+                                            ) %>% mutate(model="GPT-2")
+
+all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
+  write_rds(here("Analysis/models/maze_pre_brms.rds"))
+
+```
+
+```{r run brms, eval=F}
+# Note this was actually run on a cluster with a separate script that just had this
+data=readRDS("maze_pre_brms.rds")
+ priors <- c(
+      set_prior("normal(1000, 1000)", class="Intercept"),
+      set_prior("normal(0, 500)", class="b"),
+      set_prior("normal(0, 500)", class="sd"),
+      set_prior("lkj(1)",       class="cor"))
+run_model <- function(model_spec,data,file_name){
+  d <- data %>% filter(model==model_spec)
+  m1 <- brm(rt ~ surprisal * len + freq * len +
+              prev_surp * prev_len + prev_freq * prev_len+
+              (surprisal * len + freq * len +
+              prev_surp * prev_len + prev_freq * prev_len|subject)+(1|Word_ID),
+            data=d,
+            file=file_name,
+            prior=priors,
+            control=list(adapt_delta=.95))
+}
+
+run_model("5-gram",data,"5-gram_maze")
+run_model("GRNN",data,"grnn_maze")
+run_model("TXL",data,"txl_maze")
+run_model("GPT-2",data,"gpt_maze")
+
+```
+
+```{r summary}
+
+show_summary <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=1),
+           upper=round(.upper, digits=1),
+           E=round(.value, digits=1),
+           `Estimate`=str_c(E," [",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           ) %>% 
+    select(Term, `Estimate`)
+
+  stats
+}
+```
+
+```{r process-maze-models, eval=F}
+ngram <- read_rds(here("Analysis/models/5-gram_maze.rds")) %>% show_summary() %>% mutate(model="5-gram")
+txl <- read_rds(here("Analysis/models/txl_maze.rds")) %>% show_summary() %>% mutate(model="TXL")
+grnn <- read_rds(here("Analysis/models/grnn_maze.rds")) %>% show_summary() %>% mutate(model="GRNN")
+gpt <- read_rds(here("Analysis/models/gpt_maze.rds")) %>% show_summary() %>% mutate(model="GPT-2")
+
+summ <- ngram %>% union(txl) %>% union(grnn) %>% union(gpt) %>% write_rds(here("Analysis/models/brms_maze_summ.rds"))
+
+```
+
+# Maze Model comparison
+
+I'm not sure if this is what we want, but these are the P-values from anova's comparison the value of adding one set of surprisal predictors to a model that already has another.
+(So smaller values mean that the column  has a lot of predictive utility beyond the row label).
+
+I'm not sure what the best presentational format is -- or whether we want F values or p values or what.
+I also not sure how to arrange the table (should I transpose?) or label it so which one is better is clear.
+
+This is all on meaned data with no mixed effects and predictors of freq x length and surprisal(s) for word and previous word.
+
+```{r}
+select_data <- read_rds(here("Data/maze_pre_error.rds")) %>%  select(-starts_with("past2"), -starts_with("past3")) %>% 
+  filter(across(everything(),~!is.na(.x)))
+
+item_mean_data <- select_data %>% 
+  group_by(across(c(-subject, -rt))) %>% 
+  summarize(rt=mean(rt)) %>% 
+  ungroup()
+
+no_surp <- "rt ~ freq_center * length_center + past_freq_center * past_length_center"
+ngram <- "+ ngram_center + past_ngram_center"
+grnn <- "+ grnn_center + past_grnn_center"
+txl <- "+ txl_center + past_txl_center"
+gpt <- "+ gpt_center + past_gpt_center"
+
+do_comparison <- function(m1, m2){
+  # is there benefit of m2 above m1
+  if (m1==m2) {return("")}
+  lm1 <- lm(str_c(no_surp,m1), data=item_mean_data)
+  lm2 <- lm(str_c(no_surp,m1,m2), data=item_mean_data)
+  a <- anova(lm1,lm2)
+  fval <- anova(lm1,lm2)$F[2] %>% round() %>%  as.character
+  pval <- anova(lm1,lm2)$`Pr(>F)`[2] %>% pvalue(add_p=T) %>% as.character()
+  return(str_c(fval," (",pval,")"))
+}
+
+log_like <- function(m1){
+    lm(str_c(no_surp,m1), data=item_mean_data) %>% logLik() %>% round() %>% as.character()
+
+}
+
+tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
+             `over GRNN`=grnn,`over TXL`=txl,`over GPT-2`=gpt) %>% 
+   mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
+          `Log Lik`=map_chr(model_str, log_like)) %>% 
+   select(-model_str) %>% write_rds(here("Analysis/maze_model_compare.rds")) %>%  knitr::kable()
 
@@
 
-```{r formulae, echo=T}
+```{r spr formulae, echo=T}
 #no hierarchical, has freq x len interaction
@@
 ```{r, eval=T}
+all_mean_spr <-  read_rds(here("Analysis/models/meaned_spr.rds"))
 
@@
 
-# Maze LMs
-
-```{r for brms}
-
-ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_center, prev_surp=past_ngram_center, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
-                                            ) %>% mutate(model="5-gram") 
-grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_center, prev_surp=past_grnn_center, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
-                                            )%>% mutate(model="GRNN")
-
-txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_center, prev_surp=past_txl_center, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
-                                            ) %>% mutate(model="TXL")
-
-gpt_data <- labelled_pre_error %>% select(rt, surprisal=gpt_center, prev_surp=past_gpt_center, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
-                                            ) %>% mutate(model="GPT-2")
-
-all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
-  write_rds(here("Analysis/models/maze_pre_brms.rds"))
-
-```
-
-```{r run brms, eval=F}
-# Note this was actually run on a cluster with a separate script that just had this
-data=readRDS("maze_pre_brms.rds")
- priors <- c(
-      set_prior("normal(1000, 1000)", class="Intercept"),
-      set_prior("normal(0, 500)", class="b"),
-      set_prior("normal(0, 500)", class="sd"),
-      set_prior("lkj(1)",       class="cor"))
-run_model <- function(model_spec,data,file_name){
-  d <- data %>% filter(model==model_spec)
-  m1 <- brm(rt ~ surprisal * len + freq * len +
-              prev_surp * prev_len + prev_freq * prev_len+
-              (surprisal * len + freq * len +
-              prev_surp * prev_len + prev_freq * prev_len|subject)+(1|Word_ID),
-            data=d,
-            file=file_name,
-            prior=priors,
-            control=list(adapt_delta=.95))
-}
-
-run_model("5-gram",data,"5-gram_maze")
-run_model("GRNN",data,"grnn_maze")
-run_model("TXL",data,"txl_maze")
-run_model("GPT-2",data,"gpt_maze")
-
-```
-
-```{r summary}
-
-show_summary <- function(model){
-  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
-
-  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
-    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
-    group_by(.variable) %>% 
-    summarize(pct_above_0=mean(above_0)) %>% 
-    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
-    left_join(intervals, by=".variable") %>% 
-    mutate(lower=round(.lower, digits=1),
-           upper=round(.upper, digits=1),
-           E=round(.value, digits=1),
-           `Estimate`=str_c(E," [",lower,", ", upper,"]"),
-           Term=str_sub(.variable, 3, -1),
-           ) %>% 
-    select(Term, `Estimate`)
-
-  stats
-}
-```
-
-```{r process-maze-models, eval=F}
-ngram <- read_rds(here("Analysis/models/5-gram_maze.rds")) %>% show_summary() %>% mutate(model="5-gram")
-txl <- read_rds(here("Analysis/models/txl_maze.rds")) %>% show_summary() %>% mutate(model="TXL")
-grnn <- read_rds(here("Analysis/models/grnn_maze.rds")) %>% show_summary() %>% mutate(model="GRNN")
-gpt <- read_rds(here("Analysis/models/gpt_maze.rds")) %>% show_summary() %>% mutate(model="GPT-2")
-
-summ <- ngram %>% union(txl) %>% union(grnn) %>% union(gpt) %>% write_rds(here("Analysis/models/brms_maze_summ.rds"))
-
-```
-
 # SPR LMs
 
-```{r spr brms}
+```{r spr brms, eval=F}
+
+select_spr <- read_rds(here("Data/clean_spr.rds")) %>% 
+  filter(across(everything(),~!is.na(.x))) %>% 
+  ungroup()
+
+item_mean_spr <- select_spr %>% 
+  group_by(across(c(-subject, -rt))) %>% 
+  summarize(rt=mean(rt)) %>% 
+  ungroup()
 
@@
 
-all_mean_spr <- ngram_mean_data %>% union(grnn_mean_data) %>% union(txl_mean_data) %>% union(gpt_mean_data)
+all_mean_spr_center <- ngram_mean_data %>% union(grnn_mean_data) %>% union(txl_mean_data) %>% union(gpt_mean_data) %>% write_rds(here("Analysis/models/mean_spr_center.rds"))
 
@@
 
+all_mean_spr_center <- read_rds(here("Analysis/models/mean_spr_center.rds"))
 run_lm <- function(data,modelspec){
@@
 
-m_ngram <- run_lm(all_mean_spr,"5-gram")
-m_grnn <- run_lm(all_mean_spr, "GRNN")
-m_txl <- run_lm(all_mean_spr,"TXL")
-m_gpt <- run_lm(all_mean_spr,"GPT-2")
+m_ngram <- run_lm(all_mean_spr_center,"5-gram")
+m_grnn <- run_lm(all_mean_spr_center, "GRNN")
+m_txl <- run_lm(all_mean_spr_center,"TXL")
+m_gpt <- run_lm(all_mean_spr_center,"GPT-2")
 
@@
 
-data=readRDS("spr_pre_brms.rds")
+#data=readRDS("spr_pre_brms.rds")
+data=read_rds(here("Analysis/models/spr_pre_brms.rds"))
  priors <- c(
-      set_prior("normal(300, 100)", class="Intercept"),
       set_prior("normal(0, 5)", class="b"),
-      set_prior("normal(0, 5)", class="sd"),
-      set_prior("lkj(1)",       class="cor"))
+      set_prior("exponential(.1)", class="sd", group="subject", coef="Intercept"),
+      set_prior("exponential(.1)", class="sd", group="Word_ID", coef="Intercept"),
+      set_prior("exponential(1)", class="sd"),
+      set_prior("lkj(2)",       class="cor"),
+      set_prior("exponential(.1)", class="sigma")
+      )
 run_model <- function(model_spec,data,file_name){
-  d <- data %>% filter(model==model_spec)
-  m1 <- brm(rt ~ surp_center * length_center + length_center * freq_center +
+  d <- data %>% filter(model==model_spec) %>% 
+    mutate(center_rt=scale(rt, center=T, scale=F))
+  m1 <- brm(center_rt ~ surp_center * length_center + length_center * freq_center +
                   past_surp_center * past_length_center + past_length_center * past_freq_center + 
@@
                   past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
-              (surp_center * length_center + length_center * freq_center +
-                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
-                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
-                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center |subject)+(1|Word_ID),
+              (surp_center + length_center + freq_center|subject)+(1|Word_ID),
             data=d,
@@
 
-run_model("5-gram",data,"5-gram_spr")
-run_model("GRNN",data,"grnn_spr")
+#run_model("5-gram",data,"5-gram_spr")
+#run_model("GRNN",data,"grnn_spr")
 run_model("TXL",data,"txl_spr")
-run_model("GPT-2",data,"gpt_spr")
+#run_model("GPT-2",data,"gpt_spr")
 
 ```
+
+```{r diagnose-unhappiness}
+
+sad <- read_rds(here("Analysis/models/gpt_spr.rds"))
+
+summary(sad)
+
+d <- data %>% filter(model=="GPT-2") %>% select(-rt, -Word_ID, -model, -subject)
+
+cor(d) %>%as.tibble() %>%  mutate(across(where(is.numeric), ~round(.x,2))) %>% View()
+```
+
+Really struggling with the intercept for population & subject level -- maybe sd was intercept needs to be a lot bigger?
+
+# SPR Model comparison
+
+I'm not sure if this is what we want, but these are the P-values from anova's comparison the value of adding one set of surprisal predictors to a model that already has another.
+(So smaller values mean that the column  has a lot of predictive utility beyond the row label).
+
+I'm not sure what the best presentational format is -- or whether we want F values or p values or what.
+I also not sure how to arrange the table (should I transpose?) or label it so which one is better is clear.
+
+This is all on meaned data with no mixed effects and predictors of freq x length and surprisal(s) for word and previous word.
+
+```{r}
+select_spr <- read_rds(here("Data/clean_spr.rds")) %>% 
+  filter(across(everything(),~!is.na(.x))) %>% 
+  ungroup()
+
+item_mean_spr <- select_spr %>% 
+  group_by(across(c(-subject, -rt))) %>% 
+  summarize(rt=mean(rt)) %>% 
+  ungroup()
+
+no_surp <- "rt ~ freq_center * length_center + 
+            past_freq_center * past_length_center + 
+            past2_freq_center * past2_length_center + 
+            past3_freq_center * past3_length_center "
+ngram <- "+ ngram_center + past_ngram_center + past2_ngram_center + past3_ngram_center"
+grnn <- "+ grnn_center + past_grnn_center +past2_grnn_center +past3_grnn_center"
+txl <- "+ txl_center + past_txl_center +past2_txl_center + past3_txl_center"
+gpt <- "+ gpt_center + past_gpt_center + past2_gpt_center + + past3_gpt_center"
+
+do_comparison <- function(m1, m2){
+  # is there benefit of m2 above m1
+  if (m1==m2) {return("")}
+  lm1 <- lm(str_c(no_surp,m1), data=item_mean_spr)
+  lm2 <- lm(str_c(no_surp,m1,m2), data=item_mean_spr)
+  a <- anova(lm1,lm2)
+  fval <- anova(lm1,lm2)$F[2] %>% round() %>%  as.character
+  pval <- anova(lm1,lm2)$`Pr(>F)`[2] %>% pvalue(add_p=T) %>% as.character()
+  return(str_c(fval," (",pval,")"))
+}
+
+log_like <- function(m1){
+    lm(str_c(no_surp,m1), data=item_mean_spr) %>% logLik() %>% round() %>% as.character()
+
+}
+
+tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
+             `over GRNN`=grnn,`over TXL`=txl,`over GPT-2`=gpt) %>% 
+   mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
+          `Log Lik`=map_chr(model_str, log_like)) %>% 
+   select(-model_str)%>% write_rds(here("Analysis/spr_model_compare.rds")) %>% kable()
+
+```

### 2021-07-14 Papers/Paper/manuscript.Rmd
work on nested model comparisons

@@
 
-For model comparison, we fit models with only frequency and length as predictors, as well as models that also had one or more sources of surprisal.
-We centered all effects.
-We included a random intercept for word identity, but no further random effects so the models would fit in lme4 [@batesFittingLinearMixedeffects2015].
-TODO!!!
+For model comparison, we took by-item averaged data to aid in fast model fitting.
+We included frequency, length, and their interaction to all models.
+Then we fit models with either 1 or 2 sources of surprisal using lm and assessed the effect of adding the second surprisal source with an anova.
+We used predictors for the current and past word and centered all effects.
 
@@
 For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct.
-TODO model differences!!!
+To account for spill over effects known to exist in SPR, we analyse predictors at the current word as well as the past 3 words for all models.
+TODO differences for linear model!
 
@@
 There is no relationship between previous word surprisal and RT.
-TODO replace graph!
 
 ```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:gam-cap)"}
-labelled_pre_error <- read_rds(here("Data/maze_pre_error.rds"))
-
-ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
-                                            ) %>% mutate(model="5-gram") 
-grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
-                                            )%>% mutate(model="GRNN")
-
-txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
-                                            ) %>% mutate(model="TXL")
-
-all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% 
-  select(surprisal, prev_surp,model) %>% 
-  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
-  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
 
@@
 
-dens2 <-   ggplot(all_data, aes(x=value))+
+all_mean_data <-  read_rds(here("Analysis/models/meaned_maze.rds"))
+
+dens2 <-   ggplot(all_mean_data, aes(x=value))+
   geom_density(fill="gray",)+
@@
 
-As a last analysis, we checked which of our surprisal models had the best fit using a nested model comparison. 
-We found that all surprisal sources provide predictive value over none, but that the information provided by the Ngram model does not provide additional value to a model that already has GRNN in it.
-TXL and GRNN appear to contain some complementary predictive value.
+As a last analysis, we checked which of our surprisal models had the best fit using a nested model comparison shown in Table \@ref(tab:maze-compare).
+We assess the benefits of adding each model's predictions as a second surprisal source to see which models pick up on information not contained in another model.
+In terms of log likelihoods, GPT-2 is better than GRNN is better than TXL is better than 5-gram.
+Model comparisons tell a similar story; GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complemetary information over GPT-2.
+TXL provides a lot over 5-gram, and 5-gram provides little over any model.
+```{r maze-compare}
 
-```{r,include=F, eval=F}
-d_lm <- labelled_pre_error %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
-                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
-  summarize(mean_rt=mean(rt))
+read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="t",caption="Results of model comparisons on Maze data.
+Each row shows the additional predictive value gained from adding that model to another model.
+F values and p values from anova tests are reported.
+We also report log likelihoods of models with only one surprisal source.")
 
-no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
-
-ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
-
-grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
-
-txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
-
-all_surp <- lmer(mean_rt ~ ngram_center + past_c_ngram + 
-                   grnn_center + past_c_grnn +
-                   txl_center + past_c_txl +
-                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
-
-ngram_grnn <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
-                   grnn_center + past_c_grnn +
-                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
-
-ngram_txl <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
-                   txl_center + past_c_txl +
-                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
-
-grnn_txl <-  lmer(mean_rt ~ grnn_center + past_c_grnn +
-                   txl_center + past_c_txl +
-                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
-
-no_lag <-  lmer(mean_rt ~ ngram_center +
-                   grnn_center +
-                   txl_center + 
-                   freq_center*length_center+(1|word), data=d_lm)
-
-no_surp_lag <-  lmer(mean_rt ~ ngram_center +  
-                   grnn_center + 
-                   txl_center +                    freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)
-
-  only_surp_lag <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
-                   grnn_center + past_c_grnn +
-                   txl_center + past_c_txl +
-                   freq_center*length_center+(1|word), data=d_lm)
-```
-
-```{r, include=F, eval=F}
-anova(all_surp, txl_only)
-anova(all_surp, grnn_only)
-anova(all_surp, ngram_only)
-
-anova(txl_only, no_surp)
-anova(grnn_only, no_surp)
-anova(ngram_only, no_surp)
-
-anova(all_surp, ngram_grnn)
-anova(all_surp, ngram_txl)
-anova(all_surp, grnn_txl)
-
-anova(ngram_grnn, ngram_only)
-anova(ngram_grnn, grnn_only)
-anova(ngram_txl, ngram_only)
-anova(ngram_txl, txl_only)
-anova(grnn_txl, grnn_only)
-anova(grnn_txl, txl_only)
-
-anova(all_surp, no_lag)
-
-anova(all_surp, no_surp_lag)
-
-anova(no_surp_lag, no_lag)
-
-anova(all_surp, only_surp_lag)
-
-anova(only_surp_lag, no_lag)
 ```
@@
 
-(ref:spr-gam-cap) GAM predictions of reaction time (RT) for SPR data from @futrellNaturalStoriesCorpus2020 as a function of either current word surprisal (top) or previous word surprisal (bottom).
+(ref:spr-gam-cap) GAM predictions of reaction time (RT) for SPR data from @futrellNaturalStoriesCorpus2020 as a function of current word surprisal (top) or the surprisal of an earlier word, up to 3 words back.
 Density of data is shown along the x-axis.
-TODO replace graph!
 
@@
 p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
-plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))
+plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .2))
 
@@
 
-We ran nested model comparison on the SPR data as well, and found that all sources of surprisal have predictive value, but the GRNN model provides the most benefit.
-Adding other models to GRNN is marginally beneficial, but TXL and N-gram together are more predictive than either alone.
+Similarly to the model comparison for Maze, we also conduct a model comparison on the SPR data shown in Table \@ref(tab:spr-compare).
+In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small.
+Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN.
+Overall, F-values here are much lower than for Maze.
+TODO how to interpret!
 
-```{r, eval=F}
+```{r spr-compare}
 
-no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
+read_rds(here("Analysis/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="t",caption="Results of model comparisons on SPR data.
+Each row shows the additional predictive value gained from adding that model to another model.
+F values and p values from anova tests are reported.
+We also report log likelihoods of models with only one surprisal source.")
 
-ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-all_surp <- lmer(mean_rt ~ ngram_center + past_c_ngram + 
-                   grnn_center + past_c_grnn +
-                   txl_center + past_c_txl +
-                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-ngram_grnn <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
-                   grnn_center + past_c_grnn +
-                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-ngram_txl <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
-                   txl_center + past_c_txl +
-                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-grnn_txl <-  lmer(mean_rt ~ grnn_center + past_c_grnn +
-                   txl_center + past_c_txl +
-                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-no_lag <-  lmer(mean_rt ~ ngram_center +
-                   grnn_center +
-                   txl_center + 
-                   freq_center*length_center+(1|word), data=d.byword)
-
-no_surp_lag <-  lmer(mean_rt ~ ngram_center +  
-                   grnn_center + 
-                   txl_center +                    freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d.byword)
-
-  only_surp_lag <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
-                   grnn_center + past_c_grnn +
-                   txl_center + past_c_txl +
-                   freq_center*length_center+(1|word), data=d.byword)
 ```
-
-```{r, eval=F}
-anova(txl_only, no_surp)
-anova(grnn_only, no_surp)
-anova(ngram_only, no_surp)
-
-anova(all_surp, ngram_grnn)
-anova(all_surp, ngram_txl)
-anova(all_surp, grnn_txl)
-
-anova(ngram_grnn, ngram_only)
-anova(ngram_grnn, grnn_only)
-anova(ngram_txl, ngram_only)
-anova(ngram_txl, txl_only)
-anova(grnn_txl, grnn_only)
-anova(grnn_txl, txl_only)
-```
-
 # Discussion

### 2021-07-20 Analysis/models.Rmd
add spr models, move floats

@@
  priors <- c(
+      set_prior("normal(350,50)", class="Intercept"),
       set_prior("normal(0, 5)", class="b"),
-      set_prior("exponential(.1)", class="sd", group="subject", coef="Intercept"),
-      set_prior("exponential(.1)", class="sd", group="Word_ID", coef="Intercept"),
-      set_prior("exponential(1)", class="sd"),
-      set_prior("lkj(2)",       class="cor"),
-      set_prior("exponential(.1)", class="sigma")
+      set_prior("exponential(.01)", class="sd", group="subject", coef="Intercept"),
+      set_prior("exponential(.01)", class="sd", group="Word_ID", coef="Intercept"),
+     # set_prior("exponential(1)", class="sd"),
+     # set_prior("lkj(2)",       class="cor"),
+      set_prior("exponential(.01)", class="sigma")
       )
 run_model <- function(model_spec,data,file_name){
-  d <- data %>% filter(model==model_spec) %>% 
-    mutate(center_rt=scale(rt, center=T, scale=F))
-  m1 <- brm(center_rt ~ surp_center * length_center + length_center * freq_center +
+  d <- data %>% filter(model==model_spec)
+  m1 <- brm(rt ~ surp_center * length_center + length_center * freq_center +
                   past_surp_center * past_length_center + past_length_center * past_freq_center + 
@@
                   past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
-              (surp_center + length_center + freq_center|subject)+(1|Word_ID),
+              (1|subject)+(1|Word_ID),
             data=d,
@@
 
-```{r diagnose-unhappiness}
+```{r diagnose-unhappiness, eval=F}
 
-sad <- read_rds(here("Analysis/models/gpt_spr.rds"))
+sad <- read_rds(here("Analysis/models/txl_spr.rds"))
 
@@
 
-d <- data %>% filter(model=="GPT-2") %>% select(-rt, -Word_ID, -model, -subject)
+```
+```{r summary-spr}
 
-cor(d) %>%as.tibble() %>%  mutate(across(where(is.numeric), ~round(.x,2))) %>% View()
+show_summary_spr <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=1),
+           upper=round(.upper, digits=1),
+           E=round(.value, digits=1),
+           `Estimate`=str_c(E," [",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           ) %>% 
+    select(Term, `Estimate`)
+
+  stats
+}
+```
+
+```{r process-spr-models, eval=F}
+ngram_spr <- read_rds(here("Analysis/models/5-gram_spr.rds")) %>% show_summary_spr() %>% mutate(model="5-gram")
+txl_spr <- read_rds(here("Analysis/models/txl_spr.rds")) %>% show_summary_spr() %>% mutate(model="TXL")
+grnn_spr <- read_rds(here("Analysis/models/grnn_spr.rds")) %>% show_summary_spr() %>% mutate(model="GRNN")
+gpt_spr <- read_rds(here("Analysis/models/gpt_spr.rds")) %>% show_summary_spr() %>% mutate(model="GPT-2")
+
+summ_spr <- ngram_spr %>% union(txl_spr) %>% union(grnn_spr) %>% union(gpt_spr) %>% write_rds(here("Analysis/models/brms_spr_summ.rds"))
+
 ```

### 2021-07-20 Papers/Paper/manuscript.Rmd
add spr models, move floats

@@
  - \usepackage{setspace}\singlespacing
+ - \renewcommand{\textfraction}{0.05}
+ - \renewcommand{\topfraction}{0.8}
+ - \renewcommand{\bottomfraction}{0.8}
+ - \renewcommand{\floatpagefraction}{1}
+ - \setcounter{topnumber}{3}
+ - \setcounter{bottomnumber}{3}
+ - \setcounter{totalnumber}{4}
 
@@
 
-```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:diagram-cap)"}
+```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:diagram-cap)"}
 knitr::include_graphics(here("Papers/maze_diagram_cropped.pdf"))
@@
 To account for spill over effects known to exist in SPR, we analyse predictors at the current word as well as the past 3 words for all models.
-TODO differences for linear model!
+For linear models, we centered all predictors.
+We were unable to fit the full mixed effects model, so we used the best model that reasonably converged which had a by-subject random intercept and a per-word-token random intercept.
+We used regularizing priors(normal(350,50) for intercept, normal(0,5) for beta, exponential(.01) for sd and sigma).
+Models were run in brm [@burknerAdvancedBayesianMultilevel2018].
 
@@
 
-```{r errors, out.height="25%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:error-cap)"}
+```{r errors, out.height="25%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:error-cap)"}
 error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
@@
 
-```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:gam-cap)"}
+```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:gam-cap)"}
 
@@
 
-knitr::kable(summ, format="latex", position="t",caption="Predictions from fitted Bayesian regression models.
+knitr::kable(summ, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models.
 All terms were centered, but not rescaled.
@@
 
-read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="t",caption="Results of model comparisons on Maze data.
+read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data.
 Each row shows the additional predictive value gained from adding that model to another model.
@@
 
-```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos='t', fig.cap="(ref:spr-gam-cap)"}
+```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
 gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
@@
 
-```{r spr-table, eval=F}
+```{r spr-table}
 
-all_tidy <- read_rds(here("Analysis/SPR_lmer_summ.rds")) %>% 
-  select(Term=term, estimate, std.error, type) %>% 
-  mutate(Ci_high=estimate+1.96*std.error,
-         Ci_low=estimate-1.96*std.error,
-         Range=str_c(round(estimate,1)," [",round(Ci_low, 1),", ",round(Ci_high,1),"]")) %>% 
-  select(Term, type, Range) %>% 
-  pivot_wider(id_cols=Term, names_from=type, values_from=Range) %>% 
+summ_spr <- read_rds(here("Analysis/models/brms_spr_summ.rds")) %>% 
+  pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
   mutate(Term=factor(Term, 
-                        levels=c("(Intercept)", 
+                        levels=c("Intercept", 
                                  "surp_center",
@@
                                  "freq_center",
-                                 "freq_center:length_center",
-                                 "past_c_surp",
-                                 "past_c_length",
-                                 "past_c_freq",
-                                 "past_c_freq:past_c_length")
+                                 "surp_center:length_center",
+                                 "length_center:freq_center",
+                                 "past_surp_center",
+                                 "past_length_center",
+                                 "past_freq_center",
+                                 "past_surp_center:past_length_center",
+                                 "past_length_center:past_freq_center",
+                                 "past2_surp_center",
+                                 "past2_length_center",
+                                 "past2_freq_center",
+                                 "past2_surp_center:past2_length_center",
+                                 "past2_length_center:past2_freq_center",
+                                 "past3_surp_center",
+                                 "past3_length_center",
+                                 "past3_freq_center",
+                                 "past3_surp_center:past3_length_center",
+                                 "past3_length_center:past3_freq_center"
+                                 )
                                  )) %>% 
   arrange(Term) %>% 
-  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency",
-                   "Freq x Length", "Past Surprisal",
-                  "Past Length", "Past Freq", "Past Freq x Length"))
+  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency","Surp x Length", "Freq x Length",
+                "Past Surprisal", "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length",
+                "2Past Surprisal", "2Past Length", "2Past Freq", "2Past Surp x Length", "2Past Freq x Length",
+                "3Past Surprisal", "3Past Length", "3Past Freq", "3Past Surp x Length", "3Past Freq x Length"))
 
-knitr::kable(all_tidy, format="latex", position="t",caption="Predictions from fitted linear regression models.
+knitr::kable(summ_spr, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models for SPR data.
 All terms were centered, but not rescaled.
@@
 Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")
+
 ```
@@
 Note that the y-axis is fairly narrow, and so the predicted effect of changes in surprisal is fairly small.
-This is confirmed by linear models (see Table \@ref(tab:spr-table)); note that due to the size of the data, these models were fit differently than the A-maze models due to computational constraints.
-Surprisal and length effects are evident for both the current and past word, but we do not see frequency effects.
-The effects are much smaller than for A-maze, although some of this may be due to noise from spillover from prior words or differences in model structure.
+This is confirmed by linear models (see Table \@ref(tab:spr-table)).
+Surprisal and length effects are evident for the current word, and most models shows surprisal effects from the past word, but we do not see frequency effects.
+The effects are much smaller than for A-maze, even though this model accounts for spillover effects from more previous words.
 
-Lastly, we are interested to know which of the models is better at predicting human behavior data, and whether they are complementary.
-I fit a series of models including the frequency and length predictors, and some number of past or present surprisal sources.
-This allows us to tell whether different models of surprisal are "picking up" on different things that affect reading time.
-
-Similarly to the model comparison for Maze, we also conduct a model comparison on the SPR data shown in Table \@ref(tab:spr-compare).
+Similarly to the model comparison for Maze, we also conducted a model comparison on the SPR data shown in Table \@ref(tab:spr-compare).
 In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small.
@@
 Overall, F-values here are much lower than for Maze.
-TODO how to interpret!
 
@@
 
-read_rds(here("Analysis/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="t",caption="Results of model comparisons on SPR data.
+read_rds(here("Analysis/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on SPR data.
 Each row shows the additional predictive value gained from adding that model to another model.

### 2021-07-22 Analysis/models.Rmd
different attempts to get something reasonable for SPR

@@
 title: "Natural Stories model analyses"
-output: html_document
+output: 
+  html_document:
+    toc: true
 ---
@@
 
+```{r, eval=F}
+
+#data=readRDS("spr_pre_brms.rds")
+data=read_rds(here("Analysis/models/spr_pre_brms.rds"))
+
+run_model_lmer <- function(model_spec,data){
+  d <- data %>% filter(model==model_spec)
+  m1 <- lmer(rt ~ surp_center * length_center + length_center * freq_center +
+                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
+                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
+                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
+              (surp_center + length_center +  freq_center +
+                  past_surp_center + past_length_center + past_freq_center ||subject)+(1|Word_ID),
+            data=d) 
+  m1
+}
+
+run_model_lmer("GRNN",data) %>% write_rds(here("Analysis/models/grnn_spr_lmer.rds"))
+run_model_lmer("5-gram",data) %>% write_rds(here("Analysis/models/5-gram_spr_lmer.rds"))
+run_model_lmer("TXL",data) %>% write_rds(here("Analysis/models/txl_spr_lmer.rds"))
+run_model_lmer("GPT-2",data) %>% write_rds(here("Analysis/models/gpt_spr_lmer.rds"))
+
+```
+## Summaries of LMER 
+These have some heirarchical effects
+```{r}
+read_rds(here("Analysis/models/grnn_spr_lmer.rds")) %>% summary()
+read_rds(here("Analysis/models/5-gram_spr_lmer.rds")) %>% summary()
+read_rds(here("Analysis/models/txl_spr_lmer.rds")) %>% summary()
+read_rds(here("Analysis/models/gpt_spr_lmer.rds")) %>% summary()
+
+```
+
 ```{r diagnose-unhappiness, eval=F}

### 2021-07-28 Analysis/models.Rmd
update spr models, integrate into manuscript

@@
 These have some heirarchical effects
-```{r}
-read_rds(here("Analysis/models/grnn_spr_lmer.rds")) %>% summary()
-read_rds(here("Analysis/models/5-gram_spr_lmer.rds")) %>% summary()
-read_rds(here("Analysis/models/txl_spr_lmer.rds")) %>% summary()
-read_rds(here("Analysis/models/gpt_spr_lmer.rds")) %>% summary()
 
-```
-
-```{r diagnose-unhappiness, eval=F}
-
-sad <- read_rds(here("Analysis/models/txl_spr.rds"))
-
-summary(sad)
-
-```
 ```{r summary-spr}
@@
 show_summary_spr <- function(model){
-  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
 
-  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
-    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
-    group_by(.variable) %>% 
-    summarize(pct_above_0=mean(above_0)) %>% 
-    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
-    left_join(intervals, by=".variable") %>% 
-    mutate(lower=round(.lower, digits=1),
-           upper=round(.upper, digits=1),
-           E=round(.value, digits=1),
-           `Estimate`=str_c(E," [",lower,", ", upper,"]"),
-           Term=str_sub(.variable, 3, -1),
+  stats <- summary(model)[["coefficients"]] %>% 
+    as_tibble(rownames="Term") %>% 
+    mutate(width=1.97*`Std. Error`,
+           lower=round(Estimate-width,1),
+           upper=round(Estimate+width,1),
+           E=round(Estimate, digits=1),
+           `Estimate`=str_c(E," [",lower,", ", upper,"]")
            ) %>% 
-    select(Term, `Estimate`)
+    select(Term, Estimate)
 
@@
 ```{r process-spr-models, eval=F}
-ngram_spr <- read_rds(here("Analysis/models/5-gram_spr.rds")) %>% show_summary_spr() %>% mutate(model="5-gram")
-txl_spr <- read_rds(here("Analysis/models/txl_spr.rds")) %>% show_summary_spr() %>% mutate(model="TXL")
-grnn_spr <- read_rds(here("Analysis/models/grnn_spr.rds")) %>% show_summary_spr() %>% mutate(model="GRNN")
-gpt_spr <- read_rds(here("Analysis/models/gpt_spr.rds")) %>% show_summary_spr() %>% mutate(model="GPT-2")
+ngram_spr <- read_rds(here("Analysis/models/5-gram_spr_lmer.rds")) %>% show_summary_spr() %>% mutate(model="5-gram")
+txl_spr <- read_rds(here("Analysis/models/txl_spr_lmer.rds")) %>% show_summary_spr() %>% mutate(model="TXL")
+grnn_spr <- read_rds(here("Analysis/models/grnn_spr_lmer.rds")) %>% show_summary_spr() %>% mutate(model="GRNN")
+gpt_spr <- read_rds(here("Analysis/models/gpt_spr_lmer.rds")) %>% show_summary_spr() %>% mutate(model="GPT-2")
 
-summ_spr <- ngram_spr %>% union(txl_spr) %>% union(grnn_spr) %>% union(gpt_spr) %>% write_rds(here("Analysis/models/brms_spr_summ.rds"))
+summ_spr <- ngram_spr %>% union(txl_spr) %>% union(grnn_spr) %>% union(gpt_spr) %>% write_rds(here("Analysis/models/lmer_spr_summ.rds"))
 

### 2021-07-28 Papers/Paper/manuscript.Rmd
update spr models, integrate into manuscript

@@
 For linear models, we centered all predictors.
-We were unable to fit the full mixed effects model, so we used the best model that reasonably converged which had a by-subject random intercept and a per-word-token random intercept.
-We used regularizing priors(normal(350,50) for intercept, normal(0,5) for beta, exponential(.01) for sd and sigma).
-Models were run in brm [@burknerAdvancedBayesianMultilevel2018].
+We were unable to fit the full mixed effects model.
+The best model we could fit had by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and  a per-word-token random intercept, fit with lme4, as this structure did not fit reliably in brms.
 
@@
 
-summ_spr <- read_rds(here("Analysis/models/brms_spr_summ.rds")) %>% 
+summ_spr <- read_rds(here("Analysis/models/lmer_spr_summ.rds")) %>% 
   pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
   mutate(Term=factor(Term, 
-                        levels=c("Intercept", 
+                        levels=c("(Intercept)", 
                                  "surp_center",
@@
 
-knitr::kable(summ_spr, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models for SPR data.
+knitr::kable(summ_spr, format="latex", position="ht",caption="Predictions from fitted regression models for SPR data.
 All terms were centered, but not rescaled.
 Units are in ms.
-Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")
+Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.
+Uncertainty interval is +/- 1.97 standard error.")
 

### 2021-08-03 Analysis/models.Rmd
add scatter and R^2, update text to reflect R's comments

@@
   filter(after_mistake==0) %>% 
-  select(subject, word_num, word, rt, sentence, type)
+  select(subject, word_num, word, rt, sentence, type) %>% write_rds(here("Data/maze_pre_error.rds"))
 
@@
 
+r_squared <- function(m1){
+    lm1 <- lm(str_c(no_surp,m1), data=item_mean_data) %>% summary()
+    lm1$r.squared %>% round(2) %>% as.character()
+}
+
 tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
@@
    mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
-          `Log Lik`=map_chr(model_str, log_like)) %>% 
+          `Log Lik`=map_chr(model_str, log_like),
+          r_squared=map_chr(model_str, r_squared)) %>% 
    select(-model_str) %>% write_rds(here("Analysis/maze_model_compare.rds")) %>%  knitr::kable()
@@
 
-d.good <- d.limited %>% filter(correct>4) %>% rename(rt=RT) %>% filter(rt>100 & rt<5000)
+d.good <- d.limited %>% filter(correct>4) %>% rename(rt=RT) %>% filter(rt>100 & rt<5000) %>% write_rds(here("Data/SPR/first.rds"))
+
 ```
@@
   return(str_c(fval," (",pval,")"))
+  #return(a)
 }
 
+#do_comparison(ngram, grnn)
 log_like <- function(m1){
@@
 
+r_squared <- function(m1){
+    lm1 <- lm(str_c(no_surp,m1), data=item_mean_spr) %>% summary()
+    lm1$r.squared %>% round(2) %>% as.character()
+}
+
 tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
@@
    mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
-          `Log Lik`=map_chr(model_str, log_like)) %>% 
+          `Log Lik`=map_chr(model_str, log_like),
+          `r_squared`=map_chr(model_str, r_squared)) %>% 
    select(-model_str)%>% write_rds(here("Analysis/spr_model_compare.rds")) %>% kable()
@@
 ```
+
+# R-squared
+
+```{r}
+spr <- read_rds(here("Analysis/spr_model_compare.rds"))
+
+maze <- read_rds(here("Analysis/maze_model_compare.rds"))
+```
+
+# Maze and SPR scatter
+
+```{r}
+tokenization <- here("Data/SPR/all_stories.tok.txt") %>% 
+  read_delim(delim="\t") %>% 
+  mutate(Word_In_Story_Num=zone,
+         Story_Num=item)
+
+labs <- read_rds(here("Prep_code/natural_stories_surprisals.rds")) %>% 
+  left_join(read_delim(here("Materials/natural_stories_sentences.tsv"), delim="\t")) %>% 
+  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())
+
+spr <- read_rds(here("Data/SPR/first.rds")) %>% left_join(tokenization) %>% left_join(labs) %>% select(WorkerId,rt,Story_Num,Sentence_Num, word_num)
+
+maze <- read_rds(here("Data/maze_pre_error.rds")) %>% left_join(labs) %>% select(subject, rt, Story_Num, Sentence_Num, word_num)
+```
+
+Abnormally low and high times were excluded, then for each word, an average was taken across subjects.
+For sentence level, these averages were then summed across the sentence.
+(Because of A-maze sentence discontinuation and general outlier handling, summing the average is better than averaging the sum.)
+
+Word level correlation between maze and spr
+```{r}
+
+summ_spr <- spr %>% group_by(Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt))
+
+summ <- maze %>% group_by(Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% left_join(summ_spr)
+
+ggplot(summ, aes(x=maze, y=spr))+
+  geom_point(alpha=.02)+
+  coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
+  labs(title="Word level correlation in ms",subtitle="Axes trimmed")
+
+cor(summ$spr,summ$maze, use="complete.obs")
+
+```
+
+sentence level correlation between maze and spr
+
+Note that a lot of this is driven purely by sentence length, which isn't interesting. 
+```{r}
+sentence <- summ %>% group_by(Story_Num, Sentence_Num) %>% 
+  summarize(maze=sum(maze)/1000,
+            spr=sum(spr)/1000)
+
+ggplot(sentence, aes(x=maze, y=spr))+geom_point(alpha=.1)+
+  labs(title="Sentence level correlation in s")
+
+cor(sentence$spr,sentence$maze, use="complete.obs")
+```
+
+Averaging across words at the sentence level removes the sentence-length effect.
+
+```{r}
+sentence_avg <- summ %>% group_by(Story_Num, Sentence_Num) %>% 
+  summarize(maze=mean(maze),
+            spr=mean(spr))
+ggplot(sentence_avg, aes(x=maze, y=spr))+geom_point(alpha=.1)+
+    coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
+  labs(title="Sentence level correlation in ms")
+
+cor(sentence_avg$spr,sentence_avg$maze, use="complete.obs")
+
+```
+
+As a baseline, we can do all the same things on split-halfs of each data set.
+We split the subjects in half.
+(Note that this means that each half has the same # of subjects, but may not have the same #s of each story)
+
+Within spr correlations are higher than spr-maze.
+
+```{r spr-split-half}
+set.seed(42)
+spr_subs <- spr$WorkerId %>% unique()
+length(spr_subs)
+#180 participants
+
+spr1 <- sample(spr_subs, 90)
+
+spr_split <- spr %>% 
+  mutate(half=ifelse(WorkerId %in% spr1, "spr1","spr2")) %>% 
+  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt)) %>% 
+  pivot_wider(names_from=half, values_from=spr)
+
+ggplot(spr_split, aes(x=spr1, y=spr2))+
+  geom_point(alpha=.02)+
+  coord_cartesian(xlim=c(0,600),ylim=c(0,600))+
+  labs(title="Word level correlation in ms",subtitle="Axes trimmed")
+
+cor(spr_split$spr1,spr_split$spr2, use="complete.obs")
+spr_sentence <- spr_split %>% group_by(Story_Num, Sentence_Num) %>% 
+  summarize(spr1=sum(spr1)/1000,
+            spr2=sum(spr2)/1000)
+
+ggplot(spr_sentence, aes(x=spr1, y=spr2))+geom_point(alpha=.1)+
+  labs(title="Sentence level correlation in s")
+
+cor(spr_sentence$spr1,spr_sentence$spr2, use="complete.obs")
+
+spr_sentence_avg <- spr_split%>% group_by(Story_Num, Sentence_Num) %>% 
+  summarize(spr1=mean(spr1),
+            spr2=mean(spr2))
+ggplot(spr_sentence_avg, aes(x=spr1, y=spr2))+geom_point(alpha=.1)+
+    coord_cartesian(xlim=c(0,600),ylim=c(0,600))+
+  labs(title="Sentence level correlation in ms")
+
+cor(spr_sentence_avg$spr1,spr_sentence_avg$spr2, use="complete.obs")
+```
+
+Same things for maze.
+Note that with fewer subjects, and given only pre-error data, there's a lot more chance of emptiness.
+I think the weak sentence average correlation is from places where some sentences didn't go as long in one of the halves (b/c low data).
+
+```{r maze-split-half}
+set.seed(42)
+maze_subs <- maze$subject %>% unique()
+length(maze_subs)
+#180 participants
+
+maze1 <- sample(maze_subs, 31)
+
+maze_split <- maze %>% 
+  mutate(half=ifelse(subject %in% maze1, "maze1","maze2")) %>% 
+  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% 
+  pivot_wider(names_from=half, values_from=maze)
+
+ggplot(maze_split, aes(x=maze1, y=maze2))+
+  geom_point(alpha=.02)+
+  coord_cartesian(xlim=c(0,1500),ylim=c(0,1500))+
+  labs(title="Word level correlation in ms",subtitle="Axes trimmed")
+
+cor(maze_split$maze1,maze_split$maze2, use="complete.obs")
+maze_sentence <- maze_split %>% group_by(Story_Num, Sentence_Num) %>% 
+  summarize(maze1=sum(maze1)/1000,
+            maze2=sum(maze2)/1000)
+
+ggplot(maze_sentence, aes(x=maze1, y=maze2))+geom_point(alpha=.1)+
+  labs(title="Sentence level correlation in s")
+
+cor(maze_sentence$maze1,maze_sentence$maze2, use="complete.obs")
+
+maze_sentence_avg <- maze_split%>% group_by(Story_Num, Sentence_Num) %>% 
+  summarize(maze1=mean(maze1),
+            maze2=mean(maze2))
+ggplot(maze_sentence_avg, aes(x=maze1, y=maze2))+geom_point(alpha=.1)+
+    coord_cartesian(xlim=c(0,1500),ylim=c(0,1500))+
+  labs(title="Sentence level correlation in ms")
+
+cor(maze_sentence_avg$maze1,maze_sentence_avg$maze2, use="complete.obs")
+```

### 2021-08-03 Papers/Paper/manuscript.Rmd
add scatter and R^2, update text to reflect R's comments

@@
 We adapt A-maze for use on longer passages and test it on the Natural Stories corpus.
-We find that people can comprehend what they read during the Maze task.
+We find that people can comprehend these longer text passages during the Maze task.
 Moreover, the  task yields useable reaction time data with word predictability effects that are linear in the surprisal of the current word, with little spillover effect from the surprisal of the previous word.
@@
 This is an indication of spillover from the previous word.
-It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020].
+It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
 Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization.
@@
 If participants make a mistake, the sentence discontinues.
+Two versions of the maze task exist: lexical or L-maze where the distractors are nonce words and grammatical or G-maze where the distractors are real words of the language that don't fit in the context of the sentence so far.
 Theoretically, participants must fully integrate each word into the sentence in order to confidently select it.
-This idea is supported by studies finding strongly localized effects [@witzelComparisonsOnlineReading2012a].
+This idea is supported by studies finding strongly localized effects for G-maze which is more sensitive than L-maze [@witzelComparisonsOnlineReading2012a].
 
-The downside of Maze is that materials are effort-intensive to construct because of the need to select infelicitious words as distractors for each spot of each sentence; this may explain why the Maze task was not widely adopted.
+The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitious words as distractors for each spot of each sentence; this may explain why the Maze task was not widely adopted.
 @boyceMazeMadeEasy2020 demonstrate a way to automatically generate Maze distractors by using NLP language models to find words that are high-surprisal in the context of the target sentence.
-The quality of these A-Maze distractors is not up to that of hand-generated distractors, but @boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzelComparisonsOnlineReading2012a.
-A-maze out performed a SPR control in detecting and localizing expected slowdown effects.
+The quality of these A-maze distractors is not up to that of hand-generated G-maze distractors, but @boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzelComparisonsOnlineReading2012a.
+A-maze outperformed L-maze and an SPR control in detecting and localizing expected slowdown effects.
 @sloggettAmazeAnyOther2020 also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.
@@
 Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants.
-If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on basically any naturalistic text.
+If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on a wide variety of naturalistic texts.
 The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
@@
 To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions.
+See the Appendix for an except of one of the stories and its corresponding comprehension questions.
 All materials are available at TODO NEW REPO.
@@
 
+```{r}
+surps <- read_rds(here("Prep_code/natural_stories_surprisals.rds"))
+
+not_first <- surps %>% filter(Word_In_Sentence_Num!=1)
+
+#nrow(not_first)
+
+not_na <- not_first %>% filter(across(ends_with("surp"),~!is.na(.x)),
+                               !is.na(freq))
+
+not_ngram <- not_na %>%   filter(ngram_token_count==1)
+
+#nrow(not_ngram)
+
+not_other <- not_ngram %>% 
+filter(txl_token_count==1)%>% 
+  filter(grnn_token_count==1)
+#nrow(not_other)
+
+not_gpt <- not_other %>% filter(gpt_token_count==1)
+
+#nrow(not_gpt)
+
+```
 To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.
@@
 We got per-word surprisals for each of 4 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].
+
+We exclude the first word of every sentence because it has a different distractor, which leaves `r nrow(not_first)` words.
+Then we exclude words for which we don't have surprisal or frequency information, leaving `r nrow(not_na)` words.
+We additionally exclude words that were treated as being composed of multiple tokens.
+This is primarily words that have punctuation, where the punctuation is treated as a separate token.
+While surprisals should be additive, the summing the surprisals over these parts gives unreasonable responses sometimes.
+For instance, in one story the word king!' is given a surprisal of 64 by GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, 'What is the use of flying any higher?
+This victory is in the bag and I am king!').
+To avoid these outliers we exclude all words that a model treated as multi-token, leaving `r nrow(not_gpt)` words.
+(While GPT-2 using byte-pair encoding that can split up words into multiple parts, excluding words it split up only excludes 30 words that were not already excluded by other models.)
+
 For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
-
-We only included words that were treated as a single token by every model and for which we had frequency information.
-This mainly excluded words with punctuation as well as uncommon or proper nouns.
 We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
 We exclude words where mistakes occurred or which occurred after a mistake in the same sentence.
-We only analysed words where we had values for all predictors.
+We only analysed words where we had values for all predictors, which means that if the previous word was unknown to a model, this word will be excluded because we're missing values for a lagged predictor.
 
@@
 We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions.
-We collapse the reading times across subjects by modelling the mean RT for each word.
+To minimize the effect of repeated measures on confidence about inferred curve shapes, we collapse the reading times across subjects by modelling the mean RT for each word.
 
@@
 
-In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran.
-They recruited 181 participants, most of whom read 5 of the stories.
+In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran in 2011.
+They recruited 181 participants from Amazon Mechanical Turk, most of whom read 5 of the stories.
 After reading each story, each participant answered 6 binary-choice comprehension questions.
@@
 Note that these differences cannot be directly attributed to methods, as the participant populations differed.
+While both studies were conducted on Mturk, the quality of Mturk data has decreased from 2011 when the SPR was collected to 2020 when the A-maze was collected.
 
@@
 We built linear models with surprisal, frequency, length and surprisal x length and frequency x length effects from the current and previous words as predictors.
+We centered, but did not rescale, the predictors.
 
@@
 Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN.
-Overall, F-values here are much lower than for Maze.
+The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020.
+This contrasts with the Maze results, where the 5-gram model has the worst fit and does not provide additional predictive value to the other models.
+
+```{r}
+maze <- read_rds(here("Analysis/maze_model_compare.rds"))
+spr <- read_rds(here("Analysis/spr_model_compare.rds"))
+```
+
+As an overall measure of fit to data, we calculate multiple R-squared for the single surprisal source models for both A-maze and SPR.
+The models predict A-maze better with R-squared values ranging from `r maze %>% filter(Model=="5-gram") %>% pull(r_squared)` for the 5-gram model to `r maze %>% filter(Model=="GPT-2") %>% pull(r_squared)` for GPT-2.
+Whereas for SPR, the R-squared values range from from `r spr %>% filter(Model=="5-gram") %>% pull(r_squared)` to `r spr %>% filter(Model=="GPT-2") %>% pull(r_squared)`.
+This suggests that the effect size differences are not due merely to the larger overall reading time for A-maze, but that instead A-maze is more sensitive to surprisal and length effects.
 
@@
 
-This extreme incrementality may make the Maze task a good target for comparisons between human RTs and predictions of surprisal from neural language models because spillover effects would not need to be considered as much [@wilcoxPredictivePowerNeural2020].
+Actually the potential applicability and value is much broader, for any question where nailing down the locus on incremental processing difficulty is important, little spillover is very valuable.
+
+This extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty and thus benefits from the lack of spillover.
 It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more.
@@
 While the error-correction paradigm is crucial to running long materials, it also provides some benefits even on shorter materials when using A-maze with variable participant populations.
-The error-correction compensates for some of the shortcomings of A-Maze identified in @boyceMazeMadeEasy2020; poor distractors might still cause participants to make unavoidable mistakes, but they will still see the rest of the sentence, which may reduce frustration.
-Whether this also solves the data loss problem from the researcher perspective within a sentence depends on whether post-mistake data are high-quality and trustworthy; this is a hard-to-assess question of potential interest.
+The error-correction compensates for some of the shortcomings of A-maze identified in @boyceMazeMadeEasy2020; poor distractors might still cause participants to make unavoidable mistakes, but they will still see the rest of the sentence, which may reduce frustration.
+One questions that remains is whether this post-mistake data is useable, which depends on whether RTs from a few words after a mistake show any differences from RTs before any mistakes.
+Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is a  hard-to-assess question of potential interest.
 
@@
 \endgroup
+
+# Appendix 
+The beginning of one of the stories. This is the first 200 words of a 1000 word story.
+
+Tulip mania was a period in the Dutch Golden Age during which contract prices for bulbs of the recently introduced tulip reached extraordinarily high levels and then suddenly collapsed.
+At the peak of tulip mania in February sixteen thirty-seven, tulip contracts sold for more than ten times the annual income of a skilled craftsman.
+It is generally considered the first recorded economic bubble.
+The tulip, introduced to Europe in the mid sixteenth century from the Ottoman Empire, became very popular in the United Provinces, which we now know as the Netherlands.
+Tulip cultivation in the United Provinces is generally thought to have started in earnest around fifteen ninety-three, after the Flemish botanist Charles de l'Ecluse had taken up a post at the University of Leiden and established a botanical garden, which is famous as one of the oldest in the world.
+There, he planted his collection of tulip bulbs that the Emperor's ambassador sent to him from Turkey, which were able to tolerate the harsher conditions of the northern climate.
+It was shortly thereafter that the tulips began to grow in popularity.
+The flower rapidly became a coveted luxury item and a status symbol, and a profusion of varieties followed.
+
+The first 2 out of the 6 comprehension questions.
+
+When did tulip mania reach its peak? 1630's, 1730's
+
+From which country did tulips come to Europe? Turkey, Egypt

### 2021-08-11 Analysis/models.Rmd
tweaks

@@
 
-```{r participants}
+# Maze Prep
+
+```{r maze-participants}
 
@@
 
-<!--# Overview
-
-`r data %>% select(subject) %>% unique() %>% nrow()` participants read naturalistic stories from the natural stories corpus. Each participant read 1 story.
-
-We exclude
-
-- participants who do not report English as a native language (`r data_filt %>% select(subject) %>% unique() %>% nrow()` remaining)
-- participants who do not get 80% of the words correct (`r data_low_error %>% select(subject) %>% unique() %>% nrow()` remaining)
-- practice items (`r data_low_error %>% nrow()` words remaining)
-- words that were wrong or were within two after a mistake (`r data_error_free %>% nrow()` words remaining)
-- the first word of every sentence (didn't have a real distractor, RT is measured slightly differently) (`r data_no_first %>% nrow()` words remaining)
-- words with RTs <100 or >5000 (<100 we think is likely a recording error, or at least not reading the words at all, >5000 is likely getting distracted) (`r data_ready %>% nrow()` words remaining)
-
-Within the filtered data, each story was read between `r min(data_stories$n)` and `r max(data_stories$n)` times, for an average of `r mean(data_stories$n)`.
-
-We also do the analyses on only the words before mistakes (per sentence) (`r data_pre_error %>% nrow()` words)
-
-From the modelling side:
-(After attempts without doing this filtering) we only include words which are single token and known words in each of the models vocabularies.
-We also only include words with frequencies.
-This is roughly equivalent to excluding words with punctuation.
-
-We use as predictors:
-
-- length in characters of stripped word
-- unigram frequency of word.
-Frequencies for words are calculated using word_tokenize on the gulordava train data and counting up instances.
-(This tends to tokenize off punctuation, but is capitalization sensitive).
-Frequencies are represented as log2 of the expected occurances in 1 billion words.
-
-Surprisals are measured in bits.
-
-- ngram (5-gram KN smoothed)
-- GRNN
-- Transformer-XL
-
-For GAM models, we center length and frequency but not surprisal.
-We want surprisal interpretable, but we also will be plotting it (at least for the bootstrapping) at length and frequencies set to 0, so they need to be centered.
-(Not sure this last piece is actually true/matters).
--->
-
-# Maze Prep
 ```{r labels}
@@
 
-```{r maze}
+```{r}
 
@@
 
-```{r item-means}
+```{r maze-item-means}
 
@@
 
-```{r for gam}
-
-ngram_data <- select_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
-                                            ) %>% mutate(model="5-gram") 
-grnn_data <- select_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
-                                            )%>% mutate(model="GRNN")
-
-txl_data <- select_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
-                                            ) %>% mutate(model="TXL")
-
-gpt_data <- select_data %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
-                                            freq=freq_center, len=length_center,
-                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
-                                            ) %>% mutate(model="GPT-2")
-
+```{r maze-prep-gam}
+#on meaned data
 ngram_mean_data <- item_mean_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
@@
 
-all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
-  select(surprisal, prev_surp,model) %>% 
-  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
-  mutate(s=ifelse(name=="surprisal", "Current","Previous"))
-
 all_mean_data <- ngram_mean_data %>% union(grnn_mean_data) %>% union(txl_mean_data) %>% union(gpt_mean_data) %>% 
@@
 
-```{r formulae, echo=T}
+```{r maze-formulae, echo=T}
 #no hierarchical, has freq x len interaction
@@
 
-formula_heirarchy <- rt ~ s(surprisal, bs="cr", k=20)+
-                    s(surprisal, by=subject, bs="cr", k=20)+
-                     ti(freq, bs="cr") + 
-                     ti(len, bs="cr")+
-                     s(prev_surp, bs="cr", k=20)+
-                      s(prev_surp, by=subject, bs="cr", k=20)+
-                     ti(prev_freq, bs="cr")+
-                     ti(prev_len, bs="cr")
-
 ```
@@
 
+## Plots
+
 ```{r plots, eval=T}
-plot(gam_gpt_1, select=4)
-plot(gam_grnn_1, select=4)
-
-plot(gam_txl_1, select=4)
-plot(gam_ngram_1, select=4)
-
-```
-
-It looks like the effect is ~ 0 for most of the space where there's actually data, and it only goes off the rails in the region where there isn't data.
-
-Below is all words with centered frequency < -6 and centered_length > 4. Most of these are I think are recognizable as words. 
+message("GPT")
+plot(gam_gpt_1)
+message("GRNN")
+plot(gam_grnn_1)
+message("TXL")
+plot(gam_txl_1)
+message("NGRAM")
+plot(gam_ngram_1)
+
+```
+
+For the length: frequency interaction, it looks like the effect is ~ 0 for most of the space where there's actually data, and it only goes off the rails in the region where there isn't data.
+
+<!--Below is all words with centered frequency < -6 and centered_length > 4. Most of these are I think are recognizable as words. -->
 ```{r}
-select_data %>% select(word,  freq_center, length_center, sentence) %>% unique() %>% 
-  arrange(freq_center) %>% 
-  filter(freq_center< -6 & length_center > 4) %>% knitr::kable()
+# select_data %>% select(word,  freq_center, length_center, sentence) %>% unique() %>% 
+#   arrange(freq_center) %>% 
+#   filter(freq_center< -6 & length_center > 4) %>% knitr::kable()
 ```
@@
 
+# Maze LM with freq**2 effects
+
+Based on GAM viz, we try models with freq**2 effects on item mean data.
+
+(Tried this with orthagonal polynomials, but I get ridiculous predictions where both the effects for 1st and 2nd order are huge with huge std error, so probably cancelling out?).
+
+Instead, square and then center raw frequency. This also finds the effects, and seems a little absurd, since it's finding big effects, in opposite directions for squared and not squared.
+
+```{r}
+select_data <- read_rds(here("Data/maze_pre_error.rds")) %>%  select(-starts_with("past2"), -starts_with("past3")) %>% 
+  filter(across(everything(),~!is.na(.x)))
+
+item_mean_data <- select_data %>% 
+  group_by(across(c(-subject, -rt))) %>% 
+  summarize(rt=mean(rt)) %>% 
+  ungroup() %>% 
+  mutate(freq_sq_raw=freq**2,
+         past_freq_sq_raw=past_freq**2,
+         freq_sq=freq_sq_raw-mean(freq_sq_raw, na.rm=T),
+         past_freq_sq=past_freq_sq_raw - mean(past_freq_sq_raw, na.rm=T))
+
+no_surp_sq <- "rt ~  freq_sq+ freq_center*length_center + past_freq_sq + past_freq_center*past_length_center"
+no_surp_poly <- "rt ~ poly(freq_center,2) + length_center+ freq_center:length_center + poly(past_freq_center,2)+past_length_center + past_freq_center:past_length_center"
+ngram <- "+ ngram_center + past_ngram_center + ngram_center:length_center +   
+  past_ngram_center:past_length_center"
+grnn <- "+ grnn_center + past_grnn_center+ grnn_center:length_center +   
+  past_grnn_center:past_length_center"
+txl <- "+ txl_center + past_txl_center + txl_center:length_center +   
+  past_txl_center:past_length_center"
+gpt <- "+ gpt_center + past_gpt_center + gpt_center:length_center +   
+  past_gpt_center:past_length_center"
+
+lm(str_c(no_surp_sq,ngram), data=item_mean_data) %>% summary()
+
+lm(str_c(no_surp_sq,grnn), data=item_mean_data) %>% summary()
+lm(str_c(no_surp_sq,txl), data=item_mean_data) %>% summary()
+lm(str_c(no_surp_sq,gpt), data=item_mean_data) %>% summary()
+
+lm(str_c(no_surp_poly,ngram), data=item_mean_data) %>% summary()
+lm(str_c(no_surp_poly,grnn), data=item_mean_data) %>% summary()
+lm(str_c(no_surp_poly,txl), data=item_mean_data) %>% summary()
+lm(str_c(no_surp_poly,gpt), data=item_mean_data) %>% summary()
+```
+
 # SPR Prep
@@
 
-first <- d.raw %>% select(WorkerId,correct,item) %>% unique() %>% 
-  group_by(WorkerId,correct) %>% 
+first <- d.raw %>% select(WorkerId,item) %>% unique() %>% 
+  group_by(WorkerId) %>% 
   mutate(num=row_number()) %>% 
@@
 
-d.limited <- d.raw %>% inner_join(first, by=c("WorkerId","correct","item"))
+d.limited <- d.raw %>% inner_join(first, by=c("WorkerId","item"))
 
@@
 
-```{r for spr gam, eval=F}
+```{r for spr gam, eval=T}
 
@@
                      ti(length_center, bs="cr")+
+                     ti(freq_center,bs="cr")+
                      ti(freq_center,length_center, bs="cr")+
@@
 
-```{r spr meaned data, eval=F}
+```{r spr meaned data, cache=T, eval=T}
 gam_ngram_1 <- gam(spr_formula_interact, data=ngram_mean_spr, method="REML")
@@
 
+```
+
+```{r spr-gam-plot}
+message("GPT")
+plot(gam_gpt_1)
+message("GRNN")
+plot(gam_grnn_1)
+message("TXL")
+plot(gam_txl_1)
+message("NGRAM")
+plot(gam_ngram_1)
 ```
@@
 
-```{r spr brms, eval=F}
+```{r, eval=F}
 
@@
 summary(m_gpt)
-```
-
-```{r, eval=F}
-
-#data=readRDS("spr_pre_brms.rds")
-data=read_rds(here("Analysis/models/spr_pre_brms.rds"))
- priors <- c(
-      set_prior("normal(350,50)", class="Intercept"),
-      set_prior("normal(0, 5)", class="b"),
-      set_prior("exponential(.01)", class="sd", group="subject", coef="Intercept"),
-      set_prior("exponential(.01)", class="sd", group="Word_ID", coef="Intercept"),
-     # set_prior("exponential(1)", class="sd"),
-     # set_prior("lkj(2)",       class="cor"),
-      set_prior("exponential(.01)", class="sigma")
-      )
-run_model <- function(model_spec,data,file_name){
-  d <- data %>% filter(model==model_spec)
-  m1 <- brm(rt ~ surp_center * length_center + length_center * freq_center +
-                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
-                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
-                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
-              (1|subject)+(1|Word_ID),
-            data=d,
-            file=file_name,
-            prior=priors,
-            control=list(adapt_delta=.95))
-}
-
-#run_model("5-gram",data,"5-gram_spr")
-#run_model("GRNN",data,"grnn_spr")
-run_model("TXL",data,"txl_spr")
-#run_model("GPT-2",data,"gpt_spr")
-
 ```
@@
 ## Summaries of LMER 
-These have some heirarchical effects
+These have some hierarchical effects
 
@@
 ```
-
-Really struggling with the intercept for population & subject level -- maybe sd was intercept needs to be a lot bigger?
 
@@
     lm1 <- lm(str_c(no_surp,m1), data=item_mean_spr) %>% summary()
-    lm1$r.squared %>% round(2) %>% as.character()
+    lm1$r.squared %>% round(3) %>% as.character()
 }
@@
 ```
-
-# R-squared
+# SPR LM with freq**2 effects
 
 ```{r}
-spr <- read_rds(here("Analysis/spr_model_compare.rds"))
-
-maze <- read_rds(here("Analysis/maze_model_compare.rds"))
+select_spr <- read_rds(here("Data/clean_spr.rds")) %>% 
+  filter(across(everything(),~!is.na(.x))) %>% 
+  ungroup()
+
+item_mean_spr <- select_spr %>% 
+  group_by(across(c(-subject, -rt))) %>% 
+  summarize(rt=mean(rt)) %>% 
+  ungroup()
+
+no_surp_poly <- "rt ~ poly(freq_center,2) + length_center+ freq_center:length_center + poly(past_freq_center,2)+past_length_center + past_freq_center:past_length_center+
+poly(past2_freq_center,2)+past2_length_center + past2_freq_center:past2_length_center+
+poly(past3_freq_center,2)+past3_length_center + past3_freq_center:past3_length_center"
+ngram <- "+ ngram_center + past_ngram_center + past2_ngram_center + past3_ngram_center+
+ngram_center:length_center + past_ngram_center:past_length_center+ past2_ngram_center:past2_length_center+past3_ngram_center:past3_length_center"
+grnn <- "+ grnn_center + past_grnn_center +past2_grnn_center +past3_grnn_center +grnn_center:length_center + past_grnn_center:past_length_center+ past2_grnn_center:past2_length_center+past3_grnn_center:past3_length_center"
+txl <- "+ txl_center + past_txl_center +past2_txl_center + past3_txl_center +
+txl_center:length_center + past_txl_center:past_length_center+ past2_txl_center:past2_length_center+past3_txl_center:past3_length_center"
+gpt <- "+ gpt_center + past_gpt_center + past2_gpt_center + + past3_gpt_center+
+gpt_center:length_center + past_gpt_center:past_length_center+ past2_gpt_center:past2_length_center+past3_gpt_center:past3_length_center"
+
+lm(str_c(no_surp_poly,ngram), data=item_mean_spr) %>% summary()
+lm(str_c(no_surp_poly,grnn), data=item_mean_spr) %>% summary()
+lm(str_c(no_surp_poly,txl), data=item_mean_spr) %>% summary()
+lm(str_c(no_surp_poly,gpt), data=item_mean_spr) %>% summary()
 ```
@@
 
-Word level correlation between maze and spr
-```{r}
-
-summ_spr <- spr %>% group_by(Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt))
-
-summ <- maze %>% group_by(Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% left_join(summ_spr)
-
-ggplot(summ, aes(x=maze, y=spr))+
-  geom_point(alpha=.02)+
-  coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
-  labs(title="Word level correlation in ms",subtitle="Axes trimmed")
-
-cor(summ$spr,summ$maze, use="complete.obs")
-
-```
-
-sentence level correlation between maze and spr
-
-Note that a lot of this is driven purely by sentence length, which isn't interesting. 
-```{r}
-sentence <- summ %>% group_by(Story_Num, Sentence_Num) %>% 
-  summarize(maze=sum(maze)/1000,
-            spr=sum(spr)/1000)
-
-ggplot(sentence, aes(x=maze, y=spr))+geom_point(alpha=.1)+
-  labs(title="Sentence level correlation in s")
-
-cor(sentence$spr,sentence$maze, use="complete.obs")
-```
-
-Averaging across words at the sentence level removes the sentence-length effect.
-
-```{r}
-sentence_avg <- summ %>% group_by(Story_Num, Sentence_Num) %>% 
-  summarize(maze=mean(maze),
-            spr=mean(spr))
-ggplot(sentence_avg, aes(x=maze, y=spr))+geom_point(alpha=.1)+
-    coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
-  labs(title="Sentence level correlation in ms")
-
-cor(sentence_avg$spr,sentence_avg$maze, use="complete.obs")
-
-```
-
-As a baseline, we can do all the same things on split-halfs of each data set.
-We split the subjects in half.
-(Note that this means that each half has the same # of subjects, but may not have the same #s of each story)
-
-Within spr correlations are higher than spr-maze.
+We want to compare w/i Maze, w/i SPR and between Maze and SPR. Maze is the smaller dataset (63 subjects), so we split half for each story.
+
+For SPR, we subsample the 165 subjects down to 63 (to match the number of subjects in Maze), and again split half for each story.
+
+For Maze-SPR comparison, we compare each of the Maze half-SPR half and then average the correlations across these 4 pairs.
+
+### SPR-SPR
 
@@
 set.seed(42)
-spr_subs <- spr$WorkerId %>% unique()
-length(spr_subs)
+spr_subs <- spr %>% filter(!is.na(Story_Num)) %>% select(WorkerId, Story_Num) %>% unique() %>% sample_n(63) %>%  group_by(Story_Num) %>%
+  mutate(half=ifelse(row_number() <= .5*n(),"spr1", "spr2"))
 #180 participants
 
-spr1 <- sample(spr_subs, 90)
-
 spr_split <- spr %>%
-  mutate(half=ifelse(WorkerId %in% spr1, "spr1","spr2")) %>% 
+  inner_join(spr_subs) %>% 
   group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt)) %>% 
@@
 ```
-
-Same things for maze.
-Note that with fewer subjects, and given only pre-error data, there's a lot more chance of emptiness.
-I think the weak sentence average correlation is from places where some sentences didn't go as long in one of the halves (b/c low data).
+### Maze-Maze
 
@@
 set.seed(42)
-maze_subs <- maze$subject %>% unique()
-length(maze_subs)
-#180 participants
-
-maze1 <- sample(maze_subs, 31)
+maze_subs <- maze %>% select(subject, Story_Num) %>% unique() %>% sample_n(63)%>% group_by(Story_Num) %>%
+  mutate(half=ifelse(row_number() <= .5*n(),"maze1", "maze2"))
 
 maze_split <- maze %>% 
-  mutate(half=ifelse(subject %in% maze1, "maze1","maze2")) %>% 
+  left_join(maze_subs) %>% 
   group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% 
@@
 ```
+
+### Maze-SPR
+For maze-spr comparisons we take the 4 pairs (maze1 with spr1; maze1 with spr2, etc) We overplot all the combinations (not sure this is right), and take the average of the 4 correlations.
+
+```{r maze-spr}
+summ_spr <- spr %>% inner_join(spr_subs)%>% group_by(Story_Num, Sentence_Num, word_num,half) %>% summarize(spr=mean(rt, na.rm=T)) %>% rename(spr_half=half) %>% filter(!is.na(spr))
+
+summ <- maze %>% inner_join(maze_subs) %>% group_by(Story_Num, Sentence_Num, word_num, half) %>% summarize(maze=mean(rt, na.rm=T)) %>% rename(maze_half=half) %>% filter(!is.na(maze))%>% inner_join(summ_spr)
+
+summ11 <- summ %>% filter(maze_half=="maze1" & spr_half=="spr1")
+summ12 <- summ %>% filter(maze_half=="maze1" & spr_half=="spr2")
+summ21 <- summ %>% filter(maze_half=="maze2" & spr_half=="spr1")
+summ22 <- summ %>% filter(maze_half=="maze2" & spr_half=="spr2")
+
+ggplot(summ, aes(x=maze, y=spr))+
+  geom_point(alpha=.02)+
+  coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
+  labs(title="Word level correlation in ms",subtitle="Axes trimmed")
+
+mean(c(cor(summ11$spr,summ11$maze, use="complete.obs"),
+  cor(summ12$spr,summ12$maze, use="complete.obs"),
+  cor(summ21$spr,summ21$maze, use="complete.obs"),
+  cor(summ22$spr,summ22$maze, use="complete.obs")))
+
+sentence <- summ %>% group_by(Story_Num, Sentence_Num, maze_half, spr_half) %>% 
+  summarize(maze=sum(maze)/1000,
+            spr=sum(spr)/1000)
+
+sentence11 <- sentence %>% filter(maze_half=="maze1" & spr_half=="spr1")
+sentence12 <- sentence %>% filter(maze_half=="maze1" & spr_half=="spr2")
+sentence21 <- sentence %>% filter(maze_half=="maze2" & spr_half=="spr1")
+sentence22 <- sentence %>% filter(maze_half=="maze2" & spr_half=="spr2")
+
+ggplot(sentence, aes(x=maze, y=spr))+geom_point(alpha=.1)+
+  labs(title="Sentence level correlation in s")
+
+ mean(c(cor(sentence11$spr,sentence11$maze, use="complete.obs"),
+  cor(sentence12$spr,sentence12$maze, use="complete.obs"),
+  cor(sentence21$spr,sentence21$maze, use="complete.obs"),
+  cor(sentence22$spr,sentence22$maze, use="complete.obs")))
+
+sentence_avg <- summ %>% group_by(Story_Num, Sentence_Num, maze_half, spr_half) %>% 
+  summarize(maze=mean(maze),
+            spr=mean(spr))
+
+sentence_avg11 <- sentence_avg %>% filter(maze_half=="maze1" & spr_half=="spr1")
+sentence_avg12 <- sentence_avg %>% filter(maze_half=="maze1" & spr_half=="spr2")
+sentence_avg21 <- sentence_avg %>% filter(maze_half=="maze2" & spr_half=="spr1")
+sentence_avg22 <- sentence_avg %>% filter(maze_half=="maze2" & spr_half=="spr2")
+
+ggplot(sentence_avg, aes(x=maze, y=spr))+geom_point(alpha=.1)+
+    coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
+  labs(title="Sentence level correlation in ms")
+
+  mean(c(cor(sentence_avg11$spr,sentence_avg11$maze, use="complete.obs"),
+  cor(sentence_avg12$spr,sentence_avg12$maze, use="complete.obs"),
+  cor(sentence_avg21$spr,sentence_avg21$maze, use="complete.obs"),
+  cor(sentence_avg22$spr,sentence_avg22$maze, use="complete.obs")))
+```

### 2021-08-11 Papers/Paper/manuscript.Rmd
tweaks

@@
 
+As described in the Results section, we base all models only on data from participants who choose correctly on at least 80% of the words in the Maze task.
+
 ```{r}
@@
 
+## SPR-Maze correlation
+We additionally compare how correlated the Maze and SPR results are to each other, in comparison to within-Maze and within-SPR correlations.
+For Maze, within each story, we randomly split subjects into two halves.
+Within each half, we calculate a per-word average for each word and then a per-sentence average RT across word averages.
+We calculate a within-Maze correlation between these two halves.
+To avoid differences due to dataset size, we downsample the SPR data choosing a number of participants equal to the number we have for Maze.
+We then use the same procedure on this subset as for Maze to get a within-SPR correlation.
+For between Maze-SPR correlation, we take the average correlation across each of the 4 pairs of Maze half and SPR half.
+
 # Results
@@
 
-Many participants completed the Maze task with a high degree of accuracy and also answers the comprehension questions correctly.
+Many participants completed the Maze task with a high degree of accuracy and also answered the comprehension questions correctly.
 
@@
 We counted how many of the binary-choice comprehension questions each participant got right (out of 6).
-As seen in Figure \@ref(fig:errors)B, most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions.
+As seen in Figure \@ref(fig:errors)B, most participants who were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions.
 Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet.
@@
 Units are in ms.
-Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")
+Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.
+Interval is 2.5th quantile to 97.5th quantile of model draws.")
 ```
@@
 ```
+
+```{r}
+tokenization <- here("Data/SPR/all_stories.tok.txt") %>% 
+  read_delim(delim="\t") %>% 
+  mutate(Word_In_Story_Num=zone,
+         Story_Num=item)
+
+labs <- read_rds(here("Prep_code/natural_stories_surprisals.rds")) %>% 
+  left_join(read_delim(here("Materials/natural_stories_sentences.tsv"), delim="\t")) %>% 
+  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())
+
+spr <- read_rds(here("Data/SPR/first.rds")) %>% left_join(tokenization) %>% left_join(labs) %>% select(WorkerId,rt,Story_Num,Sentence_Num, word_num)
+
+maze <- read_rds(here("Data/maze_pre_error.rds")) %>% left_join(labs) %>% select(subject, rt, Story_Num, Sentence_Num, word_num)
+```
+
+```{r spr-split-half}
+set.seed(42)
+spr_subs <- spr %>% filter(!is.na(Story_Num)) %>% select(WorkerId, Story_Num) %>% unique() %>% sample_n(63) %>%  group_by(Story_Num) %>%
+  mutate(half=ifelse(row_number() <= .5*n(),"spr1", "spr2"))
+#180 participants
+
+spr_split <- spr %>%
+  inner_join(spr_subs) %>% 
+  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt)) %>% 
+  pivot_wider(names_from=half, values_from=spr)
+
+spr_sentence_avg <- spr_split%>% group_by(Story_Num, Sentence_Num) %>% 
+  summarize(spr1=mean(spr1),
+            spr2=mean(spr2))
+
+spr_spr <- cor(spr_sentence_avg$spr1,spr_sentence_avg$spr2, use="complete.obs")
+```
+
+```{r maze-split-half}
+set.seed(42)
+maze_subs <- maze %>% select(subject, Story_Num) %>% unique() %>% sample_n(63)%>% group_by(Story_Num) %>%
+  mutate(half=ifelse(row_number() <= .5*n(),"maze1", "maze2"))
+
+maze_split <- maze %>% 
+  left_join(maze_subs) %>% 
+  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% 
+  pivot_wider(names_from=half, values_from=maze)
+
+maze_sentence_avg <- maze_split%>% group_by(Story_Num, Sentence_Num) %>% 
+  summarize(maze1=mean(maze1),
+            maze2=mean(maze2))
+
+maze_maze <- cor(maze_sentence_avg$maze1,maze_sentence_avg$maze2, use="complete.obs")
+```
+
+```{r maze-spr}
+summ_spr <- spr %>% inner_join(spr_subs)%>% group_by(Story_Num, Sentence_Num, word_num,half) %>% summarize(spr=mean(rt, na.rm=T)) %>% rename(spr_half=half) %>% filter(!is.na(spr))
+
+summ <- maze %>% inner_join(maze_subs) %>% group_by(Story_Num, Sentence_Num, word_num, half) %>% summarize(maze=mean(rt, na.rm=T)) %>% rename(maze_half=half) %>% filter(!is.na(maze))%>% inner_join(summ_spr)
+
+sentence_avg <- summ %>% group_by(Story_Num, Sentence_Num, maze_half, spr_half) %>% 
+  summarize(maze=mean(maze),
+            spr=mean(spr))
+
+sentence_avg11 <- sentence_avg %>% filter(maze_half=="maze1" & spr_half=="spr1")
+sentence_avg12 <- sentence_avg %>% filter(maze_half=="maze1" & spr_half=="spr2")
+sentence_avg21 <- sentence_avg %>% filter(maze_half=="maze2" & spr_half=="spr1")
+sentence_avg22 <- sentence_avg %>% filter(maze_half=="maze2" & spr_half=="spr2")
+
+spr_maze <-   mean(c(cor(sentence_avg11$spr,sentence_avg11$maze, use="complete.obs"),
+  cor(sentence_avg12$spr,sentence_avg12$maze, use="complete.obs"),
+  cor(sentence_avg21$spr,sentence_avg21$maze, use="complete.obs"),
+  cor(sentence_avg22$spr,sentence_avg22$maze, use="complete.obs")))
+```
+
+(ref:scatter) Correlation between SPR and Maze data. RTs were averaged across participants per word and then averaged together within each sentence. RTs in ms.
+
+```{r scatter, out.height="20%", fig.width=6, fig.height=2, fig.pos="ht", fig.cap="(ref:scatter)"}
+spr_collapse <- spr %>% 
+  group_by(Story_Num, Sentence_Num,word_num) %>% 
+  summarize(spr=mean(rt, na.rm=T)) %>% 
+  group_by(Story_Num, Sentence_Num) %>%
+  summarize(spr=mean(spr))
+
+both <- maze %>% 
+  group_by(Story_Num, Sentence_Num, word_num) %>% 
+  summarize(maze=mean(rt, na.rm=T)) %>% 
+  group_by(Story_Num, Sentence_Num) %>% 
+  summarize(maze=mean(maze)) %>% 
+  inner_join(spr_collapse)
+ggplot(both, aes(x=maze, y=spr))+geom_point(alpha=.5, size=1)+labs(x="A-Maze", y="SPR")+coord_fixed(ratio=1)
+```
+
+An additional comparison between SPR and Maze is how correlated their reading times are; that is, do the methods pick up on the same effects.
+If so, we would expect that the sentences that take longer to read in one method also take longer in the other.
+We calculate the average RT at the sentence level (see Methods for details).
+The correlation between Maze and SPR is `r round(spr_maze,2)`, compared to `r round(spr_spr,2)` within SPR and `r round(maze_maze,2)` within Maze.
+See Figure \@ref(fig:scatter) for a visual comparison of overall Maze versus SPR RTs.
+
 # Discussion

### 2021-08-18 Papers/Paper/manuscript.Rmd
work on paper

@@
 ---
-title             : "A-maze of Natural Stories: Texts are comprehensible during the Maze task"
+title             : "A-maze of Natural Stories: Comprehension and surprisal in the Maze task TODO title"
 shorttitle        : "A-maze of Natural Stories"
@@
 abstract: |
-  We find support for the localization of reading time effects during the Maze task, as well as extending the range of materials Maze is suitable for.
-How long it takes to read a word in a sentence is reflective of how hard it is to identify and integrate the word in the surrounding context.
-Techniques that slow down the reading process and localize the processing time for each word are useful to understanding the time course of language processing.
+   TODO needs work
    A-maze is a new method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities.
@@
 Moreover, the  task yields useable reaction time data with word predictability effects that are linear in the surprisal of the current word, with little spillover effect from the surprisal of the previous word.
-This expands the types of effects that can be studied with A-maze, showing it to be a a versatile alternative to eye-tracking and self-paced reading.
+This expands the types of effects that can be studied with A-maze, showing it to be a a versatile alternative to eye-tracking and self-paced reading.We find support for the localization of reading time effects during the Maze task, as well as extending the range of materials Maze is suitable for.
+How long it takes to read a word in a sentence is reflective of how hard it is to identify and integrate the word in the surrounding context.
+Techniques that slow down the reading process and localize the processing time for each word are useful to understanding the time course of language processing.
 
@@
 ```
-# Intro
-It's remarkable how flexible we are when reading; while we do occasionally stumble when we read something unexpected, we often are able to read slightly unexpected things without a problem.
-However, these expectations shape how fast we read, even if we don't notice a stumble, unexpected words take longer to process as they force us to rebuild our burgeoning mental model of the sentence.
-Fortunately for fluent readers and unfortunately for studying language, this process of adjustment is very quick, which makes measures of reading time messy.
+# Introduction
 
-Measures of online reading are one way to understand language and how the mind processes language.
-Many theories of language structure and language processing ground out in predictions about the difficulty of processing words [@bartekSearchOnlineLocality2011].
-For instance, the subject v object relative debate includes theories that make fine-grained predictions about which word is how slow -- this needs localized methods to adjudicate it [@grodnerConsequencesSerialNature2005; @staubEyeMovementsProcessing2010; @traxlerProcessingSubjectObject2002].
-Other theories such as noisy channel processing also need support from localized word-by-word results [@levyIntegratingSurprisalUncertaininput2011].
+Sometimes we stumble in reading, where we expected a sentence to go in one direction and it did not, and then are at a parsing dead-end where we have to read a sentence again.
+The words that cause stumbles are the outliers on a distribution of words difficulty to read, within fluent reading, there is variation in how easy words are to read and process in their context.
+We're able to read many long or surprising words without noticing a problem, but these less expected words do take longer to process as we rebuild our burgeoning mental model of the sentence.
+Fortunately for fluent readers and unfortunately for studying language, this process of adjustment is very quick, with millisecond scale differences in reading time between words.
 
-Incremental processing methods such as self-paced reading or eye-tracking measure how long someone spends looking at one word before moving on and use that as a proxy for how difficult word was in context.
-How unexpected a word is correlates with how long it takes to read it and move on [@raynerEffectsFrequencyPredictability2004; @klieglLengthFrequencyPredictability2004].
-However, the two major methods of measuring incremental processing both suffer from imprecise localization.
-In eye-tracking, people read naturally which involves skipping words, jumping ahead and looking back, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982; @raynerEyeMovementsReading1998].
-Even when reading order is controlled in self-paced reading, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006].
+We don't have a thorough understanding of how the mind processes language, but we can learn something about what is going on from measuring how long it takes in different circumstances.
+This relies on a linking hypothesis that processing time reflects something about how much processing is going on, even if we don't know whether this is character recognition, memory retrieval, or parsing.
+This is enough to compare theories that predict the relative difficulty, and thus relative processing time, of two words in different contexts.
 
-One effect of this lack of localization is that reading time for a word is dependent not only on how unexpected a word is, but also how unexpected the previous word is.
-This is an indication of spillover from the previous word.
-It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
-Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization.
-In addition to suprisal predicting RT, word length and word's overall frequency are also often found to be predictive [@klieglLengthFrequencyPredictability2004].
+For instance, different theories about what makes object relative clauses harder to understand than subject relative clauses make different predictions about which words are the loci of the overall difficulty and slower reading times associated with object relatives [@grodnerConsequencesSerialNature2005; @staubEyeMovementsProcessing2010; @traxlerProcessingSubjectObject2002].
+Measures of reading time can also inform theories about the time course of processing (i.e.
+which steps are parallel versus serial, @bartekSearchOnlineLocality2011) or the functional form of relationships between word characteristics and processing time [@smithEffectWordPredictability2013].
 
-An alternative method that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing [@freedman85; @forsterMazeTaskMeasuring2009].
+Some of these theories rely on being able to attribute processing slowdowns to a particular word.
+Determining that object relatives are overall slower that subject relatives is easy, even an imprecise measure of reading time will determine that the same set of words in a different order took longer to read on a sentence level.
+However, many language processing theories make specific (and contrasting) theories about which words in a sentence should be harder to process.
+To adjudicate these theories, we want methods that are *localized*, where it is easy to determine which word is responsible for an observed slow-down in reading time.
+Ideally, a longer reading time on a word would be an indication of that word's increased difficulty, and not the lingering signal of a prior word's increased difficulty.
+When the signal isn't localized, advanced analysis techniques may be required to disentangle the slow-downs [@shainDeconvolutionalTimeSeries2018].
+
+## Incremental processing methods
+
+Behavioral methods that measure how long it takes to read each word in a sentence are referred to as incremental processing methods, and their dependent measures are called reading times or reaction times, both abbreviated RT.
+The two most commonly used methods are eye-tracking and self-paced reading, but both of these suffer from a lack of localization.
+In eye-tracking, participants read a text on a screen naturally, while their eye-movements are recorded [@raynerEyeMovementsReading1998].
+The downside of this natural reading is that people often skip short words and look ahead or look back as they read, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982].
+Self-paced reading is a more controlled process where a participant sees one word on screen at a time and presses a button to see the next word instead.
+Even in this method, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006].
+Both these methods are prone to spillover, as the time to plan motor movements like saccades or button presses is substantial compared to the time it takes to recognize a word, these movements are often initiated before a word is fully processed, so by the time any late processing difficulties occur, the reader may already be several words along.
+
+### the Maze task
+An alternative method that is designed to increase localization at the expense of naturalness is the Maze task [@freedman85; @forsterMazeTaskMeasuring2009].
 In the Maze task, participants see two words at a time, a correct word that continues the sentence, and a distractor which does not.
-Participants must choose the correct word, and their reaction time (RT) is the dependent measure.
-If participants make a mistake, the sentence discontinues.
-Two versions of the maze task exist: lexical or L-maze where the distractors are nonce words and grammatical or G-maze where the distractors are real words of the language that don't fit in the context of the sentence so far.
-Theoretically, participants must fully integrate each word into the sentence in order to confidently select it.
-This idea is supported by studies finding strongly localized effects for G-maze which is more sensitive than L-maze [@witzelComparisonsOnlineReading2012a].
+Participants must choose the correct word, and their time to selection is treated as the reaction time (RT).
+@forsterMazeTaskMeasuring2009 introduces two versions of the Maze task: lexical L-maze where the distractors are nonce words and grammatical G-maze where the distractors are real words that don't fit with the context of the sentence so far.
+Theoretically, participants must fully integrate each word into the sentence in order to confidently select it; this may require mentally reparsing previous material in order to allow the integration and selection of a disambiguating word.
+@forsterMazeTaskMeasuring2009 call this forced incremental processing to distinguish from other incremental processing methods where words can be passively read before later committing to a parse.
+This idea of strong localization is supported by studies finding strongly localized effects for G-maze which is more sensitive than L-maze [@witzelComparisonsOnlineReading2012a].
 
-The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitious words as distractors for each spot of each sentence; this may explain why the Maze task was not widely adopted.
-@boyceMazeMadeEasy2020 demonstrate a way to automatically generate Maze distractors by using NLP language models to find words that are high-surprisal in the context of the target sentence.
-The quality of these A-maze distractors is not up to that of hand-generated G-maze distractors, but @boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzelComparisonsOnlineReading2012a.
+The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitious words as distractors for each spot of each sentence; this may explain why the Maze task has not been widely adopted.
+@boyceMazeMadeEasy2020 demonstrate a way to automatically generate Maze distractors by using NLP language models to find words that are high-surprisal in the context of the target sentence, and thus likely to be judged infelicitous by human readers.
+@boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzelComparisonsOnlineReading2012a.
 A-maze outperformed L-maze and an SPR control in detecting and localizing expected slowdown effects.
@@
 
+Another recent variant of the Maze task is interpolated or I-maze, which uses a mix of real word distractors (generated via the A-maze process) and nonce word distractors [@vaniUsingInterpolatedMaze2021].
+The presence of real word distractors encourages close attention to the sentential context, while nonce words can be used as distractors where the word in the sentence is itself ungrammatical or highly unexpected.
+@vaniUsingInterpolatedMaze2021 used I-maze to compare object- and subject- relative clauses and @wilcoxTargetedAssessmentIncremental2021 used it to compare differences in RT for pairs of grammatical and ungrammatical sentences NLP language models predictions.
+
+## How to compare localization & effects, (old school freq/length), define surprisal and lang models
+TODO!!!
+One measure of how localized a method is is to look at how much overall effects on each word stay on that word, versus appearing on downstream words. 
+There are a few general effects for how properties of words tend to effect their reading time. 
+The classic pair are frequency and length -- longer and less common words take longer to process. @klieglLengthFrequencyPredictability2004
+Another measure is predictability, or how expected a word is in context, with more predictable words being read faster @raynerEffectsFrequencyPredictability2004.
+A word can have a high surprisal for lots of reasons -- it could be low frequency, or semantically unexpected, or the start of a low-frequency syntactic construction, or a word that disambiguates prior words to the less common parse.
+Many targeted effects of interest contribute to how predictable or unpredictable a word is.
+If effects of predictability are strong and well-localized, it's a good sign.
+Because predictability is made up of all these components, how do we assess how predictable a word is?
+One way of doing this is used language models that are trained to predict the next word.
+Predictability is often measured in terms of bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time etc).
+RTs for eye-tracking and SPR show a linear relationship between a words surprisal and RT It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
+ Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization.
+
+## why current experiment
+TODO!!!
+Maze is great, but there hasn't been a study looking at these coarse effects.
+We're interested in whether there's spillover ,and how similar it is to SPR.
+It's only used on targeted/constructed materials of independent sentences.
+
+So we want to do it on some longer natural text and look for these underlying patterns. & see how well participants can read and understand longer texts.
+
+The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers.
+At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions.
+Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants.
+If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on a wide variety of naturalistic texts.
+The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
+
+We tweak the A-maze task to accomodate these longer passages and then have participants read them and also compare with  @futrellNaturalStoriesCorpus2020 collected self-paced reading time data on this corpus, which we compare the A-maze results to.
+
+Using this tweak, we test A-maze on the passages from the Natural Stories corpus.
+Participants were able to read and understand these long passages using A-maze with the error correction paradigm, and their RT profiles show a similar pattern to SPR but with less noise and spillover
+
+# Error-correction maze
+<!-- GOOD -->
+(ref:diagram-cap) Schematic of error-correction Maze.
+A participant reads a sentence word by word, choosing the correct word at each time point (selections marked in blue ovals).
+When they make a mistake, an error message is displayed, so they try again and continue with the sentence.
+
+```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:diagram-cap)"}
+knitr::include_graphics(here("Papers/maze_diagram_cropped.pdf"))
+```
+*mention the whole terminate on errors thing*
 A-maze is a potentially powerful addition to the psycholinguists toolkit.
@@
 
-Using this tweak, we test A-maze on the passages from the Natural Stories corpus.
-The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers.
-At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions.
-Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants.
-If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on a wide variety of naturalistic texts.
-The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
-@futrellNaturalStoriesCorpus2020 collected self-paced reading time data on this corpus, which we compare the A-maze results to.
-Participants were able to read and understand these long passages using A-maze with the error correction paradigm, and their RT profiles showed a linear relationship to the surprisal of the words.
-
-# Error-correction maze
-<!-- GOOD -->
-(ref:diagram-cap) Schematic of error-correction Maze.
-A participant reads a sentence word by word, choosing the correct word at each time point (selections marked in blue ovals).
-When they make a mistake, an error message is displayed, so they try again and continue with the sentence.
-
-```{r diagram, out.height="25%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:diagram-cap)"}
-knitr::include_graphics(here("Papers/maze_diagram_cropped.pdf"))
-```
 One advantage of the Maze task is that it forces incremental processing and automatically excludes inattentive participants by terminating a sentence when a participant makes an error.
@@
 
+```{r}
+
+data_long <- data_filt %>% filter(type!="practice") %>% filter(rt>5000)
+data_no_prac <- data_filt %>% filter(type!="practice")
+```
+
 Participant accuracy reflects both how well participants can navigate the task and what quality the auto-generated distractors are.
 We calculated the per-word error rate for each participant and graphed it against their average reaction time.
-(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) make relatively few errors, with some reaching 99% accuracy.
+(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded; this excluded `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) make relatively few errors, with some reaching 99% accuracy.
 This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story.

### 2021-08-18 Papers/Paper/manuscript.Rmd
more edits

@@
 
+TODO overall: go through Roger's comments!
+TODO fix tense!
+
 Sometimes we stumble in reading, where we expected a sentence to go in one direction and it did not, and then are at a parsing dead-end where we have to read a sentence again.
@@
 The downside of this natural reading is that people often skip short words and look ahead or look back as they read, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982].
-Self-paced reading is a more controlled process where a participant sees one word on screen at a time and presses a button to see the next word instead.
+Self-paced reading (SPR) is a more controlled process where a participant sees one word on screen at a time and presses a button to see the next word instead.
 Even in this method, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006].
@@
 
-## How to compare localization & effects, (old school freq/length), define surprisal and lang models
-TODO!!!
-One measure of how localized a method is is to look at how much overall effects on each word stay on that word, versus appearing on downstream words. 
-There are a few general effects for how properties of words tend to effect their reading time. 
-The classic pair are frequency and length -- longer and less common words take longer to process. @klieglLengthFrequencyPredictability2004
-Another measure is predictability, or how expected a word is in context, with more predictable words being read faster @raynerEffectsFrequencyPredictability2004.
-A word can have a high surprisal for lots of reasons -- it could be low frequency, or semantically unexpected, or the start of a low-frequency syntactic construction, or a word that disambiguates prior words to the less common parse.
-Many targeted effects of interest contribute to how predictable or unpredictable a word is.
-If effects of predictability are strong and well-localized, it's a good sign.
-Because predictability is made up of all these components, how do we assess how predictable a word is?
-One way of doing this is used language models that are trained to predict the next word.
-Predictability is often measured in terms of bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time etc).
-RTs for eye-tracking and SPR show a linear relationship between a words surprisal and RT It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
- Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization.
+## Frequency, length, and surprisal effects
+One way of assessing how much a method localized RT effects is to look at how strongly properties of a word correlate with the RT of that word compared with RTs of downstream words later in the sentence.
+A couple of properties known to influence reading time are a word's length and overall frequency in a language, as longer and less common words take longer to process [@klieglLengthFrequencyPredictability2004].
+Another measure is predictability, or how expected a word is given the context, with more predictable words being read faster [@raynerEffectsFrequencyPredictability2004].
+A word can have low predictability for a number of reasons: it could be low frequency, semantically unexpected, the start of a low-frequency syntactic construction, or a word that disambiguates prior words to the less common parse.
+Many targeted effects of interest are essentially looking at specific features that contribute to how predictable or unpredictable a word is.
+This means that how good a method is at detecting and localizing effects of predictability is a key aspect of how useful it is as measure of incremental processing.
+With al these factors potentially contributing to predictability, how is predictability as a whole measured?
+The typical method is to use language models that are trained on large corpora of language to predict what word comes next in a sentence.
+A variety of pre-trained models exist, that vary in their internals, but all of them will generate assessments of how predictable words are.
+Predictability is often represented in terms of bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time etc).
+The functional form of the relationship between RTs from eye-tracking and SPR corpora and the predictability of the words is linear in terms of surprisal [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
+Due to spillover effects, this linear relationship also holds between the surprisal of a previous word and the RT on the current word.
 
-## why current experiment
-TODO!!!
-Maze is great, but there hasn't been a study looking at these coarse effects.
-We're interested in whether there's spillover ,and how similar it is to SPR.
-It's only used on targeted/constructed materials of independent sentences.
-
-So we want to do it on some longer natural text and look for these underlying patterns. & see how well participants can read and understand longer texts.
+## Current experiment
+The Maze task has thus far only been used on constructed sentences focusing on targeted effects and not on the sorts of long naturalistic passages used to assess the relationship between RT and surprisal.
+We want to test whether participants can read and understand stories while doing the demanding Maze task and also whether the RT profiles from the Maze task are similar to those from other methods.
 
@@
 At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions.
-Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants.
-If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on a wide variety of naturalistic texts.
+Taken together, these properties are a severe test of A-maze, as these features may make it harder to auto-generate appropriate distractors and require focus from participants.
+If participants can succeed at the Maze task on Natural Stories, we think they are likely to succeed on a wide variety of naturalistic texts.
 The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
 
-We tweak the A-maze task to accomodate these longer passages and then have participants read them and also compare with  @futrellNaturalStoriesCorpus2020 collected self-paced reading time data on this corpus, which we compare the A-maze results to.
-
-Using this tweak, we test A-maze on the passages from the Natural Stories corpus.
-Participants were able to read and understand these long passages using A-maze with the error correction paradigm, and their RT profiles show a similar pattern to SPR but with less noise and spillover
+We tweak the A-maze task to accomodate these longer passages and then have participants read the passages in the Maze.
+We compare our A-maze results with SPR data collected on this corpus by @futrellNaturalStoriesCorpus2020.
+We find that participants were able to read and understand these long passages using A-maze, and their RT profiles show a similar pattern to SPR but with less noise and spillover
 
@@
 ```
-*mention the whole terminate on errors thing*
-A-maze is a potentially powerful addition to the psycholinguists toolkit.
-However, like the other Maze tasks it has been limited in its application to single-sentence items probing minimal comparisons in constructed sentences.
-This limits its useful, as some important questions such as comparing human data to language models or studying discourse effects require processing times over multi-sentence passages.
-Therefore, it's important to expand the Maze task to these types of materials and verify that it finds comparable patterns to other methods.
 
-While the issue of needing to generate distractors for long passages is solved with A-maze, another problem with Maze remains.
-In particular, because Maze tasks discontinue after participants make mistakes, the farther into an item a word is, the fewer participants see it.
-This makes it hard to run long materials using Maze, and prevents Maze from being used on long text passages where SPR or eye-tracking could be used.
-We resolve this issue by introducing a new paradigm for running Maze tasks where participants correct their errors and continue reading rather than discontinuing after a mistake.
+One of the benefits of the Maze task is that it forces incremental processing by having participants make an active choice about what the next word is.
+But what if they choose incorrectly?
+In the traditional Maze paradigm, there isn't a way to handle this; a mistake indicates they're not understanding the sentence properly, so the sentence ends, and the participant moves on to the next item [@forsterMazeTaskMeasuring2009].
+An advantage of this is that participants who contribute RT data are very likely to have understand the sentence up to that point (it's hard to choose correctly in G-maze otherwise).
+This contrasts with other methods, where determining whether participants are paying attention usually requires separate comprehension check questions, that are usually not used for Maze.
 
-One advantage of the Maze task is that it forces incremental processing and automatically excludes inattentive participants by terminating a sentence when a participant makes an error.
-Thus, only participants who have processed the sentence and are paying attention contribute data at critical regions later in the sentence.
-However, this means we don't have data after a participant makes a mistake in an item.
+However, terminating sentences on errors means that we don't have data after a participant makes a mistake in an item.
 In traditional G-maze tasks, with hand-crafted distractors and attentive participants, this is a small issue.
@@
 
-This situation could be improved by auto-generating better distractors or hand-replacing problematic ones, but that does not solve the fundamental problem.
+This situation could be improved by auto-generating better distractors or hand-replacing problematic ones, but that does not solve the fundamental problem with long items.
 Well-chosen distractors and attentive participants will reduce the error rate, but the error rate will still compound over long materials.
-For instance, with a 1% error rate, `r round(.99**15*100)`% would complete each 15-word sentence, but only `r round(.99**50*100)`% of participants would complete a 50 word vignette, and `r round(.99**200*100)`% a 200 word passages.
+For instance, with a 1% error rate, `r round(.99**15*100)`%  of participants would complete each 15-word sentence, but only `r round(.99**50*100)`% would complete a 50 word vignette, and `r round(.99**200*100)`% would complete a 200 word passage.
 In order to run longer materials, we need something to do when participants make a mistake, other than terminate the entire item.
@@
 
-This variant of Maze expands the types of materials that can be used with maze to include arbitrarily long passages and cushions the impact of occasional problematic distractors.
+This variant of Maze expands the types of materials that can be used with Maze to include arbitrarily long passages and cushions the impact of occasional problematic distractors.
+Error-correction Maze is a change in experimental procedure, and is independent of what types of distractors are used.
+This error-correction presentation is used here with A-maze, but would also work with G-maze or I-maze.
 
@@
 ## Materials
-We took the texts of the Natural Stories corpus [@futrellNaturalStoriesCorpus2020], split them into sentences, and ran the sentences through the A-maze generation process.
-We follow the A-maze generation process outlined in @boyceMazeMadeEasy2020, although we use an updated version of the codebase with fixes some issues identified in that paper (code at https://github.com/vboyce/Maze).
-Additionally, the capability to appropriately handle a wider variety of punctuation (needed for this corpus) was added.
-We took the auto-generated distractors as they were, without checking them for quality.
+We used the texts from the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and generated A-maze distractors for them.
 We used the original comprehension questions provided in the Natural Stories corpus.
@@
 
+To generate distractors, we first split the corpora up into sentences, and then ran the sentences through the A-maze generation process.
+We used an updated version of the codebase from @boyceMazeMadeEasy2020.
+This newer version had the capability to match the greater variety of punctuation present in this corpus (updated auto-generation code at https://github.com/vboyce/Maze).
+We took the auto-generated distractors as they were, without checking them for quality.
+
 ## Participants
-We recruited 100 participants from Amazon Mechanical Turk, and paid each participant $3.50 for roughly 20 minutes of work.
+We recruited 100 participants from Amazon Mechanical Turk in April 2020, and paid each participant $3.50 for roughly 20 minutes of work.
 We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
@@
 Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading the main story in the A-maze task.
-After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, saw an debriefing and were given a code to enter for payment.
+After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, and were debriefed, before getting a code to enter for payment.
 The experiment was implemented in Ibex (https://github.com/addrummond/ibex) and the experimental code is available at REPO.
@@
 
-We conducted data processing and analyses using `r cite_r("r-references.bib", withhold=F, pkgs=package_list)`.
+TODO top level description
 
+In order to measure the relationship between a word's properties and it's RT, we fit models using surprisal, frequency, and length as predictors of RT.
+We considered these predictors from both the current and past word to look for the possibility of spill over effects.
+We fit generalized additive models (GAMs) to look at the functional form of the surprisal-RT relationship and linear models (LMs) to look at the size of the effects.
 As described in the Results section, we base all models only on data from participants who choose correctly on at least 80% of the words in the Maze task.
+
+ TODO FIX ME We conducted data processing and analyses using `r cite_r("r-references.bib", withhold=F, pkgs=package_list)`.
 
@@
 ```
-To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.
+### Predictors 
+We created a set of predictor variables of frequency, word length, and surprisals from 4 language models.
 For length, we used the length in characters excluding end punctuation.
 For unigram frequency, we tokenized the training data from @gulordava18 and tallied up instances.
-We then used the log2 frequency of the expected occurances in 1 billion words as the model predictor, so higher values indicate higher log frequencies.
+We then rescaled it to be the log2 frequency of expected occurrences in 1 billion words as the model predictor, so higher values indicate higher log frequencies.
 We got per-word surprisals for each of 4 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].
+These models cover a range of common architectures.
+We can infer robustness if models provide similar results, but also compare between models to look at how well each model fits human data.
+For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
 
-We exclude the first word of every sentence because it has a different distractor, which leaves `r nrow(not_first)` words.
-Then we exclude words for which we don't have surprisal or frequency information, leaving `r nrow(not_na)` words.
-We additionally exclude words that were treated as being composed of multiple tokens.
-This is primarily words that have punctuation, where the punctuation is treated as a separate token.
-While surprisals should be additive, the summing the surprisals over these parts gives unreasonable responses sometimes.
-For instance, in one story the word king!' is given a surprisal of 64 by GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, 'What is the use of flying any higher?
-This victory is in the bag and I am king!').
-To avoid these outliers we exclude all words that a model treated as multi-token, leaving `r nrow(not_gpt)` words.
+### Exclusions
+We exclude the first word of every sentence because it had an x-x-x distractor, which leaves `r nrow(not_first)` words.
+We exclude words for which we don't have surprisal or frequency information, leaving `r nrow(not_na)` words.
+We additionally exclude words that any model treated as being composed of multiple tokens.
+While surprisals should be additive, the summing the surprisals over these tokens gives some unreasonable responses.
+For instance, in one story the word king!\' is given a surprisal of 64 by GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, \'What is the use of flying any higher?
+This victory is in the bag and I am king!\').
+To avoid these outliers we exclude all words that any model treated as multi-token, leaving `r nrow(not_gpt)` words.
+This primarily excludes words with punctuation.
 (While GPT-2 using byte-pair encoding that can split up words into multiple parts, excluding words it split up only excludes 30 words that were not already excluded by other models.)
 
-For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
 We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
@@
 
+### Model specification
 For generalized additive models, we centered but did not rescale the length and frequency predictors, but left surprisal uncentered for interpretability.
@@
 To minimize the effect of repeated measures on confidence about inferred curve shapes, we collapse the reading times across subjects by modelling the mean RT for each word.
+TODO edit for bootstrapping
 
@@
 We used predictors for the current and past word and centered all effects.
+TODO cite LM
 
@@
 As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020].
+
 For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct.
@@
 We calculated the per-word error rate for each participant and graphed it against their average reaction time.
-(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded; this excluded `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) make relatively few errors, with some reaching 99% accuracy.
+(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded; this excluded `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) made relatively few errors, with some reaching 99% accuracy.
 This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story.
-These careful participants took around 1 second for each word selection, which is much slower than other paradigms.
+These careful participants took around 1 second for each word selection, which is much slower than in eye-tracking or SPR.
 Another cluster of participants (in red) sped through the task, seemingly clicking randomly.
@@
 
-We use task performance as our exclusion metric because it is more fine-grained and only analyze data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
+We use task performance as our exclusion metric for A-maze because it is more fine-grained and only analyze data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
 For the SPR comparison, we follow @futrellNaturalStoriesCorpus2020's criteria and exclude participants who got less than 5 of the questions correct.
@@
 
+TODO remind why we care so much about surprisal
+
 We fitted generalized additive models to test whether the RTs from the Maze experiment showed a linear relationship with surprisal and whether the effect was limited the current word or had spilled over from the prior word.
 For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.
+(See Methods for details on exclusions.)
 
@@
 For each of the 4 language models used, there is a linear relationship between current word surprisal and RT (at least when there is enough data).
-There is no relationship between previous word surprisal and RT.
+The relationship between previous word surprisal and RT is much flatter.
 
@@
 All of the models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data.
-All of the models show a flat relationship between previous word surprisal and RT.
-This is a sign of localization as the previous word's surprisal is not affecting RT, only the word's own suprisal is.
+All of the models show a flatter relationship between previous word surprisal and RT.
+This is a sign of localization as the previous word's surprisal is not affecting RT much.
 The linear relationship matches that found with other methodologies.
@@
 As we can see in Table \@ref(tab:pre-error), we find large effects of surprisal and length, but minimal effects of frequency.
-These effects are larger than what is usually reported in other methods CITE, but this could be due to the overall slowness of the method.
 The lack of frequency effects is somewhat surprising, but consistent with @shainLargescaleStudyEffects2019.
@@
 We assess the benefits of adding each model's predictions as a second surprisal source to see which models pick up on information not contained in another model.
-In terms of log likelihoods, GPT-2 is better than GRNN is better than TXL is better than 5-gram.
-Model comparisons tell a similar story; GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complemetary information over GPT-2.
+GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complemetary information over GPT-2.
 TXL provides a lot over 5-gram, and 5-gram provides little over any model.
+Log likelihood, which is a measure of fit to data, shows  a similar hierarchy, as GPT-2 is better than GRNN is better than TXL is better than 5-gram.
 ```{r maze-compare}
@@
 Each row shows the additional predictive value gained from adding that model to another model.
-F values and p values from anova tests are reported.
+F values and p values from anova tests between 1-surprisal-source and 2-source models are reported.
 We also report log likelihoods of models with only one surprisal source.")
@@
 Similarly to the model comparison for Maze, we also conducted a model comparison on the SPR data shown in Table \@ref(tab:spr-compare).
+Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN.
 In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small.
-Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN.
 The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020.
@@
 
+## maze works
+
+## strong effects of suprisal, mostly localized
+
+## better than SPR
+
+## correlated w SPR, probably the same effects, just stronger & less noisy
+
+## Why you care - broad applications that might want this method
+
+## shill for error correction
 We introduced a tweak to the paradigm for displaying Maze tasks that makes it useable for multi-sentence materials.

### 2021-08-19 Analysis/models.Rmd
Roger added bootstrapping to GAMs

@@
 
+```{r bootstrap code}
+require(OpenImageR)
+gam_boot <- function(dat,formula,data_frames_for_prediction,N=100,method="REML",...) {
+  models <- lapply(1:N,function(seed) {
+    set.seed(seed)
+    #dat[[weights_name]] <- tabulate(sample(1:nrow(dat),floor(nrow(dat)/5),replace=TRUE),nbins=nrow(dat))
+    dat$mgcv_weights <- tabulate(sample(1:nrow(dat),nrow(dat),replace=TRUE),nbins=nrow(dat))
+    #print(head(dat$mgcv_weights))
+    #print(head(dat[[weights_name]]))
+    #m <- gam(formula,data=dat,weights=dat[[weights_name]],...)
+    m <- gam(formula,data=dat,weights=mgcv_weights,method=method,...)
+    #print(m$deviance)
+    return(m)
+  })
+  results <- lapply(names(data_frames_for_prediction),function(term) {
+    predictions <-sapply(models,function(m)
+      predict(m,newdata=data_frames_for_prediction[[term]],type="terms",terms=term))
+    #print(dim(predictions))
+    return(predictions)
+  })
+  #print(length(results))
+  names(results) <- names(data_frames_for_prediction)
+  return(results)
+}
+
+summarize_gam_boot_results <- function(gam_boot_results,data_frames_for_prediction,term_names,model_name,point_estimation_function=mean,quantiles=c(0.025,0.975)) {
+  result <- lapply(1:length(data_frames_for_prediction),function(i) {
+    #print(term_names[[i]])
+    res <- tibble(x=data_frames_for_prediction[[i]][[term_names[[i]]]],
+                  rt=apply(gam_boot_results[[i]],1,point_estimation_function),
+                  CI_lower=apply(gam_boot_results[[i]],1,function(x) quantile(x,0.1)),
+                  CI_upper=apply(gam_boot_results[[i]],1,function(x) quantile(x,0.9)),
+                  model=model_name,
+                  term=term_names[[i]])
+    names(res)[names(res)=="x"] <- term_names[[i]]
+    return(res)
+  })
+  return(result)
+}
+
+dat_for_surprisal_prediction <- expand_grid(surprisal=seq(0,25,by=0.01),freq=0,len=0,prev_surp=0,prev_freq=0,prev_len=0)
+dat_for_freq_prediction <- expand_grid(surprisal=0,freq=seq(-9.5,7,by=0.01),len=0,prev_surp=0,prev_freq=0,prev_len=0)
+dat_for_length_prediction <- expand_grid(surprisal=0,freq=0,len=seq(-3.4,10.6,by=0.01),prev_surp=0,prev_freq=0,prev_len=0)
+dat_for_prev_surp_prediction <- expand_grid(surprisal=0,freq=0,len=0,prev_surp=seq(0,25,by=0.01),prev_freq=0,prev_len=0)
+dat_for_prev_freq_prediction <- expand_grid(surprisal=0,freq=0,len=0,prev_surp=0,prev_freq=seq(-9.5,7,by=0.01),prev_len=0)
+dat_for_prev_len_prediction <- expand_grid(surprisal=0,freq=0,len=0,prev_surp=0,prev_freq=0,prev_len=seq(-3.4,10.6,by=0.01))
+#predict(m,newdata=dat_for_surprisal_prediction,type="terms",terms='s(surprisal)')
+#predict(m,newdata=dat_for_freq_prediction,type="terms",terms='ti(freq)')
+term_names <- c("surprisal","freq","len","prev_surp","prev_freq","prev_len")
+dats_for_prediction <- list(
+  's(surprisal)'=dat_for_surprisal_prediction,
+  'ti(freq)'=dat_for_freq_prediction,
+  'ti(len)'=dat_for_length_prediction,
+  's(prev_surp)'=dat_for_prev_surp_prediction,
+  'ti(prev_freq)'=dat_for_prev_freq_prediction,
+  'ti(prev_len)'=dat_for_prev_len_prediction)
+N_replicates <- 101
+res_ngram <- gam_boot(ngram_mean_data,formula_no_interact,dats_for_prediction,N=N_replicates)
+bootstrapped_gam_predictions_ngram <- summarize_gam_boot_results(res_ngram,dats_for_prediction,term_names,"5-gram")
+res_grnn <- gam_boot(grnn_mean_data,formula_no_interact,dats_for_prediction,N=N_replicates)
+bootstrapped_gam_predictions_grnn <- summarize_gam_boot_results(res_grnn,dats_for_prediction,term_names,"GRNN")
+res_txl <- gam_boot(txl_mean_data,formula_no_interact,dats_for_prediction,N=N_replicates)
+bootstrapped_gam_predictions_txl <- summarize_gam_boot_results(res_txl,dats_for_prediction,term_names,"Transformer-XL")
+res_gpt <- gam_boot(gpt_mean_data,formula_no_interact,dats_for_prediction,N=N_replicates)
+bootstrapped_gam_predictions_gpt <- summarize_gam_boot_results(res_gpt,dats_for_prediction,term_names,"GPT-2")
+all_bootstrapped_predictions <- list(ngram=bootstrapped_gam_predictions_ngram,
+                                     grnn=bootstrapped_gam_predictions_grnn,
+                                     txl=bootstrapped_gam_predictions_txl,
+                                     gpt=bootstrapped_gam_predictions_gpt)
+all_surprisal_predictions <- rbind(bootstrapped_gam_predictions_ngram[[1]],
+                                   rename(bootstrapped_gam_predictions_ngram[[4]],surprisal=prev_surp),
+                                   bootstrapped_gam_predictions_grnn[[1]],
+                                   rename(bootstrapped_gam_predictions_grnn[[4]],surprisal=prev_surp),
+                                   bootstrapped_gam_predictions_txl[[1]],
+                                   rename(bootstrapped_gam_predictions_txl[[4]],surprisal=prev_surp),
+                                   bootstrapped_gam_predictions_gpt[[1]],
+                                   rename(bootstrapped_gam_predictions_gpt[[4]],surprisal=prev_surp)
+                                   ) %>%
+  mutate(s=ifelse(term=="surprisal","Current","Previous")) %>%
+  select(-term)
+
+write_rds(all_surprisal_predictions,"bootstrapped_GAM_surprisal_predictions.rds")
+```
+
+```{r example bootstrap results}
+# as an example, plot effect for surprisal
+summaries <- bootstrapped_gam_predictions_grnn
+ggplot(summaries[[1]],aes(x=surprisal,y=rt)) + geom_line() +
+  #geom_rug(aes(x=grnn_mean_data$surprisal),sides='b') +
+  geom_line(aes(x=summaries[[1]]$surprisal,y=summaries[[1]]$CI_lower),linetype="dashed") +
+  geom_line(aes(x=summaries[[1]]$surprisal,y=summaries[[1]]$CI_upper),linetype="dashed") +
+  ylim(c(-200,600)) +
+  xlab("surprisal") +
+  ylab("RT contribution") +
+  theme(aspect.ratio=1)
+ggsave("/tmp/bootstrap.pdf")
+```
+
+```{r example non-bootstrap results}
+m_no_bootstrap <- gam(formula_no_interact,data=grnn_mean_data)
+plot(m_no_bootstrap,select=1,ylim=c(-200,600))
+pdf("/tmp/no_bootstrap.pdf")
+plot(m_no_bootstrap,select=1,ylim=c(-200,600))
+dev.off()
+```
+
 ```{r plots, eval=T}
@@
 
-gam1 <- read_rds(here("Analysis/models/gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+#gam1 <- read_rds(here("Analysis/models/gam_predictions.rds")) %>%  
+gam1 <- read_rds(here("Analysis/bootstrapped_GAM_surprisal_predictions.rds")) %>%  
+  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+

### 2021-08-20 Papers/Paper/manuscript.Rmd
paper edits

@@
 We compare our A-maze results with SPR data collected on this corpus by @futrellNaturalStoriesCorpus2020.
-We find that participants were able to read and understand these long passages using A-maze, and their RT profiles show a similar pattern to SPR but with less noise and spillover
+We find that participants were able to read and understand these long passages using A-maze, and the A-maze RT profiles show similar relationships to surprisal, word length, and word frequency as those seen with SPR, but with less noise and spillover.
 
@@
 
-Many participants completed the Maze task with a high degree of accuracy and also answered the comprehension questions correctly.
+We first looked at participants overall performance on the Maze task and comrehension questions, in order to see if participants were able to read and understand a story in the Maze task.
+Many participants were able to complete the Maze task with a high degree of accuracy and answer the comprehension questions correctly.
 
@@
 
-Participant accuracy reflects both how well participants can navigate the task and what quality the auto-generated distractors are.
+Participant accuracy, or how often participants choose the correct word over the distractor, reflects both the quality of the distractors and the focus and skill of the participant.
 We calculated the per-word error rate for each participant and graphed it against their average reaction time.
@@
 ## RT and surprisal
+
+TODO: question: Does the current order of all Maze then all SPR results make sense? Or would be it better to have Maze-GAM, SPR-GAM, Maze-LM, SPR-LM, etc?
 
@@
 
-TODO remind why we care so much about surprisal
+Given that the Maze task worked, we next consider the relationship between surprisals and Maze RTs.
+Surprisal, a measure of overall word predictability in context, is linearly related to RT in eye-tracking and SPR.
+If Maze is measuring the same language processes, we would expect to see a linear relationship in the Maze task.
+Additionally, surprisal can be used to assess spillover, but comparing the predictive value of current and past word surprisal on current word reading time.
 
-We fitted generalized additive models to test whether the RTs from the Maze experiment showed a linear relationship with surprisal and whether the effect was limited the current word or had spilled over from the prior word.
+To assess the shape of the RT-surprisal relationship, we fit generalized additive models (GAMs).
 For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.
-(See Methods for details on exclusions.)
+We use surprisals generated by 4 different language models for robustness.
+(See Methods for details on language models, exclusions, and model fit.)
 
@@
 
-The smooths for the current and previous words surprisals are shown in Figure \@ref(fig:gam).
+The GAMs predicted relationship between current and previous word surprisals are shown in Figure \@ref(fig:gam); this is the main effect of either current or previous surprisal for an average length and frequency word, if the other surprisal value is 0 (TODO Not sure this is the correct interpretation!
+-- https://rdrr.io/cran/tidymv/man/get_gam_predictions.html is the function used for prediction).
 Note that for each of the models, high-surprisal words are rare, with much of the data for words between 0 and 15 bits of surprisal.
@@
 This is a sign of localization as the previous word's surprisal is not affecting RT much.
-The linear relationship matches that found with other methodologies.
+The linear relationship matches that found in SPR and eye-tracking.
+TODO comment on freq, length effects / non-linearities!
 
-Given that the GAM models show a roughly linear relationship, we fit mixed linear models to quantify the influence of surprisal, frequency and word length. 
-We built linear models with surprisal, frequency, length and surprisal x length and frequency x length effects from the current and previous words as predictors.
-We centered, but did not rescale, the predictors.
+Given that the GAM models show a roughly linear relationship between RT and surprisal, we are justified in using surprisal as a predictor in a linear model to quantify its effect.
+In addition to surprisal, we also include the commonly studied effects of frequency and length as well as surprisal x length and frequency x length interactions.
+For all of these, we include the predictors for the current and previous word, and we centered, but did not rescale, all predictors.
+(See Methods for more details on these predictors and model fit process.)
 
@@
 
-As we can see in Table \@ref(tab:pre-error), we find large effects of surprisal and length, but minimal effects of frequency.
+As shown in Table \@ref(tab:pre-error), we found large effects of surprisal and length, but minimal effects of frequency.
 The lack of frequency effects is somewhat surprising, but consistent with @shainLargescaleStudyEffects2019.
-Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word.
+Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word, an indication that spillover is limited and effects are strongly localized.
 
@@
 
+TODO mention at sentence level to control for spillover profiles and so single words are too noisy
+
+Add interpretation
+
 # Discussion
 
-## maze works
+We introduce error-correction Maze, a tweak on the presentation of Maze materials that makes Maze feasible for multi-sentence passages.
+We used A-maze distractors and this error-correction presentation to gather data on participants reading the Natural Stories in the Maze.
+We found that participants could read and comprehend the 1000 word stories, despite the slowness and added overhead of reading in the Maze task.
+This expands the domain of materials used with Maze beyond targeted single-sentence items to longer, naturalistic texts with sentence-to-sentence coherency.
 
-## strong effects of suprisal, mostly localized
-
-## better than SPR
-
-## correlated w SPR, probably the same effects, just stronger & less noisy
-
-## Why you care - broad applications that might want this method
-
-## shill for error correction
-We introduced a tweak to the paradigm for displaying Maze tasks that makes it useable for multi-sentence materials.
-In this error-correction Maze paradigm, participants read all the words because they can correct their mistakes and move on.
-We tested this method on the Natural Stories corpus, showing that, despite the oddness of the task, participants can read and understand a 1000 word story in this method.
-We additionally showed that the RTs generated in the Maze task show a linear relationship between the RT of a word and it's surprisal, but no relationship with the surprisal of the previous word.
-This provides additional evidence for the argument that the Maze task forces very incremental processing [@forsterMazeTaskMeasuring2009].
-
-Actually the potential applicability and value is much broader, for any question where nailing down the locus on incremental processing difficulty is important, little spillover is very valuable.
-
-This extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty and thus benefits from the lack of spillover.
-It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more.
-Comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex.
+We took advantage of the pre-existing SPR corpus on Natural Stories to compare the RT profiles between Maze and SPR.
+While these are not completely comparably as the subject populations differed, we are able to get a sense of the relationship between SPR and Maze results.
+Maze and SPR pick up on similar features in words, as shown by the correlations between Maze and SPR, and the fact that surprisal and length are predictive of RTs in both.
+However, the effect sizes of RT and surprisal are much larger is Maze, and the effects are more strongly localized on a single word.
+Further comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex.
 @bartekSearchOnlineLocality2011].
 
-While the error-correction paradigm is crucial to running long materials, it also provides some benefits even on shorter materials when using A-maze with variable participant populations.
-The error-correction compensates for some of the shortcomings of A-maze identified in @boyceMazeMadeEasy2020; poor distractors might still cause participants to make unavoidable mistakes, but they will still see the rest of the sentence, which may reduce frustration.
-One questions that remains is whether this post-mistake data is useable, which depends on whether RTs from a few words after a mistake show any differences from RTs before any mistakes.
+We find good localization on Maze, but it's worth noting that it's not absolute.
+Some of the models show small but statistically reliable relationships with surprisal from a past word.
+On the whole, however, these results support the idea that Maze forces language processing to be close to word-by-word, and that it's a relatively safe assumption that the RT of a word primarily reflects its own properties and not those of earlier words.
+
+This extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty.
+
+## Remarks about error-correction maze
+
+With that in mind, we have a few remarks on the benefits of error-correction Maze and some avenues for further work on this method.
+Error-correction Maze makes for a smoother process.
+It dampens the effect of rare poor A-maze distractors; a distractor that fits the context might still cause participants to make an unavoidable mistake, but the mistake is less disruptive, as they still see the rest of the sentence.
+An open question for researchers is whether data from after a participant makes a mistake is useable, that is, does it show the same profile as pre-errors words, or are there traces from recovering from the mistake?
 Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is a hard-to-assess question of potential interest.
 
-Even if post-mistake data is not analysed, it can be used to distinguish between errors due to inattentive participants or merely specific bad distractors early in the sentence.
+Even if post-mistake data is not included in analyses, having this complete information can distinguish between errors due to inattentive participants and errors from specific bad distractors.
 Researchers can calculate a per-word error rate for each participant; high per-word error rates are consistent with guessing, low error rates with errors clustered early the sentence are consistent with poor distractors.
@@
 
-Between the distractor auto-generation process introduced in @boyceMazeMadeEasy2020 and the error-correction paradigm introduced here, the Maze paradigm is now easy to use on a wide range of materials.
-The ease of use, large effects, and forced incrementality make Maze a good complement to existing incremental processing methods.
-We encourage researchers to consider Maze as an option for doing incremental processing work.
+Between error-correcting Maze and other innovations on the Maze paradigm such as A-maze and I-maze, the Maze task is versatile and can be used or adapted for a wide range of materials and questions of interest.
+We encourage researchers to use Maze as an incremental processing method, alone or in comparison with other methods.
+
 \newpage

### 2021-08-23 Analysis/models.Rmd
Roger completed bootstrapping work for models, both Maze and SPR

@@
 
-```{r bootstrap code}
+```{r maze bootstrap code}
 require(OpenImageR)
@@
   result <- lapply(1:length(data_frames_for_prediction),function(i) {
-    #print(term_names[[i]])
+    #print(cat(i,"\t",term_names[[i]]))
     res <- tibble(x=data_frames_for_prediction[[i]][[term_names[[i]]]],
                   rt=apply(gam_boot_results[[i]],1,point_estimation_function),
-                  CI_lower=apply(gam_boot_results[[i]],1,function(x) quantile(x,0.1)),
-                  CI_upper=apply(gam_boot_results[[i]],1,function(x) quantile(x,0.9)),
+                  CI_lower=apply(gam_boot_results[[i]],1,function(x) quantile(x,quantiles[1])),
+                  CI_upper=apply(gam_boot_results[[i]],1,function(x) quantile(x,quantiles[2])),
                   model=model_name,
-                  term=term_names[[i]])
+                  term=term_names[[i]]
+                  )
     names(res)[names(res)=="x"] <- term_names[[i]]
@@
 
+write_rds(all_bootstrapped_predictions,"bootstrapped_GAM_predictions.rds")
 write_rds(all_surprisal_predictions,"bootstrapped_GAM_surprisal_predictions.rds")
@@
 
+```{r spr bootstrapping}
+surprisal_range <- seq(0,25,by=0.01)
+freq_range <- seq(-9.5,7,by=0.01)
+len_range <- seq(-3.4,10.6,by=0.01)
+surprisal_terms <- c("surp","past_surp","past2_surp","past3_surp")
+freq_terms <- c("freq_center","past_freq_center","past2_freq_center","past3_freq_center")
+len_terms <- c("length_center","past_length_center","past2_length_center","past3_length_center")
+
+dat_for_surp_prediction <- expand_grid(surp=seq(0,25,by=0.01),freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
+dat_for_freq_prediction <- expand_grid(surp=0,freq_center=seq(-9.5,7,by=0.01),length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
+dat_for_length_prediction <- expand_grid(surp=0,freq_center=0,length_center=seq(-3.4,10.6,by=0.01),past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
+dat_for_past_surp_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=seq(0,25,by=0.01),past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
+dat_for_past_freq_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=seq(-9.5,7,by=0.01),past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
+dat_for_past_len_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=seq(-3.4,10.6,by=0.01),past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
+dat_for_past2_surp_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=seq(0,25,by=0.01),past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
+dat_for_past2_freq_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=seq(-9.5,7,by=0.01),past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=0)
+dat_for_past2_len_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=seq(-3.4,10.6,by=0.01),past3_surp=0,past3_freq_center=0,past3_length_center=0)
+dat_for_past3_surp_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=seq(0,25,by=0.01),past3_freq_center=0,past3_length_center=0)
+dat_for_past3_freq_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=seq(-9.5,7,by=0.01),past3_length_center=0)
+dat_for_past3_len_prediction <- expand_grid(surp=0,freq_center=0,length_center=0,past_surp=0,past_freq_center=0,past_length_center=0,past2_surp=0,past2_freq_center=0,past2_length_center=0,past3_surp=0,past3_freq_center=0,past3_length_center=seq(-3.4,10.6,by=0.01))
+
+term_names <- c("surp","freq_center","length_center","past_surp","past_freq_center","past_length_center","past2_surp","past2_freq_center","past2_length_center","past3_surp","past3_freq_center","past3_length_center")
+dats_for_prediction <- list(
+  's(surp)'=dat_for_surp_prediction,
+  'ti(freq_center)'=dat_for_freq_prediction,
+  'ti(length_center)'=dat_for_length_prediction,
+  's(past_surp)'=dat_for_past_surp_prediction,
+  'ti(past_freq_center)'=dat_for_past_freq_prediction,
+  'ti(past_length_center)'=dat_for_past_len_prediction,
+  's(past2_surp)'=dat_for_past2_surp_prediction,
+  'ti(past2_freq_center)'=dat_for_past2_freq_prediction,
+  'ti(past2_length_center)'=dat_for_past2_len_prediction,
+  's(past3_surp)'=dat_for_past3_surp_prediction,
+  'ti(past3_freq_center)'=dat_for_past3_freq_prediction,
+  'ti(past3_length_center)'=dat_for_past3_len_prediction)
+
+N_replicates <- 101
+res_ngram_spr <- gam_boot(ngram_mean_spr,spr_formula_interact,dats_for_prediction,N=N_replicates)
+bootstrapped_gam_predictions_ngram_spr <- summarize_gam_boot_results(res_ngram_spr,dats_for_prediction,term_names,"5-gram")
+res_grnn_spr <- gam_boot(grnn_mean_spr,spr_formula_interact,dats_for_prediction,N=N_replicates)
+bootstrapped_gam_predictions_grnn_spr <- summarize_gam_boot_results(res_grnn_spr,dats_for_prediction,term_names,"GRNN")
+res_txl_spr <- gam_boot(txl_mean_spr,spr_formula_interact,dats_for_prediction,N=N_replicates)
+bootstrapped_gam_predictions_txl_spr <- summarize_gam_boot_results(res_txl_spr,dats_for_prediction,term_names,"Transformer-XL")
+res_gpt_spr <- gam_boot(gpt_mean_spr,spr_formula_interact,dats_for_prediction,N=N_replicates)
+bootstrapped_gam_predictions_gpt_spr <- summarize_gam_boot_results(res_gpt_spr,dats_for_prediction,term_names,"GPT-2")
+all_bootstrapped_predictions_spr <- list(ngram=bootstrapped_gam_predictions_ngram_spr,
+                                     grnn=bootstrapped_gam_predictions_grnn_spr,
+                                     txl=bootstrapped_gam_predictions_txl_spr,
+                                     gpt=bootstrapped_gam_predictions_gpt_spr)
+write_rds(all_bootstrapped_predictions_spr,"bootstrapped_spr_GAM_predictions.rds")
+
+term_conversions <- c("surp"="Current","past_surp"="Previous","past2_surp"="2Previous","past3_surp"="3Previous")
+all_surprisal_predictions_spr <- rbind(rename(bootstrapped_gam_predictions_ngram_spr[[1]],surprisal=surp),
+                                   rename(bootstrapped_gam_predictions_ngram_spr[[4]],surprisal=past_surp),
+                                   rename(bootstrapped_gam_predictions_ngram_spr[[7]],surprisal=past2_surp),
+                                   rename(bootstrapped_gam_predictions_ngram_spr[[10]],surprisal=past3_surp),
+                                   rename(bootstrapped_gam_predictions_grnn_spr[[1]],surprisal=surp),
+                                   rename(bootstrapped_gam_predictions_grnn_spr[[4]],surprisal=past_surp),
+                                   rename(bootstrapped_gam_predictions_grnn_spr[[7]],surprisal=past2_surp),
+                                   rename(bootstrapped_gam_predictions_grnn_spr[[10]],surprisal=past3_surp),
+                                   rename(bootstrapped_gam_predictions_txl_spr[[1]],surprisal=surp),
+                                   rename(bootstrapped_gam_predictions_txl_spr[[4]],surprisal=past_surp),
+                                   rename(bootstrapped_gam_predictions_txl_spr[[7]],surprisal=past2_surp),
+                                   rename(bootstrapped_gam_predictions_txl_spr[[10]],surprisal=past3_surp),
+                                   rename(bootstrapped_gam_predictions_gpt_spr[[1]],surprisal=surp),
+                                   rename(bootstrapped_gam_predictions_gpt_spr[[4]],surprisal=past_surp),
+                                   rename(bootstrapped_gam_predictions_gpt_spr[[7]],surprisal=past2_surp),
+                                   rename(bootstrapped_gam_predictions_gpt_spr[[10]],surprisal=past3_surp)
+                                   ) %>%
+  mutate(s=factor(term_conversions[term],levels=term_conversions)) %>%
+  select(-term)
+write_rds(all_surprisal_predictions_spr,"bootstrapped_spr_GAM_surprisal_predictions.rds")
+
+```
+
 ```{r spr-gam-plot}
@@
 
-gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+#gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  
+gam1 <- read_rds(here("Analysis/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  
+  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+

### 2021-08-23 Papers/Paper/manuscript.Rmd

@@
 ---
-title             : "A-maze of Natural Stories: Comprehension and surprisal in the Maze task TODO title"
+title             : "A-maze of Natural Stories: Comprehension and surprisal in the Maze task"
 shorttitle        : "A-maze of Natural Stories"
@@
 # Introduction
-
-TODO overall: go through Roger's comments!
-TODO fix tense!
 
@@
 
-TODO top level description
-
 In order to measure the relationship between a word's properties and it's RT, we fit models using surprisal, frequency, and length as predictors of RT.
@@
 
-## Reading stories in the Maze task
+## Reading in the Maze
 
@@
 
-TODO: question: Does the current order of all Maze then all SPR results make sense? Or would be it better to have Maze-GAM, SPR-GAM, Maze-LM, SPR-LM, etc?
+Given that the Maze task worked, we next consider the relationship between surprisals and Maze RTs.
+Surprisal, a measure of overall word predictability in context, is linearly related to RT in eye-tracking and SPR.
+If Maze is measuring the same language processes, we would expect to see a linear relationship in the Maze task.
+Additionally, surprisal can be used to assess spillover, but comparing the predictive value of current and past word surprisal on current word reading time.
+
+### GAMs
 
@@
 
-Given that the Maze task worked, we next consider the relationship between surprisals and Maze RTs.
-Surprisal, a measure of overall word predictability in context, is linearly related to RT in eye-tracking and SPR.
-If Maze is measuring the same language processes, we would expect to see a linear relationship in the Maze task.
-Additionally, surprisal can be used to assess spillover, but comparing the predictive value of current and past word surprisal on current word reading time.
-
 To assess the shape of the RT-surprisal relationship, we fit generalized additive models (GAMs).
@@
 
-(ref:gam-cap) GAM predictions of reaction time (RT) as a function of either current word surprisal (top) or previous word surprisal (bottom).
+(ref:gam-cap) GAM predictions of Maze reaction time (RT) as a function of either current word surprisal (top) or previous word surprisal (bottom).
 Density of data is shown along the x-axis.
@@
 This is a sign of localization as the previous word's surprisal is not affecting RT much.
-The linear relationship matches that found in SPR and eye-tracking.
+
 TODO comment on freq, length effects / non-linearities!
+
+(ref:spr-gam-cap) GAM predictions of reaction time (RT) for SPR data from @futrellNaturalStoriesCorpus2020 as a function of current word surprisal (top) or the surprisal of an earlier word, up to 3 words back.
+Density of data is shown along the x-axis.
+
+```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
+gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+  geom_line()+
+  geom_ribbon(alpha=.3)+
+  facet_grid(s~model)+
+  coord_cartesian(xlim=c(0,28))+
+  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
+
+ all_mean_spr <-  read_rds(here("Analysis/models/meaned_spr.rds"))
+
+dens2 <-   ggplot(all_mean_spr, aes(x=value))+
+  geom_density(fill="gray",)+
+  facet_grid(.~model)+
+  labs(x="Surprisal (bits)", y="")+
+    coord_cartesian(xlim=c(0,28))+
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
+plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .2))
+
+```
+
+As a comparison, we also ran GAMs on the Self-Paced Reading data collected by @futrellNaturalStoriesCorpus2020.
+Previous work such as @smithEffectWordPredictability2013 has found positive relationships between RT and the surprisal of earlier words for SPR, so we include predictors from the 3 prior words.
+The predicted relationship between surprisals and RT is shown in Figure \@ref(fig:spr-gam); note that the y-axis range is muuch narrower than for Maze.
+Both current and previous word surprisals have a roughly linear positive relationship to RT.
+The surprisal of the word two back also has an influence in some models.
+
+Comparing the Maze and SPR smooths, we see that both show a linear relationship, but Maze has much larger and more localized effects of surprisal.
+
+### Effect sizes
 
@@
 
-As shown in Table \@ref(tab:pre-error), we found large effects of surprisal and length, but minimal effects of frequency.
+The Maze linear model effects are shown in Table \@ref(tab:pre-error).
+we found large effects of surprisal and length, but minimal effects of frequency.
 The lack of frequency effects is somewhat surprising, but consistent with @shainLargescaleStudyEffects2019.
 Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word, an indication that spillover is limited and effects are strongly localized.
-
-As a last analysis, we checked which of our surprisal models had the best fit using a nested model comparison shown in Table \@ref(tab:maze-compare).
-We assess the benefits of adding each model's predictions as a second surprisal source to see which models pick up on information not contained in another model.
-GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complemetary information over GPT-2.
-TXL provides a lot over 5-gram, and 5-gram provides little over any model.
-Log likelihood, which is a measure of fit to data, shows  a similar hierarchy, as GPT-2 is better than GRNN is better than TXL is better than 5-gram.
-```{r maze-compare}
-
-read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data.
-Each row shows the additional predictive value gained from adding that model to another model.
-F values and p values from anova tests between 1-surprisal-source and 2-source models are reported.
-We also report log likelihoods of models with only one surprisal source.")
-
-```
-
-## Comparison with SPR
-
-(ref:spr-gam-cap) GAM predictions of reaction time (RT) for SPR data from @futrellNaturalStoriesCorpus2020 as a function of current word surprisal (top) or the surprisal of an earlier word, up to 3 words back.
-Density of data is shown along the x-axis.
-
-```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
-gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
-  geom_line()+
-  geom_ribbon(alpha=.3)+
-  facet_grid(s~model)+
-  coord_cartesian(xlim=c(0,28))+
-  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
-
- all_mean_spr <-  read_rds(here("Analysis/models/meaned_spr.rds"))
-
-dens2 <-   ggplot(all_mean_spr, aes(x=value))+
-  geom_density(fill="gray",)+
-  facet_grid(.~model)+
-  labs(x="Surprisal (bits)", y="")+
-    coord_cartesian(xlim=c(0,28))+
-  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
-
-p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
-plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .2))
-
-```
 
@@
 
-As a comparison, we ran the same types of models on the Self-Paced Reading data collected by @futrellNaturalStoriesCorpus2020.
-As shown in Figure \@ref(fig:spr-gam), there is a roughly linear but fairly flat relationship between RT and surprisal.
-Note that the y-axis is fairly narrow, and so the predicted effect of changes in surprisal is fairly small.
-This is confirmed by linear models (see Table \@ref(tab:spr-table)).
-Surprisal and length effects are evident for the current word, and most models shows surprisal effects from the past word, but we do not see frequency effects.
-The effects are much smaller than for A-maze, even though this model accounts for spillover effects from more previous words.
+We ran similar models for SPR, although to account for known spillover effects, we consider predictors from the current and 3 previous words.
+Due to issues fitting models, the details of the models differed (see Methods).
+The SPR coefficients are shown in Table \@ref(tab:spr-table).
+Surprisal, length, and frequency effects are all evident for the current word and surprisal and frequency show effects from the previous word as well.
 
-Similarly to the model comparison for Maze, we also conducted a model comparison on the SPR data shown in Table \@ref(tab:spr-compare).
-Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN.
+Whereas Maze showed surprisal effects in the 10 to 25 ms/bit range and length effects 15-20 ms/character range, SPR effects are about 1/10 this size.
+This difference in effect is disproportionate to the overall speed of the methods; Maze is 2-3 times as slow as SPR.
+
+In addition, the relative effects of these different predictors shows somewhat different patterns; in SPR, frequency effects are roughly as big as surprisal effects, but this is not true in Maze.
+In SPR, length effects exceed surprisal effects, in Maze, this varies depending on the model.
+
+TODO: is there a way to make the tables more readable / highlight the key rows?
+Maybe stick the full tables in the appendix and show select versions omitting interactions and 2- 3- past words in the main text?
+Or maybe it's fine and people can read?
+
+### Model comparison
+
+We use 4 different models for surprisal.
+This provided robustness to the above analyses, but it also allows us to look at which models have the best fit to human Maze and SPR data.
+For each data type, we assessed each models fit using log likelihood and R-squared.
+Then we did a nested model comparison, looking at whether a model with multiple surprisal predictors (ex, GRNN and GPT) had a better fit than a model with only one (ex GRNN alone).
+This tells us whether a model contains predictive value beyond that contained in another model.
+As shown in Table \@ref(tab:maze-compare), GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complementary information over GPT-2.
+TXL provides a lot over 5-gram, and 5-gram provides little over any model.
+The single-model measures of log likelihood confirm this hierarchy, as GPT-2 is better than GRNN is better than TXL is better than 5-gram.
+```{r maze-compare}
+
+read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data.
+Each row shows the additional predictive value gained from adding that model to another model.
+F values and p values from anova tests between 1-surprisal-source and 2-source models are reported.
+We also report log likelihoods of models with only one surprisal source.")
+
+```
+
+We followed the same process for the SPR data with results shown in Table \@ref(tab:spr-compare).
+For SPR, GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN.
 In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small.
 The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020.
-This contrasts with the Maze results, where the 5-gram model has the worst fit and does not provide additional predictive value to the other models.
+This contrasts with the Maze results, where the 5-gram model had the worst fit and did not provide additional predictive value to the other models.
 
@@
 ```
+
+### Maze - SPR correlation
 
@@
 
-An additional comparison between SPR and Maze is how correlated their reading times are; that is, do the methods pick up on the same effects.
-If so, we would expect that the sentences that take longer to read in one method also take longer in the other.
-We calculate the average RT at the sentence level (see Methods for details).
+As a last analysis, we measured how correlated SPR and Maze are in their reading times; that is, did the methods pick up on the same effects.
+If so, sentences that take longer to read in one method would also take longer in the other.
+We calculate the average RT at the sentence level because differences in spillover will effect sentence level RT less than RT for individual words (see Methods for details).
 The correlation between Maze and SPR is `r round(spr_maze,2)`, compared to `r round(spr_spr,2)` within SPR and `r round(maze_maze,2)` within Maze.
@@
 
-TODO mention at sentence level to control for spillover profiles and so single words are too noisy
-
-Add interpretation
+The fact that SPR data is about as correlated with Maze as with another sample of SPR data provides some confirmation that Maze and SPR are measuring the same thing.
 
@@
 However, the effect sizes of RT and surprisal are much larger is Maze, and the effects are more strongly localized on a single word.
+
+Some language models are better or worse on predicting human data; however, we find preliminary evidence that how predictive a language model is of human data may depend on what task was used to collect the RTs, with the 5-gram model relatively better at predicting SPR data.
 Further comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex.

### 2021-09-16 Analysis/models.Rmd
add bootstraps to ms

@@
 
-## Plots
-
-```{r maze bootstrap code}
+```{r maze bootstrap code, eval=F}
 require(OpenImageR)
@@
 
-```{r example bootstrap results}
+```{r example bootstrap results,eval=F}
 # as an example, plot effect for surprisal
@@
 
-```{r example non-bootstrap results}
+```{r example non-bootstrap results, eval=F}
 m_no_bootstrap <- gam(formula_no_interact,data=grnn_mean_data)
@@
 
-```{r plots, eval=T}
+```{r plots, eval=F}
 message("GPT")
@@
 
-```{r spr bootstrapping}
+```{r spr bootstrapping, eval=F}
 surprisal_range <- seq(0,25,by=0.01)

### 2021-09-16 Papers/Paper/manuscript.Rmd
add bootstraps to ms

@@
 abstract: |
-   TODO needs work
-   A-maze is a new method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities.
-We adapt A-maze for use on longer passages and test it on the Natural Stories corpus.
-We find that people can comprehend these longer text passages during the Maze task.
-Moreover, the  task yields useable reaction time data with word predictability effects that are linear in the surprisal of the current word, with little spillover effect from the surprisal of the previous word.
-This expands the types of effects that can be studied with A-maze, showing it to be a a versatile alternative to eye-tracking and self-paced reading.We find support for the localization of reading time effects during the Maze task, as well as extending the range of materials Maze is suitable for.
-How long it takes to read a word in a sentence is reflective of how hard it is to identify and integrate the word in the surrounding context.
-Techniques that slow down the reading process and localize the processing time for each word are useful to understanding the time course of language processing.
+   Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing.
+A-maze is a recent method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities in individual sentences.
+We adapted A-maze for use on longer passages and tested it on the Natural Stories corpus.
+Participants were able to comprehend these longer text passages that they read via the Maze task.
+Moreover, the Maze task yielded useable reaction time data with word predictability effects that are linear in the surprisal of the current word, the same pattern found with other tasks.
+Crucially, Maze reaction times show a tight relationship with properties of the current word, but little spillover of effects from previous words.
+This superior localization is an advantage of Maze compared with other methods.
+Overall, we expand the scope of experimental materials, and thus theoretical questions, that can be coupled with the Maze task.
 
@@
 We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions.
-To minimize the effect of repeated measures on confidence about inferred curve shapes, we collapse the reading times across subjects by modelling the mean RT for each word.
-TODO edit for bootstrapping
+To account for by-subject variation, we do bootstrapping ...
+TODO specs for bootstrapping
 
@@
 We included frequency, length, and their interaction to all models.
-Then we fit models with either 1 or 2 sources of surprisal using lm and assessed the effect of adding the second surprisal source with an anova.
+Then we fit models with either 1 or 2 sources of surprisal using lm [@batesFittingLinearMixedeffects2015] and assessed the effect of adding the second surprisal source with an anova.
 We used predictors for the current and past word and centered all effects.
-TODO cite LM
 
@@
 
-(ref:gam-cap) GAM predictions of Maze reaction time (RT) as a function of either current word surprisal (top) or previous word surprisal (bottom).
+(ref:gam-cap) GAM results for the effect on Maze reaction time (RT) from either current word surprisal (top) or previous word surprisal (bottom).
 Density of data is shown along the x-axis.
@@
 The relationship between previous word surprisal and RT is much flatter.
+TODO these are bootstrapped 95% CI?
 
@@
 
-gam1 <- read_rds(here("Analysis/models/gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+gam1 <- read_rds(here("Analysis/bootstrapped_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+
@@
   coord_cartesian(xlim=c(0,28))+
-  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
+  labs(x="Surprisal (bits)", y="Effect on Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
 
@@
 
-The GAMs predicted relationship between current and previous word surprisals are shown in Figure \@ref(fig:gam); this is the main effect of either current or previous surprisal for an average length and frequency word, if the other surprisal value is 0 (TODO Not sure this is the correct interpretation!
--- https://rdrr.io/cran/tidymv/man/get_gam_predictions.html is the function used for prediction).
+The relationships between current and previous word surprisals and RT are shown in Figure \@ref(fig:gam); this is the main effect of either current or previous surprisal.
 Note that for each of the models, high-surprisal words are rare, with much of the data for words between 0 and 15 bits of surprisal.
@@
 
-(ref:spr-gam-cap) GAM predictions of reaction time (RT) for SPR data from @futrellNaturalStoriesCorpus2020 as a function of current word surprisal (top) or the surprisal of an earlier word, up to 3 words back.
+(ref:spr-gam-cap) GAM results for the effect on SPR RT data [@futrellNaturalStoriesCorpus2020] as a function of current word surprisal (top) or the surprisal of an earlier word, up to 3 words back.
 Density of data is shown along the x-axis.
+TODO these are bootstrapped 95% CI?
 
 ```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
-gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+gam1 <- read_rds(here("Analysis/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+
@@
 Previous work such as @smithEffectWordPredictability2013 has found positive relationships between RT and the surprisal of earlier words for SPR, so we include predictors from the 3 prior words.
-The predicted relationship between surprisals and RT is shown in Figure \@ref(fig:spr-gam); note that the y-axis range is muuch narrower than for Maze.
+The  relationship between surprisals and RT is shown in Figure \@ref(fig:spr-gam); note that the y-axis range is much narrower than for Maze.
 Both current and previous word surprisals have a roughly linear positive relationship to RT.

### 2021-09-21 Papers/Paper/manuscript.Rmd
Roger added bootstrap details

@@
 ### Model specification
-For generalized additive models, we centered but did not rescale the length and frequency predictors, but left surprisal uncentered for interpretability.
-We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions.
-To account for by-subject variation, we do bootstrapping ...
-TODO specs for bootstrapping
+To infer the shape of the relationship between our predictor variables and RTs, we fitted generalized additive models (GAMs) using `R`'s `mgcv` package to predict the mean RT (after exclusions) for each word, averaging across participants from whom we we obtained an unexcluded RT for that word.
+We centered but did not rescale the length and frequency predictors, and left surprisal uncentered for interpretability.
+We used smooth terms (`mgcv`'s `s()`) for surprisal and tensor product terms (`mgcv`'s `ti()`) for frequency-by-length effects and interactions.
+We use restricted maximum likelihood (REML) smoothing for parameter estimation.
+To more fully account for the uncertainty in the smoothing parameter estimates, we fitted 101 bootstrap replicates of each GAM model; in Figures \@ref(fig:gam) and \@ref(fig:spr-gam), the best-fit lines derive from the mean estimated effect size across the bootstrap replicates, and the shaded areas indicate a 95\% bootstrap confidence interval on this effect size (the boundaries are the 2.5\% and 97.5\% quantiles of the bootstrapped replicates).
+
+To account for by-subject variation, we do bootstrapping ... TODO specs for bootstrapping
 

### 2022-03-01 Analysis/models.Rmd
lots of paper editing

@@
 knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
+library(plyr)
 library(tidyverse)
@@
 library(scales)
+library(ggstance)
 library(here)
@@
 
+# Messing around with graphs
+```{r play}
+
+show_range<- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+    left_join(intervals, by=".variable")
+
+  stats
+}
+ngram <- read_rds(here("Analysis/models/5-gram_maze.rds")) %>% show_range() %>% mutate(model="5-gram")
+txl <- read_rds(here("Analysis/models/txl_maze.rds")) %>% show_range() %>% mutate(model="TXL")
+grnn <- read_rds(here("Analysis/models/grnn_maze.rds")) %>% show_range() %>% mutate(model="GRNN")
+gpt <- read_rds(here("Analysis/models/gpt_maze.rds")) %>% show_range() %>% mutate(model="GPT-2")
+
+test <- ngram %>% union(txl) %>% union(grnn) %>% union(gpt) %>% write_rds(here("Analysis/models/brms_maze_summ_plot.rds"))
+
+show_range_spr <- function(model){
+
+  stats <- summary(model)[["coefficients"]] %>% 
+    as_tibble(rownames="Term") %>% 
+    mutate(width=1.97*`Std. Error`,
+           lower=Estimate-width,1,
+           upper=Estimate+width,1,
+           E=Estimate)
+
+  stats
+}
+ngram_spr <- read_rds(here("Analysis/models/5-gram_spr_lmer.rds")) %>% show_range_spr() %>% mutate(model="5-gram")
+txl_spr <- read_rds(here("Analysis/models/txl_spr_lmer.rds")) %>% show_range_spr() %>% mutate(model="TXL")
+grnn_spr <- read_rds(here("Analysis/models/grnn_spr_lmer.rds")) %>% show_range_spr() %>% mutate(model="GRNN")
+gpt_spr <- read_rds(here("Analysis/models/gpt_spr_lmer.rds")) %>% show_range_spr() %>% mutate(model="GPT-2")
+
+range_spr <- ngram_spr %>% union(txl_spr) %>% union(grnn_spr) %>% union(gpt_spr) %>% write_rds(here("Analysis/models/lmer_spr_summ_plot.rds"))
+```
+
+```{r}
+
+test_plot_maze <- read_rds(here("Analysis/models/brms_maze_summ_plot.rds")) %>%  mutate(term=str_sub(.variable, 3, -1)) %>% 
+   mutate(term=as.factor(term),
+          type="Maze",
+          Upper=.upper,
+          Lower=.lower,
+          Estimate=.value,
+          time=ifelse(str_detect(term, "prev"),"Previous","Current"),
+          term2=mapvalues(term, from=c("Intercept", 
+                                 "surprisal",
+                                 "len",
+                                 "freq",
+                                 "surprisal:len",
+                                 "len:freq",
+                                 "prev_surp",
+                                 "prev_len",
+                                 "prev_freq",
+                                 "prev_surp:prev_len",
+                                 "prev_len:prev_freq"), to=c("Intercept", "Surprisal", "Length", "Frequency",
+                  "Surp x Length", "Freq x Length", "Surprisal",
+                  "Length", "Frequency", "Surp x Length", "Freq x Length")),
+          Term=fct_relevel(term2,"Intercept", "Surprisal", "Length", "Frequency", "Surp x Length", "Freq x Length" )) %>% 
+  filter(Term!="Intercept") %>% select(model, Term, type, Upper, Lower, Estimate,time)
+
+test_plot_spr <- read_rds(here("Analysis/models/lmer_spr_summ_plot.rds")) %>% mutate(term=as.factor(Term),
+          type="SPR",
+          Upper=upper,
+          Lower=lower,
+          time=case_when(
+            str_detect(term, "past2")~"2Previous",
+            str_detect(term, "past3")~"3Previous",
+            str_detect(term, "past")~"Previous",
+            T ~ "Current"),
+          term2=mapvalues(term, from=c("(Intercept)", 
+                                 "surp_center",
+                                 "length_center",
+                                 "freq_center",
+                                 "surp_center:length_center",
+                                 "length_center:freq_center",
+                                 "past_surp_center",
+                                 "past_length_center",
+                                 "past_freq_center",
+                                 "past_surp_center:past_length_center",
+                                 "past_length_center:past_freq_center",
+                                 "past2_surp_center",
+                                 "past2_length_center",
+                                 "past2_freq_center",
+                                 "past2_surp_center:past2_length_center",
+                                 "past2_length_center:past2_freq_center",
+                                 "past3_surp_center",
+                                 "past3_length_center",
+                                 "past3_freq_center",
+                                 "past3_surp_center:past3_length_center",
+                                 "past3_length_center:past3_freq_center"
+                                 ), to=c("Intercept", "Surprisal", "Length", "Frequency","Surp x Length", "Freq x Length",
+                "Surprisal", "Length", "Frequency", "Surp x Length", "Freq x Length",
+                "Surprisal", "Length", "Frequency", "Surp x Length", "Freq x Length",
+                "Surprisal", "Length", "Frequency", "Surp x Length", "Freq x Length")),
+          Term=fct_relevel(term2,"Intercept", "Surprisal", "Length", "Frequency", "Surp x Length", "Freq x Length" )) %>% 
+  filter(Term!="Intercept") %>% select(model, Term, type, Upper, Lower, Estimate, time)
+
+test_plot <- test_plot_maze %>% union(test_plot_spr) %>% write_rds(here("Analysis/models/maze_spr_for_graph.rds"))
+
+test_plot %>% filter(time %in% c("Current", "Previous")) %>% ggplot(aes(y=Term, x=Estimate, group=model, color=model, shape=model))+
+  ggstance::geom_pointrangeh(aes(xmin = Lower, xmax = Upper), position=position_dodge(width=.6))+facet_grid(type~time)+geom_vline(xintercept=0)+scale_color_brewer(palette="Dark2")+theme(legend.position = "bottom")+labs(x="Coefficient estimate", y="Term", color="Model")
+```
+
 # Maze Model comparison

### 2022-03-01 Analysis/models.Rmd
Roger's in-progress version of models.Rmd

--- 2021-09-16 Analysis/models.Rmd
@@
 
+formula3_interact <- rt ~ ti(surprisal, bs="cr", k=20)+
+                     ti(freq, bs="cr") + 
+                     ti(len, bs="cr")+
+                     ti(freq,len, bs="cr")+
+                     ti(surprisal,len, bs="cr")+
+                     ti(surprisal,freq, bs="cr")+
+                     ti(prev_surp, bs="cr", k=20)+
+                     ti(prev_freq, bs="cr")+
+                     ti(prev_len, bs="cr")+
+                     ti(prev_freq, prev_len, bs="cr") +
+                     ti(prev_surp, prev_len, bs="cr") +
+                     ti(prev_surp, prev_freq, bs="cr")
+
+formula2_interact <- rt ~ 
+                     ti(freq,len, bs="cr")+
+                     ti(surprisal,len, bs="cr")+
+                     ti(surprisal,freq, bs="cr")+
+                     ti(prev_freq, prev_len, bs="cr") +
+                     ti(prev_surp, prev_len, bs="cr") +
+                     ti(prev_surp, prev_freq, bs="cr")
 ```
@@
 gam_gpt_1 <- gam(formula_interact, data=gpt_mean_data, method="REML")
+gam_gpt_1a <- gam(formula_interact, data=filter(gpt_mean_data,len<9), method="REML")
+gam_gpt_2 <- gam(formula2_interact, data=gpt_mean_data, method="REML")
+gam_gpt_3 <- gam(formula3_interact, data=gpt_mean_data, method="REML")
+print(plot(getViz(gam_gpt_1x)))
+print(plot(getViz(gam_gpt_3)))
 
@@
 gam_txl_2 <- gam(formula_no_interact, data=txl_mean_data, method="REML")
-gam_gpt_2 <- gam(formula_no_interact, data=gpt_mean_data, method="REML")
+gam_gpt_no_interact <- gam(formula_no_interact, data=gpt_mean_data, method="REML")
 
@@
 write_rds(all_surprisal_predictions,"bootstrapped_GAM_surprisal_predictions.rds")
+
+dats_for_predictionX <- list(
+  'ti(surprisal)'=dat_for_surprisal_prediction,
+  'ti(freq)'=dat_for_freq_prediction,
+  'ti(len)'=dat_for_length_prediction,
+  'ti(prev_surp)'=dat_for_prev_surp_prediction,
+  'ti(prev_freq)'=dat_for_prev_freq_prediction,
+  'ti(prev_len)'=dat_for_prev_len_prediction)
+system.time(res_gptX <- gam_boot(gpt_mean_data,formula2_interact,dats_for_predictionX,N=N_replicates))
+bootstrapped_gam_predictions_gptX <- summarize_gam_boot_results(res_gpt,dats_for_predictionX,term_names,"GPT-2")
+write_rds(res_gptX,"res_gptX.rds")
+write_rds(bootstrapped_gam_predictions_gptX,"bootstrapped_gam_predictions_gptX.rds")
 ```
@@
 # as an example, plot effect for surprisal
+summaries <- bootstrapped_gam_predictions_gptX
 summaries <- bootstrapped_gam_predictions_grnn

### 2022-03-01 Papers/Paper/manuscript.Rmd

@@
 authornote: |
-  TODO
+   We thank the AMLAP 2020 audience, the Computational Psycholinguistics Lab at MIT, the Language and Cognition Lab at Stanford, the QuantLang Lab at UC Irvine, and Mike Frank for feedback on this work.
 
@@
 Participants were able to comprehend these longer text passages that they read via the Maze task.
-Moreover, the Maze task yielded useable reaction time data with word predictability effects that are linear in the surprisal of the current word, the same pattern found with other tasks.
+Moreover, the Maze task yielded useable reaction time data with word predictability effects that were linearly related to surprisal, the same pattern found with other incremental methods.
 Crucially, Maze reaction times show a tight relationship with properties of the current word, but little spillover of effects from previous words.
 This superior localization is an advantage of Maze compared with other methods.
-Overall, we expand the scope of experimental materials, and thus theoretical questions, that can be coupled with the Maze task.
-
-keywords: TODO
-wordcount: TODO
+Overall, we expanded the scope of experimental materials, and thus theoretical questions, that can be studied with the Maze task.
 
@@
 library(tidymv)
-
+library(kableExtra)
 r_refs("r-references.bib", append=F)
@@
 ```
+
 # Introduction
 
-Sometimes we stumble in reading, where we expected a sentence to go in one direction and it did not, and then are at a parsing dead-end where we have to read a sentence again.
-The words that cause stumbles are the outliers on a distribution of words difficulty to read, within fluent reading, there is variation in how easy words are to read and process in their context.
-We're able to read many long or surprising words without noticing a problem, but these less expected words do take longer to process as we rebuild our burgeoning mental model of the sentence.
-Fortunately for fluent readers and unfortunately for studying language, this process of adjustment is very quick, with millisecond scale differences in reading time between words.
+Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing.
+Maze is a potentially promising method, and it's feasibility has recently increased due to the introduction of A-Maze [@boyceMazeMadeEasy2020].
+We tested A-maze on a naturalistic corpora to learn whether participants could understand stories read in the demanding Maze task and whether the reaction time profiles from the Maze task matched those from other methods.
+To provide context, we first lay out the role of incremental processing methods in theory testing, describe a few common methods, and review some key influences on reading time.
 
-We don't have a thorough understanding of how the mind processes language, but we can learn something about what is going on from measuring how long it takes in different circumstances.
-This relies on a linking hypothesis that processing time reflects something about how much processing is going on, even if we don't know whether this is character recognition, memory retrieval, or parsing.
-This is enough to compare theories that predict the relative difficulty, and thus relative processing time, of two words in different contexts.
+## Why measure reading time?
 
-For instance, different theories about what makes object relative clauses harder to understand than subject relative clauses make different predictions about which words are the loci of the overall difficulty and slower reading times associated with object relatives [@grodnerConsequencesSerialNature2005; @staubEyeMovementsProcessing2010; @traxlerProcessingSubjectObject2002].
+Occasionally we read an unexpected word, cannot make sense of it and have to read the sentence again.
+The words that cause these stumbles are the outliers on a distribution of words varying in how easy they are to process in context.
+Even when we don't notice a problem, less expected words still take longer to process as we (re)construct a mental model of the sentence.
+Fortunately for reading and unfortunately for studying language, this process of adjustment is very quick, with millisecond scale differences in reading time between words.
+
+If we assume that processing time reflects something about the processing, be it character recognition, memory retrieval, parsing, or something else, then we can learn something about how the mind processes language from measuring how long the processing takes in different circumstances.
+With this assumption, psycholinguists can test theories that predict the relative difficulty of two constructions by measuring the relative processing time of these words in context.
+
+For instance, theories about what makes object relative clauses harder to understand than subject relative clauses make different predictions about which words are the loci of the overall difficulty and slower reading times associated with object relatives [@grodnerConsequencesSerialNature2005; @staubEyeMovementsProcessing2010; @traxlerProcessingSubjectObject2002].
 Measures of reading time can also inform theories about the time course of processing (i.e.
@@
 Some of these theories rely on being able to attribute processing slowdowns to a particular word.
-Determining that object relatives are overall slower that subject relatives is easy, even an imprecise measure of reading time will determine that the same set of words in a different order took longer to read on a sentence level.
-However, many language processing theories make specific (and contrasting) theories about which words in a sentence should be harder to process.
-To adjudicate these theories, we want methods that are *localized*, where it is easy to determine which word is responsible for an observed slow-down in reading time.
+Determining that object relatives are overall slower that subject relatives is easy.
+Even an imprecise measure of reading time will determine that the same set of words in a different order took longer to read at a sentence level.
+However, many language processing theories make specific (and contrasting) theories about which words in a sentence are harder to process.
+To adjudicate these theories, we want methods that are *localized*, so it is easy to determine which word is responsible for an observed slow-down in reading time.
 Ideally, a longer reading time on a word would be an indication of that word's increased difficulty, and not the lingering signal of a prior word's increased difficulty.
@@
 
-## Incremental processing methods
+## Eye-tracking and Self-paced reading
 
@@
 The two most commonly used methods are eye-tracking and self-paced reading, but both of these suffer from a lack of localization.
+
 In eye-tracking, participants read a text on a screen naturally, while their eye-movements are recorded [@raynerEyeMovementsReading1998].
-The downside of this natural reading is that people often skip short words and look ahead or look back as they read, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982].
+The downside of natural reading is that people often skip short words and look ahead or look back as they read, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982].
+
 Self-paced reading (SPR) is a more controlled process where a participant sees one word on screen at a time and presses a button to see the next word instead.
-Even in this method, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006].
-Both these methods are prone to spillover, as the time to plan motor movements like saccades or button presses is substantial compared to the time it takes to recognize a word, these movements are often initiated before a word is fully processed, so by the time any late processing difficulties occur, the reader may already be several words along.
+Even in SPR, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006].
 
-### the Maze task
-An alternative method that is designed to increase localization at the expense of naturalness is the Maze task [@freedman85; @forsterMazeTaskMeasuring2009].
-In the Maze task, participants see two words at a time, a correct word that continues the sentence, and a distractor which does not.
+Both these methods are prone to spillover.
+The time to plan motor movements like saccades or button presses is substantial compared to the time it takes to recognize a word, these movements are often initiated before a word is fully processed.
+By the time any late processing difficulties occur, the reader may already be several words along.
+
+## Maze
+The Maze task is an alternative method that is designed to increase localization at the expense of naturalness [@freedman85; @forsterMazeTaskMeasuring2009].
+In the Maze task, participants see two words at a time: a correct word that continues the sentence, and a distractor which does not.
 Participants must choose the correct word, and their time to selection is treated as the reaction time (RT).
-@forsterMazeTaskMeasuring2009 introduces two versions of the Maze task: lexical L-maze where the distractors are nonce words and grammatical G-maze where the distractors are real words that don't fit with the context of the sentence so far.
-Theoretically, participants must fully integrate each word into the sentence in order to confidently select it; this may require mentally reparsing previous material in order to allow the integration and selection of a disambiguating word.
-@forsterMazeTaskMeasuring2009 call this forced incremental processing to distinguish from other incremental processing methods where words can be passively read before later committing to a parse.
-This idea of strong localization is supported by studies finding strongly localized effects for G-maze which is more sensitive than L-maze [@witzelComparisonsOnlineReading2012a].
+@forsterMazeTaskMeasuring2009 introduced two versions of the Maze task: lexical L-maze where the distractors are nonce words and grammatical G-maze where the distractors are real words that don't fit with the context of the sentence.
+Theoretically, participants must fully integrate each word into the sentence in order to confidently select it, which may require mentally reparsing previous material in order to allow the integration and selection of a disambiguating word.
+@forsterMazeTaskMeasuring2009 call this need for full integration "forced incremental processing" to distinguish from other incremental processing methods where words can be passively read before later committing to a parse.
+This idea of strong localization is supported by studies finding strongly localized effects for G-maze [@witzelComparisonsOnlineReading2012a; @boyceMazeMadeEasy2020].
 
-The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitious words as distractors for each spot of each sentence; this may explain why the Maze task has not been widely adopted.
-@boyceMazeMadeEasy2020 demonstrate a way to automatically generate Maze distractors by using NLP language models to find words that are high-surprisal in the context of the target sentence, and thus likely to be judged infelicitous by human readers.
-@boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzelComparisonsOnlineReading2012a.
-A-maze outperformed L-maze and an SPR control in detecting and localizing expected slowdown effects.
+The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitous words as distractors for each spot of each sentence.
+This burdensome preparation may explain why the Maze task has not been widely adopted.
+@boyceMazeMadeEasy2020 demonstrated a way to automatically generate Maze distractors by using language models from Natural Language Processing to find words that are high surprisal in the context of the target sentence, and thus likely to be judged infelicitous by human readers.
+@boyceMazeMadeEasy2020 call Maze with automatically generated distractors A-maze.
+In a comparison, A-maze distractors had similar results to the hand-generated G-maze distractors from @witzelComparisonsOnlineReading2012a and A-maze outperformed L-maze and an SPR control in detecting and localizing expected slowdown effects.
 @sloggettAmazeAnyOther2020 also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.
 
-Another recent variant of the Maze task is interpolated or I-maze, which uses a mix of real word distractors (generated via the A-maze process) and nonce word distractors [@vaniUsingInterpolatedMaze2021].
+Another recent variant of the Maze task is interpolated I-maze, which uses a mix of real word distractors (generated via the A-maze process) and nonce word distractors [@vaniUsingInterpolatedMaze2021; @wilcoxTargetedAssessmentIncremental2021].
 The presence of real word distractors encourages close attention to the sentential context, while nonce words can be used as distractors where the word in the sentence is itself ungrammatical or highly unexpected.
-@vaniUsingInterpolatedMaze2021 used I-maze to compare object- and subject- relative clauses and @wilcoxTargetedAssessmentIncremental2021 used it to compare differences in RT for pairs of grammatical and ungrammatical sentences NLP language models predictions.
 
-## Frequency, length, and surprisal effects
-One way of assessing how much a method localized RT effects is to look at how strongly properties of a word correlate with the RT of that word compared with RTs of downstream words later in the sentence.
-A couple of properties known to influence reading time are a word's length and overall frequency in a language, as longer and less common words take longer to process [@klieglLengthFrequencyPredictability2004].
-Another measure is predictability, or how expected a word is given the context, with more predictable words being read faster [@raynerEffectsFrequencyPredictability2004].
-A word can have low predictability for a number of reasons: it could be low frequency, semantically unexpected, the start of a low-frequency syntactic construction, or a word that disambiguates prior words to the less common parse.
+## Measuring localization: Frequency, length, and surprisal effects
+
+Localized measures can be used to attribute processing difficulty to individual words; however, to determine if a method is localized requires knowing how hard the words were to process.
+One approach is to look at properties of words that are known to influence reading times across methods such as eye-tracking and SPR.
+Longer words and lower frequency words tend to take longer to process [@klieglLengthFrequencyPredictability2004], as do less predictable words [@raynerEffectsFrequencyPredictability2004].
+
+A word can be unpredictable for a variety of reasons: it could be low frequency, semantically unexpected, the start of a low-frequency syntactic construction, or a word that disambiguates prior words to a less common parse.
 Many targeted effects of interest are essentially looking at specific features that contribute to how predictable or unpredictable a word is.
-This means that how good a method is at detecting and localizing effects of predictability is a key aspect of how useful it is as measure of incremental processing.
-With al these factors potentially contributing to predictability, how is predictability as a whole measured?
-The typical method is to use language models that are trained on large corpora of language to predict what word comes next in a sentence.
-A variety of pre-trained models exist, that vary in their internals, but all of them will generate assessments of how predictable words are.
-Predictability is often represented in terms of bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time etc).
-The functional form of the relationship between RTs from eye-tracking and SPR corpora and the predictability of the words is linear in terms of surprisal [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
-Due to spillover effects, this linear relationship also holds between the surprisal of a previous word and the RT on the current word.
+Thus incremental processing methods that are sensitive to predictability are useful for testing linguistic theories that make predictions about what words are unexpected.
+
+The overall predictability of a word in a context can be estimated using language models that are trained on large corpora of language to predict what word comes next in a sentence.
+A variety of pre-trained models exist, with varied internal architectures and training methods, but all of them generate measures of predictability.
+Predictability is often measured in bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time etc).
+
+The functional form of the relationship between RTs from eye-tracking and SPR studies and the predictability of the words is linear in terms of surprisal [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
+Predictability reliable correlates with reading time in the range of surprisals found in natural-sounding texts, not just for words that are extremely expected or unexpected [@smithEffectWordPredictability2013].
+If Maze RTs reflect the same processing as other methods, we expect to find a similar linear relationship with surprisal.
 
 ## Current experiment
-The Maze task has thus far only been used on constructed sentences focusing on targeted effects and not on the sorts of long naturalistic passages used to assess the relationship between RT and surprisal.
-We want to test whether participants can read and understand stories while doing the demanding Maze task and also whether the RT profiles from the Maze task are similar to those from other methods.
 
-The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers.
-At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions.
-Taken together, these properties are a severe test of A-maze, as these features may make it harder to auto-generate appropriate distractors and require focus from participants.
-If participants can succeed at the Maze task on Natural Stories, we think they are likely to succeed on a wide variety of naturalistic texts.
-The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension.
+The Maze task has thus far primarily been used on constructed sentences focusing on targeted effects and not on the long naturalistic passages used to assess the relationship between RT and surprisal.
+We tested how A-maze performs on longer naturalistic corpora.
 
-We tweak the A-maze task to accomodate these longer passages and then have participants read the passages in the Maze.
-We compare our A-maze results with SPR data collected on this corpus by @futrellNaturalStoriesCorpus2020.
-We find that participants were able to read and understand these long passages using A-maze, and the A-maze RT profiles show similar relationships to surprisal, word length, and word frequency as those seen with SPR, but with less noise and spillover.
+The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages of roughly 1000 words each which are designed to read fluently to native speakers.
+At the same time, the passages contain copious punctuation, quoted speech, proper nouns, and low frequency grammatical constructions.
+The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we used to assess comprehension.
 
-# Error-correction maze
-<!-- GOOD -->
+We tweaked the A-maze task to accommodate these longer passages and then had participants read the passages in the Maze.
+We compared our A-maze results with SPR data collected on the Natural Stories corpus by @futrellNaturalStoriesCorpus2020.
+The A-maze data and SPR comparison data allow us to address the following questions:
+
+  1. Do participants engage successfully with the A-maze task?
+  2. Is A-maze as reliable as SPR?
+  3. Are the effects of surprisal on RT linear for the A-maze task?
+  4. Does A-maze have less spillover than SPR?
+  5. Which language model best fits the RT data from A-maze and SPR?
+
+# Methods
+
+We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the Maze paradigm.
+The materials, data, and analysis code are all available at REPO.
+
+## Task: Error-correction Maze
+
 (ref:diagram-cap) Schematic of error-correction Maze.
 A participant reads a sentence word by word, choosing the correct word at each time point (selections marked in blue ovals).
-When they make a mistake, an error message is displayed, so they try again and continue with the sentence.
+When they make a mistake, an error message is displayed, so the participant can try again and continue with the sentence.
 
@@
 
+In order to support longer materials, we tweaked the Maze task, creating a new variant called error-correction Maze.
+
 One of the benefits of the Maze task is that it forces incremental processing by having participants make an active choice about what the next word is.
-But what if they choose incorrectly?
-In the traditional Maze paradigm, there isn't a way to handle this; a mistake indicates they're not understanding the sentence properly, so the sentence ends, and the participant moves on to the next item [@forsterMazeTaskMeasuring2009].
-An advantage of this is that participants who contribute RT data are very likely to have understand the sentence up to that point (it's hard to choose correctly in G-maze otherwise).
-This contrasts with other methods, where determining whether participants are paying attention usually requires separate comprehension check questions, that are usually not used for Maze.
+But what happens if they choose incorrectly?
+In the traditional Maze paradigm, any mistake ends the sentence, and the participant moves on to the next item [@forsterMazeTaskMeasuring2009].
+An advantage of this is that participants who contribute RT data are very likely to have understood the sentence up to that point.
+This contrasts with other methods, where determining whether participants are paying attention usually requires separate comprehension check questions, usually not used for Maze.
 
-However, terminating sentences on errors means that we don't have data after a participant makes a mistake in an item.
-In traditional G-maze tasks, with hand-crafted distractors and attentive participants, this is a small issue.
+However, terminating sentences on errors means that we don't have RTs for words after a participant makes a mistake in an item.
+In traditional G-maze tasks, with hand-crafted distractors and attentive participants, errors are rare and data loss is a small issue.
 However, this data loss is much worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020].
@@
 
-This situation could be improved by auto-generating better distractors or hand-replacing problematic ones, but that does not solve the fundamental problem with long items.
-Well-chosen distractors and attentive participants will reduce the error rate, but the error rate will still compound over long materials.
+The high error rates could be improved by auto-generating better distractors or hand-replacing problematic ones, but that does not solve the fundamental problem with long items.
+Well-chosen distractors and attentive participants reduce the error rate, but the error rate will still compound over long materials.
 For instance, with a 1% error rate, `r round(.99**15*100)`%  of participants would complete each 15-word sentence, but only `r round(.99**50*100)`% would complete a 50 word vignette, and `r round(.99**200*100)`% would complete a 200 word passage.
-In order to run longer materials, we need something to do when participants make a mistake, other than terminate the entire item.
+In order to run longer materials, we needed something to do when participants made a mistake other than terminate the entire item.
 
-To resolve this, we introduce an *error-correction* variant of Maze shown in Figure \@ref(fig:diagram).
-When a participant makes an error, we present them with an error message and wait until they select the correct option, before continuing the sentence as normal.
-We make this "error-correction" Maze available as an option in a modification of the Ibex Maze implementation introduced in @boyceMazeMadeEasy2020 (https://github.com/vboyce/Ibex-with-Maze).
+As a solution, we introduce an *error-correction* variant of Maze shown in Figure \@ref(fig:diagram).
+When a participant makes an error, they see an error message and must try again to select the correct option, before continuing the sentence as normal.
+We make error-correction Maze available as an option in a modification of the Ibex Maze implementation introduced in @boyceMazeMadeEasy2020 (https://github.com/vboyce/Ibex-with-Maze).
 The code records both the RT to the first click and also the total RT until the correct answer is selected as separate values.
 
-This variant of Maze expands the types of materials that can be used with Maze to include arbitrarily long passages and cushions the impact of occasional problematic distractors.
+Error-correction Maze expands the types of materials that can be used with Maze to include arbitrarily long passages and cushions the impact of occasional problematic distractors.
 Error-correction Maze is a change in experimental procedure, and is independent of what types of distractors are used.
-This error-correction presentation is used here with A-maze, but would also work with G-maze or I-maze.
+This error-correction presentation is used here with A-maze, but could also be used with G-maze or I-maze.
 
-# Methods
+## Materials
+We used the texts from the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and their corresponding comprehension questions.
+To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions.
+See Appendix A for an excerpt of one of the stories.
+
+To generate distractors, we first split the corpora up into sentences, and then ran the sentences through the A-maze generation process.
+We used an updated version of the codebase from @boyceMazeMadeEasy2020 which had the capability to match the greater variety of punctuation present in the Natural Stories corpus (updated auto-generation code at https://github.com/vboyce/Maze).
+We took the auto-generated distractors as they were, without checking  for quality.
+
+## Participants
 
@@
   left_join(data_error_summ, by=c("subject", "num_correct")) %>% 
-  mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct")) %>% 
-  group_by(num_correct,accurate)
+  filter(pct_correct>.8)
 
@@
   mutate(sent.correct=ifelse(total_mistake==0,1,0))
+
+spr <- read_rds(here("Data/SPR/first.rds")) %>% filter(correct>4) %>% select(WorkerId) %>% unique()
 ```
 
-We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the Maze paradigm.
-
-## Materials
-We used the texts from the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and generated A-maze distractors for them.
-We used the original comprehension questions provided in the Natural Stories corpus.
-To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions.
-See the Appendix for an except of one of the stories and its corresponding comprehension questions.
-All materials are available at TODO NEW REPO.
-
-To generate distractors, we first split the corpora up into sentences, and then ran the sentences through the A-maze generation process.
-We used an updated version of the codebase from @boyceMazeMadeEasy2020.
-This newer version had the capability to match the greater variety of punctuation present in this corpus (updated auto-generation code at https://github.com/vboyce/Maze).
-We took the auto-generated distractors as they were, without checking them for quality.
-
-## Participants
 We recruited 100 participants from Amazon Mechanical Turk in April 2020, and paid each participant $3.50 for roughly 20 minutes of work.
 We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
+After examining the participants performance on the task (see Results for details), we excluded data from participants with less than 80% accuracy, removing participants whose behavior was consistent with random guessing.
+After this exclusion, `r some %>% select(subject) %>% unique() %>% nrow()` participants were left.
 
@@
 Participants first gave their informed consent and saw task instructions.
-Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading the main story in the A-maze task.
-After the story, they answered the 6 main comprehension questions, commented on their experience, answered optional demographic questions, and were debriefed, before getting a code to enter for payment.
-The experiment was implemented in Ibex (https://github.com/addrummond/ibex) and the experimental code is available at REPO.
+Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading one main story in the A-maze task.
+After the story, they answered 6 comprehension questions, commented on their experience, answered optional demographic questions, were debriefed, and were given a code to enter for payment.
+The experiment was implemented in Ibex (https://github.com/addrummond/ibex).
 
-## Models
+## Self-paced reading comparison
+
+In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran in 2011.
+They recruited 181 participants from Amazon Mechanical Turk, most of whom read 5 of the stories.
+After reading each story, each participant answered 6 binary-choice comprehension questions.
+For comparability with A-maze, we analyze only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct, leaving `r spr %>% nrow()`.
+
+## SPR-Maze correlation
+We compared the correlations between the Maze and SPR RTs to within-Maze and within-SPR correlations.
+For Maze, within each story, we randomly split subjects into two halves.
+Within each half, we calculated a per-word average RT for each word and then a per-sentence average RT across word averages.
+We calculated a within-Maze correlation between these two halves.
+
+For this comparison, we downsampled the SPR data choosing a number of participants equal to the number we have for Maze to avoid differences due to dataset size.
+We then used the same procedure to get a within-SPR correlation.
+For between Maze-SPR correlation, we took the average correlation across each of the 4 pairs of Maze half and SPR half.
 
@@
 package_list <- c("tidyverse","brms","rstan","papaja",
-                  "lme4", "mgcv", "tidybayes")
+                  "lme4", "mgcv", "tidybayes", "here", "patchwork","cowplot","gridExtra","broom.mixed","tidymv","kableExtra")
+
+a <- cite_r("r-references.bib", pkgs=package_list, withhold=F, footnote=T)
+
 ```
 
-In order to measure the relationship between a word's properties and it's RT, we fit models using surprisal, frequency, and length as predictors of RT.
-We considered these predictors from both the current and past word to look for the possibility of spill over effects.
-We fit generalized additive models (GAMs) to look at the functional form of the surprisal-RT relationship and linear models (LMs) to look at the size of the effects.
-As described in the Results section, we base all models only on data from participants who choose correctly on at least 80% of the words in the Maze task.
+## Modeling approach
 
- TODO FIX ME We conducted data processing and analyses using `r cite_r("r-references.bib", withhold=F, pkgs=package_list)`.
+Our analytic questions required multiple modeling approaches.
+To look at the functional form of the relationship between surprisal and RT data, we fit Generalized Additive Models (GAMs) to allow for non-linear relationships.
+GAMs are harder to interpret than linear models, so to measure effect sizes and assess spillover, we used linear mixed models.
+Finally, in order to determine which language model best predicts the RT data, we fit additional linear models with predictors from multiple language models to look at their relative contributions.
+All these models used surprisal, frequency, and length as predictors for RT.
+We considered these predictors from both the current and past word to account for the possibility of spillover effects in A-maze.
+For SPR comparisons, we included predictors from the current and past three words to account for known spillover effects.
+We conducted data processing and analyses using `r a$r`.
+
+`r a$pkgs`
 
@@
 ```
+
 ### Predictors
+
 We created a set of predictor variables of frequency, word length, and surprisals from 4 language models.
@@
 For unigram frequency, we tokenized the training data from @gulordava18 and tallied up instances.
-We then rescaled it to be the log2 frequency of expected occurrences in 1 billion words as the model predictor, so higher values indicate higher log frequencies.
-We got per-word surprisals for each of 4 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].
-These models cover a range of common architectures.
-We can infer robustness if models provide similar results, but also compare between models to look at how well each model fits human data.
-For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.
+We then rescaled the word counts to get the log2 frequency of occurrences per 1 billion words, so higher values indicate higher log frequencies.
+We got per-word surprisals for each of 4 different language models, covering a range of common architectures: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].
+For all of these predictors, we used both the predictor at the current word as well as lagged predictors from the previous word.
 
 ### Exclusions
-We exclude the first word of every sentence because it had an x-x-x distractor, which leaves `r nrow(not_first)` words.
-We exclude words for which we don't have surprisal or frequency information, leaving `r nrow(not_na)` words.
-We additionally exclude words that any model treated as being composed of multiple tokens.
-While surprisals should be additive, the summing the surprisals over these tokens gives some unreasonable responses.
-For instance, in one story the word king!\' is given a surprisal of 64 by GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, \'What is the use of flying any higher?
+
+We excluded the first word of every sentence because it had an x-x-x distractor, leaving `r nrow(not_first)` words.
+We excluded words for which we didn't have surprisal or frequency information, leaving `r nrow(not_na)` words.
+We additionally excluded words that any model treated as being composed of multiple tokens (primarily words with punctuation),  leaving `r nrow(not_gpt)` words[^1].
+We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
+We exclude RTs from words where mistakes occurred or which occurred after a mistake in the same sentence.
+We only analyzed words where we had values for all predictors, which meant that if the previous word was unknown to a model, the word was excluded because of missing values for a lagged predictor.
+
+[^1]:Surprisals should be additive, but summing the surprisals for multi-token words gave some unreasonable responses.
+For instance, in one story the word king!\' has a surprisal of 64 under GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, \'What is the use of flying any higher?
 This victory is in the bag and I am king!\').
-To avoid these outliers we exclude all words that any model treated as multi-token, leaving `r nrow(not_gpt)` words.
-This primarily excludes words with punctuation.
-(While GPT-2 using byte-pair encoding that can split up words into multiple parts, excluding words it split up only excludes 30 words that were not already excluded by other models.)
-
-We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
-We exclude words where mistakes occurred or which occurred after a mistake in the same sentence.
-We only analysed words where we had values for all predictors, which means that if the previous word was unknown to a model, this word will be excluded because we're missing values for a lagged predictor.
+While GPT-2 using byte-pair encoding that can split up words into multiple parts, excluding words it split up only excluded 30 words that were not already excluded by other models.
 
@@
 
-To account for by-subject variation, we do bootstrapping ... TODO specs for bootstrapping
+For linear models, we centered all predictors.
+We modeled the main effects of surprisal, length, and frequency as well as surprisal x length and frequency x length interactions.
+For the A-maze data, we used maximal mixed effects, including by-subject slopes and a per-word-token random intercept [@barrRandomEffectsStructure2013].
+We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations) and ran models with brm [@burknerAdvancedBayesianMultilevel2018].
 
-For linear models, we centered all predictors.
-We used full mixed effects, including by-subject slopes and a per-word-token random intercept [@barrRandomEffectsStructure2013].
-We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations).
-Models were run in brm [@burknerAdvancedBayesianMultilevel2018].
+For linear models of the SPR data, we were unable to fit the full mixed effects structure.
+The best model we could fit had by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and a per-word-token random intercept, fit with lme4 [@batesFittingLinearMixedeffects2015], as this structure did not fit reliably in brm.
 
 For model comparison, we took by-item averaged data to aid in fast model fitting.
-We included frequency, length, and their interaction to all models.
+We included frequency, length, and their interaction in all models.
 Then we fit models with either 1 or 2 sources of surprisal using lm [@batesFittingLinearMixedeffects2015] and assessed the effect of adding the second surprisal source with an anova.
-We used predictors for the current and past word and centered all effects.
-
-## Self-paced reading comparison
-
-In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran in 2011.
-They recruited 181 participants from Amazon Mechanical Turk, most of whom read 5 of the stories.
-After reading each story, each participant answered 6 binary-choice comprehension questions.
-As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020].
-
-For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct.
-To account for spill over effects known to exist in SPR, we analyse predictors at the current word as well as the past 3 words for all models.
-For linear models, we centered all predictors.
-We were unable to fit the full mixed effects model.
-The best model we could fit had by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and  a per-word-token random intercept, fit with lme4, as this structure did not fit reliably in brms.
-
-## SPR-Maze correlation
-We additionally compare how correlated the Maze and SPR results are to each other, in comparison to within-Maze and within-SPR correlations.
-For Maze, within each story, we randomly split subjects into two halves.
-Within each half, we calculate a per-word average for each word and then a per-sentence average RT across word averages.
-We calculate a within-Maze correlation between these two halves.
-To avoid differences due to dataset size, we downsample the SPR data choosing a number of participants equal to the number we have for Maze.
-We then use the same procedure on this subset as for Maze to get a within-SPR correlation.
-For between Maze-SPR correlation, we take the average correlation across each of the 4 pairs of Maze half and SPR half.
 
@@
 
-## Reading in the Maze
+## Do participants engage successfully?
 
 (ref:error-cap) A.
-Correlation between a participant's accuracy on the Maze task (fraction of words selected correctly) and their average reaction time (in ms).
-Many participants (marked in green) chose the correct word >80% of the time; others (in red) appear to be randomly guessing and were excluded from further analysis.
+Participant's accuracy on the Maze task (fraction of words selected correctly) versus their average reaction time (in ms).
+Many participants (marked in green) chose the correct word >80% of the time; others (in red) appear to be randomly guessing.
 B.
 Performance on the comprehension questions.
-Participants with low accuracy also performed poorly on comprehension questions; Participants with >80% task accuracy tended to do well; their performance was roughly comparable to the performance of SPR participants from  @futrellNaturalStoriesCorpus2020 on their first stories.
+Participants with low accuracy performed poorly on comprehension questions; Participants with >80% task accuracy tended to do well; their performance was roughly comparable to the performance of SPR participants from  @futrellNaturalStoriesCorpus2020 on their first stories.
 
@@
 
-We first looked at participants overall performance on the Maze task and comrehension questions, in order to see if participants were able to read and understand a story in the Maze task.
-Many participants were able to complete the Maze task with a high degree of accuracy and answer the comprehension questions correctly.
+Our first question was whether participants could engage successfully with the  error-correction Maze task.
+We assessed engagement by looking at participants' accuracy on the Maze task and performance on the comprehension questions.
 
@@
 
-Participant accuracy, or how often participants choose the correct word over the distractor, reflects both the quality of the distractors and the focus and skill of the participant.
-We calculated the per-word error rate for each participant and graphed it against their average reaction time.
-(To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded; this excluded `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) made relatively few errors, with some reaching 99% accuracy.
-This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story.
+Accuracy, or how often a participant chose the correct word over the distractor, reflects both the quality of the distractors and the focus and skill of the participant.
+We calculated the per-word accuracy rate for each participant and compared it against their average reaction time [^2].
+As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) made relatively few errors, with some reaching 99% accuracy.
+This high performance confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story.
 These careful participants took around 1 second for each word selection, which is much slower than in eye-tracking or SPR.
+
+[^2]: To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.
+This exclusion removed `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.
+
 Another cluster of participants (in red) sped through the task, seemingly clicking randomly.
 This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.
+We believe the high level of random guessing is an artifact of the subject population [@hauser2018, and we expect that following current recommendations for participant recruitment, such as using qualification cutoffs or another recruitment site would result in fewer participants answering randomly [@eyal2021;@peer2017].
 
-Another check is whether participants comprehended the story.
-We counted how many of the binary-choice comprehension questions each participant got right (out of 6).
-As seen in Figure \@ref(fig:errors)B, most participants who were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions.
-Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet.
+To determine comprehension accuracy, we counted how many of the binary-choice comprehension questions each participant got right (out of 6).
+As seen in Figure \@ref(fig:errors)B, most participants who were accurate on the task also did well on comprehension questions, while participants who were at chance on the task were also at chance on the comprehension questions.
+Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the Internet.
 We can't rule out that some participants may have been able to guess the answers without understanding the story.
-Nonetheless, this provides preliminary evidence that people can understand and remember details of stories they read during the Maze task.
+Nonetheless, the accurate answers provide preliminary evidence that people can understand and remember details of stories they read during the Maze task.
+
 The comprehension question performance of accurate Maze participants is broadly similar to the performance of SPR participants from @futrellNaturalStoriesCorpus2020 on the first story read.
 Overall, `r comp_tally %>% filter(source=="all Maze") %>% pull(good)`% of Maze participants got 5 or 6 questions right (`r comp_tally %>% filter(source=="<80% correct") %>% pull(good)`% of low-accuracy participants and `r comp_tally %>% filter(source==">80% correct") %>% pull(good)`% of high-accuracy participants) compared to `r comp_tally %>% filter(source=="SPR") %>% pull(good)`% of all SPR reads and `r comp_tally %>% filter(source=="SPR 1st story") %>% pull(good)`% of 1st SPR reads.
-Note that these differences cannot be directly attributed to methods, as the participant populations differed.
-While both studies were conducted on Mturk, the quality of Mturk data has decreased from 2011 when the SPR was collected to 2020 when the A-maze was collected.
+These differences cannot be directly attributed to methods, as the participant populations differed.
+While both studies were conducted on Mturk, the quality of Mturk data has decreased from 2011 when the SPR was collected to 2020 when the A-maze was collected [@chmielewski2020].
 
-We use task performance as our exclusion metric for A-maze because it is more fine-grained and only analyze data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
-For the SPR comparison, we follow @futrellNaturalStoriesCorpus2020's criteria and exclude participants who got less than 5 of the questions correct.
+For the remainder of the analyses, we use task performance as our exclusion metric for A-maze because it is more fine-grained and only analyze data from participants with at least 80% accuracy (in the gap between high-performers and low-performers).
+For the SPR comparison, we follow @futrellNaturalStoriesCorpus2020's criteria and exclude participants who got less than 5 of the comprehension questions correct.
 
-## RT and surprisal
-
-Given that the Maze task worked, we next consider the relationship between surprisals and Maze RTs.
-Surprisal, a measure of overall word predictability in context, is linearly related to RT in eye-tracking and SPR.
-If Maze is measuring the same language processes, we would expect to see a linear relationship in the Maze task.
-Additionally, surprisal can be used to assess spillover, but comparing the predictive value of current and past word surprisal on current word reading time.
-
-### GAMs
-
-```{r}
-subject_att <- data_error_summ %>% 
-  select(subject, is.attentive)
-
-data_low_error <- data_filt %>% 
-  left_join(subject_att, by="subject") %>% 
-  filter(is.attentive==1) %>% 
-  filter(type!="practice")
-
-  data_error_free <- data_low_error %>% 
-    mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
-    group_by(sentence, subject) %>% fill(word_num_mistake) %>% ungroup() %>% 
-    mutate(after_mistake=word_num-word_num_mistake,
-           after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
-    filter(correct=="yes") %>% 
-    filter(!after_mistake %in% c(1,2))
-
-data_no_first <- data_error_free %>% filter(word_num!=0)
-
-data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
-  select(subject, word_num, word, rt, sentence, type)
-
-data_post_only <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
-  filter(!after_mistake==0) %>% 
-    select(subject, word_num, word, rt, sentence, type)
-
-data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
-  filter(after_mistake==0) %>% 
-  select(subject, word_num, word, rt, sentence, type)
-
-data_stories <- data_ready %>% select(type, subject) %>% 
-  unique() %>% 
-  group_by(type) %>% 
-  tally()
-
-data_anything_goes <- data_filt %>% 
-  filter(type!="practice") %>% 
-  filter(correct=="yes") %>% 
-    select(subject, word_num, word, rt, sentence, type)
-
-```
-
-To assess the shape of the RT-surprisal relationship, we fit generalized additive models (GAMs).
-For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.
-We use surprisals generated by 4 different language models for robustness.
-(See Methods for details on language models, exclusions, and model fit.)
-
-(ref:gam-cap) GAM results for the effect on Maze reaction time (RT) from either current word surprisal (top) or previous word surprisal (bottom).
-Density of data is shown along the x-axis.
-For each of the 4 language models used, there is a linear relationship between current word surprisal and RT (at least when there is enough data).
-The relationship between previous word surprisal and RT is much flatter.
-TODO these are bootstrapped 95% CI?
-
-```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:gam-cap)"}
-
-gam1 <- read_rds(here("Analysis/bootstrapped_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
-  geom_line()+
-  geom_ribbon(alpha=.3)+
-  #facet_grid(s~model, scales="free_y")+
-  facet_grid(s~model)+
-  coord_cartesian(xlim=c(0,28))+
-  labs(x="Surprisal (bits)", y="Effect on Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
-
-all_mean_data <-  read_rds(here("Analysis/models/meaned_maze.rds"))
-
-dens2 <-   ggplot(all_mean_data, aes(x=value))+
-  geom_density(fill="gray",)+
-  facet_grid(.~model)+
-  labs(x="Surprisal (bits)", y="")+
-    coord_cartesian(xlim=c(0,28))+
-  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
-
-p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
-plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))
-
-```
-
-The relationships between current and previous word surprisals and RT are shown in Figure \@ref(fig:gam); this is the main effect of either current or previous surprisal.
-Note that for each of the models, high-surprisal words are rare, with much of the data for words between 0 and 15 bits of surprisal.
-All of the models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data.
-All of the models show a flatter relationship between previous word surprisal and RT.
-This is a sign of localization as the previous word's surprisal is not affecting RT much.
-
-TODO comment on freq, length effects / non-linearities!
-
-(ref:spr-gam-cap) GAM results for the effect on SPR RT data [@futrellNaturalStoriesCorpus2020] as a function of current word surprisal (top) or the surprisal of an earlier word, up to 3 words back.
-Density of data is shown along the x-axis.
-TODO these are bootstrapped 95% CI?
-
-```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
-gam1 <- read_rds(here("Analysis/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
-  geom_line()+
-  geom_ribbon(alpha=.3)+
-  facet_grid(s~model)+
-  coord_cartesian(xlim=c(0,28))+
-  labs(x="Surprisal (bits)", y="Effect on Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
-
- all_mean_spr <-  read_rds(here("Analysis/models/meaned_spr.rds"))
-
-dens2 <-   ggplot(all_mean_spr, aes(x=value))+
-  geom_density(fill="gray",)+
-  facet_grid(.~model)+
-  labs(x="Surprisal (bits)", y="")+
-    coord_cartesian(xlim=c(0,28))+
-  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
-
-p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
-plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .2))
-
-```
-
-As a comparison, we also ran GAMs on the Self-Paced Reading data collected by @futrellNaturalStoriesCorpus2020.
-Previous work such as @smithEffectWordPredictability2013 has found positive relationships between RT and the surprisal of earlier words for SPR, so we include predictors from the 3 prior words.
-The  relationship between surprisals and RT is shown in Figure \@ref(fig:spr-gam); note that the y-axis range is much narrower than for Maze.
-Both current and previous word surprisals have a roughly linear positive relationship to RT.
-The surprisal of the word two back also has an influence in some models.
-
-Comparing the Maze and SPR smooths, we see that both show a linear relationship, but Maze has much larger and more localized effects of surprisal.
-
-### Effect sizes
-
-Given that the GAM models show a roughly linear relationship between RT and surprisal, we are justified in using surprisal as a predictor in a linear model to quantify its effect.
-In addition to surprisal, we also include the commonly studied effects of frequency and length as well as surprisal x length and frequency x length interactions.
-For all of these, we include the predictors for the current and previous word, and we centered, but did not rescale, all predictors.
-(See Methods for more details on these predictors and model fit process.)
-
-```{r pre-error}
-summ <- read_rds(here("Analysis/models/brms_maze_summ.rds")) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
-  mutate(Term=factor(Term, 
-                        levels=c("Intercept", 
-                                 "surprisal",
-                                 "len",
-                                 "freq",
-                                 "surprisal:len",
-                                 "len:freq",
-                                 "prev_surp",
-                                 "prev_len",
-                                 "prev_freq",
-                                 "prev_surp:prev_len",
-                                 "prev_len:prev_freq")
-                                 )) %>% 
-  arrange(Term) %>% 
-  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency",
-                  "Surp x Length", "Freq x Length", "Past Surprisal",
-                  "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length"))
-
-knitr::kable(summ, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models.
-All terms were centered, but not rescaled.
-Units are in ms.
-Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.
-Interval is 2.5th quantile to 97.5th quantile of model draws.")
-```
-
-The Maze linear model effects are shown in Table \@ref(tab:pre-error).
-we found large effects of surprisal and length, but minimal effects of frequency.
-The lack of frequency effects is somewhat surprising, but consistent with @shainLargescaleStudyEffects2019.
-Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word, an indication that spillover is limited and effects are strongly localized.
-
-```{r spr-table}
-
-summ_spr <- read_rds(here("Analysis/models/lmer_spr_summ.rds")) %>% 
-  pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
-  mutate(Term=factor(Term, 
-                        levels=c("(Intercept)", 
-                                 "surp_center",
-                                 "length_center",
-                                 "freq_center",
-                                 "surp_center:length_center",
-                                 "length_center:freq_center",
-                                 "past_surp_center",
-                                 "past_length_center",
-                                 "past_freq_center",
-                                 "past_surp_center:past_length_center",
-                                 "past_length_center:past_freq_center",
-                                 "past2_surp_center",
-                                 "past2_length_center",
-                                 "past2_freq_center",
-                                 "past2_surp_center:past2_length_center",
-                                 "past2_length_center:past2_freq_center",
-                                 "past3_surp_center",
-                                 "past3_length_center",
-                                 "past3_freq_center",
-                                 "past3_surp_center:past3_length_center",
-                                 "past3_length_center:past3_freq_center"
-                                 )
-                                 )) %>% 
-  arrange(Term) %>% 
-  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency","Surp x Length", "Freq x Length",
-                "Past Surprisal", "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length",
-                "2Past Surprisal", "2Past Length", "2Past Freq", "2Past Surp x Length", "2Past Freq x Length",
-                "3Past Surprisal", "3Past Length", "3Past Freq", "3Past Surp x Length", "3Past Freq x Length"))
-
-knitr::kable(summ_spr, format="latex", position="ht",caption="Predictions from fitted regression models for SPR data.
-All terms were centered, but not rescaled.
-Units are in ms.
-Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.
-Uncertainty interval is +/- 1.97 standard error.")
-
-```
-
-We ran similar models for SPR, although to account for known spillover effects, we consider predictors from the current and 3 previous words.
-Due to issues fitting models, the details of the models differed (see Methods).
-The SPR coefficients are shown in Table \@ref(tab:spr-table).
-Surprisal, length, and frequency effects are all evident for the current word and surprisal and frequency show effects from the previous word as well.
-
-Whereas Maze showed surprisal effects in the 10 to 25 ms/bit range and length effects 15-20 ms/character range, SPR effects are about 1/10 this size.
-This difference in effect is disproportionate to the overall speed of the methods; Maze is 2-3 times as slow as SPR.
-
-In addition, the relative effects of these different predictors shows somewhat different patterns; in SPR, frequency effects are roughly as big as surprisal effects, but this is not true in Maze.
-In SPR, length effects exceed surprisal effects, in Maze, this varies depending on the model.
-
-TODO: is there a way to make the tables more readable / highlight the key rows?
-Maybe stick the full tables in the appendix and show select versions omitting interactions and 2- 3- past words in the main text?
-Or maybe it's fine and people can read?
-
-### Model comparison
-
-We use 4 different models for surprisal.
-This provided robustness to the above analyses, but it also allows us to look at which models have the best fit to human Maze and SPR data.
-For each data type, we assessed each models fit using log likelihood and R-squared.
-Then we did a nested model comparison, looking at whether a model with multiple surprisal predictors (ex, GRNN and GPT) had a better fit than a model with only one (ex GRNN alone).
-This tells us whether a model contains predictive value beyond that contained in another model.
-As shown in Table \@ref(tab:maze-compare), GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complementary information over GPT-2.
-TXL provides a lot over 5-gram, and 5-gram provides little over any model.
-The single-model measures of log likelihood confirm this hierarchy, as GPT-2 is better than GRNN is better than TXL is better than 5-gram.
-```{r maze-compare}
-
-read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data.
-Each row shows the additional predictive value gained from adding that model to another model.
-F values and p values from anova tests between 1-surprisal-source and 2-source models are reported.
-We also report log likelihoods of models with only one surprisal source.")
-
-```
-
-We followed the same process for the SPR data with results shown in Table \@ref(tab:spr-compare).
-For SPR, GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN.
-In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small.
-The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020.
-This contrasts with the Maze results, where the 5-gram model had the worst fit and did not provide additional predictive value to the other models.
-
-```{r}
-maze <- read_rds(here("Analysis/maze_model_compare.rds"))
-spr <- read_rds(here("Analysis/spr_model_compare.rds"))
-```
-
-As an overall measure of fit to data, we calculate multiple R-squared for the single surprisal source models for both A-maze and SPR.
-The models predict A-maze better with R-squared values ranging from `r maze %>% filter(Model=="5-gram") %>% pull(r_squared)` for the 5-gram model to `r maze %>% filter(Model=="GPT-2") %>% pull(r_squared)` for GPT-2.
-Whereas for SPR, the R-squared values range from from `r spr %>% filter(Model=="5-gram") %>% pull(r_squared)` to `r spr %>% filter(Model=="GPT-2") %>% pull(r_squared)`.
-This suggests that the effect size differences are not due merely to the larger overall reading time for A-maze, but that instead A-maze is more sensitive to surprisal and length effects.
-
-```{r spr-compare}
-
-read_rds(here("Analysis/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on SPR data.
-Each row shows the additional predictive value gained from adding that model to another model.
-F values and p values from anova tests are reported.
-We also report log likelihoods of models with only one surprisal source.")
-
-```
-
-### Maze - SPR correlation
+## Is A-maze reliable?
 
@@
   inner_join(spr_collapse)
-ggplot(both, aes(x=maze, y=spr))+geom_point(alpha=.5, size=1)+labs(x="A-Maze", y="SPR")+coord_fixed(ratio=1)
+ggplot(both, aes(x=maze, y=spr))+geom_point(alpha=.5, size=1)+labs(x="A-Maze", y="SPR")+coord_fixed(ratio=1)+geom_smooth(method="lm")
 ```
 
-As a last analysis, we measured how correlated SPR and Maze are in their reading times; that is, did the methods pick up on the same effects.
-If so, sentences that take longer to read in one method would also take longer in the other.
-We calculate the average RT at the sentence level because differences in spillover will effect sentence level RT less than RT for individual words (see Methods for details).
-The correlation between Maze and SPR is `r round(spr_maze,2)`, compared to `r round(spr_spr,2)` within SPR and `r round(maze_maze,2)` within Maze.
+Our second question was whether A-maze is reliable.
+To assess reliability, we conducted split-half comparisons looking at the correlations between and within SPR and A-maze.
+If the methods picked up on the same effects, we would expect them to be correlated, with sentences that took longer to read in one method also taking longer in the other.
+We calculated the average RT at the sentence level to reduce variability from spillover patterns.
+The correlation between Maze and SPR was `r round(spr_maze,2)`, compared to `r round(spr_spr,2)` within SPR and `r round(maze_maze,2)` within Maze.
 See Figure \@ref(fig:scatter) for a visual comparison of overall Maze versus SPR RTs.
+SPR data is about as correlated with Maze as with another sample of SPR data which provides some evidence that Maze and SPR are measuring the same effects.
 
-The fact that SPR data is about as correlated with Maze as with another sample of SPR data provides some confirmation that Maze and SPR are measuring the same thing.
+## Are the effects of surprisal linear?
+
+We next considered the relationship between surprisal and Maze RT.
+Surprisal, a measure of overall word predictability in context, is linearly related to RT in eye-tracking and SPR [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
+If Maze is measuring the same language processes, we would expect to see a linear relationship between surprisal and Maze RT.
+
+```{r}
+subject_att <- data_error_summ %>% 
+  select(subject, is.attentive)
+
+data_low_error <- data_filt %>% 
+  left_join(subject_att, by="subject") %>% 
+  filter(is.attentive==1) %>% 
+  filter(type!="practice")
+
+  data_error_free <- data_low_error %>% 
+    mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
+    group_by(sentence, subject) %>% fill(word_num_mistake) %>% ungroup() %>% 
+    mutate(after_mistake=word_num-word_num_mistake,
+           after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
+    filter(correct=="yes") %>% 
+    filter(!after_mistake %in% c(1,2))
+
+data_no_first <- data_error_free %>% filter(word_num!=0)
+
+data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  select(subject, word_num, word, rt, sentence, type)
+
+data_post_only <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  filter(!after_mistake==0) %>% 
+    select(subject, word_num, word, rt, sentence, type)
+
+data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
+  filter(after_mistake==0) %>% 
+  select(subject, word_num, word, rt, sentence, type)
+
+data_stories <- data_ready %>% select(type, subject) %>% 
+  unique() %>% 
+  group_by(type) %>% 
+  tally()
+
+data_anything_goes <- data_filt %>% 
+  filter(type!="practice") %>% 
+  filter(correct=="yes") %>% 
+    select(subject, word_num, word, rt, sentence, type)
+
+```
+
+To assess the shape of the RT-surprisal relationship, we fit generalized additive models (GAMs).
+For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.
+We used surprisals generated by 4 different language models for robustness.
+(See Methods for details on language models, exclusions, and model fit.)
+
+(ref:gam-cap) GAM results for the effect of current word surprisal (top) or previous word surprisal (bottom) on Maze reaction time (RT).
+Density of data is shown along the x-axis.
+The best-fit lines is from the mean estimated effect size across the bootstrap replicates, and the shaded areas indicate a 95\% bootstrap confidence interval on this effect size.
+For each of the 4 language models used, there is a linear relationship between current word surprisal and RT.
+The relationship between previous word surprisal and RT is much flatter.
+
+```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:gam-cap)"}
+
+gam1 <- read_rds(here("Analysis/bootstrapped_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+  geom_line()+
+  geom_ribbon(alpha=.3)+
+  #facet_grid(s~model, scales="free_y")+
+  facet_grid(s~model)+
+  coord_cartesian(xlim=c(0,28))+
+  labs(x="Surprisal (bits)", y="Effect on Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
+
+all_mean_data <-  read_rds(here("Analysis/models/meaned_maze.rds"))
+
+dens2 <-   ggplot(all_mean_data, aes(x=value))+
+  geom_density(fill="gray",)+
+  facet_grid(.~model)+
+  labs(x="Surprisal (bits)", y="")+
+    coord_cartesian(xlim=c(0,28))+
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
+plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))
+
+```
+
+The main effects of current and previous word surprisals on RT are shown in Figure \@ref(fig:gam).
+Note that for each of the models, high-surprisal words are rare, with much of the data from words with between 0 and 15 bits of surprisal.
+All 4 models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data.
+
+(ref:spr-gam-cap) GAM results for the effect of current word surprisal (top) or the surprisal of an earlier word, up to 3 words back on SPR RT data [@futrellNaturalStoriesCorpus2020].
+Density of data is shown along the x-axis.
+The best-fit lines is from the mean estimated effect size across the bootstrap replicates, and the shaded areas indicate a 95\% bootstrap confidence interval on this effect size.
+
+```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
+gam1 <- read_rds(here("Analysis/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+  geom_line()+
+  geom_ribbon(alpha=.3)+
+  facet_grid(s~model)+
+  coord_cartesian(xlim=c(0,28))+
+  labs(x="Surprisal (bits)", y="Effect on Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))
+
+ all_mean_spr <-  read_rds(here("Analysis/models/meaned_spr.rds"))
+
+dens2 <-   ggplot(all_mean_spr, aes(x=value))+
+  geom_density(fill="gray",)+
+  facet_grid(.~model)+
+  labs(x="Surprisal (bits)", y="")+
+    coord_cartesian(xlim=c(0,28))+
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
+plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .2))
+
+```
+
+As a comparison, we also ran GAMs on the SPR data collected by @futrellNaturalStoriesCorpus2020.
+Previous work such as @smithEffectWordPredictability2013 has found positive relationships between RT and the surprisal of earlier words for SPR, so we include predictors from the current and the 3 prior words.
+The relationship between surprisals and RT is shown in Figure \@ref(fig:spr-gam); note that the y-axis range is much narrower than for Maze.
+Both current and previous word surprisals have a roughly linear positive relationship to RT.
+The surprisal of the word two back also has an influence in some models.
+
+Comparing Maze and SPR, we see that both show a linear relationship, but Maze has much larger effects of surprisal on the current word.
+
+## Does A-maze have less spillover?
+
+One of the main claimed advantages of the Maze task is that it has better localization and less spillover than SPR.
+We examined how much spillover A-maze and SPR each had by fitting linear models with predictors from current and previous words.
+Large effects from previous words are evidence for spillover; effects of the current word that dwarf any lagged effects is evidence for localization.
+
+We modeled reading time as a function of surprisal, frequency, and length as well as surprisal x length and frequency x length interactions.
+For all of these, we included the predictors for the current and previous word, and we centered, but did not rescale, all predictors.
+(See Methods for more details on these predictors and model fit process.) As with the GAM models, we used surprisal calculations from 4 different language models for robustness.
+
+(ref:coeffs) Point estimates and 95% credible intervals for coefficients predicted by fitted Bayesian regression models predicting A-maze RT.
+Units are in ms.
+Surprisal is per bit, length per character, and frequency per $log_2$ occurrence per billion words.
+
+```{r coeffs-maze, out.height="35%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:coeffs)"}
+
+test_plot <- read_rds(here("Analysis/models/maze_spr_for_graph.rds"))
+
+# test_plot %>% filter(time %in% c("Current", "Previous")) %>% ggplot(aes(y=Term, x=Estimate, group=model, color=model, shape=model))+
+#   ggstance::geom_pointrangeh(aes(xmin = Lower, xmax = Upper), position=position_dodge(width=.6))+
+#   facet_grid(time~type)+
+#   geom_vline(xintercept=0)+
+#   scale_color_brewer(palette="Dark2")+
+#   theme(legend.position = "right")+
+#   labs(x="Coefficient estimate", y="", color="Model", shape="Model")
+
+test_plot %>% filter(type=="Maze") %>% ggplot(aes(y=Term, x=Estimate, group=model, color=model, shape=model))+
+  ggstance::geom_pointrangeh(aes(xmin = Lower, xmax = Upper), position=position_dodge2(width=.6, reverse=T))+
+  facet_grid(.~time)+
+  geom_vline(xintercept=0)+
+ggthemes::scale_color_solarized(accent="violet")+
+  theme(legend.position = "right")+
+  labs(x="Coefficient estimates for Maze data", y="", color="Model", shape="Model")
+
+```
+
+The Maze linear model effects are shown in Figure \@ref(fig:coeffs-maze) (See also Appendix B for a table of effects).
+Across all models, there were consistent large effects of length and surprisal at the current word, but minimal effects of frequency.
+The lack of frequency effects is unexpected, but consistent with @shainLargescaleStudyEffects2019.
+There was a small interaction between surprisal and length at the current word.
+
+Crucially, the effects of previous word predictors are close to zero, and much smaller than the effects of surprisal and length of the current word, an indication that spillover is limited and effects are strongly localized.
+
+(ref:coeffs-spr) Point estimates and 95% credible intervals for coefficients predicted by fitted Bayesian regression models predicting SPR RT.
+Units are in ms.
+Surprisal is per bit, length per character, and frequency per $log_2$ occurrence per billion words.
+
+```{r coeffs-spr, out.height="35%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:coeffs-spr)"}
+test_plot %>% filter(type=="SPR") %>% mutate(time=factor(time, levels=c("Current","Previous","2Previous","3Previous")))%>% ggplot(aes(y=Term, x=Estimate, group=model, color=model, shape=model))+
+  ggstance::geom_pointrangeh(aes(xmin = Lower, xmax = Upper), position=position_dodge2(width=.6, reverse=T))+
+  facet_grid(.~time)+
+  geom_vline(xintercept=0)+
+ggthemes::scale_color_solarized(accent="violet")+
+  theme(legend.position = "right")+
+  labs(x="Coefficient estimate for SPR data", y="", color="Model", shape="Model")
+
+```
+
+We ran similar models for SPR, although to account for known spillover effects, we consider predictors from the current and 3 previous words.
+Due to issues fitting models, the details of the models differed (see Methods).
+The SPR coefficients are shown in Figure \@ref(fig:coeffs-spr) (see also Appendix B for a table of coefficients).
+Surprisal, length, and frequency effects are all evident for the current word and surprisal and frequency show effects from the previous word as well.
+Unlike for Maze, with SPR there is not a clear diminishing of the size of the effects as one goes from current word to prior word predictors.
+
+Whereas Maze showed surprisal effects in the 10 to 25 ms/bit range and length effects 15 to 20 ms/character range, SPR effects are about 1-2 ms per bit or character.
+This difference in effect size is disproportionate to the overall speed of the methods; the predicted intercept for the Maze task was roughly 880 ms and for SPR was roughly 360 ms.
+Thus Maze is is 2-3 times as slow as SPR but has roughly 10 times larger effects.
+
+## Which language model fits best?
+
+Our last analysis question is whether some of the language models fit the human RT data better than others.
+We assessed each model's fit to A-maze data using log likelihood and R-squared.
+Then we did a nested model comparison, looking at whether a model with multiple surprisal predictors (ex, GRNN and GPT-2) had a better fit than a model with only one (ex GRNN alone).
+
+As shown in Table \@ref(tab:maze-compare), GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complementary information over GPT-2.
+TXL provides a lot over 5-gram, and 5-gram provides little over any model.
+The single-model measures of log likelihood confirm this hierarchy, as GPT-2 is better than GRNN is better than TXL is better than 5-gram.
+
+```{r maze-compare}
+
+read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data.
+Each row shows the additional predictive value gained from adding that model to another model.
+F values and p values from ANOVA tests between 1-surprisal-source and 2-source models are reported.
+We also report log likelihoods of models with only one surprisal source and the r-squared correlation between the model's predictions and the data.")
+
+```
+
+We followed the same process for the SPR data with results shown in Table \@ref(tab:spr-compare).
+For SPR, GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN.
+In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small.
+The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020, and contrasts with the Maze results, where the 5-gram model had the worst fit and did not provide additional predictive value over the other models.
+
+```{r}
+maze <- read_rds(here("Analysis/maze_model_compare.rds"))
+spr <- read_rds(here("Analysis/spr_model_compare.rds"))
+```
+
+As an overall measure of fit to data, we calculate multiple R-squared for the single surprisal source models for both A-maze and SPR.
+The models predict A-maze better than SPR with R-squared values for A-maze ranging from `r maze %>% filter(Model=="5-gram") %>% pull(r_squared)` for the 5-gram model to `r maze %>% filter(Model=="GPT-2") %>% pull(r_squared)` for GPT-2.
+For SPR, the R-squared values range from from `r spr %>% filter(Model=="5-gram") %>% pull(r_squared)` to `r spr %>% filter(Model=="GPT-2") %>% pull(r_squared)`.
+This pattern suggests that the effect size differences are not due merely to the larger overall reading time for A-maze, but that instead A-maze is more sensitive to surprisal and length effects.
+
+```{r spr-compare}
+
+read_rds(here("Analysis/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on SPR data.
+Each row shows the additional predictive value gained from adding that model to another model.
+F values and p values from ANOVA tests between 1-surprisal-source and 2-source models are reported.
+We also report log likelihoods of models with only one surprisal source and the r-squared correlation between the model's predictions and the data.")
+
+```
 
@@
 
-We introduce error-correction Maze, a tweak on the presentation of Maze materials that makes Maze feasible for multi-sentence passages.
-We used A-maze distractors and this error-correction presentation to gather data on participants reading the Natural Stories in the Maze.
-We found that participants could read and comprehend the 1000 word stories, despite the slowness and added overhead of reading in the Maze task.
-This expands the domain of materials used with Maze beyond targeted single-sentence items to longer, naturalistic texts with sentence-to-sentence coherency.
+We introduced error-correction Maze, a tweak on the presentation of Maze materials that makes Maze feasible for multi-sentence passages.
+We then used A-maze distractors and the error-correction Maze presentation to gather data on participants reading stories from the Natural Stories corpus in the Maze.
+As laid out in the Introduction, this current study addressed five main questions.
 
-We took advantage of the pre-existing SPR corpus on Natural Stories to compare the RT profiles between Maze and SPR.
-While these are not completely comparably as the subject populations differed, we are able to get a sense of the relationship between SPR and Maze results.
-Maze and SPR pick up on similar features in words, as shown by the correlations between Maze and SPR, and the fact that surprisal and length are predictive of RTs in both.
-However, the effect sizes of RT and surprisal are much larger is Maze, and the effects are more strongly localized on a single word.
+First, we found that participants could read and comprehend the 1000 word stories, despite the slowness and added overhead of reading in the Maze task.
+This result expands the domain of materials usable with Maze beyond targeted single-sentence items to longer, naturalistic texts with sentence-to-sentence coherency.
 
-Some language models are better or worse on predicting human data; however, we find preliminary evidence that how predictive a language model is of human data may depend on what task was used to collect the RTs, with the 5-gram model relatively better at predicting SPR data.
-Further comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex.
+Second, we took advantage of the pre-existing SPR corpus on Natural Stories to compare the RT profiles between Maze and SPR.
+Maze and SPR pick up on similar features in words, as shown by the high correlations between Maze and SPR RTs on the sentence level.
+
+Third, we addressed whether the A-maze RT for a word showed a linear relationship with that word's surprisal.
+We found that A-maze RTs are linearly related to surprisal, matching the functional profile found with other incremental processing methods.
+
+Fourth, we compared the spillover profiles between Maze and SPR.
+For Maze, we found large effects of the current words surprisal and length, which dwarfed any spillover effects from previous word predictors.
+In contrast, for SPR, we found effects of roughly equal sizes from the current and previous words.
+Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus and thus generating less noisy data.
+
+Lastly, we examined how different language models fare at predicting human RT data.
+We found that overall, the models were more predictive of the A-maze data than SPR data; however, the hierarchy of the model's predictive performance also differed between the A-maze and SPR datasets.
+This difference suggests that how predictive a language model is of RTs may depend on the task.
+Further comparisons between different processing methods on the same materials could be useful for identifying how task demands influence language processing [ex.
 @bartekSearchOnlineLocality2011].
 
-We find good localization on Maze, but it's worth noting that it's not absolute.
-Some of the models show small but statistically reliable relationships with surprisal from a past word.
-On the whole, however, these results support the idea that Maze forces language processing to be close to word-by-word, and that it's a relatively safe assumption that the RT of a word primarily reflects its own properties and not those of earlier words.
+Overall, A-maze has good localization, although some  models showed small but statistically reliable effects of the past word.
+On the whole, however, our results support the idea that Maze forces language processing to be close to word-by-word, and thus the Maze task can be used under the assumption that the RT of a word primarily reflects its own properties and not those of earlier words.
 
-This extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty.
+## Limitations
 
-## Remarks about error-correction maze
+While we expect these patterns of results reflect features of the A-maze task, the effects could be moderated by quirks of the materials or the participant population.
+We excluded a large number of participants for having low accuracy on the task and appearing to guess randomly.
+We compared RTs collected on the A-maze task to SPR RTs previously collected on the same corpus, but we did not randomly assign participants to SPR and Maze conditions.
+This study suggests that A-maze is a localized and widely-usable method, but only broader applications can confirm these findings.
 
-With that in mind, we have a few remarks on the benefits of error-correction Maze and some avenues for further work on this method.
-Error-correction Maze makes for a smoother process.
-It dampens the effect of rare poor A-maze distractors; a distractor that fits the context might still cause participants to make an unavoidable mistake, but the mistake is less disruptive, as they still see the rest of the sentence.
-An open question for researchers is whether data from after a participant makes a mistake is useable, that is, does it show the same profile as pre-errors words, or are there traces from recovering from the mistake?
-Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is a hard-to-assess question of potential interest.
+## Future directions
 
-Even if post-mistake data is not included in analyses, having this complete information can distinguish between errors due to inattentive participants and errors from specific bad distractors.
-Researchers can calculate a per-word error rate for each participant; high per-word error rates are consistent with guessing, low error rates with errors clustered early the sentence are consistent with poor distractors.
-This per-word error rate metric measures participants task accuracy and provides a convenient and clear-cut way of controlling data quality after the fact.
+Error-correction Maze  reduces perverse incentives from the desire to complete the task quickly compared to traditional Maze, but clicking randomly is still more efficient than doing the task.
+In discussing this work, we received the suggestion that one way to further disincentivize random clicking would be to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms) before correcting their mistake.
+This delay would make randomly hitting buttons slower than doing the task as intended.
 
-Error-correction Maze also starts to reduce perverse incentives from the desire to complete the task quickly.
-With traditional Maze, clicking randomly will likely lead to a mistake, which will cause a participant to skip ahead to the next sentence.
-With error-correction Maze, randomly jamming buttons takes more effort, but is still faster than doing the task.
-In discussing this work, we received the suggestion that one way to disincentivize random clicking is to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms or 1 sec) before being able to correct their mistake.
-This seems like a promising improvement that could be worth implementing and testing.
+Error-correction Maze records RTs for words after a participant makes a mistake in the sentence.
+In our analyses, we excluded this post-error data, but we believe it is an open question whether data from after a participant makes a mistake is usable.
+That is, does it show the same profile as RTs from pre-error words, or are there traces from recovering from the mistake, and if so, how long do these effects take to fade?
+Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is hard-to-assess; if it can be used, it would make the Maze task more data efficient.
 
-Between error-correcting Maze and other innovations on the Maze paradigm such as A-maze and I-maze, the Maze task is versatile and can be used or adapted for a wide range of materials and questions of interest.
-We encourage researchers to use Maze as an incremental processing method, alone or in comparison with other methods.
+The Maze task is versatile and can be used or adapted for a wide range of materials and questions of interest.
+The extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty.
+We encourage researchers to use Maze as an incremental processing method, alone or in comparison with other methods, and we suggest that the error-correction mode be the default choice for presenting Maze materials.
 
@@
 
-# Appendix 
-The beginning of one of the stories. This is the first 200 words of a 1000 word story.
+# Appendix A
+The beginning of one of the stories. This excerpt is the first 200 words of a 1000 word story.
 
@@
 From which country did tulips come to Europe? Turkey, Egypt
+
+# Appendix B
+
+```{r pre-error}
+summ <- read_rds(here("Analysis/models/brms_maze_summ.rds")) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
+  mutate(Term=factor(Term, 
+                        levels=c("Intercept", 
+                                 "surprisal",
+                                 "len",
+                                 "freq",
+                                 "surprisal:len",
+                                 "len:freq",
+                                 "prev_surp",
+                                 "prev_len",
+                                 "prev_freq",
+                                 "prev_surp:prev_len",
+                                 "prev_len:prev_freq")
+                                 )) %>% 
+  arrange(Term) %>% 
+  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency",
+                  "Surp x Length", "Freq x Length", "Past Surprisal",
+                  "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length"))
+
+knitr::kable(summ, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models.
+All terms were centered, but not rescaled.
+Units are in ms.
+Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.
+Interval is 2.5th quantile to 97.5th quantile of model draws.")%>%   kable_styling(font_size = 10)
+```
+
+```{r spr-table}
+
+summ_spr <- read_rds(here("Analysis/models/lmer_spr_summ.rds")) %>% 
+  pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
+  mutate(Term=factor(Term, 
+                        levels=c("(Intercept)", 
+                                 "surp_center",
+                                 "length_center",
+                                 "freq_center",
+                                 "surp_center:length_center",
+                                 "length_center:freq_center",
+                                 "past_surp_center",
+                                 "past_length_center",
+                                 "past_freq_center",
+                                 "past_surp_center:past_length_center",
+                                 "past_length_center:past_freq_center",
+                                 "past2_surp_center",
+                                 "past2_length_center",
+                                 "past2_freq_center",
+                                 "past2_surp_center:past2_length_center",
+                                 "past2_length_center:past2_freq_center",
+                                 "past3_surp_center",
+                                 "past3_length_center",
+                                 "past3_freq_center",
+                                 "past3_surp_center:past3_length_center",
+                                 "past3_length_center:past3_freq_center"
+                                 )
+                                 )) %>% 
+  arrange(Term) %>% 
+  mutate(Term=c("Intercept", "Surprisal", "Length", "Frequency","Surp x Length", "Freq x Length",
+                "Past Surprisal", "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length",
+                "2Past Surprisal", "2Past Length", "2Past Freq", "2Past Surp x Length", "2Past Freq x Length",
+                "3Past Surprisal", "3Past Length", "3Past Freq", "3Past Surp x Length", "3Past Freq x Length"))
+
+knitr::kable(summ_spr, format="latex", position="ht",caption="Predictions from fitted regression models for SPR data.
+All terms were centered, but not rescaled.
+Units are in ms.
+Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.
+Uncertainty interval is +/- 1.97 standard error.") %>%   kable_styling(font_size = 10)
+
+```

### 2022-08-10 Analysis/models.Rmd
Roger touched

@@
 gam_gpt_3 <- gam(formula3_interact, data=gpt_mean_data, method="REML")
-print(plot(getViz(gam_gpt_1x)))
-print(plot(getViz(gam_gpt_3)))
 
@@
 gam_gpt_no_interact <- gam(formula_no_interact, data=gpt_mean_data, method="REML")
+
+#print(plot(getViz(gam_gpt_1a)))
+print(plot(getViz(gam_gpt_no_interact)))
+print(plot(getViz(gam_gpt_2)))
+summary(gam_gpt_3)
 

### 2022-08-10 Analysis/models.Rmd
Added some more formulae including test for non-linearity in surprisal

@@
 ```{r maze-formulae, echo=T}
+
+#no hierarchical, has freq x len interaction
+formula_no_interact_linear_surprisal <- rt ~ surprisal +
+                     ti(freq, bs="cr") + 
+                     ti(len, bs="cr")+
+                     ti(freq,len, bs="cr")+
+                     prev_surp +
+                     ti(prev_freq, bs="cr")+
+                     ti(prev_len, bs="cr")+
+                     ti(prev_freq, prev_len, bs="cr")
+
+# no hierarchical, NO freq x len interaction
+formula_no_interact_plus_linear_surprisal <- rt ~ surprisal +
+                     s(surprisal, bs="cr", k=20)+
+                     ti(freq, bs="cr") + 
+                     ti(len, bs="cr")+
+                     prev_surp +
+                     s(prev_surp, bs="cr", k=20)+
+                     ti(prev_freq, bs="cr")+
+                     ti(prev_len, bs="cr")
+
 #no hierarchical, has freq x len interaction
@@
 
+gam_gpt_3_hf <- gam(formula3_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
+gam_gpt_3_lf <- gam(formula3_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")
+
+gam_gpt_2_hf <- gam(formula2_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
+gam_gpt_2_lf <- gam(formula2_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")
+
 gam_ngram_2 <- gam(formula_no_interact, data=ngram_mean_data, method="REML")
@@
 gam_gpt_no_interact <- gam(formula_no_interact, data=gpt_mean_data, method="REML")
+gam_gpt_no_interact_plus_linear_surprisal <- gam(formula_no_interact_plus_linear_surprisal, data=gpt_mean_data, method="REML")
 
@@
 
+summary(getViz(gam_gpt_3_hf))
+print(plot(getViz(gam_gpt_3_hf)))
+print(plot(getViz(gam_gpt_2_hf)))
+
+summary(getViz(gam_gpt_3_lf))
+print(plot(getViz(gam_gpt_3_lf)))
+
+summary(gam_gpt_no_interact_plus_linear_surprisal)
 ```

### 2022-08-11 Papers/Paper/manuscript.Rmd
Roger starting to edit ms

@@
 
-Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing.
+Two chief results of human language processing research are that comprehension is highly incremental and that comprehension difficulty is differential and localized.
+Incrementality in comprehension means that our minds do not wait for large stretches of linguistic input to accrue; rather, we eagerly analyze each moment of input and rapidly integrate it into context [@marslen-wilson:1975; @tanenhaus-etal:1995].
+Differential and localized processing difficulty means that different inputs in context present different processing demands during comprehension, and that by and large these demands are met relatively quickly by the mind once they are presented [@levy:2008].
+XXX Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing.
 Maze is a potentially promising method, and it's feasibility has recently increased due to the introduction of A-Maze [@boyceMazeMadeEasy2020].

### 2022-08-12 Papers/Paper/manuscript.Rmd
Roger revised intro

--- 2022-03-01 Papers/Paper/manuscript.Rmd
@@
 
-Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing.
+Two chief results of human language processing research are that comprehension is highly incremental and that comprehension difficulty is differential and localized.
+Incrementality in comprehension means that our minds do not wait for large stretches of linguistic input to accrue; rather, we eagerly analyze each moment of input and rapidly integrate it into context [@marslen-wilson:1975].
+Differential and localized processing difficulty means that different inputs in context present different processing demands during comprehension [@levy:2008].
+Due to incrementality these differential processing demands are, by and large, met relatively quickly by the mind once they are presented, and they can be measured in both brain [@kutas-hillyard:1980; @osterhout-holcomb:1992jml] and behavioral [@raynerEyeMovementsReading1998; @mitchell:2004online-methods] responses.
+These measurements often have low signal-to-noise ratio, and many methods require bringing participants into the lab and often require cumbersome equipment.
+However, they can provide considerable insight into how language processing unfolds in real time.
+Developing more sensitive methods that can easily be used with remote participants is thus of considerable interest.
+
+Word-by-word reading or response times are among the most widely used behavioral measurements in language comprehension and give relatively direct insight into processing difficulty.
+The Maze task [@freedman85; @forsterMazeTaskMeasuring2009], which involves collecting participants' response times in a repeated two-alternative forced-choice between a word that fits the preceding linguistic context and a distractor that doesn't, has recently been proposed as a high-sensitivity method that can easily be used remotely.
+@boyceMazeMadeEasy2020 introduced several implementational innovations that made it easier for researchers to use Maze, and showed that for several controlled syntactic processing contrasts [@witzelComparisonsOnlineReading2012a] Maze offers better statistical power than self-paced reading, the other word-by-word response time method easy to use remotely.
+Maze has since had rapid uptake in the language processing community [@chacon2021limits; @ungerer2021using; @orth2022processing] (other cites).
+
+However, there is increasing interest in collecting data during comprehension of more naturalistic materials such as stories and news articles [@demberg-keller:2008; @luke-christianson:2016-limits; @futrellNaturalStoriesCorpus2020], which offer potentially improved ecological validity over repeated presentation of isolated sentences out of context.
+These more materials require maintaining and integrating discourse dependencies and other type of information over longer stretches of time and linguistic material.
+Previous work leaves unclear whether the Maze task would be feasible for this purpose: the increased task demands might interfere with the demands presented by these more naturalistic materials, and vice versa.
+In this paper we report a new modification of the Maze task and show that it makes reading of extended, naturalistic texts feasible.
+We also analyze the resulting reaction time profiles and show that they provide strong signal regarding the probabilistic relationship between a word and the context in which it appears, and that the systematic relationship between word surprisal and response time observed in other reading paradigms [@smithEffectWordPredictability2013] also arises in the Maze task.
+<!-- Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing.
 Maze is a potentially promising method, and it's feasibility has recently increased due to the introduction of A-Maze [@boyceMazeMadeEasy2020].
-We tested A-maze on a naturalistic corpora to learn whether participants could understand stories read in the demanding Maze task and whether the reaction time profiles from the Maze task matched those from other methods.
-To provide context, we first lay out the role of incremental processing methods in theory testing, describe a few common methods, and review some key influences on reading time.
+-->
+In the remainder of the Introduction, we lay out the role of incremental processing methods in theory testing, describe a few common methods, and review some key influences on reading time.
+We then proceed to present our modified "error-correction Maze" paradigm, our experiment, and the results of our analyses of the resulting data.
 

### 2022-08-13 Papers/Paper/manuscript.Rmd
add appendix graph and table

@@
 ```
+
+# Appendix C
+
+```{r gam-linearity-table}
+
+gam_linear <- read_rds(here("Analysis/models/gams_linearity_test.rds")) %>% 
+  mutate(Term=factor(Term, 
+                        levels=c("s(surprisal)",
+                                 "ti(freq)",
+                                 "ti(len)",
+                                 "s(prev_surp)",
+                                 "ti(prev_freq)",
+                                 "ti(prev_len)",
+                                 "(Intercept)",
+                                 "surprisal",
+                                 "prev_surp"
+                                 )
+                                 )) %>% 
+  arrange(Term) %>% filter(Term %in% c("s(surprisal)", "s(prev_surp)", "surprisal","prev_surp")) %>% 
+  mutate(Term=c("Spline Surprisal", "Spline Past Surprisal", "Linear Surprisal", "Linear Past Surprisal"))
+
+knitr::kable(gam_linear, format="latex", position="ht",caption="Comparison of significances for linear and spline terms of suprisals from a GAM.
+We fit GAM models with current and past word surprisal as parametric terms, current and past word surprisal and spline terms, and current and past frequence and length as tensors to predict reading time.
+Here we show the estimated pvalues for the linear and spline surprisal terms at current and past words.
+The spline terms account for any non-linear surprisal effects.") %>%   kable_styling(font_size = 10)
+
+```
+
+# Appendix D
+(ref:gam-grid) CAPTION GOES HERE
+
+```{r gam-grid, out.height="100%",  fig.pos="ht", fig.cap="(ref:gam-grid)"}
+
+knitr::include_graphics(here("Analysis/models/gam_grid.png"))
+
+```

### 2022-08-15 Papers/Paper/manuscript.Rmd
Roger updated manuscript, finished pass over intro

--- 2022-08-12 Papers/Paper/manuscript.Rmd
@@
 
-However, there is increasing interest in collecting data during comprehension of more naturalistic materials such as stories and news articles [@demberg-keller:2008; @luke-christianson:2016-limits; @futrellNaturalStoriesCorpus2020], which offer potentially improved ecological validity over repeated presentation of isolated sentences out of context.
+However, there is increasing interest in collecting data during comprehension of more naturalistic materials such as stories and news articles [@demberg-keller:2008; @luke-christianson:2016-limits; @futrellNaturalStoriesCorpus2020], which offer potentially improved ecological validity and larger data scale in comparison with repeated presentation of isolated sentences out of context.
 These more materials require maintaining and integrating discourse dependencies and other type of information over longer stretches of time and linguistic material.
@@
 In this paper we report a new modification of the Maze task and show that it makes reading of extended, naturalistic texts feasible.
-We also analyze the resulting reaction time profiles and show that they provide strong signal regarding the probabilistic relationship between a word and the context in which it appears, and that the systematic relationship between word surprisal and response time observed in other reading paradigms [@smithEffectWordPredictability2013] also arises in the Maze task.
+We also analyze the resulting reaction time profiles and show that they provide strong signal regarding the probabilistic relationship between a word and the context in which it appears, and that the systematic linear relationship between word surprisal and response time observed in other reading paradigms [@smithEffectWordPredictability2013] also arises in the Maze task.
 <!-- Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing.
@@
 -->
-In the remainder of the Introduction, we lay out the role of incremental processing methods in theory testing, describe a few common methods, and review some key influences on reading time.
+In the remainder of the Introduction, we lay out the role of RT-based methods in theory testing, describe a few common methods, and review some key influences on reading time.
 We then proceed to present our modified "error-correction Maze" paradigm, our experiment, and the results of our analyses of the resulting data.
 
-## Why measure reading time?
+## Why measure RTs?
 
-Occasionally we read an unexpected word, cannot make sense of it and have to read the sentence again.
+A major feature of human language is that not all sentences or utterances are equally easy to successfully comprehend.
+Sometimes this is mostly or entirely due to the linguistic structure of the sentence: for example, *The rat that the cat that the dog chased killed ate the cheese* is more difficult than *The rat that was killed by the cat that was chased by the dog ate the cheese* even though the meaning of the two sentences is (near-)identical.
+Sometimes the source of difficulty can be a mismatch between expectations set up by the context and the word choice in an utterance: for example, the question *Is the cup red?* may be confusing in a context containing more than one cup.
+Psycholinguistic theories may differ in their ability to predict what is easy and what is hard.
+One of the most powerful methods for studying these differential difficulty effects is to turn over control of presentation of the linguistic material to comprehender, and to measure what she takes time on.
+For this purpose, taking measurements from experimental participants during reading, a widespread, highly practiced skill in diverse populations around the world, is of unparalleled value.
+
+To a first approximation, everyday reading (when the reader's goal is to understand a text's overall content) is *progressive*: we read documents, paragraphs, and sentences from beginning to end.
+The reader encounters each word with the benefit of the preceding linguistic context.
+Incrementality in reading involves successively processing each word encountered and integrating it into the context.
+For a skilled reader experienced with the type of text being read, most words are easy enough that the subjective experience of reading the text is of smooth, continuously unfolding understanding as we construct a mental model of what is being described.
+But occasionally a word may be sufficiently surprising or otherwise difficult to reconcile with the context that it disrupts comprehension to the level of conscious awareness: in the sentence *I take my coffee with cream and chamomile*, for example, the last word is likely to do so.
+Behaviorally, this disruption typically manifests as a slowdown or longer *reading time* (RT) on the word itself, on the immediately following words, or in other forms such as regressive eye movements back to earlier parts of the text to check the context.
+
+In fact, RTs and other measures that capture processing disruption vary substantially with the difficulty of words in their context below the level of conscious awareness as well, with millisecond scale differences in reading time between words.
+<!-- Occasionally we read an unexpected word, cannot make sense of it and have to read the sentence again.
 The words that cause these stumbles are the outliers on a distribution of words varying in how easy they are to process in context.
@@
 Fortunately for reading and unfortunately for studying language, this process of adjustment is very quick, with millisecond scale differences in reading time between words.
+-->
+That is, the differential difficulty or processing load posed by various parts of a text is to a considerable extent *localizable* to specific words in their context.
+For this reason, RTs have proven a highly valuable measure for testing the predictions of psycholinguistic theory, ranging from theories of character recognition, memory retrieval, parsing, and beyond.
 
-If we assume that processing time reflects something about the processing, be it character recognition, memory retrieval, parsing, or something else, then we can learn something about how the mind processes language from measuring how long the processing takes in different circumstances.
-With this assumption, psycholinguists can test theories that predict the relative difficulty of two constructions by measuring the relative processing time of these words in context.
-
-For instance, theories about what makes object relative clauses harder to understand than subject relative clauses make different predictions about which words are the loci of the overall difficulty and slower reading times associated with object relatives [@grodnerConsequencesSerialNature2005; @staubEyeMovementsProcessing2010; @traxlerProcessingSubjectObject2002].
-Measures of reading time can also inform theories about the time course of processing (i.e.
+For instance, competing theories about why certain types of object-extracted relative clauses, like *the lawyer that the banker irritated*, are harder to understand than the corresponding subject-extracted relative clauses, like *the lawyer that irritated the banker*, make different predictions about which words are the loci of the overall difficulty and slower RTs associated with object relatives [@grodnerConsequencesSerialNature2005; @staubEyeMovementsProcessing2010; @traxlerProcessingSubjectObject2002].
+RT measures can potentially also inform theories about the time course of processing (i.e.
 which steps are parallel versus serial, @bartekSearchOnlineLocality2011) or the functional form of relationships between word characteristics and processing time [@smithEffectWordPredictability2013].
@@
 Determining that object relatives are overall slower that subject relatives is easy.
-Even an imprecise measure of reading time will determine that the same set of words in a different order took longer to read at a sentence level.
+Even an imprecise RT measure will determine that the same set of words in a different order took longer to read at a sentence level.
 However, many language processing theories make specific (and contrasting) theories about which words in a sentence are harder to process.
-To adjudicate these theories, we want methods that are *localized*, so it is easy to determine which word is responsible for an observed slow-down in reading time.
-Ideally, a longer reading time on a word would be an indication of that word's increased difficulty, and not the lingering signal of a prior word's increased difficulty.
+To adjudicate among these theories, we want methods that are *well-localized*, so it is easy to determine which word is responsible for an observed RT slow-down.
+Ideally, longer RT on a word would be an indication of that word's increased difficulty, and not the lingering signal of a prior word's increased difficulty.
 When the signal isn't localized, advanced analysis techniques may be required to disentangle the slow-downs [@shainDeconvolutionalTimeSeries2018].
@@
 
-Behavioral methods that measure how long it takes to read each word in a sentence are referred to as incremental processing methods, and their dependent measures are called reading times or reaction times, both abbreviated RT.
-The two most commonly used methods are eye-tracking and self-paced reading, but both of these suffer from a lack of localization.
+<!-- #Behavioral methods that measure how long it takes to read each word in a sentence are referred to as incremental processing methods, and their dependent measures are called reading times or reaction times, both abbreviated RT.
+-->
+The two most commonly used behavioral methods for studying incremental language processing during reading are tracking eye movements and self-paced reading.
+While both of these have proven powerful and highly flexible, they both have important limitations as well.
 
-In eye-tracking, participants read a text on a screen naturally, while their eye-movements are recorded [@raynerEyeMovementsReading1998].
-The downside of natural reading is that people often skip short words and look ahead or look back as they read, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982].
+In eye-tracking, participants read a text on a screen naturally, while their saccadic eye movements are recorded on a computer-connected camera that is calibrated so that the researcher can reconstruct with high precision where the participant's gaze falls on the screen at all times [@raynerEyeMovementsReading1998].
+From these eye movements can be reconstructed various position-specific reading time measures such as *gaze duration* (the total amount of time the eyes spend on a word the first time it is fixated, or zero if the eye skipped the word the first time it was approached from the left) and *total viewing time* (the total amount of time that the word is fixated).
+Eye tracking data collected with state-of-the-art high-precision recording equipment offers relatively good signal-to-noise ratio, but the difficulty presented by a word can still *spill over* into reading measures on subsequent words, a dynamic that can make it hard to isolate the source of an effect of potential theoretical interest [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982].
+Additionally, the equipment is expensive and data collection is laborious and must occur in-lab.
 
-Self-paced reading (SPR) is a more controlled process where a participant sees one word on screen at a time and presses a button to see the next word instead.
-Even in SPR, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006].
-
-Both these methods are prone to spillover.
-The time to plan motor movements like saccades or button presses is substantial compared to the time it takes to recognize a word, these movements are often initiated before a word is fully processed.
-By the time any late processing difficulties occur, the reader may already be several words along.
+Self-paced reading (SPR; @mitchell:1984) is a somewhat less natural paradigm in which the participant manually control the visual presentation of the text by pressing a button.
+In its generally preferred variant, moving-window self-paced reading, words are revealed one at a time or one group at a time, wihth every press of the button masks the currently presented word (group) and simultaneously reveals the next.
+The time spent between button presses is the unique RT measure for that word (group).
+Self-paced reading requires no special equipment and can be delivered remotely, but the measurements are noisier and even more prone to spillover [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006; @smithEffectWordPredictability2013].
 
@@
 The Maze task is an alternative method that is designed to increase localization at the expense of naturalness [@freedman85; @forsterMazeTaskMeasuring2009].
-In the Maze task, participants see two words at a time: a correct word that continues the sentence, and a distractor which does not.
-Participants must choose the correct word, and their time to selection is treated as the reaction time (RT).
-@forsterMazeTaskMeasuring2009 introduced two versions of the Maze task: lexical L-maze where the distractors are nonce words and grammatical G-maze where the distractors are real words that don't fit with the context of the sentence.
-Theoretically, participants must fully integrate each word into the sentence in order to confidently select it, which may require mentally reparsing previous material in order to allow the integration and selection of a disambiguating word.
+In the Maze task, participants must repeatedly choose between two simultaneously presented options: a correct word that continues the sentence, and a distractor string which does not.
+Participants must choose the correct word, and their time to selection is treated as the reaction time, or RT.
+(We intentionally overload the abbreviation "RT" and use it for Maze reaction times as well as reading times from eye tracking and SPR, because the desirable properties of reading times turn out to hold for Maze reaction times as well.) @forsterMazeTaskMeasuring2009 introduced two versions of the Maze task: lexical "L"-maze where the distractors are non-word strings, and grammatical "G"-maze where the distractors are real words that don't fit with the context of the sentence.
+In theory, participants must fully integrate each word into the sentence in order to confidently select it, which may require mentally reparsing previous material in order to allow the integration and selection of a disambiguating word.
 @forsterMazeTaskMeasuring2009 call this need for full integration "forced incremental processing" to distinguish from other incremental processing methods where words can be passively read before later committing to a parse.
@@
 Another recent variant of the Maze task is interpolated I-maze, which uses a mix of real word distractors (generated via the A-maze process) and nonce word distractors [@vaniUsingInterpolatedMaze2021; @wilcoxTargetedAssessmentIncremental2021].
-The presence of real word distractors encourages close attention to the sentential context, while nonce words can be used as distractors where the word in the sentence is itself ungrammatical or highly unexpected.
+The presence of real word distractors encourages close attention to the sentential context, while nonce words can be used as distractors where the word in the sentence is itself ungrammatical or highly unexpected, and/or it is important that the predictability of the distractor in the context is perfectly well-balanced (at zero) across all experimental conditions.
 
@@
 
-The functional form of the relationship between RTs from eye-tracking and SPR studies and the predictability of the words is linear in terms of surprisal [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016].
-Predictability reliable correlates with reading time in the range of surprisals found in natural-sounding texts, not just for words that are extremely expected or unexpected [@smithEffectWordPredictability2013].
+The functional form of the relationship between RTs from eye-tracking and SPR studies and the predictability of the words is linear in terms of surprisal [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016], even when two important context-invariant word features known to influence RTs, length and frequency, are controlled for.
+Predictability reliably correlates with reading time over a wide range of surprisals found in natural-sounding texts, not just for words that are extremely expected or unexpected [@smithEffectWordPredictability2013].
 If Maze RTs reflect the same processing as other methods, we expect to find a similar linear relationship with surprisal.
@@
 The Maze task has thus far primarily been used on constructed sentences focusing on targeted effects and not on the long naturalistic passages used to assess the relationship between RT and surprisal.
-We tested how A-maze performs on longer naturalistic corpora.
+We tested how A-maze performs on longer naturalistic corpora and compared it with self-paced reading (SPR), with the following main questions in mind:
 
-The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages of roughly 1000 words each which are designed to read fluently to native speakers.
+  1. Do participants engage with these longer passage successfully with the A-maze task?
+  2. Is A-maze as powerful and reliable a method as SPR for these longer passages?
+  3. What is the functional form between word surprisal and RT for the A-maze task?
+  4. Does A-maze have less spillover than SPR?
+  5. What types of context-driven expectations, as operationalized in competing computational language models, are deployed to determine RTs in A-maze and SPR?
+
+We used the Natural Stories corpus [@futrellNaturalStoriesCorpus2020], which consists of 10 passages of roughly 1000 words each which are designed to read fluently to native speakers.
 At the same time, the passages contain copious punctuation, quoted speech, proper nouns, and low frequency grammatical constructions.
@@
 We tweaked the A-maze task to accommodate these longer passages and then had participants read the passages in the Maze.
-We compared our A-maze results with SPR data collected on the Natural Stories corpus by @futrellNaturalStoriesCorpus2020.
-The A-maze data and SPR comparison data allow us to address the following questions:
-
-  1. Do participants engage successfully with the A-maze task?
-  2. Is A-maze as reliable as SPR?
-  3. Are the effects of surprisal on RT linear for the A-maze task?
-  4. Does A-maze have less spillover than SPR?
-  5. Which language model best fits the RT data from A-maze and SPR?
+We compare our A-maze results with SPR data collected on the Natural Stories corpus by @futrellNaturalStoriesCorpus2020.
 
@@
 We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
-After examining the participants performance on the task (see Results for details), we excluded data from participants with less than 80% accuracy, removing participants whose behavior was consistent with random guessing.
+After examining participants' performance on the task (see Results for details), we excluded data from participants with less than 80% accuracy, removing participants whose behavior was consistent with random guessing.
 After this exclusion, `r some %>% select(subject) %>% unique() %>% nrow()` participants were left.
@@
 
-## Is A-maze reliable?
+## How do A-maze and SPR compare in power and reliability?
 
@@
 SPR data is about as correlated with Maze as with another sample of SPR data which provides some evidence that Maze and SPR are measuring the same effects.
+The superior within-method split-half correlation we see for Maze relative to SPR, despite the smaller number of participants, suggests that it is the more powerful of the two methods (higher signal-to-noise ratio), consistent with the findings of @boyceMazeMadeEasy2020 for factorial experimental designs with isolated-sentence presentation.
 

### 2022-08-15 Analysis/models.Rmd

@@
 gam_gpt_1 <- gam(formula_interact, data=gpt_mean_data, method="REML")
-gam_gpt_1a <- gam(formula_interact, data=filter(gpt_mean_data,len<9), method="REML")
+#gam_gpt_1a <- gam(formula_interact, data=filter(gpt_mean_data,len<9), method="REML")
 gam_gpt_2 <- gam(formula2_interact, data=gpt_mean_data, method="REML")
@@
 
-gam_gpt_3_hf <- gam(formula3_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
-gam_gpt_3_lf <- gam(formula3_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")
-
-gam_gpt_2_hf <- gam(formula2_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
-gam_gpt_2_lf <- gam(formula2_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")
+#gam_gpt_3_hf <- gam(formula3_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
+#gam_gpt_3_lf <- gam(formula3_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")
+
+#gam_gpt_2_hf <- gam(formula2_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
+#gam_gpt_2_lf <- gam(formula2_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")
 
@@
 summary(gam_gpt_no_interact_plus_linear_surprisal)
+```
+## Extract linearity comparisons
+
+```{r}
+make_gam_linear_summary <- function(data){
+  model <- gam(formula_no_interact_plus_linear_surprisal, data=data, method="REML")
+  summ <- summary(model)
+  stable <- summ$s.table %>% as_tibble(rownames="Term") %>% rename(pvalue=`p-value`) %>% select(Term, pvalue)
+  ptable <- summ$p.table %>% as_tibble(rownames="Term") %>% rename(pvalue=`Pr(>|t|)`) %>% select(Term, pvalue)
+  return(union(stable,ptable))
+}
+
+gpt_linear <- make_gam_linear_summary(gpt_mean_data) %>% mutate(model="GPT-2")
+ngram_linear <- make_gam_linear_summary(ngram_mean_data) %>% mutate(model="5-gram")
+txl_linear <-  make_gam_linear_summary(txl_mean_data) %>% mutate(model="TXL")
+grnn_linear <-  make_gam_linear_summary(grnn_mean_data) %>% mutate(model="GRNN")
+
+all_linear <- gpt_linear %>% union(ngram_linear) %>% union(txl_linear) %>% union(grnn_linear) %>% mutate(pvalue=pvalue(pvalue,accuracy=.0001,add_p=T),model=factor(model, levels=c("5-gram", "GRNN", "TXL","GPT-2")))%>% arrange(model) %>%
+  pivot_wider(names_from=model, values_from=pvalue) %>% write_rds(here("Analysis/models/gams_linearity_test.rds"))
+```
+
+## Pretty interaction pics
+```{r fig.width=6, fig.height=8,}
+
+#gam_gpt_3 <- gam(formula3_interact, data=gpt_mean_data, method="REML")
+
+#gam_gpt_interact <- getViz(gam_gpt_3) %>% write_rds(here("Analysis/models/gam_gpt_interact.rds"))
+
+# can't figure out how to adjust the number of rows cols using the default, so we instead copy their internal functions off github and hack
+
+addDefaultLayers <- function( .l ){
+    .cl <- paste(.l$type, collapse = '')
+
+    if( grepl("Check", .cl) ){ # [A] Checking plots
+      if("qgam" %in% mcls){ # a.1 Quantile GAMS
+        .l <- switch(.cl, 
+                     "Check0DScalarNumeric" = .l + l_hist() + l_vline(),
+                     "Check0DScalarFactor" = .l + l_hist() + l_vline(),
+                     "Check0DScalarLogical" = .l + l_hist() + l_vline(),
+                     "Check0DVectorNumeric" = .l + l_hist(),
+                     "Check0DVectorFactor" = .l + l_hist(),
+                     "Check0DVectorLogical" = .l + l_hist(),
+                     "Check1DNumeric" =  .l + l_gridQCheck1D() + l_rug(), 
+                     "Check1DFactor" = .l + l_gridQCheck1D() + l_rug(),
+                     "Check1DLogical" = .l + l_gridQCheck1D() + l_rug(),
+                     "Check2DNumericNumeric" = .l + l_gridQCheck2D(),
+                     "Check2DFactorFactor" = .l + l_gridQCheck2D() + l_rug(),
+                     "Check2DFactorNumeric" = .l + l_gridQCheck2D() + l_rug(),
+                     .l
+        )
+      } else { # a.2 Standard GAMS
+        .l <- switch(.cl, 
+                     "Check0DScalarNumeric" = .l + l_hist() + l_vline(),
+                     "Check0DScalarFactor" = .l + l_hist() + l_vline(),
+                     "Check0DScalarLogical" = .l + l_hist() + l_vline(),
+                     "Check0DVectorNumeric" = .l + l_hist(),
+                     "Check0DVectorFactor" = .l + l_hist(),
+                     "Check0DVectorLogical" = .l + l_hist(),
+                     "Check1DNumeric" =  .l + l_dens2D("cond") + l_gridCheck1D(mean, showReps = FALSE), 
+                     "Check1DFactor" = .l + l_gridCheck1D(showReps = FALSE) + l_rug(),
+                     "Check1DLogical" = .l + l_gridCheck1D(showReps = FALSE) + l_rug(),
+                     "Check2DNumericNumeric" = .l + l_gridCheck2D(),
+                     "Check2DFactorFactor" = .l + l_gridCheck2D() + l_rug(),
+                     "Check2DFactorNumeric" = .l + l_gridCheck2D() + l_rug(),
+                     .l
+        )
+      }
+    } else { # [B] Smooth effect plots
+
+    .l <- switch(.cl, 
+                 "singleIndex1D" = .l + l_fitLine() + l_ciLine() + l_rug(), 
+                 "fs1D" = .l + l_fitLine() + theme(legend.position="none"),
+                 "1D" = .l + l_fitLine() + l_ciLine() + l_rug(),
+                 "2D" = .l + l_fitRaster() + l_fitContour(), 
+                 "MD" = .l + l_fitRaster() + l_fitContour(), 
+                 "MDslice" = .l + l_fitRaster() + l_fitContour(), 
+                 "sos0" = .l + l_fitRaster() + l_fitContour(), 
+                 "sos1" = .l + l_fitRaster() + l_fitContour(),
+                 "randomEffect" = .l + l_fitLine() + l_ciLine() + l_points(), 
+                 "MultiRandomEffect" = .l + l_points(), 
+                 "mrf" = .l + l_poly(), 
+                 "PtermNumeric" = .l + l_fitLine() + l_ciLine() + l_rug(),
+                 "PtermFactor" = .l + l_ciBar() + l_fitPoints() + l_rug(),
+                 "PtermLogical" = .l + l_ciBar() + l_fitPoints() + l_rug(), 
+                 "PtermMatrixNumeric" = .l + l_fitLine() + l_ciLine(),
+                 "Multi1D" = .l + l_fitLine(),
+                 "Multi2D" = .l + l_fitRaster() + l_fitContour(),
+                 "MultiPtermNumeric" = .l + l_ciBar() + l_fitPoints(),
+                 "MultiPtermFactor" = .l + l_ciBar() + l_fitPoints(), 
+                 "ALE1DNumeric" = .l + l_fitLine() + l_ciLine() + l_rug(),
+                 "ALE1DFactor" = .l + l_ciBar() + l_fitPoints() + l_rug(),
+                  .l
+                 )
+
+    }
+
+    return( .l )
+  }
+
+g <- read_rds(here("Analysis/models/gam_gpt_interact.rds")) %>% plot()
+g$plots <- lapply(g$plots,   function(.l){
+                         addDefaultLayers(.l)
+                       })
+
+blah <-  arrangeGrob(grobs=read_rds(here("Analysis/models/draft_grid.rds")), ncol=3, )
+
+ggsave(here("Analysis/models/gam_grid.png"), blah)
+
 ```

### 2022-08-15 Papers/Paper/manuscript.Rmd
Roger's changes & merge with Veronica's changes

@@
 Accuracy, or how often a participant chose the correct word over the distractor, reflects both the quality of the distractors and the focus and skill of the participant.
-We calculated the per-word accuracy rate for each participant and compared it against their average reaction time [^2].
-As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) made relatively few errors, with some reaching 99% accuracy.
+We calculated the per-word accuracy rate for each participant and compared it against their average reaction time.[^2] As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) made relatively few errors, with some reaching 99% accuracy.
 This high performance confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story.
@@
 This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.
-We believe the high level of random guessing is an artifact of the subject population [@hauser2018, and we expect that following current recommendations for participant recruitment, such as using qualification cutoffs or another recruitment site would result in fewer participants answering randomly [@eyal2021;@peer2017].
+We believe the high level of random guessing is an artifact of the subject population [@hauser2018], and we expect that following current recommendations for participant recruitment, such as using qualification cutoffs or another recruitment site would result in fewer participants answering randomly [@eyal2021;@peer2017].
 

### 2022-08-16 Papers/Paper/manuscript.Rmd
Roger updated ms

@@
 
-To assess the shape of the RT-surprisal relationship, we fit generalized additive models (GAMs).
+Due to previous reports of a length--frequency interaction in RT measures [@kliegl-etal:2006], before pursuing our primary question of the functional form of the surprisal--RT relationship, as an exploratory measure we fitted generalized additive models (GAMs) with not only the main effects but also the two-way interactions between surprisal, length, and frequency, for the current word and for the previous word.
+This analysis revealed significant effects of current-word and previous-word surprisal, current-word and previous-word length, and significant interactions of current-word frequency by length and frequency by surprisal.
+The other main effects and interactions did not reach statistical significance.^[These are results from `mgcv`'s `summary()`; the $p$-values are approximate.] Appendix C provides tables and plots of these effects and interactions for GPT-2.
+The interactions can be summarized as long low-frequency words and surprising, high-frequency words as having especially long RTs; and surprising, low-frequency words as having shorter RTs than would otherwise be predicted.
+However, these effects are small in terms of variance explained compared to the current-word surprisal effect, which is by far the largest single effect in the model.
+For simplicity we therefore set aside the interaction terms involving surprisal for the remainder of this analysis.
+
+To assess the shape of the RT-surprisal relationship, we then fitted generalized additive models (GAMs).
 For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.
@@
 All 4 models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data.
+To determine the goodness of fit of a model in which word probability effects on RT are taken to be linear in surprisal, we also fitted GAM models with both parametric linear and nonparametric non-linear terms for surprisal; for all but the n-gram model, these analyses supported a linear effect of surprisal (Appendix D).
 
@@
 
+We use `mgcv`'s `ti()` tensor interaction terms to test all main effects and two-way interactions among frequency, length, and surprisal for the current word and for the previous word.
+These effects are visualized in Figure \@ref(fig:gam-grid) and `mgcv`'s approximate significance levels are give in Table XXX insert table XXX.
+Based on these approximate significance levels, the main effects of current and previous word surprisal and length are significant, as are the current-word frequency-by-length and frequency-by-surprisal interactions; other terms are not statistically significant.
+These significant interactions can be summarized as especially long, infrequent words being especially slow to select; especially frequent and surprising words being especially slow to select; and especially infrequent and surprising words being less slow to select than a main-effects-only model would predict.
+The data driving these interactions are in the sparse tails of the word length and surprisal distributions, and as the $F$ statistics in Table XXX show, their variance explained is small relative to the large effect of current-word surprisal, so in the main-text analysis we set these interactions aside.
+
+XXX Veronica could we show the "approximate significance of smooth terms" results for gam_gpt_3 in a table in this Appendix as well?
+
+(ref:gam-grid) Generalized Additive Model main effects and two-way interactions among frequency, length, and surprisal for A-Maze reading of the Natural Stories corpus.
+Confidence bands do not take into account the uncertainty associated with `mgcv` hyperparameter estimation (@wood:2006GAMs).
+
+```{r gam-grid, out.height="100%",  fig.pos="ht", fig.cap="(ref:gam-grid)"}
+
+knitr::include_graphics(here("Analysis/models/gam_grid.png"))
+
+```
+
+# Appendix D
+
+The `mgcv` package's implementation of Generalized Additive Models [@wood:2006GAMs] allows linear and nonparametric spline effects of the same continuous predictor to be entered simultaneously into a model.
+Doing so associates only the nonlinear part of the effect to the spline term, allowing for approximate statistical testing of the linear and non-linear components of the effect respectively.
+We thus test for whether the effect of surprisal on A-Maze RTs is best described as linear or includes a non-linear component, using the `mgcv` formula:
+
+`rt ~ surprisal + s(surprisal, bs="cr", k=20) + ti(freq, bs="cr") + ti(len, bs="cr") + prev_surp + s(prev_surp, bs="cr", k=20) + ti(prev_freq, bs="cr") + ti(prev_len, bs="cr")`
+
+The results are in Table \@ref(tab:gam-linearity-table).
+For all but the 5-gram surprisal estimate, there is overwhelming evidence for a linear contribution of current-word surprisal, but little to no evidence for a non-linear contribution.
+For the 5-gram estimate, there is overwhelming evidence for the linear term, and some evidence for a nonlinearity as well.
+Consulting \@ref(fig:gam) shows that this nonlinearity takes the form of the surprisal effect dwindling to zero in the sparse tail of high-surprisal words.
+This nonlinearity is plausibly due to the measurement error (high variance) of using counts to estimate very low multinomial probabilities.
+Taken together with the 5-gram model's inferior overall fit, we conclude from this analysis that the evidence is quite strong that, like self-paced reading and eye tracking, A-Maze reading of naturalistic texts exhibits a linear effect of surprisal on RTs.
+
 ```{r gam-linearity-table}
@@
 ```
-
-# Appendix D
-(ref:gam-grid) CAPTION GOES HERE
-
-```{r gam-grid, out.height="100%",  fig.pos="ht", fig.cap="(ref:gam-grid)"}
-
-knitr::include_graphics(here("Analysis/models/gam_grid.png"))
-
-```

### 2022-08-17 Analysis/models.Rmd
add table for app C; futz with table locations

@@
   pivot_wider(names_from=model, values_from=pvalue) %>% write_rds(here("Analysis/models/gams_linearity_test.rds"))
+```
+
+```{r}
+gam_gpt_3 <- gam(formula3_interact, data=gpt_mean_data, method="REML")
+  stable <- summary(gam_gpt_3)$s.table %>% as_tibble(rownames="Term") %>% rename(pvalue=`p-value`, `F-statistic`=`F`) %>% select(Term, `F-statistic`, pvalue) %>% mutate(`F-statistic`=signif(`F-statistic`,digits=4), pvalue=pvalue(pvalue,accuracy=.0001,add_p=T)) %>%
+    write_rds(here("Analysis/models/gam_gpt_3_summ.rds"))
+
 ```

### 2022-08-17 Papers/Paper/manuscript.Rmd
add table for app C; futz with table locations

@@
 
+Full numerical results from the fitted regression models are shown in Table \@ref(tab:pre-error) for A-maze and in Table \@ref(tab:spr-table) for SPR.
+
 ```{r pre-error}
@@
 
-knitr::kable(summ, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models.
+knitr::kable(summ, format="latex", position="!h",caption="Predictions from fitted Bayesian regression models.
 All terms were centered, but not rescaled.
@@
 
-knitr::kable(summ_spr, format="latex", position="ht",caption="Predictions from fitted regression models for SPR data.
+knitr::kable(summ_spr, format="latex", position="!h",caption="Predictions from fitted regression models for SPR data.
 All terms were centered, but not rescaled.
@@
 We use `mgcv`'s `ti()` tensor interaction terms to test all main effects and two-way interactions among frequency, length, and surprisal for the current word and for the previous word.
-These effects are visualized in Figure \@ref(fig:gam-grid) and `mgcv`'s approximate significance levels are give in Table XXX insert table XXX.
+These effects are visualized in Figure \@ref(fig:gam-grid) and `mgcv`'s approximate significance levels are give in Table \@ref(tab:gam-interact-table).
 Based on these approximate significance levels, the main effects of current and previous word surprisal and length are significant, as are the current-word frequency-by-length and frequency-by-surprisal interactions; other terms are not statistically significant.
 These significant interactions can be summarized as especially long, infrequent words being especially slow to select; especially frequent and surprising words being especially slow to select; and especially infrequent and surprising words being less slow to select than a main-effects-only model would predict.
-The data driving these interactions are in the sparse tails of the word length and surprisal distributions, and as the $F$ statistics in Table XXX show, their variance explained is small relative to the large effect of current-word surprisal, so in the main-text analysis we set these interactions aside.
-
-XXX Veronica could we show the "approximate significance of smooth terms" results for gam_gpt_3 in a table in this Appendix as well?
+The data driving these interactions are in the sparse tails of the word length and surprisal distributions, and as the $F$ statistics in Table \@ref(tab:gam-interact-table) show, their variance explained is small relative to the large effect of current-word surprisal, so in the main-text analysis we set these interactions aside.
+
+```{r gam-interact-table}
+
+gam_interact <- read_rds(here("Analysis/models/gam_gpt_3_summ.rds"))
+
+knitr::kable(gam_interact, format="latex", position="ht",caption="Significance of Generalized Additive Model main effects and two-way interactions among frequency, length, and surprisal for A-Maze reading of the Natural Stories corpus.") %>%   kable_styling(font_size = 10)
+
+```
 
@@
 
-```{r gam-grid, out.height="100%",  fig.pos="ht", fig.cap="(ref:gam-grid)"}
+```{r gam-grid, out.height="90%",  fig.pos="ht", fig.cap="(ref:gam-grid)"}
 

### 2022-08-18 Papers/Paper/manuscript.Rmd
small tweaks by roger

@@
 We then rescaled the word counts to get the log2 frequency of occurrences per 1 billion words, so higher values indicate higher log frequencies.
-We got per-word surprisals for each of 4 different language models, covering a range of common architectures: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].
+We got per-word surprisals for each of 4 different language models, covering a range of common architectures: a Kneser-Ney smoothed 5-gram; the long short-term memory recurrent neural network model of @gulordava18, which we refer to as GRNNN; Transformer-XL [@daiTransformerXLAttentiveLanguage2019]; and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].
 For all of these predictors, we used both the predictor at the current word as well as lagged predictors from the previous word.

### 2022-08-19 Provo/Analysis/read_results.R
remove provo expt

@@ -1,72 +1,0 @@
-library(tidyverse)
-library(stringr)
-library(urltools)
-#function reads in data from a maze file plus demographic, other questions.
-read_in_data <- function(filename){
-  #reads data in generically because different format for different controllers
-#filename="pilot_1_raw"
-  data <- read_csv(filename, comment="#", col_names=c("time", "MD5", "controller", "item", "elem",
-                                                      "type", "group", "col_8", "col_9", "col_10", "col_11", "col_12", "col_13", "col_14"), col_types=cols(
-                                                        time=col_integer(),
-                                                        MD5=col_character(),
-                                                        controller=col_character(),
-                                                        item=col_integer(),
-                                                        elem=col_integer(),
-                                                        type=col_character(),
-                                                        group=col_integer(),
-                                                        col_8=col_character(),
-                                                        col_9=col_character(),
-                                                        col_10=col_character(),
-                                                        col_11=col_character(),
-                                                        col_12=col_character(),
-                                                        col_13=col_character(),
-                                                        col_14=col_character()
-                                                      )) %>% mutate_all(url_decode) #deal with %2C issue
-  
-  order_seen <- data %>% 
-    select(time, MD5, group) %>% 
-    filter(!is.na(group)) %>% 
-    unique() %>% 
-    group_by(time, MD5) %>% 
-    mutate(trial_num = 1:n()) %>% 
-    type_convert()
-  
-  #split off non-maze, participant level stuff and process them
-  other <- filter(data, controller %in% c("Form", "Question")) %>% 
-    select(time, MD5, col_8, col_9) %>% 
-    group_by(time, MD5) %>%
-    pivot_wider(names_from=col_8, values_from=col_9) %>% 
-    rename(gender=`Please select your gender.`,
-           native=`Are you a native speaker of English?`,
-           residence=`Do you currently reside in the United States?`) %>% 
-    type_convert()
-  #TODO:read for comments
-  
-  #take the Maze task results, relabel and type appropriately
-  maze<- filter(data, controller=="Maze") %>% 
-    select(time, MD5, type, group, word_num=col_8, word=col_9, distractor=col_10, on_right=col_11, correct=col_12, rt=col_13, sentence=col_14) %>% 
[…]

### 2022-08-19 Provo/provo_materials/experiment.js

@@ -1,151 +1,0 @@
-var shuffleSequence = seq("code", "setcounter", "welcome", "intro", "intro-practice",  followEachWith("practice", "sep"), "end-practice",randomize(anyOf(startsWith("critical")),
-"instructions2", anyOf("questionnaire"),"topic","debriefing");
-
-var showProgressBar =false;
-
-var defaults = [
-		"Question", {
-				as: ["yes", "no"],
-        presentAsScale: false,
-        presentHorizontally: false,
-    },
-    "Maze", {redo: true},
-];
-
-var code = Math.floor(Math.random()*100000000);
-var sendingResultsMessage = "The results are now being transferred.  Please wait.";
-var completionMessage = "Thank you for your participation.  The results were successfully transmitted.  Your participation code is: " + code.toString();
-var completionErrorMessage = "The transmission of the results failed.  Please contact online_experiment@mit.edu and retry the transmission again by clicking the link.  Your participation code is: " + code.toString();
-
-var items = [
-    ["code", "DashedSentence", {s:code.toString(), mode:"speeded acceptability", wordTime:1}],
-	["setcounter", "__SetCounter__", { }],
-    ["welcome", "Message", {html:'<table width="100%"><tr><td valign="top" align="right">Department of Brain and Cognitive Sciences<br>Massachusetts Institute of Technology<br>77 Massachusetts Avenue<br>Cambridge, MA 02139-4307, USA</tr></table>\
-<h2>Thank you very much for your participation!</h2><p>This task is part of a MIT scientific research project. Your decision to complete this task is voluntary. There is no way for us to identify you. The only information we will have, in addition to your responses, is the time at which you completed the survey. The results of the research may be presented at scientific meetings or published in scientific journals. Clicking on the link below indicates that you are at least 18 years of age and agree to complete this task voluntarily.'}],
-    ["topic", "Form", {html:'Thank you for completing the experiment. Very briefly, what do you think this study is about?<br/><textarea name="topic" rows="3" cols="50" autofocus="true"></textarea>'}],
-    ["debriefing", "Message", {html:'<p>Thank you.  You will receive the participation code on the next page.</p>\n\n<p>Purpose of this study (feel free to skip): We’re generally interested in how the human brain processes language. The present study is testing out a new method for studying what types of sentence constructions are easier or harder to read. Your data will help us to answer these questions.</p>'}],
[…]

### 2022-08-19 Papers/Paper/manuscript.Rmd
move file

@@
 
-comp_comp <- read_rds(here("Analysis/comp.rds")) %>% 
+comp_comp <- read_rds(here("Analysis/models/comp.rds")) %>% 
   filter(source %in% c(">80% correct","<80% correct", "SPR 1st story"))
 
-comp_tally <- read_rds(here("Analysis/comp.rds")) %>% group_by(source) %>% filter(correct>4) %>% 
+comp_tally <- read_rds(here("Analysis/models/comp.rds")) %>% group_by(source) %>% filter(correct>4) %>% 
   summarize(good=round(sum(pct)*100))

### 2022-08-19 Analysis/models.Rmd
move file

@@
 
-write_rds(all_bootstrapped_predictions,"bootstrapped_GAM_predictions.rds")
-write_rds(all_surprisal_predictions,"bootstrapped_GAM_surprisal_predictions.rds")
+write_rds(all_bootstrapped_predictions,here("Analysis/models/bootstrapped_GAM_predictions.rds"))
+write_rds(all_surprisal_predictions,here("Analysis/models/bootstrapped_GAM_surprisal_predictions.rds"))
 
@@
 #gam1 <- read_rds(here("Analysis/models/gam_predictions.rds")) %>%  
-gam1 <- read_rds(here("Analysis/bootstrapped_GAM_surprisal_predictions.rds")) %>%  
+gam1 <- read_rds(here("Analysis/models/bootstrapped_GAM_surprisal_predictions.rds")) %>%  
   ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
@@
                                      gpt=bootstrapped_gam_predictions_gpt_spr)
-write_rds(all_bootstrapped_predictions_spr,"bootstrapped_spr_GAM_predictions.rds")
+write_rds(all_bootstrapped_predictions_spr,here("Analysis/models/bootstrapped_spr_GAM_predictions.rds"))
 
@@
   select(-term)
-write_rds(all_surprisal_predictions_spr,"bootstrapped_spr_GAM_surprisal_predictions.rds")
+write_rds(all_surprisal_predictions_spr,here("Analysis/models/bootstrapped_spr_GAM_surprisal_predictions.rds"))
 
@@
 #gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  
-gam1 <- read_rds(here("Analysis/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  
+gam1 <- read_rds(here("Analysis/models/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  
   ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+

### 2022-08-19 Papers/Paper/manuscript.Rmd
move file

@@
 
-gam1 <- read_rds(here("Analysis/bootstrapped_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+gam1 <- read_rds(here("Analysis/models/bootstrapped_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+
@@
 ```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
-gam1 <- read_rds(here("Analysis/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+gam1 <- read_rds(here("Analysis/models/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+

### 2022-08-19 Analysis/models.Rmd
move files

@@
           r_squared=map_chr(model_str, r_squared)) %>% 
-   select(-model_str) %>% write_rds(here("Analysis/maze_model_compare.rds")) %>%  knitr::kable()
+   select(-model_str) %>% write_rds(here("Analysis/models/maze_model_compare.rds")) %>%  knitr::kable()
 
@@
           `r_squared`=map_chr(model_str, r_squared)) %>% 
-   select(-model_str)%>% write_rds(here("Analysis/spr_model_compare.rds")) %>% kable()
+   select(-model_str)%>% write_rds(here("Analysis/models/spr_model_compare.rds")) %>% kable()
 

### 2022-08-19 Papers/Paper/manuscript.Rmd
move files

@@
 
-read_rds(here("Analysis/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data.
+read_rds(here("Analysis/models/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data.
 Each row shows the additional predictive value gained from adding that model to another model.
@@
 ```{r}
-maze <- read_rds(here("Analysis/maze_model_compare.rds"))
-spr <- read_rds(here("Analysis/spr_model_compare.rds"))
+maze <- read_rds(here("Analysis/models/maze_model_compare.rds"))
+spr <- read_rds(here("Analysis/models/spr_model_compare.rds"))
 ```
@@
 
-read_rds(here("Analysis/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on SPR data.
+read_rds(here("Analysis/models/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on SPR data.
 Each row shows the additional predictive value gained from adding that model to another model.

### 2022-08-19 Papers/Paper/manuscript.Rmd
switch to glossa format

@@
 shorttitle        : "A-maze of Natural Stories"
-
+shortauthors: Veronica Boyce & Roger Levy
 author: 
   - name          : "Veronica Boyce"
-    affiliation   : "1"
+    affiliation   : "Stanford University"
     corresponding : yes    # Define only one corresponding author
-    address       : "450 Jane Stanford Way, Building 420, Stanford University, Stanford, CA 94305"
     email         : "vboyce@stanford.edu"
@@
   - name          : "Roger Levy"
-    affiliation   : "2"
+    affiliation   : "Massachusetts Institute of Technology"
     role:
@@
 
-authornote: |
-   We thank the AMLAP 2020 audience, the Computational Psycholinguistics Lab at MIT, the Language and Cognition Lab at Stanford, the QuantLang Lab at UC Irvine, and Mike Frank for feedback on this work.
-
+classoption: [times, biblatex]
+#
+# Possible classoptions:
+# - [times] for Times font (default if no option is chosen)
+# - [cm] for Computer Modern font
+# - [lucida] for Lucida font (not freely available)
+# - [brill] open type font, freely downloadable for non-commercial use from
+#   http://www.brill.com/about/brill-fonts; requires xetex
+# - [charis] for CharisSIL font, freely downloadable from http://software.sil.org/charis/
+# - for the Brill an CharisSIL fonts, you have to use the XeLatex
+#   typesetting engine (not pdfLatex)
+# - for headings, tables, captions, etc., Fira Sans is
+#   used: https://www.fontsquirrel.com/fonts/fira-sans
+# - [biblatex] for using biblatex (the default is natbib, DO NOT LOAD the
+#   natbib package in this file, it is loaded automatically via the document
+#   class glossa.cls)
+# - [linguex] loads the linguex example package
+#
+# !!! A note on the use of linguex: in glossed examples, the third line of the
+# example (the translation) needs to be prefixed with \glt. This is to allow a
+# first line with the name of the language and the source of the example. See
+# example (2) in the text for an illustration.
+#
+# !!! A note on the use of bibtex: for PhD dissertations to typeset correctly
+# in the references list, the Address field needs to contain the city
+# (for US cities in the format "Santa Cruz, CA")
+#
+bibliography: ["r-references.bib","refs.bib"]
+biblio-style: "apalike"
+link-citations: true
+# The bibliography style is set automatically by glossa.cls when using
+# either natbib or biblatex.
+# Pandoc's citeproc is not supported.
+#
 abstract: |
@@
 Overall, we expanded the scope of experimental materials, and thus theoretical questions, that can be studied with the Maze task.
-
-bibliography      : ["r-references.bib","refs.bib"]
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_pdf
+# Specify keywords here:
+keywords: stylesheet, glossa, article
+# To add the word count, uncomment the following option and replace the number
+# with the document word count.
+# wordcount: 1000
+output: 
+  bookdown::pdf_book:
+    base_format: rticles::glossa_article
+    number_sections: yes
 ---
@@
 ```{r setup, include = FALSE}
-knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
+knitr::opts_chunk$set(echo=F, warning=F, message=F)#dev = "png", dev.args = list(type = "cairo-png")
  library("papaja")
@@
 
-```{r errors, out.height="25%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:error-cap)"}
+```{r errors, out.width="\\textwidth", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:error-cap)"}
 error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
@@
 
-```{r gam, out.height="40%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:gam-cap)"}
+```{r gam, out.width="\\textwidth", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:gam-cap)"}
 
@@
 
-```{r spr-gam, out.height="40%", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
+```{r spr-gam, out.width="\\textwidth", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
 gam1 <- read_rds(here("Analysis/models/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
@@
 
-```{r coeffs-maze, out.height="35%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:coeffs)"}
+```{r coeffs-maze, out.width="\\textwidth", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:coeffs)"}
 
@@
 
-```{r coeffs-spr, out.height="35%", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:coeffs-spr)"}
+```{r coeffs-spr, out.width="\\textwidth", fig.width=8, fig.height=3, fig.pos="ht", fig.cap="(ref:coeffs-spr)"}
 test_plot %>% filter(type=="SPR") %>% mutate(time=factor(time, levels=c("Current","Previous","2Previous","3Previous")))%>% ggplot(aes(y=Term, x=Estimate, group=model, color=model, shape=model))+
@@
 
-```{r gam-grid, out.height="90%",  fig.pos="ht", fig.cap="(ref:gam-grid)"}
+```{r gam-grid, out.width="\\textwidth",  fig.pos="ht", fig.cap="(ref:gam-grid)"}
 
@@
 ```
+
+# Ethics and consent (optional) {#ethics-and-consent-optional .unnumbered}
+
+Research involving human subjects, human material, or human data, must
+have been performed in accordance with the Declaration of Helsinki.
+Where applicable, the studies must have been approved by an appropriate
+ethics committee. In the published paper, the authors should include a
+statement detailing this approval, including the name of the ethics
+committee and reference number of the approval. The identity of the
+research subject(s) should be anonymised whenever possible. For research
+involving human subjects, informed consent to participate in the study
+must be obtained from participants (or their legal guardian).
+
+# Funding information (optional) {#funding-information-optional .unnumbered}
+
+Should the research have received a funding grant, then the grant
+provider and grant number can be detailed in this section.
+
+# Acknowledgements (optional) {#acknowledgements-optional .unnumbered}
+
+ We thank the AMLAP 2020 audience, the Computational Psycholinguistics Lab at MIT, the Language and Cognition Lab at Stanford, the QuantLang Lab at UC Irvine, and Mike Frank for feedback on this work.
+
+# Competing interests (mandatory) {#competing-interests-mandatory .unnumbered}
+
+If any of the authors have any competing interests then these must be
+declared. Consult the [Competing
+Interests](http://www.glossa-journal.org/about/competinginterests/)
+section on the Glossa website for more information. If there are no
+competing interests to declare then the following statement should be
+present: "The author(s) has/have no competing interests to declare".
+
+# Authors' contributions (optional) {#contrib .unnumbered}
+
+Here you can provide a sentence or a short paragraph detailing the
+contribution of each author to the paper.

### 2022-08-22 Papers/Paper/manuscript.Rmd
Roger took revision pass through ms

@@
 Moreover, the Maze task yielded useable reaction time data with word predictability effects that were linearly related to surprisal, the same pattern found with other incremental methods.
-Crucially, Maze reaction times show a tight relationship with properties of the current word, but little spillover of effects from previous words.
+Crucially, Maze reaction times show a tight relationship with properties of the current word, with little spillover of effects from previous words.
 This superior localization is an advantage of Maze compared with other methods.
@@
 
-A major feature of human language is that not all sentences or utterances are equally easy to successfully comprehend.
+A major feature of human language processing is that not all sentences or utterances are equally easy to successfully comprehend.
 Sometimes this is mostly or entirely due to the linguistic structure of the sentence: for example, *The rat that the cat that the dog chased killed ate the cheese* is more difficult than *The rat that was killed by the cat that was chased by the dog ate the cheese* even though the meaning of the two sentences is (near-)identical.
@@
 Self-paced reading (SPR; @mitchell:1984) is a somewhat less natural paradigm in which the participant manually control the visual presentation of the text by pressing a button.
-In its generally preferred variant, moving-window self-paced reading, words are revealed one at a time or one group at a time, wihth every press of the button masks the currently presented word (group) and simultaneously reveals the next.
+In its generally preferred variant, moving-window self-paced reading, words are revealed one at a time or one group at a time: every press of the button masks the currently presented word (group) and simultaneously reveals the next.
 The time spent between button presses is the unique RT measure for that word (group).
@@
 Participants must choose the correct word, and their time to selection is treated as the reaction time, or RT.
-(We intentionally overload the abbreviation "RT" and use it for Maze reaction times as well as reading times from eye tracking and SPR, because the desirable properties of reading times turn out to hold for Maze reaction times as well.) @forsterMazeTaskMeasuring2009 introduced two versions of the Maze task: lexical "L"-maze where the distractors are non-word strings, and grammatical "G"-maze where the distractors are real words that don't fit with the context of the sentence.
+(We deliberately overload the abbreviation "RT" and use it for Maze reaction times as well as reading times from eye tracking and SPR, because the desirable properties of reading times turn out to hold for Maze reaction times as well.) @forsterMazeTaskMeasuring2009 introduced two versions of the Maze task: lexical "L"-maze where the distractors are non-word strings, and grammatical "G"-maze where the distractors are real words that don't fit with the context of the sentence.
 In theory, participants must fully integrate each word into the sentence in order to confidently select it, which may require mentally reparsing previous material in order to allow the integration and selection of a disambiguating word.
@@
 A variety of pre-trained models exist, with varied internal architectures and training methods, but all of them generate measures of predictability.
-Predictability is often measured in bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time etc).
+Predictability is often measured in bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time, etc.).
 
@@
 In traditional G-maze tasks, with hand-crafted distractors and attentive participants, errors are rare and data loss is a small issue.
-However, this data loss is much worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020].
+However, this data loss can be worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020].
 The high errors are likely from some combination of participants guessing randomly and from auto-generated distractors that in fact fit the sentence; as @boyceMazeMadeEasy2020 noted, some distractors, especially early in the sentence, were problematic and caused considerable data loss.
@@
 
-## SPR-Maze correlation
+## SPR--Maze correlation
 We compared the correlations between the Maze and SPR RTs to within-Maze and within-SPR correlations.
@@
 To look at the functional form of the relationship between surprisal and RT data, we fit Generalized Additive Models (GAMs) to allow for non-linear relationships.
-GAMs are harder to interpret than linear models, so to measure effect sizes and assess spillover, we used linear mixed models.
+GAM model summaries can be harder to interpret than those for linear models, so to measure effect sizes and assess spillover, we used linear mixed models.
 Finally, in order to determine which language model best predicts the RT data, we fit additional linear models with predictors from multiple language models to look at their relative contributions.
@@
 
-For linear models of the SPR data, we were unable to fit the full mixed effects structure.
-The best model we could fit had by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and a per-word-token random intercept, fit with lme4 [@batesFittingLinearMixedeffects2015], as this structure did not fit reliably in brm.
-
-For model comparison, we took by-item averaged data to aid in fast model fitting.
+For linear models of the SPR data, we were unable to fit a single model whose random effects structure was maximal with respect to all fixed-effects predictors.
+We report results for the best (in terms of having maximal random effects structure with respect to fixed effects of primary theoretical interest) single model we could fit by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and a per-word-token random intercept, fit with lme4 [@batesFittingLinearMixedeffects2015], as this model specification did not fit reliably in brm.
+
+For model comparisons, we took by-item averaged data to aid in fast model fitting.
 We included frequency, length, and their interaction in all models.
-Then we fit models with either 1 or 2 sources of surprisal using lm [@batesFittingLinearMixedeffects2015] and assessed the effect of adding the second surprisal source with an anova.
+Then we fit simple linear regression models (using `R`'s `lm()`) with either 1 or 2 sources of surprisal and assessed the effect of adding the second surprisal source with an F test (using `R`'s `anova()`).
 
@@
 Overall, `r comp_tally %>% filter(source=="all Maze") %>% pull(good)`% of Maze participants got 5 or 6 questions right (`r comp_tally %>% filter(source=="<80% correct") %>% pull(good)`% of low-accuracy participants and `r comp_tally %>% filter(source==">80% correct") %>% pull(good)`% of high-accuracy participants) compared to `r comp_tally %>% filter(source=="SPR") %>% pull(good)`% of all SPR reads and `r comp_tally %>% filter(source=="SPR 1st story") %>% pull(good)`% of 1st SPR reads.
-These differences cannot be directly attributed to methods, as the participant populations differed.
+These differences cannot necessarily be attributed to methods, as the participant populations differed.
 While both studies were conducted on Mturk, the quality of Mturk data has decreased from 2011 when the SPR was collected to 2020 when the A-maze was collected [@chmielewski2020].
@@
 
-(ref:scatter) Correlation between SPR and Maze data. RTs were averaged across participants per word and then averaged together within each sentence. RTs in ms.
+(ref:scatter) Correlation between SPR and Maze data.
+RTs (measured in milliseconds) were averaged across participants per word and then averaged together within each sentence, so that each point represents the average RT in the two methods for one sentence in the corpus.
+Presented on a fixed scale coordinate system where 1 millisecond of RT takes equal physical space on both axes.
+Line and confidence interval reflect best linear fit regression of SPR time against Maze time.
 
@@
   inner_join(spr_collapse)
-ggplot(both, aes(x=maze, y=spr))+geom_point(alpha=.5, size=1)+labs(x="A-Maze", y="SPR")+coord_fixed(ratio=1)+geom_smooth(method="lm")
+ggplot(both, aes(x=maze, y=spr)) + 
+  geom_point(alpha=.5, size=1) + 
+  labs(x="A-Maze", y="SPR") + 
+  coord_fixed(ratio=1) +
+  geom_smooth(method="lm")
+
+# mod1 <- lm(spr ~ maze,both)
+# mod2 <- lm(maze ~ spr,both)
+# slope2 <- 1/coef(mod2)[2]
+# intercept2 <- with(both,mean(spr) - slope2 * mean(maze))
+# matr <- with(both,matrix(c(var(maze),cov(spr,maze),cov(spr,maze),var(spr)),2,2))
+# slope <- eigen(matr)$vectors[2,1]/eigen(matr)$vectors[1,1]
+# intercept <- with(both,mean(spr) - slope * mean(maze))
+# ggplot(both, aes(x=maze, y=spr)) + 
+#   geom_point(alpha=.5, size=1) + 
+#   labs(x="A-Maze", y="SPR") + 
+#   #theme(aspect.ratio=1) +
+#   coord_fixed(ratio=1) +
+#   stat_ellipse(type="norm",segments=501) +
+#   stat_ellipse(type="t",color="yellow",segments=501) +
+#   #geom_smooth(method="lm") 
+#   geom_abline(intercept=coef(mod1)[1],slope=coef(mod1)[2],color="blue") + 
+#   geom_abline(intercept=intercept2,slope=slope2,color="red") + 
+#   geom_abline(intercept=intercept,slope=slope,color="green")
+#   #geom_abline(intercept=0,slope=1)
+# 
+# slope1 <- 1/coef(mod1)[2]
+# intercept1 <- with(both,mean(maze) - slope1 * mean(spr))
+# slopeA <- eigen(matr)$vectors[1,1]/eigen(matr)$vectors[2,1]
+# interceptA <- with(both,mean(maze) - slopeA * mean(spr))
+# ggplot(both, aes(x=spr, y=maze)) + 
+#   geom_point(alpha=.5, size=1) + 
+#   labs(y="A-Maze", x="SPR") + 
+#   theme(aspect.ratio=1) +
+#   #coord_fixed(ratio=1) +
+#   stat_ellipse(type="norm",segments=501) +
+#   stat_ellipse(type="t",color="yellow",segments=501) +
+#   #geom_smooth(method="lm") 
+#   geom_abline(intercept=coef(mod2)[1],slope=coef(mod2)[2],color="red") + 
+#   geom_abline(intercept=intercept1,slope=slope1,color="blue") + 
+#   geom_abline(intercept=interceptA,slope=slopeA,color="green")
+#   #geom_abline(intercept=0,slope=1)
 ```
@@
 We examined how much spillover A-maze and SPR each had by fitting linear models with predictors from current and previous words.
-Large effects from previous words are evidence for spillover; effects of the current word that dwarf any lagged effects is evidence for localization.
-
-We modeled reading time as a function of surprisal, frequency, and length as well as surprisal x length and frequency x length interactions.
+Large effects from previous words are evidence for spillover; effects of the current word dwarfing any lagged effects would be evidence for localization.
+
+We modeled reading time as a function of surprisal, frequency, and length as well as surprisal$\times$length and frequency$\times$length interactions.
 For all of these, we included the predictors for the current and previous word, and we centered, but did not rescale, all predictors.
@@
 This difference in effect size is disproportionate to the overall speed of the methods; the predicted intercept for the Maze task was roughly 880 ms and for SPR was roughly 360 ms.
-Thus Maze is is 2-3 times as slow as SPR but has roughly 10 times larger effects.
+Thus Maze is 2--3 times as slow as SPR but has roughly 10 times larger effects.
 
@@
 The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020, and contrasts with the Maze results, where the 5-gram model had the worst fit and did not provide additional predictive value over the other models.
+The fact that the neural language models, which capture deeper features of linguistic structure  [@huSystematicAssessmentSyntactic2020,@wilcox-etal:2022-using-computational-models], dominate the more superficial $n$-gram models for Maze but not SPR suggests that Maze RTs reflect richer language structure-related processes during real-time comprehension than do SPR RTs.
 
@@
 For Maze, we found large effects of the current words surprisal and length, which dwarfed any spillover effects from previous word predictors.
-In contrast, for SPR, we found effects of roughly equal sizes from the current and previous words.
-Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus and thus generating less noisy data.
+In contrast, for SPR, we found effects of roughly equal sizes from the current and previous words.[^Furthermore, the typical spillover profile for SPR data may be worse than suggested by the Natural Stories corpus SPR data: for example, @smithEffectWordPredictability2013, found that most of a word's surprisal effect showed up only one to two words downstream.] Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus, and thus generating less noisy data.
 
@@
 We found that overall, the models were more predictive of the A-maze data than SPR data; however, the hierarchy of the model's predictive performance also differed between the A-maze and SPR datasets.
-This difference suggests that how predictive a language model is of RTs may depend on the task.
-Further comparisons between different processing methods on the same materials could be useful for identifying how task demands influence language processing [ex.
+This difference suggests that how well a language model predicts human RTs may depend on task.
+Maze RTs were by far best predicted by language models that capture rich features of natural language grammatical structure, unlike SPR RTs which were nearly equally well predicted by more superficial $n$-gram models.
+These findings further add to the evidence that the Maze task is favorable for RT-based investigations of underlying linguistic processing in the human mind.
+More broadly, further comparisons between different processing methods on the same materials could be useful for a deeper understanding of how task demands influence language processing [ex.
 @bartekSearchOnlineLocality2011].
 
-Overall, A-maze has good localization, although some  models showed small but statistically reliable effects of the past word.
-On the whole, however, our results support the idea that Maze forces language processing to be close to word-by-word, and thus the Maze task can be used under the assumption that the RT of a word primarily reflects its own properties and not those of earlier words.
+Overall, A-maze has excellent good localization, although some  models showed small but statistically reliable effects of the past word.
+On the whole, our results support the idea that Maze forces language processing to be close to word-by-word, and thus the Maze task can be used under the assumption that the RT of a word primarily reflects its own properties and not those of earlier words.
 
@@
 
-Error-correction Maze  reduces perverse incentives from the desire to complete the task quickly compared to traditional Maze, but clicking randomly is still more efficient than doing the task.
+Compared to traditional Maze, error-correction Maze reduces perverse (from the experimenter's point of view) incentives to complete the task as quickly as possible.
+However, even with error-correction Maze, clicking randomly is still likely faster than doing the task.
 In discussing this work, we received the suggestion that one way to further disincentivize random clicking would be to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms) before correcting their mistake.
@@
 Error-correction Maze records RTs for words after a participant makes a mistake in the sentence.
-In our analyses, we excluded this post-error data, but we believe it is an open question whether data from after a participant makes a mistake is usable.
-That is, does it show the same profile as RTs from pre-error words, or are there traces from recovering from the mistake, and if so, how long do these effects take to fade?
-Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is hard-to-assess; if it can be used, it would make the Maze task more data efficient.
+In our analyses, we excluded these post-error data, but we believe it is an open question whether data from after a participant makes a mistake is usable.
+That is, does it show the same profile as RTs from pre-error words, or are there traces from recovering from the mistake?
+If there are, how long do these effects take to fade?
+Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is hard to assess; if it can be used, it would make the Maze task more data efficient.
 
 The Maze task is versatile and can be used or adapted for a wide range of materials and questions of interest.
-The extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty.
+Its forced incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty.
 We encourage researchers to use Maze as an incremental processing method, alone or in comparison with other methods, and we suggest that the error-correction mode be the default choice for presenting Maze materials.

### 2022-08-22 Papers/Paper/manuscript.Rmd
reorder models in figs, credit roles

@@
     email         : "vboyce@stanford.edu"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - Conceptualization
-      - Formal Analysis
-      - Investigation
-      - Software
-      - Writing - Original Draft Preparation
-      - Writing - Review & Editing
+
   - name          : "Roger Levy"
     affiliation   : "Massachusetts Institute of Technology"
-    role:
-      - Conceptualization
-      - Formal Analysis
-      - Writing - Review & Editing
 
@@
 # Specify keywords here:
-keywords: stylesheet, glossa, article
+keywords: A-Maze, reading time measure, surprisal,
 # To add the word count, uncomment the following option and replace the number
@@
 
-gam1 <- read_rds(here("Analysis/models/bootstrapped_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+gam1 <- read_rds(here("Analysis/models/bootstrapped_GAM_surprisal_predictions.rds")) %>% mutate(model=factor(model, 
+                        levels=c("5-gram",
+                                 "GRNN",
+                                 "Transformer-XL",
+                                 "GPT-2"))) %>%
+ ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+
@@
 
-all_mean_data <-  read_rds(here("Analysis/models/meaned_maze.rds"))
+all_mean_data <-  read_rds(here("Analysis/models/meaned_maze.rds")) %>% mutate(model=factor(model, 
+                        levels=c("5-gram",
+                                 "GRNN",
+                                 "TXL",
+                                 "GPT-2")))
 
@@
     coord_cartesian(xlim=c(0,28))+
-  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
+  theme(axis.text.y = element_blank(), strip.text=element_blank(), 
+        axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))
 
@@
 ```{r spr-gam, out.width="\\textwidth", fig.width=8, fig.height=4, fig.pos="ht", fig.cap="(ref:spr-gam-cap)"}
-gam1 <- read_rds(here("Analysis/models/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
+gam1 <- read_rds(here("Analysis/models/bootstrapped_spr_GAM_surprisal_predictions.rds")) %>%
+  mutate(model=factor(model, 
+                        levels=c("5-gram",
+                                 "GRNN",
+                                 "Transformer-XL",
+                                 "GPT-2"))) %>%
+  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
   geom_line()+
@@
 
- all_mean_spr <-  read_rds(here("Analysis/models/meaned_spr.rds"))
+all_mean_spr <-  read_rds(here("Analysis/models/meaned_spr.rds")) %>% mutate(model=factor(model, 
+                        levels=c("5-gram",
+                                 "GRNN",
+                                 "TXL",
+                                 "GPT-2")))
 
@@
 
-test_plot <- read_rds(here("Analysis/models/maze_spr_for_graph.rds"))
+test_plot <- read_rds(here("Analysis/models/maze_spr_for_graph.rds"))%>% mutate(model=factor(model, 
+                        levels=c("5-gram",
+                                 "GRNN",
+                                 "TXL",
+                                 "GPT-2")))
 
@@
 ```{r pre-error}
-summ <- read_rds(here("Analysis/models/brms_maze_summ.rds")) %>% pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
+summ <- read_rds(here("Analysis/models/brms_maze_summ.rds")) %>%
+  pivot_wider(names_from="model", values_from=c(`Estimate`)) %>% 
   mutate(Term=factor(Term, 
@@
                   "Surp x Length", "Freq x Length", "Past Surprisal",
-                  "Past Length", "Past Freq", "Past Surp x Length", "Past Freq x Length"))
+                  "Past Length", "Past Freq", "Past Surp x Length", 
+                "Past Freq x Length")) %>%
+  select(Term, `5-gram`, GRNN, `Transformer-XL`=TXL,`GPT-2`)
 
@@
                 "2Past Surprisal", "2Past Length", "2Past Freq", "2Past Surp x Length", "2Past Freq x Length",
-                "3Past Surprisal", "3Past Length", "3Past Freq", "3Past Surp x Length", "3Past Freq x Length"))
+                "3Past Surprisal", "3Past Length", "3Past Freq", "3Past Surp x Length", "3Past Freq x Length")) %>%
+  select(Term, `5-gram`, GRNN, `Transformer-XL`=TXL,`GPT-2`)
 
@@
   arrange(Term) %>% filter(Term %in% c("s(surprisal)", "s(prev_surp)", "surprisal","prev_surp")) %>% 
-  mutate(Term=c("Spline Surprisal", "Spline Past Surprisal", "Linear Surprisal", "Linear Past Surprisal"))
+  mutate(Term=c("Spline Surprisal", "Spline Past Surprisal", "Linear Surprisal", "Linear Past Surprisal")) %>%
+  select(Term, `5-gram`, GRNN, `Transformer-XL`=TXL,`GPT-2`)
 
@@
 
-# Ethics and consent (optional) {#ethics-and-consent-optional .unnumbered}
+<!--# Ethics and consent (optional) {#ethics-and-consent-optional .unnumbered}
 
@@
 must be obtained from participants (or their legal guardian).
-
-# Funding information (optional) {#funding-information-optional .unnumbered}
+-->
+<!--# Funding information (optional) {#funding-information-optional .unnumbered}
 
 Should the research have received a funding grant, then the grant
-provider and grant number can be detailed in this section.
-
-# Acknowledgements (optional) {#acknowledgements-optional .unnumbered}
+provider and grant number can be detailed in this section.-->
+
+# Acknowledgements {#acknowledgements-optional .unnumbered}
 
@@
 
-# Competing interests (mandatory) {#competing-interests-mandatory .unnumbered}
-
-If any of the authors have any competing interests then these must be
-declared. Consult the [Competing
-Interests](http://www.glossa-journal.org/about/competinginterests/)
-section on the Glossa website for more information. If there are no
-competing interests to declare then the following statement should be
-present: "The author(s) has/have no competing interests to declare".
-
-# Authors' contributions (optional) {#contrib .unnumbered}
-
-Here you can provide a sentence or a short paragraph detailing the
-contribution of each author to the paper.
+# Competing interests {#competing-interests-mandatory .unnumbered}
+
+The authors have no competing interests to declare.
+
+# Authors' contributions {#contrib .unnumbered}
+Veronica Boyce contributed Conceptualization, Formal Analysis, Investigation, Methodology, Software, and Writing - Original Draft Preparation.
+Roger Levy contributed Conceptualization, Formal Analysis, Funding Acquisition, Methodology, Supervision, and Writing - Review & Editing.

### 2022-08-24 Papers/Paper/manuscript.Rmd
fix small errors, fuss with bib

@@
 
-  - name          : "Roger Levy"
+  - name          : "Roger P. Levy"
     affiliation   : "Massachusetts Institute of Technology"
@@
  - \usepackage{setspace}\singlespacing
- - \renewcommand{\textfraction}{0.05}
- - \renewcommand{\topfraction}{0.8}
- - \renewcommand{\bottomfraction}{0.8}
+ - \renewcommand{\textfraction}{0.00}
+ - \renewcommand{\topfraction}{1}
+ - \renewcommand{\bottomfraction}{1}
  - \renewcommand{\floatpagefraction}{1}
@@
 # Specify keywords here:
-keywords: A-Maze, reading time measure, surprisal,
+keywords: A-Maze, self-paced reading, incremental processing, reading time measure, surprisal, naturalistic text,
 # To add the word count, uncomment the following option and replace the number
@@
  library("papaja")
+library("bookdown")
+library("rticles")
  library(here)
@@
 library(tidymv)
+library(mgcViz)
 library(kableExtra)
-r_refs("r-references.bib", append=F)
+#r_refs("r-references.bib", append=F)
 theme_set(theme_bw())
@@
 
-However, there is increasing interest in collecting data during comprehension of more naturalistic materials such as stories and news articles [@demberg-keller:2008; @luke-christianson:2016-limits; @futrellNaturalStoriesCorpus2020], which offer potentially improved ecological validity and larger data scale in comparison with repeated presentation of isolated sentences out of context.
-These more materials require maintaining and integrating discourse dependencies and other type of information over longer stretches of time and linguistic material.
+However, there is increasing interest in collecting data during comprehension of more naturalistic materials such as stories and news articles [@demberg-keller:2008; @lukeLimitsLexicalPrediction2016; @futrellNaturalStoriesCorpus2020], which offer potentially improved ecological validity and larger scale data in comparison with repeated presentation of isolated sentences out of context.
+These more naturalistic materials require maintaining and integrating discourse dependencies and other types of information over longer stretches of time and linguistic material.
 Previous work leaves unclear whether the Maze task would be feasible for this purpose: the increased task demands might interfere with the demands presented by these more naturalistic materials, and vice versa.
@@
 Psycholinguistic theories may differ in their ability to predict what is easy and what is hard.
-One of the most powerful methods for studying these differential difficulty effects is to turn over control of presentation of the linguistic material to comprehender, and to measure what she takes time on.
+One of the most powerful methods for studying these differential difficulty effects is let the comprehender control the pace of presentation of the linguistic material, and to measure what she takes time on.
 For this purpose, taking measurements from experimental participants during reading, a widespread, highly practiced skill in diverse populations around the world, is of unparalleled value.
@@
 In eye-tracking, participants read a text on a screen naturally, while their saccadic eye movements are recorded on a computer-connected camera that is calibrated so that the researcher can reconstruct with high precision where the participant's gaze falls on the screen at all times [@raynerEyeMovementsReading1998].
-From these eye movements can be reconstructed various position-specific reading time measures such as *gaze duration* (the total amount of time the eyes spend on a word the first time it is fixated, or zero if the eye skipped the word the first time it was approached from the left) and *total viewing time* (the total amount of time that the word is fixated).
+These eye movements can be used to reconstruct various position-specific reading time measures such as *gaze duration* (the total amount of time the eyes spend on a word the first time it is fixated, or zero if the eye skipped the word the first time it was approached from the left) and *total viewing time* (the total amount of time that the word is fixated).
 Eye tracking data collected with state-of-the-art high-precision recording equipment offers relatively good signal-to-noise ratio, but the difficulty presented by a word can still *spill over* into reading measures on subsequent words, a dynamic that can make it hard to isolate the source of an effect of potential theoretical interest [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982].
@@
 
-Self-paced reading (SPR; @mitchell:1984) is a somewhat less natural paradigm in which the participant manually control the visual presentation of the text by pressing a button.
+Self-paced reading (SPR; @mitchell:1984) is a somewhat less natural paradigm in which the participant manually controls the visual presentation of the text by pressing a button.
 In its generally preferred variant, moving-window self-paced reading, words are revealed one at a time or one group at a time: every press of the button masks the currently presented word (group) and simultaneously reveals the next.
@@
 The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitous words as distractors for each spot of each sentence.
-This burdensome preparation may explain why the Maze task has not been widely adopted.
+This burdensome preparation may explain why the Maze task was not widely adopted.
 @boyceMazeMadeEasy2020 demonstrated a way to automatically generate Maze distractors by using language models from Natural Language Processing to find words that are high surprisal in the context of the target sentence, and thus likely to be judged infelicitous by human readers.
@@
 
-Another recent variant of the Maze task is interpolated I-maze, which uses a mix of real word distractors (generated via the A-maze process) and nonce word distractors [@vaniUsingInterpolatedMaze2021; @wilcoxTargetedAssessmentIncremental2021].
-The presence of real word distractors encourages close attention to the sentential context, while nonce words can be used as distractors where the word in the sentence is itself ungrammatical or highly unexpected, and/or it is important that the predictability of the distractor in the context is perfectly well-balanced (at zero) across all experimental conditions.
+Another recent variant of the Maze task is interpolated I-maze, which uses a mix of real word distractors (generated via the A-maze process) and non-word distractors [@vaniUsingInterpolatedMaze2021; @wilcoxTargetedAssessmentIncremental2021].
+The presence of real word distractors encourages close attention to the sentential context, while non-words can be used as distractors where the word in the sentence is itself ungrammatical or highly unexpected, and/or it is important that the predictability of the distractor in the context is perfectly well-balanced (at zero) across all experimental conditions.
 
@@
 
-  1. Do participants engage with these longer passage successfully with the A-maze task?
+  1. Do participants engage with these longer passages successfully with the A-maze task?
   2. Is A-maze as powerful and reliable a method as SPR for these longer passages?
@@
 We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the Maze paradigm.
-The materials, data, and analysis code are all available at REPO.
+The materials, data, and analysis code are all available at https://github.com/vboyce/natural-stories-maze.
 
@@
 After reading each story, each participant answered 6 binary-choice comprehension questions.
-For comparability with A-maze, we analyze only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct, leaving `r spr %>% nrow()`.
+For comparability with A-maze, we analyze only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct, leaving `r spr %>% nrow()` SPR participants.
 
@@
 ```{r}
-package_list <- c("tidyverse","brms","rstan","papaja",
-                  "lme4", "mgcv", "tidybayes", "here", "patchwork","cowplot","gridExtra","broom.mixed","tidymv","kableExtra")
+package_list <- c("tidyverse","brms","rstan","bookdown", "rticles", "papaja",
+                  "lme4", "mgcv", "mgcViz",
+ "tidybayes", "here", "patchwork","cowplot","gridExtra","broom.mixed","tidymv","kableExtra")
 
@@
 We then rescaled the word counts to get the log2 frequency of occurrences per 1 billion words, so higher values indicate higher log frequencies.
-We got per-word surprisals for each of 4 different language models, covering a range of common architectures: a Kneser-Ney smoothed 5-gram; the long short-term memory recurrent neural network model of @gulordava18, which we refer to as GRNNN; Transformer-XL [@daiTransformerXLAttentiveLanguage2019]; and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].
+We got per-word surprisals for each of 4 different language models, covering a range of common architectures: a Kneser-Ney smoothed 5-gram; the long short-term memory recurrent neural network model of @gulordava18, which we refer to as GRNN; Transformer-XL [@daiTransformerXLAttentiveLanguage2019]; and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].
 For all of these predictors, we used both the predictor at the current word as well as lagged predictors from the previous word.
@@
 ### Model specification
-To infer the shape of the relationship between our predictor variables and RTs, we fitted generalized additive models (GAMs) using `R`'s `mgcv` package to predict the mean RT (after exclusions) for each word, averaging across participants from whom we we obtained an unexcluded RT for that word.
+To infer the shape of the relationship between our predictor variables and RTs, we fit generalized additive models (GAMs) using `R`'s `mgcv` package to predict the mean RT (after exclusions) for each word, averaging across participants from whom we obtained an unexcluded RT for that word.
 We centered but did not rescale the length and frequency predictors, and left surprisal uncentered for interpretability.
@@
 We use restricted maximum likelihood (REML) smoothing for parameter estimation.
-To more fully account for the uncertainty in the smoothing parameter estimates, we fitted 101 bootstrap replicates of each GAM model; in Figures \@ref(fig:gam) and \@ref(fig:spr-gam), the best-fit lines derive from the mean estimated effect size across the bootstrap replicates, and the shaded areas indicate a 95\% bootstrap confidence interval on this effect size (the boundaries are the 2.5\% and 97.5\% quantiles of the bootstrapped replicates).
+To more fully account for the uncertainty in the smoothing parameter estimates, we fit 101 bootstrap replicates of each GAM model; in Figures \@ref(fig:gam) and \@ref(fig:spr-gam), the best-fit lines derive from the mean estimated effect size across the bootstrap replicates, and the shaded areas indicate a 95\% bootstrap confidence interval on this effect size (the boundaries are the 2.5\% and 97.5\% quantiles of the bootstrapped replicates).
 
 For linear models, we centered all predictors.
-We modeled the main effects of surprisal, length, and frequency as well as surprisal x length and frequency x length interactions.
+We modeled the main effects of surprisal, length, and frequency as well as surprisal-by-length and frequency-by-length interactions.
 For the A-maze data, we used maximal mixed effects, including by-subject slopes and a per-word-token random intercept [@barrRandomEffectsStructure2013].
-We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations) and ran models with brm [@burknerAdvancedBayesianMultilevel2018].
+We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations) and ran models with brm [@R-brms_b].
 
 For linear models of the SPR data, we were unable to fit a single model whose random effects structure was maximal with respect to all fixed-effects predictors.
-We report results for the best (in terms of having maximal random effects structure with respect to fixed effects of primary theoretical interest) single model we could fit by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and a per-word-token random intercept, fit with lme4 [@batesFittingLinearMixedeffects2015], as this model specification did not fit reliably in brm.
+We report results for the best (in terms of having maximal random effects structure with respect to fixed effects of primary theoretical interest) single model we could fit: by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and a per-word-token random intercept, fit with lme4 [@R-lme4], as this model specification did not fit reliably in brm.
 
@@
 [^2]: To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.
-This exclusion removed `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.
+This exclusion removed `r nrow(data_long)` words, or `r round(nrow(data_long)/nrow(data_no_prac)*100, digits=2)`% of trials.
 
@@
 
-Due to previous reports of a length--frequency interaction in RT measures [@kliegl-etal:2006], before pursuing our primary question of the functional form of the surprisal--RT relationship, as an exploratory measure we fitted generalized additive models (GAMs) with not only the main effects but also the two-way interactions between surprisal, length, and frequency, for the current word and for the previous word.
+Due to previous reports of a length--frequency interaction in RT measures [@kliegl-etal:2006], before pursuing our primary question of the functional form of the surprisal--RT relationship, as an exploratory measure we fit generalized additive models (GAMs) with not only the main effects but also the two-way interactions between surprisal, length, and frequency, for the current word and for the previous word.
 This analysis revealed significant effects of current-word and previous-word surprisal, current-word and previous-word length, and significant interactions of current-word frequency by length and frequency by surprisal.
@@
 
-To assess the shape of the RT-surprisal relationship, we then fitted generalized additive models (GAMs).
+To assess the shape of the RT-surprisal relationship, we then fit generalized additive models (GAMs).
 For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.
@@
 All 4 models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data.
-To determine the goodness of fit of a model in which word probability effects on RT are taken to be linear in surprisal, we also fitted GAM models with both parametric linear and nonparametric non-linear terms for surprisal; for all but the n-gram model, these analyses supported a linear effect of surprisal (Appendix D).
+To determine the goodness of fit of a model in which word probability effects on RT are taken to be linear in surprisal, we also fit GAM models with both parametric linear and nonparametric non-linear terms for surprisal; for all but the 5-gram model, these analyses supported a linear effect of surprisal (Appendix D).
 
@@
 
-(ref:coeffs-spr) Point estimates and 95% credible intervals for coefficients predicted by fitted Bayesian regression models predicting SPR RT.
+(ref:coeffs-spr) Point estimates and 95% confidence intervals (+/- 1.97 standard error) for coefficients predicted by fitted regression models predicting SPR RT.
 Units are in ms.
@@
 
-Whereas Maze showed surprisal effects in the 10 to 25 ms/bit range and length effects 15 to 20 ms/character range, SPR effects are about 1-2 ms per bit or character.
+Whereas Maze showed surprisal effects in the 10 to 25 ms/bit range and length effects in the 15 to 20 ms/character range, SPR effects are about 1-2 ms per bit or character.
 This difference in effect size is disproportionate to the overall speed of the methods; the predicted intercept for the Maze task was roughly 880 ms and for SPR was roughly 360 ms.
@@
 
-As shown in Table \@ref(tab:maze-compare), GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and TXL and a little complementary information over GPT-2.
-TXL provides a lot over 5-gram, and 5-gram provides little over any model.
-The single-model measures of log likelihood confirm this hierarchy, as GPT-2 is better than GRNN is better than TXL is better than 5-gram.
+As shown in Table \@ref(tab:maze-compare), GPT-2 provides a lot of additional predictive value over each other model, GRNN provides a lot over 5-gram and Transformer-XL and a little complementary information over GPT-2.
+Transformer-XL provides a lot over 5-gram, and 5-gram provides little over any model.
+The single-model measures of log likelihood confirm this hierarchy, as GPT-2 is better than GRNN is better than Transformer-XL is better than 5-gram.
 
@@
 We followed the same process for the SPR data with results shown in Table \@ref(tab:spr-compare).
-For SPR, GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN.
-In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small.
+For SPR, GPT-2 and 5-gram models contain some value over each other model, which is less clear for Transformer-XL and GRNN.
+In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than Transformer-XL, although differences are small.
 The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020, and contrasts with the Maze results, where the 5-gram model had the worst fit and did not provide additional predictive value over the other models.
-The fact that the neural language models, which capture deeper features of linguistic structure  [@huSystematicAssessmentSyntactic2020,@wilcox-etal:2022-using-computational-models], dominate the more superficial $n$-gram models for Maze but not SPR suggests that Maze RTs reflect richer language structure-related processes during real-time comprehension than do SPR RTs.
+The fact that the neural language models, which capture deeper features of linguistic structure  [@huSystematicAssessmentSyntactic2020; @wilcox-etal:2022-using-computational-models], dominate the more superficial 5-gram models for Maze but not SPR suggests that Maze RTs reflect richer language structure-related processes during real-time comprehension than do SPR RTs.
 
@@
 Fourth, we compared the spillover profiles between Maze and SPR.
-For Maze, we found large effects of the current words surprisal and length, which dwarfed any spillover effects from previous word predictors.
-In contrast, for SPR, we found effects of roughly equal sizes from the current and previous words.[^Furthermore, the typical spillover profile for SPR data may be worse than suggested by the Natural Stories corpus SPR data: for example, @smithEffectWordPredictability2013, found that most of a word's surprisal effect showed up only one to two words downstream.] Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus, and thus generating less noisy data.
+For Maze, we found large effects of the current word's surprisal and length, which dwarfed any spillover effects from previous word predictors.
+In contrast, for SPR, we found effects of roughly equal sizes from the current and previous words.
+^[Furthermore, the typical spillover profile for SPR data may be worse than suggested by the Natural Stories corpus SPR data: for example, @smithEffectWordPredictability2013 found that most of a word's surprisal effect showed up only one to two words downstream.] Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus, and thus generating less noisy data.
 
@@
 This difference suggests that how well a language model predicts human RTs may depend on task.
-Maze RTs were by far best predicted by language models that capture rich features of natural language grammatical structure, unlike SPR RTs which were nearly equally well predicted by more superficial $n$-gram models.
+Maze RTs were by far best predicted by language models that capture rich features of natural language grammatical structure, unlike SPR RTs which were nearly equally well predicted by more superficial 5-gram models.
 These findings further add to the evidence that the Maze task is favorable for RT-based investigations of underlying linguistic processing in the human mind.
@@
 
-Overall, A-maze has excellent good localization, although some  models showed small but statistically reliable effects of the past word.
+Overall, A-maze has excellent localization, although some  models showed small but statistically reliable effects of the past word.
 On the whole, our results support the idea that Maze forces language processing to be close to word-by-word, and thus the Maze task can be used under the assumption that the RT of a word primarily reflects its own properties and not those of earlier words.
@@
 In discussing this work, we received the suggestion that one way to further disincentivize random clicking would be to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms) before correcting their mistake.
-This delay would make randomly hitting buttons slower than doing the task as intended.
+This delay would make randomly hitting buttons slower than doing the task as intended, and we have made delaying after wrong presses an option in the error-correction Maze implementation at https://github.com/vboyce/Ibex-with-Maze.
 
@@
 For the 5-gram estimate, there is overwhelming evidence for the linear term, and some evidence for a nonlinearity as well.
-Consulting \@ref(fig:gam) shows that this nonlinearity takes the form of the surprisal effect dwindling to zero in the sparse tail of high-surprisal words.
+Consulting Figure \@ref(fig:gam) shows that this nonlinearity takes the form of the surprisal effect dwindling to zero in the sparse tail of high-surprisal words.
 This nonlinearity is plausibly due to the measurement error (high variance) of using counts to estimate very low multinomial probabilities.
@@
 
-<!--# Ethics and consent (optional) {#ethics-and-consent-optional .unnumbered}
-
-Research involving human subjects, human material, or human data, must
-have been performed in accordance with the Declaration of Helsinki.
-Where applicable, the studies must have been approved by an appropriate
-ethics committee. In the published paper, the authors should include a
-statement detailing this approval, including the name of the ethics
-committee and reference number of the approval. The identity of the
-research subject(s) should be anonymised whenever possible. For research
-involving human subjects, informed consent to participate in the study
-must be obtained from participants (or their legal guardian).
--->
-<!--# Funding information (optional) {#funding-information-optional .unnumbered}
-
-Should the research have received a funding grant, then the grant
-provider and grant number can be detailed in this section.-->
+# Ethics and consent {#ethics-and-consent-optional .unnumbered}
+
+This research was approved by MIT’s Committee on the Use of Humans as Experimental Subjects and run under protocol number 1605559077.
+
+# Funding information {#funding-information-optional .unnumbered}
+
+RPL acknowledges support from NSF grant BCS-2121074, NIH grant U01-NS121471, and the MIT–IBM Artificial Intelligence Research Lab.
 
@@
 # Authors' contributions {#contrib .unnumbered}
-Veronica Boyce contributed Conceptualization, Formal Analysis, Investigation, Methodology, Software, and Writing - Original Draft Preparation.
-Roger Levy contributed Conceptualization, Formal Analysis, Funding Acquisition, Methodology, Supervision, and Writing - Review & Editing.
+VB contributed Conceptualization, Formal Analysis, Investigation, Methodology, Software, and Writing - Original Draft Preparation.
+RPL contributed Conceptualization, Formal Analysis, Funding Acquisition, Methodology, Supervision, and Writing - Review & Editing.

### 2022-08-24 Papers/Paper/manuscript.Rmd
futzing to get word count

@@
 # Specify keywords here:
-keywords: A-Maze, self-paced reading, incremental processing, reading time measure, surprisal, naturalistic text,
+keywords: A-Maze, self-paced reading, incremental processing, surprisal, naturalistic text
 # To add the word count, uncomment the following option and replace the number
 # with the document word count.
-# wordcount: 1000
+wordcount: 9061
 output: 
@@
     number_sections: yes
+ #papaja::apa6_pdf
 ---

### 2022-08-24 Papers/Paper/manuscript.Rmd
Roger's edits

@@
 @boyceMazeMadeEasy2020 introduced several implementational innovations that made it easier for researchers to use Maze, and showed that for several controlled syntactic processing contrasts [@witzelComparisonsOnlineReading2012a] Maze offers better statistical power than self-paced reading, the other word-by-word response time method easy to use remotely.
-Maze has since had rapid uptake in the language processing community [@chacon2021limits; @ungerer2021using; @orth2022processing] (other cites).
+Maze has since had rapid uptake in the language processing community [@chacon2021limits; @ungerer2021using; @orth2022processing;@van-lieburg-etal:2022-using-the-maze-task;@levinson:2022-beyond-surprising].
 
@@
 A variety of pre-trained models exist, with varied internal architectures and training methods, but all of them generate measures of predictability.
-Predictability is often measured in bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time, etc.).
+Predictability is often measured in bits of surprisal, which is the negative log probability of a word (@hale:2001; @levy:2008; 1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time, etc.).
 
@@
 
-Due to previous reports of a length--frequency interaction in RT measures [@kliegl-etal:2006], before pursuing our primary question of the functional form of the surprisal--RT relationship, as an exploratory measure we fit generalized additive models (GAMs) with not only the main effects but also the two-way interactions between surprisal, length, and frequency, for the current word and for the previous word.
+To assess the shape of the RT-surprisal relationship, we then fit generalized additive models (GAMs).[^14] For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.
+We used surprisals generated by 4 different language models for robustness.
+(See Methods for details on language models, exclusions, and model fit.)
+
+[^14]: Due to previous reports of a length--frequency interaction in RT measures [@kliegl-etal:2006], before pursuing our primary question of the functional form of the surprisal--RT relationship, as an exploratory measure we fit generalized additive models (GAMs) with not only the main effects but also the two-way interactions between surprisal, length, and frequency, for the current word and for the previous word.
 This analysis revealed significant effects of current-word and previous-word surprisal, current-word and previous-word length, and significant interactions of current-word frequency by length and frequency by surprisal.
-The other main effects and interactions did not reach statistical significance.^[These are results from `mgcv`'s `summary()`; the $p$-values are approximate.] Appendix C provides tables and plots of these effects and interactions for GPT-2.
+The other main effects and interactions did not reach statistical significance.
+(These are results from `mgcv`'s `summary()`; the $p$-values are approximate.) Appendix C provides tables and plots of these effects and interactions for GPT-2.
 The interactions can be summarized as long low-frequency words and surprising, high-frequency words as having especially long RTs; and surprising, low-frequency words as having shorter RTs than would otherwise be predicted.
@@
 For simplicity we therefore set aside the interaction terms involving surprisal for the remainder of this analysis.
-
-To assess the shape of the RT-surprisal relationship, we then fit generalized additive models (GAMs).
-For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded.
-We used surprisals generated by 4 different language models for robustness.
-(See Methods for details on language models, exclusions, and model fit.)
 

### 2022-08-25 Papers/Paper/manuscript.Rmd
update word count
now with anonymity

@@
 shorttitle        : "A-maze of Natural Stories"
-shortauthors: Veronica Boyce & Roger Levy
+shortauthors: "" #Veronica Boyce & Roger Levy
 author: 
-  - name          : "Veronica Boyce"
-    affiliation   : "Stanford University"
+  - name          : "Anonymous submission" #Veronica Boyce"
+    affiliation   : "" #Stanford University"
     corresponding : yes    # Define only one corresponding author
-    email         : "vboyce@stanford.edu"
-
-  - name          : "Roger P. Levy"
-    affiliation   : "Massachusetts Institute of Technology"
-
-affiliation:
-  - id            : "1"
-    institution   : "Stanford University"
-  - id            : "2"
-    institution   : "Massachusetts Institute of Technology"
+    email         : "" #vboyce@stanford.edu"
+
+ # - name          : "Roger P. Levy"
+ #   affiliation   : "Massachusetts Institute of Technology"
+
 header-includes:
@@
 # with the document word count.
-wordcount: 9061
+wordcount: 9069
 output: 
@@
     number_sections: yes
- #papaja::apa6_pdf
+# papaja::apa6_pdf
 ---
@@
 We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the Maze paradigm.
-The materials, data, and analysis code are all available at https://github.com/vboyce/natural-stories-maze.
+The materials, data, and analysis code are all available at https://osf.io/gfs87/?view_only=c04ea31c7ea34d9ba84eb106f88b564a <!--https://github.com/vboyce/natural-stories-maze-->.
 
@@
 When a participant makes an error, they see an error message and must try again to select the correct option, before continuing the sentence as normal.
-We make error-correction Maze available as an option in a modification of the Ibex Maze implementation introduced in @boyceMazeMadeEasy2020 (https://github.com/vboyce/Ibex-with-Maze).
+We make error-correction Maze available as an option in a modification of the Ibex Maze implementation introduced in @boyceMazeMadeEasy2020 (link obscured for anonymity) <!--(https://github.com/vboyce/Ibex-with-Maze)-->.
 The code records both the RT to the first click and also the total RT until the correct answer is selected as separate values.
@@
 To generate distractors, we first split the corpora up into sentences, and then ran the sentences through the A-maze generation process.
-We used an updated version of the codebase from @boyceMazeMadeEasy2020 which had the capability to match the greater variety of punctuation present in the Natural Stories corpus (updated auto-generation code at https://github.com/vboyce/Maze).
+We used an updated version of the codebase from @boyceMazeMadeEasy2020 which had the capability to match the greater variety of punctuation present in the Natural Stories corpus (updated auto-generation code at link obscured for anonymity <!--https://github.com/vboyce/Maze-->).
 We took the auto-generated distractors as they were, without checking  for quality.
@@
 In discussing this work, we received the suggestion that one way to further disincentivize random clicking would be to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms) before correcting their mistake.
-This delay would make randomly hitting buttons slower than doing the task as intended, and we have made delaying after wrong presses an option in the error-correction Maze implementation at https://github.com/vboyce/Ibex-with-Maze.
+This delay would make randomly hitting buttons slower than doing the task as intended, and we have made delaying after wrong presses an option in the error-correction Maze implementation at link obscured for anonymity <!--https://github.com/vboyce/Ibex-with-Maze-->.
 

### 2022-08-25 Papers/Cover_letter/glossapx_cover_1.tex
draft cover letter

\documentclass{stanfordletter}
%\makelabels
\begin{document}
	\name{Veronica Boyce}
	\signature{Veronica Boyce \\ Roger Levy}

	\begin{letter}{Fernanda Ferreira and Brian Dillon \\ Co-Editors-in-Chief \\ Glossa Psycholinguistics}

		\opening{Dear Professor Ferreira and Professor Dillon,} 
		 This submission contains the manuscript ``A-maze of Natural Stories: Comprehension and surprisal in the Maze task'' which we are submitting to Glossa Psycholinguistics for publication. This reports a variation on the Maze task designed to accommodate longer naturalistic texts and an experiment using this method where participants read passages from the Natural Stories corpus (Futrell et al 2020). We present an analysis of the reading time characteristics of the data, focusing on surprisal effects, and we compare to self-paced-reading data collected on the same materials by Futrell et al 2020.

		 Researchers with relevant expertise and interest who therefore might be appropriate reviewers include Jeffrey Witzel, Kenneth Forster, Viktor Kuperman, Dustin Chac{\'o}n, and Masaya Yoshida.

		 This manuscript is not under review at any other venue. Please let us know if you need any further information in
		 connection with this submission. We look forward to hearing from you regarding the manuscript!
		\closing{Sincerely,}

	\end{letter}

\end{document}

### 2022-08-25 Papers/Cover_letter/glossapx_cover_1.tex
Roger's edits

@@
 
-		\opening{Dear Professor Ferreira and Professor Dillon,} 
-		 This submission contains the manuscript ``A-maze of Natural Stories: Comprehension and surprisal in the Maze task'' which we are submitting to Glossa Psycholinguistics for publication.
-This reports a variation on the Maze task designed to accommodate longer naturalistic texts and an experiment using this method where participants read passages from the Natural Stories corpus (Futrell et al 2020).
-We present an analysis of the reading time characteristics of the data, focusing on surprisal effects, and we compare to self-paced-reading data collected on the same materials by Futrell et al 2020.
+          \opening{Dear Professor Ferreira and Professor Dillon,} This
+          submission contains the manuscript ``A-maze of Natural
+          Stories: Comprehension and surprisal in the Maze task''
+          which we are submitting to Glossa Psycholinguistics for
+          publication. The manuscript reports a variation on the Maze
+          task (Forster et al., 2009; Boyce et al., 2020) designed to
+          accommodate longer naturalistic texts and an experiment
+          using this method where participants read passages from the
+          Natural Stories corpus (Futrell et al 2020). We present an
+          analysis of the reading time characteristics of the data,
+          focusing on surprisal effects, and we compare to
+          self-paced-reading data collected on the same materials by
+          Futrell et al 2020. We find that in our new Maze task
+          variant participants can successfully read and comprehend
+          the longer Natural Stories corpus texts, that the task
+          yields RTs that are better localized and more sensitive than
+          self-paced reading RTs, and that the same linear effect of
+          word surprisal on RTs previously documented in self-paced
+          reading and eye tracking of naturalistic texts is obtained
+          in our maze RT data. We believe this paper will be of broad
+          methodological relevance, and also potential theoretical
+          interest, for the psycholinguistics community.
 
-		 Researchers with relevant expertise and interest who therefore might be appropriate reviewers include Jeffrey Witzel, Kenneth Forster, Viktor Kuperman, Dustin Chac{\'o}n, and Masaya Yoshida.
+          Researchers with relevant expertise and interest who
+          therefore might be appropriate reviewers include Jeffrey
+          Witzel, Kenneth Forster, Viktor Kuperman, Dustin Chac{\'o}n,
+          and Masaya Yoshida.
 
-		 This manuscript is not under review at any other venue. Please let us know if you need any further information in
-		 connection with this submission. We look forward to hearing from you regarding the manuscript!
+          This manuscript is not under review at any other
+          venue. Please let us know if you need any further
+          information in connection with this submission. We look
+          forward to hearing from you regarding the manuscript!
           \closing{Sincerely,}

### 2022-10-26 Papers/Paper/manuscript.Rmd
Roger implemented edits

@@
 A word can be unpredictable for a variety of reasons: it could be low frequency, semantically unexpected, the start of a low-frequency syntactic construction, or a word that disambiguates prior words to a less common parse.
-Many targeted effects of interest are essentially looking at specific features that contribute to how predictable or unpredictable a word is.
-Thus incremental processing methods that are sensitive to predictability are useful for testing linguistic theories that make predictions about what words are unexpected.
+Many targeted effects of interest can thus be potentially accomodated theoretically as specific features that influence word predictability.[^Of course, not all effects can necessarily be reduced to word predictability effects, and effects that *cannot* be reduced to word predictability may be of particular theoretical interest.
+Candidates include, for example, memory-based effects [@lewis-etal:2006; @levy-fedorenko-gibson:2013jml], noisy-channel error identification [@levy-etal:2009pnas], and the magnitude of processing difficulty in garden-path resolution [@van-schijndel-linzen-2021:single-stage-models; @wilcoxTargetedAssessmentIncremental2021].] Thus incremental processing methods that are sensitive to predictability are useful for testing linguistic theories that make predictions about what words are unexpected.
 
@@
 The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020, and contrasts with the Maze results, where the 5-gram model had the worst fit and did not provide additional predictive value over the other models.
-The fact that the neural language models, which capture deeper features of linguistic structure  [@huSystematicAssessmentSyntactic2020; @wilcox-etal:2022-using-computational-models], dominate the more superficial 5-gram models for Maze but not SPR suggests that Maze RTs reflect richer language structure-related processes during real-time comprehension than do SPR RTs.
+While the nature of the generalizations made by these neural network-based models are not fully understood, controlled tests have suggested that their next-word predictions often reflect deeper features of linguistic structure [@warstadt-etal:2020-BLiMP; @huSystematicAssessmentSyntactic2020], such as subject--verb agreement [@marvin-linzen:2018-targeted] and wh-dependencies [@wilcox-etal:2022-using-computational-models], and are sensitive over longer context windows, than n-gram models.
+The fact that the neural language models, dominate the 5-gram models for Maze but not SPR thus suggests that Maze RTs may be more sensitive than SPR RTs to richer language structure-related processes during real-time comprehension.
 
@@
 This difference suggests that how well a language model predicts human RTs may depend on task.
-Maze RTs were by far best predicted by language models that capture rich features of natural language grammatical structure, unlike SPR RTs which were nearly equally well predicted by more superficial 5-gram models.
+Maze RTs were by far best predicted by neural network language models, whereas SPR RTs were predicted nearly as well by 5-gram models.
+Our understanding of the linguistic generalization capabilities and performance of these neural network models is still limited, but controlled tests in the NLP literature that analyze their behavior on classic psycholinguistics paradigms [@linzen-etal:2016tacl; @futrell-etal:2019-neural-language-models] suggest more human-like performance than n-gram models are capable of.
 These findings further add to the evidence that the Maze task is favorable for RT-based investigations of underlying linguistic processing in the human mind.

### 2022-10-26 Papers/Paper/manuscript.Rmd
first pass at edits for R&R

@@
 
-However, there is increasing interest in collecting data during comprehension of more naturalistic materials such as stories and news articles [@demberg-keller:2008; @lukeLimitsLexicalPrediction2016; @futrellNaturalStoriesCorpus2020], which offer potentially improved ecological validity and larger scale data in comparison with repeated presentation of isolated sentences out of context.
+There is increasing interest in collecting data during comprehension of more naturalistic materials such as stories and news articles [@demberg-keller:2008; @lukeLimitsLexicalPrediction2016; @futrellNaturalStoriesCorpus2020], which offer potentially improved ecological validity and larger scale data in comparison with repeated presentation of isolated sentences out of context.
 These more naturalistic materials require maintaining and integrating discourse dependencies and other types of information over longer stretches of time and linguistic material.
@@
 
-For instance, competing theories about why certain types of object-extracted relative clauses, like *the lawyer that the banker irritated*, are harder to understand than the corresponding subject-extracted relative clauses, like *the lawyer that irritated the banker*, make different predictions about which words are the loci of the overall difficulty and slower RTs associated with object relatives [@grodnerConsequencesSerialNature2005; @staubEyeMovementsProcessing2010; @traxlerProcessingSubjectObject2002].
+For instance, competing theories about why certain types of object-extracted relative clauses, like *the lawyer that the banker irritated*, are harder to understand than the corresponding subject-extracted relative clauses, like *the lawyer that irritated the banker*, make different predictions about which words are the loci of the overall difficulty and slower RTs associated with object relatives  [@traxlerProcessingSubjectObject2002].
+Dependency-locality theory [@grodnerConsequencesSerialNature2005] predicts that the locus of difficulty on object relatives is on the verb of the embedded clause (*irritated* in *the lawyer that the banker irritated*).
+In contrast, surprisal theory predicts the locus of difficulty will be on the article *the* at the start of the embedded clause [@staubEyeMovementsProcessing2010].
 RT measures can potentially also inform theories about the time course of processing (i.e.
@@
 Even an imprecise RT measure will determine that the same set of words in a different order took longer to read at a sentence level.
-However, many language processing theories make specific (and contrasting) theories about which words in a sentence are harder to process.
+However, many language processing theories make specific (and contrasting) predictions about which words in a sentence are harder to process.
 To adjudicate among these theories, we want methods that are *well-localized*, so it is easy to determine which word is responsible for an observed RT slow-down.
-Ideally, longer RT on a word would be an indication of that word's increased difficulty, and not the lingering signal of a prior word's increased difficulty.
+Ideally, a longer RT on a word would be an indication of that word's increased difficulty, and not the lingering signal of a prior word's increased difficulty.
 When the signal isn't localized, advanced analysis techniques may be required to disentangle the slow-downs [@shainDeconvolutionalTimeSeries2018].
@@
 In eye-tracking, participants read a text on a screen naturally, while their saccadic eye movements are recorded on a computer-connected camera that is calibrated so that the researcher can reconstruct with high precision where the participant's gaze falls on the screen at all times [@raynerEyeMovementsReading1998].
-These eye movements can be used to reconstruct various position-specific reading time measures such as *gaze duration* (the total amount of time the eyes spend on a word the first time it is fixated, or zero if the eye skipped the word the first time it was approached from the left) and *total viewing time* (the total amount of time that the word is fixated).
+These eye movements can be used to reconstruct various position-specific reading time measures such as *gaze duration* (the total amount of time the eyes spend on a word the first time it is fixated before saccading to a later word) and *total viewing time* (the total amount of time that the word is fixated).
+If the eyes skipped the word the first time it was approached to the left, the trial is generally excluded.
 Eye tracking data collected with state-of-the-art high-precision recording equipment offers relatively good signal-to-noise ratio, but the difficulty presented by a word can still *spill over* into reading measures on subsequent words, a dynamic that can make it hard to isolate the source of an effect of potential theoretical interest [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982].
+Short words such as articles and pronouns are often not fixated directly which makes it harder to study the processing of these words with eye-tracking.
 Additionally, the equipment is expensive and data collection is laborious and must occur in-lab.
@@
 
+The Maze task has much less face-validity than eye-tracking or even SPR; repeated forced-choice selections does not look very similar to normal reading.
+Despite this, @forsterMazeTaskMeasuring2009 report that "At a phenomenological level, participants typically report that they feel as if they are reading the sentence relatively
+naturally and that the correct alternative seems to “leap out” at them, so that they do not have to inspect the incorrect
+alternative very carefully, if at all." This suggests that the Maze task is not about conscious comparison between two potential continuations and instead relies on the same language processing facilities shared by other methods.
+Maze is not trying to study the process of normal reading, but instead to get underlying language processing, using an artificial method to slow down and localize language processing.
+
 The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitous words as distractors for each spot of each sentence.
@@
 A variety of pre-trained models exist, with varied internal architectures and training methods, but all of them generate measures of predictability.
-Predictability is often measured in bits of surprisal, which is the negative log probability of a word (@hale:2001; @levy:2008; 1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time, etc.).
+Predictability is often measured in bits of surprisal, which is the negative log probability of a word [@hale:2001, @levy:2008].
+1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time, etc.
 
@@
 If Maze RTs reflect the same processing as other methods, we expect to find a similar linear relationship with surprisal.
+
+Frequency is often found to be a predictor of reading time in SPR and eye-tracking experiments.
+It is unclear why we find frequency effects on RT when surprisal effects are accounted for.
+In their discussion, @smith-levy:2008 posit that the role of frequency could diminish or disappear as measurements of surprisal improve (as they have over the past decade).
+Additionally, @smith-levy:2008 hypothesize that the effects of frequency (in addition to those of surprisal) may be the result of quick, heuristic responses to the words based on the more easily available unigram frequency before more fine-grained context-sensitive surprisal is available.
+An alternative is that frequency effects don't actually exist, but are an artifact of how RT data is analysed [@shainLargescaleStudyEffects2019].
 
@@
 
-We used the Natural Stories corpus [@futrellNaturalStoriesCorpus2020], which consists of 10 passages of roughly 1000 words each which are designed to read fluently to native speakers.
-At the same time, the passages contain copious punctuation, quoted speech, proper nouns, and low frequency grammatical constructions.
+We used the Natural Stories corpus [@futrellNaturalStoriesCorpus2020], a set of 10 passages designed to read fluently to native speakers.
+Each passage is roughly 1000 words long.
+The passages contain copious punctuation, quoted speech, proper nouns, and low frequency grammatical constructions.
 The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we used to assess comprehension.
@@
 
-# Methods
-
-We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the Maze paradigm.
-The materials, data, and analysis code are all available at https://osf.io/gfs87/?view_only=c04ea31c7ea34d9ba84eb106f88b564a <!--https://github.com/vboyce/natural-stories-maze-->.
-
-## Task: Error-correction Maze
+# Error-correct Maze: Adapting the Maze paradigm for long passages
 
@@
 This error-correction presentation is used here with A-maze, but could also be used with G-maze or I-maze.
+
+# Methods
+
+We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the Maze paradigm.
+The materials, data, and analysis code are all available at https://osf.io/gfs87/?view_only=c04ea31c7ea34d9ba84eb106f88b564a <!--https://github.com/vboyce/natural-stories-maze-->.
 
@@
 Participants first gave their informed consent and saw task instructions.
-Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading one main story in the A-maze task.
+Then they read a short practice story in the Maze paradigm and answered 2 binary-choice practice comprehension questions, before reading one main story in the error-correction A-maze task.
 After the story, they answered 6 comprehension questions, commented on their experience, answered optional demographic questions, were debriefed, and were given a code to enter for payment.
@@
 We compared the correlations between the Maze and SPR RTs to within-Maze and within-SPR correlations.
-For Maze, within each story, we randomly split subjects into two halves.
+For Maze, within each story, we split the data in half, randomly assigning subjects into two equal groups.
 Within each half, we calculated a per-word average RT for each word and then a per-sentence average RT across word averages.
@@
 For this comparison, we downsampled the SPR data choosing a number of participants equal to the number we have for Maze to avoid differences due to dataset size.
-We then used the same procedure to get a within-SPR correlation.
+We then used the same split-half procedure to get a within-SPR correlation.
 For between Maze-SPR correlation, we took the average correlation across each of the 4 pairs of Maze half and SPR half.
@@
 Our analytic questions required multiple modeling approaches.
-To look at the functional form of the relationship between surprisal and RT data, we fit Generalized Additive Models (GAMs) to allow for non-linear relationships.
+To look at the functional form of the relationship between surprisal and RT data, we fit Generalized Additive Models (GAMs) to allow for non-linear relationships [@wood:2006GAMs].
 GAM model summaries can be harder to interpret than those for linear models, so to measure effect sizes and assess spillover, we used linear mixed models.
@@
 
-We excluded the first word of every sentence because it had an x-x-x distractor, leaving `r nrow(not_first)` words.
+In the Maze task, the first word of every sentence is paired with a nonce (x-x-x) distractor rather than a real word (as there is no context to use to distinguish between real words); due to this difference, we excluded the first word of every sentence, leaving `r nrow(not_first)` words.
 We excluded words for which we didn't have surprisal or frequency information, leaving `r nrow(not_na)` words.
@@
 
-## How do A-maze and SPR compare in power and reliability?
+## How do A-maze and SPR compare in reliability?
 
@@
 Across all models, there were consistent large effects of length and surprisal at the current word, but minimal effects of frequency.
-The lack of frequency effects is unexpected, but consistent with @shainLargescaleStudyEffects2019.
+This lack of frequency effects is inconsistent with the results usually reported for SPR and eye-tracking.
 There was a small interaction between surprisal and length at the current word.
@@
 Maze and SPR pick up on similar features in words, as shown by the high correlations between Maze and SPR RTs on the sentence level.
+The correlation within Maze is higher than Maze to SPR correlation or SPR-SPR correlations which is evidence that Maze is less noisy that SPR and may pick up on deeper linguistic features than SPR.
 
@@
 ^[Furthermore, the typical spillover profile for SPR data may be worse than suggested by the Natural Stories corpus SPR data: for example, @smithEffectWordPredictability2013 found that most of a word's surprisal effect showed up only one to two words downstream.] Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus, and thus generating less noisy data.
+We do not find frequency effects on the Maze data, but we do on the SPR data.
+This could be explained if frequency effects are a first rough approximation of in-context predictability, before the fuller context-sensitive surprisal information is available.
+In this case, faster methods like eye-tracking and SPR would show frequency effects (in addition to surprisal), but slower methods like Maze would not as the additional demands slow down the response, allowing more contextual information to be used.
+While this is a difference between Maze and other incremental processing methods, we do not consider it a flaw, since there is not strong theoretical justification for the presence of frequency effects (in addition to surprisal effects) in language processing.
+These differences do point to the importance of understanding task demands of different incremental processing methods.
 
@@
 On the whole, our results support the idea that Maze forces language processing to be close to word-by-word, and thus the Maze task can be used under the assumption that the RT of a word primarily reflects its own properties and not those of earlier words.
+Correlations between Maze and SPR indicate that Maze is picking up on similar patterns, but with less noise, than SPR.
 
@@
 
-Compared to traditional Maze, error-correction Maze reduces perverse (from the experimenter's point of view) incentives to complete the task as quickly as possible.
+Compared to traditional Maze,in error-correction Maze, participants' incentives to finish quickly are in less conflict with the experimenter's desire that participants do the task as intended.
 However, even with error-correction Maze, clicking randomly is still likely faster than doing the task.

### 2022-10-26 Papers/Paper/manuscript.Rmd
fix footnote

@@
 A word can be unpredictable for a variety of reasons: it could be low frequency, semantically unexpected, the start of a low-frequency syntactic construction, or a word that disambiguates prior words to a less common parse.
-Many targeted effects of interest can thus be potentially accomodated theoretically as specific features that influence word predictability.[^Of course, not all effects can necessarily be reduced to word predictability effects, and effects that *cannot* be reduced to word predictability may be of particular theoretical interest.
+Many targeted effects of interest can thus be potentially accommodated theoretically as specific features that influence word predictability.
+^[Of course, not all effects can necessarily be reduced to word predictability effects, and effects that *cannot* be reduced to word predictability may be of particular theoretical interest.
 Candidates include, for example, memory-based effects [@lewis-etal:2006; @levy-fedorenko-gibson:2013jml], noisy-channel error identification [@levy-etal:2009pnas], and the magnitude of processing difficulty in garden-path resolution [@van-schijndel-linzen-2021:single-stage-models; @wilcoxTargetedAssessmentIncremental2021].] Thus incremental processing methods that are sensitive to predictability are useful for testing linguistic theories that make predictions about what words are unexpected.

### 2022-10-27 Papers/Paper/manuscript.Rmd
minor tweaks

@@
 If the eyes skipped the word the first time it was approached to the left, the trial is generally excluded.
-Eye tracking data collected with state-of-the-art high-precision recording equipment offers relatively good signal-to-noise ratio, but the difficulty presented by a word can still *spill over* into reading measures on subsequent words, a dynamic that can make it hard to isolate the source of an effect of potential theoretical interest [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982].
+Eye tracking data collected with state-of-the-art high-precision recording equipment offers relatively good signal-to-noise ratio, but the difficulty presented by a word can still *spill over* into reading measures on subsequent words, a dynamic that can make it hard to isolate the source of an effect of potential theoretical interest [@raynerEffectsFrequencyPredictability2004; @levy-etal:2009pnas; @frazierMakingCorrectingErrors1982].
 Short words such as articles and pronouns are often not fixated directly which makes it harder to study the processing of these words with eye-tracking.
@@
 
-# Error-correct Maze: Adapting the Maze paradigm for long passages
+# Error-correction Maze
 
@@
 
-We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the Maze paradigm.
+We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the error-correction Maze paradigm.
 The materials, data, and analysis code are all available at https://osf.io/gfs87/?view_only=c04ea31c7ea34d9ba84eb106f88b564a <!--https://github.com/vboyce/natural-stories-maze-->.
@@
 Our analytic questions required multiple modeling approaches.
-To look at the functional form of the relationship between surprisal and RT data, we fit Generalized Additive Models (GAMs) to allow for non-linear relationships [@wood:2006GAMs].
+To look at the functional form of the relationship between surprisal and RT data, we fit Generalized Additive Models (GAMs) to allow for non-linear relationships [@wood:2017GAMs].
 GAM model summaries can be harder to interpret than those for linear models, so to measure effect sizes and assess spillover, we used linear mixed models.
@@
 (ref:gam-grid) Generalized Additive Model main effects and two-way interactions among frequency, length, and surprisal for A-Maze reading of the Natural Stories corpus.
-Confidence bands do not take into account the uncertainty associated with `mgcv` hyperparameter estimation (@wood:2006GAMs).
+Confidence bands do not take into account the uncertainty associated with `mgcv` hyperparameter estimation [@wood:2017GAMs].
 
@@
 
-The `mgcv` package's implementation of Generalized Additive Models [@wood:2006GAMs] allows linear and nonparametric spline effects of the same continuous predictor to be entered simultaneously into a model.
+The `mgcv` package's implementation of Generalized Additive Models [@wood:2017GAMs] allows linear and nonparametric spline effects of the same continuous predictor to be entered simultaneously into a model.
 Doing so associates only the nonlinear part of the effect to the spline term, allowing for approximate statistical testing of the linear and non-linear components of the effect respectively.

### 2022-10-31 Papers/Paper/manuscript.Rmd
Roger in the middle of changes

--- 2022-10-26 Papers/Paper/manuscript.Rmd
@@
 
-The Maze task has much less face-validity than eye-tracking or even SPR; repeated forced-choice selections does not look very similar to normal reading.
+The Maze task has less face-validity than eye-tracking or even SPR; repeated forced-choice selections does not seem very similar to normal reading.
 Despite this, @forsterMazeTaskMeasuring2009 report that "At a phenomenological level, participants typically report that they feel as if they are reading the sentence relatively
 naturally and that the correct alternative seems to “leap out” at them, so that they do not have to inspect the incorrect
-alternative very carefully, if at all." This suggests that the Maze task is not about conscious comparison between two potential continuations and instead relies on the same language processing facilities shared by other methods.
-Maze is not trying to study the process of normal reading, but instead to get underlying language processing, using an artificial method to slow down and localize language processing.
-
-The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitous words as distractors for each spot of each sentence.
+alternative very carefully, if at all." This suggests that the Maze task may rely on the same language processing facilities tapped into by other reading methods.
+Thus, using Maze may not be the best paradigm for studying the process of normal reading, but may be perfectly good or even superior for getting at underlying language processing.
+
+However, G-maze materials are effort-intensive to construct because of the need to select infelicitous words as distractors for each spot of each sentence.
 This burdensome preparation may explain why the Maze task was not widely adopted.
@@
 
-Frequency is often found to be a predictor of reading time in SPR and eye-tracking experiments.
-It is unclear why we find frequency effects on RT when surprisal effects are accounted for.
-In their discussion, @smith-levy:2008 posit that the role of frequency could diminish or disappear as measurements of surprisal improve (as they have over the past decade).
-Additionally, @smith-levy:2008 hypothesize that the effects of frequency (in addition to those of surprisal) may be the result of quick, heuristic responses to the words based on the more easily available unigram frequency before more fine-grained context-sensitive surprisal is available.
+The role of word frequency is worth special note.
+Although the facilitative effect of word frequency on reading measures has been known for decades, it is theoretically unclear why this effect should exist in any model that is sensitive to conditional word probability:
+
+@smith-levy:2008 noted this puzzle and posited that the role of frequency could diminish or disappear as measurements of surprisal improve (as they have over the past decade), or alternatively that the effects of frequency (in addition to those of surprisal) may be the result of quick, heuristic responses to the words based on the more easily available unigram frequency before more fine-grained context-sensitive surprisal is available.
 An alternative is that frequency effects don't actually exist, but are an artifact of how RT data is analysed [@shainLargescaleStudyEffects2019].

### 2022-10-31 Papers/Paper/manuscript.Rmd
Roger's edits

@@
 Dependency-locality theory [@grodnerConsequencesSerialNature2005] predicts that the locus of difficulty on object relatives is on the verb of the embedded clause (*irritated* in *the lawyer that the banker irritated*).
-In contrast, surprisal theory predicts the locus of difficulty will be on the article *the* at the start of the embedded clause [@staubEyeMovementsProcessing2010].
+In contrast, surprisal theory predicts the locus of difficulty will be on the article *the* at the start of the embedded clause [@staubEyeMovementsProcessing2010; @vaniUsingInterpolatedMaze2021].
 RT measures can potentially also inform theories about the time course of processing (i.e.
@@
 The role of word frequency is worth special note.
-Although the facilitative effect of word frequency on reading measures has been known for decades, it is theoretically unclear why this effect should exist in any model that is sensitive to conditional word probability:
-
-@smith-levy:2008 noted this puzzle and posited that the role of frequency could diminish or disappear as measurements of surprisal improve (as they have over the past decade), or alternatively that the effects of frequency (in addition to those of surprisal) may be the result of quick, heuristic responses to the words based on the more easily available unigram frequency before more fine-grained context-sensitive surprisal is available.
-An alternative is that frequency effects don't actually exist, but are an artifact of how RT data is analysed [@shainLargescaleStudyEffects2019].
+Although the facilitative effect of word frequency on reading measures has been known for decades, this effect remains a theoretical puzzle.
+Sensitivity to a word's probability in context can be derived from any of a number of optimization principles [@smithEffectWordPredictability2013; @levy:2013sentenceProcessing], and word frequency is a context-insensitive estimate of word probability.
+However, it is not straightforward why this effect should exist in models in which a role is played by *contextually-conditioned* word probability, which reading and other language processing measures are known to be sensitive to: conditioning on context should render the context-insensitive probability estimate irrelevant for optimal processing.
+@smith-levy:2008 noted this puzzle and posited two alternative explanations.
+First is estimation error of conditional word probability in computationally implemented language models used at the time, in which case the role of frequency could diminish or disappear as measurements of surprisal improve (as they have over the past decade); consistent with this hypothesis is the recent report of @shainLargescaleStudyEffects2019 finding no word frequency effects in reading using more recent language models.
+Second, the effects of frequency (in addition to those of surprisal) could be the result of quick, heuristic responses to the words based on the more easily available unigram frequency before more fine-grained context-sensitive surprisal becomes available within real-time language processing mechanisms.
 
@@
   1. Do participants engage with these longer passages successfully with the A-maze task?
-  2. Is A-maze as powerful and reliable a method as SPR for these longer passages?
+  2. Is A-maze as sensitive a method as SPR for these longer passages?
   3. What is the functional form between word surprisal and RT for the A-maze task?
@@
 
-## How do A-maze and SPR compare in reliability?
+## How do A-maze and SPR compare in sensitivity?
 
@@
 
-Our second question was whether A-maze is reliable.
-To assess reliability, we conducted split-half comparisons looking at the correlations between and within SPR and A-maze.
+Our second question was a comparison of the sensitivity (the signal-to-noise ratio) of A-maze and SPR.
+To assess sensitivity, we conducted split-half comparisons looking at the correlations between and within SPR and A-maze.
 If the methods picked up on the same effects, we would expect them to be correlated, with sentences that took longer to read in one method also taking longer in the other.
@@
 SPR data is about as correlated with Maze as with another sample of SPR data which provides some evidence that Maze and SPR are measuring the same effects.
-The superior within-method split-half correlation we see for Maze relative to SPR, despite the smaller number of participants, suggests that it is the more powerful of the two methods (higher signal-to-noise ratio), consistent with the findings of @boyceMazeMadeEasy2020 for factorial experimental designs with isolated-sentence presentation.
+The superior within-method split-half correlation we see for Maze relative to SPR, despite the smaller number of participants, suggests that it is the more sensitive of the two methods (higher signal-to-noise ratio), consistent with the findings of @boyceMazeMadeEasy2020 for factorial experimental designs with isolated-sentence presentation.
 
@@
 Across all models, there were consistent large effects of length and surprisal at the current word, but minimal effects of frequency.
-This lack of frequency effects is inconsistent with the results usually reported for SPR and eye-tracking.
+This lack of frequency effects differs from the results usually reported for SPR and eye-tracking (though see @shainLargescaleStudyEffects2019).
 There was a small interaction between surprisal and length at the current word.
@@
 While the nature of the generalizations made by these neural network-based models are not fully understood, controlled tests have suggested that their next-word predictions often reflect deeper features of linguistic structure [@warstadt-etal:2020-BLiMP; @huSystematicAssessmentSyntactic2020], such as subject--verb agreement [@marvin-linzen:2018-targeted] and wh-dependencies [@wilcox-etal:2022-using-computational-models], and are sensitive over longer context windows, than n-gram models.
-The fact that the neural language models, dominate the 5-gram models for Maze but not SPR thus suggests that Maze RTs may be more sensitive than SPR RTs to richer language structure-related processes during real-time comprehension.
+The fact that the neural language models dominate the 5-gram models for Maze but not SPR thus suggests that Maze RTs may be more sensitive than SPR RTs to richer language structure-related processes during real-time comprehension.
 
@@
 Maze and SPR pick up on similar features in words, as shown by the high correlations between Maze and SPR RTs on the sentence level.
-The correlation within Maze is higher than Maze to SPR correlation or SPR-SPR correlations which is evidence that Maze is less noisy that SPR and may pick up on deeper linguistic features than SPR.
+The correlation within Maze is higher than Maze to SPR correlation or SPR-SPR correlations, which is evidence that Maze is less noisy than SPR.
 
@@
 In this case, faster methods like eye-tracking and SPR would show frequency effects (in addition to surprisal), but slower methods like Maze would not as the additional demands slow down the response, allowing more contextual information to be used.
-While this is a difference between Maze and other incremental processing methods, we do not consider it a flaw, since there is not strong theoretical justification for the presence of frequency effects (in addition to surprisal effects) in language processing.
-These differences do point to the importance of understanding task demands of different incremental processing methods.
+While this is a difference between Maze and other incremental processing methods, we do not consider it a flaw for Maze---indeed, for researchers interested in focusing on context-contingent language processing, it may suggest an advantage for the Maze task.
+Regardless, these differences highlight the importance of understanding task demands of different incremental processing methods.
 
@@
 Maze RTs were by far best predicted by neural network language models, whereas SPR RTs were predicted nearly as well by 5-gram models.
-Our understanding of the linguistic generalization capabilities and performance of these neural network models is still limited, but controlled tests in the NLP literature that analyze their behavior on classic psycholinguistics paradigms [@linzen-etal:2016tacl; @futrell-etal:2019-neural-language-models] suggest more human-like performance than n-gram models are capable of.
+Our understanding of the linguistic generalization capabilities and performance of these neural network models is still limited, and there are cases where they are known to make more superficial, non-human-like generalizations [@mccoy-etal-2019-right; @chaves-2020-dont], but controlled tests in the NLP literature that analyze their behavior on classic psycholinguistics paradigms [@linzen-etal:2016tacl; @futrell-etal:2019-neural-language-models; @warstadt-etal:2020-BLiMP; @wilcox-etal:2022-using-computational-models] suggest more human-like performance than n-gram models are capable of.
 These findings further add to the evidence that the Maze task is favorable for RT-based investigations of underlying linguistic processing in the human mind.
@@
 
-Overall, A-maze has excellent localization, although some models showed small but statistically reliable effects of the past word.
+Overall, A-maze has excellent localization, although some models showed small but statistically significant effects of the past word.
 On the whole, our results support the idea that Maze forces language processing to be close to word-by-word, and thus the Maze task can be used under the assumption that the RT of a word primarily reflects its own properties and not those of earlier words.
-Correlations between Maze and SPR indicate that Maze is picking up on similar patterns, but with less noise, than SPR.
+Correlation analysis between Maze and SPR suggest that Maze is picking up on many of the same patterns as does SPR, but with less noise.
 

### 2022-11-02 Papers/Paper/manuscript.Rmd
redo diff, word count
deanon version for psyarxiv

@@
 author: 
-  - name          : "Anonymous submission" #Veronica Boyce"
-    affiliation   : "" #Stanford University"
+  - name          : "Veronica Boyce"
+    affiliation   : "Stanford University"
     corresponding : yes    # Define only one corresponding author
-    email         : "" #vboyce@stanford.edu"
-
- # - name          : "Roger P. Levy"
- #   affiliation   : "Massachusetts Institute of Technology"
+    email         : "vboyce@stanford.edu"
+
+  - name          : "Roger P. Levy"
+    affiliation   : "Massachusetts Institute of Technology"
 
@@
 # with the document word count.
-wordcount: 9069
+wordcount: 9594
 output: 
@@
 We also analyze the resulting reaction time profiles and show that they provide strong signal regarding the probabilistic relationship between a word and the context in which it appears, and that the systematic linear relationship between word surprisal and response time observed in other reading paradigms [@smithEffectWordPredictability2013] also arises in the Maze task.
-<!-- Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing.
-Maze is a potentially promising method, and it's feasibility has recently increased due to the introduction of A-Maze [@boyceMazeMadeEasy2020].
--->
+
 In the remainder of the Introduction, we lay out the role of RT-based methods in theory testing, describe a few common methods, and review some key influences on reading time.
@@
 In fact, RTs and other measures that capture processing disruption vary substantially with the difficulty of words in their context below the level of conscious awareness as well, with millisecond scale differences in reading time between words.
-<!-- Occasionally we read an unexpected word, cannot make sense of it and have to read the sentence again.
-The words that cause these stumbles are the outliers on a distribution of words varying in how easy they are to process in context.
-Even when we don't notice a problem, less expected words still take longer to process as we (re)construct a mental model of the sentence.
-Fortunately for reading and unfortunately for studying language, this process of adjustment is very quick, with millisecond scale differences in reading time between words.
--->
 That is, the differential difficulty or processing load posed by various parts of a text is to a considerable extent *localizable* to specific words in their context.
@@
 
-<!-- #Behavioral methods that measure how long it takes to read each word in a sentence are referred to as incremental processing methods, and their dependent measures are called reading times or reaction times, both abbreviated RT.
--->
 The two most commonly used behavioral methods for studying incremental language processing during reading are tracking eye movements and self-paced reading.
@@
 When a participant makes an error, they see an error message and must try again to select the correct option, before continuing the sentence as normal.
-We make error-correction Maze available as an option in a modification of the Ibex Maze implementation introduced in @boyceMazeMadeEasy2020 (link obscured for anonymity) <!--(https://github.com/vboyce/Ibex-with-Maze)-->.
+We make error-correction Maze available as an option in a modification of the Ibex Maze implementation introduced in @boyceMazeMadeEasy2020 <!--(link obscured for anonymity)--> (https://github.com/vboyce/Ibex-with-Maze).
 The code records both the RT to the first click and also the total RT until the correct answer is selected as separate values.
@@
 We constructed A-maze distractors for the Natural Stories corpus [@futrellNaturalStoriesCorpus2020] and recruited 100 crowd-sourced participants to each read a story in the error-correction Maze paradigm.
-The materials, data, and analysis code are all available at https://osf.io/gfs87/?view_only=c04ea31c7ea34d9ba84eb106f88b564a <!--https://github.com/vboyce/natural-stories-maze-->.
+The materials, data, and analysis code are all available at <!--https://osf.io/gfs87/?view_only=c04ea31c7ea34d9ba84eb106f88b564a--> https://github.com/vboyce/natural-stories-maze.
 
@@
 To generate distractors, we first split the corpora up into sentences, and then ran the sentences through the A-maze generation process.
-We used an updated version of the codebase from @boyceMazeMadeEasy2020 which had the capability to match the greater variety of punctuation present in the Natural Stories corpus (updated auto-generation code at link obscured for anonymity <!--https://github.com/vboyce/Maze-->).
+We used an updated version of the codebase from @boyceMazeMadeEasy2020 which had the capability to match the greater variety of punctuation present in the Natural Stories corpus (updated auto-generation code at <!--link obscured for anonymity-->https://github.com/vboyce/Maze).
 We took the auto-generated distractors as they were, without checking  for quality.
@@
 In discussing this work, we received the suggestion that one way to further disincentivize random clicking would be to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms) before correcting their mistake.
-This delay would make randomly hitting buttons slower than doing the task as intended, and we have made delaying after wrong presses an option in the error-correction Maze implementation at link obscured for anonymity <!--https://github.com/vboyce/Ibex-with-Maze-->.
+This delay would make randomly hitting buttons slower than doing the task as intended, and we have made delaying after wrong presses an option in the error-correction Maze implementation at <!--link obscured for anonymity-->https://github.com/vboyce/Ibex-with-Maze.
 

### 2023-03-29 Papers/Paper/manuscript.Rmd
formatting fussing for journal

@@
 
-classoption: [times, biblatex]
+classoption: [times]
 #
@@
 bibliography: ["r-references.bib","refs.bib"]
-biblio-style: "apalike"
 link-citations: true
@@
 # Specify keywords here:
-keywords: A-Maze, self-paced reading, incremental processing, surprisal, naturalistic text
+keywords: A-maze, self-paced reading, incremental processing, surprisal, naturalistic text
 # To add the word count, uncomment the following option and replace the number
@@
 wordcount: 9594
+csl               : "apa7.csl"
 output: 
     bookdown::pdf_book:
-    base_format: rticles::glossa_article
-    number_sections: yes
-# papaja::apa6_pdf
+     base_format: rticles::glossapx_article
+ #papaja::apa6_pdf
 ---
@@
 Differential and localized processing difficulty means that different inputs in context present different processing demands during comprehension [@levy:2008].
-Due to incrementality these differential processing demands are, by and large, met relatively quickly by the mind once they are presented, and they can be measured in both brain [@kutas-hillyard:1980; @osterhout-holcomb:1992jml] and behavioral [@raynerEyeMovementsReading1998; @mitchell:2004online-methods] responses.
+Due to incrementality, these differential processing demands are, by and large, met relatively quickly by the mind once they are presented, and they can be measured in both brain [@kutas-hillyard:1980; @osterhout-holcomb:1992jml] and behavioral [@raynerEyeMovementsReading1998; @mitchell:2004online-methods] responses.
 These measurements often have low signal-to-noise ratio, and many methods require bringing participants into the lab and often require cumbersome equipment.
@@
 The Maze task [@freedman85; @forsterMazeTaskMeasuring2009], which involves collecting participants' response times in a repeated two-alternative forced-choice between a word that fits the preceding linguistic context and a distractor that doesn't, has recently been proposed as a high-sensitivity method that can easily be used remotely.
-@boyceMazeMadeEasy2020 introduced several implementational innovations that made it easier for researchers to use Maze, and showed that for several controlled syntactic processing contrasts [@witzelComparisonsOnlineReading2012a] Maze offers better statistical power than self-paced reading, the other word-by-word response time method easy to use remotely.
+@boyceMazeMadeEasy2020 introduced several implementational innovations that made it easier for researchers to use Maze, and showed that for several controlled syntactic processing contrasts [@witzelComparisonsOnlineReading2012a].
+Maze offers better statistical power than self-paced reading, the other word-by-word response time method easy to use remotely.
 Maze has since had rapid uptake in the language processing community [@chacon2021limits; @ungerer2021using; @orth2022processing;@van-lieburg-etal:2022-using-the-maze-task;@levinson:2022-beyond-surprising].
@@
 
-In the remainder of the Introduction, we lay out the role of RT-based methods in theory testing, describe a few common methods, and review some key influences on reading time.
+In the remainder of the introduction, we lay out the role of RT-based methods in theory testing, describe a few common methods, and review some key influences on reading time.
 We then proceed to present our modified "error-correction Maze" paradigm, our experiment, and the results of our analyses of the resulting data.
@@
 Psycholinguistic theories may differ in their ability to predict what is easy and what is hard.
-One of the most powerful methods for studying these differential difficulty effects is let the comprehender control the pace of presentation of the linguistic material, and to measure what she takes time on.
+One of the most powerful methods for studying these differential difficulty effects is to let the comprehender control the pace of presentation of the linguistic material, and to measure what she takes time on.
 For this purpose, taking measurements from experimental participants during reading, a widespread, highly practiced skill in diverse populations around the world, is of unparalleled value.
@@
 
-## Eye-tracking and Self-paced reading
+## Eye-tracking and self-paced reading
 
@@
 
-Self-paced reading (SPR; @mitchell:1984) is a somewhat less natural paradigm in which the participant manually controls the visual presentation of the text by pressing a button.
+Self-paced reading (SPR) is a somewhat less natural paradigm in which the participant manually controls the visual presentation of the text by pressing a button [@mitchell:1984].
 In its generally preferred variant, moving-window self-paced reading, words are revealed one at a time or one group at a time: every press of the button masks the currently presented word (group) and simultaneously reveals the next.
@@
 In theory, participants must fully integrate each word into the sentence in order to confidently select it, which may require mentally reparsing previous material in order to allow the integration and selection of a disambiguating word.
-@forsterMazeTaskMeasuring2009 call this need for full integration "forced incremental processing" to distinguish from other incremental processing methods where words can be passively read before later committing to a parse.
+@forsterMazeTaskMeasuring2009 call this need for full integration "forced incremental sentence processing" in their title (p.
+163) to distinguish from other incremental processing methods where words can be passively read before later committing to a parse.
 This idea of strong localization is supported by studies finding strongly localized effects for G-maze [@witzelComparisonsOnlineReading2012a; @boyceMazeMadeEasy2020].
@@
 naturally and that the correct alternative seems to “leap out” at them, so that they do not have to inspect the incorrect
-alternative very carefully, if at all." This suggests that the Maze task may rely on the same language processing facilities tapped into by other reading methods.
+alternative very carefully, if at all." (p.
+164).
+This suggests that the Maze task may rely on the same language processing facilities tapped into by other reading methods.
 Thus, using Maze may not be the best paradigm for studying the process of normal reading, but may be perfectly good or even superior for getting at underlying language processing.
@@
 A word can be unpredictable for a variety of reasons: it could be low frequency, semantically unexpected, the start of a low-frequency syntactic construction, or a word that disambiguates prior words to a less common parse.
-Many targeted effects of interest can thus be potentially accommodated theoretically as specific features that influence word predictability.
-^[Of course, not all effects can necessarily be reduced to word predictability effects, and effects that *cannot* be reduced to word predictability may be of particular theoretical interest.
+Many targeted effects of interest can thus be potentially accommodated theoretically as specific features that influence word predictability.^[Of course, not all effects can necessarily be reduced to word predictability effects, and effects that *cannot* be reduced to word predictability may be of particular theoretical interest.
 Candidates include, for example, memory-based effects [@lewis-etal:2006; @levy-fedorenko-gibson:2013jml], noisy-channel error identification [@levy-etal:2009pnas], and the magnitude of processing difficulty in garden-path resolution [@van-schijndel-linzen-2021:single-stage-models; @wilcoxTargetedAssessmentIncremental2021].] Thus incremental processing methods that are sensitive to predictability are useful for testing linguistic theories that make predictions about what words are unexpected.
@@
 A variety of pre-trained models exist, with varied internal architectures and training methods, but all of them generate measures of predictability.
-Predictability is often measured in bits of surprisal, which is the negative log probability of a word [@hale:2001, @levy:2008].
+Predictability is often measured in bits of surprisal, which is the negative log probability of a word [@hale:2001; @levy:2008].
 1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time, etc.
@@
 Well-chosen distractors and attentive participants reduce the error rate, but the error rate will still compound over long materials.
-For instance, with a 1% error rate, `r round(.99**15*100)`%  of participants would complete each 15-word sentence, but only `r round(.99**50*100)`% would complete a 50 word vignette, and `r round(.99**200*100)`% would complete a 200 word passage.
+For instance, with a 1% error rate, `r round(.99**15*100)`%  of participants would complete each 15-word sentence, but only `r round(.99**50*100)`% would complete a 50-word vignette, and `r round(.99**200*100)`% would complete a 200-word passage.
 In order to run longer materials, we needed something to do when participants made a mistake other than terminate the entire item.
@@
 We excluded data from those who did not report English as their native language, leaving `r data_filt %>% select(subject) %>% unique() %>% nrow()` participants.
-After examining participants' performance on the task (see Results for details), we excluded data from participants with less than 80% accuracy, removing participants whose behavior was consistent with random guessing.
+After examining participants' performance on the task (see results for details), we excluded data from participants with less than 80% accuracy, removing participants whose behavior was consistent with random guessing.
 After this exclusion, `r some %>% select(subject) %>% unique() %>% nrow()` participants were left.
@@
 
+foo <- a$pkgs[1] |> str_sub(3,-3)
 ```
@@
 For SPR comparisons, we included predictors from the current and past three words to account for known spillover effects.
-We conducted data processing and analyses using `r a$r`.
-
-`r a$pkgs`
+We conducted data processing and analyses using R Version 4.2.2 [@R-base] [^papaja_pkg_citations].
+
+`r foo`
 
@@
 We excluded words for which we didn't have surprisal or frequency information, leaving `r nrow(not_na)` words.
-We additionally excluded words that any model treated as being composed of multiple tokens (primarily words with punctuation),  leaving `r nrow(not_gpt)` words[^1].
-We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
+We additionally excluded words that any model treated as being composed of multiple tokens (primarily words with punctuation),  leaving `r nrow(not_gpt)` words.[^1] We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted).
 We exclude RTs from words where mistakes occurred or which occurred after a mistake in the same sentence.
@@
 [^1]:Surprisals should be additive, but summing the surprisals for multi-token words gave some unreasonable responses.
-For instance, in one story the word king!\' has a surprisal of 64 under GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, \'What is the use of flying any higher?
+For instance, in one story the word "king!\'" has a surprisal of 64 under GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, \'What is the use of flying any higher?
 This victory is in the bag and I am king!\').
@@
 (ref:error-cap) A.
-Participant's accuracy on the Maze task (fraction of words selected correctly) versus their average reaction time (in ms).
+Accuracy on the Maze task (fraction of words selected correctly) versus their average reaction time (in ms).
 Many participants (marked in green) chose the correct word >80% of the time; others (in red) appear to be randomly guessing.
@@
 This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.
-We believe the high level of random guessing is an artifact of the subject population [@hauser2018], and we expect that following current recommendations for participant recruitment, such as using qualification cutoffs or another recruitment site would result in fewer participants answering randomly [@eyal2021;@peer2017].
+We believe the high level of random guessing is an artifact of the subject population [@hauser2018], and we expect that following current recommendations for participant recruitment, such as using qualification cutoffs or another recruitment site, would result in fewer participants answering randomly [@eyal2021;@peer2017].
 
@@
   geom_point(alpha=.5, size=1) + 
-  labs(x="A-Maze", y="SPR") + 
+  labs(x="A-maze", y="SPR") + 
   coord_fixed(ratio=1) +
@@
 We used surprisals generated by 4 different language models for robustness.
-(See Methods for details on language models, exclusions, and model fit.)
+(See methods for details on language models, exclusions, and model fit.)
 
@@
 Density of data is shown along the x-axis.
-The best-fit lines is from the mean estimated effect size across the bootstrap replicates, and the shaded areas indicate a 95\% bootstrap confidence interval on this effect size.
+The best-fit lines are from the mean estimated effect size across the bootstrap replicates, and the shaded areas indicate a 95\% bootstrap confidence interval on this effect size.
 For each of the 4 language models used, there is a linear relationship between current word surprisal and RT.
@@
 Density of data is shown along the x-axis.
-The best-fit lines is from the mean estimated effect size across the bootstrap replicates, and the shaded areas indicate a 95\% bootstrap confidence interval on this effect size.
+The best-fit lines are from the mean estimated effect size across the bootstrap replicates, and the shaded areas indicate a 95\% bootstrap confidence interval on this effect size.
 
@@
 For all of these, we included the predictors for the current and previous word, and we centered, but did not rescale, all predictors.
-(See Methods for more details on these predictors and model fit process.) As with the GAM models, we used surprisal calculations from 4 different language models for robustness.
+(See methods for more details on these predictors and model fit process.) As with the GAM models, we used surprisal calculations from 4 different language models for robustness.
 
@@
 
-The Maze linear model effects are shown in Figure \@ref(fig:coeffs-maze) (See also Appendix B for a table of effects).
+The Maze linear model effects are shown in Figure \@ref(fig:coeffs-maze) (see also Appendix B for a table of effects).
 Across all models, there were consistent large effects of length and surprisal at the current word, but minimal effects of frequency.
-This lack of frequency effects differs from the results usually reported for SPR and eye-tracking (though see @shainLargescaleStudyEffects2019).
+This lack of frequency effects differs from the results usually reported for SPR and eye-tracking [though see @shainLargescaleStudyEffects2019].
 There was a small interaction between surprisal and length at the current word.
@@
 
-(ref:coeffs-spr) Point estimates and 95% confidence intervals (+/- 1.97 standard error) for coefficients predicted by fitted regression models predicting SPR RT.
+(ref:coeffs-spr) Point estimates and 95% confidence intervals (+/-1.97 standard error) for coefficients predicted by fitted regression models predicting SPR RT.
 Units are in ms.
@@
 We ran similar models for SPR, although to account for known spillover effects, we consider predictors from the current and 3 previous words.
-Due to issues fitting models, the details of the models differed (see Methods).
+Due to issues fitting models, the details of the models differed (see methods).
 The SPR coefficients are shown in Figure \@ref(fig:coeffs-spr) (see also Appendix B for a table of coefficients).
@@
 
-Whereas Maze showed surprisal effects in the 10 to 25 ms/bit range and length effects in the 15 to 20 ms/character range, SPR effects are about 1-2 ms per bit or character.
+Whereas Maze showed surprisal effects in the 10 to 25 ms/bit range and length effects in the 15 to 20 ms/character range, SPR effects are about 1 to 2 ms per bit or character.
 This difference in effect size is disproportionate to the overall speed of the methods; the predicted intercept for the Maze task was roughly 880 ms and for SPR was roughly 360 ms.
@@
 We assessed each model's fit to A-maze data using log likelihood and R-squared.
-Then we did a nested model comparison, looking at whether a model with multiple surprisal predictors (ex, GRNN and GPT-2) had a better fit than a model with only one (ex GRNN alone).
+Then we did a nested model comparison, looking at whether a model with multiple surprisal predictors (ex.
+GRNN and GPT-2) had a better fit than a model with only one (ex.
+GRNN alone).
 
@@
 
-read_rds(here("Analysis/models/maze_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data.
+read_rds(here("Analysis/models/maze_model_compare.rds")) |> mutate(`Log Lik`=str_replace(`Log Lik`,"-","--")) |> 
+knitr::kable(format="latex", position="ht",caption="Results of model comparisons on Maze data.
 Each row shows the additional predictive value gained from adding that model to another model.
@@
 The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020, and contrasts with the Maze results, where the 5-gram model had the worst fit and did not provide additional predictive value over the other models.
-While the nature of the generalizations made by these neural network-based models are not fully understood, controlled tests have suggested that their next-word predictions often reflect deeper features of linguistic structure [@warstadt-etal:2020-BLiMP; @huSystematicAssessmentSyntactic2020], such as subject--verb agreement [@marvin-linzen:2018-targeted] and wh-dependencies [@wilcox-etal:2022-using-computational-models], and are sensitive over longer context windows, than n-gram models.
+While the nature of the generalizations made by these neural network-based models is not fully understood, controlled tests have suggested that their next-word predictions often reflect deeper features of linguistic structure [@warstadt-etal:2020-BLiMP; @huSystematicAssessmentSyntactic2020], such as subject--verb agreement [@marvin-linzen:2018-targeted] and wh-dependencies [@wilcox-etal:2022-using-computational-models], and are sensitive over longer context windows, than n-gram models.
 The fact that the neural language models dominate the 5-gram models for Maze but not SPR thus suggests that Maze RTs may be more sensitive than SPR RTs to richer language structure-related processes during real-time comprehension.
@@
 
-read_rds(here("Analysis/models/spr_model_compare.rds")) %>% knitr::kable(format="latex", position="ht",caption="Results of model comparisons on SPR data.
+read_rds(here("Analysis/models/spr_model_compare.rds")) |> mutate(`Log Lik`=str_replace(`Log Lik`,"-","--")) |>  knitr::kable(format="latex", position="ht",caption="Results of model comparisons on SPR data.
 Each row shows the additional predictive value gained from adding that model to another model.
@@
 We then used A-maze distractors and the error-correction Maze presentation to gather data on participants reading stories from the Natural Stories corpus in the Maze.
-As laid out in the Introduction, this current study addressed five main questions.
+As laid out in the introduction, this current study addressed five main questions.
 
@@
 Maze and SPR pick up on similar features in words, as shown by the high correlations between Maze and SPR RTs on the sentence level.
-The correlation within Maze is higher than Maze to SPR correlation or SPR-SPR correlations, which is evidence that Maze is less noisy than SPR.
+The correlation within Maze is higher than the Maze to SPR correlation or SPR-SPR correlations, which is evidence that Maze is less noisy than SPR.
 
@@
 For Maze, we found large effects of the current word's surprisal and length, which dwarfed any spillover effects from previous word predictors.
-In contrast, for SPR, we found effects of roughly equal sizes from the current and previous words.
-^[Furthermore, the typical spillover profile for SPR data may be worse than suggested by the Natural Stories corpus SPR data: for example, @smithEffectWordPredictability2013 found that most of a word's surprisal effect showed up only one to two words downstream.] Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus, and thus generating less noisy data.
+In contrast, for SPR, we found effects of roughly equal sizes from the current and previous words^[Furthermore, the typical spillover profile for SPR data may be worse than suggested by the Natural Stories corpus SPR data: for example, @smithEffectWordPredictability2013 found that most of a word's surprisal effect showed up only one to two words downstream.].
+Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus, and thus generating less noisy data.
 We do not find frequency effects on the Maze data, but we do on the SPR data.
@@
 In this case, faster methods like eye-tracking and SPR would show frequency effects (in addition to surprisal), but slower methods like Maze would not as the additional demands slow down the response, allowing more contextual information to be used.
-While this is a difference between Maze and other incremental processing methods, we do not consider it a flaw for Maze---indeed, for researchers interested in focusing on context-contingent language processing, it may suggest an advantage for the Maze task.
+While this is a difference between Maze and other incremental processing methods, we do not consider it a flaw for Maze &ndash; indeed, for researchers interested in focusing on context-contingent language processing, it may suggest an advantage for the Maze task.
 Regardless, these differences highlight the importance of understanding task demands of different incremental processing methods.
@@
 On the whole, our results support the idea that Maze forces language processing to be close to word-by-word, and thus the Maze task can be used under the assumption that the RT of a word primarily reflects its own properties and not those of earlier words.
-Correlation analysis between Maze and SPR suggest that Maze is picking up on many of the same patterns as does SPR, but with less noise.
+Correlation analysis between Maze and SPR suggests that Maze is picking up on many of the same patterns as does SPR, but with less noise.
 
@@
 
-Compared to traditional Maze,in error-correction Maze, participants' incentives to finish quickly are in less conflict with the experimenter's desire that participants do the task as intended.
+Compared to traditional Maze, in error-correction Maze, participants' incentives to finish quickly are in less conflict with the experimenter's desire that participants do the task as intended.
 However, even with error-correction Maze, clicking randomly is still likely faster than doing the task.
-In discussing this work, we received the suggestion that one way to further disincentivize random clicking would be to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms) before correcting their mistake.
+In discussing this work, we received the suggestion that one way to further disincentivize random clicking would be to add a pause when a participant makes a mistake, forcing them to wait some short period of time, such as 500ms, before correcting their mistake.
 This delay would make randomly hitting buttons slower than doing the task as intended, and we have made delaying after wrong presses an option in the error-correction Maze implementation at <!--link obscured for anonymity-->https://github.com/vboyce/Ibex-with-Maze.
@@
 
+# Data Accessibility {.unnumbered}
+Data and materials are available at https://github.com/vboyce/natural-stories-maze. 
+# Ethics and consent {#ethics-and-consent-optional .unnumbered}
+
+This research was approved by MIT’s Committee on the Use of Humans as Experimental Subjects and run under protocol number 1605559077.
+
+# Funding information {#funding-information-optional .unnumbered}
+
+RPL acknowledges support from NSF grant BCS-2121074, NIH grant U01-NS121471, and the MIT–IBM Artificial Intelligence Research Lab.
+
+# Acknowledgements {#acknowledgements-optional .unnumbered}
+
+We thank the AMLAP 2020 audience, the Computational Psycholinguistics Lab at MIT, the Language and Cognition Lab at Stanford, the QuantLang Lab at UC Irvine, and Mike Frank for feedback on this work.
+
+# Competing interests {#competing-interests-mandatory .unnumbered}
+
+The authors have no competing interests to declare.
+
+# Authors' contributions {#contrib .unnumbered}
+VB contributed Conceptualization, Formal Analysis, Investigation, Methodology, Software, and Writing - Original Draft Preparation.
+RPL contributed Conceptualization, Formal Analysis, Funding Acquisition, Methodology, Supervision, and Writing - Review & Editing.
+
 \newpage
@@
 
-# Appendix A
+\newpage
+
+# Appendix A {.unnumbered}
 The beginning of one of the stories. This excerpt is the first 200 words of a 1000 word story.
@@
 
-# Appendix B
+# Appendix B {.unnumbered}
 
@@
 
-# Appendix C
+# Appendix C {.unnumbered}
 
@@
 
-# Appendix D
+# Appendix D {.unnumbered}
 
@@
 ```
-
-# Ethics and consent {#ethics-and-consent-optional .unnumbered}
-
-This research was approved by MIT’s Committee on the Use of Humans as Experimental Subjects and run under protocol number 1605559077.
-
-# Funding information {#funding-information-optional .unnumbered}
-
-RPL acknowledges support from NSF grant BCS-2121074, NIH grant U01-NS121471, and the MIT–IBM Artificial Intelligence Research Lab.
-
-# Acknowledgements {#acknowledgements-optional .unnumbered}
-
-We thank the AMLAP 2020 audience, the Computational Psycholinguistics Lab at MIT, the Language and Cognition Lab at Stanford, the QuantLang Lab at UC Irvine, and Mike Frank for feedback on this work.
-
-# Competing interests {#competing-interests-mandatory .unnumbered}
-
-The authors have no competing interests to declare.
-
-# Authors' contributions {#contrib .unnumbered}
-VB contributed Conceptualization, Formal Analysis, Investigation, Methodology, Software, and Writing - Original Draft Preparation.
-RPL contributed Conceptualization, Formal Analysis, Funding Acquisition, Methodology, Supervision, and Writing - Review & Editing.
